### 2023-05-03 manuscript/index.qmd
friendship ended with pdflatex, now quarto is my best friend

# Abstract {.unnumbered}

The literature around the study of neural field dynamics is centered around the convergence of discrete neural networks to a smooth continuum in some $p$-dimensional space (@Bressloff_2012, @schmutz2023convergence).

In this thesis, we show that this dimensionality is arbitrary, and demonstrate examples of equivalent dynamics emerging from neural fields in different dimensions. We derive method to reduce dimensionality of any neural field by employing fractal mappings, and study properties that these must satisfy.

### 2023-05-03 manuscript/intro.qmd
friendship ended with pdflatex, now quarto is my best friend

# Introduction

Rp smooth, dynamics work
-> [0,1], do the dynamics work ?
there are non-continuous embeddings, so neural field disc.
is not continuous -> does it break the dynamics ?
                  -> are dynamics well-defined in [0,1] ?
                     [cite Spatiotemporal dynamics of continuum neural fields : in litterature
					   the kernel is always continuous. ]
				  -> here the "field" is nowhere continuous (fractal), but is it still well-defined
				  -> numerical answers

why low-rank ? -> it is easy to simulate, we just use it to illustrate our examples @schmutz2023convergence

do not ask this question "can we write a neural field ?" 
but rather this       -> "what neural field can we write given a regularity ?"

here we have the first example of two equivalent dynamics. one with high regularity and dimension.,
the other with low regularity and dim.

why are we doing this ?
dimensionality of activity = rank(connectivity)
normally, we would think that if we can write a neural field equation, then the embedding space if of dimension rank(connectivity).
-> but we have shown that is not the case ! with the correct mapping, we can write a neural field in 1D, even though the dimensionality of the activity is 2D !
-> not even a tradeoff. IDEA : smooth function in 1D. when we map back to 2D using recusive local, the mapped space is everywhere discontinuous, but (TEST THIS !!) the dynamics are the same. -> THE DIMENSION OF THE EMBEDDING IS UNRELATED TO THE DIMENSION OF THE ACTIVITY. Rather, we should ask in which dimension/embedding the connectivity kernel is going to have a nice expression (e.g. continuous etc)

@pham2016existence

![connectivity](figures/connectivity_binning.png)

### 2023-05-08 manuscript/chapters/1_intro.qmd
prrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrr

# Introduction

::: hidden
{{< include ../macros.tex >}}
:::

## Networks of rate neurons

Networks of rate neurons model the time evolution of the potential $h_i(t) \,[\textrm{V}]$ associated to $N$ individual neurons labeled $i\in\{1,\cdots,N\}$. In their most general form, a connectivity matrix $J_{ij}$ describes the amplitude of the recurrent currents $I^\text{rec}_{ij} = J_{ij} \phi(h_j(t)) \,[\textrm{A}]$ emitted by neuron $j$ and received by neuron $i$. The (monotonically increasing) function $\phi(h)$ has units of $\textrm{Hz}$, and describes the spike firing rate of a given neuron as a function of its potential. In the absence of recurrent currents, the potential of each neuron decays exponentially with time constant $\tau \,[\textrm{s}]$. Neurons can also receive external currents $I^\text{ext}_i(t) \,[\textrm{A}]$. Conversion of current to voltage happens through the resistivity factor $R_i \,[\textrm{Ohm}]$. @eq-rnn writes out the full equation a rate neural network.

$$
\tau \dot h_i(t) = \underbrace{-h_i(t)}_{\text{exponential decay}} + R_i \sum_{j=1}^{N} \underbrace{J_{ij} \phi(h_j(t))}_{I^\text{rec}_i(t)} + R_i I^{\text{ext}}_i(t)
$$ {#eq-rnn}

This model can intuitively be seen as the long time limit of a network of spiking neurons, in which neurons emit current spikes (modeled as $\delta$-functions) at some time-variable rate $\phi(h_i(t))$. The convergence is formally proven in @schmutz2023convergence.

In the following, we set $\tau=1$ and $R_i=1$ for all neurons $i$, since these just correspond to rescaling time and currents. Additionally, we consider $I^\text{ext}_i(t)=0$ to simplify the model.

## Convergence to a neural field equation

Neural field equations emerge from the structure of rate neural networks as the number of neurons $N$ becomes large. Positions $\vec \xi_i \in \Rp$ are attributed to each neuron, and define the *embedding space*. If the embedding is well-chosen, then as $N \rightarrow \infty$ neurons close in the embedding space tend to have similar potentials $h_i(t)$, such that these tend to a continuum $h(t, \vec z)$, where $\vec z \in \Rp$.

![Every neuron in a rate neural network has a position in the $p$-dimensional embedding space](figures/neural_field_discrete.png)

We write down the neural field equation @eq-nf, in which $\rho(\vec z)$ measures the density of neurons inside the embedding space. The connectivity kernel $w(\vec z, \vec y)$ can be seen as the "infinite $N$ limit" of the connectivity matrix $J_{ij}$, in the sense that $w(\vec z, \vec y)$ describes the influence of the neuron population in the box at $[\vec y, \vec y + \d \vec y]$ on the neuron population $[\vec z, \vec z + \d \vec z]$.

$$
\partial_t h(t, \vec z) = -h(t, \vec z) + \int_{\mathbb{R}^p} w(\vec z, \vec y) \phi(h(t, \vec y)) \rho(\mathrm d \vec y)
$$ {#eq-nf}

![The connectivity kernel defines the influence between neural densities in the embedding space](figures/connectivity_kernel.png)

In the litterature, the embedding space is often thought of as being closely related to the physical positions of the neurons @Bressloff2012. Rigorous proofs of convergence are still an active area of study @chevallier2018mean and are not the subject of this report.

The results we present here should hold for any neural field, but for the sake of argument we shall consider a low-rank rate neural network, for which there are informal proofs of convergence to a neural field equation  @schmutz2023convergence \[TODO : reformulate this\]. Low-rank neural networks hold their name from the rank of their connectivity matrix, defined in @eq-lrj.

$$
J_{ij} = \frac 1N \sum_{\mu=1}^p F_{\mu,i} G_{\mu,j}
$$ {#eq-lrj}

In the case that $\vec F_i = (F_{1,i},\cdots,F_{p,i}) \in \Rp$ sample from some $p$-dimensional distribution $\rho$, the vectors $\vec F_i$ define a "natural embedding" of neuron $i$ in $\Rp$. When the number of neurons increases, the numeric density of neurons in the embedding approaches the actual probability distribution $\rho$, and we can intuitively see the convergence towards a connectivity kernel $w(\vec z, \vec y) = \sum_{\mu=1}^p z_\mu y_\mu$.

## A one-dimensional neural field

The paper @jabin2022meanfield provides some motivation for the existence of an equivalent neural field in one dimension. Their main result is that under some boundedness assumptions of the connectivity, the limiting equation describing the dynamics of the system incorporates only an integral over $[0,1]$. In our low-rank case, the weights scale as $J_{ij} \sim \frac 1N$, and so we can expect that it should be possible to perform a change of variables from $\Rp$ to $[0,1]$.

The idea is that we define a measurable mapping $S : \Rp \mapsto [0,1]$ and a measure $\mu : [0,1] \mapsto \mathbb{R}^+$ on $([0,1], \mathcal B([0,1]))$, such that we can write the change of variables in the integral :

$$
\int_{\mathbb{R}^p} w(\vec z, \vec y) \phi(h(t, \vec y)) \rho(\mathrm d \vec y) = \int_{[0,1]} [w(\vec z, \cdot) \phi(h(t, \cdot)) \rho(\cdot)] \circ S^{-1} \; \mathrm d \mu
$$

Informally, we would like to define the one-dimensional field $\tilde h(\alpha, t) = h(S^{-1}(\alpha), t)$, connectivity $\tilde w(\alpha, \beta) = w(S^{-1}(\alpha), S^{-1}(\beta))$ and density $\tilde \rho(\alpha) = \rho(S^{-1}(\alpha))$, so that we can write the neural field in the $[0,1]$ embedding :

$$
\partial_t \tilde h(t, \alpha) = -\tilde h(t, \alpha) + \int_{\mathbb{R}^p} \tilde w(\alpha, \beta) \phi(\tilde h(t, \beta)) \tilde \rho(\mathrm d \beta)
$$ {#eq-nf1d}

This change of variable seems counter-intuitive : how is it possible to map a higher dimension to a lower dimension without losing information ? And even if mathematically such a change of variables is possible, is it actually feasible to simulate it numerically ?

## Numerical simulations of neural fields

The goal of this report is to not only give an intuition on the mappings between $[0,1]$ and $\Rp$ embedding spaces, but also show that the mathematical construction works numerically.

We distinguish two types of numerical simulations for the neural field in @eq-nf :

1. A neuron-based simulation of $N$ neurons. Each neuron samples the probability distribution $\rho(z_1, \cdots, z_p)$ independantly. This case then just reduces to simulating the network of rate neurons, which follows @eq-rnn. In other words, we are simulating the full network of rate neurons, and the simulation converges to the true mathematical neural field equation as $N$ becomes large.

2. A discretization-based simulation. In the $\Rp$ space, we discretize each dimension, forming a grid of $p$-dimensional bins. Each bin effectively corresponds to an infinite population of neurons, which are approximated to be the same (this is a mean-field approximation). This works because by continuity of the field, neurons which are close in $\Rp$ also have similar potentials. As we increase the number of bins, the simulation becomes more precise and the dynamics approach that of the true mathematical neural field.

With this in mind, we can try to understand why simulating a neural field in $[0,1]$ might be difficult. Suppose we take bins in the $\Rp$ space, and map each bin to a random position in $[0,1]$. Intuively, this won't work (in the sense that the dynamics in $[0,1]$ will not be the same as the dynamics in $\Rp$), because the hypothesis of continuity is not longer true : the mapping is random, so the nearby populations in $[0,1]$ will no longer have similar potentials, and the error of the mean-field approximation will no longer vanish as we take increasingly many bins.

\[TODO : an illustration for this\]

In some sense, we would like to find a mapping that converses "locality" : populations that are close in $\Rp$ should also be close in $[0,1]$, or at least there shouldn't be "too many large discontinuities". @sec-mapping is dedicated to the discussion of these mappings, and gives various numerical demonstrations of the resulting dynamics.

<!--

neurons in some Rp space
N -> infty
spatial structure
	=> close neurons -> similar potential
	=> continuum of neurons
	=> neural field

"natural embedding" (here in Rp) is NOT UNIQUE

we show : 
* we can embed in a space of lower dimension.
* not just a math abstraction (solution of the 1d neural field is well-defined)
* not only well-defined, but actually works numerically. we assert the dynamics in [0,1] are the same

[how do we actually "convince ourselves" that is works numerically ?
-> little intro on "discretizing a neural field", either by sampling the PDF, or by discretization
-> particle (=neurons)-based simulation vs grid (=bins)/density (TODO)-based simulation
(with the discretization approach we already have the "infinite number of neurons" approximation inside each bin)

def sampling : tirer au sort un neurone selon la pdf
def discretization : discretize pdf regularly, "bins"
]

introduce "RandomMapping", to show why this mapping is not trivial

why it doesnt work : even in a little bin, we will have many "wildly different neurons", in the sense scattered far away in 2D

what we want in [0,1] : as bins become smaller, the "populations inside the bin" have more and more similar potentials. in the random case, never happens.
if in 2D, they are close, then they have the same potential. close in 2D ?=> close in 1D -> "locality" (how do formulate a mapping which has this property ?) (something close to an "isometric mapping", distances are kind of conserved)

(we show we can afford "some" discontinuity)
 -->

### 2023-05-08 manuscript/chapters/2_neural_fields.qmd
prrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrr

# A simple model for a neural field in $\mathbb R^2$

## Building up a neural field in $\mathbb R^2$

p=1 then p=2, show simulations

## Characterizing dynamics

introduce overlap, and in our case the low-rank model the kappa

overlaps as an approximation of the "dynamics" (latent space), in the case phi=linear

warning : we need absolutely to distinguish the latent and the embedding space

## A cycling neural field

why ? dynamics are more interesting

1. delayed neural field
2. rolled neural field

show what the analytical kappa is

introduce both, derive kappa again

simulate cycling in p=2

### 2023-05-08 manuscript/chapters/3_dynamics_1d.qmd
prrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrr

# Mapping $\mathbb R^2$ to $[0,1]$ and resulting dynamics in $[0,1]$ {#sec-mapping}

## Which types of mappings

peano vs cantor mappings

linear projections don't work ofc

## Sequence of mappings and their limit

intuition for reshape, and the fractals

expose different types of mapping

## Numerically implementing mappings to $[0,1]$

how to we actually simulate the mappings in one-dimension

go into detail on how numerically S is a reordering of the bins

## Simulating the $[0,1]$ neural field

how we numerically implemement a simulation on a fractal

averaging of 2^n neurons

### 2023-05-08 manuscript/index.qmd
prrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrr

@@
-# Abstract {.unnumbered}
+# Index {.unnumbered .hidden}
 
-The literature around the study of neural field dynamics is centered around the convergence of discrete neural networks to a smooth continuum in some $p$-dimensional space (@Bressloff_2012, @schmutz2023convergence).
-
-In this thesis, we show that this dimensionality is arbitrary, and demonstrate examples of equivalent dynamics emerging from neural fields in different dimensions.
-We derive method to reduce dimensionality of any neural field by employing fractal mappings, and study properties that these must satisfy.
+\[TODO : a nice illustration for this, maybe some cool animation ?\]

### 2023-05-31 notebooks/neurodyn/_anim.py
latent space anim

@@ -6,6 +6,7 @@ from pathlib import Path
 from math import ceil
 from ._rnn import LowRankRNN, Result
 from ._plot import plot_overlap_trajectory, scale_lightness
+from ._overlap import overlap
 
 __all__ = ['animate1d', 'animate2d']
 
@@ -65,7 +66,7 @@ def animate1d(rnn: LowRankRNN, res: Result, outpath: Path, time_stride: int = 1,
 		plt.close()
 
 
-def animate2d(rnn: LowRankRNN, res: Result, outpath: Path, time_stride: int = 1, Nmax = 1500):
+def animate2d_old(rnn: LowRankRNN, res: Result, outpath: Path, time_stride: int = 1, Nmax = 1500):
 	idt = 0
 	activity = rnn.phi(res.h)
 	cmap = mpl.colormaps['RdBu_r']
@@ -149,6 +150,73 @@ def animate2d(rnn: LowRankRNN, res: Result, outpath: Path, time_stride: int = 1,
 			# update the scatter
 			# things['sc'].set_facecolors([ scale_lightness(c[:3], 0.7) for c in cmap(activity[:Nmax, idt]) ])
 			things['sc'].set_array(activity[:Nmax, idt])
+
+			pbar.update(1)
+
+		ani = animation.FuncAnimation(fig, update, frames=range(0, len(res.t), time_stride))
+		ani.save(outpath, writer='ffmpeg', fps=15)
+
+		plt.close()
+
+
+def animate2d(rnn: LowRankRNN, res: Result, outpath: Path, time_stride: int = 1, Nmax = 1500):
+	idt = 0
+	activity = rnn.phi(res.h)
+	cmap = mpl.colormaps['RdBu_r']
+
+	fig, axes = plt.subplots(ncols=2, figsize=(9, 4), layout='constrained')
+
+	# 2d plotting
+
+	axes[1].set_xlabel('Overlap $m_0(t)$')
+	axes[1].set_ylabel('Overlap $m_1(t)$')
+	axes[1].set_title('Latent trajectory')
+	axes[1].set_aspect('equal')
+	axes[1].set_xlim((0,1))
+	axes[1].set_ylim((0,1))
+	from matplotlib_tufte import breathe
+	breathe(axes[1])
+
+	m = overlap(rnn, res.h)
+	phaseline, = axes[1].plot(m[0, :1], m[1, :1])
[…]

### 2023-05-31 notebooks/neurodyn/_mapping.py

@@ -10,10 +10,10 @@ __all__ = [
 	'Box',
 	'Mapping', 'BinMapping',
 	'RecursiveLocalMapping',
-	'ReshapeMapping',
+	'ReshapeMapping', 'ColumnMapping',  # alias
 	'DiagonalMapping',
 	'SzudzikMapping',
-	'RecursiveFarMapping',
+	'RecursiveFarMapping', 'AntiZMapping',  # alias
 	'ZMapping',
 	'RandomMapping',
 	'LinearMapping',
@@ -413,6 +413,8 @@ class ReshapeMapping(BinMapping):
 	def __str__(self) -> str:
 		return f'ReshapeMapping{{nx={self.nx}, ny={self.ny}}}'
 
+ColumnMapping = ReshapeMapping
+
 
 class DiagonalMapping(BinMapping):
 	"""Implements [Cantor mapping](https://en.wikipedia.org/wiki/Pairing_function)
@@ -607,6 +609,8 @@ class RecursiveFarMapping(BinMapping):
 
 	def __str__(self) -> str:
 		return f'RecursiveFarMapping{{nrec={self.nrec}}}'
+
+AntiZMapping = RecursiveFarMapping
 
 
 class RandomMapping(BinMapping):

### 2023-05-31 manuscript/chapters/1_intro.qmd
yay for more writing !

@@
 
-::: hidden
+::: {.content-hidden}
 {{< include ../macros.tex >}}
@@
 
-This model can intuitively be seen as the long time limit of a network of spiking neurons, in which neurons emit current spikes (modeled as $\delta$-functions) at some time-variable rate $\phi(h_i(t))$.
+This model can intuitively be seen as the $N \rightarrow \infty$ limit of a network of spiking neurons, in which neurons emit current spikes (modeled as $\delta$-functions) at some time-variable rate $\phi(h_i(t))$.
 The convergence is formally proven in @schmutz2023convergence.
@@
 In the case that $\vec F_i = (F_{1,i},\cdots,F_{p,i}) \in \Rp$ sample from some $p$-dimensional distribution $\rho$, the vectors $\vec F_i$ define a "natural embedding" of neuron $i$ in $\Rp$.
-When the number of neurons increases, the numeric density of neurons in the embedding approaches the actual probability distribution $\rho$, and we can intuitively see the convergence towards a connectivity kernel $w(\vec z, \vec y) = \sum_{\mu=1}^p z_\mu y_\mu$.
+When the number of neurons increases, the numeric density of neurons in the embedding approaches the actual probability distribution $\rho$, and we can intuitively see the convergence towards a connectivity kernel $w(\vec z, \vec y) = \sum_{\mu=1}^p z_\mu g_\mu(y_\mu)$.
 
@@
 $$
-\partial_t \tilde h(t, \alpha) = -\tilde h(t, \alpha) + \int_{\mathbb{R}^p} \tilde w(\alpha, \beta) \phi(\tilde h(t, \beta)) \tilde \rho(\mathrm d \beta)
+\partial_t \tilde h(t, \alpha) = -\tilde h(t, \alpha) + \int_{[0,1]} \tilde w(\alpha, \beta) \phi(\tilde h(t, \beta)) \tilde \rho(\mathrm d \beta)
 $$ {#eq-nf1d}
@@
 Suppose we take bins in the $\Rp$ space, and map each bin to a random position in $[0,1]$.
-Intuively, this won't work (in the sense that the dynamics in $[0,1]$ will not be the same as the dynamics in $\Rp$), because the hypothesis of continuity is not longer true : the mapping is random, so the nearby populations in $[0,1]$ will no longer have similar potentials, and the error of the mean-field approximation will no longer vanish as we take increasingly many bins.
+Intuitively, this won't work (in the sense that the dynamics in $[0,1]$ will not be the same as the dynamics in $\Rp$), because the hypothesis of continuity is not longer true : the mapping is random, so the nearby populations in $[0,1]$ will no longer have similar potentials, and the error of the mean-field approximation will no longer vanish as we take increasingly many bins.
 
@@
 
-In some sense, we would like to find a mapping that converses "locality" : populations that are close in $\Rp$ should also be close in $[0,1]$, or at least there shouldn't be "too many large discontinuities".
+In some sense, we would like to find a mapping that conserves "locality" : populations that are close in $\Rp$ should also be close in $[0,1]$, or at least there shouldn't be "too many large discontinuities".
 @sec-mapping is dedicated to the discussion of these mappings, and gives various numerical demonstrations of the resulting dynamics.

### 2023-05-31 manuscript/chapters/2_neural_fields.qmd
yay for more writing !

@@
-# A simple model for a neural field in $\mathbb R^2$
+# A simple model for a neural field in 2D
 
-## Building up a neural field in $\mathbb R^2$
+## Networks of neurons to neural fields
 
-p=1 then p=2, show simulations
+no external current,
+
+general low-rank neural fields, F and G notation
+
+$$
+J_{ij} = \sum_{\mu=1}^p F_{\mu,i} G_{\mu,j}
+$$
+
+show how the embedding becomes smooth
+
+study of fixed points
+
+- $h_i = \xi_{\mu,i} \quad \forall \mu = 1,\cdots,p$ (stable)
+- $h_i=0$ (unstable)
+
+the fixed points also become more and more stable as N->oo (study of stability with eigenvalues)
+
+## Building up to a neural field in $\mathbb R^2$
+
+Introduce the neural field in 1D, then 2D
+
+Neurons are at position $\vec z = (z_1, \cdots, z_p) \in \mathbb{R}^p$, distributed according to the distribution $\frac{\mathrm{exp}(\frac12\sum_{\mu=1}^p z_\mu^2)}{(2\pi)^{p/2}} \mathrm{d} z_1 \cdots \mathrm d z_p = \rho(\mathrm d \vec z)$
+
+The RNN potential becomes $h(t, \vec z)$ and evolves according to
+
+$$
+\begin{aligned}
+\partial_t h(t, \vec z) &= -h(t, \vec z) + \int_{\mathbb{R}^p} w(\vec z, \vec y) \phi(h(t, \vec y)) \rho(\mathrm d \vec y),
+\quad
+w(\vec z, \vec y) = \sum_{\mu=1}^p \tilde \phi (y_\mu) z_\mu \\
+&= -h(t, \vec z) + \sum_{\mu=1}^p z_\mu m_\mu(t)
+\end{aligned}
+$$
+
+study of fixed points (and convergence of the fixed points)
+
+simulation strategies : grid and sampling
 
@@
 
+### Overlap variables
+
+networks of neurons :
+
+$$
+m_\mu(t) = \frac 1N \sum_{i=1}^N \tilde \phi(\xi_{\mu,i}) \phi(h_i(t))
+$$
+
+neural field :
+
+$$
+m_\mu(t) = \int_{\mathbb R^p} \tilde \phi(y_\mu) \phi(h(t,\vec y)) \rho(\mathrm d \vec y)
+$$
+
+### Low-rank neural fields as $p$-dimensional closed systems
+
 introduce overlap, and in our case the low-rank model the kappa
 
-overlaps as an approximation of the "dynamics" (latent space), in the case phi=linear
+overlaps as an approximation of the “dynamics” (latent space), in the case phi=linear
+
+The dynamics of $h(t, \vec z)$ are in a subsystem of dimension $p$ spanned by the ONB of functions $\{e_\mu(\vec z) = z_\mu | \mu=1,\cdots,p\}$, with the scalar product $\langle f, g \rangle = \int_{\mathbb R^p} f(\vec y) g(\vec y) \rho(\mathrm d \vec y)$.
+
+We decompose $h(t, \vec z) = h^\perp(t, \vec z) + \sum_{\mu=1}^p \kappa_\mu(t) z_\mu$, and write the system of equations for $\mu=1,\cdots,p$.
+
+$$
+\dot \kappa_\mu(t) = -\kappa_\mu(t) + \int_{\mathbb{R}^p} \tilde\phi(y_\mu) \phi(h(t, \vec y)) \rho(\mathrm d \vec y) = -\kappa_\mu(t) + m_\mu(t)
+$$
+
+with initial conditions
+
+$$
+\kappa_\mu(0) = \int_{\mathbb{R}^p} y_\mu h(0, \vec y) \rho(\mathrm d \vec y)
+$$
+
+and the orthogonal component evolves according to
+
+$$
+h^\perp(t, \vec z) = h^\perp(0, \vec z) \mathrm e^{-t},
+\quad
+h^\perp(0, \vec z) = h(0, \vec z) - \sum_{\mu=1}^p \kappa_\mu(0) z_\mu
+$$
+
+### Dynamics in the overlap latent space
 
 warning : we need absolutely to distinguish the latent and the embedding space
+
+Relation between $p$-dimensional closed system and neural field equation in $\mathbb R^p$
+
+$$
+\begin{aligned}
+\partial_t h(t, \vec z) &= \partial_t h^\perp(t, \vec z) + \sum_{\mu=1}^p z_\mu \dot \kappa_\mu(t) \\
+&= -h^\perp(t, \vec z) - \sum_{\mu=1}^p z_\mu \kappa_\mu(t) + \sum_{\mu=1}^p z_\mu m_\mu(t) \\
+&= -h(t, \vec z) + \sum_{\mu=1}^p z_\mu m_\mu(t)
+\end{aligned}
+$$
+
+we use the latent space to assert dynamics are the same
 
@@
 simulate cycling in p=2
+
+recurrent current picture

### 2023-05-31 manuscript/chapters/3_mappings.qmd
yay for more writing !

# Mapping the neural field in 2D to a 1D embedding

## Simple example : population dynamics

the case of population dynamics can easily be embedded in 1D

## Neural field in $[0,1]$

Define a measurable mapping $S : \mathbb{R}^p \rightarrow [0, 1]$.

Let $\mu : [0,1] \rightarrow [0, \infty]$ be a measure on $([0,1], \mathcal B([0,1]))$.

Then we can write

$$
\partial_t h(t, \vec z) = -h(t, \vec z) + \int_{[0,1]} [w(\vec z, \cdot) \phi(h(t, \cdot)) \rho(\cdot)] \circ S^{-1} \; \mathrm d \mu
$$

## From $\R^2$ to $[0,1]^2$

Defining the change of variables $u_\mu=\mathrm{CDF}(y_\mu), v_\mu=\mathrm{CDF}(z_\mu)$, the neural field becomes

$$
\partial_t h_U(t, \vec v) = -h_U(t, \vec v) + \int_{[0,1]^p} w_U(\vec v, \vec u) \phi(h_U(t, \vec u)) \mathrm d \vec u, \quad
w_U(\vec v, \vec u) = w(\mathrm{CDF}^{-1}(\vec v), \mathrm{CDF}^{-1}(\vec u)), \quad h_U(t, \vec v) = h(t, \mathrm{CDF}^{-1}(\vec v))
$$

We now have a neural field equation on a unit (hyper)cube, and uniform sampling.

## Mappings of $[0,1]^2$ to $[0,1]$

### Types of mappings

peano vs cantor mappings

linear projections don’t work ofc

We consider $(x,y) \in [0,1]²$. Given $n \in \mathbb N$, we write a (truncated) binary expansion of $x$ as (and similarly for $y$)

$$
b^x_k = \mathrm{Ind}\left\{2^{k-1}x - \lfloor2^{k-1}x\rfloor \geq \frac{1}{2}\right\}, \text{ such that } x= \sum_{k=1}^{n} b^x_k 2^{-k} = 0.b^x_1b^x_2\cdots b^x_{n}
$$

#### Z-mapping

Also called Z-order curve, Morton mapping. It conserves locality 2D -> 1D.

$$
z = Z(x,y) = 0.b^x_1 b^y_1 b^x_2 b^y_2 \cdots b^x_n b^y_n = \sum_{k=1}^{n} b^x_k 2^{1-2k} + b^y_k 2^{-2k}
$$

Its inverse is :

$$
\begin{aligned}
x = Z^{-1}_x(z) &= 0.b_1 b_3 \cdots b_{2n-1} = \sum_{k=1}^n b_{2k-1} 2^{-k} \\
y = Z^{-1}_y(z) &= 0.b_2 b_4 \cdots b_{2n} = \sum_{k=1}^n b_{2k} 2^{-k}
\end{aligned}
$$

#### Column mapping

Also called "Reshape mapping" in the code. It converges to a projection on the $x$-axis

$$
z = C(x,y) = 0.b^x_1 b^x_2 \cdots b^x_n b^y_1 b^y_2 \cdots b^y_n
$$

Its inverse is :

$$
\begin{aligned}
x = C^{-1}_x(z) &= 0.b_1 b_2 \cdots b_n = \sum_{k=1}^n b_{k} 2^{-k} \\
y = C^{-1}_y(z) &= 0.b_{n+1} b_{n+2} \cdots b_{2n} = \sum_{k=1}^n b_{n+k} 2^{-k}
\end{aligned}
$$

#### Anti-Z mapping

Also called in code "Far mapping" (initially I didn't see the link with the Z-order curve), because it destroys locality 2D -> 1D.

$$
z = A(x,y) = 0.b^x_n b^y_n b^x_{n-1} b^y_{n-1} \cdots b^x_1 b^y_1 = \sum_{k=1}^{n} b^x_{n+1-k} 2^{1-2k} + b^y_{n+1-k} 2^{-2k}
$$

Its inverse is :

$$
\begin{aligned}
x = A^{-1}_x(z) &= 0.b_{2n-1} b_{2n-3} \cdots b_{1} = \sum_{k=1}^n b_{2(n-k)-1} 2^{-k} \\
y = A^{-1}_y(z) &= 0.b_{2n} b_{2n-2} \cdots b_{2} = \sum_{k=1}^n b_{2(n+1-k)} 2^{-k}
\end{aligned}
$$

it does not converve

TODO : 3D vis of the mappings

### Sequence of mappings and their limit

intuition for reshape, and the fractals

expose different types of mapping

argue over binary expansion

### Numerical implementation

how to we actually simulate the mappings in one-dimension

how to we actually simulate the mappings in one-dimension

go into detail on how numerically S is a reordering of the bins

we use the fact we have a compact set onto another [0,1]² -> [0,1], so we can write a binary expansion of (x, y) -> z.
x has n bits, y has n bits, z has 2n bits.
an arbitrary mapping can then be written as a recombination of the bits of x and y (and any other constant) into 2n bits
this is good because with floating point we lose precision, here we work with integers (~ fixed point precision), so the implementations of the mappings are exact.

the locality of 2D -> 1D is not important (i.e neurons can be repeated in 1D without too much trouble). locality 1D -> 2D is much more important (a neuron in 1D must refer to only one (not multiple) neurons in 2D)

## Reinterpretation of simulation strategies

### Bin averaging samples

give an intuition, show the figure and some simulations to show that it works

### Direct remapping of the grid

We define $\vec z(\alpha) = S(\alpha)$ the 2D point corresponding to the mapping $\alpha$. (details : we need a bounding box to map back). Defining a matrix $Z_{\mu,\alpha} = \vec z(\alpha)_\mu$, we can write down a numerical PDF $\tilde \rho(Z_{1,\alpha},\cdots,Z_{p,\alpha}) = \tilde \rho(Z_{:,\alpha}) = \frac{\rho(Z_{:,\alpha})}{\sum_\beta \rho(Z_{:,\beta})}$ and the following patterns to simulate the embedded $[0,1]$ neural field as a low-rank RNN.

$$
\tilde F_{\mu,\alpha} = Z_{\mu,\alpha}, \quad \tilde G_{\mu,\alpha} = \tilde\phi(Z_{\mu,\alpha}), \quad \tilde J_{\alpha,\beta}=\tilde \rho(Z_{:,\beta}) \sum_{\mu=1}^p \tilde F_{\mu,\alpha} \tilde G_{\mu,\beta}
$$

Doing this is equivalent to the formulation of the binned connectivity matrix.

## Numerical simulations of the neural field in $[0,1]$

how we numerically implemement a simulation on a fractal

averaging of 2ˆn neurons

see sampling.ipynb

### 2023-05-31 manuscript/chapters/4_locality.qmd
yay for more writing !

# Locality and numerical convergence

## Analytical equivalence of the dynamics

composition of functions

bijectivity (actually injection in one sense)

## Locality

the notion of infinite number of discontinuities, but the size counts too

plot distance 2D vs distance 1D
* "close to the line y=sqrt2 x" (but fractals dont get arbitrarily close)
* "size of the discontinuities"
* L1 distance

“resilient to some discontinuity”
the notion of infinite number of discontinuities, but the size counts too

Total variation vs bining variation

histogram of discontinuities ?

$$
\frac{1}{2^n}\sum_{i=1}^{2^n} \sup_{z,z' \in \{\frac{k}{4^n} | k= 1, \cdots, 2^n\}} |f^n_1(\tfrac{i-1}{2^n}+z) - f^n_1(\tfrac{i-1}{2^n}+z')| + |f^n_2(\tfrac{i-1}{2^n}+z) - f^n_2(\tfrac{i-1}{2^n}+z')|
$$

plot the bins in the 2D scatter, so we can see what we are doing.

it is different to TV, TV is projecting all the jumps

## Numerical feasibility

failed notions of discontinuity

locality might be a side-effect from bijections

measure for V

proof that numerical convergence

## Details

ironing out details : here we considered 2d -> 1d, but possible to repeat the application of the mapping so that Rp -> Rp-1 -> .. -> 1d

also should work with any neural field, here low-rank cycling was just for the purpose of demonstration

### 2023-05-31 manuscript/chapters/6_appendix.qmd
yay for more writing !

# Appendix

::: {.content-hidden}
{{< include ../macros.tex >}}
:::

## Fixed point study in the network of neurons

We recall the equation for the low-rank network of neurons :

$$
\dot h_i(t) = -h_i(t) + \frac 1N \sum_{\mu=1}^p \sum_{j=1}^N F_{\mu i} G_{\mu j} \phi(h_j(t))
$$

The low-rank vectors are in general obtained by sampling a distribution $\vec{\xi} = (\xi_1, \cdots, \xi_p) \sim \rho$. Every neuron $i$ samples this distribution independantly, and low-rank vectors are obtained by passing the samples through functions $f_\mu(\vec{\xi})$ and $g_\mu(\vec{\xi})$.

We note that in the main text, the distribution is a multivariate normal with unit covariance and zero mean, such that it factors out as $\rho(\xi_1, \cdots, \xi_p) = \prod_{\mu=1}^p \mathcal{N}(\xi_\mu)$. Furthermore, the functions are component-wise $f_\mu(\xi_1, \cdots, \xi_p) = \xi_\mu$ and $g_\mu(\xi_1, \cdots, \xi_p) = \tilde \phi(\xi_\mu) = \frac{\phi(\xi_\mu) - \avg{\phi(\xi_\mu)}}{\mathrm{Var}[\phi(\xi_\mu)]}$.

$$
F_{\mu i} = f_\mu(\xi_{1 i}, \cdots, \xi_{p i}), \, G_{\mu i} = g_\mu(\xi_{1 i}, \cdots, \xi_{p i})
$$

### Derivation of fixed points

In order to find the fixed points, one has to solve in general $\dot h_i(t) = 0$ for all $i = 1, \cdots, N$. Because of the nonlinearity $\phi(h_j(t))$, this is difficult, but we can guess the fixed points $h^\star_i$ and verify them by substituting into the equation that $\dot h_i(t) = 0$.

#### Zero fixed point

A trivial fixed point is $h^\star_i = 0$ for all $i$, which we prove to be a fixed point as $N \to \infty$.

$$
\begin{aligned}
\dot h_i(t) &= \frac 1N \sum_{\mu=1}^p \sum_{j=1}^N F_{\mu i} G_{\mu j} \phi(0) \\
&= \frac{\phi(0)}{N} \sum_{\mu=1}^p \sum_{j=1}^N F_{\mu i} G_{\mu j} \\
&= \frac{\phi(0)}{N} \sum_{\mu=1}^p \sum_{j \neq i}^N F_{\mu i} G_{\mu j} + \underbrace{\frac{\phi(0)}{N} \sum_{\mu=1}^p F_{\mu i} G_{\mu i}}_{\bO(\frac 1N)} \\
&= \phi(0) \sum_{\mu=1}^p \frac 1N \sum_{j \neq i}^N F_{\mu i} G_{\mu j} + \bO(\tfrac 1N) \\
&= \phi(0) \sum_{\mu=1}^p F_{\mu i} \frac{N-1}{N} \frac{1}{N-1} \sum_{j \neq i}^N G_{\mu j} + \bO(\tfrac 1N) \\
&\stackrel{N \gg 1}{\approx} \phi(0) \sum_{\mu=1}^p F_{\mu i} \avg{g_\mu(\vec{\xi})}
\end{aligned}
$$

<!-- TODO : fixme. the F_i should not be in the integral -->

To write the last line, we make a slight abuse of notation in order to avoid notation problems when taking the $N \to \infty$ limit of $F_{\mu i}$.

In the case of the network studied in this work, the find that

$$
\avg{g_\mu(\vec{\xi})} = \avg{\tilde \phi(\xi_\mu)} = \avg{\frac{\phi(\xi_\mu) - \avg{\phi(\xi_\mu)}}{\mathrm{Var}[\phi(\xi_\mu)]}} = 0
$$

Thus, $h_i^\star=0$ is a fixed point in the limit $N \to \infty$.

#### Pattern fixed point

Another set of fixed points is $h^\star_i = F_{\nu i}$ for all $\nu = 1, \cdots, p$. We refer to these as the "patterns" in the context of learning the network "rembers" (converges to) these vectors. We verify the fixed point similarly as before.

$$
\begin{aligned}
\dot h_i(t) &= -F_{\nu i} + \frac 1N \sum_{\mu=1}^p \sum_{j=1}^N F_{\mu i} G_{\mu j} \phi(F_{\nu j}) \\
&= -F_{\nu i} + \frac 1N \sum_{\mu=1}^p \sum_{j \neq i}^N F_{\mu i} G_{\mu j} \phi(F_{\nu j}) + \bO(\tfrac 1N) \\
&= -F_{\nu i} + \frac 1N \sum_{j \neq i}^N F_{\nu i} G_{\nu j} \phi(F_{\nu j}) + \frac 1N \sum_{\mu \neq \nu}^p \sum_{j \neq i}^N F_{\mu i} G_{\mu j} \phi(F_{\nu j}) + \bO(\tfrac 1N) \\
&\stackrel{N \gg 1}{\approx} -F_{\nu i} + F_{\nu i} \avg{g_\nu(\vec{\xi_j}) \phi(\xi_{\nu j})} + \sum_{\mu \neq \nu}^p F_{\mu i} \avg{g_\mu(\vec{\xi_j}) \phi(\xi_{\nu j})}
\end{aligned}
$$

The term $\avg{g_\mu(\vec{\xi_j}) \phi(\xi_{\nu j})}$ vanishes in the case of a factorizable distribution and when the functions $g_\mu$ are component-wise.

$$
\begin{aligned}
\avg{g_\mu(\vec{\xi_j}) \phi(\xi_{\nu j})} &= 
\int_{\Rp} g_\mu(\xi_1, \cdots, \xi_p) \phi(\xi_\nu) \rho(\d \vec{\xi}) \\
&= \int_{\Rp} g_\mu(\xi_\mu) \phi(\xi_\nu) \prod_{\gamma=1}^p \rho_\gamma(\d \xi_\gamma) \\
&= \int_{\R} g_\mu(\xi_\mu) \rho_\mu(\d \xi_\mu) \int_{\R} \phi(\xi_\nu) \rho_\nu(\d \xi_\nu) \\
&= \avg{g_\mu(\xi_\mu)} \avg{\phi(\xi_\nu)}
\end{aligned}
$$

As established in the previous section, in the case of the network studied in this work, $\avg{g_\mu(\xi_\mu)} = 0$.

The term $\avg{g_\nu(\vec{\xi_j}) \phi(\xi_{\nu j})}$ is more difficult, but we can prove that in our case it equals one.

$$
\begin{aligned}
\avg{g_\nu(\vec{\xi_j}) \phi(\xi_{\nu j})} &=
\avg{\frac{\phi(\xi_{\mu j}) - \avg{\phi(\xi_{\mu j})}}{\mathrm{Var}[\phi(\xi_{\mu j})]} \phi(\xi_{\mu j})} \\
&= \avg{\frac{\phi(\xi)^2 - \avg{\phi(\xi)} \phi(\xi) }{\mathrm{Var}[\phi(\xi)]}} \\
&= \frac{\avg{\phi(\xi)^2} - \avg{\avg{\phi(\xi)} \phi(\xi)} }{\mathrm{Var}[\phi(\xi)]} \\
&= \frac{\avg{\phi(\xi)^2} - \avg{\phi(\xi)}^2}{\mathrm{Var}[\phi(\xi)]} \\
&= \frac{\mathrm{Var}[\phi(\xi)]}{\mathrm{Var}[\phi(\xi)]} \\
&= 1
\end{aligned}
$$

Therefore, the set of $h_i^\star = F_{\nu i}$ for all $\nu = 1, \cdots, p$ are fixed points in the limit of large $N$, under the assumptions that the distribution is factorizable, the functions $g_\mu$ are component-wise, and $\avg{g_\nu(\xi) \phi(\xi)}=1$, which is the case in the setup of this work.

### Fixed points of the neural field

The derivation of the fixed points for the neural field equation follows the same arguments as the fixed points in the network of neurons.

As $N \to \infty$, the fixed points $h_i^\star$ become functions $h^\star(\xi_1, \cdots, \xi_p)$ on the $\Rp$ space :

* $h_i^\star = 0 \to h^\star(\xi_1, \cdots, \xi_p) = 0$
* $h_i^\star = F_{\nu i} = \xi_{\nu i} \to h^\star(\xi_1, \cdots, \xi_p) = \xi_\nu$ for all $\nu = 1, \cdots, p$

<!-- TODO : add derivation for neural field ? -->

### Stability of the fixed points

<!-- TODO -->

## Behavior of $\kappa_\mu(t)$ in the cyclic network

kappa in the delayed, rolled and cyclic network

relation to overlaps

## "Mean patterns" in the binned connectivity

derivation of $\Jab$ and mean patterns

For finite number of recursive quadrant iterations $n$, we can do a "mean-field approximation" inside each of the $4^n$ segments. Let $\alpha = \{i_1,\cdots,i_{|\alpha|}\}$ be the multi-index corresponding to all neurons of which the embedding in $\mathbb R^p$ gets mapped to the segment $\alpha$ in $[0,1]$. Let $H_\alpha(t) = \frac 1 {|\alpha|} \sum_{i \in \alpha} h_i(t)$ be the (mean) RNN potential of the segment $\alpha$. The connectivity matrix $\tilde J_{\alpha,\beta}$ satisfies

$$
\dot H_\alpha(t) = -H_\alpha(t) + \sum_{\beta \in \text{segments of length } 4^{-n}} \tilde J_{\alpha,\beta} \phi(H_\beta(t))
$$

By substituting the original $h_i(t)$, we find the correct rescaling is given by

$$
\tilde J_{\alpha,\beta} = \frac 1 {|\alpha|} \sum_{i \in \alpha} \sum_{j \in \beta} J_{ij}
$$

In the low-rank case, we have

$$
J_{ij} = \frac 1N \sum_{\mu=1}^p F_{\mu,i} G_{\mu,j}
$$

We can define the "mean pattern" inside each bin :

$$
\tilde F_{\mu,\alpha} = \frac{1}{|\alpha|} \sum_{i \in \alpha} F_{\mu,i}, \; \tilde G_{\mu,\alpha} = \frac{1}{|\alpha|} \sum_{i \in \alpha} G_{\mu,i}
$$

Then the connectivity matrix is given by, noting $N = \sum_{\alpha} |\alpha|$,

$$
\tilde J_{\alpha,\beta} = \frac{|\beta|}{\sum_{\beta'} |\beta'|} \left( \sum_{\mu=1}^p \tilde F_{\mu,\alpha} \tilde G_{\mu,\beta} - \delta_{\alpha,\beta} \underbrace{\sum_{\mu=1}^p \sum_{i \in \alpha} \frac{F_{\mu,i}}{|\alpha|} \frac{G_{\mu,i}}{|\alpha|}}_{\gamma_{\alpha}} \right)
$$

Recurrent current is given by

$$
I^\text{rec}_\alpha(t) = \sum_\beta \tilde J_{\alpha,\beta} \phi(H_\beta(t)) = \frac{1}{\sum_{\beta'} |\beta'|} \left( \sum_\beta |\beta| \sum_{\mu=1}^p \tilde F_{\mu,\alpha} \tilde G_{\mu,\beta} \phi(H_\beta(t)) - |\alpha| \tilde \gamma_\alpha \phi(H_\alpha(t)) \right)
$$

The overlaps can be computed as

$$
\tilde m_\mu(t) = \frac{1}{\sum_{\alpha} |\alpha|} \sum_\alpha |\alpha| G_{\mu,\alpha} \phi(H_\alpha(t))
$$

### 2023-06-04 manuscript/chapters/2_neural_fields.qmd
this takes so much time nyaaa

@@
 # A simple model for a neural field in 2D
+
+::: {.content-hidden}
+{{< include ../macros.tex >}}
+:::
 
@@
 
-no external current,
+### Case study : low-rank networks of neurons
 
-general low-rank neural fields, F and G notation
+start with the study of network of neurons, no external current
+
+[TODO MATH : show self-connections $\sim \frac 1N$, so we ignore them]
+
+general setting of low-rank networks
 
 $$
-J_{ij} = \sum_{\mu=1}^p F_{\mu,i} G_{\mu,j}
+\begin{aligned}
+\dot h_i(t) = -h_i(t) + \frac 1N \sum_{\mu=1}^p \sum_{j=1}^N F_{\mu i} G_{\mu j} \phi(h_j(t)) \\
+F_{\mu i} = f_\mu(z_{1 i}, \cdots, z_{p i}),\, G_{\mu i} = g_\mu(z_{1 i}, \cdots, z_{p i}),\, \vec{z_i} \stackrel{\text{iid}}{\sim} \rho
+\end{aligned}
 $$
 
-show how the embedding becomes smooth
+interpretation : each neuron has a position $\vec{z_i} \in \Rp$. $J_{ij} = \frac 1N \sum_{\mu=1}^p = F_{\mu i} G_{\mu j}$ is the weight of recurrent currents.
+
+our setting is :
+
+$$
+\begin{aligned}
+f_\mu(z_1, \cdots, z_p) = z_\mu \\
+g_\mu(z_1, \cdots, z_p) = \tilde \phi(z_\mu) = \frac{\phi(z_\mu) - \avg{\phi(z_\mu)}}{\mathrm{Var}[\phi(z_\mu)]} \\
+\phi(z) = \frac{1}{1+\mathrm{e}^{-z}} \\
+\vec{z} = (z_1, \cdots, z_p) \sim \rho(z_1, \cdots, z_p) = \prod_{\mu=1}^p \mathcal{N}(z_\mu),\, \mathcal{N}(z) = \frac{1}{\sqrt{2 \pi}} \mathrm{e}^{-\frac 12 z^2} \\
+\end{aligned}
+$$
 
@@
 
-- $h_i = \xi_{\mu,i} \quad \forall \mu = 1,\cdots,p$ (stable)
-- $h_i=0$ (unstable)
+* $h_i = z_{\mu,i} \quad \forall \mu = 1,\cdots,p$ (stable)
+* $h_i=0$ (unstable)
 
@@
 
-## Building up to a neural field in $\mathbb R^2$
+talk about since $\lambda_\nu = \avg{g^\prime_\nu(z_\nu) f_\nu(z_\nu)}-1$, the sharper the gain, the more stable to fixed points (since the bump of $\delta \phi$ becomes more concentrated)
+
+[TODO FIG : draw a flow field of in the latent space]
+[TODO FIG : numerical eigenvalues plot]
+
+### Emergence of structure in the networks of neurons
+
+show simulations, first in $p=1$, then $p=2$.
+
+![simulation in 1D, p=1](figures/embedding_d=1_h0=0_phi=sigmoid.mp4)
+
+![simulation in 2D, p=2](figures/embedding_d=2_h0=0.mp4)
+
+[TODO FIG : scatterplot of the network, with animation]
+
+show it converges to the fixed points $z_1$ and $z_2$
+
+the structure of the scatterplot emerges naturally as a smooth manifold in $\R^2$
+
+### Direct simulation of the low-rank neural field
 
@@
 \quad
-w(\vec z, \vec y) = \sum_{\mu=1}^p \tilde \phi (y_\mu) z_\mu \\
-&= -h(t, \vec z) + \sum_{\mu=1}^p z_\mu m_\mu(t)
+w(\vec z, \vec y) = \sum_{\mu=1}^p \tilde \phi (y_\mu) z_\mu
 \end{aligned}
@@
 
-study of fixed points (and convergence of the fixed points)
+repeat the fixed points are now the generalization, functions $h=z_1$ and $h=z_2$
 
 simulation strategies : grid and sampling
+
+[TODO : show the connectivity with grid sampling. mention that the problems with the compact support get adressed later in the next chapter, when we do $\R^2 \mapsto [0,1]^2$]
 
@@
 
+note that :
+
+$$
+\partial_t h(t, \vec z) = -h(t, \vec z) + \sum_{\mu=1}^p z_\mu m_\mu(t)
+$$
+
+[TODO : i'm not even sure that introducing the overlaps in necessary. we might just include them for completeness, and because they are relevant in papers on low-rank stuff]
+
 ### Low-rank neural fields as $p$-dimensional closed systems
 
-introduce overlap, and in our case the low-rank model the kappa
+introduce kappa, and in our case the low-rank model the kappa
 
-overlaps as an approximation of the “dynamics” (latent space), in the case phi=linear
-
-The dynamics of $h(t, \vec z)$ are in a subsystem of dimension $p$ spanned by the ONB of functions $\{e_\mu(\vec z) = z_\mu | \mu=1,\cdots,p\}$, with the scalar product $\langle f, g \rangle = \int_{\mathbb R^p} f(\vec y) g(\vec y) \rho(\mathrm d \vec y)$.
-
-We decompose $h(t, \vec z) = h^\perp(t, \vec z) + \sum_{\mu=1}^p \kappa_\mu(t) z_\mu$, and write the system of equations for $\mu=1,\cdots,p$.
+we define
 
 $$
-\dot \kappa_\mu(t) = -\kappa_\mu(t) + \int_{\mathbb{R}^p} \tilde\phi(y_\mu) \phi(h(t, \vec y)) \rho(\mathrm d \vec y) = -\kappa_\mu(t) + m_\mu(t)
+\kappa_\mu(t) = \avg{f_\mu, h(t, \cdot)} = \int_{\Rp} y_\mu h(0, \vec y) \rho(\mathrm d \vec y)
 $$
 
-with initial conditions
-
+with these definition, we have in our case
 $$
-\kappa_\mu(0) = \int_{\mathbb{R}^p} y_\mu h(0, \vec y) \rho(\mathrm d \vec y)
+\begin{aligned}
+\dot \kappa_\mu(t) &= -\kappa_\mu(t) + \int_{\Rp} \tilde\phi(y_\mu) \phi(h(t, \vec y)) \rho(\mathrm d \vec y) = -\kappa_\mu(t) + m_\mu(t) \\
+\kappa_\mu(0) &= \int_{\mathbb{R}^p} y_\mu h(0, \vec y) \rho(\mathrm d \vec y)
+\end{aligned}
 $$
 
-and the orthogonal component evolves according to
+overlaps as an approximation of the “dynamics” (latent space), in the case $\phi=\mathrm{linear}$.
+since in our case $\phi$ is a sigmoid, it can be considered linear in the range $h \sim \pm 3$.
+we therefore use the computed overlaps as to show the latent space
 
-$$
-h^\perp(t, \vec z) = h^\perp(0, \vec z) \mathrm e^{-t},
-\quad
-h^\perp(0, \vec z) = h(0, \vec z) - \sum_{\mu=1}^p \kappa_\mu(0) z_\mu
-$$
+[TODO MATH : actually we could also juste compute the kappa as a function of time, and study that]
 
@@
 
+we introduce $\delta$ the delay, and rolling $\mu+1$, with convention $p+1=1$
+
+$$
+\partial_t h(t, \vec z) = -h(t, \vec z) + \int_{\mathbb{R}^p} \sum_{\mu=1}^p \tilde \phi (y_\mu) z_{\mu+1} \phi(h(t - \delta, \vec y)) \rho(\mathrm d \vec y)
+$$
+
+note this is now a DDE instead of a PDE
+
 1. delayed neural field
@@
 
-show what the analytical kappa is
+[TODO : is this useful ? I derived this because it shows we need both delay and rolling, but might be too much]
 
-introduce both, derive kappa again
+we can show that in this case
+
+$$
+\dot \kappa_\mu(t) = -\kappa_\mu(t) + m_{\mu+1}(t - \delta)
+$$
+
+then in the case that $\phi=linear \implies \kappa_\mu = m_\mu$, and initial conditions $m_\mu(t<0) = m_\mu(0)$ (constant inititial condition), $m_1(0) = 1$ and $m_2(0) = 0$, we can solve the first "cycle" ($0 < t < \delta$) analytically :
+
+$$
+\begin{aligned}
+\kappa_1(t) &= \mathrm{e}^{-t} \\
+\kappa_2(t) &= 1 - \mathrm{e}^{-t} \\
+\end{aligned}
+$$
+
+then one can repeat this for the next cycles. basically each cycle is applying an exponential convolution filter
+
+[TODO MATH : the derivation of this, we have an exponential convolution filter]
+
+![](figures/kappa_cycling_p=2.tmp.png)
 
@@
 
-recurrent current picture
+[TODO : might be good to show a simulation where we explicitly simulate $\kappa$ ? but the goal here is just to "prove" the cycling behavior]
+
+![](figures/embedding_d=2_h0=0_cycling_delta=6.mp4)
+
+![Intuition for the cycling behavior by plotting the recurrent currents](figures/cycling_recurrent.png)
+
+## From $\R^2$ to $[0,1]^2$
+
+Defining the change of variables $u_\mu=\mathrm{CDF}(y_\mu), v_\mu=\mathrm{CDF}(z_\mu)$, the neural field becomes
+
+$$
+\begin{aligned}
+& \partial_t h_U(t, \vec v) = -h_U(t, \vec v) + \int_{[0,1]^p} w_U(\vec v, \vec u) \phi(h_U(t, \vec u)) \mathrm d \vec u \\
+& \text{where } w_U(\vec v, \vec u) = w(\mathrm{CDF}^{-1}(\vec v), \mathrm{CDF}^{-1}(\vec u)), \, h_U(t, \vec v) = h(t, \mathrm{CDF}^{-1}(\vec v))
+\end{aligned}
+$$
+
+![normal distribution CDF](figures/1920px-Normal_Distribution_CDF.svg.png)
+
+We now have a neural field equation on a unit (hyper)cube, and uniform sampling.
+
+the problem of sampling all of $\R^2$ is now "hidden away" in the CDF, since it goes to $\pm \infty$ at 0 and 1.
+
+note : here we use the fact that the components are independant. therefore we can define the total as the product, i.e.
+
+$$
+\mathrm{CDF}(z_1, z_2) = \int_{-\infty}^{z_1} \rho_1(d y_1) \int_{-\infty}^{z_2} \rho_2(\d y_2) = v_1 v_2
+$$
+
+we would not be able to do this is the coordinates were coupled, then we could define still define the CDF, but we cannot factorize it into 2 different components
+
+![](figures/sampling_R2_01.png)
+
+[TODO : does this deserve its own section ? maybe mention it later, we can give a picture of the "isolines" of the 2D gaussian CDF]
+
+## Numerical simulation of neural fields
+
+### Direct sampling
+
+we simulate the network of neurons
+
+### Bin averaging samples
+
+![](figures/sampling_R2_nogrid_kernel.png)
+
+### Grid over $[0,1]^2$
+
+We define $\vec z(\alpha) = S(\alpha)$ the 2D point corresponding to the mapping $\alpha$.
+(details : we need a bounding box to map back).
+Defining a matrix $Z_{\mu,\alpha} = \vec z(\alpha)_\mu$, we can write down a numerical PDF $\tilde \rho(Z_{1,\alpha},\cdots,Z_{p,\alpha}) = \tilde \rho(Z_{:,\alpha}) = \frac{\rho(Z_{:,\alpha})}{\sum_\beta \rho(Z_{:,\beta})}$ and the following patterns to simulate the embedded $[0,1]$ neural field as a low-rank RNN.
+
+$$
+\tilde F_{\mu,\alpha} = Z_{\mu,\alpha}, \quad \tilde G_{\mu,\alpha} = \tilde\phi(Z_{\mu,\alpha}), \quad \tilde J_{\alpha,\beta}=\tilde \rho(Z_{:,\beta}) \sum_{\mu=1}^p \tilde F_{\mu,\alpha} \tilde G_{\mu,\beta}
+$$
+
+Doing this is equivalent to the formulation of the binned connectivity matrix.
+
+![](figures/sampling_R2_01_kernel.png)

### 2023-06-04 manuscript/chapters/3_mappings.qmd
this takes so much time nyaaa

@@
 # Mapping the neural field in 2D to a 1D embedding
+
+::: {.content-hidden}
+{{< include ../macros.tex >}}
+:::
 
@@
 
+[TODO MOTIVATION : is this really useful ? the point i'm trying to make here is the embedding problem arises when you have an infinite number]
+
 the case of population dynamics can easily be embedded in 1D
+
+$$
+\begin{aligned}
+\rho(\vec z) &= \sum_{a \in P} p_{a} \delta(\vec{z_a} - \vec{z}) \\
+\text{normalization : } \int_{\Rp} \rho(\d \vec{z}) &= \sum_{a} p_a = 1
+\end{aligned}
+$$
+
+we can explicitly formulate the embedding with a new 1D PDF (arrange the populations on the line, their probability is their segment length)
+
+$$
+\gamma(\alpha) = \sum_{a \in P} p_a \mathbb{1}(C_{a-1} \leq \alpha < C_{a}),\, C_a=\sum_{b=1}^a p_b
+$$
+
+the connectivity matrix in the 1D model becomes
+
+$$
+J_{ab} = \sum_{\mu=1}^p f_\mu(\vec{z_a}) g_\mu(\vec{z_b})
+$$
+
+[TODO MATH : flesh this out]
+
+case 2 populations, we get the ising model (binary patterns)
+
+the point i try to make is that for discrete distributions (i can enumerate the populations), it is easy to embed in 1D. the problem comes when the distribution is continuous.
+
+this also helps give an intuition that effectively, when we do the binning in $[0,1]^2$, each bin represents a population
 
@@
 
-Define a measurable mapping $S : \mathbb{R}^p \rightarrow [0, 1]$.
+Define a measurable and bijective mapping $S : \mathbb{R}^p \rightarrow [0, 1]$.
 
-Let $\mu : [0,1] \rightarrow [0, \infty]$ be a measure on $([0,1], \mathcal B([0,1]))$.
+Let $\lambda : [0,1]^2 \rightarrow [0, \infty]$ be a measure.
 
-Then we can write
+We define $\tilde h(t, \alpha) = h_U(t, S^{-1}(\alpha))$, $\tilde w(\alpha, \beta) = w_U(S^{-1}(\alpha), S^{-1}(\beta))$
 
 $$
-\partial_t h(t, \vec z) = -h(t, \vec z) + \int_{[0,1]} [w(\vec z, \cdot) \phi(h(t, \cdot)) \rho(\cdot)] \circ S^{-1} \; \mathrm d \mu
+\partial_t \tilde h(t, \alpha) = -\tilde h(t, \alpha) + \int_{[0,1]} \tilde w(\alpha, \beta) \tilde\phi(h(t, \beta)) \lambda \circ S^{-1}(\d \beta)
 $$
 
-## From $\R^2$ to $[0,1]^2$
+![](figures/mappings_h1d_state.png)
 
-Defining the change of variables $u_\mu=\mathrm{CDF}(y_\mu), v_\mu=\mathrm{CDF}(z_\mu)$, the neural field becomes
-
-$$
-\partial_t h_U(t, \vec v) = -h_U(t, \vec v) + \int_{[0,1]^p} w_U(\vec v, \vec u) \phi(h_U(t, \vec u)) \mathrm d \vec u, \quad
-w_U(\vec v, \vec u) = w(\mathrm{CDF}^{-1}(\vec v), \mathrm{CDF}^{-1}(\vec u)), \quad h_U(t, \vec v) = h(t, \mathrm{CDF}^{-1}(\vec v))
-$$
-
-We now have a neural field equation on a unit (hyper)cube, and uniform sampling.
+![](figures/mapping_recursive_quadrant.mp4)
 
@@
 
-linear projections don’t work ofc
+linear projections don’t work ofc.
+
+Note : why are not using the CDF to embed in 1D ? because it's not bijective !
 
@@
 $$
+
+![](figures/mapping_demo_indices.png)
+
+[TODO : do we mention the random mapping ?]
 
@@
 $$
+
+note that in other definitions (and in code) it is defined by $0.b^y_1 b^x_1 b^y_2 b^x_2 \cdots$. This difference is not fundamentally important though.
 
@@
 
-argue over binary expansion
+argue over binary expansion, and especially how the fact that
+
+* it converges forces the least significant bits of $z_1$ and $z_2$ to go to the lsb of $\alpha$
+* it's a bijection means "both bits of x and y" need to be used in the expansion
 
@@
 
-how to we actually simulate the mappings in one-dimension
+go into detail on how numerically S is a reordering of the bins
 
-how to we actually simulate the mappings in one-dimension
+what we are doing in reality is assigning indices to squares of size $2^{-n} \times 2^{-n}$, and mapping them to indices representing segments of length $4^{-n}$
 
-go into detail on how numerically S is a reordering of the bins
+[TODO FIG : drawing of the square-to-segment mapping, something like
+
+![Alt text](figures/Screenshot_2023-06-03_19-59-49.png)]
 
@@
 
-## Reinterpretation of simulation strategies
-
-### Bin averaging samples
-
-give an intuition, show the figure and some simulations to show that it works
-
-### Direct remapping of the grid
-
-We define $\vec z(\alpha) = S(\alpha)$ the 2D point corresponding to the mapping $\alpha$.
-(details : we need a bounding box to map back).
-Defining a matrix $Z_{\mu,\alpha} = \vec z(\alpha)_\mu$, we can write down a numerical PDF $\tilde \rho(Z_{1,\alpha},\cdots,Z_{p,\alpha}) = \tilde \rho(Z_{:,\alpha}) = \frac{\rho(Z_{:,\alpha})}{\sum_\beta \rho(Z_{:,\beta})}$ and the following patterns to simulate the embedded $[0,1]$ neural field as a low-rank RNN.
-
-$$
-\tilde F_{\mu,\alpha} = Z_{\mu,\alpha}, \quad \tilde G_{\mu,\alpha} = \tilde\phi(Z_{\mu,\alpha}), \quad \tilde J_{\alpha,\beta}=\tilde \rho(Z_{:,\beta}) \sum_{\mu=1}^p \tilde F_{\mu,\alpha} \tilde G_{\mu,\beta}
-$$
-
-Doing this is equivalent to the formulation of the binned connectivity matrix.
-
-## Numerical simulations of the neural field in $[0,1]$
+## Coarse-graining the neural field in $[0,1]$
 
@@
 
-averaging of 2ˆn neurons
+In order to simulate the neural field numerically, we discretize by taking a grid of $4^n$ bins in the $[0,1]^2$ space.
 
-see sampling.ipynb
+$$
+\int_{[0,1]^2} w_U(\vec v, \vec u) \phi(h_U(t, \vec u)) \mathrm d \vec u \rightarrow \sum_{j=1}^N \underbrace{w_U(\vec u_i, \vec u_j)}_{J_{ij}} \phi(\underbrace{h_U(t, \vec u_j)}_{h_j(t)})
+$$
+
+Each bin therefore corresponds to a population (density) indexed by $i = \{1\cdots4^n\}$. Population $i$ interacts with other populations through the (numerically finite) connectivity $J_{ij}$.
+
+To simulate the neural field in $[0,1]$, we use the mapping $S(\vec u_i) \stackrel{\text{notation}}= S(i) = \alpha_i$ to embed the populations in 1D.
+
+One trivial mapping would be $S(i) = i / 4^n$, which is just ordering the populations in [0,1] in the same order they appear inside the bins.
+More elaborate mappings are simply *permutations* of the numerically binned populations set $\{1\cdots4^n\}$, and the resulting connectivity matrix is then given by $J_{S(i),S(j)}$ [^1].
+
+However, the connectivity matrix is invariant to permutations of neurons, and without further modification, the dynamics emerging from $J_{S(i),S(j)}$ are the same as the ones emerging from $J_{ij}$.
+
+In the same way that discretizing the smooth 2D field works *because of* the smoothness, we discretize the 1D field emerging from the mapping.
+If the mapping has "nice properties" (e.g.
+"locality"), this discretization should work in a similar manner.
+
+For this, we consider the $4^n$ mapped populations $\alpha_i$ to be the "mathematically exact" 1D field, and we "discretize" the 1D field by taking a grid of $2^n$ bins in $[0,1]$.
+In effect, we average bins of $2^n$ populations, in order to obtain $4^n/2^n = 2^n$ coarse-grained populations in $[0,1]$.
+We call this "downsampling" [^2].
+
+The main question is now, what conditions must be imposed on the mapping $S : [0,1]^2 \rightarrow [0,1]$ (numerically : a permutation), such that it is *resilient* to this downsampling, in other words that the dynamics obtained from the connectivity in 1D are the same as the dynamics in 2D [^3].
+
+In the following, this assertion is made by comparing the trajectories $(m_0(t),m_1(t))$ in phase space (also called latent space).
+If the mapping is "resilient", small differences between the trajectories vanish as $n \rightarrow \infty$ [^4].
+
+[^1] : I am glossing over the details of how we bin the grid of $4^n$ elements in $[0,1]^2$ space.
+Numerically, there is an order in which we enumerate each of the $4^n$ bins (the most obvious is column-wise enumeration, i.e.
+``ReshapeMapping.inverse_samples()``).
+As described, this ordering is abitrary, and one should always consider the enumeration of the bins to be in random order.
+
+[^2] : The reason for which we take bins of size $2^n$, is because $2^n$ is the maximum size a bin can have which spans in only one direction in the 2D embedding.
+For instance, with the ``ReshapeMapping`` a column contains $2^n$ populations, all with the same $\xi_0$ position : visually the bin is vertical.
+With the ``DiagonalMapping``, the largest sequence of populations which "go in the same direction" (TODO : clarify this) is of size $2^n$, and spans the diagonal (1,0)->(0,1).
+Intuitively, this averaging can be seen as "destroying fast and large oscillations".
+
+[^3] : The connectivity $\tilde J_{\alpha\beta}$ in 1D can be obtained by computing the "average low-rank vectors". See ``draft.md`` for more details.
+
+[^4] : Each point in $(m_0,m_1)$ fully encodes the state of a linear low-rank RNN.
+We can prove (see my own handwritten notes) that $m_\mu(t) \approx \kappa_\mu(t)$ when $\phi(h)$ is close to linear.
+In our case, $\phi(h)$ is a sigmoid (I found a sigmoid has better convergence properties than a linear function), and can be considered linear as long as $-3 \lesssim h(t) \lesssim 3$.
+
+![](figures/coarsegraining.png)

### 2023-06-04 manuscript/chapters/4_locality.qmd
this takes so much time nyaaa

@@
+::: {.content-hidden}
+{{< include ../macros.tex >}}
+:::
+
 # Locality and numerical convergence
@@
 
+TODO : write the proof of this
+
 ## Locality
+
+### Intuition behind the notion of locality
+
+![](figures/distance_2d_1d.png)
+
+localité :
+1. proche in 2D -> proche in 1D (OK projection (obv., x est proche), OK zorder, PAS random)
+2. proche in 1D -> proche in 2D (PAS projection, PAS random, OK zorder)
+
+[TODO EXPLAIN : explain this better. basically the thing is we AVERAGE in 1D, so this better be meaningful]
 
@@
 
-“resilient to some discontinuity”
-the notion of infinite number of discontinuities, but the size counts too
+"resilient to some discontinuity" the notion of infinite number of discontinuities, but the size counts too
 
-Total variation vs bining variation
+![histogram of discontinuities](figures/image_20230517_2.png)
 
-histogram of discontinuities ?
+### The total variation attempt
+
+Let $S^{-1} : \mathbb R \rightarrow \mathbb R^2$, let $|| \cdot ||$ be a norm in $\mathbb R^2$ (any norm is OK, since all equivalent (i.e.
+equal to a constant)).
+Here we use the $\ell^1$ norm (absolute value).
 
 $$
-\frac{1}{2^n}\sum_{i=1}^{2^n} \sup_{z,z' \in \{\frac{k}{4^n} | k= 1, \cdots, 2^n\}} |f^n_1(\tfrac{i-1}{2^n}+z) - f^n_1(\tfrac{i-1}{2^n}+z')| + |f^n_2(\tfrac{i-1}{2^n}+z) - f^n_2(\tfrac{i-1}{2^n}+z')|
+\begin{aligned}
+\mathrm{TV}(S^{-1}) &= \sup_{P \in \mathcal P} \sum_{i=1}^{n_P} || S^{-1}(z_{i}) - S^{-1}(z_{i-1}) ||_1 \\
+&= \sup_{P \in \mathcal P} \sum_{i=1}^{n_P} | S^{-1}_1(z_i) - S^{-1}_1(z_{i-1})| + | f_2(z_i) - f_2(z_{i-1}) |
+\end{aligned}
 $$
 
-plot the bins in the 2D scatter, so we can see what we are doing.
+this basically just projects the discontinuities on x and y.
 
-it is different to TV, TV is projecting all the jumps
+in a way since you need to go to each point only once, so there is only so much you can do inside the $[0,1]^2$. the tiny jumps still add up, and all mappings seem to grow $TV$ in the same way
+
+![](figures/image_20230517.png)
+
+-> what we really are expression, is the fact that doing the averaging in 1D is not a big problem when the mapping is good.
+
+### Mathematical formulation of average binned variation
+
+variation inside the bins, weak version
+
+$$
+\begin{aligned}
+V_n(S^{-1}) &= \frac{1}{2^n}\sum_{i=1}^{2^n} \sup_{\delta,\delta' \in [0, \tfrac{1}{2^n}]^2} |S_1(\tfrac{i-1}{2^n}+\delta) - S_1(\tfrac{i-1}{2^n}+\delta')| + |S_2(\tfrac{i-1}{2^n}+\delta) - S_2(\tfrac{i-1}{2^n}+\delta')| \\
+&= \frac{1}{2^n}\sum_{i=1}^{2^n} \sup_{\alpha, \alpha^\prime \in [\tfrac{i-1}{2^n}, \tfrac{i}{2^n}]^2} \norm{S^{-1}(\alpha^\prime) - S^{-1}(\alpha)}_1
+\end{aligned}
+$$
+
+variation inside the bins, strong version.
+basically we take the max over all binning methods that have bins smaller than $\delta$.
+Let $P_\delta$ be a permutation of $[0,1]$ with $n_{P_\delta}$ elements $\{\alpha_1 \leq \alpha_2 \leq \cdots \alpha_{n_{P_\delta}}\}$, where $\alpha_{i}-\alpha_{i-1} \leq \delta$, and $\mathcal{P}_\delta$ the set of such permutations.
+
+$$
+\begin{aligned}
+V_\delta(S^{-1}) &= \sup_{P_\delta \in \mathcal{P}_\delta} \delta \sum_{i=1}^{n_{P_\delta}} \sup_{\alpha, \alpha^\prime \in [\alpha_{i-1}, \alpha_{i}]^2} \norm{S^{-1}(\alpha^\prime) - S^{-1}(\alpha)}_1 \\
+V &= \limsup_{\delta \to 0} V_\delta
+\end{aligned}
+$$
+
+evidently,
+
+$$
+0 \leq V_n \leq V_\delta
+$$
+
+[TODO FIG : show the range of each bin, how each bin contains $2^n$ smaller bins. plot the bins in the 2D scatter, so we can see what we are doing]
+
+[TODO MATH : the exact norm is not important. all norms are equivalent]
+
+[TODO NOTE : in the definition of $V$, we use the converged version of $S$, instead of a finite truncation $S^n$.
+we need to convince people this is not a big mistake, since $S^n \to S$, and we can take both limits (binning and $S^n$) "at the same time"]
+
+### Average variation of mappings
+
+Numerical demonstration
+
+![](figures/v_num.png)
+
+In order to compute (a bound on) $V_n$ for some mappings, we consider the binary expansion of $\alpha$ and $\alpha^\prime$ which appear in the supremum.
+For clarity and to represent actual numerical implementations, we furthermore truncate the binary expansion to $2m$ bits (on most machines, $2m=64$ bits).
+
+$$
+\begin{aligned}
+\alpha &= 0. b_1 b_2 \cdots b_{2m-1} b_{2m}\\
+\alpha^\prime &= 0. b_1^\prime b_2^\prime \cdots b_{2m-1}^\prime b_{2m}^\prime \\
+\end{aligned}
+$$
+
+Furthermore, since $\alpha \in \left[\frac{i-1}{2^n}, \frac{i}{2^n} \right],\, i \in \{1, \cdots, 2^n\}$, we may express the first $n$ (where $n < 2m$) bits of $\alpha$ with the bits of $i=\sum_{k=0}^{n-1} i_k 2^k$ (we ignore $i=2^n$ for the sake of simplicity, which would introduce one more bit, $n+1$ in total).
+
+$$
+\begin{aligned}
+\alpha &= 0. i_{n-1} i_{n-2} \cdots i_1 i_0 b_n b_{n+1} \cdots b_{2m-1} b_{2m}\\
+\alpha^\prime &= 0. i_{n-1} i_{n-2} \cdots i_1 i_0 b^\prime_n b^\prime_{n+1} \cdots b^\prime_{2m-1} b^\prime_{2m}
+\end{aligned}
+$$
+
+#### Z-mapping
+
+Let us consider the Z-mapping, and the two terms compositing its inverse. In the case of $\alpha, \alpha^\prime$ above, we have if $n$ is even :
+
+$$
+\begin{aligned}
+Z^{-1}_1(\alpha) &= 0.\underbrace{i_{n-1} i_{n-3}}_\text{$n/2$ bits} \cdots i_1 b_n b_{n+2} \cdots b_{2m-1} \\
+Z^{-1}_2(\alpha) &= 0.\underbrace{i_{n-2} i_{n-4}}_\text{$n/2$ bits} \cdots i_0 b_{n+1} b_{n+3} \cdots b_{2m}
+\end{aligned}
+$$
+
+If $n$ is odd, we have :
+
+$$
+\begin{aligned}
+Z^{-1}_1(\alpha) &= 0.\underbrace{i_{n-1} i_{n-3}}_\text{$(n+1)/2$ bits} \cdots i_0 b_n b_{n+2} \cdots b_{2m-1} \\
+Z^{-1}_2(\alpha) &= 0.\underbrace{i_{n-2} i_{n-4}}_\text{$(n-1)/2$ bits} \cdots i_1 b_{n+1} b_{n+3} \cdots b_{2m}
+\end{aligned}
+$$
+
+We only consider even $n$ in the following, as the proof remains very similar.
+
+The core of the proof relies on the fact the first $n/2$ bits of the difference of the inverses cancel out when computing the average binned variation.
+
+$$
+\begin{aligned}
+|Z^{-1}_1(\alpha) - Z^{-1}_1(\alpha^\prime)| &= 0.0 \cdots 0 (b_n b_{n+2} \cdots b_{2m-1} - b^\prime_n b^\prime_{n+2} \cdots b^\prime_{2m-1}) \\
+&= 2^{-n/2} \sum_{k=1}^{\tfrac{2m-n+1}{2}} (b_{n+2k-1} - b^\prime_{n+2k-1}) 2^{-k} \\
+&\leq 2^{-n/2} \\
+|Z^{-1}_2(\alpha) - Z^{-1}_2(\alpha^\prime)| &= 2^{-n/2} \sum_{k=1}^{\tfrac{2m-n+1}{2}} (b_{n+2k} - b^\prime_{n+2k}) 2^{-k} \\
+&\leq 2^{-n/2}
+\end{aligned}
+$$
+
+Putting everything together, we get an upper bound for the average binned variation.
+
+$$
+V_n(Z^{-1}) \leq \frac{1}{2^n} \sum_{i=1}^{2^n} 2^{-n/2} + 2^{-n/2} = 2^{-n/2+1} \xrightarrow{n \to \infty} 0
+$$
+
+#### Column mapping
+
+We now consider the column mapping. Since here we consider finite-precision $\alpha$ and $\alpha^\prime$, the invere is well-defined and we have
+
+$$
+\begin{aligned}
+C^{-1}_1(\alpha) &= 0.i_{n-1} i_{n-2} \cdots i_1 i_0 b_n b_{n+1} \cdots b_m \\
+C^{-1}_2(\alpha) &= 0.b_{m+1} b_{m+2} \cdots b_{2m}
+\end{aligned}
+$$
+
+The first component can be bounded using the same trick from the previous proof :
+
+$$
+\begin{aligned}
+|C^{-1}_1(\alpha) - C^{-1}_1(\alpha^\prime)| &= 2^{-n} \sum_{k=1}^{m-n+1} (b_{n+k-1} - b^\prime_{n+k-1}) 2^{-k} \\
+&\leq 2^{-n}
+\end{aligned}
+$$
+
+However now problems arise when we compute the averge binned variation, since the second component is not well-behaved.
+In particular, since we take the $\sup_{\alpha,\alpha^\prime \in [\tfrac{i-1}{2^n},\tfrac{i}{2^n}]^2}$, we can pick the specific values such that $b_k=1$ and $b^\prime_k=0$ for all $k \in \{m+1, \cdots 2m\}$.
+
+$$
+\begin{aligned}
+|C^{-1}_2(\alpha) - C^{-1}_2(\alpha^\prime)| &= \sum_{k=m+1}^{2m} (b_k - b^\prime_k) 2^{-(k-m)} \\
+&= \sum_{k=m+1}^{2m} 2^{-(k-m)} \\
+&> \frac 12
+\end{aligned}
+$$
+
+Putting both together, we can write :
+
+$$
+\begin{aligned}
+V_n(C^{-1}) &= \frac{1}{2^n}\sum_{i=1}^{2^n} \sup_{\alpha, \alpha^\prime \in [\tfrac{i-1}{2^n}, \tfrac{i}{2^n}]^2} |C^{-1}_1(\alpha^\prime) - C_1^{-1}(\alpha)| + |C^{-1}_2(\alpha^\prime) - C_2^{-1}(\alpha)| \\
+&\geq \frac{1}{2^n}\sum_{i=1}^{2^n} \sup_{\alpha, \alpha^\prime \in [\tfrac{i-1}{2^n}, \tfrac{i}{2^n}]^2} |C^{-1}_1(\alpha^\prime) - C_1^{-1}(\alpha)| + \sup_{\alpha, \alpha^\prime \in [\tfrac{i-1}{2^n}, \tfrac{i}{2^n}]^2} |C^{-1}_2(\alpha^\prime) - C_2^{-1}(\alpha)| \\
+&\geq \frac 12 + \bO(2^{-n})
+\end{aligned}
+$$
+
+Therefore we have found a lower bound, which proves the average binned variation does not converge to zero (if it converges at all).
+
+$$
+V_n(C^{-1}) > \frac 12 \implies \lim_{n \to \infty} V_n > \frac 12
+$$
 
@@
 
-failed notions of discontinuity
+numerically, we have finitely many bits, and cannot fully represent the fractal. we therefore need the finite approximations to be already good.
 
@@
 
-measure for V
+Before we make the proof of numerical feasibility.
 
-proof that numerical convergence
+In the evaluation of the evolution equation, the integral term in each coordinate $\alpha$ is problematic.
 
-## Details
+Let us define (recall) the numerical and analytical integral (we leave a $w_U(\cdot, \beta)$ as implicit function of the first argument $\alpha$).
+The intuition behind this proof of convergence is that the variation inside of each numerical bin vanishes as we increase the number of iterations $n$ (even though the structure becomes fractal !).
+The variation comes from the coarse-graining, where we go from $4^n$ to $2^n$ samples.
 
-ironing out details : here we considered 2d -> 1d, but possible to repeat the application of the mapping so that Rp -> Rp-1 -> .. -> 1d
+We define $\tilde h_i(t) = \tilde h(t, \tfrac{i-1}{2^n}) = \tilde h(t, \beta_i),\,\beta_i=\tfrac{i-1}{2^n},\,i\in\{1,\cdots,2^n\}$.
+
+$$
+\begin{aligned}
+\mathrm{NI}_n &= \frac{1}{2^n} \sum_{i=1}^{2^n} \tilde w(\cdot, \beta_i) \phi(\tilde h_i) \\
+\mathrm{AI} &= \int_{[0,1]} \tilde w(\cdot, \beta) \phi(\tilde h(t, \beta)) \lambda \circ S^{-1}(\d \beta)
+\end{aligned}
+$$
+
+We wish to prove that $|\mathrm{NI}_n - \mathrm{AI}| \xrightarrow{n \to \infty} 0$ (implicitly : this is valid for all $\alpha \in [0,1]$).
+
+The first part of the proof involves splitting the integral on $[0,1]$ into $2^n$ integral over segments of length $2^{-n}$, and using the triangle inequality.
+
+$$
+\begin{aligned}
+|\mathrm{NI}_n - \mathrm{AI}| &= \left| \frac{1}{2^n} \sum_{i=1}^{2^n} \tilde w(\cdot, \beta_i) \phi(\tilde h_i) - \int_{0}^{1} \tilde w(\cdot, \beta) \phi(\tilde h(t, \beta)) \lambda \circ S^{-1}(\d \beta) \right| \\
+&= \left| \frac{1}{2^n} \sum_{i=1}^{2^n} \tilde w(\cdot, \beta_i) \phi(\tilde h_i) - \sum_{i=1}^{2^n} \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \tilde w(\cdot, \beta) \phi(\tilde h(t, \beta)) \lambda \circ S^{-1}(\d \beta) \right| \\
+&= \left| \sum_{i=1}^{2^n} \frac{1}{2^n} \tilde w(\cdot, \beta_i) \phi(\tilde h_i) - \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \tilde w(\cdot, \beta) \phi(\tilde h(t, \beta)) \lambda \circ S^{-1}(\d \beta) \right| \\
+&\leq \sum_{i=1}^{2^n} \left| \frac{1}{2^n} \tilde w(\cdot, \beta_i) \phi(\tilde h_i) - \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \tilde w(\cdot, \beta) \phi(\tilde h(t, \beta)) \lambda \circ S^{-1}(\d \beta) \right| \\
+\end{aligned}
+$$
+
+We now want to prove that evaluating the integrand at $\beta_i = \tfrac{i-1}{2^n}$ gives a good approximation of the integral.
+Since $\int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \lambda \circ S^{-1}(\d \beta) = \tfrac{1}{2^n}$ [TODO MATH : is this true ?], we write
+
+$$
+\begin{aligned}
+|\mathrm{NI}_n - \mathrm{AI}| &\leq \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \left[ \tilde w(\cdot, \beta_i) \phi(\tilde h_i) - \tilde w(\cdot, \beta) \phi(\tilde h(t, \beta)) \right] \lambda \circ S^{-1}(\d \beta) \right| \\
+&= \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \left[ w_U(\cdot, S^{-1}(\beta_i)) \phi(h_U(t, S^{-1}(\beta_i))) - w_U(\cdot, S^{-1}(\beta)) \phi(h_U(t, S^{-1}(\beta))) \right] \lambda \circ S^{-1}(\d \beta) \right|
+\end{aligned}
+$$
+
+The main source of variance inside the integrand is the term $S^{-1}$, since $w_U$ and $h_U$ are "nice" functions.
+We express this regularity by the added assumption that $w_U(\cdot, \vec z) h_U(t, \vec z)$ is Lipschitz in $\vec z$, with Lipschitz constant $L(\cdot)$, and $\norm{\cdot}$ the corresponding norm on $[0,1]^2$.
+
+Once we have applied the Lipschitz property, we make use of our intuition of "locality" : points close in $[0,1]$ should be (on average) mapped to points close in $[0,1]^2$.
+Therefore, taking the sup inside each $2^{-n}$ segment does not introduce a too large error.
+
+[TODO MATH : prove that it is indeed Lipschitz.
+The main problem I've enountered with this is that these functions are $\textrm{regular function} \circ \textrm{CDF}^{-1}$, and the inverse CDF goes to infinity at 0 and 1].
+
+$$
+\begin{aligned}
+|\mathrm{NI}_n - \mathrm{AI}| &\leq L(\cdot) \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \lambda \circ S^{-1}(\d \beta) \right| \\
+&\leq L(\cdot) \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \sup_{\beta \in [\tfrac{i-1}{2^n}, \tfrac{i}{2^n}]} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \lambda \circ S^{-1}(\d \beta^\prime) \right| \\
+&= L(\cdot) \frac{1}{2^n} \sum_{i=1}^{2^n} \sup_{\beta \in [\tfrac{i-1}{2^n}, \tfrac{i}{2^n}]} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \\
+&\leq L(\cdot) \frac{1}{2^n} \sum_{i=1}^{2^n} \sup_{\beta, \beta^\prime \in [\tfrac{i-1}{2^n}, \tfrac{i}{2^n}]^2} \norm{S^{-1}(\beta^\prime) - S^{-1}(\beta)} \\
+&= L(\cdot) V_n \\
+&\xrightarrow{n \to \infty} 0
+\end{aligned}
+$$
+
+We see that it is the notion of "locality" of the mapping $S$ that allowed us to write the convergence of the numerical integral to the analytical value.
+In doing so, we exploited the fact that the "variance" inside each small segment in $[0,1]$ is small on average, and that the errors vanish as we take finer and finer bins.
+
+## Generalizing to $[0,1]^p \mapsto [0,1]$
+
+ironing out details : here we considered $[0,1]^2 \mapsto [0,1]$, but the mappings can be generalized to any number of dimensions.
+
+Example for the Z-mapping. Let $\vec \alpha = (\alpha_1, \cdots, \alpha_p)$, and $\alpha_\mu = \sum_{k=1}^{n} b^\mu_k 2^{-k}$. Then
+
+$$
+Z^n(\alpha_1, \cdots, \alpha_p) = \sum_{k=1}^{n} \sum_{\mu=1}^p b^{\mu}_k 2^{-(p(k-1)+\mu)}
+$$
 

### 2023-06-04 manuscript/chapters/6_appendix.qmd
this takes so much time nyaaa

@@
 
-::: {.content-hidden}
+::: content-hidden
 {{< include ../macros.tex >}}
@@
 
-The low-rank vectors are in general obtained by sampling a distribution $\vec{\xi} = (\xi_1, \cdots, \xi_p) \sim \rho$.
-Every neuron $i$ samples this distribution independantly, and low-rank vectors are obtained by passing the samples through functions $f_\mu(\vec{\xi})$ and $g_\mu(\vec{\xi})$.
+The low-rank vectors are in general obtained by sampling a distribution $\vec{z} = (z_1, \cdots, z_p) \sim \rho$.
+Every neuron $i$ samples this distribution independently, and low-rank vectors are obtained by passing the samples through functions $f_\mu(\vec{z})$ and $g_\mu(\vec{z})$.
 
-We note that in the main text, the distribution is a multivariate normal with unit covariance and zero mean, such that it factors out as $\rho(\xi_1, \cdots, \xi_p) = \prod_{\mu=1}^p \mathcal{N}(\xi_\mu)$.
-Furthermore, the functions are component-wise $f_\mu(\xi_1, \cdots, \xi_p) = \xi_\mu$ and $g_\mu(\xi_1, \cdots, \xi_p) = \tilde \phi(\xi_\mu) = \frac{\phi(\xi_\mu) - \avg{\phi(\xi_\mu)}}{\mathrm{Var}[\phi(\xi_\mu)]}$.
+We note that in the main text, the distribution is a multivariate normal with unit covariance and zero mean, such that it factors out as $\rho(z_1, \cdots, z_p) = \prod_{\mu=1}^p \mathcal{N}(z_\mu)$.
+Furthermore, the functions are component-wise $f_\mu(z_1, \cdots, z_p) = z_\mu$ and $g_\mu(z_1, \cdots, z_p) = \tilde \phi(z_\mu) = \frac{\phi(z_\mu) - \avg{\phi(z_\mu)}}{\mathrm{Var}[\phi(z_\mu)]}$.
 
 $$
-F_{\mu i} = f_\mu(\xi_{1 i}, \cdots, \xi_{p i}), \, G_{\mu i} = g_\mu(\xi_{1 i}, \cdots, \xi_{p i})
+F_{\mu i} = f_\mu(z_{1 i}, \cdots, z_{p i}), \, G_{\mu i} = g_\mu(z_{1 i}, \cdots, z_{p i})
 $$
@@
 In order to find the fixed points, one has to solve in general $\dot h_i(t) = 0$ for all $i = 1, \cdots, N$.
-Because of the nonlinearity $\phi(h_j(t))$, this is difficult, but we can guess the fixed points $h^\star_i$ and verify them by substituting into the equation that $\dot h_i(t) = 0$.
+Because of the non-linearity $\phi(h_j(t))$, this is difficult, but we can guess the fixed points $h^\star_i$ and verify them by substituting into the equation that $\dot h_i(t) = 0$.
 
@@
 &= \phi(0) \sum_{\mu=1}^p F_{\mu i} \frac{N-1}{N} \frac{1}{N-1} \sum_{j \neq i}^N G_{\mu j} + \bO(\tfrac 1N) \\
-&\stackrel{N \gg 1}{\approx} \phi(0) \sum_{\mu=1}^p F_{\mu i} \avg{g_\mu(\vec{\xi})}
+&\stackrel{N \gg 1}{\approx} \phi(0) \sum_{\mu=1}^p F_{\mu i} \avg{g_\mu(\vec{z})}
 \end{aligned}
 $$
-
-<!-- TODO : fixme. the F_i should not be in the integral -->
 
@@
 $$
-\avg{g_\mu(\vec{\xi})} = \avg{\tilde \phi(\xi_\mu)} = \avg{\frac{\phi(\xi_\mu) - \avg{\phi(\xi_\mu)}}{\mathrm{Var}[\phi(\xi_\mu)]}} = 0
+\avg{g_\mu(\vec{z})} = \avg{\tilde \phi(z_\mu)} = \avg{\frac{\phi(z_\mu) - \avg{\phi(z_\mu)}}{\mathrm{Var}[\phi(z_\mu)]}} = 0
 $$
@@
 Another set of fixed points is $h^\star_i = F_{\nu i}$ for all $\nu = 1, \cdots, p$.
-We refer to these as the "patterns" in the context of learning the network "rembers" (converges to) these vectors.
+We refer to these as the "patterns" in the context of learning the network "remembers" (converges to) these vectors.
 We verify the fixed point similarly as before.
@@
 &= -F_{\nu i} + \frac 1N \sum_{j \neq i}^N F_{\nu i} G_{\nu j} \phi(F_{\nu j}) + \frac 1N \sum_{\mu \neq \nu}^p \sum_{j \neq i}^N F_{\mu i} G_{\mu j} \phi(F_{\nu j}) + \bO(\tfrac 1N) \\
-&\stackrel{N \gg 1}{\approx} -F_{\nu i} + F_{\nu i} \avg{g_\nu(\vec{\xi_j}) \phi(\xi_{\nu j})} + \sum_{\mu \neq \nu}^p F_{\mu i} \avg{g_\mu(\vec{\xi_j}) \phi(\xi_{\nu j})}
+&\stackrel{N \gg 1}{\approx} -F_{\nu i} + F_{\nu i} \avg{g_\nu(\vec{z_j}) \phi(z_{\nu j})} + \sum_{\mu \neq \nu}^p F_{\mu i} \avg{g_\mu(\vec{z_j}) \phi(z_{\nu j})}
 \end{aligned}
@@
 
-The term $\avg{g_\mu(\vec{\xi_j}) \phi(\xi_{\nu j})}$ vanishes in the case of a factorizable distribution and when the functions $g_\mu$ are component-wise.
+The term $\avg{g_\mu(\vec{z_j}) \phi(z_{\nu j})}$ vanishes in the case of a factorizable distribution and when the functions $g_\mu$ are component-wise.
 
@@
 \begin{aligned}
-\avg{g_\mu(\vec{\xi_j}) \phi(\xi_{\nu j})} &= 
-\int_{\Rp} g_\mu(\xi_1, \cdots, \xi_p) \phi(\xi_\nu) \rho(\d \vec{\xi}) \\
-&= \int_{\Rp} g_\mu(\xi_\mu) \phi(\xi_\nu) \prod_{\gamma=1}^p \rho_\gamma(\d \xi_\gamma) \\
-&= \int_{\R} g_\mu(\xi_\mu) \rho_\mu(\d \xi_\mu) \int_{\R} \phi(\xi_\nu) \rho_\nu(\d \xi_\nu) \\
-&= \avg{g_\mu(\xi_\mu)} \avg{\phi(\xi_\nu)}
+\avg{g_\mu(\vec{z_j}) \phi(z_{\nu j})} &= 
+\int_{\Rp} g_\mu(z_1, \cdots, z_p) \phi(z_\nu) \rho(\d \vec{z}) \\
+&= \int_{\Rp} g_\mu(z_\mu) \phi(z_\nu) \prod_{\gamma=1}^p \rho_\gamma(\d z_\gamma) \\
+&= \int_{\R} g_\mu(z_\mu) \rho_\mu(\d z_\mu) \int_{\R} \phi(z_\nu) \rho_\nu(\d z_\nu) \\
+&= \avg{g_\mu(z_\mu)} \avg{\phi(z_\nu)}
 \end{aligned}
@@
 
-As established in the previous section, in the case of the network studied in this work, $\avg{g_\mu(\xi_\mu)} = 0$.
+As established in the previous section, in the case of the network studied in this work, $\avg{g_\mu(z_\mu)} = 0$.
 
-The term $\avg{g_\nu(\vec{\xi_j}) \phi(\xi_{\nu j})}$ is more difficult, but we can prove that in our case it equals one.
+The term $\avg{g_\nu(\vec{z_j}) \phi(z_{\nu j})}$ is more difficult, but we can prove that in our case it equals one.
 
@@
 \begin{aligned}
-\avg{g_\nu(\vec{\xi_j}) \phi(\xi_{\nu j})} &=
-\avg{\frac{\phi(\xi_{\mu j}) - \avg{\phi(\xi_{\mu j})}}{\mathrm{Var}[\phi(\xi_{\mu j})]} \phi(\xi_{\mu j})} \\
-&= \avg{\frac{\phi(\xi)^2 - \avg{\phi(\xi)} \phi(\xi) }{\mathrm{Var}[\phi(\xi)]}} \\
-&= \frac{\avg{\phi(\xi)^2} - \avg{\avg{\phi(\xi)} \phi(\xi)} }{\mathrm{Var}[\phi(\xi)]} \\
-&= \frac{\avg{\phi(\xi)^2} - \avg{\phi(\xi)}^2}{\mathrm{Var}[\phi(\xi)]} \\
-&= \frac{\mathrm{Var}[\phi(\xi)]}{\mathrm{Var}[\phi(\xi)]} \\
+\avg{g_\nu(\vec{z_j}) \phi(z_{\nu j})} &=
+\avg{\frac{\phi(z_{\mu j}) - \avg{\phi(z_{\mu j})}}{\mathrm{Var}[\phi(z_{\mu j})]} \phi(z_{\mu j})} \\
+&= \avg{\frac{\phi(z)^2 - \avg{\phi(z)} \phi(z) }{\mathrm{Var}[\phi(z)]}} \\
+&= \frac{\avg{\phi(z)^2} - \avg{\avg{\phi(z)} \phi(z)} }{\mathrm{Var}[\phi(z)]} \\
+&= \frac{\avg{\phi(z)^2} - \avg{\phi(z)}^2}{\mathrm{Var}[\phi(z)]} \\
+&= \frac{\mathrm{Var}[\phi(z)]}{\mathrm{Var}[\phi(z)]} \\
 &= 1
@@
 
-Therefore, the set of $h_i^\star = F_{\nu i}$ for all $\nu = 1, \cdots, p$ are fixed points in the limit of large $N$, under the assumptions that the distribution is factorizable, the functions $g_\mu$ are component-wise, and $\avg{g_\nu(\xi) \phi(\xi)}=1$, which is the case in the setup of this work.
+Therefore, the set of $h_i^\star = F_{\nu i}$ for all $\nu = 1, \cdots, p$ are fixed points in the limit of large $N$, under the assumptions that the distribution is factorizable, the functions $g_\mu$ are component-wise, and $\avg{g_\nu(z) \phi(z)}=1$, which is the case in the setup of this work.
 
@@
 
-As $N \to \infty$, the fixed points $h_i^\star$ become functions $h^\star(\xi_1, \cdots, \xi_p)$ on the $\Rp$ space :
+As $N \to \infty$, the fixed points $h_i^\star$ become functions $h^\star(z_1, \cdots, z_p)$ on the $\Rp$ space :
 
-* $h_i^\star = 0 \to h^\star(\xi_1, \cdots, \xi_p) = 0$
-* $h_i^\star = F_{\nu i} = \xi_{\nu i} \to h^\star(\xi_1, \cdots, \xi_p) = \xi_\nu$ for all $\nu = 1, \cdots, p$
+-   $h_i^\star = 0 \to h^\star(z_1, \cdots, z_p) = 0$
+-   $h_i^\star = F_{\nu i} = z_{\nu i} \to h^\star(z_1, \cdots, z_p) = z_\nu$ for all $\nu = 1, \cdots, p$
 
-<!-- TODO : add derivation for neural field ? -->
+[TODO : add derivation for neural field ?]
 
@@
 
-<!-- TODO -->
+We now study the stability of the fixed points. In the original equation, we substitute $h_i(t) = h_i^\star + \delta_i(t)$, where $\delta_i(t)$ is a small perturbation.
 
-## Behavior of $\kappa_\mu(t)$ in the cyclic network
+$$
+\begin{aligned}
+\dot h_i(t) &= \dot \delta_i(t) \\
+&= -(h_i^\star + \delta_i(t)) + \sum_{j=1}^N J_{ij} \phi(h_j^\star + \delta_j(t)) \\
+&= -(h_i^\star + \delta_i(t)) + \sum_{j=1}^N J_{ij} (\phi(h_j^\star) + \partial \phi(h_j^\star) \delta_j(t) + \bO(\delta_j(t)^2)) \\
+&= \underbrace{-h_i^\star + \sum_{j=1}^N J_{ij} \phi(h_j^\star)}_\text{$=0$ by definition of the fixed point} - \delta_i(t) + \sum_{j=1}^N J_{ij} \partial \phi(h_j^\star) \delta_j(t) + \bO(\norm{\vec\delta(t)}^2) \\
+&= - \delta_i(t) + \sum_{j=1}^N J_{ij} \partial \phi(h_j^\star) \delta_j(t) + \bO(\norm{\vec\delta(t)}^2) \\
+&= \sum_{j=1}^N (\underbrace{-\mathrm{Id}_{ij} + \underbrace{J_{ij} \partial \phi(h_j^\star)}_{J_{ij}^\prime}}_{K_{ij}}) \delta_j(t) + \bO(\norm{\vec\delta(t)}^2)
+\end{aligned}
+$$
 
-kappa in the delayed, rolled and cyclic network
+In matrix notation $\vec\delta(t) = (\delta_1(t), \cdots, \delta_N(t))$, this becomes
 
-relation to overlaps
+$$
+\begin{aligned}
+&\dot{\vec\delta}(t) = K \vec\delta(t) \\
+\implies &\vec\delta(t) = \mathrm{exp}(Kt) \vec\delta(0)
+\end{aligned}
+$$
+
+The fixed point is stable if $\vec\delta(t) \to 0$ as $t \to \infty$, in other words if all the eigenvalues of $K$ are negative.
+
+Before we study the spectrum of $K$, we cite the following result : if $\lambda$ is an eigenvalue of $A$, then $\lambda-1$ is an eigenvalue of $A-\mathrm{Id}$.
+This follows simply from the equation for the eigenvalues $\lambda_i$ and eigenvectors $\vec{v_i}$ of $A$.
+If $A \vec{v_i} = \lambda_i \vec{v_i}$, then $(A-\mathrm{Id})\vec{v_i}=\lambda_i \vec{v_i} - \vec{v_i} = (\lambda_i - 1)\vec{v_i}$.
+Therefore to study $K$ it suffices to study the eigenvalues of $J^\prime$.
+
+Since $J^\prime$ is low-rank, its image is spanned by the vectors $\{\vec{F_\nu} = (F_{\nu 1}, \cdots, F_{\nu N}) | \nu=1,\cdots,p\}$, which motivates the Ansatz of eigenvectors $\vec{v_\nu} = \vec{F_\nu}$ in the $N \to \infty$ limit.
+
+$$
+\begin{aligned}
+(J^\prime \vec{F_\nu})_i &= \frac 1N \sum_{\mu=1}^p \sum_{j=1}^N F_{\mu i} \underbrace{G_{\mu j} \partial \phi(h_j^\star)}_{G^\prime_{\mu j}} F_{\nu j} \\
+&= \frac 1N \sum_{\mu \neq \nu}^p \sum_{j=1}^N F_{\mu i} G^\prime_{\mu j} F_{\nu j} + \frac 1N \sum_{j=1}^N F_{\nu i} G^\prime_{\nu j} F_{\nu j} \\
+&\stackrel{N \gg 1}{\approx} \sum_{\mu \neq \nu}^p F_{\mu i} \avg{g^\prime_\mu(\vec{z}) f_\nu(\vec{z})} + F_{\nu i} \avg{g^\prime_\nu(\vec{z}) f_\nu(\vec{z})}
+\end{aligned}
+$$
+
+We study each term separately.
+Additionally (and similarly as before), we make the assumption that the functions $g_\mu$ and the functions $f_\mu$ are component-wise, and that the distribution $\rho$ factorises.
+
+$$
+\begin{aligned}
+\avg{g^\prime_\mu(\vec{z}) f_\nu(\vec{z})} &=
+\int_{\Rp} g^\prime_\mu(\vec z) f_\nu(\vec z) \rho(\d \vec z) \\
+&= \int_{\R} \int_{\R} g^\prime_\mu(z_\mu) f_\nu(z_\nu) \rho_\mu(\d z_\mu) \rho_\nu(\d z_\nu) \\
+&= \int_{\R} g^\prime_\mu(z_\mu) \rho_\mu(\d z_\mu) \int_{\R} f_\nu(z_\nu) \rho_\nu(\d z_\nu) \\
+&= \avg{g^\prime_\mu(z_\mu)} \avg{f_\nu(z_\nu)}
+\end{aligned}
+$$
+
+In our case, $\avg{f_\nu(z_\nu)} = \avg{z_\nu} = \int_{\R} z_\nu \mathcal{N}(\d z_\nu) = 0$ and so the term vanishes. Therefore, we are left with
+
+$$
+\begin{aligned}
+&(J^\prime \vec{F_\nu})_i \stackrel{N \gg 1}{\approx} F_{\nu i} \avg{g^\prime_\nu(z_\nu) f_\nu(z_\nu)} \\
+\implies &J^\prime \vec{F_\nu} = \avg{g^\prime_\nu(z_\nu) f_\nu(z_\nu)} \vec{F_\nu} \\
+\implies &K \vec{F_\nu} = (\underbrace{\avg{g^\prime_\nu(z_\nu) f_\nu(z_\nu)}-1}_\text{eigenvalue $\lambda_\nu$}) \vec{F_\nu} \\
+\end{aligned}
+$$
+
+The term $\avg{g^\prime_\nu(z_\nu) f_\nu(z_\nu)}$ does not vanish in general because of the coupling between the components.
+The integral is too complicated to be analytically solved, but we provide a numerical estimation in our case ($\rho(z_1, \cdots, z_p)=\prod_{\mu=1}^p\mathcal{N}(z_\mu)$, $f_\mu(z_\mu)=z_\mu$, $g_\mu(z_\mu)=\tilde\phi(z_\mu)$) and with fixed points $h_i^\star=0$ and $h_i^\star=F_{\nu i}$.
+
+$$
+\begin{aligned}
+\phi(h)&=\frac{1}{1+\mathrm{e}^{-h}} \\
+&\implies
+\begin{cases}
+h_i^\star=0 &\implies \lambda_\nu = \avg{\tilde\phi(z) \partial \phi(0) z} - 1 \approx 0.19061 \pm 0.00014 \\
+h_i^\star=F_{\nu i} &\implies \lambda_\nu = \avg{\tilde\phi(z) \partial \phi(z) z} - 1 \approx -0.28090 ± 0.00006
+\end{cases}
+\end{aligned}
+$$
+
+When $\phi$ is the logistic function, the pattern fixed points are stable, whereas the zero fixed point is unstable.
+
+We now consider a Taylor expansion of the rate function $\phi(h)$ up to order 3, and cite results obtained from symbolic computation and known moments of the gaussian distribution.
+
+$$
+\begin{aligned}
+	\phi(h)&= c_0 + c_1 h,\, c_0, c_1 \in \R^2 \\
+	&\implies
+	\lambda_\nu = \avg{\frac{z}{c_{1}} c_{1} z} - 1 = 0 \text{ for any fixed point}\\
+	\phi(h)&= c_0 + c_1 h + c_2 h^2,\, c_0, c_1, c_2 \in \R^3 \\
+	&\implies
+	\begin{cases}
+		h_i^\star=0 &\implies \begin{aligned}[t]
+			\lambda_\nu &= \avg{\frac{c_{1} z + c_{2} z^{2} - c_{2}}{c_{1}^{2} + 2 c_{2}^{2}} (c_1) z} - 1 \\
+			&= - \frac{2 c_{2}^{2}}{c_{1}^{2} + 2 c_{2}^{2}}
+		\end{aligned} \\
+		h_i^\star=F_{\nu i} &\implies \begin{aligned}[t]
+			\lambda_\nu &= \avg{\frac{c_{1} z + c_{2} z^{2} - c_{2}}{c_{1}^{2} + 2 c_{2}^{2}} (c_{1} + 2 c_{2} z) z} - 1 \\
+			&= \frac{2 c_{2}^{2}}{c_{1}^{2} + 2 c_{2}^{2}}
+		\end{aligned}
+	\end{cases} \\
+	\phi(h)&= c_0 + c_1 h + c_2 h^2 + c_3 h^3,\, c_0, c_1, c_2, c_3 \in \R^4 \\
+	&\implies
+	\begin{cases}
+		h_i^\star=0 &\implies \begin{aligned}[t]
+			\lambda_\nu &= \avg{\frac{c_{1} z + c_{2} z^{2} - c_{2} + c_{3} z^{3}}{c_{1}^{2} + 6 c_{1} c_{3} + 2 c_{2}^{2} + 15 c_{3}^{2}} (c_{1}) z} - 1 \\
+			&= \frac{- 3 c_{1} c_{3} - 2 c_{2}^{2} - 15 c_{3}^{2}}{c_{1}^{2} + 6 c_{1} c_{3} + 2 c_{2}^{2} + 15 c_{3}^{2}}
+		\end{aligned} \\
+		h_i^\star=F_{\nu i} &\implies \begin{aligned}[t]
+			\lambda_\nu &= \avg{\frac{c_{1} z + c_{2} z^{2} - c_{2} + c_{3} z^{3}}{c_{1}^{2} + 6 c_{1} c_{3} + 2 c_{2}^{2} + 15 c_{3}^{2}}
+			(c_{1} + 2 c_{2} z + 3 c_{3} z^{2}) z} - 1 \\
+			&= \frac{2 \cdot \left(3 c_{1} c_{3} + c_{2}^{2} + 15 c_{3}^{2}\right)}{c_{1}^{2} + 6 c_{1} c_{3} + 2 c_{2}^{2} + 15 c_{3}^{2}}
+		\end{aligned}
+	\end{cases}
+\end{aligned}
+$$
+
+What a mess ! We summarize the results from the Taylor expansion :
+
+* When $\phi$ is linear, we see that all the eigenvalues are zero, therefore the first order stability is unconclusive.
+Numerically, we see that when $\phi$ is linear, the convergence to the fixed points is very slow.
+* When $\phi$ is quadratic, the zero fixed point is stable, and the pattern fixed point is unstable, regardless of the sign of the coefficients.
+* When $\phi$ is cubic, the sign of the eigenvalues depends primarily on $c_3$ (negative $c_1$ breaks the assumption that $\phi$ must be monotonically increasing).
+The intuition is that we require $c_1$ to be large enough (in other words, the growth of the firing rate when $h$ becomes positive) and $c_3 < 0$ such that $3 c_1 c_3 + 15 c_3^2 < 0$.
+  The magnitude of the second order term $c_2$ only contributes to adding instability to the pattern fixed point, and stability to the zero fixed point, which leads to the conclusion it is best set to $c_2=0$ if we want the network to be stable at the pattern fixed points.
+
+[TODO FIG : might be nice to have a phase plot of $c_1, c_3$ and the value of the eigenvalue to illustrate the previous illustration]
+
+The study of the Taylor expansion therefore instructs us that stability is improved when $\phi$ grows rapidly around $h=0$, and that even powers (symmetries of $\phi$ around the y-axis) deteriorate the stability of the pattern fixed point.
+This helps us understand why the logistic function $\phi(h)=\frac{1}{1+\mathrm{e}^{-h}}=\tfrac 12 + \tfrac h4 - \tfrac{h^3}{48} + \bO(h^5)$ has good convergence properties : we have that $c_1$ is large enough, $c_3$ is negative, and $c_2$ is zero.
+
+We finish the study of eigenvalues by noting that in the $N \to \infty$ limit, the remaining $N-p$ eigenvalues of $J^\prime$ are zero.
+The corresponding eigenvectors span the orthogonal complement of the vectors $\vec{G^\prime_\nu} = (G^\prime_{\nu 1}, \cdots, G^\prime_{\nu N})$, formally $\vec{v} \in \mathrm{span}(\vec{G^\prime_1}, \cdots,\vec{G^\prime_p})^\perp$.
+
+In conclusion, the spectrum of $K$ is composed of $p$ eigenvalues $\lambda_\nu = \avg{g_\nu(z_\nu) f_\nu(z_\nu) \partial \phi(h^\star)}-1$, and $N-p$ eigenvalues $\lambda_\perp = -1$.
+For the network considered in this work, the interpretation of this is that components of the field aligned with the patterns $z_\nu$ converge to the fixed point, and the orthogonal components decay as $t \to \infty$.
+The fixed point $h_i^\star=0$ is therefore unstable, and the pattern fixed points $h_i^\star = z_{\nu i}$ are stable.
+We also note that since the normal distribution is centered around zero and symmetric, $h_i^\star = -z_{\nu i}$ are also stable fixed points.
+
+## Behaviour of $\kappa_\mu(t)$ in the cyclic neural field
+
+### Definitions of dynamics
+
+[TODO MATH : derivation of this ? from the general setting to our case]
+
+We recall the definition of the low-rank variables $\kappa_\mu(t)$.
+
+The dynamics of $h(t, \vec z)$ are in a subsystem of dimension $p$ spanned by the ONB of functions $\{e_\mu(\vec z) = z_\mu | \mu=1,\cdots,p\}$, with the scalar product $\avg{f, g} = \int_{\mathbb R^p} f(\vec y) g(\vec y) \rho(\mathrm d \vec y)$.
+
+We decompose $h(t, \vec z) = h^\perp(t, \vec z) + \sum_{\mu=1}^p \kappa_\mu(t) z_\mu$, and write the system of equations for $\mu=1,\cdots,p$.
+
+$$
+\dot \kappa_\mu(t) = -\kappa_\mu(t) + \int_{\mathbb{R}^p} \tilde\phi(y_\mu) \phi(h(t, \vec y)) \rho(\mathrm d \vec y) = -\kappa_\mu(t) + m_\mu(t)
+$$
+
+with initial conditions
+
+$$
+\kappa_\mu(0) = \int_{\mathbb{R}^p} y_\mu h(0, \vec y) \rho(\mathrm d \vec y)
+$$
+
+and the orthogonal component evolves according to
+
+$$
+h^\perp(t, \vec z) = h^\perp(0, \vec z) \mathrm e^{-t},
+\quad
+h^\perp(0, \vec z) = h(0, \vec z) - \sum_{\mu=1}^p \kappa_\mu(0) z_\mu
+$$
+
+### Cyclic dynamics in the $p=2$ cyclic neural field
+
+[TODO MATH : derivation of delayed ?]
+
+[TODO MATH : derivation of cyclic ?]
+
+In the case of the cyclic neural field, the evolution equation for $\kappa_\nu(t)$ is modified, where $\delta$ is the delay, and with the convention $\mu+1 = 1$ if $\mu=p$.
+
+$$
+\dot \kappa_\mu(t) = -\kappa_\mu(t) + m_{\mu+1}(t - \delta)
+$$
+
+### Relation to overlaps $m_\mu(t)$
+
+[TODO]
 
 ## "Mean patterns" in the binned connectivity
+
+[TODO]
 

### 2023-06-07 manuscript/appendices/A_fixed_points.qmd
id rather not lose my changes now lmao

--- 2023-06-04 manuscript/chapters/6_appendix.qmd
@@
-# Appendix
-
 ::: content-hidden
@@
 
-## Fixed point study in the network of neurons
+# Fixed point study in the network of neurons
 
@@
 
-### Derivation of fixed points
+## Derivation of fixed points
 
@@
 
-#### Zero fixed point
+### Zero fixed point
 
@@
 
-#### Pattern fixed point
+### Pattern fixed point
 
@@
 
-### Fixed points of the neural field
+## Fixed points of the neural field
 
@@
 
-### Stability of the fixed points
+## Stability of the fixed points
 
@@
 We also note that since the normal distribution is centered around zero and symmetric, $h_i^\star = -z_{\nu i}$ are also stable fixed points.
-
-## Behaviour of $\kappa_\mu(t)$ in the cyclic neural field
-
-### Definitions of dynamics
-
-[TODO MATH : derivation of this ? from the general setting to our case]
-
-We recall the definition of the low-rank variables $\kappa_\mu(t)$.
-
-The dynamics of $h(t, \vec z)$ are in a subsystem of dimension $p$ spanned by the ONB of functions $\{e_\mu(\vec z) = z_\mu | \mu=1,\cdots,p\}$, with the scalar product $\avg{f, g} = \int_{\mathbb R^p} f(\vec y) g(\vec y) \rho(\mathrm d \vec y)$.
-
-We decompose $h(t, \vec z) = h^\perp(t, \vec z) + \sum_{\mu=1}^p \kappa_\mu(t) z_\mu$, and write the system of equations for $\mu=1,\cdots,p$.
-
-$$
-\dot \kappa_\mu(t) = -\kappa_\mu(t) + \int_{\mathbb{R}^p} \tilde\phi(y_\mu) \phi(h(t, \vec y)) \rho(\mathrm d \vec y) = -\kappa_\mu(t) + m_\mu(t)
-$$
-
-with initial conditions
-
-$$
-\kappa_\mu(0) = \int_{\mathbb{R}^p} y_\mu h(0, \vec y) \rho(\mathrm d \vec y)
-$$
-
-and the orthogonal component evolves according to
-
-$$
-h^\perp(t, \vec z) = h^\perp(0, \vec z) \mathrm e^{-t},
-\quad
-h^\perp(0, \vec z) = h(0, \vec z) - \sum_{\mu=1}^p \kappa_\mu(0) z_\mu
-$$
-
-### Cyclic dynamics in the $p=2$ cyclic neural field
-
-[TODO MATH : derivation of delayed ?]
-
-[TODO MATH : derivation of cyclic ?]
-
-In the case of the cyclic neural field, the evolution equation for $\kappa_\nu(t)$ is modified, where $\delta$ is the delay, and with the convention $\mu+1 = 1$ if $\mu=p$.
-
-$$
-\dot \kappa_\mu(t) = -\kappa_\mu(t) + m_{\mu+1}(t - \delta)
-$$
-
-### Relation to overlaps $m_\mu(t)$
-
-[TODO]
-
-## "Mean patterns" in the binned connectivity
-
-[TODO]
-
-derivation of $\Jab$ and mean patterns
-
-For finite number of recursive quadrant iterations $n$, we can do a "mean-field approximation" inside each of the $4^n$ segments.
-Let $\alpha = \{i_1,\cdots,i_{|\alpha|}\}$ be the multi-index corresponding to all neurons of which the embedding in $\mathbb R^p$ gets mapped to the segment $\alpha$ in $[0,1]$.
-Let $H_\alpha(t) = \frac 1 {|\alpha|} \sum_{i \in \alpha} h_i(t)$ be the (mean) RNN potential of the segment $\alpha$.
-The connectivity matrix $\tilde J_{\alpha,\beta}$ satisfies
-
-$$
-\dot H_\alpha(t) = -H_\alpha(t) + \sum_{\beta \in \text{segments of length } 4^{-n}} \tilde J_{\alpha,\beta} \phi(H_\beta(t))
-$$
-
-By substituting the original $h_i(t)$, we find the correct rescaling is given by
-
-$$
-\tilde J_{\alpha,\beta} = \frac 1 {|\alpha|} \sum_{i \in \alpha} \sum_{j \in \beta} J_{ij}
-$$
-
-In the low-rank case, we have
-
-$$
-J_{ij} = \frac 1N \sum_{\mu=1}^p F_{\mu,i} G_{\mu,j}
-$$
-
-We can define the "mean pattern" inside each bin :
-
-$$
-\tilde F_{\mu,\alpha} = \frac{1}{|\alpha|} \sum_{i \in \alpha} F_{\mu,i}, \; \tilde G_{\mu,\alpha} = \frac{1}{|\alpha|} \sum_{i \in \alpha} G_{\mu,i}
-$$
-
-Then the connectivity matrix is given by, noting $N = \sum_{\alpha} |\alpha|$,
-
-$$
-\tilde J_{\alpha,\beta} = \frac{|\beta|}{\sum_{\beta'} |\beta'|} \left( \sum_{\mu=1}^p \tilde F_{\mu,\alpha} \tilde G_{\mu,\beta} - \delta_{\alpha,\beta} \underbrace{\sum_{\mu=1}^p \sum_{i \in \alpha} \frac{F_{\mu,i}}{|\alpha|} \frac{G_{\mu,i}}{|\alpha|}}_{\gamma_{\alpha}} \right)
-$$
-
-Recurrent current is given by
-
-$$
-I^\text{rec}_\alpha(t) = \sum_\beta \tilde J_{\alpha,\beta} \phi(H_\beta(t)) = \frac{1}{\sum_{\beta'} |\beta'|} \left( \sum_\beta |\beta| \sum_{\mu=1}^p \tilde F_{\mu,\alpha} \tilde G_{\mu,\beta} \phi(H_\beta(t)) - |\alpha| \tilde \gamma_\alpha \phi(H_\alpha(t)) \right)
-$$
-
-The overlaps can be computed as
-
-$$
-\tilde m_\mu(t) = \frac{1}{\sum_{\alpha} |\alpha|} \sum_\alpha |\alpha| G_{\mu,\alpha} \phi(H_\alpha(t))
-$$

### 2023-06-07 manuscript/appendices/C_mean_patterns.qmd
id rather not lose my changes now lmao

--- 2023-05-31 manuscript/chapters/6_appendix.qmd
@@
-# Appendix
+# "Mean patterns" in the binned connectivity
 
-::: {.content-hidden}
-{{< include ../macros.tex >}}
-:::
-
-## Fixed point study in the network of neurons
-
-We recall the equation for the low-rank network of neurons :
-
-$$
-\dot h_i(t) = -h_i(t) + \frac 1N \sum_{\mu=1}^p \sum_{j=1}^N F_{\mu i} G_{\mu j} \phi(h_j(t))
-$$
-
-The low-rank vectors are in general obtained by sampling a distribution $\vec{\xi} = (\xi_1, \cdots, \xi_p) \sim \rho$.
-Every neuron $i$ samples this distribution independantly, and low-rank vectors are obtained by passing the samples through functions $f_\mu(\vec{\xi})$ and $g_\mu(\vec{\xi})$.
-
-We note that in the main text, the distribution is a multivariate normal with unit covariance and zero mean, such that it factors out as $\rho(\xi_1, \cdots, \xi_p) = \prod_{\mu=1}^p \mathcal{N}(\xi_\mu)$.
-Furthermore, the functions are component-wise $f_\mu(\xi_1, \cdots, \xi_p) = \xi_\mu$ and $g_\mu(\xi_1, \cdots, \xi_p) = \tilde \phi(\xi_\mu) = \frac{\phi(\xi_\mu) - \avg{\phi(\xi_\mu)}}{\mathrm{Var}[\phi(\xi_\mu)]}$.
-
-$$
-F_{\mu i} = f_\mu(\xi_{1 i}, \cdots, \xi_{p i}), \, G_{\mu i} = g_\mu(\xi_{1 i}, \cdots, \xi_{p i})
-$$
-
-### Derivation of fixed points
-
-In order to find the fixed points, one has to solve in general $\dot h_i(t) = 0$ for all $i = 1, \cdots, N$.
-Because of the nonlinearity $\phi(h_j(t))$, this is difficult, but we can guess the fixed points $h^\star_i$ and verify them by substituting into the equation that $\dot h_i(t) = 0$.
-
-#### Zero fixed point
-
-A trivial fixed point is $h^\star_i = 0$ for all $i$, which we prove to be a fixed point as $N \to \infty$.
-
-$$
-\begin{aligned}
-\dot h_i(t) &= \frac 1N \sum_{\mu=1}^p \sum_{j=1}^N F_{\mu i} G_{\mu j} \phi(0) \\
-&= \frac{\phi(0)}{N} \sum_{\mu=1}^p \sum_{j=1}^N F_{\mu i} G_{\mu j} \\
-&= \frac{\phi(0)}{N} \sum_{\mu=1}^p \sum_{j \neq i}^N F_{\mu i} G_{\mu j} + \underbrace{\frac{\phi(0)}{N} \sum_{\mu=1}^p F_{\mu i} G_{\mu i}}_{\bO(\frac 1N)} \\
-&= \phi(0) \sum_{\mu=1}^p \frac 1N \sum_{j \neq i}^N F_{\mu i} G_{\mu j} + \bO(\tfrac 1N) \\
-&= \phi(0) \sum_{\mu=1}^p F_{\mu i} \frac{N-1}{N} \frac{1}{N-1} \sum_{j \neq i}^N G_{\mu j} + \bO(\tfrac 1N) \\
-&\stackrel{N \gg 1}{\approx} \phi(0) \sum_{\mu=1}^p F_{\mu i} \avg{g_\mu(\vec{\xi})}
-\end{aligned}
-$$
-
-<!-- TODO : fixme. the F_i should not be in the integral -->
-
-To write the last line, we make a slight abuse of notation in order to avoid notation problems when taking the $N \to \infty$ limit of $F_{\mu i}$.
-
-In the case of the network studied in this work, the find that
-
-$$
-\avg{g_\mu(\vec{\xi})} = \avg{\tilde \phi(\xi_\mu)} = \avg{\frac{\phi(\xi_\mu) - \avg{\phi(\xi_\mu)}}{\mathrm{Var}[\phi(\xi_\mu)]}} = 0
-$$
-
-Thus, $h_i^\star=0$ is a fixed point in the limit $N \to \infty$.
-
-#### Pattern fixed point
-
-Another set of fixed points is $h^\star_i = F_{\nu i}$ for all $\nu = 1, \cdots, p$.
-We refer to these as the "patterns" in the context of learning the network "rembers" (converges to) these vectors.
-We verify the fixed point similarly as before.
-
-$$
-\begin{aligned}
-\dot h_i(t) &= -F_{\nu i} + \frac 1N \sum_{\mu=1}^p \sum_{j=1}^N F_{\mu i} G_{\mu j} \phi(F_{\nu j}) \\
-&= -F_{\nu i} + \frac 1N \sum_{\mu=1}^p \sum_{j \neq i}^N F_{\mu i} G_{\mu j} \phi(F_{\nu j}) + \bO(\tfrac 1N) \\
-&= -F_{\nu i} + \frac 1N \sum_{j \neq i}^N F_{\nu i} G_{\nu j} \phi(F_{\nu j}) + \frac 1N \sum_{\mu \neq \nu}^p \sum_{j \neq i}^N F_{\mu i} G_{\mu j} \phi(F_{\nu j}) + \bO(\tfrac 1N) \\
-&\stackrel{N \gg 1}{\approx} -F_{\nu i} + F_{\nu i} \avg{g_\nu(\vec{\xi_j}) \phi(\xi_{\nu j})} + \sum_{\mu \neq \nu}^p F_{\mu i} \avg{g_\mu(\vec{\xi_j}) \phi(\xi_{\nu j})}
-\end{aligned}
-$$
-
-The term $\avg{g_\mu(\vec{\xi_j}) \phi(\xi_{\nu j})}$ vanishes in the case of a factorizable distribution and when the functions $g_\mu$ are component-wise.
-
-$$
-\begin{aligned}
-\avg{g_\mu(\vec{\xi_j}) \phi(\xi_{\nu j})} &= 
-\int_{\Rp} g_\mu(\xi_1, \cdots, \xi_p) \phi(\xi_\nu) \rho(\d \vec{\xi}) \\
-&= \int_{\Rp} g_\mu(\xi_\mu) \phi(\xi_\nu) \prod_{\gamma=1}^p \rho_\gamma(\d \xi_\gamma) \\
-&= \int_{\R} g_\mu(\xi_\mu) \rho_\mu(\d \xi_\mu) \int_{\R} \phi(\xi_\nu) \rho_\nu(\d \xi_\nu) \\
-&= \avg{g_\mu(\xi_\mu)} \avg{\phi(\xi_\nu)}
-\end{aligned}
-$$
-
-As established in the previous section, in the case of the network studied in this work, $\avg{g_\mu(\xi_\mu)} = 0$.
-
-The term $\avg{g_\nu(\vec{\xi_j}) \phi(\xi_{\nu j})}$ is more difficult, but we can prove that in our case it equals one.
-
-$$
-\begin{aligned}
-\avg{g_\nu(\vec{\xi_j}) \phi(\xi_{\nu j})} &=
-\avg{\frac{\phi(\xi_{\mu j}) - \avg{\phi(\xi_{\mu j})}}{\mathrm{Var}[\phi(\xi_{\mu j})]} \phi(\xi_{\mu j})} \\
-&= \avg{\frac{\phi(\xi)^2 - \avg{\phi(\xi)} \phi(\xi) }{\mathrm{Var}[\phi(\xi)]}} \\
-&= \frac{\avg{\phi(\xi)^2} - \avg{\avg{\phi(\xi)} \phi(\xi)} }{\mathrm{Var}[\phi(\xi)]} \\
-&= \frac{\avg{\phi(\xi)^2} - \avg{\phi(\xi)}^2}{\mathrm{Var}[\phi(\xi)]} \\
-&= \frac{\mathrm{Var}[\phi(\xi)]}{\mathrm{Var}[\phi(\xi)]} \\
-&= 1
-\end{aligned}
-$$
-
-Therefore, the set of $h_i^\star = F_{\nu i}$ for all $\nu = 1, \cdots, p$ are fixed points in the limit of large $N$, under the assumptions that the distribution is factorizable, the functions $g_\mu$ are component-wise, and $\avg{g_\nu(\xi) \phi(\xi)}=1$, which is the case in the setup of this work.
-
-### Fixed points of the neural field
-
-The derivation of the fixed points for the neural field equation follows the same arguments as the fixed points in the network of neurons.
-
-As $N \to \infty$, the fixed points $h_i^\star$ become functions $h^\star(\xi_1, \cdots, \xi_p)$ on the $\Rp$ space :
-
-* $h_i^\star = 0 \to h^\star(\xi_1, \cdots, \xi_p) = 0$
-* $h_i^\star = F_{\nu i} = \xi_{\nu i} \to h^\star(\xi_1, \cdots, \xi_p) = \xi_\nu$ for all $\nu = 1, \cdots, p$
-
-<!-- TODO : add derivation for neural field ? -->
-
-### Stability of the fixed points
-
-<!-- TODO -->
-
-## Behavior of $\kappa_\mu(t)$ in the cyclic network
-
-kappa in the delayed, rolled and cyclic network
-
-relation to overlaps
-
-## "Mean patterns" in the binned connectivity
+[TODO]
 

### 2023-06-07 manuscript/chapters/1_intro.qmd
id rather not lose my changes now lmao

@@
 
-## Networks of rate neurons
+## Neural field models in neuroscience
 
-Networks of rate neurons model the time evolution of the potential $h_i(t) \,[\textrm{V}]$ associated to $N$ individual neurons labeled $i\in\{1,\cdots,N\}$.
-In their most general form, a connectivity matrix $J_{ij}$ describes the amplitude of the recurrent currents $I^\text{rec}_{ij} = J_{ij} \phi(h_j(t)) \,[\textrm{A}]$ emitted by neuron $j$ and received by neuron $i$.
-The (monotonically increasing) function $\phi(h)$ has units of $\textrm{Hz}$, and describes the spike firing rate of a given neuron as a function of its potential.
-In the absence of recurrent currents, the potential of each neuron decays exponentially with time constant $\tau \,[\textrm{s}]$.
-Neurons can also receive external currents $I^\text{ext}_i(t) \,[\textrm{A}]$.
-Conversion of current to voltage happens through the resistivity factor $R_i \,[\textrm{Ohm}]$.
-@eq-rnn writes out the full equation a rate neural network.
+Early recordings of the cat's somatosensory @Mou57 visual cortices suggested that the cortical sheet is organized in vertical columns of functionally similar neurons.
 
-$$
-\tau \dot h_i(t) = \underbrace{-h_i(t)}_{\text{exponential decay}} + R_i \sum_{j=1}^{N} \underbrace{J_{ij} \phi(h_j(t))}_{I^\text{rec}_i(t)} + R_i I^{\text{ext}}_i(t)
-$$ {#eq-rnn}
+The putative columnar organization of the cortex motivated the design of spatially structured models of neural population dynamics, namely neural field models [@WilCow73;@Nun74;@Ama77], where the spatial dimensions were nothing but the two dimensions of the cortical sheet.
 
-This model can intuitively be seen as the $N \rightarrow \infty$ limit of a network of spiking neurons, in which neurons emit current spikes (modeled as $\delta$-functions) at some time-variable rate $\phi(h_i(t))$.
-The convergence is formally proven in @schmutz2023convergence.
+As models of spatiotemporal neural population dynamics on the cortical sheet, neural field models can be used to explain experimentally observed patterns of cortical activity, such as travelling waves in visual [@SatNau12;@MulRey14], somatosensory [@PetHah03;@FerBol06], motor [@RubRob06;@TakKim15], and hippocampal [@LubSia09;@PatFuj12;@PatSch13] cortices (see @MulCha18 for a review), and are used to model large-scale brain signals, such as electroencephalography (EEG) recordings @Bre17.
+See also @DipRan18 for an example of neural field modelling of 2-photon calcium imaging recording of visual cortex @DipRan18.
 
-In the following, we set $\tau=1$ and $R_i=1$ for all neurons $i$, since these just correspond to rescaling time and currents. Additionally, we consider $I^\text{ext}_i(t)=0$ to simplify the model.
+More recent recording methods, especially in rodents, have provided evidence of low-dimensional organization of neuronal activity that does not depend on the physical locations of neurons on the cortical sheet.
+For example, contrary to cat's visual cortex, the orientation selectivity of pyramidal neurons in rat's visual cortex does not depend on their locations, that is, orientation selectivity is heterogeneous at any given location @OhkChu05.
+Moreover, the activity of interneurons in any small volume of visual cortex is also heterogeneous and seems to be structured by a low-dimensional transcriptomic manifold of fine cell subtypes @BugDuf22.
+The local functional heterogeneity of neuronal activity in cortex challenges the old concept of functionally homogeneous cortical columns and, with it, classical neural field models.
 
-## Convergence to a neural field equation
+While, historically, the spatial dimensions in neural field models have been the two dimensions of the cortical sheet (with the notable exception of the ring models for orientation selectivity in visual cortex @BenBar95 and the head-direction system @Zha96), the ``space'' in neural field models does not need to represent physical space but can represent any suitable abstract embedding space.
+Neural field models with abstract embedding spaces could constitute a generalization of classical neural field theory for cortical networks that are not solely structured by distances of neurons on the cortical sheet.
+
+If we consider neural fields in abstract embedding spaces, which embedding space should we choose?
+More fundamentally, given some spatio-temporal dynamics, is there a ``natural'' choice for the embedding space?
+
+The goal of this Master Thesis is to show that the answer to the second question is highly nontrivial because even the dimensionality of the embedding space, for some given spatio-temporal dynamics, is not clearly defined if we don't specify the regularity (the ``smoothness'') of the neural fields along the spatial dimensions.
+
+In particular, we show that for any two-dimensional neural field equation (that is, with two spatial dimensions), there is a one-dimensional neural field equation on the interval, where the embedding space is simply the interval $[0,1]$, from which the solution to the two-dimensional equation can be fully reconstructed.
+The mapping from the two-dimensional equation to the one-dimensional equation is done via well-known measurable bijections for the square $[0,1]^2$ and the interval $[0,1]$.
+Importantly, these bijections are not (and can't be) diffeomorphisms.
+Also, we show that despite the fact that the connectivity kernel of the one-dimensional equation is not continuous, it is sufficiently regular for its solution to be numerically approximated using standard mesh-based simulations.
+
+Our result shows that the notion of (spatial) dimensionality in neural field dynamics is not well-defined if not associated with a constraint on the regularity of the neural fields (or the connectivity kernel).
+Thereby, this work, by studying a simple toy model, illustrates the importance of the analytic notion of regularity when neural population dynamics over abstract continuous spaces are considered.
+
+This work was inspired by recent mathematical development in mean-field theory involving the theory of graphons \cite{JabPoy21}.
+
+## Neural fields as the limit of spatially-structured networks of neurons
+
+<!-- Valentin Schmutz: Which structure?
+At this point, the reader can't guess that you are talking about spatial structure.
+For the logic of this section, you should first gently introduce the notion of spatial structure.
+Nicole : the papers you give for the introduction should give a good understanding of the spatial structure -->
 
 Neural field equations emerge from the structure of rate neural networks as the number of neurons $N$ becomes large.
-Positions $\vec \xi_i \in \Rp$ are attributed to each neuron, and define the *embedding space*.
+Positions $\vec z_i \in \Rp$ are attributed to each neuron, and define the *embedding space*.
 If the embedding is well-chosen, then as $N \rightarrow \infty$ neurons close in the embedding space tend to have similar potentials $h_i(t)$, such that these tend to a continuum $h(t, \vec z)$, where $\vec z \in \Rp$.
@@
 
-In the litterature, the embedding space is often thought of as being closely related to the physical positions of the neurons @Bressloff2012.
-Rigorous proofs of convergence are still an active area of study @chevallier2018mean and are not the subject of this report.
+In the litterature, the embedding space is often thought of as being closely related to the physical positions of the neurons @Bre11.
+Rigorous proofs of convergence are still an active area of study @CheDua19 and are not the subject of this report.
 
-The results we present here should hold for any neural field, but for the sake of argument we shall consider a low-rank rate neural network, for which there are informal proofs of convergence to a neural field equation  @schmutz2023convergence \[TODO : reformulate this\].
-Low-rank neural networks hold their name from the rank of their connectivity matrix, defined in @eq-lrj.
+## Motivation for a one-dimensional neural field
 
-$$
-J_{ij} = \frac 1N \sum_{\mu=1}^p F_{\mu,i} G_{\mu,j}
-$$ {#eq-lrj}
+[TODO : extra motivation, see Valentin's intro]
 
-In the case that $\vec F_i = (F_{1,i},\cdots,F_{p,i}) \in \Rp$ sample from some $p$-dimensional distribution $\rho$, the vectors $\vec F_i$ define a "natural embedding" of neuron $i$ in $\Rp$.
-When the number of neurons increases, the numeric density of neurons in the embedding approaches the actual probability distribution $\rho$, and we can intuitively see the convergence towards a connectivity kernel $w(\vec z, \vec y) = \sum_{\mu=1}^p z_\mu g_\mu(y_\mu)$.
-
-## A one-dimensional neural field
-
-The paper @jabin2022meanfield provides some motivation for the existence of an equivalent neural field in one dimension.
+The paper @JabPoy21 provides some motivation for the existence of an equivalent neural field in one dimension.
 Their main result is that under some boundedness assumptions of the connectivity, the limiting equation describing the dynamics of the system incorporates only an integral over $[0,1]$.
@@
 
-The idea is that we define a measurable mapping $S : \Rp \mapsto [0,1]$ and a measure $\mu : [0,1] \mapsto \mathbb{R}^+$ on $([0,1], \mathcal B([0,1]))$, such that we can write the change of variables in the integral :
+The idea is that we define a measurable mapping $S : \Rp \mapsto [0,1]$ and a measure $\lambda : [0,1]^2 \mapsto \mathbb{R}^+$, such that we can write the change of variables in the integral :
 
 $$
-\int_{\mathbb{R}^p} w(\vec z, \vec y) \phi(h(t, \vec y)) \rho(\mathrm d \vec y) = \int_{[0,1]} [w(\vec z, \cdot) \phi(h(t, \cdot)) \rho(\cdot)] \circ S^{-1} \; \mathrm d \mu
+\int_{\mathbb{R}^p} w(\vec z, \vec y) \phi(h(t, \vec y)) \rho(\mathrm d \vec y) = \int_{[0,1]} [w(\vec z, \cdot) \phi(h(t, \cdot)) \rho(\cdot) \lambda(\cdot)] \circ S^{-1}(\d \alpha)
 $$
@@
 Each neuron samples the probability distribution $\rho(z_1, \cdots, z_p)$ independantly.
-This case then just reduces to simulating the network of rate neurons, which follows @eq-rnn.
+This case then just reduces to simulating the network of rate neurons.
 In other words, we are simulating the full network of rate neurons, and the simulation converges to the true mathematical neural field equation as $N$ becomes large.
@@
 
-In some sense, we would like to find a mapping that conserves "locality" : populations that are close in $\Rp$ should also be close in $[0,1]$, or at least there shouldn't be "too many large discontinuities".
-@sec-mapping is dedicated to the discussion of these mappings, and gives various numerical demonstrations of the resulting dynamics.
+In some sense, we would like to find a mapping that conserves "locality" : populations that are close in $[0,1]$ should represent nearby populations in $\Rp$, or at least there shouldn't be "too many large discontinuities".
+
+## Summary of contents
+
+@sec-nf introduces a toy model for a neural field in $p$ dimensions, and we discuss interpretations of numerical simulation schemes.
+
+@sec-mappings is dedicated to the discussion of mappings from 2 dimensions to 1 dimension, we give intuitions accompanied by simulations, and discuss in detail how "locality" emerges from these mappings.
+
+Finally, @sec-locality tries to formalize the intuitive results obtained in the previous chapters.
+We build a metric to measure the "discretization error", and provide the key arguments for a proof of convergence of the numerical approximations.
 

### 2023-06-07 manuscript/chapters/2_neural_fields.qmd
id rather not lose my changes now lmao

@@
-# A simple model for a neural field in 2D
+# Neural field toy model and simulations {#sec-nf}
 
@@
 
-### Case study : low-rank networks of neurons
+### Low-rank network of neurons
+
+The results we present in this thesis should hold for any neural field, but for the sake of argument we shall consider a low-rank rate neural network, for which there are informal proofs of convergence to a neural field equation  @schmutz2023convergence \[TODO : reformulate this\].
+Low-rank neural networks hold their name from the rank of their connectivity matrix, defined in @eq-lrj.
+
+$$
+J_{ij} = \frac 1N \sum_{\mu=1}^p F_{\mu,i} G_{\mu,j}
+$$ {#eq-lrj}
+
+In the case that $\vec F_i = (F_{1,i},\cdots,F_{p,i}) \in \Rp$ sample from some $p$-dimensional distribution $\rho$, the vectors $\vec F_i$ define a "natural embedding" of neuron $i$ in $\Rp$.
+When the number of neurons increases, the numeric density of neurons in the embedding approaches the actual probability distribution $\rho$, and we can intuitively see the convergence towards a connectivity kernel $w(\vec z, \vec y) = \sum_{\mu=1}^p z_\mu g_\mu(y_\mu)$.
 
@@
 \begin{aligned}
-\dot h_i(t) = -h_i(t) + \frac 1N \sum_{\mu=1}^p \sum_{j=1}^N F_{\mu i} G_{\mu j} \phi(h_j(t)) \\
-F_{\mu i} = f_\mu(z_{1 i}, \cdots, z_{p i}),\, G_{\mu i} = g_\mu(z_{1 i}, \cdots, z_{p i}),\, \vec{z_i} \stackrel{\text{iid}}{\sim} \rho
+\dot h_i(t) =& -h_i(t) + \frac 1N \sum_{\mu=1}^p \sum_{j=1}^N F_{\mu i} G_{\mu j} \phi(h_j(t)) \\
+F_{\mu i} =& f_\mu(z_{1 i}, \cdots, z_{p i}), \\
+G_{\mu i} =& g_\mu(z_{1 i}, \cdots, z_{p i}), \\
+&\vec{z_i} \stackrel{\text{i.i.d.}}{\sim} \rho
 \end{aligned}
@@
 
+### Gaussian $\mathcal{N}(\vec 0, \mathbb{1}_N)$ low-rank network of neurons
+
+#### Description
+
 our setting is :
@@
 \begin{aligned}
-f_\mu(z_1, \cdots, z_p) = z_\mu \\
-g_\mu(z_1, \cdots, z_p) = \tilde \phi(z_\mu) = \frac{\phi(z_\mu) - \avg{\phi(z_\mu)}}{\mathrm{Var}[\phi(z_\mu)]} \\
-\phi(z) = \frac{1}{1+\mathrm{e}^{-z}} \\
-\vec{z} = (z_1, \cdots, z_p) \sim \rho(z_1, \cdots, z_p) = \prod_{\mu=1}^p \mathcal{N}(z_\mu),\, \mathcal{N}(z) = \frac{1}{\sqrt{2 \pi}} \mathrm{e}^{-\frac 12 z^2} \\
+f_\mu(z_1, \cdots, z_p) &= z_\mu \\
+g_\mu(z_1, \cdots, z_p) &= \tilde \phi(z_\mu) = \frac{\phi(z_\mu) - \avg{\phi(z_\mu)}}{\mathrm{Var}[\phi(z_\mu)]} \\
+\phi(z) &= \frac{1}{1+\mathrm{e}^{-z}} \\
+\vec{z} &= (z_1, \cdots, z_p) \sim \rho(z_1, \cdots, z_p) = \prod_{\mu=1}^p \mathcal{N}(z_\mu),\\
+&\text{where}\ \mathcal{N}(z) = \frac{1}{\sqrt{2 \pi}} \mathrm{e}^{-\frac 12 z^2}
 \end{aligned}
@@
 
-study of fixed points
+#### Study of fixed points and their stability
 
@@
 
-[TODO FIG : draw a flow field of in the latent space]
-[TODO FIG : numerical eigenvalues plot]
+![spectrum of $K$ for $N=64$](figures/spectrum_N=64.png)
 
@@
 
-![simulation in 2D, p=2](figures/embedding_d=2_h0=0.mp4)
-
-[TODO FIG : scatterplot of the network, with animation]
+![simulation in 2D, p=2. TODO redo this one with only scatter and latent-](figures/embedding_d=2_h0=0.mp4)
 
@@
 
-### Direct simulation of the low-rank neural field
+introduce the neural field as a "smooth fitting of the scatterplot"
 
-Introduce the neural field in 1D, then 2D
+eq of neural field
+
+### Simulations of the low-rank neural field
+
+we use the RK4 integration scheme, and the integral is decoupled into product of 1D integrals, each is estimated using a dot product
+
+#### Sampling method (Monte-Carlo integration)
 
@@
 \begin{aligned}
-\partial_t h(t, \vec z) &= -h(t, \vec z) + \int_{\mathbb{R}^p} w(\vec z, \vec y) \phi(h(t, \vec y)) \rho(\mathrm d \vec y),
+\partial_t h(t, \vec z) &= -h(t, \vec z) + \int_{\Rp} w(\vec z, \vec y) \phi(h(t, \vec y)) \rho(\mathrm d \vec y),
 \quad
@@
 
-simulation strategies : grid and sampling
+simulation : when we simulate a network of neurons with $N \gg 1$, we are approximating the integral using a monte carlo method.
 
-[TODO : show the connectivity with grid sampling. mention that the problems with the compact support get adressed later in the next chapter, when we do $\R^2 \mapsto [0,1]^2$]
+Monte-carlo estimation of the integral is
 
-## Characterizing dynamics
+$$
+\begin{aligned}
+\mathcal{I}(\vec z) &= \int_{\Rp} \underbrace{w(\vec z, \vec y) \phi(h(t, \vec y))}_{I(\vec z, \vec y)} \rho(\d \vec y) \\
+\hat{\mathcal{I}}(\vec z) &= \frac 1N \sum_{i=1}^N I(\vec z, \vec{y_i}), \, \vec{y_i} \stackrel{\text{i.i.d.}}{\sim} \rho
+\end{aligned}
+$$
+
+Numerically, we are estimating $\hat{\mathcal{I}}(\vec z)$, and we can show the convergence rate is in $\bO(1/\sqrt{N})$, since
+
+$$
+\begin{aligned}
+\mathrm{Var}\left[\hat{\mathcal{I}}(\vec z) \right] &= \frac 1N \mathrm{Var}_{\vec y \sim \rho}[I(\vec z, \vec y)]
+\end{aligned}
+$$
+
+Detail : strictly speaking, MC samples new points every timestep, here we sample once at t=0, then propagate these points
+
+#### Grid method (Trapezoidal integration)
+
+This method describes the trapezoidal method of integration.
+We state the result that the convergence rate is $\bO(N^{-2/p})$.
+For $p \geq 4$, MC shows better scaling behavior (but prefactors of error can change which method is better).
+
+In our case, since the distribution $\rho$ can be factorized, we can decouple the $p$-dimensional integral into the product of $p$ one-dimensional integrals.
+Then the error estimation of the integral scales as $\bO(N^{-2})$.
+For this reason, we use the grid method for the remainder of this thesis.
+
+## Characterizing dynamics of the low-rank neural field
 
@@
 
-[TODO : i'm not even sure that introducing the overlaps in necessary. we might just include them for completeness, and because they are relevant in papers on low-rank stuff]
+[TODO : i'm not even sure that introducing the overlaps in necessary.
+we might just include them for completeness, and because they are relevant in papers on low-rank stuff.
+in the rest, just use the kappa for the latent space]
+
+[TODO : do not expand mathematically, say this is well-known and given for context. the fixed point and stability study is new though]
 
 ### Low-rank neural fields as $p$-dimensional closed systems
+
+cite @veltz2009localglobal for this.
+
+on explicite les equations du paper dans notre cas
+
+introduce this a just the projection of the field on the fixed points
 
@@
 
-with these definition, we have in our case
+with these definitions, we have in our case
 $$
@@
 
-overlaps as an approximation of the “dynamics” (latent space), in the case $\phi=\mathrm{linear}$.
-since in our case $\phi$ is a sigmoid, it can be considered linear in the range $h \sim \pm 3$.
-we therefore use the computed overlaps as to show the latent space
+and the orthogonal component evolves according to
 
-[TODO MATH : actually we could also juste compute the kappa as a function of time, and study that]
+$$
+h^\perp(t, \vec z) = h^\perp(0, \vec z) \mathrm e^{-t},
+\quad
+h^\perp(0, \vec z) = h(0, \vec z) - \sum_{\mu=1}^p \kappa_\mu(0) z_\mu
+$$
 
-### Dynamics in the overlap latent space
+todo : put accent on the fact the RHS can be computed using all the kappas, kappa -> h -> integral = m -> RHS
+
+conceptually : do not forget that the kappa is only relevant in this particular toy model, in general we do not have these kappa, and the study of low-dim fields is interesting.
+If we always had the kappa, then of course we can only look at the kappa, since much simpler (a 1D manifold (line) in a Rp space, instead of a Rp volume in a Rp space).
 
@@
 
-Relation between $p$-dimensional closed system and neural field equation in $\mathbb R^p$
-
+<!-- Relation between $p$-dimensional closed system and neural field equation in $\mathbb R^p$
 $$
@@
 \end{aligned}
-$$
+$$ -->
+
+overlaps as an approximation of the “dynamics” (latent space), in the case $\phi=\mathrm{linear}$.
+since in our case $\phi$ is a sigmoid, it can be considered linear in the range $h \sim \pm 3$.
+we therefore use the computed overlaps as to show the latent space
 
@@
 
+<!-- [TODO FIG : draw a flow field of in the latent space. -> actually, might be difficult when there is delay] -->
+
 ## A cycling neural field
+
+### Intuition behind the cycling
 
@@
 
+![Intuition for the cycling behavior by plotting the recurrent currents](figures/cycling_recurrent.png)
+
 note this is now a DDE instead of a PDE
+
+![](figures/embedding_d=2_h0=0_cycling_delta=6.mp4)
+
+### Cycling behavior of the latent variables
 
@@
 
-then one can repeat this for the next cycles. basically each cycle is applying an exponential convolution filter
+then one can repeat this for the next cycles. basically each cycle is applying an exponential convolution filter [TODO : trim this, don't go into detail]
 
-[TODO MATH : the derivation of this, we have an exponential convolution filter]
+<!-- [TODO MATH : the derivation of this, we have an exponential convolution filter] -->
 
-![](figures/kappa_cycling_p=2.tmp.png)
+![simulate cycling in p=2](figures/kappa_cycling_p=2.tmp.png)
 
-simulate cycling in p=2
+## Mapping $\R^2$ to $[0,1]^2$ using the CDF
 
-[TODO : might be good to show a simulation where we explicitly simulate $\kappa$ ? but the goal here is just to "prove" the cycling behavior]
+<!-- ### Using the CDF -->
 
-![](figures/embedding_d=2_h0=0_cycling_delta=6.mp4)
+later we will map [0,1]² -> [0,1], so we first want to map R² -> [0,1]².
 
-![Intuition for the cycling behavior by plotting the recurrent currents](figures/cycling_recurrent.png)
-
-## From $\R^2$ to $[0,1]^2$
+there are many ways to do this using functions R -> [0,1], but a practical choice is the inverse CDF, which has the added benefit of absorbing the density in the kernel, in the sense that the integral over [0,1]² becomes "weighted" by a uniform distribution (lesbesgue measure $\lambda$)
 
@@
 
-![](figures/sampling_R2_01.png)
+![samples in $\Rp$ and corresponding samples in $[0,1]$](figures/sampling_R2_01.png)
 
-[TODO : does this deserve its own section ? maybe mention it later, we can give a picture of the "isolines" of the 2D gaussian CDF]
+give a picture of the "isolines" of the 2D gaussian CDF
 
-## Numerical simulation of neural fields
+furthermore numerically, mapping R² -> [0,1]² adresses the problem of "putting a finite grid on the infinite R²".
+(previously we would have had to make the approximation of compact support, then sample the PDF on the grid and numerically renormalize.)
 
-### Direct sampling
+<!-- ### Consequences on simulations
+
+#### Direct sampling
 
@@
 
-### Bin averaging samples
-
-![](figures/sampling_R2_nogrid_kernel.png)
-
-### Grid over $[0,1]^2$
+#### Grid over $[0,1]^2$
 
@@
 
-Doing this is equivalent to the formulation of the binned connectivity matrix.
+Doing this is equivalent, but more memory-efficient to the formulation of the binned connectivity matrix.
 
-![](figures/sampling_R2_01_kernel.png)
+![](figures/sampling_R2_01_kernel.png) -->

### 2023-06-07 manuscript/chapters/3_mappings.qmd
id rather not lose my changes now lmao

@@
-# Mapping the neural field in 2D to a 1D embedding
+# Mapping neural fields to lower-dimensional embeddings {#sec-mappings}
 
@@
 
-## Simple example : population dynamics
+## Population dynamics : an introductory example of a 1D embedding
 
-[TODO MOTIVATION : is this really useful ? the point i'm trying to make here is the embedding problem arises when you have an infinite number]
+the case of population dynamics can easily be embedded in 1D (this is the trivial case)
 
-the case of population dynamics can easily be embedded in 1D
-
-$$
-\begin{aligned}
+<!--
+$$\begin{aligned}
 \rho(\vec z) &= \sum_{a \in P} p_{a} \delta(\vec{z_a} - \vec{z}) \\
 \text{normalization : } \int_{\Rp} \rho(\d \vec{z}) &= \sum_{a} p_a = 1
-\end{aligned}
-$$
+\end{aligned}$$
+we can explicitly formulate the embedding with a new 1D PDF (arrange the populations on the line, their probability is their segment length)
+$$\gamma(\alpha) = \sum_{a \in P} p_a \mathbb{1}(C_{a-1} \leq \alpha < C_{a}),\, C_a=\sum_{b=1}^a p_b$$
+the connectivity matrix in the 1D model becomes
+$$J_{ab} = \sum_{\mu=1}^p f_\mu(\vec{z_a}) g_\mu(\vec{z_b})$$
+[TODO MATH : flesh this out]
+[TODOSIM : would this be interesting to simulate ?]
+case 2 populations, we get the ising model (binary patterns) -->
 
-we can explicitly formulate the embedding with a new 1D PDF (arrange the populations on the line, their probability is their segment length)
-
-$$
-\gamma(\alpha) = \sum_{a \in P} p_a \mathbb{1}(C_{a-1} \leq \alpha < C_{a}),\, C_a=\sum_{b=1}^a p_b
-$$
-
-the connectivity matrix in the 1D model becomes
-
-$$
-J_{ab} = \sum_{\mu=1}^p f_\mu(\vec{z_a}) g_\mu(\vec{z_b})
-$$
-
-[TODO MATH : flesh this out]
-
-case 2 populations, we get the ising model (binary patterns)
-
-the point i try to make is that for discrete distributions (i can enumerate the populations), it is easy to embed in 1D. the problem comes when the distribution is continuous.
+TODO : a nice drawing with motivation
 
@@
 
-## Neural field in $[0,1]$
+the point i try to make is that for discrete distributions (i can enumerate the HOMOGENEOUS populations of neurons), it is easy to embed in 1D.
+the problem comes when the distribution is continuous -> we have an infinite amount of homogeneous populations.
 
-Define a measurable and bijective mapping $S : \mathbb{R}^p \rightarrow [0, 1]$.
+we know how to map finite populations to [0,1], we now need a mapping that is good when the number of populations tends to infty
+
+idea : we need a sequence of mappings ("permutations", this expresses J perm-invar) acting on a finite number of populations, that has a "well defined convergence"
+
+## Defining the neural field in $[0,1]$
+
+Define a measurable and bijective mapping $S : [0,1]^2 \rightarrow [0, 1]$.
 
@@
 
-We define $\tilde h(t, \alpha) = h_U(t, S^{-1}(\alpha))$, $\tilde w(\alpha, \beta) = w_U(S^{-1}(\alpha), S^{-1}(\beta))$
+We define $\tilde w(\alpha, \beta) = w_U(S^{-1}(\alpha), S^{-1}(\beta))$ + intial condition
+
+<!-- $\tilde h(t, \alpha) = h_U(t, S^{-1}(\alpha))$ -->
 
 $$
-\partial_t \tilde h(t, \alpha) = -\tilde h(t, \alpha) + \int_{[0,1]} \tilde w(\alpha, \beta) \tilde\phi(h(t, \beta)) \lambda \circ S^{-1}(\d \beta)
+\partial_t \tilde h(t, \alpha) = -\tilde h(t, \alpha) + \int_{[0,1]} \tilde w(\alpha, \beta) \phi(\tilde h(t, \beta)) \lambda \circ S^{-1}(\d \beta)
 $$
 
-![](figures/mappings_h1d_state.png)
-
-![](figures/mapping_recursive_quadrant.mp4)
+[TODO FIG : a nice drawing of this, also to give an intuition for $\lambda$]
 
@@
 
-### Types of mappings
+cite @kharazishvili_strange_2017 peano vs cantor mappings
 
-peano vs cantor mappings
+### Continuous mappings
 
@@
 
-We consider $(x,y) \in [0,1]²$. Given $n \in \mathbb N$, we write a (truncated) binary expansion of $x$ as (and similarly for $y$)
+> there is no homeomorphism
+  (continuous bijective function)
+  between [0,1]² and [0,1]
+
+### Bijective mappings
+
+We consider $(v_1, v_2) \in [0,1]^2$. Given $n \in \mathbb N$, we write a (truncated) binary expansion of $v_1$ and $v_2$.
 
 $$
-b^x_k = \mathrm{Ind}\left\{2^{k-1}x - \lfloor2^{k-1}x\rfloor \geq \frac{1}{2}\right\}, \text{ such that } x= \sum_{k=1}^{n} b^x_k 2^{-k} = 0.b^x_1b^x_2\cdots b^x_{n}
+v_1=\sum_{l=1}^{n} b^1_l 2^{-l} = 0.b^1_1b^1_2\cdots b^1_{n} \iff b^1_l = \mathrm{Ind}\left\{2^{l-1}v_1 - \lfloor2^{l-1}v_1\rfloor \geq \frac{1}{2}\right\}
 $$
+
+The intuition behind this expansion is we take a segment, we split it in two, write 0 to go left, write one to go right, then recurse.
+Then combine the bits of $v_1$ and $v_2$ to get the position in the $[0,1]^2$ embedding.
+
+[TODO FIG : splitting the segment recursively locates the position of the component]
+
+We define the mapping $S^n$ that takes the $n$-bit truncations of $v_1$ and $v_2$ as input, and outputs a $2n$ bit truncation of $\alpha=S(v_1, v_2)$, where $S$ is the $n\to\infty$ limit of the sequence $S^n$.
+
+$$
+\begin{alignat}{2}
+S^n : &\{0, \tfrac{1}{2^n}, \cdots, 1-\tfrac{1}{2^n}\}^2 &&\mapsto \{0, \tfrac{1}{4^n}, \cdots, 1-\tfrac{1}{4^n}\} \\
+&(0.b^1_1 b^1_2 \cdots b^1_n, 0.b^2_1 b^2_2 \cdots b^2_n) &&\mapsto 0.b_1 b_2 \cdots b_{2n}
+\end{alignat}
+$$
+
+In general, each bit $b^k$ of the image would be a binary function of the $n$ bits of $v_1$ and the $n$ bits of $v_2$, implicitly $b_k = b_k(b^1_1, b^1_2, \cdots, b^1_n, b^2_1, b^2_2, \cdots, b^2_n)$.
+With this, we are able to formulate any mapping of $4^n$ distinct elements (here : squares of size $2^{-n} \times 2^{-n}$) onto $4^n$ elements (here : segments of length $4^{-n}$).
+
+The intuition behind the use of $2n$ bits for the image is that we take $2^n \times 2^n = 4^n$ squares in $[0,1]^2$ and map those to $4^n$ segments in $[0,1]$.
+In an informal way, we map $n+n$ bits onto $2n$ bits.
 
@@
 
-[TODO : do we mention the random mapping ?]
+In the following, we will restrict ourselves to mappings where the functions $b_k(\{b^1_l\}_{l\in\{1,\cdots,n\}}, \{b^1_l\}_{l\in\{1,\cdots,n\}})$ are defined in such a way that each bit of the inputs is used once, and only once.
+Therefore, for each bit $k$ there is a function $b_k(b^{\mu_k}_{l_k})$ acting on only one bit.
+Since this function is a binary, the only allowed definitions are $b_k(b^{\mu_k}_{l_k}) = b^{\mu_k}_{l_k}$ (identity) or $b_k(b^{\mu_k}_{l_k}) = 1-b^{\mu_k}_{l_k}$ (inversion).
+We furthermore restrict the formulation by allowing only identity functions[^inversion_restriction].
 
-#### Z-mapping
+[^inversion_restriction]: This last restriction is not strictly necessary for the following, however it helps to have a simple model to build a good understanding of the problem
 
-Also called Z-order curve, Morton mapping. It conserves locality 2D -> 1D.
+All this mathematical formalism expresses, in essence, that the mapping $S^n$ is just a *reordering* of the input bits.
+The geometric view of this is that $S^n$ takes $4^n$ points embedded in $[0,1]^2$, and reorders them onto $4^n$ points in $[0,1]$.
+Therefore, there is a one-to-one mapping between the points on the square and the points on the segment, and the mapping $S^n$, with the restrictions is bijective.
+
+[TODO FIG : mapping of points to segment]
+
+These simplifications might seem too restrictive, however we will show in the rest of this thesis that this formulation is flexible enough to express mapping which have good properties such as "locality", all while being simple enough to help build an intuition.
+
+Random mapping. mentionned in the introduction. basically a random permutation of bits. not even stable formulation for fixed $n$.
+
+Anti-Z mapping. Also called in code "Far mapping" (initially I didn't see the link with the Z-order curve), because it destroys locality 2D -> 1D.
 
 $$
-z = Z(x,y) = 0.b^x_1 b^y_1 b^x_2 b^y_2 \cdots b^x_n b^y_n = \sum_{k=1}^{n} b^x_k 2^{1-2k} + b^y_k 2^{-2k}
+\alpha = A(v_1,v_2) = 0.b^1_n b^2_n b^1_{n-1} b^2_{n-1} \cdots b^1_1 b^2_1 = \sum_{k=1}^{n} b^1_{n+1-k} 2^{1-2k} + b^2_{n+1-k} 2^{-2k}
 $$
@@
 \begin{aligned}
-x = Z^{-1}_x(z) &= 0.b_1 b_3 \cdots b_{2n-1} = \sum_{k=1}^n b_{2k-1} 2^{-k} \\
-y = Z^{-1}_y(z) &= 0.b_2 b_4 \cdots b_{2n} = \sum_{k=1}^n b_{2k} 2^{-k}
+v_1 = A^{-1}_1(\alpha) &= 0.b_{2n-1} b_{2n-3} \cdots b_{1} = \sum_{k=1}^n b_{2(n-k)-1} 2^{-k} \\
+v_2 = A^{-1}_2(\alpha) &= 0.b_{2n} b_{2n-2} \cdots b_{2} = \sum_{k=1}^n b_{2(n+1-k)} 2^{-k}
 \end{aligned}
@@
 
-note that in other definitions (and in code) it is defined by $0.b^y_1 b^x_1 b^y_2 b^x_2 \cdots$. This difference is not fundamentally important though.
+it does not converge
 
-#### Column mapping
-
-Also called "Reshape mapping" in the code. It converges to a projection on the $x$-axis
+Column mapping. Also called "Reshape mapping" in the code. It converges to a projection on the $x$-axis
 
 $$
-z = C(x,y) = 0.b^x_1 b^x_2 \cdots b^x_n b^y_1 b^y_2 \cdots b^y_n
+\alpha = C(v_1,v_2) = 0.b^1_1 b^1_2 \cdots b^1_n b^2_1 b^2_2 \cdots b^2_n
 $$
@@
 \begin{aligned}
-x = C^{-1}_x(z) &= 0.b_1 b_2 \cdots b_n = \sum_{k=1}^n b_{k} 2^{-k} \\
-y = C^{-1}_y(z) &= 0.b_{n+1} b_{n+2} \cdots b_{2n} = \sum_{k=1}^n b_{n+k} 2^{-k}
+v_1 = C^{-1}_1(\alpha) &= 0.b_1 b_2 \cdots b_n = \sum_{k=1}^n b_{k} 2^{-k} \\
+v_2 = C^{-1}_2(\alpha) &= 0.b_{n+1} b_{n+2} \cdots b_{2n} = \sum_{k=1}^n b_{n+k} 2^{-k}
 \end{aligned}
@@
 
-#### Anti-Z mapping
-
-Also called in code "Far mapping" (initially I didn't see the link with the Z-order curve), because it destroys locality 2D -> 1D.
+Z-mapping. Also called Z-order curve, Morton mapping. It conserves locality 2D -> 1D.
 
 $$
-z = A(x,y) = 0.b^x_n b^y_n b^x_{n-1} b^y_{n-1} \cdots b^x_1 b^y_1 = \sum_{k=1}^{n} b^x_{n+1-k} 2^{1-2k} + b^y_{n+1-k} 2^{-2k}
+\alpha = Z(v_1,v_2) = 0.b^1_1 b^2_1 b^1_2 b^2_2 \cdots b^1_n b^2_n = \sum_{k=1}^{n} b^1_k 2^{1-2k} + b^2_k 2^{-2k}
 $$
@@
 \begin{aligned}
-x = A^{-1}_x(z) &= 0.b_{2n-1} b_{2n-3} \cdots b_{1} = \sum_{k=1}^n b_{2(n-k)-1} 2^{-k} \\
-y = A^{-1}_y(z) &= 0.b_{2n} b_{2n-2} \cdots b_{2} = \sum_{k=1}^n b_{2(n+1-k)} 2^{-k}
+v_1 = Z^{-1}_1(\alpha) &= 0.b_1 b_3 \cdots b_{2n-1} = \sum_{k=1}^n b_{2k-1} 2^{-k} \\
+v_2 = Z^{-1}_2(\alpha) &= 0.b_2 b_4 \cdots b_{2n} = \sum_{k=1}^n b_{2k} 2^{-k}
 \end{aligned}
@@
 
-it does not converve
+note that in other definitions (and in code) it is defined by $0.b^2_1 b^1_1 b^2_2 b^1_2 \cdots$. This difference is not fundamentally important though.
 
-TODO : 3D vis of the mappings
+[TODO FIG : 3D vis of the mapping]
 
@@
 
-intuition for reshape, and the fractals
+the sequence in [0,1] traces a path in [0,1]²
 
-expose different types of mapping
+intuition for column.
 
-argue over binary expansion, and especially how the fact that
+the limit of pointwise convergent mappings and measurable (which they are) is measurable.
 
-* it converges forces the least significant bits of $z_1$ and $z_2$ to go to the lsb of $\alpha$
-* it's a bijection means "both bits of x and y" need to be used in the expansion
+argue over binary expansion, and especially how the fact that it converges forces the least significant bits of $z_1$ and $z_2$ to go to the lsb of $\alpha$.
+this imposes some structure of what mappings are good
+
+z-mapping can in this context be seen as a "canonical representation" of the pointwise convergent, bijective, local mapping.
 
@@
 
-go into detail on how numerically S is a reordering of the bins
-
 what we are doing in reality is assigning indices to squares of size $2^{-n} \times 2^{-n}$, and mapping them to indices representing segments of length $4^{-n}$
 
-[TODO FIG : drawing of the square-to-segment mapping, something like
+[TODO FIG : finite n is population dynamics. drawing of the square-to-segment mapping, something like
 
@@
 x has n bits, y has n bits, z has 2n bits.
+
 an arbitrary mapping can then be written as a recombination of the bits of x and y (and any other constant) into 2n bits
@@
 locality 1D -> 2D is much more important (a neuron in 1D must refer to only one (not multiple) neurons in 2D)
+
+![](figures/mappings_h1d_state.png)
+
+![simulation of 2D field, and mapping of the dynamics to the 1D field](figures/mapping_recursive_quadrant.mp4)
 
@@
 
-[^1] : I am glossing over the details of how we bin the grid of $4^n$ elements in $[0,1]^2$ space.
+[^1]: I am glossing over the details of how we bin the grid of $4^n$ elements in $[0,1]^2$ space.
 Numerically, there is an order in which we enumerate each of the $4^n$ bins (the most obvious is column-wise enumeration, i.e.
@@
 
-[^2] : The reason for which we take bins of size $2^n$, is because $2^n$ is the maximum size a bin can have which spans in only one direction in the 2D embedding.
+[^2]: The reason for which we take bins of size $2^n$, is because $2^n$ is the maximum size a bin can have which spans in only one direction in the 2D embedding.
 For instance, with the ``ReshapeMapping`` a column contains $2^n$ populations, all with the same $\xi_0$ position : visually the bin is vertical.
@@
 
-[^3] : The connectivity $\tilde J_{\alpha\beta}$ in 1D can be obtained by computing the "average low-rank vectors". See ``draft.md`` for more details.
+[^3]: The connectivity $\tilde J_{\alpha\beta}$ in 1D can be obtained by computing the "average low-rank vectors". See ``draft.md`` for more details.
 
-[^4] : Each point in $(m_0,m_1)$ fully encodes the state of a linear low-rank RNN.
+[^4]: Each point in $(m_0,m_1)$ fully encodes the state of a linear low-rank RNN.
 We can prove (see my own handwritten notes) that $m_\mu(t) \approx \kappa_\mu(t)$ when $\phi(h)$ is close to linear.
@@
 
+### Reinterpretation of sampling and binning methods
+
+reintrpretation of the simulation methods
+
+average low-rank vectors
+
+derivation of $\Jab$ and mean patterns
+
+For finite number of recursive quadrant iterations $n$, we can do a "mean-field approximation" inside each of the $4^n$ segments.
+Let $\alpha = \{i_1,\cdots,i_{|\alpha|}\}$ be the multi-index corresponding to all neurons of which the embedding in $\mathbb R^p$ gets mapped to the segment $\alpha$ in $[0,1]$.
+Let $H_\alpha(t) = \frac 1 {|\alpha|} \sum_{i \in \alpha} h_i(t)$ be the (mean) RNN potential of the segment $\alpha$.
+The connectivity matrix $\tilde J_{\alpha,\beta}$ satisfies
+
+$$
+\dot H_\alpha(t) = -H_\alpha(t) + \sum_{\beta \in \text{segments of length } 4^{-n}} \tilde J_{\alpha,\beta} \phi(H_\beta(t))
+$$
+
+By substituting the original $h_i(t)$, we find the correct rescaling is given by
+
+$$
+\tilde J_{\alpha,\beta} = \frac 1 {|\alpha|} \sum_{i \in \alpha} \sum_{j \in \beta} J_{ij}
+$$
+
+In the low-rank case, we have
+
+$$
+J_{ij} = \frac 1N \sum_{\mu=1}^p F_{\mu,i} G_{\mu,j}
+$$
+
+We can define the "mean pattern" inside each bin :
+
+$$
+\tilde F_{\mu,\alpha} = \frac{1}{|\alpha|} \sum_{i \in \alpha} F_{\mu,i}, \; \tilde G_{\mu,\alpha} = \frac{1}{|\alpha|} \sum_{i \in \alpha} G_{\mu,i}
+$$
+
+Then the connectivity matrix is given by, noting $N = \sum_{\alpha} |\alpha|$,
+
+$$
+\tilde J_{\alpha,\beta} = \frac{|\beta|}{\sum_{\beta'} |\beta'|} \left( \sum_{\mu=1}^p \tilde F_{\mu,\alpha} \tilde G_{\mu,\beta} - \delta_{\alpha,\beta} \underbrace{\sum_{\mu=1}^p \sum_{i \in \alpha} \frac{F_{\mu,i}}{|\alpha|} \frac{G_{\mu,i}}{|\alpha|}}_{\gamma_{\alpha}} \right)
+$$
+
+![](figures/sampling_R2_nogrid_kernel.png)
+
+Grid method : just do $\tilde w = w_U \circ S^{-1}$
+
+## Simulations of neural fields in 3D, 2D and 1D
+
+### Application of $S$ : from $[0,1]^2$ to $[0,1]$
+
 ![](figures/coarsegraining.png)
+
+### Iterative application of $S$ : from $[0,1]^3$ to $[0,1]$
+
+#### Continuity on one axis and fractal on the other : from $[0,1]^3$ to $[0,1]^2$
+
+#### From fractal to fractal : $[0,1]^2$ to $[0,1]$
+
+![Effective sampling points after averaging](figures/mapping_3d_2d_1d.png)
+
+![Equivalence of dynamics, from 3D to 2D to 1D](figures/embedding_3d_2d_1d.mp4)

### 2023-06-07 manuscript/chapters/4_locality.qmd
id rather not lose my changes now lmao

@@
 
-# Locality and numerical convergence
+# Locality and numerical convergence {#sec-locality}
 
@@
 
-composition of functions
+### The case $S$ measurable and bijective
 
-bijectivity (actually injection in one sense)
+In the previous sections, it has been obtained that a sufficient criterion for equivalent dynamics in the 2D and the 1D embedding, is that the mapping $S$ is bijective (or at least, injective from 2D in 1D, but we will use the slightly stronger assumption of bijectivity).
 
-TODO : write the proof of this
+We furthermore assume that the mapping $S$ is measurable.
+This is not a strong assumption, since the limit of pointwise convergent measurable functions is also measurable.
+In our case, we have written finite-precision expansions $S^n$, which by being sums of piecewise constant functions, are also measurable.
+Measurability of $S$ by definition allows us to write $\lambda \circ S^{-1}$ [TODO MATH : write this better.
+the inverse image of a set is measurable.], which defines a measure of the 1D space by first going to 2D space, then taking the measure there.
+
+[TODO FIG : $\lambda \circ S^{-1}$ defines a measure in 1D by going through 2D first ![](figures/Screenshot_2023-06-05_10-12-36.png)]
+
+With this in mind, let us prove that by defining the initial condition $\tilde h(t=0, \alpha) = h_U(t=0, S^{-1}(\alpha))$ and connectivity $\tilde w(\alpha, \beta) = w_U(S^{-1}(\alpha), S^{-1}(\beta)$, then the solution $h_U(t, \vec v)$ of the neural field equation in $[0,1]^2$ uniquely defines the solution of the neural field equation in $[0,1]$ by writing $\tilde h(t, \alpha) = h_U(t, \cdot) \circ S^{-1}(\alpha)$.
+
+We write the neural field equation in 1D and the initial condition.
+
+$$
+\begin{aligned}
+\partial_t \tilde h(t, \alpha) &= -\tilde h(t, \alpha) + \int_{[0,1]} \tilde w(\alpha, \beta) \phi(\tilde h(t, \beta)) \lambda \circ S^{-1}(\d \beta) \\
+\tilde h(t=0, \alpha) &= h_U(t=0, S^{-1}(\alpha))
+\end{aligned}
+$$
+
+We now substitute $\tilde h(t) = h_U(t) \circ S^{-1}$, and write $\vec{v} = S^{-1}(\alpha),\, \vec{u} = S^{-1}(\beta)$.
+
+$$
+\begin{aligned}
+\partial_t h_U(t, S^{-1}(\alpha)) &= -h_U(t, S^{-1}(\alpha)) + \int_{[0,1]} w(S^{-1}(\alpha), S^{-1}(\beta)) \phi(h_U(t, S^{-1}(\beta))) \lambda \circ S^{-1}(\d \beta) \\
+\iff \partial_t h_U(t, \vec v) &= -h_U(t, \vec v) + \int_{[0,1]} w(\vec v, S^{-1}(\beta)) \phi(h_U(t, S^{-1}(\beta))) \lambda \circ S^{-1}(\d \beta) \\
+&= -h_U(t, \vec v) + \int_{[0,1]^2} w(\vec v, \vec u) \phi(h_U(t, \vec u)) \lambda(\d \vec u)
+\end{aligned}
+$$
+
+The measure $\lambda : [0,1]^2 \mapsto \R^+$ simply plays the role of a uniform distribution on $[0,1]^2$ [TODO MATH : $\lambda$ is the lebesgue measure ?], and we simply write
+
+$$
+\partial_t h_U(t, \vec v) = -h_U(t, \vec v) + \int_{[0,1]^2} w(\vec v, \vec u) \phi(h_U(t, \vec u)) \d \vec u
+$$
+
+Therefore, if $h_U(t, \vec v)$ solves the neural field equation in $[0,1]^2$, then $\tilde h(t) = h_U(t) \circ S^{-1}$ solves the equivalent neural field equation in $[0,1]$.
+
+[TODO MATH : do i need to state a theorem of existence and uniqueness of the solution ?]
+
+### Can the bijectivity condition be relaxed ?
+
+[TODO : go into more detail on the injectivity, surjectivity.
+how the fact that $S$ is injective forces us to have that we can't map 2 neurons to the same positino.
+surjecivity is not that important, we could duplicate a neuron and adjust the weights]
 
@@
 We now want to prove that evaluating the integrand at $\beta_i = \tfrac{i-1}{2^n}$ gives a good approximation of the integral.
-Since $\int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \lambda \circ S^{-1}(\d \beta) = \tfrac{1}{2^n}$ [TODO MATH : is this true ?], we write
+Since $\int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \lambda \circ S^{-1}(\d \beta) = \tfrac{1}{2^n}$ [TODO MATH : is this true ?
+this is obvious when we consider the binary expansion of $S^{-1}$], we write
 
@@
 We express this regularity by the added assumption that $w_U(\cdot, \vec z) h_U(t, \vec z)$ is Lipschitz in $\vec z$, with Lipschitz constant $L(\cdot)$, and $\norm{\cdot}$ the corresponding norm on $[0,1]^2$.
+[TODO : argue on the niceness of $I_rec$, add $\phi$]
 

### 2023-06-07 notebooks/neurodyn/_overlap.py

@@ -4,7 +4,7 @@ import numpy as np
 import warnings
 from ._rnn import LowRankRNN, BinMappedRNN, DenseRNN
 
-__all__ = ['overlap']
+__all__ = ['overlap', 'projection']
 
 def overlap(rnn: LowRankRNN | BinMappedRNN | DenseRNN, h: np.ndarray) -> np.ndarray:
 	"""Compute overlap
@@ -27,5 +27,30 @@ def overlap(rnn: LowRankRNN | BinMappedRNN | DenseRNN, h: np.ndarray) -> np.ndar
 	elif isinstance(rnn, DenseRNN):
 		warnings.warn(f'attempting to compute overlap on a DenseRNN {rnn}, unless rnn.G is set, this will probably not work')
 		return np.einsum('im,i...->m...', rnn.G, rnn.phi(h)) / rnn.N
+	else:
+		raise RuntimeError(f'cannot compute overlap on {rnn} of type {type(rnn)}')
+
+
+def projection(rnn: LowRankRNN | BinMappedRNN | DenseRNN, h: np.ndarray) -> np.ndarray:
+	"""Compute the projection of the neural field onto the fixed point
+
+	Parameters
+	----------
+	rnn : LowRankRNN | BinMappedRNN
+	h : np.ndarray
+		neuron potentials, shape (N, T)
+
+	Returns
+	-------
+	np.ndarray
+		overlap with patterns, shape (p, T)
+	"""
+	if isinstance(rnn, LowRankRNN):
+		return np.einsum('im,i...->m...', rnn.F, h) / rnn.N
+	elif isinstance(rnn, BinMappedRNN):
+		return np.einsum('am,a...,a->m...', rnn.F, h, rnn.bincounts) / rnn.bintotal
+	elif isinstance(rnn, DenseRNN):
+		warnings.warn(f'attempting to compute overlap on a DenseRNN {rnn}, unless rnn.G is set, this will probably not work')
+		return np.einsum('im,i...->m...', rnn.F, h) / rnn.N
 	else:
 		raise RuntimeError(f'cannot compute overlap on {rnn} of type {type(rnn)}')

### 2023-06-11 manuscript/chapters/3_mappings.qmd
hehe thats a lot of changes

@@
 {{< include ../macros.tex >}}
+
+```{python}
+import matplotlib as mpl, matplotlib.pyplot as plt, matplotlib.colors as mplcolors, matplotlib.patheffects as pe, matplotlib.ticker as ticker
+import numpy as np
+import sys
+sys.path.append('../../notebooks')  # import neurodyn
+from neurodyn import *
+from matplotlib_tufte import *
+setup()
+
+mpl.use('svg')
+```
 :::
+
+In the previous chapter, we have established a simple toy model for a neural field in $p$ dimensions, and have shown how it can be embedded in $[0,1]^p$.
+In this chapter, we discuss in-depth what types of mappings can be used to map the $p$-dimensional space to $[0,1]$, and in particular first study the case $p=2$.
+Before getting into too much detail, let us give some insight as to why this problem is not trivial.
 
@@
 
-the case of population dynamics can easily be embedded in 1D (this is the trivial case)
+### Embedding a finite number of populations in $[0,1]$
 
-<!--
-$$\begin{aligned}
-\rho(\vec z) &= \sum_{a \in P} p_{a} \delta(\vec{z_a} - \vec{z}) \\
-\text{normalization : } \int_{\Rp} \rho(\d \vec{z}) &= \sum_{a} p_a = 1
-\end{aligned}$$
-we can explicitly formulate the embedding with a new 1D PDF (arrange the populations on the line, their probability is their segment length)
-$$\gamma(\alpha) = \sum_{a \in P} p_a \mathbb{1}(C_{a-1} \leq \alpha < C_{a}),\, C_a=\sum_{b=1}^a p_b$$
-the connectivity matrix in the 1D model becomes
-$$J_{ab} = \sum_{\mu=1}^p f_\mu(\vec{z_a}) g_\mu(\vec{z_b})$$
-[TODO MATH : flesh this out]
-[TODOSIM : would this be interesting to simulate ?]
-case 2 populations, we get the ising model (binary patterns) -->
+Let us for a moment consider another type of model networks of neurons, that of homoegeneous population dynamics, that describes how $N_\text{pop}$ different populations interact with each other.
+These types of models emerge when networks of neurons exhibit a finite number of "neuron types".
+We then group together similar neurons, and make the approximation that all neurons inside the population have the same potential, and the same interaction weights with other neurons.
+In effet, every population then represents in actuality an infinity of discrete identical neurons.
 
-TODO : a nice drawing with motivation
+The result of this is that the population dynamics can be written with the same equation as a finite network of neurons, but it already represents the $N \to \infty$ limit.
+We write the potential of the populations $H_a(t)$ and the population connectivity $\tilde J_{ab}$, where $a,b \in \{1, \cdots, N_\text{pop}\}^2$ in @eq-pop-dynamics.
 
-this also helps give an intuition that effectively, when we do the binning in $[0,1]^2$, each bin represents a population
+$$
+\dot H_a(t) = -H_a(t) + \sum_{b=1}^{N_\text{pop}} \tilde J_{ab} \phi(H_b(t))
+$$ {#eq-pop-dynamics}
 
-the point i try to make is that for discrete distributions (i can enumerate the HOMOGENEOUS populations of neurons), it is easy to embed in 1D.
-the problem comes when the distribution is continuous -> we have an infinite amount of homogeneous populations.
+We don't show the derivation of population dynamics here (see @GerKis14, chapter 12.3 for a detailed analysis), but we point out the implicit scaling behavior of $\tilde J_{ab}$.
+Letting $i$ be a neuron of population $a$ and $j$ be neuron of population $b$, the connectivity between the two neurons is $J_{ij}$ and the population connectivity scales with the number of neurons in the presynaptic population $|b|$.
+Furthermore, we make the reasonable assumption that $J_{ij}$ scales as $J^0_{ij} / N$, where $J^0_{ij} \sim \bO(1)$.
 
-we know how to map finite populations to [0,1], we now need a mapping that is good when the number of populations tends to infty
+$$
+\tilde J_{ab} = \frac{|b|}{N} J^0_{ij}
+$$
 
-idea : we need a sequence of mappings ("permutations", this expresses J perm-invar) acting on a finite number of populations, that has a "well defined convergence"
+As $N\to\infty$, the ratio $\tfrac{|b|}{N}$ converges to a constant $p_b$, the probability of the population, normalized such that $\sum_{a=1}^{N_\text{pop}} p_a = 1$.
+It expresses the probability of uniformly sampling a neuron belonging to population $b$.
+In effect, $p_b$ can be interpreted as the "weight" of the population.
+
+So far, we have not considered that the populations are embedded in some abstract space ; but the notion of population weights gives rise to the obvious embedding in $[0,1]$ given by the cumulative sums of the population weights, illustrated in @fig-popdynamics-01.
+We define (somewhat informally) the connectivity kernel $w(\alpha, \beta) = J^0_{ab}$ for $\alpha, \beta$ in the regions corresponding to populations $a$ and $b$ respectively[^poplocation].
+
+[^poplocation]: Formally, we would write the cumulative sum $K_a = \sum_{b=1}^a p_b$.
+Then, we have $\alpha \in [K_{a-1}, K_{a})$ and $\beta \in [K_{b-1}, K_{b})$.
+We abusively wrote $J^0_{ab}$ since all neurons in each population are identical, therefore we can index into the connectivity $J^0_{ij}$ with the population indices $a,b$ instead of neuron indices $i,j$.
+
+![(TODO) Embedding of the population dynamics in $[0,1]$](figures/photo_2023-06-10_19-26-49.jpg){#fig-popdynamics-01}
+
+Defining the neural field $H(\alpha, t) = H_a(t)$, we can write a "neural field equation" on the $[0,1]$ space.
+
+$$
+\begin{aligned}
+\partial_t H(\alpha, t) &= -H(\alpha, t) + \int_0^1 w(\alpha, \beta) \phi(H(\beta, t)) \d \beta \\
+&= -H(\alpha, t) + \sum_{b=1}^{N_\text{pop}} p_b w(\alpha, b) \phi(H_b(t)) \\
+\iff \partial_t H_a(t) &= -H(a, t) + \sum_{b=1}^{N_\text{pop}} p_b w(a, b) \phi(H_b(t)) \\
+&= -H(a, t) + \sum_{b=1}^{N_\text{pop}} p_b J^0_{ab} \phi(H_b(t)) \\
+&= -H(a, t) + \sum_{b=1}^{N_\text{pop}} \tilde J^0_{ab} \phi(H_b(t)) 
+\end{aligned}
+$$
+
+We hereby showed (having taken many writing liberties, but this is just to give an idea) how the neural field reduces back to the original population dynamics, which was in itself already the $N\to\infty$ limit of the network.
+The case of population dynamics therefore shows a trivial example of an embedding in $[0,1]$.
+
+### Population dynamics with an infinite number of populations
+
+When taken in isolation, the previous developments might sound a little bit silly, though they serve to give an impression of the problem at hand, and an interpretation of the rest of this work.
+For a finite number of populations $N_\text{pop}$, the population dynamics is perfectly defined, however what happens as $N_\text{pop} \to \infty$ ?
+The core of the question lies in the fact that we can easily *enumerate* a finite number of populations, but the sum $\sum_{b=1}^{N_\text{pop}}$ must have a clearly defined interpration when this number tends to infinity.
+
+In the light of the toy model presented in @sec-nf, we can reinterpret the density $\rho$ of the neural field equation.
+Every point $\vec z \in \R^p$ (let us consider $p=2$ for the rest) describes a infinite population of homogeneous neurons, associated with a probability (density) $\rho(\vec z)$.
+Therefore, in order to write an equivalent neural field in $[0,1]$, we need a way to *enumerate* the populations.
+This is done using a mapping $S : [0,1]^2 \mapsto [0,1]$, where we traverse the $[0,1]^2$ space by following the order in the image $[0,1]$.
+
+Additionally, the notion of population dynamics helps give another interpretation of the grid integration method given in @sec-gridmethod.
+When we discretize the $[0,1]^2$ space, we are effectively performing a coarse-graining approximation and simulating a finite number of populations.
+For every bin localized by $(v_1, v_2)$, we are sampling the connectivity kernel $w_U((v_1, v_2), \cdot)$ and 2D neural field $h_U((v_1, v_2), t)$ (see @fig-grid-sampling).
+In the rest of the chapter, we refer to these 2D bins as "square populations", and to the corresponding 1D bins as "segment populations".
+
+![Every point of the grid effectively performs coarse-graining, and defines a point where we sample the connectivity kernel $w_U$ and the 2D neural field $h_U$](figures/photo_2023-06-10_21-13-23.jpg){#fig-grid-sampling}
+
+For a finite number of populations $N_\text{pop}$, applying a bijective mapping $S$ will evidently give an equivalent neural field, because the connectivity matrix is permutation-invariant, and the resulting connectivity matrix $J_{S(a),S(b)}$ will therefore give equivalent dynamics.
+As we increase $N_\text{pop}$ by taking finer and finer grids, the image of the grids define a sequence of mappings $S^n : \text{grid in}\ [0,1]^2 \mapsto \text{grid in}\ [0,1]$, which we aim to study in the rest of this chapter.
+
+> How do we define a sequence of mappings $S^n$ such that the limiting $S : [0,1]^2 \mapsto [0,1]$ can express a 1D neural field with identical dynamics ?
 
@@
 
-Define a measurable and bijective mapping $S : [0,1]^2 \rightarrow [0, 1]$.
+We start by writing the neural field equation in the $[0,1]$ space.
+For this, we define a measurable and bijective mapping $S : [0,1]^2 \rightarrow [0, 1]$.
+Furthermore, we let $\lambda : [0,1]^2 \rightarrow [0, \infty]$ be a measure on the unit square (for instance, the Lebesgue measure).
+Taking the composition $\lambda \circ S^{-1}$ then defines a measure on $[0,1]$, which  keeps track of how 2D surfaces are transformed to 1D lengths through the mapping $S$.
 
-Let $\lambda : [0,1]^2 \rightarrow [0, \infty]$ be a measure.
+The connectivity kernel $\tilde w$ defines the interactions between the populations in the $[0,1]$ embedding, which we define in relation to the uniform kernel on $[0,1]^2$ as :
 
-We define $\tilde w(\alpha, \beta) = w_U(S^{-1}(\alpha), S^{-1}(\beta))$ + intial condition
+$$
+\tilde w(\alpha, \beta) = w_U(S^{-1}(\alpha), S^{-1}(\beta))
+$$ {#eq-kernel-1d}
 
-<!-- $\tilde h(t, \alpha) = h_U(t, S^{-1}(\alpha))$ -->
+Defining the initial condition for the 1D neural field as $\tilde h(\alpha, t=0) = h_U(S^{-1}(\alpha), t=0)$, we have fully defined the equation of evolution in the one-dimensional embedding in @eq-nf-1d.
 
@@
 \partial_t \tilde h(t, \alpha) = -\tilde h(t, \alpha) + \int_{[0,1]} \tilde w(\alpha, \beta) \phi(\tilde h(t, \beta)) \lambda \circ S^{-1}(\d \beta)
+$$ {#eq-nf-1d}
+
+The definition of the connectivity kernel in @eq-kernel-1d might seem innocent, but the key element here is the inclusion of the inverse $S^{-1}$.
+We recall that in the toy model, the original connectivity kernel is written as $w(\vec z, \vec y) = \sum_{\mu=1}^p \tilde \phi(y_\mu) z_\mu$.
+This kernel is highly regular, by which we mean that by its continuity in each of its arguments, populations nearby in the embedding will "sense" similar connectivity functions.
+Since it is also differentiable, we can write this as a Taylor expansion in the first argument :
+
+$$
+\begin{aligned}
+w(\vec z + \vec \epsilon, \vec y) &= w(\vec z, \vec y) + \sum_{\mu=1}^p \epsilon_\mu \frac{\partial w}{\partial z_\mu} + \bO(\norm{\vec \epsilon}^2) \\
+&= w(\vec z, \vec y) + \sum_{\mu=1}^p \epsilon_\mu \phi(y_\mu) + \bO(\norm{\vec \epsilon}^2) \\
+&\approx w(\vec z, \vec y) 
+\end{aligned}
 $$
 
-[TODO FIG : a nice drawing of this, also to give an intuition for $\lambda$]
+When we do the change of variables with the CDF, we don't encounter problems because the gaussian (inverse) CDF is differentiable and its derivative is continuous[^inversecdf] ; therefore $w_U(\vec v, \vec u) = w(\textrm{CDF}^{-1}(\vec v), \textrm{CDF}^{-1}(\vec u))$ is still highly regular.
+
+[^inversecdf]: Strictly speaking, we have the problem that $\textrm{CDF}^{-1}(v) \xrightarrow{v \to 1} \infty$, and so the derivative diverges when $v \to 1$ (and similarly, when $v \to 0$).
+However this is not a problem, since these limits correspond to $z \to \pm \infty$, where the density vanishes.
+
+Therefore, we can see why the composition of $w_U$ with a possibly non-differentiable mapping $S$ might be problematic.
+Since the only thing we know about $S$ is that it is measurable and bijective, *there is no guarantee that the regularity of the kernel is conserved through the composition with the mapping $S$.*
 
@@
 
-cite @kharazishvili_strange_2017 peano vs cantor mappings
+### Differentiable mappings
 
-### Continuous mappings
+Let us start our search for a mapping $S^{-1}$ (more precisely, we need $S^{-1} : [0,1] \mapsto [0,1]^2$) that is continuous, motivated by the need to conserve the regularity of the kernel.
+
+We start the discussion on the search of an adequate mapping by citing a result from @kharazishvili_strange_2017.
 
@@
 
-### Bijective mappings
+### Sequences of bijective mappings
+
+#### Mappings as binary expansions {#sec-mappings-binary}
 
@@
 $$
-v_1=\sum_{l=1}^{n} b^1_l 2^{-l} = 0.b^1_1b^1_2\cdots b^1_{n} \iff b^1_l = \mathrm{Ind}\left\{2^{l-1}v_1 - \lfloor2^{l-1}v_1\rfloor \geq \frac{1}{2}\right\}
-$$
+\begin{aligned}
+v^{(n)}_1 &= \sum_{l=1}^{n} b^1_l 2^{-l} = 0.b^1_1b^1_2\cdots b^1_{n} \iff b^1_l = \mathrm{Ind}\left\{2^{l-1}v_1 - \lfloor2^{l-1}v_1\rfloor \geq \frac{1}{2}\right\} \\
+v^{(n)}_2 &= \sum_{l=1}^{n} b^2_l 2^{-l} = 0.b^2_1b^2_2\cdots b^2_{n} \iff b^2_l = \mathrm{Ind}\left\{2^{l-1}v_2 - \lfloor2^{l-1}v_2\rfloor \geq \frac{1}{2}\right\}
+\end{aligned}
+$$ {#eq-trunc}
 
 The intuition behind this expansion is we take a segment, we split it in two, write 0 to go left, write one to go right, then recurse.
-Then combine the bits of $v_1$ and $v_2$ to get the position in the $[0,1]^2$ embedding.
+Then combine the bits of $v^{(n)}_1$ and $v^{(n)}_2$ to get the position in the $[0,1]^2$ embedding.
 
-[TODO FIG : splitting the segment recursively locates the position of the component]
+![Splitting $[0,1]$ recursively gives the binary bit expansion of $v_1$](figures/photo_2023-06-10_17-21-28.jpg)
 
-We define the mapping $S^n$ that takes the $n$-bit truncations of $v_1$ and $v_2$ as input, and outputs a $2n$ bit truncation of $\alpha=S(v_1, v_2)$, where $S$ is the $n\to\infty$ limit of the sequence $S^n$.
+We define the mapping $S^n$ that takes the $n$-bit truncations $v^{(n)}_1$ and $v^{(n)}_2$ as input, and outputs a $2n$ bit truncation $\alpha^{(n)}$ of $\alpha=S(v_1, v_2)$, where $S$ is the $n\to\infty$ limit of the sequence $S^n$.
 
@@
 S^n : &\{0, \tfrac{1}{2^n}, \cdots, 1-\tfrac{1}{2^n}\}^2 &&\mapsto \{0, \tfrac{1}{4^n}, \cdots, 1-\tfrac{1}{4^n}\} \\
-&(0.b^1_1 b^1_2 \cdots b^1_n, 0.b^2_1 b^2_2 \cdots b^2_n) &&\mapsto 0.b_1 b_2 \cdots b_{2n}
+&(v^{(n)}_1,v^{(n)}_2)=(0.b^1_1 b^1_2 \cdots b^1_n, 0.b^2_1 b^2_2 \cdots b^2_n) &&\mapsto \alpha^{(n)}=0.b_1 b_2 \cdots b_{2n}
 \end{alignat}
-$$
+$$ {#eq-Sn}
 
@@
 
-![](figures/mapping_demo_indices.png)
+In the following, we will restrict ourselves to mappings where the functions $b_k(\{b^1_l\}_{l\in\{1,\cdots,n\}}, \{b^2_l\}_{l\in\{1,\cdots,n\}})$ are defined in such a way that each bit of the inputs is used once, and only once.
+Therefore, for each bit $k$ there is a function $b_k(b^{\mu_k}_{l_k})$ acting on only one bit of the input.
+$\mu_k \in \{1,2\}$ is the input component and $l_k \in \{1,\cdots,n\}$ the bit number corresponding to $k$.
+Since this function is binary, the only allowed definitions are
 
-In the following, we will restrict ourselves to mappings where the functions $b_k(\{b^1_l\}_{l\in\{1,\cdots,n\}}, \{b^1_l\}_{l\in\{1,\cdots,n\}})$ are defined in such a way that each bit of the inputs is used once, and only once.
-Therefore, for each bit $k$ there is a function $b_k(b^{\mu_k}_{l_k})$ acting on only one bit.
-Since this function is a binary, the only allowed definitions are $b_k(b^{\mu_k}_{l_k}) = b^{\mu_k}_{l_k}$ (identity) or $b_k(b^{\mu_k}_{l_k}) = 1-b^{\mu_k}_{l_k}$ (inversion).
-We furthermore restrict the formulation by allowing only identity functions[^inversion_restriction].
+* $b_k(b^{\mu_k}_{l_k}) = b^{\mu_k}_{l_k}$ (identity),
+* $b_k(b^{\mu_k}_{l_k}) = 1-b^{\mu_k}_{l_k}$ (inversion),
+* $b_k(b^{\mu_k}_{l_k}) = 0$ (constant zero),
+* $b_k(b^{\mu_k}_{l_k}) = 1$ (constant one).
 
-[^inversion_restriction]: This last restriction is not strictly necessary for the following, however it helps to have a simple model to build a good understanding of the problem
+The constant functions are ignored, since they effectively just restrict the output domain to a subset of $[0,1]$ ; and since only $2n$ bits are allowed for the output, they are a "waste" by ignoring information of the input.
+Because the don't add information to the output (they just "swap" the values of the output bits at known locations), inversion functions are also not considered.
+
+<!-- [^inversion_restriction]: This last restriction is not strictly necessary for the following, however it helps to have a simple model to build a good understanding of the problem -->
 
 All this mathematical formalism expresses, in essence, that the mapping $S^n$ is just a *reordering* of the input bits.
-The geometric view of this is that $S^n$ takes $4^n$ points embedded in $[0,1]^2$, and reorders them onto $4^n$ points in $[0,1]$.
-Therefore, there is a one-to-one mapping between the points on the square and the points on the segment, and the mapping $S^n$, with the restrictions is bijective.
+$S^n$ takes $4^n$ points embedded in $[0,1]^2$, and reorders them onto $4^n$ points in $[0,1]$.
 
-[TODO FIG : mapping of points to segment]
+<!-- ```{dot}
+digraph {
+	subgraph {
+		rank=same
+		ordering="in"
+		node [shape=square];
+		edge [style=invis]           // for all edges, invisible links
+		// we use the invisible edges to establish their sequence (kludge)
+		cdots1 [label=⋯ shape=plaintext]
+		cdots2 [label=⋯ shape=plaintext]
+		b11 [label=b¹₁ shape=rect]
+		b12 [label=b¹₂ shape=rect]
+		b1n1 [label=b¹ₙ₋₁ shape=rect]
+		b11 -> b12 -> cdots1 -> b1n1 -> b1n -> b21 -> b22 -> cdots2 -> b2n
+	}
+
+	subgraph {
+		rank=same
+		ordering="in"
+		edge [style=invis]           // for all edges, invisible links
+		// we use the invisible edges to establish their sequence (kludge)
+		b1 -> b2 -> b3 -> b4 -> b5 -> b6 -> cdots0 -> b_2n1 -> b_2n
+	}
+
+	b11 -> b1
+	b21 -> b2
+	b12 -> b3
+	b22 -> b4
+	b1n -> b_2n1
+	b2n -> b_2n
+}
+``` -->
+
+Therefore, there is a one-to-one mapping between the points on the square and the points on the segment, and the mapping $S^n$ is bijective.
 
@@
 
-Random mapping. mentionned in the introduction. basically a random permutation of bits. not even stable formulation for fixed $n$.
+TODO : might need to talk about how $S(v^{(n)}_1,v^{(n)}_2) = S^n(v^{(n)}_1,v^{(n)}_2)$ in appendix ?
+so basically the limit $n\to\infty$ is the limit where we take infinite precision in the input.
+Or we can drop the $n$ on the inputs, $\alpha^{(n)} = S^n(v^{(n)}_1,v^{(n)}_2) = S^n(v_1,v_2)$ (i think this true just for pointwise convergent mappings)
 
-Anti-Z mapping. Also called in code "Far mapping" (initially I didn't see the link with the Z-order curve), because it destroys locality 2D -> 1D.
+TODO NOTE : talk about how we center the bins instead of taking the corners.
+this is not super important in the n -> oo limit, but numerically at 0 and 1 the CDF-1 goes to infty.
+at infinity, the connectivity kernel is infinite but we have zero neurons (zero measure) (linear growth vs exponential tail of the gaussina) : effectively the recurrent current is zero and we can safely ignore these effects.
+
+#### A link with space-filling curves
+
+With this formalism in mind, we give a geometric intrpretation to the mappings $S^n$.
+The core idea is that the populations in $[0,1]$ have a clear ordering to them, and we can enumerate them simply by their position in the embedding.
+Since we constructed $S^n$ to be bijective, this enumeration traces a path in the corresponding populations of the $[0,1]^2$ embedding.
+
+We illustrate this by introducing the "column" mapping.
 
 $$
-\alpha = A(v_1,v_2) = 0.b^1_n b^2_n b^1_{n-1} b^2_{n-1} \cdots b^1_1 b^2_1 = \sum_{k=1}^{n} b^1_{n+1-k} 2^{1-2k} + b^2_{n+1-k} 2^{-2k}
-$$
+\alpha^{(n)} = C(v^{(n)}_1,v^{(n)}_2) = 0.b^1_1 b^1_2 \cdots b^1_n b^2_1 b^2_2 \cdots b^2_n
+$$ {#eq-column}
 
-Its inverse is :
+The intuition behind this mapping is we enumerate the populations on a grid column by column, letting the index of the rows vary the fastest.
+We visualize this mapping in @fig-col, by coloring the square populations by the position of the corresponding $[0,1]$ population, and additionally drawing lines between the points to help guide the eye.
+
+```{python}
+#| label: fig-col
+#| fig-cap: Column mapping for $n=2$. Numbers represent the ordering of the populations in the $[0,1]$ embedding, neighbouring populations are connected by lines.
+#| echo: false
+def plot_trace(mapping_cls, nrec, do_trace: bool = True, do_numbers: bool = True):
+	F = ColumnMapping.new_nrec(nrec).inverse_samples(centered=True)
+	mapping = mapping_cls.new_nrec(nrec)
+	alpha = mapping(F)
+
+	fig, ax = plt.subplots(facecolor='#FFFFFF00')
+	im = ax.imshow(alpha.reshape((2**nrec, 2**nrec)).T, origin='lower', vmin=0, vmax=1, extent=(0,1,0,1), interpolation='none')
+	mapping_letter = { ZMapping: 'Z', ColumnMapping: 'C', AntiZMapping: 'A' }.get(mapping_cls, 'S')
+	plt.colorbar(im, ax=ax, label=f'$\\alpha={mapping_letter}^{{n}}(v_1, v_2)$')
+	if do_trace:
+		# plot the trace
+		ax.plot(*F[np.argsort(alpha)].T)
+		# hide trace at numbering location
+		ax.scatter(*F.T, s=250, c=alpha, cmap='viridis', vmin=0, vmax=1, zorder=10)  
+	# add the bin numbers
+	if do_numbers:
+		for xy, i in zip(F[np.argsort(alpha)], range(1, 4**nrec+1)):
+			ax.annotate(i, xy=xy, va='center_baseline', ha='center', zorder=20, color='white', fontweight='bold',
+				path_effects=[pe.withStroke(linewidth=1, foreground='black')])
+	ax.set_aspect('equal')
+	ax.set_xlabel('$v_1$'); ax.set_ylabel('$v_2$')
+	ax.set_xlim((0,1)); ax.set_ylim((0,1))
+	ax.xaxis.set_major_locator(ticker.MultipleLocator(1/2**2))
+	ax.yaxis.set_major_locator(ticker.MultipleLocator(1/2**2))
+	# for x in np.linspace(0, 1, 2**nrec+1): ax.axvline(x, ymin=0, ymax=1, color='tab:gray', linewidth=0.5, zorder=0, clip_on=False)
+	# for y in np.linspace(0, 1, 2**nrec+1): ax.axhline(y, xmin=0, xmax=1, color='tab:gray', linewidth=0.5, zorder=0, clip_on=False)
+	breathe()
+	return fig, ax
+
+fig, ax = plot_trace(ColumnMapping, nrec=2)
+display(fig)
+```
+
+For finite $n$, we write the inverse of the column mapping in @eq-column-inverse.
 
@@
 \begin{aligned}
-v_1 = A^{-1}_1(\alpha) &= 0.b_{2n-1} b_{2n-3} \cdots b_{1} = \sum_{k=1}^n b_{2(n-k)-1} 2^{-k} \\
-v_2 = A^{-1}_2(\alpha) &= 0.b_{2n} b_{2n-2} \cdots b_{2} = \sum_{k=1}^n b_{2(n+1-k)} 2^{-k}
+v^{(n)}_1 = C^{-1}_1(\alpha^{(n)}) &= 0.b_1 b_2 \cdots b_n = \sum_{k=1}^n b_{k} 2^{-k} \\
+v^{(n)}_2 = C^{-1}_2(\alpha^{(n)}) &= 0.b_{n+1} b_{n+2} \cdots b_{2n} = \sum_{k=1}^n b_{n+k} 2^{-k}
+\end{aligned}
+$$ {#eq-column-inverse}
+
+This way of enumerating the square populations might be familiar to people who have worked with numerical arrays.
+In such (contiguous) arrays, the values of are laid out in memory in "row-major" order (C-style) or "column-major" order (Fortran-style) ; and we can go from the 1D memory representation to a 2D matrix representation by using "reshape" operations.
+Effectively, this is how this mapping is implemented in code, using `numpy.ravel_multi_index` (2D to 1D) and `numpy.unravel_index` (1D to 2D).
+
+#### "Local" space-filling curves {#sec-local}
+
+Research in computer science and efficient data structures has given rise to other ways of storing 2-dimensional information in a 1-dimensional memory.
+A problem in GPU computing is that of texture locality : sampling of textures often involves reading data corresponding to a small 2D region, and therefore in order to improve speed (minimize cache misses), the 2D data should be packed close together in the 1D memory.
+One way of ordering the 2D texture data in the 1D memory is by using the Z-order curve (also known as "Morton mapping" or "Lebesgue curve" @Sagan1994).
+In this way, points close in 2D *tend* to be tightly packed in 1D.
+We can define the Z-order curve (Z-mapping) in the following way[^zcurve] :
+
+[^zcurve]: The Z-curve is often defined in the transpose manner, putting first the bits of $v_2$ (vertical component) then the bits of $v_1$, instead of $v_1$ then $v_2$.
+Doing this, we get the "upright" Z shapes, as [shown here](https://en.wikipedia.org/wiki/Z-order_curve#/media/File:Four-level_Z.svg).
+
+$$
+\alpha^{(n)} = Z(v^{(n)}_1,v^{(n)}_2) = 0.b^1_1 b^2_1 b^1_2 b^2_2 \cdots b^1_n b^2_n = \sum_{k=1}^{n} b^1_k 2^{1-2k} + b^2_k 2^{-2k}
+$$ {#eq-z}
+
+And its inverse is :
+
+$$
+\begin{aligned}
+v^{(n)}_1 = Z^{-1}_1(\alpha^{(n)}) &= 0.b_1 b_3 \cdots b_{2n-1} = \sum_{k=1}^n b_{2k-1} 2^{-k} \\
+v^{(n)}_2 = Z^{-1}_2(\alpha^{(n)}) &= 0.b_2 b_4 \cdots b_{2n} = \sum_{k=1}^n b_{2k} 2^{-k}
+\end{aligned}
+$$ {#eq-z-inverse}
+
+@fig-z is a visualization for the Z-mapping when $n=2$, and shows how it is composed of the "Z" shapes giving it its name.
+
+```{python}
+#| label: fig-z
+#| fig-cap: Z-mapping for $n=2$. Numbers represent the ordering of the populations in the $[0,1]$ embedding, neighbouring populations are connected by lines.
+#| echo: false
+fig, ax = plot_trace(ZMapping, nrec=2)
+display(fig)
+```
+
+To illustrate this notion of locality, we take a few consecutive populations in 1D and show where the corresponding populations fall in the 2D embedding.
+This is done in @fig-locality, in which we see that the Z-mapping seems to conserve locality from 1D to 2D : populations close in 1D seem to be close in 2D.
+
+We also see that the Column mapping is not local in this sense, since there is a large variation along the vertial direction, which seems to always be "of order one".
+This can be explained by looking at the expression for $C^{-1}_2(\alpha^{(n)}) = 0.b_{n+1} b_{n+2} \cdots b_{2n}$ : small variations of the order of $2^{-n}$ in the value of $\alpha$ always result in variations of order one in $C^{-1}_2$.
+
+```{python}
+#| label: fig-locality
+#| fig-cap: Locality of the mappings. The 2D populations corresponding to three small segments of 1D populations of length $\tfrac{1}{16}$ are shown. The figures are generated with $n=6$.
+#| echo: false
+##| fig-subcap:
+##|   - Z mapping
+##|   - Column mapping
+##| layout-ncol: 2
+fig, axes = plt.subplot_mosaic('aa\nbc', facecolor='#FFFFFF00', figsize=(6, 4), height_ratios=[1, 5], subplot_kw=dict(facecolor='#FFFFFF00'))
+nrec = 6
+
+cmap = mpl.colormaps['viridis']
+cmap.set_bad(color='#FFFFFF00')
+
+alpha = np.linspace(0, 1, 4**nrec+1)
+mask1d = (0.2 <= alpha) & (alpha < 0.2+1/4**2)
+mask1d |= (0.8 <= alpha) & (alpha < 0.8+1/4**2)
+mask1d |= (0.5 <= alpha) & (alpha < 0.5+1/4**2)
+alpha[~mask1d] = np.nan
+axes['a'].imshow(alpha[None, :], vmin=0, vmax=1, extent=((0,1,0,0.2)), interpolation='none', cmap=cmap)
+axes['a'].set_aspect('auto')
+axes['a'].spines.left.set_visible(False)
+axes['a'].yaxis.set_visible(False)
+axes['a'].set_title('Populations in 1D', fontsize='medium')
+axes['a'].set_xlabel('$\\alpha$')
+breathe(axes['a'])
+
+for mapping_cls, ax in zip((ZMapping, ColumnMapping), (axes['b'], axes['c'])):
+	F = ColumnMapping.new_nrec(nrec).inverse_samples(centered=True)
+	mapping = mapping_cls.new_nrec(nrec)
+	alpha = mapping(F)
+	mask1d = (0.2 <= alpha) & (alpha < 0.2+1/4**2)
+	mask1d |= (0.8 <= alpha) & (alpha < 0.8+1/4**2)
+	mask1d |= (0.5 <= alpha) & (alpha < 0.5+1/4**2)
+	alpha[~mask1d] = np.nan
+	ax.imshow(alpha.reshape((2**nrec, 2**nrec)).T, origin='lower', vmin=0, vmax=1, extent=(0,1,0,1), interpolation='none', cmap=cmap)
+	ax.set_xlabel('$v_1=S_1^{{-1}}(\\alpha)$'); ax.set_ylabel('$v_2=S_2^{{-1}}(\\alpha)$')
+	ax.set_xlim((0,1)); ax.set_ylim((0,1))
+	ax.xaxis.set_major_locator(ticker.MultipleLocator(1/2**2))
+	ax.yaxis.set_major_locator(ticker.MultipleLocator(1/2**2))
+	ax.set_title(f'To 2D with {mapping_cls.__name__}', fontsize='medium')
+	breathe(ax)
+
+display(fig)
+```
+
+The attentive reader might however object this treatment of locality, since in some sense the Column mapping is local, but from 2D to 1D.
+This can be seen from the definition of $\alpha^{(n)}=C(v^{(n)}_1,v^{(n)}_2) = 0.b^1_1 b^1_2 \cdots b^1_n b^2_1 b^2_2 \cdots b^2_n$, since small variations of order $2^{-n}$ in $v^{(n)}_1$ result in small variations also of order $2^{-n}$ in $\alpha^{(n)}$.
+Similarly, small variations of order $2^{-n}$ in $v^{(n)}_1$ result in even smaller variations also of order $2^{-2n}$ in $\alpha^{(n)}$ !
+We therefore seem to have two notions of locality : from 1D to 2D, and from 2D to 1D ; and the Column mapping shows that they are not equivalent.
+
+However, we argue that in our case only the notion of locality from 1D to 2D matters.
+In essence, the dynamics in the 2D embeddeding are already known, and we ask the question of whether it is possible to write an equivalent neural field 1D (in the sense that it has identical latent trajectories).
+Without repeating the intuition given in @sec-intro-simnf, if neighbouring populations in 1D have similar potentials, then this allows us to write a neural field equation in $[0,1]$.
+Therefore we would like to have the property that populations close in 1D are (inversely) mapped to populations close in 2D, which already have similar potentials.
+
+<!-- note for me : find a formulation for the recursive local mapping. it is not trivial, since it has
+
+```
+n=1
+0 0 -> 00
+1 0 -> 01
+1 1 -> 10
+0 1 -> 11
+
+n=2
+00 00 -> 0000
+01 00 -> 0001
+01 01 -> 0010
+00 01 -> 0011
+10 00 -> 0100
+11 00 -> 0101
+11 01 -> 0110
+10 01 -> 0111
+10 10 -> 1000
+11 10 -> 1001
+11 11 -> 1010
+10 11 -> 1011
+00 10 -> 1100
+01 10 -> 1101
+01 11 -> 1110
+00 11 -> 1111
+```
+
+this might be nice as an appendix, to show that our formulation is very simplistic, and more complex formulations can also have nice properties
+
+this also shows our proof is "incomplete", because it relies on the fact we have a permutation of the bits, although it is not strictly necessary. we have a sufficient, but not necessary proof
+
+also : this might be a type of peano curve or smth. see @Sagan1994
+
+https://www.cut-the-knot.org/Curriculum/Geometry/LebesgueCurve.shtml
+-->
+
+### Sequence of mappings and their limit
+
+#### Pointwise convergence of mappings
+
+We now discuss what happens to the bijective mappings when $n\to\infty$.
+For this, let us consider again the "random mapping" considered in the introduction @sec-intro-simnf.
+We discussed how this hypothetical mapping maps each position in $[0,1]^2$ into a "random position" in $[0,1]$.
+A finite-$n$ formulation would be that the mapping corresponds to a *random permutation* of the bits.
+It is however unclear how to formally write this mapping in a way that it is well-defined for a given $n$, since by definition, it is random.
+For a given position of the input $(v_1,v_2)$, we would like that the image is stable in the $n\to\infty$ limit.
+
+$$
+S^{n}(v_1, v_2) \xrightarrow{n\to\infty} \alpha =: S(v_1, v_2)
+$$
+
+This is the condition of "pointwise convergence" of the mappings.
+We might be tempted to write that any mapping for which we can write the binary expansion $S^n$ is pointwise convergent, however we show by the example of the "anti-Z" mapping that this is not the case.
+Let us define for finite $n$
+
+$$
+\begin{aligned}
+\alpha^{(n)} &= A(v^{(n)}_1,v^{(n)}_2) = 0.b^1_n b^2_n b^1_{n-1} b^2_{n-1} \cdots b^1_1 b^2_1 = \sum_{k=1}^{n} b^1_{n+1-k} 2^{1-2k} + b^2_{n+1-k} 2^{-2k} \\
+\iff& \\
+v^{(n)}_1 &= A^{-1}_1(\alpha^{(n)}) = 0.b_{2n-1} b_{2n-3} \cdots b_{1} = \sum_{k=1}^n b_{2(n-k)-1} 2^{-k} \\
+v^{(n)}_2 &= A^{-1}_2(\alpha^{(n)}) = 0.b_{2n} b_{2n-2} \cdots b_{2} = \sum_{k=1}^n b_{2(n+1-k)} 2^{-k}
 \end{aligned}
@@
 
-it does not converge
+<!-- Anti-Z mapping. Also called in code "Far mapping" (initially I didn't see the link with the Z-order curve), because it destroys locality 2D -> 1D. -->
 
-Column mapping. Also called "Reshape mapping" in the code. It converges to a projection on the $x$-axis
+```{python}
+#| label: fig-anti-z
+#| fig-cap: Anti-Z mapping for $n=2$.
+Numbers represent the ordering of the populations in the $[0,1]$ embedding.
+We don't plot the trace, because it has many crossovers and overlaps and makes the plot hard to read.
+#| echo: false
+fig, ax = plot_trace(AntiZMapping, nrec=2, do_trace=False)
+display(fig)
+```
 
-$$
-\alpha = C(v_1,v_2) = 0.b^1_1 b^1_2 \cdots b^1_n b^2_1 b^2_2 \cdots b^2_n
-$$
+@fig-anti-z gives some insight into the problem of this mapping.
+For a fixed input $(v_1,v_2)$, we see that the output point position jumps around, and that these jumps do not decay, they are always of order 1.
+The intuition behind pointwise convergence is that as $n$ increases, the image of a point by $S^n$ is continually refined.
 
-Its inverse is :
+Another way of looking at this is to consider how $\alpha^{(n)}$ fluctuates as the precision of the input increases with $n \to \infty$.
+As we refine the input with fluctuations of order $2^{-n}$, the output $\alpha^{(n)} = b^1_n 4^{-1} + b^2_n 4^{-2} + b^1_{n-1} 4^{-3} + b^2_{n-1} 4^{-4} \cdots$ fluctuates (at least) with amplitude $\sim 4^{-2}$ independantly of $n$.
+Therefore the output never converges, and the mapping $A^{n}$ is not pointwise convergent.
+
+In our setting, how can we therefore guarantee that the mappings are pointwise convergent ?
+We need to show that small fluctuations in the input result in small fluctuations of the output.
+The binary expansion allows us to argue that for the sequence of mappings to be pointwise convergent, we need that the "least significant bits" (LSB, the bits to the right of the binary expansion) of the input are mapped to the LSB of the output.
+Conversely, the "most significant bits" (MSB, the bits to the left of the binary expansion) should be mapped to the MSB of the output.
+
+We illustrate this with the Z-mapping, written in @eq-z.
+Corrections of order $2^{-n}$ in $v_1$ and $v_2$ induce corrections of order $2^{1-2n}$ and $2^{-2n}$ in $\alpha$ respectively.
+Therefore, it is easy to see that the Z-mapping is pointwise convergent as $n\to\infty$.
+Similarly, the Column mapping in @eq-column is also pointwise convergent, since as discussed @sec-local it has locality of 2D to 1D.
+
+#### Conservation of bijectivity in the $n\to\infty$ limit
+
+However, the behavior of the Column mapping in that of the Z-mapping are very different in the $n\to\infty$ limit.
+Up until now, we have avoided this issue by considering the finite-$n$ approximation of the mappings, which by construction is bijective.
+We illustrate the infinite-$n$ (numerically, large $n$) limit of these two mappings in @fig-n-to-infty.
+
+```{python}
+#| label: fig-n-to-infty
+#| fig-cap: The $n\to\infty$ limit of mappings.
+#| fig-subcap:
+#|   - Z-mapping
+#|   - Column mapping
+#| layout-ncol: 2
+#| echo: false
+fig, ax = plot_trace(ZMapping, nrec=6, do_trace=False, do_numbers=False)
+display(fig)
+fig, ax = plot_trace(ColumnMapping, nrec=6, do_trace=False, do_numbers=False)
+display(fig)
+```
+
+We see that the Column mapping converges to projection on $v_1$, in other words $\alpha=C(v_1, v_2) = v_1$, showing that despite every $C^{n}$ being bijective, the limit of the sequence is not bijective.
+This becomes clear when looking at the binary expansion of $C$ :
 
@@
 \begin{aligned}
-v_1 = C^{-1}_1(\alpha) &= 0.b_1 b_2 \cdots b_n = \sum_{k=1}^n b_{k} 2^{-k} \\
-v_2 = C^{-1}_2(\alpha) &= 0.b_{n+1} b_{n+2} \cdots b_{2n} = \sum_{k=1}^n b_{n+k} 2^{-k}
+C(v_1, v_2) &= \lim_{n\to\infty} C^{(n)}(v_1,v_2)\\
+&= \lim_{n\to\infty} 0.b^1_1 b^1_2 \cdots b^1_n b^2_1 b^2_2 \cdots b^2_n \\
+&= \lim_{n\to\infty} 0.b^1_1 b^1_2 \cdots \\
+&= v_1
 \end{aligned}
@@
 
-Z-mapping. Also called Z-order curve, Morton mapping. It conserves locality 2D -> 1D.
+The limit is not bijective, because the bits of $v_2$ are "lost" in the $n\to\infty$ limit, and the inverse $C_2^{-1}(\alpha)$ is not well-defined, since it "takes bits from $\infty$ to $2\infty$".
 
-$$
-\alpha = Z(v_1,v_2) = 0.b^1_1 b^2_1 b^1_2 b^2_2 \cdots b^1_n b^2_n = \sum_{k=1}^{n} b^1_k 2^{1-2k} + b^2_k 2^{-2k}
-$$
+The Z-mapping however seems to converge to some sort of fractal structure, the limit is nowhere continuous and has an infinite number of discontinuities, but the size of these discontinuities is overwhelmingly small : on average, the fact that there are large discontinuities does not matter in the $n\to\infty$.
+We don't go further into detail about this in this section, but this gives some intuition for the arguments presented in @sec-locality.
 
-Its inverse is :
+We finish this discussion on the limit of the sequence of mappings by asking if there are other mappings similar to the Z-mapping, in the sense that they are pointwise convergent and the limit is bijective.
+In an informal way, we established two conditions for this in the context of the mappings described in @sec-mappings-binary :
 
-$$
-\begin{aligned}
-v_1 = Z^{-1}_1(\alpha) &= 0.b_1 b_3 \cdots b_{2n-1} = \sum_{k=1}^n b_{2k-1} 2^{-k} \\
-v_2 = Z^{-1}_2(\alpha) &= 0.b_2 b_4 \cdots b_{2n} = \sum_{k=1}^n b_{2k} 2^{-k}
-\end{aligned}
-$$
+1. condition of pointwise convergence : the LSB of the input must be mapped to the LSB of the output,
+2. condition of "no information loss" : the MSB of the input must be mapped to the MSB of the output.
 
-note that in other definitions (and in code) it is defined by $0.b^2_1 b^1_1 b^2_2 b^1_2 \cdots$. This difference is not fundamentally important though.
+These two conditions do not leave a lot of flexibility in the design of new mappings.
+We can of course define mappings where we interleave the bits in a different order, for instance $0.b^2_1 b^1_1 b^2_2 b^1_2 \cdots$ ; or mappings where we reorder the $k$ bits of the input, for instance $0.b^1_2 b^2_2 b^1_1 b^2_1 b^1_4 b^2_4 b^1_3 b^2_3 \cdots$, but the conditions restrict the reordering to be permutations of neighbouring bits.
+In essence, these mappings are not fundamentally different, and the Z-mapping can be seen as a "canonical representation" of the pointwise convergent mappings with bijective limits.
 
-[TODO FIG : 3D vis of the mapping]
+### Visualizating the mappings of the dynamic 2D neural field
 
-### Sequence of mappings and their limit
+It is time for a little treat.
+Before we argue about how to simulate the dynamics of the 1D neural field, let us simulate the dynamics of the 2D neural field, then at each timestep, we map the 2D neural field to the 1D embedding.
+This will give us a reference of how the 1D neural field is supposed to look like under the condition that its latent dynamics are identical.
 
-the sequence in [0,1] traces a path in [0,1]²
-
-intuition for column.
-
-the limit of pointwise convergent mappings and measurable (which they are) is measurable.
-
-argue over binary expansion, and especially how the fact that it converges forces the least significant bits of $z_1$ and $z_2$ to go to the lsb of $\alpha$.
-this imposes some structure of what mappings are good
-
-z-mapping can in this context be seen as a "canonical representation" of the pointwise convergent, bijective, local mapping.
-
-### Numerical implementation
-
-what we are doing in reality is assigning indices to squares of size $2^{-n} \times 2^{-n}$, and mapping them to indices representing segments of length $4^{-n}$
-
-[TODO FIG : finite n is population dynamics. drawing of the square-to-segment mapping, something like
-
-![Alt text](figures/Screenshot_2023-06-03_19-59-49.png)]
-
-we use the fact we have a compact set onto another [0,1]² -> [0,1], so we can write a binary expansion of (x, y) -> z.
-x has n bits, y has n bits, z has 2n bits.
-
-an arbitrary mapping can then be written as a recombination of the bits of x and y (and any other constant) into 2n bits
-this is good because with floating point we lose precision, here we work with integers (~ fixed point precision), so the implementations of the mappings are exact.
-
-the locality of 2D -> 1D is not important (i.e neurons can be repeated in 1D without too much trouble).
-locality 1D -> 2D is much more important (a neuron in 1D must refer to only one (not multiple) neurons in 2D)
-
-![](figures/mappings_h1d_state.png)
-
-![simulation of 2D field, and mapping of the dynamics to the 1D field](figures/mapping_recursive_quadrant.mp4)
+![(TODO) Animation showing the image of the dynamics 2D neural field through different mappings $S : [0,1]^2 \mapsto [0,1]$](figures/photo_2023-06-10_17-21-19.jpg)
 
@@
 
-### Reinterpretation of sampling and binning methods
-
 reintrpretation of the simulation methods
 
-average low-rank vectors
+-> sampling the 2D connectivity kernel
+
+-> average low-rank vectors
+
+![Alt text](figures/Screenshot_2023-06-03_19-59-49.png)]
 

### 2023-06-11 notebooks/neurodyn/_mapping.py

@@ -394,7 +394,7 @@ class RecursiveLocalMapping(BinMapping):
 		return f'RecursiveLocalMapping{{nrec={self.nrec}}}'
 
 
-class ReshapeMapping(BinMapping):
+class ColumnMapping(BinMapping):
 	"""Implements mapping column-by-column (reshape operation)
 	
 	This mapping converges to a projection on x.
@@ -413,7 +413,7 @@ class ReshapeMapping(BinMapping):
 	def __str__(self) -> str:
 		return f'ReshapeMapping{{nx={self.nx}, ny={self.ny}}}'
 
-ColumnMapping = ReshapeMapping
+ReshapeMapping = ColumnMapping
 
 
 class DiagonalMapping(BinMapping):
@@ -505,6 +505,9 @@ class SzudzikMapping(BinMapping):
 			np.vstack((fsqrt, indices - fsqrt**2 - fsqrt)).T,
 		)
 
+	def __str__(self) -> str:
+		return f'SzudzikMapping{{nxy={self.nxy}}}'
+
 
 class ZMapping(BinMapping):
 	"""Implements [Z-order curve](https://en.wikipedia.org/wiki/Z-order_curve)"""
@@ -525,12 +528,16 @@ class ZMapping(BinMapping):
 
 	def indices(self, F: np.ndarray, bbox: Box | None = None) -> np.ndarray:
 		indices2d = self.indices2d(F, bbox)
-		return ZMapping.part1by1_64(indices2d[:, 0]) | (ZMapping.part1by1_64(indices2d[:, 1]) << 1)
+		return ZMapping.part1by1_64(indices2d[:, 1]) | (ZMapping.part1by1_64(indices2d[:, 0]) << 1)
 
 	def indices_to_indices2d(self, indices: np.ndarray) -> np.ndarray:
 		indices2d = np.zeros((len(indices), 2), dtype='uint64')
-		indices2d[:, 0] = ZMapping.unpart1by1_64(indices)
-		indices2d[:, 1] = ZMapping.unpart1by1_64(indices >> 1)
+		# NOTE : we inverse x and y, because we define the mapping as
+		# alpha = 0.b^1_1 b^2_1 b^1_2 b^2_2 ...
+		# whereas the Z-mapping is defined as
+		# alpha = 0.b^2_1 b^1_1 b^2_2 b^1_2 ...
+		indices2d[:, 1] = ZMapping.unpart1by1_64(indices)
[…]

### 2023-06-11 notebooks/neurodyn/_plot.py

@@ -6,7 +6,7 @@ import colorsys
 from matplotlib.axes import Axes
 from matplotlib.figure import Figure
 from ._rnn import LowRankRNN, Result
-from ._overlap import overlap
+from ._overlap import overlap, projection
 
 __all__ = [
 	'plot_neuron_trajectory', 'plot_overlap_trajectory', 'plot_dh_hist',
@@ -41,12 +41,12 @@ def plot_overlap_trajectory(rnn: LowRankRNN, res: Result, figax: tuple[Figure, A
 	fig, ax = _unwrap_figax(figax)
 
 	ax.set_xlabel('Time $t$ [s]')
-	ax.set_ylabel('Overlap $m_{{\\mu}}$')
-	ax.set_title('Overlap trajectory')
-	m = overlap(rnn, res.h)
+	ax.set_ylabel('Latent $\\kappa_{{\\mu}}$')
+	ax.set_title('Latent trajectory')
+	kappa = projection(rnn, res.h)
 
-	for mu in range(len(m)):
-		ax.plot(res.t, m[mu], label=f'$\\mu={mu}$', **kwargs)
+	for mu in range(len(kappa)):
+		ax.plot(res.t, kappa[mu], label=f'$\\mu={mu+1}$', **kwargs)
 
 	ax.grid(visible=True, axis='y')
 	ax.legend()
@@ -57,14 +57,14 @@ def plot_overlap_trajectory(rnn: LowRankRNN, res: Result, figax: tuple[Figure, A
 def plot_overlap_phase2D(rnn: LowRankRNN, res: Result, point_start: bool = True, point_end: bool = True, lim01: bool = True, figax: tuple[Figure, Axes] | None = None, **kwargs) -> tuple[Figure, Axes]:
 	fig, ax = _unwrap_figax(figax)
 
-	ax.set_xlabel('Overlap $m_0(t)$')
-	ax.set_ylabel('Overlap $m_1(t)$')
+	ax.set_xlabel('Latent $\\kappa_1(t)$')
+	ax.set_ylabel('Latent $\\kappa_2(t)$')
 	ax.set_title('Latent trajectory')
 	ax.set_aspect('equal')
-	m = overlap(rnn, res.h)
-	line, = ax.plot(m[0], m[1], **kwargs)
[…]

### 2023-06-11 manuscript/appendices/B_zcurve.qmd
guess who just went through half a book in one day :))

::: content-hidden
{{< include ../macros.tex >}}
\def\C{\mathcal{C}}
:::

# On the quasi-bijectivity and continuity of the Z-mapping

We study the Z-curve $Z^{-1}$

## The Z-curve is not injective {#sec-zcurve-noninjective}

this is common misconception, and people write it is a bijection, when really it is not

The Z-curve : [0,1] -> [0,1]^2 is not injective since the binary expansion is ambiguous

x = 0.10000... = 0.011111...

=> alpha = 0.1y0y0y0  !=   alpha = 0.0y1y1y1y1

but the point on which the expansion is ambiguous form a set of zero measure, so the integral is still ok

on the set [0,1] \ {problems}, Z-curve is bijective

injectivity is the property that f(x) != f(y)  iff  x != y,  (equiv f(x) == f(y) iff x == y)

Z^{-1} defines a space-filling curve

not a problem because we are just repeating the same population twice

## $Z^{-1}$ defines a space-filling curve on a subset of $[0,1]$ {#sec-zcurve-lebesgue}

### The Cantor set $\C$

the cantor set has measure 0, but has the same cardinality of the continuum

### $Z^{-1} | \C$ is a surjection

### $Z^{-1} | \C$ is continuous

## Relation with the Lebesgue space-filling curve

we would also relax and say the Z-curve is discontinuous, because lim alpha ->0.5+ Z(alpha) != lim alpha -> 0.5- Z(alpha)
-> NO ! because we linearly interpolate between the quadrants. we say the curve is continuous, in the sense that it is connected. but it is allowed to jump around !

talk about the fact we have other space filling curves, such as hilbert curve or peano curve, that are continuous and surjective [0,1] -> [0,1]^2. we use the Z-curve because it's easy to implement numerically

### 2023-06-11 manuscript/chapters/3_mappings.qmd
guess who just went through half a book in one day :))

@@
 
-### Differentiable mappings
+### The search for differentiable mappings
 
-Let us start our search for a mapping $S^{-1}$ (more precisely, we need $S^{-1} : [0,1] \mapsto [0,1]^2$) that is continuous, motivated by the need to conserve the regularity of the kernel.
+We start our search of mappings by attempting to find a mapping $S$ that is differentiable.
+The easiest example of such a mapping would be a linear projection, for instance $P(v_1, v_2) = v_1$.
+This mapping is obviously continuous and differentiable in both components, and surjectively maps the unit square to the unit segment.
 
-We start the discussion on the search of an adequate mapping by citing a result from @kharazishvili_strange_2017.
+However, the obvious problem with this approach is that linear projections are not invertible.
+Hence, should we nevertheless define an inverse $P^{-1}$ that computes the preimage of $\alpha \in [0,1]$, we have that
 
-linear projections don’t work ofc.
+$$
+P^{-1}(\alpha) = \{(\alpha, v_2)\ |\ v_2 \in [0,1] \}.
+$$
 
-Note : why are not using the CDF to embed in 1D ? because it's not bijective !
+In other words, $P^{-1}(\alpha)$ defines a vertical line in the $[0,1]^2$ embedding.
+When we then take the composition $w_U \circ P^{-1}$, the regularity is destroyed, in the sense that the same population $\alpha$ in the 1D embedding will "sense" all the connectivity kernels on the vertical line.
 
-> there is no homeomorphism
-  (continuous bijective function)
-  between [0,1]² and [0,1]
+A more sophisticated attempt, in the spirit of @sec-mapping-cdf, might be to consider the joint cumulative function
+
+$$
+\alpha = \mathrm{CDF}(z_1, z_2) = \int_{-\infty}^{z_1} \int_{-\infty}^{z_2} \rho(y_1, y_2) \d y_1 \d y_2.
+$$
+
+However this again fails because the joint cumulative function is not bijective, and isolines of the CDF define diagonals in the uniform embedding space $[0,1]^2$[^jointcdf].
+
+[^jointcdf]: We note to avoid any confusion, that in @sec-mapping-cdf the CDF is taken component-wise, instead of jointly.
+The univariate gaussian CDF defines a bijection between $\R$ and $(0,1)$, but the joint CDF does not.
+
+Perhaps the linear projection or CDF examples are a bit too naive, and we need to look for more complex examples.
+We might ask the question of whether a diffeomorphism (a differentiable bijection with differentiable inverse) even exists between the unit square and the unit segment.
+Unfortunately, results from topology give a negative answer, and even worse, it can be shown that there is no homeomorphism (a continuous bijection with continuous inverse) between the unit square and the unit segment.
+This is stated by Netto's theorem, which imposes continuous bijections to conserve the dimensionality between the domain and the image.
+
+### Mappings of $[0,1]$ into $[0,1]^2$
+
+Since in the composion $w_U \circ S^{-1}$ we work with the inverse mapping, it might be judicious to directly consider functions from $[0,1]$ into $[0,1]^2$.
+We have already established that there is no homeomorphism between those spaces, and need to relax one of the conditions.
+
+Functions of $[0,1]$ into $[0,1]^2$ are extensively discussed in chapter 1 of @kharazishvili_strange_2017, and we cite two classes of such functions :
+
+1. Cantor functions are bijections between $[0,1]$ and $[0,1]^2$. By the non-existence theorem, these functions cannot be continuous.
+2.
+Peano functions are continuous surjections from $[0,1]$ to $[0,1]^2$.
+They cannot be injections.
+For reasons that will become apparent in the following sections, Peano functions are often called space-filling curves.
+
+In the light of these results, we might be tempted to search for expressions for Cantor-type functions, since having a bijection would allow us to have a direct mapping between the square and the segment populations.
+Unfortunately, these functions are discontinuous which breaks the regularity of the kernel.
+
+Peano functions seem like good candidates for the mapping $S^{-1}$, and are extensively studied in the litterature (we refer for instance to [@sagan_space-filling_1994;@Sagan1993]).
+They have the nice property that they are continuous, and are surjective, so we the segment populations are able to "reach" the entirety of the square populations.
+All things considered, non-injectivity is not a problem, since it just means that two segment populations can "refer" the same square population, which does not break the regularity of the kernel, since the two segments populations can just have their own neighbourhoods.
+Furthermore, we will see that in our examples, the non-injectivity can be ignored because the mapping is quasi-bijective.
+
+![Non-injectivity of $S^{-1}$ means that two segment populations can refer to the same square population.](figures/photo_2023-06-12_00-46-26.jpg)
 
@@
 
-The Z-mapping however seems to converge to some sort of fractal structure, the limit is nowhere continuous and has an infinite number of discontinuities, but the size of these discontinuities is overwhelmingly small : on average, the fact that there are large discontinuities does not matter in the $n\to\infty$.
-We don't go further into detail about this in this section, but this gives some intuition for the arguments presented in @sec-locality.
+The Z-mapping however seems to converge to a fractal curve.
+Numerically of course, there seem to be infinite number of discontinuities, but the size of these discontinuities is overwhelmingly small, and *on average* the size of these numerical disconuities vanishes as $n\to\infty$.
+We don't go further into detail about this in this section (in @sec-zcurve-lebesgue we develop on the convergence to a space-filling fractal curve), but this gives some intuition for the arguments presented in @sec-locality.
 
-We finish this discussion on the limit of the sequence of mappings by asking if there are other mappings similar to the Z-mapping, in the sense that they are pointwise convergent and the limit is bijective.
+We finish this discussion on the limit of the sequence of mappings by asking if there are other mappings similar to the Z-mapping, in the sense that they are pointwise convergent and the limit is bijective[^zmapping-notbijective].
 In an informal way, we established two conditions for this in the context of the mappings described in @sec-mappings-binary :
+
+[^zmapping-notbijective]: Despite what can often be found in online resources, the Z-mapping is *not* a bijection because that would violate Netto's theorem.
+The rew_ason for this is the ambiguity in the writing of binary expansions, and in @sec-zcurve-noninjective we give a proof that the Z-mapping is not injective in the $n\to\infty$ limit.
+However the Z-mapping is *quasi-bijective*, in the sense that the set of points for which the Z-mapping is not injective has zero measure, and can therefore be ignored.
 

### 2023-06-11 manuscript/unused/B_kappa.qmd
guess who just went through half a book in one day :))

--- 2023-05-31 manuscript/chapters/6_appendix.qmd
@@
-# Appendix
+# Behaviour of $\kappa_\mu(t)$ in the cyclic neural field
 
-::: {.content-hidden}
-{{< include ../macros.tex >}}
-:::
+## Definitions of dynamics
 
-## Fixed point study in the network of neurons
+[TODO MATH : derivation of this ? from the general setting to our case]
 
-We recall the equation for the low-rank network of neurons :
+We recall the definition of the low-rank variables $\kappa_\mu(t)$.
+
+The dynamics of $h(t, \vec z)$ are in a subsystem of dimension $p$ spanned by the ONB of functions $\{e_\mu(\vec z) = z_\mu | \mu=1,\cdots,p\}$, with the scalar product $\avg{f, g} = \int_{\mathbb R^p} f(\vec y) g(\vec y) \rho(\mathrm d \vec y)$.
+
+We decompose $h(t, \vec z) = h^\perp(t, \vec z) + \sum_{\mu=1}^p \kappa_\mu(t) z_\mu$, and write the system of equations for $\mu=1,\cdots,p$.
 
 $$
-\dot h_i(t) = -h_i(t) + \frac 1N \sum_{\mu=1}^p \sum_{j=1}^N F_{\mu i} G_{\mu j} \phi(h_j(t))
+\dot \kappa_\mu(t) = -\kappa_\mu(t) + \int_{\mathbb{R}^p} \tilde\phi(y_\mu) \phi(h(t, \vec y)) \rho(\mathrm d \vec y) = -\kappa_\mu(t) + m_\mu(t)
 $$
 
-The low-rank vectors are in general obtained by sampling a distribution $\vec{\xi} = (\xi_1, \cdots, \xi_p) \sim \rho$.
-Every neuron $i$ samples this distribution independantly, and low-rank vectors are obtained by passing the samples through functions $f_\mu(\vec{\xi})$ and $g_\mu(\vec{\xi})$.
-
-We note that in the main text, the distribution is a multivariate normal with unit covariance and zero mean, such that it factors out as $\rho(\xi_1, \cdots, \xi_p) = \prod_{\mu=1}^p \mathcal{N}(\xi_\mu)$.
-Furthermore, the functions are component-wise $f_\mu(\xi_1, \cdots, \xi_p) = \xi_\mu$ and $g_\mu(\xi_1, \cdots, \xi_p) = \tilde \phi(\xi_\mu) = \frac{\phi(\xi_\mu) - \avg{\phi(\xi_\mu)}}{\mathrm{Var}[\phi(\xi_\mu)]}$.
+with initial conditions
 
 $$
-F_{\mu i} = f_\mu(\xi_{1 i}, \cdots, \xi_{p i}), \, G_{\mu i} = g_\mu(\xi_{1 i}, \cdots, \xi_{p i})
+\kappa_\mu(0) = \int_{\mathbb{R}^p} y_\mu h(0, \vec y) \rho(\mathrm d \vec y)
 $$
 
-### Derivation of fixed points
-
-In order to find the fixed points, one has to solve in general $\dot h_i(t) = 0$ for all $i = 1, \cdots, N$.
-Because of the nonlinearity $\phi(h_j(t))$, this is difficult, but we can guess the fixed points $h^\star_i$ and verify them by substituting into the equation that $\dot h_i(t) = 0$.
-
-#### Zero fixed point
-
-A trivial fixed point is $h^\star_i = 0$ for all $i$, which we prove to be a fixed point as $N \to \infty$.
+and the orthogonal component evolves according to
 
 $$
-\begin{aligned}
-\dot h_i(t) &= \frac 1N \sum_{\mu=1}^p \sum_{j=1}^N F_{\mu i} G_{\mu j} \phi(0) \\
-&= \frac{\phi(0)}{N} \sum_{\mu=1}^p \sum_{j=1}^N F_{\mu i} G_{\mu j} \\
-&= \frac{\phi(0)}{N} \sum_{\mu=1}^p \sum_{j \neq i}^N F_{\mu i} G_{\mu j} + \underbrace{\frac{\phi(0)}{N} \sum_{\mu=1}^p F_{\mu i} G_{\mu i}}_{\bO(\frac 1N)} \\
-&= \phi(0) \sum_{\mu=1}^p \frac 1N \sum_{j \neq i}^N F_{\mu i} G_{\mu j} + \bO(\tfrac 1N) \\
-&= \phi(0) \sum_{\mu=1}^p F_{\mu i} \frac{N-1}{N} \frac{1}{N-1} \sum_{j \neq i}^N G_{\mu j} + \bO(\tfrac 1N) \\
-&\stackrel{N \gg 1}{\approx} \phi(0) \sum_{\mu=1}^p F_{\mu i} \avg{g_\mu(\vec{\xi})}
-\end{aligned}
+h^\perp(t, \vec z) = h^\perp(0, \vec z) \mathrm e^{-t},
+\quad
+h^\perp(0, \vec z) = h(0, \vec z) - \sum_{\mu=1}^p \kappa_\mu(0) z_\mu
 $$
 
-<!-- TODO : fixme. the F_i should not be in the integral -->
+## Cyclic dynamics in the $p=2$ cyclic neural field
 
-To write the last line, we make a slight abuse of notation in order to avoid notation problems when taking the $N \to \infty$ limit of $F_{\mu i}$.
+[TODO MATH : derivation of delayed ?]
 
-In the case of the network studied in this work, the find that
+[TODO MATH : derivation of cyclic ?]
+
+In the case of the cyclic neural field, the evolution equation for $\kappa_\nu(t)$ is modified, where $\delta$ is the delay, and with the convention $\mu+1 = 1$ if $\mu=p$.
 
 $$
-\avg{g_\mu(\vec{\xi})} = \avg{\tilde \phi(\xi_\mu)} = \avg{\frac{\phi(\xi_\mu) - \avg{\phi(\xi_\mu)}}{\mathrm{Var}[\phi(\xi_\mu)]}} = 0
+\dot \kappa_\mu(t) = -\kappa_\mu(t) + m_{\mu+1}(t - \delta)
 $$
 
-Thus, $h_i^\star=0$ is a fixed point in the limit $N \to \infty$.
+## Relation to overlaps $m_\mu(t)$
 
-#### Pattern fixed point
-
-Another set of fixed points is $h^\star_i = F_{\nu i}$ for all $\nu = 1, \cdots, p$.
-We refer to these as the "patterns" in the context of learning the network "rembers" (converges to) these vectors.
-We verify the fixed point similarly as before.
-
-$$
-\begin{aligned}
-\dot h_i(t) &= -F_{\nu i} + \frac 1N \sum_{\mu=1}^p \sum_{j=1}^N F_{\mu i} G_{\mu j} \phi(F_{\nu j}) \\
-&= -F_{\nu i} + \frac 1N \sum_{\mu=1}^p \sum_{j \neq i}^N F_{\mu i} G_{\mu j} \phi(F_{\nu j}) + \bO(\tfrac 1N) \\
-&= -F_{\nu i} + \frac 1N \sum_{j \neq i}^N F_{\nu i} G_{\nu j} \phi(F_{\nu j}) + \frac 1N \sum_{\mu \neq \nu}^p \sum_{j \neq i}^N F_{\mu i} G_{\mu j} \phi(F_{\nu j}) + \bO(\tfrac 1N) \\
-&\stackrel{N \gg 1}{\approx} -F_{\nu i} + F_{\nu i} \avg{g_\nu(\vec{\xi_j}) \phi(\xi_{\nu j})} + \sum_{\mu \neq \nu}^p F_{\mu i} \avg{g_\mu(\vec{\xi_j}) \phi(\xi_{\nu j})}
-\end{aligned}
-$$
-
-The term $\avg{g_\mu(\vec{\xi_j}) \phi(\xi_{\nu j})}$ vanishes in the case of a factorizable distribution and when the functions $g_\mu$ are component-wise.
-
-$$
-\begin{aligned}
-\avg{g_\mu(\vec{\xi_j}) \phi(\xi_{\nu j})} &= 
-\int_{\Rp} g_\mu(\xi_1, \cdots, \xi_p) \phi(\xi_\nu) \rho(\d \vec{\xi}) \\
-&= \int_{\Rp} g_\mu(\xi_\mu) \phi(\xi_\nu) \prod_{\gamma=1}^p \rho_\gamma(\d \xi_\gamma) \\
-&= \int_{\R} g_\mu(\xi_\mu) \rho_\mu(\d \xi_\mu) \int_{\R} \phi(\xi_\nu) \rho_\nu(\d \xi_\nu) \\
-&= \avg{g_\mu(\xi_\mu)} \avg{\phi(\xi_\nu)}
-\end{aligned}
-$$
-
-As established in the previous section, in the case of the network studied in this work, $\avg{g_\mu(\xi_\mu)} = 0$.
-
-The term $\avg{g_\nu(\vec{\xi_j}) \phi(\xi_{\nu j})}$ is more difficult, but we can prove that in our case it equals one.
-
-$$
-\begin{aligned}
-\avg{g_\nu(\vec{\xi_j}) \phi(\xi_{\nu j})} &=
-\avg{\frac{\phi(\xi_{\mu j}) - \avg{\phi(\xi_{\mu j})}}{\mathrm{Var}[\phi(\xi_{\mu j})]} \phi(\xi_{\mu j})} \\
-&= \avg{\frac{\phi(\xi)^2 - \avg{\phi(\xi)} \phi(\xi) }{\mathrm{Var}[\phi(\xi)]}} \\
-&= \frac{\avg{\phi(\xi)^2} - \avg{\avg{\phi(\xi)} \phi(\xi)} }{\mathrm{Var}[\phi(\xi)]} \\
-&= \frac{\avg{\phi(\xi)^2} - \avg{\phi(\xi)}^2}{\mathrm{Var}[\phi(\xi)]} \\
-&= \frac{\mathrm{Var}[\phi(\xi)]}{\mathrm{Var}[\phi(\xi)]} \\
-&= 1
-\end{aligned}
-$$
-
-Therefore, the set of $h_i^\star = F_{\nu i}$ for all $\nu = 1, \cdots, p$ are fixed points in the limit of large $N$, under the assumptions that the distribution is factorizable, the functions $g_\mu$ are component-wise, and $\avg{g_\nu(\xi) \phi(\xi)}=1$, which is the case in the setup of this work.
-
-### Fixed points of the neural field
-
-The derivation of the fixed points for the neural field equation follows the same arguments as the fixed points in the network of neurons.
-
-As $N \to \infty$, the fixed points $h_i^\star$ become functions $h^\star(\xi_1, \cdots, \xi_p)$ on the $\Rp$ space :
-
-* $h_i^\star = 0 \to h^\star(\xi_1, \cdots, \xi_p) = 0$
-* $h_i^\star = F_{\nu i} = \xi_{\nu i} \to h^\star(\xi_1, \cdots, \xi_p) = \xi_\nu$ for all $\nu = 1, \cdots, p$
-
-<!-- TODO : add derivation for neural field ? -->
-
-### Stability of the fixed points
-
-<!-- TODO -->
-
-## Behavior of $\kappa_\mu(t)$ in the cyclic network
-
-kappa in the delayed, rolled and cyclic network
-
-relation to overlaps
-
-## "Mean patterns" in the binned connectivity
-
-derivation of $\Jab$ and mean patterns
-
-For finite number of recursive quadrant iterations $n$, we can do a "mean-field approximation" inside each of the $4^n$ segments.
-Let $\alpha = \{i_1,\cdots,i_{|\alpha|}\}$ be the multi-index corresponding to all neurons of which the embedding in $\mathbb R^p$ gets mapped to the segment $\alpha$ in $[0,1]$.
-Let $H_\alpha(t) = \frac 1 {|\alpha|} \sum_{i \in \alpha} h_i(t)$ be the (mean) RNN potential of the segment $\alpha$.
-The connectivity matrix $\tilde J_{\alpha,\beta}$ satisfies
-
-$$
-\dot H_\alpha(t) = -H_\alpha(t) + \sum_{\beta \in \text{segments of length } 4^{-n}} \tilde J_{\alpha,\beta} \phi(H_\beta(t))
-$$
-
-By substituting the original $h_i(t)$, we find the correct rescaling is given by
-
-$$
-\tilde J_{\alpha,\beta} = \frac 1 {|\alpha|} \sum_{i \in \alpha} \sum_{j \in \beta} J_{ij}
-$$
-
-In the low-rank case, we have
-
-$$
-J_{ij} = \frac 1N \sum_{\mu=1}^p F_{\mu,i} G_{\mu,j}
-$$
-
-We can define the "mean pattern" inside each bin :
-
-$$
-\tilde F_{\mu,\alpha} = \frac{1}{|\alpha|} \sum_{i \in \alpha} F_{\mu,i}, \; \tilde G_{\mu,\alpha} = \frac{1}{|\alpha|} \sum_{i \in \alpha} G_{\mu,i}
-$$
-
-Then the connectivity matrix is given by, noting $N = \sum_{\alpha} |\alpha|$,
-
-$$
-\tilde J_{\alpha,\beta} = \frac{|\beta|}{\sum_{\beta'} |\beta'|} \left( \sum_{\mu=1}^p \tilde F_{\mu,\alpha} \tilde G_{\mu,\beta} - \delta_{\alpha,\beta} \underbrace{\sum_{\mu=1}^p \sum_{i \in \alpha} \frac{F_{\mu,i}}{|\alpha|} \frac{G_{\mu,i}}{|\alpha|}}_{\gamma_{\alpha}} \right)
-$$
-
-Recurrent current is given by
-
-$$
-I^\text{rec}_\alpha(t) = \sum_\beta \tilde J_{\alpha,\beta} \phi(H_\beta(t)) = \frac{1}{\sum_{\beta'} |\beta'|} \left( \sum_\beta |\beta| \sum_{\mu=1}^p \tilde F_{\mu,\alpha} \tilde G_{\mu,\beta} \phi(H_\beta(t)) - |\alpha| \tilde \gamma_\alpha \phi(H_\alpha(t)) \right)
-$$
-
-The overlaps can be computed as
-
-$$
-\tilde m_\mu(t) = \frac{1}{\sum_{\alpha} |\alpha|} \sum_\alpha |\alpha| G_{\mu,\alpha} \phi(H_\alpha(t))
-$$
+[TODO]

### 2023-06-13 manuscript/chapters/3_mappings.qmd
nearly finished chap 3 !

@@
 
-## Population dynamics : an introductory example of a 1D embedding
+## Population dynamics: an introductory example of a 1D embedding
 
@@
 
-Let us for a moment consider another type of model networks of neurons, that of homoegeneous population dynamics, that describes how $N_\text{pop}$ different populations interact with each other.
+Let us for a moment consider another type of model for networks of neurons, that of homogeneous population dynamics, that describes how $N_\text{pop}$ different populations interact with each other.
 These types of models emerge when networks of neurons exhibit a finite number of "neuron types".
-We then group together similar neurons, and make the approximation that all neurons inside the population have the same potential, and the same interaction weights with other neurons.
-In effet, every population then represents in actuality an infinity of discrete identical neurons.
+We then group similar neurons ad make the approximation that all neurons inside the population have the same potential, and the same interaction weights with other neurons.
+In effect, every population then represents in actuality an infinity of discrete identical neurons.
 
@@
 We don't show the derivation of population dynamics here (see @GerKis14, chapter 12.3 for a detailed analysis), but we point out the implicit scaling behavior of $\tilde J_{ab}$.
-Letting $i$ be a neuron of population $a$ and $j$ be neuron of population $b$, the connectivity between the two neurons is $J_{ij}$ and the population connectivity scales with the number of neurons in the presynaptic population $|b|$.
+Letting $i$ be a neuron of population $a$ and $j$ be a neuron of population $b$, the connectivity between the two neurons is $J_{ij}$ and the population connectivity scales with the number of neurons in the presynaptic population $|b|$.
 Furthermore, we make the reasonable assumption that $J_{ij}$ scales as $J^0_{ij} / N$, where $J^0_{ij} \sim \bO(1)$.
@@
 
-So far, we have not considered that the populations are embedded in some abstract space ; but the notion of population weights gives rise to the obvious embedding in $[0,1]$ given by the cumulative sums of the population weights, illustrated in @fig-popdynamics-01.
-We define (somewhat informally) the connectivity kernel $w(\alpha, \beta) = J^0_{ab}$ for $\alpha, \beta$ in the regions corresponding to populations $a$ and $b$ respectively[^poplocation].
+So far, we have not considered that the populations are embedded in some abstract space; but the notion of population weights gives rise to the obvious embedding in $[0,1]$ given by the cumulative sums of the population weights, illustrated in @fig-popdynamics-01.
+We define (somewhat informally) the connectivity kernel $w(\alpha, \beta) = J^0_{ab}$ for $\alpha, \beta$ in the regions corresponding to populations $a$ and $b$ respectively.[^poplocation]
 
@@
 For a finite number of populations $N_\text{pop}$, the population dynamics is perfectly defined, however what happens as $N_\text{pop} \to \infty$ ?
-The core of the question lies in the fact that we can easily *enumerate* a finite number of populations, but the sum $\sum_{b=1}^{N_\text{pop}}$ must have a clearly defined interpration when this number tends to infinity.
+The core of the question lies in the fact that we can easily *enumerate* a finite number of populations, but the sum $\sum_{b=1}^{N_\text{pop}}$ must have a clearly defined interpretation when this number tends to infinity.
 
-In the light of the toy model presented in @sec-nf, we can reinterpret the density $\rho$ of the neural field equation.
-Every point $\vec z \in \R^p$ (let us consider $p=2$ for the rest) describes a infinite population of homogeneous neurons, associated with a probability (density) $\rho(\vec z)$.
+In light of the toy model presented in @sec-nf, we can reinterpret the density $\rho$ of the neural field equation.
+Every point $\vec z \in \R^p$ (let us consider $p=2$ for the rest) describes an infinite population of homogeneous neurons, associated with a probability (density) $\rho(\vec z)$.
 Therefore, in order to write an equivalent neural field in $[0,1]$, we need a way to *enumerate* the populations.
@@
 For a finite number of populations $N_\text{pop}$, applying a bijective mapping $S$ will evidently give an equivalent neural field, because the connectivity matrix is permutation-invariant, and the resulting connectivity matrix $J_{S(a),S(b)}$ will therefore give equivalent dynamics.
-As we increase $N_\text{pop}$ by taking finer and finer grids, the image of the grids define a sequence of mappings $S^n : \text{grid in}\ [0,1]^2 \mapsto \text{grid in}\ [0,1]$, which we aim to study in the rest of this chapter.
+As we increase $N_\text{pop}$ by taking finer and finer grids, the image of the grids defines a sequence of mappings $S^n: \text{grid in}\ [0,1]^2 \mapsto \text{grid in}\ [0,1]$, which we aim to study in the rest of this chapter.
 
@@
 
-The connectivity kernel $\tilde w$ defines the interactions between the populations in the $[0,1]$ embedding, which we define in relation to the uniform kernel on $[0,1]^2$ as :
+The connectivity kernel $\tilde w$ defines the interactions between the populations in the $[0,1]$ embedding, which we define relating to the uniform kernel on $[0,1]^2$ as :
 
@@
 
-When we do the change of variables with the CDF, we don't encounter problems because the gaussian (inverse) CDF is differentiable and its derivative is continuous[^inversecdf] ; therefore $w_U(\vec v, \vec u) = w(\textrm{CDF}^{-1}(\vec v), \textrm{CDF}^{-1}(\vec u))$ is still highly regular.
+When we do the change of variables with the CDF, we don't encounter problems because the Gaussian (inverse) CDF is differentiable and its derivative is continuous[^inversecdf]; therefore $w_U(\vec v, \vec u) = w(\textrm{CDF}^{-1}(\vec v), \textrm{CDF}^{-1}(\vec u))$ is still highly regular.
 
 [^inversecdf]: Strictly speaking, we have the problem that $\textrm{CDF}^{-1}(v) \xrightarrow{v \to 1} \infty$, and so the derivative diverges when $v \to 1$ (and similarly, when $v \to 0$).
-However this is not a problem, since these limits correspond to $z \to \pm \infty$, where the density vanishes.
+However, this is not a problem, since these limits correspond to $z \to \pm \infty$, where the density vanishes.
 
@@
 We start our search of mappings by attempting to find a mapping $S$ that is differentiable.
-The easiest example of such a mapping would be a linear projection, for instance $P(v_1, v_2) = v_1$.
+The easiest example of such a mapping would be a linear projection, for instance, $P(v_1, v_2) = v_1$.
 This mapping is obviously continuous and differentiable in both components, and surjectively maps the unit square to the unit segment.
@@
 
-However this again fails because the joint cumulative function is not bijective, and isolines of the CDF define diagonals in the uniform embedding space $[0,1]^2$[^jointcdf].
+However, this again fails because the joint cumulative function is not bijective, and isolines of the CDF define diagonals in the uniform embedding space $[0,1]^2$.[^jointcdf]
 
 [^jointcdf]: We note to avoid any confusion, that in @sec-mapping-cdf the CDF is taken component-wise, instead of jointly.
-The univariate gaussian CDF defines a bijection between $\R$ and $(0,1)$, but the joint CDF does not.
+The univariate Gaussian CDF defines a bijection between $\R$ and $(0,1)$, but the joint CDF does not.
 
 Perhaps the linear projection or CDF examples are a bit too naive, and we need to look for more complex examples.
-We might ask the question of whether a diffeomorphism (a differentiable bijection with differentiable inverse) even exists between the unit square and the unit segment.
+We might ask the question of whether a diffeomorphism (a differentiable bijection with a differentiable inverse) even exists between the unit square and the unit segment.
 Unfortunately, results from topology give a negative answer, and even worse, it can be shown that there is no homeomorphism (a continuous bijection with continuous inverse) between the unit square and the unit segment.
@@
 
-Since in the composion $w_U \circ S^{-1}$ we work with the inverse mapping, it might be judicious to directly consider functions from $[0,1]$ into $[0,1]^2$.
+Since in the composition $w_U \circ S^{-1}$ we work with the inverse mapping, it might be judicious to directly consider functions from $[0,1]$ into $[0,1]^2$.
 We have already established that there is no homeomorphism between those spaces, and need to relax one of the conditions.
@@
 
-In the light of these results, we might be tempted to search for expressions for Cantor-type functions, since having a bijection would allow us to have a direct mapping between the square and the segment populations.
+In light of these results, we might be tempted to search for expressions for Cantor-type functions, since having a bijection would allow us to have a direct mapping between the square and the segment populations.
 Unfortunately, these functions are discontinuous which breaks the regularity of the kernel.
 
-Peano functions seem like good candidates for the mapping $S^{-1}$, and are extensively studied in the litterature (we refer for instance to [@sagan_space-filling_1994;@Sagan1993]).
-They have the nice property that they are continuous, and are surjective, so we the segment populations are able to "reach" the entirety of the square populations.
-All things considered, non-injectivity is not a problem, since it just means that two segment populations can "refer" the same square population, which does not break the regularity of the kernel, since the two segments populations can just have their own neighbourhoods.
+Peano functions seem like good candidates for the mapping $S^{-1}$, and are extensively studied in the literature (we refer for instance to [@sagan_space-filling_1994;@Sagan1993]).
+They have the nice property that they are continuous and surjective, so we the segment populations can "reach" the entirety of the square populations.
+All things considered, non-injectivity is not a problem, since it just means that two segment populations can "refer" to the same square population, which does not break the regularity of the kernel, since the two segment populations can just have their own neighborhoods.
 Furthermore, we will see that in our examples, the non-injectivity can be ignored because the mapping is quasi-bijective.
@@
 
-The intuition behind this expansion is we take a segment, we split it in two, write 0 to go left, write one to go right, then recurse.
+The intuition behind this expansion is we take a segment, split it in two, write 0 to go left, write one to go right, then recurse.
 Then combine the bits of $v^{(n)}_1$ and $v^{(n)}_2$ to get the position in the $[0,1]^2$ embedding.
@@
 In general, each bit $b^k$ of the image would be a binary function of the $n$ bits of $v_1$ and the $n$ bits of $v_2$, implicitly $b_k = b_k(b^1_1, b^1_2, \cdots, b^1_n, b^2_1, b^2_2, \cdots, b^2_n)$.
-With this, we are able to formulate any mapping of $4^n$ distinct elements (here : squares of size $2^{-n} \times 2^{-n}$) onto $4^n$ elements (here : segments of length $4^{-n}$).
+With this, we can formulate any mapping of $4^n$ distinct elements (here : squares of size $2^{-n} \times 2^{-n}$) onto $4^n$ elements (here : segments of length $4^{-n}$).
 
 The intuition behind the use of $2n$ bits for the image is that we take $2^n \times 2^n = 4^n$ squares in $[0,1]^2$ and map those to $4^n$ segments in $[0,1]$.
-In an informal way, we map $n+n$ bits onto $2n$ bits.
+Informally, we map $n+n$ bits onto $2n$ bits.
 
@@
 
-The constant functions are ignored, since they effectively just restrict the output domain to a subset of $[0,1]$ ; and since only $2n$ bits are allowed for the output, they are a "waste" by ignoring information of the input.
-Because the don't add information to the output (they just "swap" the values of the output bits at known locations), inversion functions are also not considered.
-
-<!-- [^inversion_restriction]: This last restriction is not strictly necessary for the following, however it helps to have a simple model to build a good understanding of the problem -->
+The constant functions are ignored, since they effectively just restrict the output domain to a subset of $[0,1]$; and since only $2n$ bits are allowed for the output, they are a "waste" by ignoring information of the input.
+Because they don't add information to the output (they just "swap" the values of the output bits at known locations), inversion functions are also not considered.
 
@@
 
-<!-- ```{dot}
-digraph {
-	subgraph {
-		rank=same
-		ordering="in"
-		node [shape=square];
-		edge [style=invis]           // for all edges, invisible links
-		// we use the invisible edges to establish their sequence (kludge)
-		cdots1 [label=⋯ shape=plaintext]
-		cdots2 [label=⋯ shape=plaintext]
-		b11 [label=b¹₁ shape=rect]
-		b12 [label=b¹₂ shape=rect]
-		b1n1 [label=b¹ₙ₋₁ shape=rect]
-		b11 -> b12 -> cdots1 -> b1n1 -> b1n -> b21 -> b22 -> cdots2 -> b2n
-	}
-
-	subgraph {
-		rank=same
-		ordering="in"
-		edge [style=invis]           // for all edges, invisible links
-		// we use the invisible edges to establish their sequence (kludge)
-		b1 -> b2 -> b3 -> b4 -> b5 -> b6 -> cdots0 -> b_2n1 -> b_2n
-	}
-
-	b11 -> b1
-	b21 -> b2
-	b12 -> b3
-	b22 -> b4
-	b1n -> b_2n1
-	b2n -> b_2n
-}
-``` -->
-
 Therefore, there is a one-to-one mapping between the points on the square and the points on the segment, and the mapping $S^n$ is bijective.
 
-These simplifications might seem too restrictive, however we will show in the rest of this thesis that this formulation is flexible enough to express mapping which have good properties such as "locality", all while being simple enough to help build an intuition.
+These simplifications might seem too restrictive, however, we will show in the rest of this thesis that this formulation is flexible enough to express mappings that have good properties such as "locality", all while being simple enough to help build intuitions.
 
@@
 
-With this formalism in mind, we give a geometric intrpretation to the mappings $S^n$.
+With this formalism in mind, we give a geometric interpretation to the mappings $S^n$.
 The core idea is that the populations in $[0,1]$ have a clear ordering to them, and we can enumerate them simply by their position in the embedding.
@@
 This way of enumerating the square populations might be familiar to people who have worked with numerical arrays.
-In such (contiguous) arrays, the values of are laid out in memory in "row-major" order (C-style) or "column-major" order (Fortran-style) ; and we can go from the 1D memory representation to a 2D matrix representation by using "reshape" operations.
+In such (contiguous) arrays, the values are laid out in memory in "row-major" order (C-style) or "column-major" order (Fortran-style); and we can go from the 1D memory representation to a 2D matrix representation by using "reshape" operations.
 Effectively, this is how this mapping is implemented in code, using `numpy.ravel_multi_index` (2D to 1D) and `numpy.unravel_index` (1D to 2D).
@@
 Research in computer science and efficient data structures has given rise to other ways of storing 2-dimensional information in a 1-dimensional memory.
-A problem in GPU computing is that of texture locality : sampling of textures often involves reading data corresponding to a small 2D region, and therefore in order to improve speed (minimize cache misses), the 2D data should be packed close together in the 1D memory.
+A problem in GPU computing is that of texture locality: sampling textures often involves reading data corresponding to a small 2D region, and therefore to improve speed (minimize cache misses), the 2D data should be packed close together in the 1D memory.
 One way of ordering the 2D texture data in the 1D memory is by using the Z-order curve (also known as "Morton mapping" or "Lebesgue curve" @Sagan1994).
@@
 
-[^zcurve]: The Z-curve is often defined in the transpose manner, putting first the bits of $v_2$ (vertical component) then the bits of $v_1$, instead of $v_1$ then $v_2$.
-Doing this, we get the "upright" Z shapes, as [shown here](https://en.wikipedia.org/wiki/Z-order_curve#/media/File:Four-level_Z.svg).
+[^zcurve]: The Z-curve is often defined in a transpose manner, putting first the bits of $v_2$ (vertical component) then the bits of $v_1$, instead of $v_1$ then $v_2$.
+By doing this, we get the "upright" Z shapes, as [shown here](https://en.wikipedia.org/wiki/Z-order_curve#/media/File:Four-level_Z.svg).
 
@@
 To illustrate this notion of locality, we take a few consecutive populations in 1D and show where the corresponding populations fall in the 2D embedding.
-This is done in @fig-locality, in which we see that the Z-mapping seems to conserve locality from 1D to 2D : populations close in 1D seem to be close in 2D.
+This is done in @fig-locality, in which we see that the Z-mapping seems to conserve locality from 1D to 2D: populations close in 1D seem to be close in 2D.
 
-We also see that the Column mapping is not local in this sense, since there is a large variation along the vertial direction, which seems to always be "of order one".
+We also see that the Column mapping is not local in this sense, since there is a large variation along the vertical direction, which seems to always be "of order one".
 This can be explained by looking at the expression for $C^{-1}_2(\alpha^{(n)}) = 0.b_{n+1} b_{n+2} \cdots b_{2n}$ : small variations of the order of $2^{-n}$ in the value of $\alpha$ always result in variations of order one in $C^{-1}_2$.
@@
 
-The attentive reader might however object this treatment of locality, since in some sense the Column mapping is local, but from 2D to 1D.
+The attentive reader might however object to this treatment of locality, since in some sense the Column mapping is local, but from 2D to 1D.
 This can be seen from the definition of $\alpha^{(n)}=C(v^{(n)}_1,v^{(n)}_2) = 0.b^1_1 b^1_2 \cdots b^1_n b^2_1 b^2_2 \cdots b^2_n$, since small variations of order $2^{-n}$ in $v^{(n)}_1$ result in small variations also of order $2^{-n}$ in $\alpha^{(n)}$.
 Similarly, small variations of order $2^{-n}$ in $v^{(n)}_1$ result in even smaller variations also of order $2^{-2n}$ in $\alpha^{(n)}$ !
-We therefore seem to have two notions of locality : from 1D to 2D, and from 2D to 1D ; and the Column mapping shows that they are not equivalent.
+We therefore seem to have two notions of locality: from 1D to 2D, and from 2D to 1D; and the Column mapping shows that they are not equivalent.
 
 However, we argue that in our case only the notion of locality from 1D to 2D matters.
-In essence, the dynamics in the 2D embeddeding are already known, and we ask the question of whether it is possible to write an equivalent neural field 1D (in the sense that it has identical latent trajectories).
-Without repeating the intuition given in @sec-intro-simnf, if neighbouring populations in 1D have similar potentials, then this allows us to write a neural field equation in $[0,1]$.
+In essence, the dynamics in the 2D embedding are already known, and we ask the question of whether it is possible to write an equivalent neural field 1D (in the sense that it has identical latent trajectories).
+Without repeating the intuition given in @sec-intro-simnf, if neighboring populations in 1D have similar potentials, then this allows us to write a neural field equation in $[0,1]$.
 Therefore we would like to have the property that populations close in 1D are (inversely) mapped to populations close in 2D, which already have similar potentials.
@@
 A finite-$n$ formulation would be that the mapping corresponds to a *random permutation* of the bits.
-It is however unclear how to formally write this mapping in a way that it is well-defined for a given $n$, since by definition, it is random.
+It is unclear how to formally write this mapping in a way that is well-defined for a given $n$, since by definition, it is random.
 For a given position of the input $(v_1,v_2)$, we would like that the image is stable in the $n\to\infty$ limit.
@@
 This is the condition of "pointwise convergence" of the mappings.
-We might be tempted to write that any mapping for which we can write the binary expansion $S^n$ is pointwise convergent, however we show by the example of the "anti-Z" mapping that this is not the case.
+We might be tempted to write that any mapping for which we can write the binary expansion $S^n$ is pointwise convergent, however, we show by the example of the "anti-Z" mapping that this is not the case.
 Let us define for finite $n$
@@
 
-<!-- Anti-Z mapping. Also called in code "Far mapping" (initially I didn't see the link with the Z-order curve), because it destroys locality 2D -> 1D. -->
-
 ```{python}
@@
 Numbers represent the ordering of the populations in the $[0,1]$ embedding.
-We don't plot the trace, because it has many crossovers and overlaps and makes the plot hard to read.
+We don't plot the trace, because it has many crossovers and overlaps which makes the plot hard to read.
 #| echo: false
@@
 @fig-anti-z gives some insight into the problem of this mapping.
-For a fixed input $(v_1,v_2)$, we see that the output point position jumps around, and that these jumps do not decay, they are always of order 1.
+For a fixed input $(v_1,v_2)$, we see that the output point position jumps around and that these jumps do not decay, they are always of order 1.
 The intuition behind pointwise convergence is that as $n$ increases, the image of a point by $S^n$ is continually refined.
@@
 
-In our setting, how can we therefore guarantee that the mappings are pointwise convergent ?
-We need to show that small fluctuations in the input result in small fluctuations of the output.
+In our setting, how can we guarantee that the mappings are pointwise convergent?
+We need to show that small fluctuations in the input result in small fluctuations in the output.
 The binary expansion allows us to argue that for the sequence of mappings to be pointwise convergent, we need that the "least significant bits" (LSB, the bits to the right of the binary expansion) of the input are mapped to the LSB of the output.
@@
 Therefore, it is easy to see that the Z-mapping is pointwise convergent as $n\to\infty$.
-Similarly, the Column mapping in @eq-column is also pointwise convergent, since as discussed @sec-local it has locality of 2D to 1D.
+Similarly, the Column mapping in @eq-column is also pointwise convergent, since as discussed @sec-local it has a locality of 2D to 1D.
 
-#### Conservation of bijectivity in the $n\to\infty$ limit
+#### Conservation of bijectivity in the $n\to\infty$ limit {#sec-ninfty-bijection}
 
-However, the behavior of the Column mapping in that of the Z-mapping are very different in the $n\to\infty$ limit.
+However, the behavior of the Column mapping and that of the Z-mapping are very different in the $n\to\infty$ limit.
 Up until now, we have avoided this issue by considering the finite-$n$ approximation of the mappings, which by construction is bijective.
@@
 
-We see that the Column mapping converges to projection on $v_1$, in other words $\alpha=C(v_1, v_2) = v_1$, showing that despite every $C^{n}$ being bijective, the limit of the sequence is not bijective.
+We see that the Column mapping converges to projection on $v_1$, in other words, $\alpha=C(v_1, v_2) = v_1$, showing that despite every $C^{n}$ being bijective, the limit of the sequence is not bijective.
 This becomes clear when looking at the binary expansion of $C$ :
@@
 The Z-mapping however seems to converge to a fractal curve.
-Numerically of course, there seem to be infinite number of discontinuities, but the size of these discontinuities is overwhelmingly small, and *on average* the size of these numerical disconuities vanishes as $n\to\infty$.
-We don't go further into detail about this in this section (in @sec-zcurve-lebesgue we develop on the convergence to a space-filling fractal curve), but this gives some intuition for the arguments presented in @sec-locality.
+Numerically, there seems to be an infinite number of discontinuities, but the size of these discontinuities is overwhelmingly small, and *on average* the size of these numerical discontinuities vanishes as $n\to\infty$.
+We won't go further into detail about this in this section (in @sec-zcurve-lebesgue we develop the convergence to a space-filling fractal curve), but this gives some intuition for the arguments presented in @sec-locality.
 
-We finish this discussion on the limit of the sequence of mappings by asking if there are other mappings similar to the Z-mapping, in the sense that they are pointwise convergent and the limit is bijective[^zmapping-notbijective].
-In an informal way, we established two conditions for this in the context of the mappings described in @sec-mappings-binary :
+We finish this discussion on the limit of the sequence of mappings by asking if there are other mappings similar to the Z-mapping, in the sense that they are pointwise convergent and the limit is bijective.[^zmapping-notbijective] In an informal way, we established two conditions for this in the context of the mappings described in @sec-mappings-binary :
 
 [^zmapping-notbijective]: Despite what can often be found in online resources, the Z-mapping is *not* a bijection because that would violate Netto's theorem.
-The rew_ason for this is the ambiguity in the writing of binary expansions, and in @sec-zcurve-noninjective we give a proof that the Z-mapping is not injective in the $n\to\infty$ limit.
-However the Z-mapping is *quasi-bijective*, in the sense that the set of points for which the Z-mapping is not injective has zero measure, and can therefore be ignored.
+The reason for this is the ambiguity in the writing of binary expansions, and in @sec-zcurve-noninjective we give proof that the Z-mapping is not injective in the $n\to\infty$ limit.
+However, the Z-mapping is *quasi-bijective*, in the sense that the set of points for which the Z-mapping is not injective has zero measure, and can therefore be ignored.
 
-1. condition of pointwise convergence : the LSB of the input must be mapped to the LSB of the output,
-2. condition of "no information loss" : the MSB of the input must be mapped to the MSB of the output.
+1. condition of pointwise convergence: the LSB of the input must be mapped to the LSB of the output,
+2. condition of "no information loss": the MSB of the input must be mapped to the MSB of the output.
 
 These two conditions do not leave a lot of flexibility in the design of new mappings.
-We can of course define mappings where we interleave the bits in a different order, for instance $0.b^2_1 b^1_1 b^2_2 b^1_2 \cdots$ ; or mappings where we reorder the $k$ bits of the input, for instance $0.b^1_2 b^2_2 b^1_1 b^2_1 b^1_4 b^2_4 b^1_3 b^2_3 \cdots$, but the conditions restrict the reordering to be permutations of neighbouring bits.
+We can evidently define mappings where we interleave the bits in a different order, such as $0.b^2_1 b^1_1 b^2_2 b^1_2 \cdots$; or mappings where we reorder the $k$ bits of the input, as in $0.b^1_2 b^2_2 b^1_1 b^2_1 b^1_4 b^2_4 b^1_3 b^2_3 \cdots$, but the conditions restrict the reordering to be permutations of neighboring bits.
 In essence, these mappings are not fundamentally different, and the Z-mapping can be seen as a "canonical representation" of the pointwise convergent mappings with bijective limits.
@@
 Before we argue about how to simulate the dynamics of the 1D neural field, let us simulate the dynamics of the 2D neural field, then at each timestep, we map the 2D neural field to the 1D embedding.
-This will give us a reference of how the 1D neural field is supposed to look like under the condition that its latent dynamics are identical.
+This will give us a reference of how the 1D neural field is supposed to look under the condition that its latent dynamics are identical.
 
@@
 
-how we numerically implemement a simulation on a fractal
+### Numerical aspects of implementing the mappings
 
-In order to simulate the neural field numerically, we discretize by taking a grid of $4^n$ bins in the $[0,1]^2$ space.
+In the previous sections, we have discussed what mappings have good properties, and have given a few hints as to how we can implement them numerically.
+Since the mappings we presented can be described as a binary expansion, they naturally lend themselves to numerical approximations.
+Furthermore, this is made especially easy since the mappings we defined simply are permutations of bits (in the $n < \infty$ case), and so we can write a one-to-one mapping between (unsigned integers).
+
+Without going into too much detail on the programming, we have implemented the mappings as all inheriting from the same abstract class, and each mapping simply has to define a method `indices_to_indices2d(indices: array[int, 1]) -> array[int, 2]` to go from 1D to 2D ($S^{-1}$), and a method `indices2d_to_indices(indices2d: array[int, 2]) -> array[int, 1]` to go from 2D to 1D.[^indices] The abstract class simply defines additional quality-of-life functionality, which makes use of these two methods.
+
+[^indices]: In the source code, only the method `indices(v12: array[float, 2]) -> array[int, 1]` is defined.
+This is because some mappings are implemented (for instance, the `RecursiveLocalMapping`, that we don't discuss here), which don't lend themselves to the simple permutation of bits formulation.
+
+### Numerical approximations of a fractal structure
+
+In the continuation of the simulation methods established in the @sec-simnf, we adopt a grid discretization scheme, as it naturally lends itself to be applied with the mappings.
+We however note that using binning methods, the Monte-Carlo simulation can still be applied.
+The intuition behind this is that sampling populations inside a small square of size $2^{-n} \times 2^{-n}$ gives rise to the notion of an "average connectivity kernel", and in our case the notion of "mean patterns".
+More details are given in @sec-mean-patterns.
+
+We now address the elephant in the room: how do we numerically simulate a neural field defined on -- what appears to be -- a fractal curve? Naively, we could just proceed as follows :
+
+1.
+Create a grid of $2^n \times 2^n = 4^n$ squares of size $2^{-n} \times 2^{-n}$ in the $[0,1]\times[0,1]$ embedding.
+Each square is located by a position $\vec{v_i} = (v_{1,i},v_{2,i}),\,i\in\{1,\cdots\,2^n\}$.
+2. Compute a discretized version of the 1D connectivity kernel
+   $$
+   \begin{aligned}
+   \tilde w(\alpha_i, \beta_j) &= \tilde J_{\alpha_i,\beta_j}\\
+   &= \tilde J_{S(i),S(j)}\\
+   &\stackrel{\text{def}}{=} J_{ij}\\
+   &= w_U(\vec{v_i}, \vec{v_j})
+   \end{aligned}
+   $$
+   at every square location (where we denote for simplicity of notation $S(i)=S(\vec{v_i})$).
+3. Simulate the neural field in $[0,1]$ on the corresponding discretized space with discretization step $4^{-n}$.
+
+The problem with this approach is that since the mappings are just *permutations* of the square populations, the resulting connectivity matrix $\tilde J_{\alpha_i,\beta_j}$ is just a permutation of the rows and columns of $J_{ij}$, and will therefore trivially lead to the identical neural field dynamics.
+
+In some sense, we would like to "take into account" the fact that a mapping of the square populations has been performed.
+Let us consider what happens to the mapping when $n$ is large, and in particular how $S^{n/2}$ and $S^{n}$ behave.
+Since the mappings are pointwise convergent and we are considering large $n$, we argue that both $S^{n/2}$ and $S^{n}$ are good approximations of the true mapping $S$.
+More importantly, $S^{n/2}$ can be considered as an approximation of $S^{n}$.
+This is interesting from a numerical standpoint since it is feasible to compute $S^{n}$, and the approximation $S^{n/2}$ can be obtained by "coarse-graining" $S^n$.
+Another way of thinking about this is that $S^n$ is considered to be the "mathematically exact" fractal structure and $S^{n/2}$ the numerical approximation.
+
+The way we go from $S^n$ to $S^{n/2}$ is by averaging the segment populations by batches of $2^n$, as schematized in @fig-averaging. We call this operation "downsampling" the $S^n$ mapping.
+
+![(TODO) Averaging $2^n$ of the $4^n$ segments populations of $S^n$ results in the $2^n = 4^{n/2}$ segments populations of $S^{n/2}$](figures/photo_2023-06-13_12-15-32.jpg){width=400px #fig-averaging}
+
+The geometrical intuition we can give for this is how the averaging of the segment populations leads to sampling the connectivity kernel at the "average" positions in the $[0,1]^2$ embedding.
+In @fig-averaging-geom, we color-code the square populations corresponding to the segment populations before averaging.
+
+::: {#fig-averaging-geom layout-ncol=2}
+
+![Z-mapping](figures/averaging_z.png)
+
+![Column mapping](figures/averaging_column.png)
+
+(TODO) The corresponding square populations corresponding to the segment populations before downsampling.
+:::
+
+One question that may arise is why we take $n/2$ instead of $n-1$.
+The reason for this is that we want the number of averaged segment populations to scale with $n$.
+If we had chosen $S^{n-1}$, then the number of segment populations is related via
 
 $$
-\int_{[0,1]^2} w_U(\vec v, \vec u) \phi(h_U(t, \vec u)) \mathrm d \vec u \rightarrow \sum_{j=1}^N \underbrace{w_U(\vec u_i, \vec u_j)}_{J_{ij}} \phi(\underbrace{h_U(t, \vec u_j)}_{h_j(t)})
+N_\text{pop}(n-1) = 4^{n-1} = 4^{n} / 4 = N_\text{pop}(n) / 4.
 $$
 
-Each bin therefore corresponds to a population (density) indexed by $i = \{1\cdots4^n\}$. Population $i$ interacts with other populations through the (numerically finite) connectivity $J_{ij}$.
-
-To simulate the neural field in $[0,1]$, we use the mapping $S(\vec u_i) \stackrel{\text{notation}}= S(i) = \alpha_i$ to embed the populations in 1D.
-
-One trivial mapping would be $S(i) = i / 4^n$, which is just ordering the populations in [0,1] in the same order they appear inside the bins.
-More elaborate mappings are simply *permutations* of the numerically binned populations set $\{1\cdots4^n\}$, and the resulting connectivity matrix is then given by $J_{S(i),S(j)}$ [^1].
-
-However, the connectivity matrix is invariant to permutations of neurons, and without further modification, the dynamics emerging from $J_{S(i),S(j)}$ are the same as the ones emerging from $J_{ij}$.
-
-In the same way that discretizing the smooth 2D field works *because of* the smoothness, we discretize the 1D field emerging from the mapping.
-If the mapping has "nice properties" (e.g.
-"locality"), this discretization should work in a similar manner.
-
-For this, we consider the $4^n$ mapped populations $\alpha_i$ to be the "mathematically exact" 1D field, and we "discretize" the 1D field by taking a grid of $2^n$ bins in $[0,1]$.
-In effect, we average bins of $2^n$ populations, in order to obtain $4^n/2^n = 2^n$ coarse-grained populations in $[0,1]$.
-We call this "downsampling" [^2].
-
-The main question is now, what conditions must be imposed on the mapping $S : [0,1]^2 \rightarrow [0,1]$ (numerically : a permutation), such that it is *resilient* to this downsampling, in other words that the dynamics obtained from the connectivity in 1D are the same as the dynamics in 2D [^3].
-
-In the following, this assertion is made by comparing the trajectories $(m_0(t),m_1(t))$ in phase space (also called latent space).
-If the mapping is "resilient", small differences between the trajectories vanish as $n \rightarrow \infty$ [^4].
-
-[^1]: I am glossing over the details of how we bin the grid of $4^n$ elements in $[0,1]^2$ space.
-Numerically, there is an order in which we enumerate each of the $4^n$ bins (the most obvious is column-wise enumeration, i.e.
-``ReshapeMapping.inverse_samples()``).
-As described, this ordering is abitrary, and one should always consider the enumeration of the bins to be in random order.
-
-[^2]: The reason for which we take bins of size $2^n$, is because $2^n$ is the maximum size a bin can have which spans in only one direction in the 2D embedding.
-For instance, with the ``ReshapeMapping`` a column contains $2^n$ populations, all with the same $\xi_0$ position : visually the bin is vertical.
-With the ``DiagonalMapping``, the largest sequence of populations which "go in the same direction" (TODO : clarify this) is of size $2^n$, and spans the diagonal (1,0)->(0,1).
-Intuitively, this averaging can be seen as "destroying fast and large oscillations".
-
-[^3]: The connectivity $\tilde J_{\alpha\beta}$ in 1D can be obtained by computing the "average low-rank vectors". See ``draft.md`` for more details.
-
-[^4]: Each point in $(m_0,m_1)$ fully encodes the state of a linear low-rank RNN.
-We can prove (see my own handwritten notes) that $m_\mu(t) \approx \kappa_\mu(t)$ when $\phi(h)$ is close to linear.
-In our case, $\phi(h)$ is a sigmoid (I found a sigmoid has better convergence properties than a linear function), and can be considered linear as long as $-3 \lesssim h(t) \lesssim 3$.
-
-reintrpretation of the simulation methods
-
--> sampling the 2D connectivity kernel
-
--> average low-rank vectors
-
-![Alt text](figures/Screenshot_2023-06-03_19-59-49.png)]
-
-derivation of $\Jab$ and mean patterns
-
-For finite number of recursive quadrant iterations $n$, we can do a "mean-field approximation" inside each of the $4^n$ segments.
-Let $\alpha = \{i_1,\cdots,i_{|\alpha|}\}$ be the multi-index corresponding to all neurons of which the embedding in $\mathbb R^p$ gets mapped to the segment $\alpha$ in $[0,1]$.
-Let $H_\alpha(t) = \frac 1 {|\alpha|} \sum_{i \in \alpha} h_i(t)$ be the (mean) RNN potential of the segment $\alpha$.
-The connectivity matrix $\tilde J_{\alpha,\beta}$ satisfies
+In other words, we would be averaging a constant number (4 in this case) of segment populations, independently of $n$.
+Choosing a fraction of $n$ for the approximation then allows us to ensure that we average $2^n$ populations, since
 
 $$
-\dot H_\alpha(t) = -H_\alpha(t) + \sum_{\beta \in \text{segments of length } 4^{-n}} \tilde J_{\alpha,\beta} \phi(H_\beta(t))
+N_\text{pop}(n/2) = 4^{n/2} = 4^n / 4^{n/2} = 4^{n} / 2^{n} = N_\text{pop}(n) / 2^n.
 $$
 
-By substituting the original $h_i(t)$, we find the correct rescaling is given by
+Furthermore, this number of averaged populations still vanishes, since
 
 $$
-\tilde J_{\alpha,\beta} = \frac 1 {|\alpha|} \sum_{i \in \alpha} \sum_{j \in \beta} J_{ij}
+N_\text{pop}(n/2) / N_\text{pop}(n) = 2^{-n} \xrightarrow{n\to\infty} 0.
 $$
-
-In the low-rank case, we have
-
-$$
-J_{ij} = \frac 1N \sum_{\mu=1}^p F_{\mu,i} G_{\mu,j}
-$$
-
-We can define the "mean pattern" inside each bin :
-
-$$
-\tilde F_{\mu,\alpha} = \frac{1}{|\alpha|} \sum_{i \in \alpha} F_{\mu,i}, \; \tilde G_{\mu,\alpha} = \frac{1}{|\alpha|} \sum_{i \in \alpha} G_{\mu,i}
-$$
-
-Then the connectivity matrix is given by, noting $N = \sum_{\alpha} |\alpha|$,
-
-$$
-\tilde J_{\alpha,\beta} = \frac{|\beta|}{\sum_{\beta'} |\beta'|} \left( \sum_{\mu=1}^p \tilde F_{\mu,\alpha} \tilde G_{\mu,\beta} - \delta_{\alpha,\beta} \underbrace{\sum_{\mu=1}^p \sum_{i \in \alpha} \frac{F_{\mu,i}}{|\alpha|} \frac{G_{\mu,i}}{|\alpha|}}_{\gamma_{\alpha}} \right)
-$$
-
-![](figures/sampling_R2_nogrid_kernel.png)
-
-Grid method : just do $\tilde w = w_U \circ S^{-1}$
 
@@
 
-#### Continuity on one axis and fractal on the other : from $[0,1]^3$ to $[0,1]^2$
+#### Continuity on one axis and fractal on the other: from $[0,1]^3$ to $[0,1]^2$
 

### 2023-06-13 manuscript/chapters/1_intro.qmd
hehe thats a lot of changes
intro done !

@@
 
-Early recordings of the cat's somatosensory @Mou57 visual cortices suggested that the cortical sheet is organized in vertical columns of functionally similar neurons.
+Early recordings of the cat's somatosensory visual cortices suggested that the cortical sheet is organized in vertical columns of functionally similar neurons @Mou57.
 
-The putative columnar organization of the cortex motivated the design of spatially structured models of neural population dynamics, namely neural field models [@WilCow73;@Nun74;@Ama77], where the spatial dimensions were nothing but the two dimensions of the cortical sheet.
+The putative columnar organization of the cortex motivated the design of spatially structured models of neural population dynamics, namely neural field models [@WilCow73;@Nun74;@Ama77], where the spatial dimensions corresponded to the two dimensions of the cortical sheet.
 
 As models of spatiotemporal neural population dynamics on the cortical sheet, neural field models can be used to explain experimentally observed patterns of cortical activity, such as travelling waves in visual [@SatNau12;@MulRey14], somatosensory [@PetHah03;@FerBol06], motor [@RubRob06;@TakKim15], and hippocampal [@LubSia09;@PatFuj12;@PatSch13] cortices (see @MulCha18 for a review), and are used to model large-scale brain signals, such as electroencephalography (EEG) recordings @Bre17.
-See also @DipRan18 for an example of neural field modelling of 2-photon calcium imaging recording of visual cortex @DipRan18.
+See also @DipRan18 for an example of neural field modelling of 2-photon calcium imaging recording of visual cortex.
 
 More recent recording methods, especially in rodents, have provided evidence of low-dimensional organization of neuronal activity that does not depend on the physical locations of neurons on the cortical sheet.
-For example, contrary to cat's visual cortex, the orientation selectivity of pyramidal neurons in rat's visual cortex does not depend on their locations, that is, orientation selectivity is heterogeneous at any given location @OhkChu05.
-Moreover, the activity of interneurons in any small volume of visual cortex is also heterogeneous and seems to be structured by a low-dimensional transcriptomic manifold of fine cell subtypes @BugDuf22.
-The local functional heterogeneity of neuronal activity in cortex challenges the old concept of functionally homogeneous cortical columns and, with it, classical neural field models.
+For example, contrary to the cat's visual cortex, the orientation selectivity of pyramidal neurons in a rat's visual cortex does not depend on their locations, that is, orientation selectivity is heterogeneous at any given location @OhkChu05.
+Moreover, the activity of interneurons in any small volume of the visual cortex is also heterogeneous and seems to be structured by a low-dimensional transcriptomic manifold of fine cell subtypes @BugDuf22.
+The local functional heterogeneity of neuronal activity in the cortex challenges the old concept of functionally homogeneous cortical columns and, with it, classical neural field models.
 
-While, historically, the spatial dimensions in neural field models have been the two dimensions of the cortical sheet (with the notable exception of the ring models for orientation selectivity in visual cortex @BenBar95 and the head-direction system @Zha96), the ``space'' in neural field models does not need to represent physical space but can represent any suitable abstract embedding space.
+While, historically, the spatial dimensions in neural field models have been the two dimensions of the cortical sheet (with the notable exception of the ring models for orientation selectivity in the visual cortex @BenBar95 and the head-direction system @Zha96), the "space" in neural field models does not need to represent physical space but can represent any suitable abstract embedding space.
 Neural field models with abstract embedding spaces could constitute a generalization of classical neural field theory for cortical networks that are not solely structured by distances of neurons on the cortical sheet.
@@
 If we consider neural fields in abstract embedding spaces, which embedding space should we choose?
-More fundamentally, given some spatio-temporal dynamics, is there a ``natural'' choice for the embedding space?
+More fundamentally, given some spatiotemporal dynamics, is there a "natural" choice for the embedding space?
 
-The goal of this Master Thesis is to show that the answer to the second question is highly nontrivial because even the dimensionality of the embedding space, for some given spatio-temporal dynamics, is not clearly defined if we don't specify the regularity (the ``smoothness'') of the neural fields along the spatial dimensions.
+The goal of this Master Thesis is to show that the answer to the second question is highly nontrivial because even the dimensionality of the embedding space, for some given spatiotemporal dynamics, is not clearly defined.
+This means that it is possible to find two embeddings (and connectivity kernels) of different dimensions that give rise to identical dynamics; contradicting the intuition that the embedding dimension should at least be as high as the dimensionality of the dynamics.
 
@@
 Importantly, these bijections are not (and can't be) diffeomorphisms.
-Also, we show that despite the fact that the connectivity kernel of the one-dimensional equation is not continuous, it is sufficiently regular for its solution to be numerically approximated using standard mesh-based simulations.
+Also, we show that although the connectivity kernel of the one-dimensional equation seems to have a fractal structure, it is sufficiently regular for its solution to be numerically approximated using standard mesh-based simulations.
 
-Our result shows that the notion of (spatial) dimensionality in neural field dynamics is not well-defined if not associated with a constraint on the regularity of the neural fields (or the connectivity kernel).
-Thereby, this work, by studying a simple toy model, illustrates the importance of the analytic notion of regularity when neural population dynamics over abstract continuous spaces are considered.
-
-This work was inspired by recent mathematical development in mean-field theory involving the theory of graphons \cite{JabPoy21}.
+Our results suggest that the notion of (spatial) dimensionality in neural field dynamics is not well-defined if not associated with a constraint on the regularity of the neural fields (and the connectivity kernel).
+Thereby, by studying a simple toy model, this work illustrates the importance of the analytic notion of regularity when neural population dynamics over abstract continuous spaces are considered.
 
@@
 
-<!-- Valentin Schmutz: Which structure?
-At this point, the reader can't guess that you are talking about spatial structure.
-For the logic of this section, you should first gently introduce the notion of spatial structure.
-Nicole : the papers you give for the introduction should give a good understanding of the spatial structure -->
+Let us now give a short introduction on the mathematical formulation of neural fields.
+Neural field equations emerge from the structure of networks of rate neurons, when the number of neurons $N$ becomes large.
+Positions $\vec z_i \in \Rp$ are attributed to each neuron, which defines the *embedding space*.
+Each neuron has a potential $h_i(t)$ that evolves in time, and the weight with which the neurons interact with each other is given by the components of the connectivity matrix $J_{ij} \in \R^{N \times N}$.
+Additionally, the activation function $\phi(h)$ describes the firing rate of a given neuron with potential $h$.
+Then, the product $J_{ij} \phi(h_j)$ gives the *recurrent current* of neuron $j$ on neuron $i$.
 
-Neural field equations emerge from the structure of rate neural networks as the number of neurons $N$ becomes large.
-Positions $\vec z_i \in \Rp$ are attributed to each neuron, and define the *embedding space*.
-If the embedding is well-chosen, then as $N \rightarrow \infty$ neurons close in the embedding space tend to have similar potentials $h_i(t)$, such that these tend to a continuum $h(t, \vec z)$, where $\vec z \in \Rp$.
+![(TODO add phi, add z instead of xi) Every neuron in a rate neural network has a position in the $p$-dimensional embedding space](figures/neural_field_discrete.png)
 
-![Every neuron in a rate neural network has a position in the $p$-dimensional embedding space](figures/neural_field_discrete.png)
+If the embedding is well-chosen,[^1] the connectivity matrix $J_{ij}$ approaches a connectivity kernel $w(\vec y, \vec z)$ as $N \rightarrow \infty$, and neurons close in the embedding space tend to have similar potentials, such that we can write the continous *neural field* $h(t, \vec z)$ of neuron potentials.
+Before we write down the neural field equation which describes how the field evolves in time, we need to introduce the neuron density $\rho(\vec z)$, which measures the number of neurons per unit volume of the embedding space.
+The neural field equation is then written as such:
 
-We write down the neural field equation @eq-nf, in which $\rho(\vec z)$ measures the density of neurons inside the embedding space.
-The connectivity kernel $w(\vec z, \vec y)$ can be seen as the "infinite $N$ limit" of the connectivity matrix $J_{ij}$, in the sense that $w(\vec z, \vec y)$ describes the influence of the neuron population in the box at $[\vec y, \vec y + \d \vec y]$ on the neuron population $[\vec z, \vec z + \d \vec z]$.
+[^1]: Rigorous proofs of convergence are still an active area of study @CheDua19 and are not the subject of this thesis.
 
 $$
-\partial_t h(t, \vec z) = -h(t, \vec z) + \int_{\mathbb{R}^p} w(\vec z, \vec y) \phi(h(t, \vec y)) \rho(\mathrm d \vec y)
+\partial_t h(t, \vec z) = -h(t, \vec z) + \int_{\mathbb{R}^p} w(\vec z, \vec y) \phi(h(t, \vec y)) \rho(\mathrm d \vec y).
 $$ {#eq-nf}
 
-![The connectivity kernel defines the influence between neural densities in the embedding space](figures/connectivity_kernel.png)
-
-In the litterature, the embedding space is often thought of as being closely related to the physical positions of the neurons @Bre11.
-Rigorous proofs of convergence are still an active area of study @CheDua19 and are not the subject of this report.
+![(TODO add phi, add z instead of xi) The connectivity kernel defines the influence between neural densities in the embedding space](figures/connectivity_kernel.png)
 
@@
 
-[TODO : extra motivation, see Valentin's intro]
+This work was inspired by recent mathematical development in mean-field theories involving the theory of graphons @JabPoy21.
+Their main result is that under some boundedness assumptions of the connectivity between the components of a system, the equation of evolution incorporates only an integral over $[0,1]$.
+This is indeed the case in the toy model described next chapter, where weights scale as $J_{ij} \sim \frac 1N$, which motivates the search for a formuation of a one-dimensional neural embedding.
 
-The paper @JabPoy21 provides some motivation for the existence of an equivalent neural field in one dimension.
-Their main result is that under some boundedness assumptions of the connectivity, the limiting equation describing the dynamics of the system incorporates only an integral over $[0,1]$.
-In our low-rank case, the weights scale as $J_{ij} \sim \frac 1N$, and so we can expect that it should be possible to perform a change of variables from $\Rp$ to $[0,1]$.
+The main idea is to define a measurable bijective mapping $S : \Rp \mapsto [0,1]$ and a measure $\lambda : [0,1]^2 \mapsto \mathbb{R}^+$, such that we can write a one-dimensional neural field	$\tilde h(\alpha, t)$ with a connectivity kernel loosely defined as $\tilde w = w \circ S^{-1}$.
 
-The idea is that we define a measurable mapping $S : \Rp \mapsto [0,1]$ and a measure $\lambda : [0,1]^2 \mapsto \mathbb{R}^+$, such that we can write the change of variables in the integral :
-
+<!--
 $$
@@
 $$
-
-Informally, we would like to define the one-dimensional field $\tilde h(\alpha, t) = h(S^{-1}(\alpha), t)$, connectivity $\tilde w(\alpha, \beta) = w(S^{-1}(\alpha), S^{-1}(\beta))$ and density $\tilde \rho(\alpha) = \rho(S^{-1}(\alpha))$, so that we can write the neural field in the $[0,1]$ embedding :
+Informally, we would like to define the one-dimensional field $\tilde h(\alpha, t) = h(S^{-1}(\alpha), t)$, connectivity $\tilde w(\alpha, \beta) = w(S^{-1}(\alpha), S^{-1}(\beta))$ and density $\tilde \rho(\alpha) = \rho(S^{-1}(\alpha))$, so that we can write the neural field in the $[0,1]$ embedding : -->
 
 $$
-\partial_t \tilde h(t, \alpha) = -\tilde h(t, \alpha) + \int_{[0,1]} \tilde w(\alpha, \beta) \phi(\tilde h(t, \beta)) \tilde \rho(\mathrm d \beta)
-$$ {#eq-nf1d}
+\partial_t \tilde h(t, \alpha) = -\tilde h(t, \alpha) + \int_{[0,1]} \tilde w(\alpha, \beta) \phi(\tilde h(t, \beta)) \left[\lambda \circ S^{-1}\right](\d \beta)
+$$
 
-This change of variable seems counter-intuitive : how is it possible to map a higher dimension to a lower dimension without losing information ?
-And even if mathematically such a change of variables is possible, is it actually feasible to simulate it numerically ?
+This change of variable in the kernel seems counter-intuitive : how is it possible to map a higher dimension to a lower dimension without losing information?
+And even if mathematically such a change of variables is possible, is it even feasible to simulate it numerically?
 
-## Numerical simulations of neural fields
+## Numerical simulations of neural fields {#sec-intro-simnf}
 
-The goal of this report is to not only give an intuition on the mappings between $[0,1]$ and $\Rp$ embedding spaces, but also show that the mathematical construction works numerically.
+The goal of this thesis is to not only give an intuition on the mappings between the $[0,1]$ and $\Rp$ embedding spaces, but also show that the mathematical construction works numerically.
 
-We distinguish two types of numerical simulations for the neural field in @eq-nf :
+We distinguish two types of numerical simulations for the neural field in @eq-nf:
 
@@
 Suppose we take bins in the $\Rp$ space, and map each bin to a random position in $[0,1]$.
-Intuitively, this won't work (in the sense that the dynamics in $[0,1]$ will not be the same as the dynamics in $\Rp$), because the hypothesis of continuity is not longer true : the mapping is random, so the nearby populations in $[0,1]$ will no longer have similar potentials, and the error of the mean-field approximation will no longer vanish as we take increasingly many bins.
+Intuitively, this won't work (in the sense that the dynamics in $[0,1]$ will not be the same as the dynamics in $\Rp$), because the hypothesis of continuity is not longer true: the mapping is random, so the nearby populations in $[0,1]$ will no longer have similar potentials, and the error of the mean-field approximation will no longer vanish as we take increasingly many bins.
 
-\[TODO : an illustration for this\]
+![(TODO) Good mappings from $\Rp$ to $[0,1]$ should maintain some sense of "locality"](figures/photo_2023-06-13_18-15-50.jpg){width=500px}
 
-In some sense, we would like to find a mapping that conserves "locality" : populations that are close in $[0,1]$ should represent nearby populations in $\Rp$, or at least there shouldn't be "too many large discontinuities".
+In some sense, we would like to find a mapping that conserves "locality": populations that are close in $[0,1]$ should represent nearby populations in $\Rp$, or at least there shouldn't be "too many large discontinuities".
 
@@
 We build a metric to measure the "discretization error", and provide the key arguments for a proof of convergence of the numerical approximations.
-
-<!--
-
-neurons in some Rp space
-N -> infty
-spatial structure
-	=> close neurons -> similar potential
-	=> continuum of neurons
-	=> neural field
-
-"natural embedding" (here in Rp) is NOT UNIQUE
-
-we show : 
-* we can embed in a space of lower dimension.
-* not just a math abstraction (solution of the 1d neural field is well-defined)
-* not only well-defined, but actually works numerically. we assert the dynamics in [0,1] are the same
-
-[how do we actually "convince ourselves" that is works numerically ?
--> little intro on "discretizing a neural field", either by sampling the PDF, or by discretization
--> particle (=neurons)-based simulation vs grid (=bins)/density (TODO)-based simulation
-(with the discretization approach we already have the "infinite number of neurons" approximation inside each bin)
-
-def sampling : tirer au sort un neurone selon la pdf
-def discretization : discretize pdf regularly, "bins"
-]
-
-introduce "RandomMapping", to show why this mapping is not trivial
-
-why it doesnt work : even in a little bin, we will have many "wildly different neurons", in the sense scattered far away in 2D
-
-what we want in [0,1] : as bins become smaller, the "populations inside the bin" have more and more similar potentials. in the random case, never happens.
-if in 2D, they are close, then they have the same potential.
-close in 2D ?=> close in 1D -> "locality" (how do formulate a mapping which has this property ?) (something close to an "isometric mapping", distances are kind of conserved)
-
-(we show we can afford "some" discontinuity)
- -->

### 2023-06-13 manuscript/chapters/3_mappings.qmd
intro done !

@@
 
-TODO : might need to talk about how $S(v^{(n)}_1,v^{(n)}_2) = S^n(v^{(n)}_1,v^{(n)}_2)$ in appendix ?
+TODO NOTE : might need to talk about how $S(v^{(n)}_1,v^{(n)}_2) = S^n(v^{(n)}_1,v^{(n)}_2)$ ?
 so basically the limit $n\to\infty$ is the limit where we take infinite precision in the input.
@@
 TODO NOTE : talk about how we center the bins instead of taking the corners.
-this is not super important in the n -> oo limit, but numerically at 0 and 1 the CDF-1 goes to infty.
-at infinity, the connectivity kernel is infinite but we have zero neurons (zero measure) (linear growth vs exponential tail of the gaussina) : effectively the recurrent current is zero and we can safely ignore these effects.
+this is not super important in the $n\to\infty$ limit, but numerically at 0 and 1 the $\mathrm{CDF}^{-1}$ goes to infty.
+at infinity, the connectivity kernel is infinite but we have zero neurons (zero measure, linear growth of kernel vs exponential tail of the gaussian) : effectively the recurrent current is zero and we can safely ignore these effects.
 

### 2023-06-13 manuscript/chapters/1_intro.qmd
look a kitty !!! *baps with paw*

@@
 As models of spatiotemporal neural population dynamics on the cortical sheet, neural field models can be used to explain experimentally observed patterns of cortical activity, such as travelling waves in visual [@SatNau12;@MulRey14], somatosensory [@PetHah03;@FerBol06], motor [@RubRob06;@TakKim15], and hippocampal [@LubSia09;@PatFuj12;@PatSch13] cortices (see @MulCha18 for a review), and are used to model large-scale brain signals, such as electroencephalography (EEG) recordings @Bre17.
-See also @DipRan18 for an example of neural field modelling of 2-photon calcium imaging recording of visual cortex.
+See also @DipRan18 for an example of neural field modelling of 2-photon calcium imaging recording of the visual cortex.
 
@@
 
-Let us now give a short introduction on the mathematical formulation of neural fields.
-Neural field equations emerge from the structure of networks of rate neurons, when the number of neurons $N$ becomes large.
+Let us now give a short introduction to the mathematical formulation of neural fields.
+Neural field equations emerge from the structure of networks of rate neurons when the number of neurons $N$ becomes large.
 Positions $\vec z_i \in \Rp$ are attributed to each neuron, which defines the *embedding space*.
@@
 
-If the embedding is well-chosen,[^1] the connectivity matrix $J_{ij}$ approaches a connectivity kernel $w(\vec y, \vec z)$ as $N \rightarrow \infty$, and neurons close in the embedding space tend to have similar potentials, such that we can write the continous *neural field* $h(t, \vec z)$ of neuron potentials.
+If the embedding is well-chosen,[^1] the connectivity matrix $J_{ij}$ approaches a connectivity kernel $w(\vec y, \vec z)$ as $N \rightarrow \infty$, and neurons close in the embedding space tend to have similar potentials, such that we can write the continuous *neural field* $h(t, \vec z)$ of neuron potentials.
 Before we write down the neural field equation which describes how the field evolves in time, we need to introduce the neuron density $\rho(\vec z)$, which measures the number of neurons per unit volume of the embedding space.
@@
 Their main result is that under some boundedness assumptions of the connectivity between the components of a system, the equation of evolution incorporates only an integral over $[0,1]$.
-This is indeed the case in the toy model described next chapter, where weights scale as $J_{ij} \sim \frac 1N$, which motivates the search for a formuation of a one-dimensional neural embedding.
+This is indeed the case in the toy model described next chapter, where weights scale as $J_{ij} \sim \frac 1N$, which motivates the search for a formulation of a one-dimensional neural embedding.
 
@@
 
-This change of variable in the kernel seems counter-intuitive : how is it possible to map a higher dimension to a lower dimension without losing information?
+This change of variable in the kernel seems counter-intuitive: how is it possible to map a higher dimension to a lower dimension without losing information?
 And even if mathematically such a change of variables is possible, is it even feasible to simulate it numerically?

### 2023-06-13 manuscript/chapters/2_neural_fields.qmd
hehe thats a lot of changes
guess who just went through half a book in one day :))
nearly finished chap 3 !
look a kitty !!! *baps with paw*

@@
 
+In this chapter, we introduce a toy model so that we can later study the mappings of embedding spaces.
+We stress that this toy model serves only for illustration purposes, and our results apply to any neural field.
+For this reason, we only cite or give short derivations of results used to understand the model, and the influence of the mappings that are later applied to it.
+
 ## Networks of neurons to neural fields
 
-### Low-rank network of neurons
+### Low-rank networks of neurons
 
-The results we present in this thesis should hold for any neural field, but for the sake of argument we shall consider a low-rank rate neural network, for which there are informal proofs of convergence to a neural field equation  @schmutz2023convergence \[TODO : reformulate this\].
-Low-rank neural networks hold their name from the rank of their connectivity matrix, defined in @eq-lrj.
+The toy model we introduce is an instance of a low-rank rate neural network.
+Such models hold their name from the form of their connectivity matrix, which has a rank of (at most) $p$.
+We write the connectivity matrix as
 
 $$
-J_{ij} = \frac 1N \sum_{\mu=1}^p F_{\mu,i} G_{\mu,j}
-$$ {#eq-lrj}
+J_{ij} = \frac 1N \sum_{\mu=1}^p F_{\mu i} G_{\mu j}.
+$$ {#eq-lowrankj}
 
-In the case that $\vec F_i = (F_{1,i},\cdots,F_{p,i}) \in \Rp$ sample from some $p$-dimensional distribution $\rho$, the vectors $\vec F_i$ define a "natural embedding" of neuron $i$ in $\Rp$.
-When the number of neurons increases, the numeric density of neurons in the embedding approaches the actual probability distribution $\rho$, and we can intuitively see the convergence towards a connectivity kernel $w(\vec z, \vec y) = \sum_{\mu=1}^p z_\mu g_\mu(y_\mu)$.
+The geometric view of these networks is that the recurrent currents lie in a $p$-dimensional subspace of $\R^N$ spanned by the vectors $\{\vec{F_1}, \cdots, \vec{F_p}\}$ (where $\vec{F_\mu}=(F_{\mu 1}, \cdots F_{\mu N})$), which therefore form a "natural" embedding.
 
-start with the study of network of neurons, no external current
+Recalling the notation from the introduction, $h_i(t)$ are the neuron potentials (with initial condition $h_i(0)$), $\phi : \R \mapsto \R^+$ is the monotonic increasing activation function, and we write the evolution equation for this network of neurons:
 
-[TODO MATH : show self-connections $\sim \frac 1N$, so we ignore them]
+$$
+\dot h_i(t) = \underbrace{-h_i(t)}_\text{exponential decay} + \underbrace{\frac 1N \sum_{\mu=1}^p \sum_{j=1}^{N} F_{\mu i} G_{\mu j} \phi(h_j(t))}_\text{recurrent current $I^\text{rec}_i(t)$}.
+$$ {#eq-lowrank-evolution}
 
-general setting of low-rank networks
+The exponential decay term describes how, in isolation, a neuron's potential tends to zero, biologically corresponding to a depolarized membrane.
+The recurrent current term $I^\text{rec}_i(t)$ describes the "currents" received by neuron $i$ from all the other neurons in the network.
+
+We would like to add a few remarks on the physical units of @eq-lowrank-evolution.
+Strictly speaking, a prefactor $\tau$ with units of $\mathrm{seconds}^{-1}$ should be multiplied with $\dot h_i(t)$ of units $\mathrm{volt}\cdot\mathrm{seconds}^{-1}$, such that the left hand side of the equation is consistent with the unit of voltage.
+Similarly, the activation function has units $\mathrm{seconds}^{-1}$ and the connectivity matrix units of $\mathrm{coulomb}$, and a multiplicative term $R$ with units $\mathrm{ohm}$ should multiply the recurrent currents in order to yield a voltage.
+By rescaling of time or connection weights both $\tau$ and $R$ can be set to one, and so we omit to write them.
+
+One last remark is that in our model, we omit to add any source of external currents: from the initial condition, the network is left to evolve in isolation.
+
+### Gaussian $\mathcal{N}(\vec 0, \mathbb{1}_N)$ low-rank network of neurons
+
+Let us now specify the low-rank model introduced in the previous section, and use the model introduced in @schmutz2023convergence.
+This paper defines the low-rank vectors $F_{\mu i}$ such that each component independantly samples a standard gaussian (zero mean, unit variance).
+In other words, every neuron $i$ samples a vector $\vec{F_i} = (F_{1i}, \cdots, F_{pi})$ from the $p$-dimensional gaussian ball.
+We write
+
+$$
+\vec{F_i} = \vec{z_i},\ \vec{z_i} \stackrel{\text{iid}}{\sim} \rho(z_1, \cdots, z_p),
+$$
+
+where
 
@@
 \begin{aligned}
-\dot h_i(t) =& -h_i(t) + \frac 1N \sum_{\mu=1}^p \sum_{j=1}^N F_{\mu i} G_{\mu j} \phi(h_j(t)) \\
-F_{\mu i} =& f_\mu(z_{1 i}, \cdots, z_{p i}), \\
-G_{\mu i} =& g_\mu(z_{1 i}, \cdots, z_{p i}), \\
-&\vec{z_i} \stackrel{\text{i.i.d.}}{\sim} \rho
+\rho(z_1, \cdots, z_p) &= \prod_{\mu=1}^p \mathcal{N}(z_\mu) \\
+\mathcal{N}(z) &= \frac{1}{\sqrt{2 \pi}} \mathrm{e}^{-\frac 12 z^2}.
 \end{aligned}
@@
 
-interpretation : each neuron has a position $\vec{z_i} \in \Rp$. $J_{ij} = \frac 1N \sum_{\mu=1}^p = F_{\mu i} G_{\mu j}$ is the weight of recurrent currents.
-
-### Gaussian $\mathcal{N}(\vec 0, \mathbb{1}_N)$ low-rank network of neurons
-
-#### Description
-
-our setting is :
+The vectors $G_{\mu i}$ are defined as
 
 $$
-\begin{aligned}
-f_\mu(z_1, \cdots, z_p) &= z_\mu \\
-g_\mu(z_1, \cdots, z_p) &= \tilde \phi(z_\mu) = \frac{\phi(z_\mu) - \avg{\phi(z_\mu)}}{\mathrm{Var}[\phi(z_\mu)]} \\
-\phi(z) &= \frac{1}{1+\mathrm{e}^{-z}} \\
-\vec{z} &= (z_1, \cdots, z_p) \sim \rho(z_1, \cdots, z_p) = \prod_{\mu=1}^p \mathcal{N}(z_\mu),\\
-&\text{where}\ \mathcal{N}(z) = \frac{1}{\sqrt{2 \pi}} \mathrm{e}^{-\frac 12 z^2}
-\end{aligned}
+G_{\mu i} = \tilde\phi(z_{i \mu}) \stackrel{\text{def}}{=} \frac{\phi(z_{i \mu}) - \avg{z_{i \mu}}}{\mathrm{Var}[\phi(z_{i \mu})]}.
 $$
 
-#### Study of fixed points and their stability
+This choice be elucidated in a few paragraphs.
+Finally, the activation function remains to be defined.
+For simplicity, we take it to be the logistic function, although the following results are not sensitive to this choice.
 
-* $h_i = z_{\mu,i} \quad \forall \mu = 1,\cdots,p$ (stable)
-* $h_i=0$ (unstable)
+$$
+\phi(h) = \frac{1}{1 + \mathrm{e}^h}
+$$
 
-the fixed points also become more and more stable as N->oo (study of stability with eigenvalues)
+Putting everything together, we get the following expression for the equation of evolution of the network of neurons:
 
-talk about since $\lambda_\nu = \avg{g^\prime_\nu(z_\nu) f_\nu(z_\nu)}-1$, the sharper the gain, the more stable to fixed points (since the bump of $\delta \phi$ becomes more concentrated)
+$$
+\dot h_i(t) = -h_i(t) + \frac 1N \sum_{\mu=1}^p \sum_{j=1}^{N} z_{\mu i} \tilde \phi(z_{\mu j}) \phi(h_j(t)).
+$$ {#eq-toymodel-evolution}
 
-![spectrum of $K$ for $N=64$](figures/spectrum_N=64.png)
+We note that contrary to the model introduced in @schmutz2023convergence, we ignore the self-connections term, because it introduces a vanishing correction of order $\bO(1/N)$.
+
+In order to understand this model better, we can do an analysis of its fixed points and their stability.
+Here we summarize the results from the derivation presented in @sec-fixedpoints.
+We use the notation $h^\star$ to refer to fixed points of the network.
+The fixed points solve the roots of the evolution equation:
+
+$$
+\dot h_i(t)\ \text{evaluated at}\ h_i^\star = -h^\star_i(t) + \frac 1N \sum_{\mu=1}^p \sum_{j=1}^{N} z_{\mu i} \tilde \phi(z_{\mu j}) \phi(h^\star_j(t)) = 0
+$$
+
+* $h^\star_i = z_{\mu i}$ is a stable fixed point for all $\mu \in \{1,\cdots,p\}$.
+Additionally, because the Gaussian distribution is symmetric around zero, $h_i = -z_{\mu i}$ are also stable fixed points.
+We refer to these fixed points as the "pattern" fixed points, because the network of neurons "remembers" the vectors $\vec{F_1}, \cdots, \vec{F_p}$.
+* $h^\star_i=0$ is an unstable fixed point. It corresponds to all the neuron potentials being set to zero.
+
+The analysis of stability from @sec-fixedpoints-stability can be summarized by the study of the eigenvalues of the matrix $K_{ij} = J_{ij} \partial \phi(h^\star_j) - \mathrm{Id}_{ij}$.
+
+1.
+Taking Taylor expansions of the activation functions reveals that -- at least up to order 3 -- uneven powers tend to increase stability of the pattern fixed points (and reciprocally, decrease stability of the zero fixed point), and vice-versa for the even powers (with the notable exception of the constant offset, which doesn't play a role in this analysis).
+2.
+A steeper slope at the origin of the activation function also seems to improve stability of the pattern fixed points.
+This corresponds to a sharper difference between the "inactive" (low potential) and the "active" (high potential) neurons.
+The intuition behind this is that if the slope is stronger, then the bump of $\partial \phi$ becomes "more concentrated", which helps lower the eigenvalues associated to the pattern fixed points.
+3. The spectrum of $K$ is composed of $p$ eigenvalues that depend on the fixed point, and $N-p$ eigenvalues $\lambda_\perp = -1$ corresponding to the orthogonal complement of the pattern fixed points.
+
+In our case, this analysis in Taylor expansions was sufficient to explain the observed stability resulting from a logistic activation function.
+
+We should however note that the results obtained in @sec-fixedpoints are valid in the $N \to \infty$ limit.
+A numerical analysis shows in @fig-spectrum that already at $N=64$, the analytic eigenvalues match the numerical eigenvalues closely, and this correspondance improves when we take larger $N$ (typically we will for the rest of this Thesis take $N > 10^3$).
+
+![(TODO add spectrum for zero fixed point) Spectrum of $K$ for $N=64$ for the pattern fixed point](figures/spectrum_N=64.png){#fig-spectrum}
 
@@
 
-### Simulations of the low-rank neural field
+## Simulation schemes for neural fields {#sec-simnf}
 
@@
 
-#### Sampling method (Monte-Carlo integration)
+### Sampling method (Monte-Carlo integration)
 
@@
 
-#### Grid method (Trapezoidal integration)
+### Grid method (Trapezoidal integration) {#sec-gridmethod}
 
@@
 
-## Mapping $\R^2$ to $[0,1]^2$ using the CDF
+## Mapping $\R^2$ to $[0,1]^2$ using the CDF {#sec-mapping-cdf}
 

### 2023-06-13 manuscript/chapters/3_mappings.qmd
look a kitty !!! *baps with paw*

@@
 
-We finish this discussion on the limit of the sequence of mappings by asking if there are other mappings similar to the Z-mapping, in the sense that they are pointwise convergent and the limit is bijective.[^zmapping-notbijective] In an informal way, we established two conditions for this in the context of the mappings described in @sec-mappings-binary :
+We finish this discussion on the limit of the sequence of mappings by asking if there are other mappings similar to the Z-mapping, in the sense that they are pointwise convergent and the limit is bijective.[^zmapping-notbijective] In an informal way, we established two conditions for this in the context of the mappings described in @sec-mappings-binary:
 
@@
 These two conditions do not leave a lot of flexibility in the design of new mappings.
-We can evidently define mappings where we interleave the bits in a different order, such as $0.b^2_1 b^1_1 b^2_2 b^1_2 \cdots$; or mappings where we reorder the $k$ bits of the input, as in $0.b^1_2 b^2_2 b^1_1 b^2_1 b^1_4 b^2_4 b^1_3 b^2_3 \cdots$, but the conditions restrict the reordering to be permutations of neighboring bits.
+We can evidently define mappings where we interleave the bits in a different order, such as $0.b^2_1 b^1_1 b^2_2 b^1_2 \cdots$; or mappings where we reorder the $k$ bits of the input, as in $0.b^1_2 b^2_2 b^1_1 b^2_1 b^1_4 b^2_4 b^1_3 b^2_3 \cdots$, but the conditions restrict the reordering to be permutations of neighbouring bits.
 In essence, these mappings are not fundamentally different, and the Z-mapping can be seen as a "canonical representation" of the pointwise convergent mappings with bijective limits.
@@
 
-### Numerical aspects of implementing the mappings
+### Practical aspects of implementing the mappings
 
@@
 The geometrical intuition we can give for this is how the averaging of the segment populations leads to sampling the connectivity kernel at the "average" positions in the $[0,1]^2$ embedding.
-In @fig-averaging-geom, we color-code the square populations corresponding to the segment populations before averaging.
+In @fig-averaging-geom, we colour-code the square populations corresponding to the segment populations before averaging.
 

### 2023-06-14 manuscript/chapters/2_neural_fields.qmd
meow meow i finish section 2 !!!!

@@
 
-## Networks of neurons to neural fields
+## Networks of neurons to neural fields {#sec-networks-of-neurons-to-neural-fields}
 
-### Low-rank networks of neurons
+### Low-rank networks of neurons {#sec-low-rank-network}
 
@@
 We would like to add a few remarks on the physical units of @eq-lowrank-evolution.
-Strictly speaking, a prefactor $\tau$ with units of $\mathrm{seconds}^{-1}$ should be multiplied with $\dot h_i(t)$ of units $\mathrm{volt}\cdot\mathrm{seconds}^{-1}$, such that the left hand side of the equation is consistent with the unit of voltage.
+Strictly speaking, a prefactor $\tau$ with units of $\mathrm{seconds}^{-1}$ should be multiplied with $\dot h_i(t)$ of units $\mathrm{volt}\cdot\mathrm{seconds}^{-1}$, such that the left-hand side of the equation is consistent with the unit of voltage.
 Similarly, the activation function has units $\mathrm{seconds}^{-1}$ and the connectivity matrix units of $\mathrm{coulomb}$, and a multiplicative term $R$ with units $\mathrm{ohm}$ should multiply the recurrent currents in order to yield a voltage.
-By rescaling of time or connection weights both $\tau$ and $R$ can be set to one, and so we omit to write them.
+By rescaling time or connection weights, both $\tau$ and $R$ can be set to one, and so we omit to write them.
 
@@
 Let us now specify the low-rank model introduced in the previous section, and use the model introduced in @schmutz2023convergence.
-This paper defines the low-rank vectors $F_{\mu i}$ such that each component independantly samples a standard gaussian (zero mean, unit variance).
+This paper defines the low-rank vectors $F_{\mu i}$ such that each component independently samples a standard Gaussian (zero mean, unit variance).
 In other words, every neuron $i$ samples a vector $\vec{F_i} = (F_{1i}, \cdots, F_{pi})$ from the $p$-dimensional gaussian ball.
@@
 
-In order to understand this model better, we can do an analysis of its fixed points and their stability.
+To understand this model better, we can analyze its fixed points and their stability.
 Here we summarize the results from the derivation presented in @sec-fixedpoints.
@@
 We should however note that the results obtained in @sec-fixedpoints are valid in the $N \to \infty$ limit.
-A numerical analysis shows in @fig-spectrum that already at $N=64$, the analytic eigenvalues match the numerical eigenvalues closely, and this correspondance improves when we take larger $N$ (typically we will for the rest of this Thesis take $N > 10^3$).
+A numerical analysis shows in @fig-spectrum that already at $N=64$, the analytic eigenvalues match the numerical eigenvalues closely, and this correspondence improves when we take larger $N$ (typically we will for the rest of this Thesis take $N > 10^3$).
 
@@
 
-show simulations, first in $p=1$, then $p=2$.
+In the introduction, we introduced the convergence of a network of neurons to a smooth neural field when the embedding space is well-chosen.
+For our toy model, it seems natural to try to embed the neurons according to their positions on the $p$-dimensional Gaussian ball.
 
-![simulation in 1D, p=1](figures/embedding_d=1_h0=0_phi=sigmoid.mp4)
+#### Numerical aspects of simulating a low-rank network of neurons
 
-![simulation in 2D, p=2. TODO redo this one with only scatter and latent-](figures/embedding_d=2_h0=0.mp4)
+We first demonstrate what happens with numerical simulations in the simple case of $p=1$.
+We generate a network of 50,000 neurons, where each neuron samples a one-dimensional Gaussian to get its position.
+The function $\tilde \phi$ is computed by rescaling $\phi$ with the numerically computed mean and variance.[^code1]
 
-show it converges to the fixed points $z_1$ and $z_2$
+Instead of numerically computing the full connectivity matrix $J$ (which would require storing $N^2 = 50,000^2 \sim 10^{10}$ entries, taking about 40 GiB if we use 32-bit floating point numbers...
+yikes !), we take advantage of its low-rank structure to compute recurrent currents, therefore only storing $p \times N$ numbers instead of $N^2$.[^code2]
 
-the structure of the scatterplot emerges naturally as a smooth manifold in $\R^2$
+[^code1]: In the source code, implementation for this can be found under `notebooks/neurodyn/_rnn.py`, in the method [`LowRankRNNParams.new_sampled_valentin`](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_rnn.py#L154).
 
-introduce the neural field as a "smooth fitting of the scatterplot"
+[^code2]: See the implementation in `notebooks/neurodyn/_rnn.py`, in the method [`LowRankRNN.I_rec`](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_rnn.py#L302).
 
-eq of neural field
+#### Naive and natural embeddings for $p=1$
+
+Initially, we set all the potentials to zero, then simulate the differential equation by using the Runge-Kutta of order 4 integration schema (see [`scipy.integrate.solve_ivp`](https://docs.scipy.org/doc/scipy/reference/generated/scipy.integrate.solve_ivp.html)).
+In @fig-p1, we show the resulting dynamics for a few thousand neurons.
+We order the neurons as they appear in the numerical array (a naive "random" one-dimensional embedding), and compare this to the ordering of neurons by their sampled position $z_1 \sim \mathcal{N}$.
+
+![(TODO remove overlap, change notation, color for activity) Simulation of the Gaussian network of neurons with $p=1$.
+Every neuron is represented by a dot, of which the height represents its activity level.](figures/embedding_d=1_h0=0_phi=sigmoid.mp4){width=400px #fig-p1}
+
+The results of the simulation show that the random ordering of the neurons does not yield a smooth embedding, on contrary of the smooth surface that appears when we order the neurons by their position $z_{1i}$.
+The simulation serves as a demonstration that the neurons converge to the stable fixed point $h^\star_i = z_{1i}$, as well as confirming that the zero fixed point is unstable.
+Indeed, despite the zero initialization, the simulation did not stay there because small numerical fluctuations were amplified by the instability.
+
+The observed smooth surface corresponds to the activation function $\phi(h^\star) = \phi(z_1)$.
+Therefore, the Gaussian patterns define a natural embedding of the network, and the neural field at the fixed point reads $h^\star(z_1, t) = z_1$.
+We can write the neural field equation for $p=1$ in the Gaussian embedding as
+
+$$
+\partial_t h(z_1, t) = -h(z_1, t) + \int_{\R} z_1 \tilde\phi(y_1) \phi(h(y_1, t)) \mathcal{N}(\d y_1).
+$$ {#eq-nf-gaussian-p1}
+
+#### A natural embedding for $p=2$
+
+Because @eq-nf-gaussian-p1 already involves a one-dimensional integral, the $[0,1]$ embedding can easily be found by applying rescaling methods, such as the inverse CDF method described in @sec-mapping-cdf later in this chapter.
+
+The question of finding a one-dimensional embedding becomes non-trivial when we consider $p>1$, here we take $p=2$.
+In @fig-p2, we repeat the same simulations as in the previous section.
+Again, the structure of the network emerges naturally as a smooth surface in $\R^2$.
+
+![(TODO remove overlap, change notation, scatter only) Simulation of the Gaussian network of neurons with $p=2$.
+Every neuron is represented by a dot, of which the height represents its activity level.](figures/embedding_d=2_h0=0.mp4){width=500px #fig-p2}
+
+We however note that in this case, there are two pattern fixed points that the network can converge to.
+In the example given here the network coincidentally, due to finite numerical effects, converged again to the first pattern fixed point, which we can see because the neural field is constant in the $z_2$ direction.
+We write the fixed point $h^\star(z_1, z_2, t) = z_1$.
+The neural field is now written as
+
+$$
+\begin{aligned}
+\partial_t h(z_1, z_2, t) &= -h(z_1, z_2, t)\\
+&+ \int_{\R} \left(z_1 \tilde\phi(y_1) + z_2 \tilde\phi(y_2)\right) \phi(h(y_1, y_2, t)) \mathcal{N}(\d y_1) \mathcal{N}(\d y_2).
+\end{aligned}
+$$ {#eq-nf-gaussian-p2}
+
+More generally, we can write the $p$-dimensional neural field, where $\vec z \in \Rp$:
+
+$$
+\begin{aligned}
+\partial_t h(\vec z, t) &= -h(\vec z, t) + \int_{\Rp} w(\vec z, \vec y) \phi(h(\vec y, t)) \mathcal{N}^p(\d \vec y) \\
+w(\vec z, \vec y) &= \sum_{\mu=1}^p z_\mu \tilde\phi(y_\mu) \\
+\mathcal{N}^p(\d \vec y) &= \frac{\mathrm{exp}(\frac12\sum_{\mu=1}^p z_\mu^2)}{(2\pi)^{p/2}} \mathrm{d} y_1 \cdots \mathrm d y_p.
+\end{aligned}
+$$ {#eq-nf-gaussian-p}
 
@@
 
-we use the RK4 integration scheme, and the integral is decoupled into product of 1D integrals, each is estimated using a dot product
+Now that we have introduced the analytical neural fields, let us discuss how we can simulate them numerically, more specifically how we can estimate the $p$-dimensional integral.
 
-### Sampling method (Monte-Carlo integration)
+### Sampling method (Monte Carlo integration)
 
-Neurons are at position $\vec z = (z_1, \cdots, z_p) \in \mathbb{R}^p$, distributed according to the distribution $\frac{\mathrm{exp}(\frac12\sum_{\mu=1}^p z_\mu^2)}{(2\pi)^{p/2}} \mathrm{d} z_1 \cdots \mathrm d z_p = \rho(\mathrm d \vec z)$
-
-The RNN potential becomes $h(t, \vec z)$ and evolves according to
-
-$$
-\begin{aligned}
-\partial_t h(t, \vec z) &= -h(t, \vec z) + \int_{\Rp} w(\vec z, \vec y) \phi(h(t, \vec y)) \rho(\mathrm d \vec y),
-\quad
-w(\vec z, \vec y) = \sum_{\mu=1}^p \tilde \phi (y_\mu) z_\mu
-\end{aligned}
-$$
-
-repeat the fixed points are now the generalization, functions $h=z_1$ and $h=z_2$
-
-simulation : when we simulate a network of neurons with $N \gg 1$, we are approximating the integral using a monte carlo method.
-
-Monte-carlo estimation of the integral is
+Due to the formulation of the neural field weighed by a probability distribution $\rho(y_1, \cdots, y_p)$, a natural way to estimate the integral is by Monte Carlo integration.
+The essence of this method is the use of the Central Limit Theorem, where we take $N$ independent samples $\vec{y_i} = (y_{1i}, \cdots, y_{pi})$ from the distribution $\rho$, and use them to estimate the integrand.
 
@@
 \mathcal{I}(\vec z) &= \int_{\Rp} \underbrace{w(\vec z, \vec y) \phi(h(t, \vec y))}_{I(\vec z, \vec y)} \rho(\d \vec y) \\
-\hat{\mathcal{I}}(\vec z) &= \frac 1N \sum_{i=1}^N I(\vec z, \vec{y_i}), \, \vec{y_i} \stackrel{\text{i.i.d.}}{\sim} \rho
+\hat{\mathcal{I}}(\vec z) &= \frac 1N \sum_{i=1}^N I(\vec z, \vec{y_i}), \, \vec{y_i} \stackrel{\text{iid}}{\sim} \rho
+\end{aligned}
+$$ {#eq-mc}
+
+Applying the Central Limit Theorem, we get that the estimation $\mathcal I(\vec z)$ from @eq-mc has a convergence rate of $\bO(1/\sqrt{N})$, since
+
+$$
+\begin{aligned}
+\mathrm{Var}\left[\hat{\mathcal{I}}(\vec z) \right] &= \frac 1N \mathrm{Var}_{\vec y \sim \rho}[I(\vec z, \vec y)].
 \end{aligned}
@@
 
-Numerically, we are estimating $\hat{\mathcal{I}}(\vec z)$, and we can show the convergence rate is in $\bO(1/\sqrt{N})$, since
+The attentive reader might have noticed by now that the Monte-Carlo method just reduces to simulating the network of neurons @eq-toymodel-evolution.
+Effectively, the network of neurons converges to the neural field with rate $\bO(1/\sqrt{N})$.
 
-$$
-\begin{aligned}
-\mathrm{Var}\left[\hat{\mathcal{I}}(\vec z) \right] &= \frac 1N \mathrm{Var}_{\vec y \sim \rho}[I(\vec z, \vec y)]
-\end{aligned}
-$$
-
-Detail : strictly speaking, MC samples new points every timestep, here we sample once at t=0, then propagate these points
+One small nuance we should add is that strictly speaking, Monte-Carlo integration should resample the integrand at every timestep, but the network of neurons samples the neurons once, and propagates these points through time.
 
@@
 
-This method describes the trapezoidal method of integration.
-We state the result that the convergence rate is $\bO(N^{-2/p})$.
-For $p \geq 4$, MC shows better scaling behavior (but prefactors of error can change which method is better).
+Another method to estimate an integral numerically is by regularly sampling the integrand on a fixed grid.
+We simply state the result that the convergence rate of the estimation is $\bO(N^{-2/p})$.
+Interestingly, the convergence rate depends on the dimension $p$ of the integral.
+For $p \geq 4$, Monte-Carlo integration shows better scaling behaviour (but depending on the use case, prefactors of the error can change which method is better).
 
-In our case, since the distribution $\rho$ can be factorized, we can decouple the $p$-dimensional integral into the product of $p$ one-dimensional integrals.
+In our case, since the distribution $\rho(z_1, \cdots, z_p)$ can be factorized into the product of $p$ independent Gaussian distributions, we can decouple the $p$-dimensional integral into the product of $p$ one-dimensional integrals.
 Then the error estimation of the integral scales as $\bO(N^{-2})$.
@@
 
+We now introduce some tools that will help us to understand the toy model.
+We stress that in general, other neural field models might not have this luxury, and that we only make use of these tools to help explain the dynamics of the neural fields.
+Computationally, the overlaps and projections introduced here are only computed from the results of the simulation (see [`overlap.py`](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_overlap.py) in the source code).
+
 ### Overlap variables
 
-networks of neurons :
+Overlap variables were introduced by @Gerstner_1992 in the context of the study of the Hopfield network, which can be seen as a low-rank network with discrete variables instead of our continuous $\vec{z}$.
+The overlap variables measure the correlation between the state of the network (the value of the potential $h(\vec z, t)$ at each point in space) and the pattern fixed point.
+
+For the network for neurons, we write the overlap with the pattern $\mu$ as
 
 $$
-m_\mu(t) = \frac 1N \sum_{i=1}^N \tilde \phi(\xi_{\mu,i}) \phi(h_i(t))
+\begin{aligned}
+m_\mu(t) &= \frac 1N \sum_{i=1}^N \tilde \phi(z_{\mu i}) \phi(h_i(t)) \\
+&\xrightarrow{N \to \infty} \int_{\Rp} \tilde \phi(z_\mu) \phi(h(\vec z, t)) \rho(\d \vec z)
+\end{aligned}
+$$ {#eq-overlap}
+
+We note that similarly to @Gerstner_1992, we can write the dynamics of the neural field in terms of the overlap variables:
+
+$$
+\partial_t h(\vec z, t) = -h(\vec z, t) + \sum_{\mu=1}^p z_\mu m_\mu(t).
+$$ {#eq-overlap-evolution}
+
+This formulation "hides" the integral inside the overlap variables, but one should not forget that the overlaps depend on the neural field at a given time.
+
+### Low-rank neural fields as $p$-dimensional closed systems {#sec-closed-system}
+
+@eq-overlap-evolution hints towards the possibility of writing the dynamics of the low-rank system as a closed system of $p$ variables.
+As explained in @sec-low-rank-network, the recurrent currents span a $p$-dimensional subspace, and the orthogonal component behaves independently.
+This motivates the formulation of a $p$-dimensional closed system.
+
+We cite @veltz2009localglobal, section 4.3 for the reduction of general low-rank networks to $p$-dimensional closed systems, and give the equations applied to our toy model.
+Defining the projections $\kappa_\mu(t)$ of the neural field onto the patterns
+
+$$
+\kappa_\mu(t) = \int_{\Rp} y_\mu h(\vec y, t) \rho(\d \vec y),
+$$ {#eq-kappa}
+
+we can decompose the neural field onto the basis of patterns (mathematically, the patterns form an orthonormal basis of functions):
+
+$$
+h(\vec z, t) = h^\perp(\vec z, t) + \sum_{\mu=1}^p \kappa_\mu(t) z_\mu
 $$
 
-neural field :
+The orthonormal component $h^\perp(\vec z, t)$ is independent of the rest of the system and decays exponentially.
+Then, the equations of evolution for the projections are given by the following closed system:
 
 $$
-m_\mu(t) = \int_{\mathbb R^p} \tilde \phi(y_\mu) \phi(h(t,\vec y)) \rho(\mathrm d \vec y)
-$$
-
-note that :
-
-$$
-\partial_t h(t, \vec z) = -h(t, \vec z) + \sum_{\mu=1}^p z_\mu m_\mu(t)
-$$
-
-[TODO : i'm not even sure that introducing the overlaps in necessary.
-we might just include them for completeness, and because they are relevant in papers on low-rank stuff.
-in the rest, just use the kappa for the latent space]
-
-[TODO : do not expand mathematically, say this is well-known and given for context. the fixed point and stability study is new though]
-
-### Low-rank neural fields as $p$-dimensional closed systems
-
-cite @veltz2009localglobal for this.
-
-on explicite les equations du paper dans notre cas
-
-introduce this a just the projection of the field on the fixed points
-
-introduce kappa, and in our case the low-rank model the kappa
-
-we define
-
-$$
-\kappa_\mu(t) = \avg{f_\mu, h(t, \cdot)} = \int_{\Rp} y_\mu h(0, \vec y) \rho(\mathrm d \vec y)
-$$
-
-with these definitions, we have in our case
-$$
 \begin{aligned}
-\dot \kappa_\mu(t) &= -\kappa_\mu(t) + \int_{\Rp} \tilde\phi(y_\mu) \phi(h(t, \vec y)) \rho(\mathrm d \vec y) = -\kappa_\mu(t) + m_\mu(t) \\
-\kappa_\mu(0) &= \int_{\mathbb{R}^p} y_\mu h(0, \vec y) \rho(\mathrm d \vec y)
+\dot \kappa_\mu(t) &= -\kappa_\mu(t) + \int_{\Rp} \tilde\phi(y_\mu) \phi(h(\vec y, t)) \rho(\mathrm d \vec y) = -\kappa_\mu(t) + m_\mu(t) \\
+\kappa_\mu(0) &= \int_{\mathbb{R}^p} y_\mu h(\vec y, 0) \rho(\mathrm d \vec y),
 \end{aligned}
@@
 $$
-h^\perp(t, \vec z) = h^\perp(0, \vec z) \mathrm e^{-t},
-\quad
-h^\perp(0, \vec z) = h(0, \vec z) - \sum_{\mu=1}^p \kappa_\mu(0) z_\mu
+\begin{aligned}
+h^\perp(\vec z, t) &= h^\perp(\vec z, 0) \mathrm e^{-t}\\
+h^\perp(\vec z, 0) &= h(\vec z, 0) - \sum_{\mu=1}^p \kappa_\mu(0) z_\mu.
+\end{aligned}
 $$
 
-todo : put accent on the fact the RHS can be computed using all the kappas, kappa -> h -> integral = m -> RHS
+The set of $\kappa_\mu(t)$ defines a trajectory in a $p$-dimensional *latent space*.
+We note that similarly to @eq-overlap-evolution, the overlaps intervene in the equations of evolution, and carry the information on the neural field.
 
-conceptually : do not forget that the kappa is only relevant in this particular toy model, in general we do not have these kappa, and the study of low-dim fields is interesting.
-If we always had the kappa, then of course we can only look at the kappa, since much simpler (a 1D manifold (line) in a Rp space, instead of a Rp volume in a Rp space).
-
-warning : we need absolutely to distinguish the latent and the embedding space
-
-<!-- Relation between $p$-dimensional closed system and neural field equation in $\mathbb R^p$
-$$
-\begin{aligned}
-\partial_t h(t, \vec z) &= \partial_t h^\perp(t, \vec z) + \sum_{\mu=1}^p z_\mu \dot \kappa_\mu(t) \\
-&= -h^\perp(t, \vec z) - \sum_{\mu=1}^p z_\mu \kappa_\mu(t) + \sum_{\mu=1}^p z_\mu m_\mu(t) \\
-&= -h(t, \vec z) + \sum_{\mu=1}^p z_\mu m_\mu(t)
-\end{aligned}
-$$ -->
-
-overlaps as an approximation of the “dynamics” (latent space), in the case $\phi=\mathrm{linear}$.
-since in our case $\phi$ is a sigmoid, it can be considered linear in the range $h \sim \pm 3$.
-we therefore use the computed overlaps as to show the latent space
-
-we use the latent space to assert dynamics are the same
-
-<!-- [TODO FIG : draw a flow field of in the latent space. -> actually, might be difficult when there is delay] -->
+Additionally, it can be seen from the expression of the overlaps in @eq-overlap, that in the case of a linear activation function $\phi(h) = c_0 + c_1 h$, then $m_\mu(t) = \kappa_\mu(t)$.
 
@@
 
-### Intuition behind the cycling
+In @sec-networks-of-neurons-to-neural-fields we introduced a simple toy model a neural field with a natural embedding on the $p$-dimensional Gaussian ball.
+Analytical derivations accompanied by numerical simulations showed the dynamics of this model can be summarized by the convergence to pattern fixed points.
+In this section, we modify the toy model minimally, such that we can observe more interesting behaviour in the form of cycling.
 
-why ? dynamics are more interesting
+We modify @eq-nf-gaussian-p in two ways:
 
-we introduce $\delta$ the delay, and rolling $\mu+1$, with convention $p+1=1$
+1. We define a time delay $\delta$ with which a neuron will "wait" before reacting to a change in its potential.
+2.
+We define "rolling" as the response of a neuron to shift towards the "next" pattern given its state matching a current pattern.
+This is done by adding a shift $\mu+1$ in the connectivity kernel, with the convention $p+1=1$.
+
+The resulting cycling neural field is written as such:
 
 $$
-\partial_t h(t, \vec z) = -h(t, \vec z) + \int_{\mathbb{R}^p} \sum_{\mu=1}^p \tilde \phi (y_\mu) z_{\mu+1} \phi(h(t - \delta, \vec y)) \rho(\mathrm d \vec y)
+\partial_t h(\vec z, t) = -h(\vec z, t) + \sum_{\mu=1}^p \int_{\mathbb{R}^p} z_{\mu+1} \tilde \phi (y_\mu) \phi(h(\vec y, t - \delta)) \mathcal{N}^p(\mathrm d \vec y)
+$$ {#eq-nf-cycling-p}
+
+The geometric intuition behind this formulation is that now the recurrent currents are "rotated by half a turn" (around the axis normal to the plane $(z_\mu, z_{\mu+1})$) in the embedding.
+When the network is at a pattern fixed point $\mu$, the recurrent drive will then push it towards the pattern $\mu+1$.
+@fig-cycling-irec shows the difference in the behaviour of the original and the cycling neural field.
+
+![(TODO animation in 3D, rename non-cycling to original) Intuition for the cycling behavior by considering the recurrent currents](figures/cycling_recurrent.png){#fig-cycling-irec}
+
+We note however that @eq-nf-cycling-p no longer is a partial differential equation, but rather a delayed differential equation, which are in general much more complicated to solve.
+A simplifying assumption that we make is that we extend the initial condition back in time, such that $h(\vec z, t < 0) = h(\vec z, 0)$ is a constant function.
+Numerically, this also means that we have to store a history of the neural fields instead of only the current one.
+Since the `solve_ivp` method dynamically adapts the timestep of integration, we use linear interpolation between two known points in the stored neural field states to estimate $h(\vec z, t-\delta)$ (see [`lagging.py`](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_lagging.py) and [its use](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_rnn.py#L334)).
+
+With these considerations, we simulate a $p=2$ cycling neural field with $\delta=6$ and initial condition $h(z_1, z_2, t < 0) = z_1$ in @fig-cycling-anim.
+The latent trajectory is estimated from the simulation results, and animated simultaneously to the neural field.
+We clearly see the oscillations between the patterns $\mu=1$ and $\mu=2$.
+
+![A cycling neural field with $\delta=6$ and initial condition $h(z_1, z_2, t < 0) = z_1$.
+The latent trajectory is also computed, and demonstrates the oscillatory behaviour.](figures/embedding_d=2_h0=0_cycling_delta=6.mp4){#fig-cycling-anim}
+
+Relating to @sec-closed-system, the dynamics of the projection now read:
+
+$$
+\dot \kappa_\mu(t) = -\kappa_\mu(t) + m_{\mu+1}(t - \delta).
 $$
 
-![Intuition for the cycling behavior by plotting the recurrent currents](figures/cycling_recurrent.png)
+This system of equations can be solved by steps of $\delta$, iteratively plugging in the previous solution to solve each cycle.
 
-note this is now a DDE instead of a PDE
-
-![](figures/embedding_d=2_h0=0_cycling_delta=6.mp4)
-
-### Cycling behavior of the latent variables
-
-1. delayed neural field
-2. rolled neural field
-
-[TODO : is this useful ? I derived this because it shows we need both delay and rolling, but might be too much]
-
-we can show that in this case
-
-$$
-\dot \kappa_\mu(t) = -\kappa_\mu(t) + m_{\mu+1}(t - \delta)
-$$
-
-then in the case that $\phi=linear \implies \kappa_\mu = m_\mu$, and initial conditions $m_\mu(t<0) = m_\mu(0)$ (constant inititial condition), $m_1(0) = 1$ and $m_2(0) = 0$, we can solve the first "cycle" ($0 < t < \delta$) analytically :
+<!-- When the neural field starts in the pattern $\nu$, that is $h(\vec z, 0) = z_\nu$, this translates to $\kappa_\nu(t \leq \delta) = 1$, and all other $\kappa_\mu(t \leq \delta) = 0,\,\mu\neq\nu$.
+Additionally, let us suppose that $\phi$ is linear, so that $m_\mu(t) = \kappa_\mu(t)$, and let us take $p=2$.
+Then we can solve the equation for the "first cycle" $0 < t \leq \delta$:
 
@@
 \kappa_1(t) &= \mathrm{e}^{-t} \\
-\kappa_2(t) &= 1 - \mathrm{e}^{-t} \\
+\kappa_2(t) &= 1 - \mathrm{e}^{-t}.
 \end{aligned}
@@
 
-then one can repeat this for the next cycles. basically each cycle is applying an exponential convolution filter [TODO : trim this, don't go into detail]
-
-<!-- [TODO MATH : the derivation of this, we have an exponential convolution filter] -->
-
-![simulate cycling in p=2](figures/kappa_cycling_p=2.tmp.png)
+With this, we see that the projections "alternate" back and forth with period $\delta$. -->
 
@@
 
-<!-- ### Using the CDF -->
+In preparation for mapping of $[0,1]^2$ to $[0,1]$, we first map the 2-dimensional neural field equation in $\R^2$ to $[0,1]^2$, and, in general $\Rp$ to $[0,1]^p$.
+There are many ways to do this by using functions $\R \mapsto [0,1]$, but a practical choice is the Gaussian CDF (Cumulative Density Function), because it has the benefit of absorbing the density in the kernel, in the sense that the integral $[0,1]^p$ becomes weighted by the uniform distribution.
+The one-dimensional Gaussian CDF is defined as
 
-later we will map [0,1]² -> [0,1], so we first want to map R² -> [0,1]².
+$$
+\CDF(z) = \int_{-\infty}^z \mathcal{N}(\d y)
+$$
 
-there are many ways to do this using functions R -> [0,1], but a practical choice is the inverse CDF, which has the added benefit of absorbing the density in the kernel, in the sense that the integral over [0,1]² becomes "weighted" by a uniform distribution (lesbesgue measure $\lambda$)
+and maps $\R$ to $[0,1]$,[^R-to-01] as pictured in @fig-gaussian-cdf.
 
-Defining the change of variables $u_\mu=\mathrm{CDF}(y_\mu), v_\mu=\mathrm{CDF}(z_\mu)$, the neural field becomes
+![(TODO) CDF of the Gaussian distribution](figures/1920px-Normal_Distribution_CDF.svg.png){width=400px #fig-gaussian-cdf}
+
+[^R-to-01]: Strictly speaking the image of $\R$ is in $(0,1)$, but for simplicity we consider $\R \cup \{-\infty, +\infty\}$ which maps to $[0,1]$.
+This is not fundamentally important, because numerically the infinities are never touched, and analytically the recurrent currents given by the integration vanish at infinity due to the Gaussian weighting.
+
+Defining the change of variables $v_\mu=\CDF(z_\mu), u_\mu=\CDF(y_\mu)$, we can define the uniform neural field and uniform connectivity kernel as
 
@@
 \begin{aligned}
-& \partial_t h_U(t, \vec v) = -h_U(t, \vec v) + \int_{[0,1]^p} w_U(\vec v, \vec u) \phi(h_U(t, \vec u)) \mathrm d \vec u \\
-& \text{where } w_U(\vec v, \vec u) = w(\mathrm{CDF}^{-1}(\vec v), \mathrm{CDF}^{-1}(\vec u)), \, h_U(t, \vec v) = h(t, \mathrm{CDF}^{-1}(\vec v))
+h_U(\vec v, t) &= h(\CDF^{-1}(\vec v), t) = h(\CDF^{-1}(v_1), \cdots, \CDF^{-1}(v_p)) \\
+w_U(\vec v, \vec u) &= w(\CDF^{-1}(\vec v), \CDF^{-1}(\vec u)).
 \end{aligned}
@@
 
-![normal distribution CDF](figures/1920px-Normal_Distribution_CDF.svg.png)
-
-We now have a neural field equation on a unit (hyper)cube, and uniform sampling.
-
-the problem of sampling all of $\R^2$ is now "hidden away" in the CDF, since it goes to $\pm \infty$ at 0 and 1.
-
-note : here we use the fact that the components are independant. therefore we can define the total as the product, i.e.
+After the change of variables, the neural field in $[0,1]^p$ is equivalent to the neural field in $\Rp$, and its equation reads
 
 $$
-\mathrm{CDF}(z_1, z_2) = \int_{-\infty}^{z_1} \rho_1(d y_1) \int_{-\infty}^{z_2} \rho_2(\d y_2) = v_1 v_2
+\partial_t h_U(t, \vec v) = -h_U(t, \vec v) + \int_{[0,1]^p} w_U(\vec v, \vec u) \phi(h_U(t, \vec u)) \mathrm d \vec u.
 $$
 
-we would not be able to do this is the coordinates were coupled, then we could define still define the CDF, but we cannot factorize it into 2 different components
+Defining the neural field in $[0,1]^p$ also addresses the problem that we tried to hide in @sec-gridmethod relating to the difficulties of defining a grid on $\Rp$.
+With the integral on $[0,1]^p$, we can simply use a uniform grid, which maps back to samples in $\Rp$ by using the inverse CDF, as shown in @fig-sampling-R2-012.
 
-![samples in $\Rp$ and corresponding samples in $[0,1]$](figures/sampling_R2_01.png)
+![(TODO) A grid in $[0,1]^2$ and the corresponding samples in $\R^2$.
+The numeric density of points in $\R^2$ approaches the normal distribution as the grid becomes finer.](figures/sampling_R2_01.png){#fig-sampling-R2-012}
 
-give a picture of the "isolines" of the 2D gaussian CDF
-
-furthermore numerically, mapping R² -> [0,1]² adresses the problem of "putting a finite grid on the infinite R²".
-(previously we would have had to make the approximation of compact support, then sample the PDF on the grid and numerically renormalize.)
-
-<!-- ### Consequences on simulations
-
-#### Direct sampling
-
-we simulate the network of neurons
-
-#### Grid over $[0,1]^2$
-
-We define $\vec z(\alpha) = S(\alpha)$ the 2D point corresponding to the mapping $\alpha$.
-(details : we need a bounding box to map back).
-Defining a matrix $Z_{\mu,\alpha} = \vec z(\alpha)_\mu$, we can write down a numerical PDF $\tilde \rho(Z_{1,\alpha},\cdots,Z_{p,\alpha}) = \tilde \rho(Z_{:,\alpha}) = \frac{\rho(Z_{:,\alpha})}{\sum_\beta \rho(Z_{:,\beta})}$ and the following patterns to simulate the embedded $[0,1]$ neural field as a low-rank RNN.
+We make the final remark that this method of mapping $\Rp$ to $[0,1]^p$ does not work for any neural field.
+In our case, we use the fact that the distribution factorizes into $\rho(z_1, \cdots, z_p) = \rho_1(z_1) \cdots \rho_p(z_p)$, such that the components of $\vec{v}$ are sampled independently.
+We can then define the "componentwise CDF" that relates to the total CDF via
 
 $$
-\tilde F_{\mu,\alpha} = Z_{\mu,\alpha}, \quad \tilde G_{\mu,\alpha} = \tilde\phi(Z_{\mu,\alpha}), \quad \tilde J_{\alpha,\beta}=\tilde \rho(Z_{:,\beta}) \sum_{\mu=1}^p \tilde F_{\mu,\alpha} \tilde G_{\mu,\beta}
+\begin{aligned}
+\CDF(z_1, \cdots, z_p) &= \int_{-\infty}^{z_1} \cdots \int_{-\infty}^{z_p} \rho(\d y_1, \cdots, \d y_p) \\
+&= \int_{-\infty}^{z_1} \rho_1(\d y_1) \cdots \int_{-\infty}^{z_p} \rho_p(\d y_p) \\
+&= \CDF(z_1) \cdots \CDF(z_p) \\
+&= v_1 \cdots v_p = \prod_{\mu=1}^p v_\mu.
+\end{aligned}
 $$
-
-Doing this is equivalent, but more memory-efficient to the formulation of the binned connectivity matrix.
-
-![](figures/sampling_R2_01_kernel.png) -->

### 2023-06-14 manuscript/chapters/3_mappings.qmd
meow meow i finish section 2 !!!!

@@
 For this, we define a measurable and bijective mapping $S : [0,1]^2 \rightarrow [0, 1]$.
-Furthermore, we let $\lambda : [0,1]^2 \rightarrow [0, \infty]$ be a measure on the unit square (for instance, the Lebesgue measure).
+Furthermore, we let $\lambda : [0,1]^2 \rightarrow [0, \infty]$ be a measure on the unit square (for instance, the Lebesgue measure, which plays the role of a uniform distribution).
 Taking the composition $\lambda \circ S^{-1}$ then defines a measure on $[0,1]$, which keeps track of how 2D surfaces are transformed to 1D lengths through the mapping $S$.

### 2023-06-14 manuscript/chapters/3_mappings.qmd
arruuwueughgeheg

@@
 $$
-\partial_t \tilde h(t, \alpha) = -\tilde h(t, \alpha) + \int_{[0,1]} \tilde w(\alpha, \beta) \phi(\tilde h(t, \beta)) \lambda \circ S^{-1}(\d \beta)
+\partial_t \tilde h(\alpha, t) = -\tilde h(\alpha, t) + \int_{[0,1]} \tilde w(\alpha, \beta) \phi(\tilde h(\beta, t)) \left[\lambda \circ S^{-1}\right](\d \beta)
 $$ {#eq-nf-1d}
@@
 
+TODO small intro
+
 ### Application of $S$ : from $[0,1]^2$ to $[0,1]$
 
-![](figures/coarsegraining.png)
+TODO : we run simulations and compare latent trajectories.
+first show only column and z-mapping.
+add quick intro to diagonal, col/row, reclocal, mention cannot be expressed by bit permutation.
+diagonal, col/row do not work, despite the non-zero field and recurrent currents.
+reclocal does work.
+intuition is that the recursive approach (zmapping, reclocal) seems to give locality
+
+![(TODO animation) Simulations of downsampling with different mappings from the $p=2$ embedding.](figures/coarsegraining.png){width=500px}
 
@@
 
+TODO : take cycling p=3
+
 #### Continuity on one axis and fractal on the other: from $[0,1]^3$ to $[0,1]^2$
+
+TODO : we do z-mapping on each slice $(v_2, v_3)$ of $v_1$, and downsample.
+we end up with a 2d neural field, direction $v_1$ is continuous, direction $\alpha_{23}$ is fractal.
+now the field is not continuous anymore, it is just sufficiently regular to be L1.
+this shows "local mappings conserve regularity"
+
+![(TODO 3D -> 2D only) Equivalence of dynamics, from 3D to 2D.](figures/embedding_3d_2d_1d.mp4)
+
+![(TODO 3D -> 2D only) Effective sampling points after averaging.](figures/mapping_3d_2d_1d.png){width=400px}
 
@@
 
-![Effective sampling points after averaging](figures/mapping_3d_2d_1d.png)
+TODO : we can re-apply the z-mapping in $(v_1, \alpha_{23})$, we get a fractal 1d embedding. completely fractal, but the dynamics still work, because the mapping conserves regularity.
 
-![Equivalence of dynamics, from 3D to 2D to 1D](figures/embedding_3d_2d_1d.mp4)
+![Equivalence of dynamics, from 3D to 2D to 1D.](figures/embedding_3d_2d_1d.mp4)
+
+![(TODO) Effective sampling points after averaging.](figures/mapping_3d_2d_1d.png){width=400px}

### 2023-06-14 manuscript/chapters/4_locality.qmd
arruuwueughgeheg

--- 2023-06-04 manuscript/chapters/4_locality.qmd
@@
 
-# Locality and numerical convergence
+# Locality and numerical convergence {#sec-locality}
 
-## Analytical equivalence of the dynamics
+The previous chapter gave a detailed description of the mappings $S : [0,1]^2 \mapsto [0,1]$, as well as a few numerical considerations.
+We found that mappings that yield equivalent dynamics crucially seem to have in common the property of *locality*, that is that they conserve the regularity between the domain and the image, even if the image seems to have a fractal structure.
+Intuitively, we saw that the locality seems to emerge from the mapping having overwhelmingly small variations.
 
-composition of functions
+In this chapter, we will further explore these results, and attempt to formulate a measure of this "locality".
+We start by showing that bijectivity and measurability is a sufficient condition for equivalent dynamics.
+Then, we apply the locality measure to the special cases of the Column and Z-mappings, and show how this notion can discriminate between the two.
+Finally, we show that the formulated measure is sufficient to ensure the numerical estimation of the integral on the fractal structure converges to the analytical integral; therefore, such embeddings are feasible to find and simulate numerically.
 
-bijectivity (actually injection in one sense)
+## Analytical equivalence of the dynamics when $S$ is measurable and bijective
 
-TODO : write the proof of this
+In the previous chapter, we have introduced Cantor and Peano functions, and argued that if have to chose between the two, that the non-injective Peano functions are preferable, because by Netto's theorem, they are the only ones that can be space-filling and reach all the square populations.
+Despite this, let us make the simplifying assumption that $S$ is a bijective function.
+However, as to not break Netto's theorem, we do not make any assumptions regarding its continuity.
+
+We furthermore assume that the mapping $S$ is measurable.
+This is not a strong assumption, since the limit of pointwise convergent measurable functions is also measurable.
+In our case, we have written the finite-$n$ expansions $S^n$, which by being sums of piecewise constant functions, are also measurable.
+Measurability of $S$ allows us to write $\lambda \circ S^{-1}$, since by definition the preimage of a measurable function is measurable.
+
+With these two assumptions, let us prove that by defining the initial condition and connectivity
+$$
+\begin{aligned}
+\tilde h(\alpha, t=0) &= h_U(S^{-1}(\alpha), t=0)\\
+\tilde w(\alpha, \beta) &= w_U(S^{-1}(\alpha), S^{-1}(\beta)),
+\end{aligned}
+$$
+
+then the solution $h_U(\vec v, t)$ of the neural field equation in $[0,1]^2$ uniquely defines the solution of the neural field equation in $[0,1]$ by writing
+
+$$
+\tilde h(\alpha, t) = h_U(\cdot, t) \circ S^{-1}(\alpha).
+$$ {#eq-analytical-equivalence-tildeh}
+
+We start this the one-dimensional neural field, then substitute @eq-analytical-equivalence-tildeh and $\vec{v} = S^{-1}(\alpha),\, \vec{u} = S^{-1}(\beta)$.
+
+$$
+\begin{aligned}
+\partial_t \tilde h(\alpha, t) &= -\tilde h(\alpha, t) + \int_{[0,1]} \tilde w(\alpha, \beta) \phi(\tilde h(\beta, t)) \left[\lambda \circ S^{-1}\right](\d \beta) \\
+\partial_t h_U(S^{-1}(\alpha), t) &= 
+\begin{aligned}[t]
+	&-h_U(S^{-1}(\alpha), t)\\&+ \int_{[0,1]} w(S^{-1}(\alpha), S^{-1}(\beta)) \phi(h_U(S^{-1}(\beta), t)) \left[\lambda \circ S^{-1}\right](\d \beta)
+\end{aligned}\\
+&= \begin{aligned}[t]
+	&-h_U(\vec v, t)\\&+ \int_{[0,1]} w(\vec v, S^{-1}(\beta)) \phi(h_U(S^{-1}(\beta), t)) \left[\lambda \circ S^{-1}\right](\d \beta)
+\end{aligned}\\
+&= -h_U(\vec v, t) + \int_{[0,1]^2} w(\vec v, \vec u) \phi(h_U(\vec u, t)) \lambda(\d \vec u)
+\end{aligned}
+$$
+
+In the last line, we have used the bijectivity of $S$. The Lebesgue measure $\lambda : [0,1]^2 \mapsto \R^+$ simply plays the role of a uniform distribution on $[0,1]^2$, and we simply write
+
+$$
+\partial_t h_U(\vec v, t) = -h_U(\vec v, t) + \int_{[0,1]^2} w(\vec v, \vec u) \phi(h_U(\vec u, t)) \d \vec u.
+$$
+
+Therefore, if $h_U(\vec v, t)$ solves the neural field equation in $[0,1]^2$, then $\tilde h(t) = h_U(t) \circ S^{-1}$ solves the equivalent neural field equation in $[0,1]$.
+The same reasoning applies in reverse: if $\tilde h(t, \alpha)$ solves the neural field equation in $[0,1]$, then $h_U(t) = \tilde h \circ S$ solves the neural field in $[0,1]^2$.
+
+[TODO : uniqueness of the solution ?]
 
@@
 
-### Intuition behind the notion of locality
-
-![](figures/distance_2d_1d.png)
-
-localité :
-1. proche in 2D -> proche in 1D (OK projection (obv., x est proche), OK zorder, PAS random)
-2. proche in 1D -> proche in 2D (PAS projection, PAS random, OK zorder)
-
-[TODO EXPLAIN : explain this better. basically the thing is we AVERAGE in 1D, so this better be meaningful]
-
-the notion of infinite number of discontinuities, but the size counts too
-
-plot distance 2D vs distance 1D
-* "close to the line y=sqrt2 x" (but fractals dont get arbitrarily close)
-* "size of the discontinuities"
-* L1 distance
-
-"resilient to some discontinuity" the notion of infinite number of discontinuities, but the size counts too
-
-![histogram of discontinuities](figures/image_20230517_2.png)
-
 ### The total variation attempt
 
-Let $S^{-1} : \mathbb R \rightarrow \mathbb R^2$, let $|| \cdot ||$ be a norm in $\mathbb R^2$ (any norm is OK, since all equivalent (i.e.
-equal to a constant)).
-Here we use the $\ell^1$ norm (absolute value).
+The previous chapter gave the intuition that local mappings map populations close in 1D to populations close in 2D.
+In order to verify this, we might try to compare the distance of all segment populations to the population at $\alpha=0$, to the distance of all square populations to the population at $S^{-1}(\alpha=0)$.
+This is done in @fig-distance-2d-1d, where the distance is taken to be the $\ell^2$ norm.
+If the distances matched perfectly, then we would see a line with slope $\sqrt 2$ in the graph.
+
+![(TODO random, column and zcurve only) Distance between the first population and all other populations, in 1D and 2D.
+The dotted line has a slope $\sqrt 2$.](figures/distance_2d_1d.png){#fig-distance-2d-1d}
+
+For control, we also demonstrate that the random mapping has no locality: there is no correlation between the positions in 1D and the positions in 2D.
+Both Column and Z-mappings show a clear trend, the distances in 2D are correlated to the distances in 1D.
+The notable difference is that distances for the Column mapping seem to "oscillate" wildly, while the Z-mapping seems much more regular and matches the straight line more closely.
+
+This analysis motivates the following hypothesis: the property that sets the mappings apart, is that of total variation.
+The idea is to "project and sum up" all of the variations defined by the curves as they trace a path in the $[0,1]^2$ embedding.
+As $n\to\infty$, we expect to see that the total variation of the Z-mapping grows much more slowly than that of the Column mapping.
+
+Denoting $\alpha_i,\,i \in \{1, \cdots, 4^n\}$ the positions of the grid in the $[0,1]$ embedding, we define the total variation as the projection of all the corresponding 2D segments:[^tv]
 
@@
 \begin{aligned}
-\mathrm{TV}(S^{-1}) &= \sup_{P \in \mathcal P} \sum_{i=1}^{n_P} || S^{-1}(z_{i}) - S^{-1}(z_{i-1}) ||_1 \\
-&= \sup_{P \in \mathcal P} \sum_{i=1}^{n_P} | S^{-1}_1(z_i) - S^{-1}_1(z_{i-1})| + | f_2(z_i) - f_2(z_{i-1}) |
+\mathrm{TV}(S^{-1}) &= \sum_{i=2}^{4^n} \norm{S^{-1}(\alpha_i) - S^{-1}(\alpha_{i-1})}_1 \\
+&= \sum_{i=2}^{4^n} |S_1^{-1}(\alpha_i) - S_1^{-1}(\alpha_{i-1})| + |S_2^{-1}(\alpha_i) - S_2^{-1}(\alpha_{i-1})|.
 \end{aligned}
+$$ {#eq-tv}
+
+[^tv]:
+  This definition is slightly different from other definitions of total variation, which involve a supremum over all partitions $P = \{0 \leq \alpha_1 < \cdots < \alpha_{N_p} \leq 1\}$.
+  <!-- The total variation is then defined as
   $$
+  \mathrm{TV}(S^{-1}) = \sup_{P} \sum_{i=1}^{N_P} \norm{S^{-1}(\alpha_i) - S^{-1}(\alpha_{i-1})}_1.
+  $$ -->
+  We don't use the more complex definition, because our version of total variation is just used for illustration purposes.
+Additionally, we make use of the $\ell^1$ norm because it lends itself to the nice geometric intuition of "projecting the jumps" on the axes.
+This choice of the norm is irrelevant however, because analytically all norms are equivalent up to a constant.
 
-this basically just projects the discontinuities on x and y.
+It is easy to see that $2\times 4^n$ is an upper bound of $\mathrm{TV}(S^{-1})$, since the image of $S^{-1}$ is bounded in $[0,1]^2$.
+In @fig-tv, we compute the total variation for different $n$ corresponding to different mappings.
 
-in a way since you need to go to each point only once, so there is only so much you can do inside the $[0,1]^2$. the tiny jumps still add up, and all mappings seem to grow $TV$ in the same way
+![(TODO $2\times 4^n$) Total variation for different mappings, as a function of $n$](figures/tv.png){#fig-tv}
 
-![](figures/image_20230517.png)
+Dissapointingly, all mappings seem to exhibit the same growth of total variation.
+It seems like our notion of total variation is not adequate to help discriminate between the mappings.
+We need to rethink our intuition.
 
--> what we really are expression, is the fact that doing the averaging in 1D is not a big problem when the mapping is good.
+<!-- In some sence, having the constraint that the mapping traces a path in the bounded space $[0,1]^2$ that can go through each point of the grid only once leaves little flexibility.
+For fixed $n$, there are only finitely many "large jumps" the mapping can make, and the large number of "small jumps" still add up, so that the total variation still grows at approximatively the same rate.
+-->
 
-### Mathematical formulation of average binned variation
+### Locality as vanishing average binned variation
 
-variation inside the bins, weak version
+When we do the downsampling of the mappings, we expressed that the locality helps by bounding the variation inside of each averaged bin.
+Therefore, what we would like to express, is that on average, the variation inside of each bin before downsampling is small.
+@fig-sketch-binned-variation gives an intuition of this.
+Even though the $[0,1]$ embedding might be discontinuous in places, on average nearby neurons should exhibit the same activity levels.
+
+![(TODO the variation for the lower one shoud be lesser) Intuition behind the average binned variation.
+To simplify the visualization, activity levels are used in lieu of the 2D positions.](figures/photo_2023-06-15_00-44-43.jpg){width=400px #fig-sketch-binned-variation}
+
+Given $n$, let us define the average binned variation as the maximum difference of the 2D positions associated to segment populations in the same bin of size $2^{-n}$:
 
 $$
-\begin{aligned}
-V_n(S^{-1}) &= \frac{1}{2^n}\sum_{i=1}^{2^n} \sup_{\delta,\delta' \in [0, \tfrac{1}{2^n}]^2} |S_1(\tfrac{i-1}{2^n}+\delta) - S_1(\tfrac{i-1}{2^n}+\delta')| + |S_2(\tfrac{i-1}{2^n}+\delta) - S_2(\tfrac{i-1}{2^n}+\delta')| \\
-&= \frac{1}{2^n}\sum_{i=1}^{2^n} \sup_{\alpha, \alpha^\prime \in [\tfrac{i-1}{2^n}, \tfrac{i}{2^n}]^2} \norm{S^{-1}(\alpha^\prime) - S^{-1}(\alpha)}_1
-\end{aligned}
-$$
+V_n(S^{-1}) = \frac{1}{2^n}\sum_{i=1}^{2^n} \sup_{\alpha, \alpha^\prime \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]^2} \norm{S^{-1}(\alpha^\prime) - S^{-1}(\alpha)}_1
+$$ {#eq-Vn}
 
+<!--
 variation inside the bins, strong version.
@@
 $$
+-->
 
-[TODO FIG : show the range of each bin, how each bin contains $2^n$ smaller bins. plot the bins in the 2D scatter, so we can see what we are doing]
+TODO : I have removed the definition involving the sup over all permutations, because we don't end up using it. Still mention for completeness ?
 
-[TODO MATH : the exact norm is not important. all norms are equivalent]
+TODO NOTE : in the definition of $V$, we use the converged version of $S$, instead of a finite truncation $S^n$.
+we need to convince people this is not a big mistake, since $S^n \to S$, and we can take both limits (binning and $S^n$) "at the same time".
+This might actually be adressed by a similar todo in @sec-mappings-binary
 
-[TODO NOTE : in the definition of $V$, we use the converged version of $S$, instead of a finite truncation $S^n$.
-we need to convince people this is not a big mistake, since $S^n \to S$, and we can take both limits (binning and $S^n$) "at the same time"]
+Before we move to analytical derivations applying the average variation to mappings, we can first check numerically if it is gives the expected results; that is $V_n$ vanishes a $n\to\infty$ for local mappings, whereas $V_n$ is lower bounded by a constant for non-local mappings.
+@fig-vnum shows this is indeed the case, rejoice !
 
-### Average variation of mappings
+![(TODO change title, change name of mappings) (Numerator of the) average binned variation as a function of $n$.
+The local mappings have a growth slower than the upper bound $2\times 2^n$.](figures/v_num.png){width=500px #fig-vnum}
 
-Numerical demonstration
-
-![](figures/v_num.png)
+### Computations of average variation
 
@@
 
-Putting everything together, we get an upper bound for the average binned variation.
+Putting everything together, we get a vanishing upper bound for the average binned variation.
 
@@
 C^{-1}_1(\alpha) &= 0.i_{n-1} i_{n-2} \cdots i_1 i_0 b_n b_{n+1} \cdots b_m \\
-C^{-1}_2(\alpha) &= 0.b_{m+1} b_{m+2} \cdots b_{2m}
+C^{-1}_2(\alpha) &= 0.b_{m+1} b_{m+2} \cdots b_{2m}.
 \end{aligned}
@@
 |C^{-1}_1(\alpha) - C^{-1}_1(\alpha^\prime)| &= 2^{-n} \sum_{k=1}^{m-n+1} (b_{n+k-1} - b^\prime_{n+k-1}) 2^{-k} \\
-&\leq 2^{-n}
+&\leq 2^{-n}.
 \end{aligned}
@@
 However now problems arise when we compute the averge binned variation, since the second component is not well-behaved.
-In particular, since we take the $\sup_{\alpha,\alpha^\prime \in [\tfrac{i-1}{2^n},\tfrac{i}{2^n}]^2}$, we can pick the specific values such that $b_k=1$ and $b^\prime_k=0$ for all $k \in \{m+1, \cdots 2m\}$.
+In particular, since we take the supremum on $\alpha,\alpha^\prime \in \left[\tfrac{i-1}{2^n},\tfrac{i}{2^n}\right]^2$, we can pick the specific values such that $b_k=1$ and $b^\prime_k=0$ for all $k \in \{m+1, \cdots 2m\}$.
 
@@
 \begin{aligned}
-V_n(C^{-1}) &= \frac{1}{2^n}\sum_{i=1}^{2^n} \sup_{\alpha, \alpha^\prime \in [\tfrac{i-1}{2^n}, \tfrac{i}{2^n}]^2} |C^{-1}_1(\alpha^\prime) - C_1^{-1}(\alpha)| + |C^{-1}_2(\alpha^\prime) - C_2^{-1}(\alpha)| \\
-&\geq \frac{1}{2^n}\sum_{i=1}^{2^n} \sup_{\alpha, \alpha^\prime \in [\tfrac{i-1}{2^n}, \tfrac{i}{2^n}]^2} |C^{-1}_1(\alpha^\prime) - C_1^{-1}(\alpha)| + \sup_{\alpha, \alpha^\prime \in [\tfrac{i-1}{2^n}, \tfrac{i}{2^n}]^2} |C^{-1}_2(\alpha^\prime) - C_2^{-1}(\alpha)| \\
+V_n(C^{-1}) &= \frac{1}{2^n}\sum_{i=1}^{2^n} \sup_{\alpha, \alpha^\prime \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]^2} |C^{-1}_1(\alpha^\prime) - C_1^{-1}(\alpha)| + |C^{-1}_2(\alpha^\prime) - C_2^{-1}(\alpha)| \\
+&\geq
+\begin{aligned}[t]
+	\frac{1}{2^n}\sum_{i=1}^{2^n}&
+	\sup_{\alpha, \alpha^\prime \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]^2} |C^{-1}_1(\alpha^\prime) - C_1^{-1}(\alpha)|\\
+	&+ \sup_{\alpha, \alpha^\prime \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]^2} |C^{-1}_2(\alpha^\prime) - C_2^{-1}(\alpha)|
+\end{aligned}\\
 &\geq \frac 12 + \bO(2^{-n})
@@
 |\mathrm{NI}_n - \mathrm{AI}| &\leq L(\cdot) \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \lambda \circ S^{-1}(\d \beta) \right| \\
-&\leq L(\cdot) \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \sup_{\beta \in [\tfrac{i-1}{2^n}, \tfrac{i}{2^n}]} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \lambda \circ S^{-1}(\d \beta^\prime) \right| \\
-&= L(\cdot) \frac{1}{2^n} \sum_{i=1}^{2^n} \sup_{\beta \in [\tfrac{i-1}{2^n}, \tfrac{i}{2^n}]} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \\
-&\leq L(\cdot) \frac{1}{2^n} \sum_{i=1}^{2^n} \sup_{\beta, \beta^\prime \in [\tfrac{i-1}{2^n}, \tfrac{i}{2^n}]^2} \norm{S^{-1}(\beta^\prime) - S^{-1}(\beta)} \\
+&\leq L(\cdot) \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \sup_{\beta \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \lambda \circ S^{-1}(\d \beta^\prime) \right| \\
+&= L(\cdot) \frac{1}{2^n} \sum_{i=1}^{2^n} \sup_{\beta \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \\
+&\leq L(\cdot) \frac{1}{2^n} \sum_{i=1}^{2^n} \sup_{\beta, \beta^\prime \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]^2} \norm{S^{-1}(\beta^\prime) - S^{-1}(\beta)} \\
 &= L(\cdot) V_n \\

### 2023-06-14 notebooks/neurodyn/_mapping.py

@@ -12,7 +12,7 @@ __all__ = [
 	'RecursiveLocalMapping',
 	'ReshapeMapping', 'ColumnMapping',  # alias
 	'DiagonalMapping',
-	'SzudzikMapping',
+	'SzudzikMapping', 'ColRowMapping',  # alias
 	'RecursiveFarMapping', 'AntiZMapping',  # alias
 	'ZMapping',
 	'RandomMapping',
@@ -508,6 +508,7 @@ class SzudzikMapping(BinMapping):
 	def __str__(self) -> str:
 		return f'SzudzikMapping{{nxy={self.nxy}}}'
 
+ColRowMapping = SzudzikMapping
 
 class ZMapping(BinMapping):
 	"""Implements [Z-order curve](https://en.wikipedia.org/wiki/Z-order_curve)"""

### 2023-06-14 manuscript/chapters/2_neural_fields.qmd
messy compile to PDF

@@
 
+::: {.content-visible when-format="html"}
 ![(TODO remove overlap, change notation, color for activity) Simulation of the Gaussian network of neurons with $p=1$.
 Every neuron is represented by a dot, of which the height represents its activity level.](figures/embedding_d=1_h0=0_phi=sigmoid.mp4){width=400px #fig-p1}
+:::
+
+::: {.content-visible when-format="pdf"}
+![(TODO PDF remove overlap, change notation, color for activity) Simulation of the Gaussian network of neurons with $p=1$.
+Every neuron is represented by a dot, of which the height represents its activity level.](figures/lol7.jpg){width=400px #fig-p1}
+:::
 
@@
 
+::: {.content-visible when-format="html"}
 ![(TODO remove overlap, change notation, scatter only) Simulation of the Gaussian network of neurons with $p=2$.
 Every neuron is represented by a dot, of which the height represents its activity level.](figures/embedding_d=2_h0=0.mp4){width=500px #fig-p2}
+:::
+
+::: {.content-visible when-format="pdf"}
+![(TODO PDF remove overlap, change notation, scatter only) Simulation of the Gaussian network of neurons with $p=2$.
+Every neuron is represented by a dot, of which the height represents its activity level.](figures/lol7.jpg){width=500px #fig-p2}
+:::
 
@@
 
+::: {.content-visible when-format="html"}
 ![A cycling neural field with $\delta=6$ and initial condition $h(z_1, z_2, t < 0) = z_1$.
 The latent trajectory is also computed, and demonstrates the oscillatory behaviour.](figures/embedding_d=2_h0=0_cycling_delta=6.mp4){#fig-cycling-anim}
+:::
+
+::: {.content-visible when-format="pdf"}
+![(TODO PDF) A cycling neural field with $\delta=6$ and initial condition $h(z_1, z_2, t < 0) = z_1$.
+The latent trajectory is also computed, and demonstrates the oscillatory behaviour.](figures/lol7.jpg){#fig-cycling-anim}
+:::
 

### 2023-06-14 manuscript/chapters/3_mappings.qmd
messy compile to PDF

@@
 $$
-\begin{alignat}{2}
-S^n : &\{0, \tfrac{1}{2^n}, \cdots, 1-\tfrac{1}{2^n}\}^2 &&\mapsto \{0, \tfrac{1}{4^n}, \cdots, 1-\tfrac{1}{4^n}\} \\
-&(v^{(n)}_1,v^{(n)}_2)=(0.b^1_1 b^1_2 \cdots b^1_n, 0.b^2_1 b^2_2 \cdots b^2_n) &&\mapsto \alpha^{(n)}=0.b_1 b_2 \cdots b_{2n}
-\end{alignat}
+\begin{aligned}
+S^n : &\{0, \tfrac{1}{2^n}, \cdots, 1-\tfrac{1}{2^n}\}^2 \mapsto \{0, \tfrac{1}{4^n}, \cdots, 1-\tfrac{1}{4^n}\} \\
+&(v^{(n)}_1,v^{(n)}_2)=(0.b^1_1 b^1_2 \cdots b^1_n, 0.b^2_1 b^2_2 \cdots b^2_n) \mapsto \alpha^{(n)}=0.b_1 b_2 \cdots b_{2n}
+\end{aligned}
 $$ {#eq-Sn}
@@
 
+::: {.content-visible when-format="html"}
 ![(TODO 3D -> 2D only) Equivalence of dynamics, from 3D to 2D.](figures/embedding_3d_2d_1d.mp4)
+:::
+
+::: {.content-visible when-format="pdf"}
+![(TODO PDF, 3D -> 2D only) Equivalence of dynamics, from 3D to 2D.](figures/lol7.jpg)
+:::
 
@@
 
+::: {.content-visible when-format="html"}
 ![Equivalence of dynamics, from 3D to 2D to 1D.](figures/embedding_3d_2d_1d.mp4)
+:::
+
+::: {.content-visible when-format="pdf"}
+![(TODO PDF) Equivalence of dynamics, from 3D to 2D to 1D.](figures/lol7.jpg)
+:::
 

### 2023-06-15 manuscript/appendices/A_fixed_points.qmd
look a kitty !!! *baps with paw*
pwoofwead !!!1

@@
 
-# Fixed point study in the network of neurons
+# Fixed point study in the network of neurons {#sec-fixedpoints}
 
@@
 
-In order to find the fixed points, one has to solve in general $\dot h_i(t) = 0$ for all $i = 1, \cdots, N$.
+To find the fixed points, one has to solve in general $\dot h_i(t) = 0$ for all $i = 1, \cdots, N$.
 Because of the non-linearity $\phi(h_j(t))$, this is difficult, but we can guess the fixed points $h^\star_i$ and verify them by substituting into the equation that $\dot h_i(t) = 0$.
@@
 
-In the case of the network studied in this work, the find that
+In the case of the network studied in this work, we find that
 
@@
 
--   $h_i^\star = 0 \to h^\star(z_1, \cdots, z_p) = 0$
--   $h_i^\star = F_{\nu i} = z_{\nu i} \to h^\star(z_1, \cdots, z_p) = z_\nu$ for all $\nu = 1, \cdots, p$
+* $h_i^\star = 0 \to h^\star(z_1, \cdots, z_p) = 0$
+* $h_i^\star = F_{\nu i} = z_{\nu i} \to h^\star(z_1, \cdots, z_p) = z_\nu$ for all $\nu = 1, \cdots, p$
 
-[TODO : add derivation for neural field ?]
-
-## Stability of the fixed points
+## Stability of the fixed points {#sec-fixedpoints-stability}
 
@@
 
-The fixed point is stable if $\vec\delta(t) \to 0$ as $t \to \infty$, in other words if all the eigenvalues of $K$ are negative.
+The fixed point is stable if $\vec\delta(t) \to 0$ as $t \to \infty$, in other words, if all the eigenvalues of $K$ are negative.
 
-Before we study the spectrum of $K$, we cite the following result : if $\lambda$ is an eigenvalue of $A$, then $\lambda-1$ is an eigenvalue of $A-\mathrm{Id}$.
+Before we study the spectrum of $K$, we cite the following result: if $\lambda$ is an eigenvalue of $A$, then $\lambda-1$ is an eigenvalue of $A-\mathrm{Id}$.
 This follows simply from the equation for the eigenvalues $\lambda_i$ and eigenvectors $\vec{v_i}$ of $A$.
@@
 
-We study each term separately.
-Additionally (and similarly as before), we make the assumption that the functions $g_\mu$ and the functions $f_\mu$ are component-wise, and that the distribution $\rho$ factorises.
+We study each term separately. Additionally (and similarly as before), we assume that the functions $g_\mu$ and the functions $f_\mu$ are component-wise, and that the distribution $\rho$ factorises.
 
@@
 
-We now consider a Taylor expansion of the rate function $\phi(h)$ up to order 3, and cite results obtained from symbolic computation and known moments of the gaussian distribution.
+We now consider a Taylor expansion of the rate function $\phi(h)$ up to order 3, and cite results obtained from symbolic computation and known moments of the Gaussian distribution.
 
@@
 
-* When $\phi$ is linear, we see that all the eigenvalues are zero, therefore the first order stability is unconclusive.
+* When $\phi$ is linear, we see that all the eigenvalues are zero, therefore the first order stability is inconclusive.
 Numerically, we see that when $\phi$ is linear, the convergence to the fixed points is very slow.
@@
 The intuition is that we require $c_1$ to be large enough (in other words, the growth of the firing rate when $h$ becomes positive) and $c_3 < 0$ such that $3 c_1 c_3 + 15 c_3^2 < 0$.
+  The intuition behind this is that if the slope $c_1$ is stronger, then the bump of $\partial \phi$ becomes "more concentrated", which helps lower the eigenvalues associated with the pattern fixed points.
   The magnitude of the second order term $c_2$ only contributes to adding instability to the pattern fixed point, and stability to the zero fixed point, which leads to the conclusion it is best set to $c_2=0$ if we want the network to be stable at the pattern fixed points.
@@
 The fixed point $h_i^\star=0$ is therefore unstable, and the pattern fixed points $h_i^\star = z_{\nu i}$ are stable.
-We also note that since the normal distribution is centered around zero and symmetric, $h_i^\star = -z_{\nu i}$ are also stable fixed points.
+We also note that since the normal distribution is centred around zero and symmetric, $h_i^\star = -z_{\nu i}$ are also stable fixed points.

### 2023-06-15 manuscript/chapters/1_intro.qmd
pwoofwead !!!1

@@
 As models of spatiotemporal neural population dynamics on the cortical sheet, neural field models can be used to explain experimentally observed patterns of cortical activity, such as travelling waves in visual [@SatNau12;@MulRey14], somatosensory [@PetHah03;@FerBol06], motor [@RubRob06;@TakKim15], and hippocampal [@LubSia09;@PatFuj12;@PatSch13] cortices (see @MulCha18 for a review), and are used to model large-scale brain signals, such as electroencephalography (EEG) recordings @Bre17.
-See also @DipRan18 for an example of neural field modelling of 2-photon calcium imaging recording of the visual cortex.
+See also @DipRan18 for an example of neural field modelling of calcium imaging recording of the visual cortex.
 
@@
 The goal of this Master Thesis is to show that the answer to the second question is highly nontrivial because even the dimensionality of the embedding space, for some given spatiotemporal dynamics, is not clearly defined.
-This means that it is possible to find two embeddings (and connectivity kernels) of different dimensions that give rise to identical dynamics; contradicting the intuition that the embedding dimension should at least be as high as the dimensionality of the dynamics.
+This means that it is possible to find two embeddings (and connectivity kernels) of different dimensions that give rise to identical dynamics; *contradicting the intuition that the embedding dimension should at least be as high as the dimensionality of the dynamics* (the number of variables needed to fully describe the dynamical system).
 
 In particular, we show that for any two-dimensional neural field equation (that is, with two spatial dimensions), there is a one-dimensional neural field equation on the interval, where the embedding space is simply the interval $[0,1]$, from which the solution to the two-dimensional equation can be fully reconstructed.
-The mapping from the two-dimensional equation to the one-dimensional equation is done via well-known measurable bijections for the square $[0,1]^2$ and the interval $[0,1]$.
+The mapping from the two-dimensional equation to the one-dimensional equation is done via well-known measurable bijections between the square $[0,1]^2$ and the interval $[0,1]$.
 Importantly, these bijections are not (and can't be) diffeomorphisms.
@@
 
-If the embedding is well-chosen,[^1] the connectivity matrix $J_{ij}$ approaches a connectivity kernel $w(\vec y, \vec z)$ as $N \rightarrow \infty$, and neurons close in the embedding space tend to have similar potentials, such that we can write the continuous *neural field* $h(t, \vec z)$ of neuron potentials.
+If the embedding is well-chosen,[^1] the connectivity matrix $J_{ij}$ approaches a connectivity kernel $w(\vec y, \vec z)$ as $N \rightarrow \infty$, and neurons close in the embedding space tend to have similar potentials, such that we can write the continuous *neural field* $h(\vec z, t)$ of neuron potentials.
 Before we write down the neural field equation which describes how the field evolves in time, we need to introduce the neuron density $\rho(\vec z)$, which measures the number of neurons per unit volume of the embedding space.
@@
 $$
-\partial_t h(t, \vec z) = -h(t, \vec z) + \int_{\mathbb{R}^p} w(\vec z, \vec y) \phi(h(t, \vec y)) \rho(\mathrm d \vec y).
+\partial_t h(\vec z, t) = -h(\vec z, t) + \int_{\mathbb{R}^p} w(\vec z, \vec y) \phi(h(\vec y, t)) \rho(\mathrm d \vec y).
 $$ {#eq-nf}
@@
 $$
-\partial_t \tilde h(t, \alpha) = -\tilde h(t, \alpha) + \int_{[0,1]} \tilde w(\alpha, \beta) \phi(\tilde h(t, \beta)) \left[\lambda \circ S^{-1}\right](\d \beta)
+\partial_t \tilde h(\alpha, t) = -\tilde h(\alpha, t) + \int_{[0,1]} \tilde w(\alpha, \beta) \phi(\tilde h(\beta, t)) \left[\lambda \circ S^{-1}\right](\d \beta)
 $$
@@
 
-@sec-nf introduces a toy model for a neural field in $p$ dimensions, and we discuss interpretations of numerical simulation schemes.
+@sec-nf introduces a toy model for a neural field in $p$ dimensions, and tools to study its dynamics. We also discuss interpretations of numerical simulation schemes.
 
-@sec-mappings is dedicated to the discussion of mappings from 2 dimensions to 1 dimension, we give intuitions accompanied by simulations, and discuss in detail how "locality" emerges from these mappings.
+@sec-mappings is dedicated to the discussion of mappings from 2 dimensions to 1 dimension.
+We give intuitions accompanied by simulations, and discuss in detail how "locality" emerges from these mappings.
 

### 2023-06-15 manuscript/chapters/2_neural_fields.qmd
pwoofwead !!!1

@@
 $$
-G_{\mu i} = \tilde\phi(z_{i \mu}) \stackrel{\text{def}}{=} \frac{\phi(z_{i \mu}) - \avg{z_{i \mu}}}{\mathrm{Var}[\phi(z_{i \mu})]}.
+G_{\mu i} = \tilde\phi(z_{\mu i}) \stackrel{\text{def}}{=} \frac{\phi(z_{\mu i}) - \avg{z_{\mu i}}}{\mathrm{Var}[\phi(z_{\mu i})]}.
 $$
@@
 * $h^\star_i = z_{\mu i}$ is a stable fixed point for all $\mu \in \{1,\cdots,p\}$.
-Additionally, because the Gaussian distribution is symmetric around zero, $h_i = -z_{\mu i}$ are also stable fixed points.
+Additionally, because the Gaussian distribution is symmetric around zero, $h^\star_i = -z_{\mu i}$ are also stable fixed points.
 We refer to these fixed points as the "pattern" fixed points, because the network of neurons "remembers" the vectors $\vec{F_1}, \cdots, \vec{F_p}$.
@@
 This corresponds to a sharper difference between the "inactive" (low potential) and the "active" (high potential) neurons.
-The intuition behind this is that if the slope is stronger, then the bump of $\partial \phi$ becomes "more concentrated", which helps lower the eigenvalues associated to the pattern fixed points.
 3. The spectrum of $K$ is composed of $p$ eigenvalues that depend on the fixed point, and $N-p$ eigenvalues $\lambda_\perp = -1$ corresponding to the orthogonal complement of the pattern fixed points.
@@
 We should however note that the results obtained in @sec-fixedpoints are valid in the $N \to \infty$ limit.
-A numerical analysis shows in @fig-spectrum that already at $N=64$, the analytic eigenvalues match the numerical eigenvalues closely, and this correspondence improves when we take larger $N$ (typically we will for the rest of this Thesis take $N > 10^3$).
+A numerical analysis shows in @fig-spectrum that already at $N=64$, the analytic eigenvalues match the numerical eigenvalues closely, and this correspondence improves when we take larger $N$ (typically we will for the rest of this thesis take $N > 10^3$).
 
@@
 We first demonstrate what happens with numerical simulations in the simple case of $p=1$.
-We generate a network of 50,000 neurons, where each neuron samples a one-dimensional Gaussian to get its position.
+We generate a network of 50,000 neurons, where each neuron samples a one-dimensional Gaussian to get its position in the embedding.
 The function $\tilde \phi$ is computed by rescaling $\phi$ with the numerically computed mean and variance.[^code1]
@@
 We however note that in this case, there are two pattern fixed points that the network can converge to.
-In the example given here the network coincidentally, due to finite numerical effects, converged again to the first pattern fixed point, which we can see because the neural field is constant in the $z_2$ direction.
+In the example given here, the network coincidentally converged again to the first pattern fixed point.
+The network is in the $\mu=1$ pattern, which we can see because the neural field is constant in the $z_2$ direction.
 We write the fixed point $h^\star(z_1, z_2, t) = z_1$.
@@
 \begin{aligned}
-\mathcal{I}(\vec z) &= \int_{\Rp} \underbrace{w(\vec z, \vec y) \phi(h(t, \vec y))}_{I(\vec z, \vec y)} \rho(\d \vec y) \\
+\mathcal{I}(\vec z) &= \int_{\Rp} \underbrace{w(\vec z, \vec y) \phi(h(\vec y, t))}_{I(\vec z, \vec y)} \rho(\d \vec y) \\
 \hat{\mathcal{I}}(\vec z) &= \frac 1N \sum_{i=1}^N I(\vec z, \vec{y_i}), \, \vec{y_i} \stackrel{\text{iid}}{\sim} \rho
@@
 
-One small nuance we should add is that strictly speaking, Monte-Carlo integration should resample the integrand at every timestep, but the network of neurons samples the neurons once, and propagates these points through time.
+One small nuance we should add, is that strictly speaking, Monte-Carlo integration should resample the integrand at every timestep, but the network of neurons samples the neurons once, and propagates these points through time.
 
@@
 
-Another method to estimate an integral numerically is by regularly sampling the integrand on a fixed grid.
-We simply state the result that the convergence rate of the estimation is $\bO(N^{-2/p})$.
+Another method to estimate an integral numerically is by discretizing the integrand on a regular grid.
+We simply state the well-known result that the convergence rate of the estimation is $\bO(N^{-2/p})$.
 Interestingly, the convergence rate depends on the dimension $p$ of the integral.
@@
 
-We now introduce some tools that will help us to understand the toy model.
-We stress that in general, other neural field models might not have this luxury, and that we only make use of these tools to help explain the dynamics of the neural fields.
-Computationally, the overlaps and projections introduced here are only computed from the results of the simulation (see [`overlap.py`](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_overlap.py) in the source code).
+We now introduce some tools that will help us understand the toy model.
+We stress that in general, other neural field models might not have this luxury, and that we only make use of these tools to help explain the dynamics of the simulated neural fields.
+The overlaps and projections introduced here are only computed from the results of the simulation (see [`overlap.py`](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_overlap.py) in the source code).
 
@@
 m_\mu(t) &= \frac 1N \sum_{i=1}^N \tilde \phi(z_{\mu i}) \phi(h_i(t)) \\
-&\xrightarrow{N \to \infty} \int_{\Rp} \tilde \phi(z_\mu) \phi(h(\vec z, t)) \rho(\d \vec z)
+&\xrightarrow{N \to \infty} \int_{\Rp} \tilde \phi(z_\mu) \phi(h(\vec z, t)) \mathcal{N}^p(\d \vec z)
 \end{aligned}
@@
 
-This formulation "hides" the integral inside the overlap variables, but one should not forget that the overlaps depend on the neural field at a given time.
+This formulation "hides" the integral inside the overlap variables, but one should not forget that the overlaps depend on the neural field at any given time.
 
@@
 $$
-\kappa_\mu(t) = \int_{\Rp} y_\mu h(\vec y, t) \rho(\d \vec y),
+\kappa_\mu(t) = \int_{\Rp} y_\mu h(\vec y, t) \mathcal{N}^p(\d \vec y),
 $$ {#eq-kappa}
@@
 \begin{aligned}
-\dot \kappa_\mu(t) &= -\kappa_\mu(t) + \int_{\Rp} \tilde\phi(y_\mu) \phi(h(\vec y, t)) \rho(\mathrm d \vec y) = -\kappa_\mu(t) + m_\mu(t) \\
-\kappa_\mu(0) &= \int_{\mathbb{R}^p} y_\mu h(\vec y, 0) \rho(\mathrm d \vec y),
+\dot \kappa_\mu(t) &= -\kappa_\mu(t) + \int_{\Rp} \tilde\phi(y_\mu) \phi(h(\vec y, t)) \mathcal{N}^p(\mathrm d \vec y) = -\kappa_\mu(t) + m_\mu(t) \\
+\kappa_\mu(0) &= \int_{\mathbb{R}^p} y_\mu h(\vec y, 0) \mathcal{N}^p(\mathrm d \vec y),
 \end{aligned}
@@
 
-Additionally, it can be seen from the expression of the overlaps in @eq-overlap, that in the case of a linear activation function $\phi(h) = c_0 + c_1 h$, then $m_\mu(t) = \kappa_\mu(t)$.
+Additionally, it can be seen from the expression of the overlaps in @eq-overlap, that in the case of a linear activation function $\phi(h) = c_0 + c_1 h$, we have the equality $m_\mu(t) = \kappa_\mu(t)$.
 
-## A cycling neural field
+## A cycling neural field {#sec-cyling-nf}
 
@@
 We note however that @eq-nf-cycling-p no longer is a partial differential equation, but rather a delayed differential equation, which are in general much more complicated to solve.
-A simplifying assumption that we make is that we extend the initial condition back in time, such that $h(\vec z, t < 0) = h(\vec z, 0)$ is a constant function.
+A simplifying assumption that we make is we extend the initial condition back in time, such that $h(\vec z, t < 0) = h(\vec z, 0)$ is a constant function.
 Numerically, this also means that we have to store a history of the neural fields instead of only the current one.
-Since the `solve_ivp` method dynamically adapts the timestep of integration, we use linear interpolation between two known points in the stored neural field states to estimate $h(\vec z, t-\delta)$ (see [`lagging.py`](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_lagging.py) and [its use](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_rnn.py#L334)).
+Since the [`solve_ivp`](https://docs.scipy.org/doc/scipy/reference/generated/scipy.integrate.solve_ivp.html) method dynamically adapts the timestep of integration, we use linear interpolation between two known points in the stored neural field states to estimate $h(\vec z, t-\delta)$ (see [`lagging.py`](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_lagging.py) and [its use](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_rnn.py#L334)).
 
 With these considerations, we simulate a $p=2$ cycling neural field with $\delta=6$ and initial condition $h(z_1, z_2, t < 0) = z_1$ in @fig-cycling-anim.
-The latent trajectory is estimated from the simulation results, and animated simultaneously to the neural field.
+The latent trajectory is estimated from the simulation results, and animated simultaneously with the neural field.
 We clearly see the oscillations between the patterns $\mu=1$ and $\mu=2$.
@@
 
-In preparation for mapping of $[0,1]^2$ to $[0,1]$, we first map the 2-dimensional neural field equation in $\R^2$ to $[0,1]^2$, and, in general $\Rp$ to $[0,1]^p$.
-There are many ways to do this by using functions $\R \mapsto [0,1]$, but a practical choice is the Gaussian CDF (Cumulative Density Function), because it has the benefit of absorbing the density in the kernel, in the sense that the integral $[0,1]^p$ becomes weighted by the uniform distribution.
+In preparation to the next chapter, we first map the 2-dimensional neural field equation in $\R^2$ to $[0,1]^2$, and, in general $\Rp$ to $[0,1]^p$.
+There are many ways to do this by using functions $\R \mapsto [0,1]$, but a practical choice is the Gaussian CDF (Cumulative Density Function), because it has the benefit of absorbing the density into the kernel, in the sense that the integral $[0,1]^p$ becomes weighted by the uniform distribution.
 The one-dimensional Gaussian CDF is defined as
@@
 
-We make the final remark that this method of mapping $\Rp$ to $[0,1]^p$ does not work for any neural field.
-In our case, we use the fact that the distribution factorizes into $\rho(z_1, \cdots, z_p) = \rho_1(z_1) \cdots \rho_p(z_p)$, such that the components of $\vec{v}$ are sampled independently.
+We make the final remark that this method of mapping $\Rp$ to $[0,1]^p$ by using the CDF does not work for any neural field.
+In our case, we use the fact that the distribution factorizes into $\rho(z_1, \cdots, z_p) = \rho_1(z_1) \cdots \rho_p(z_p)$, such that the components of $\vec{v}$ are independent.
 We can then define the "componentwise CDF" that relates to the total CDF via

### 2023-06-15 manuscript/chapters/3_mappings.qmd
pwoofwead !!!1

@@
 In the previous chapter, we have established a simple toy model for a neural field in $p$ dimensions, and have shown how it can be embedded in $[0,1]^p$.
-In this chapter, we discuss in-depth what types of mappings can be used to map the $p$-dimensional space to $[0,1]$, and in particular first study the case $p=2$.
+In this chapter, we discuss in-depth what types of mappings can be used to map the $p$-dimensional space to $[0,1]$, and in particular we study the case $p=2$.
 Before getting into too much detail, let us give some insight as to why this problem is not trivial.
@@
 These types of models emerge when networks of neurons exhibit a finite number of "neuron types".
-We then group similar neurons ad make the approximation that all neurons inside the population have the same potential, and the same interaction weights with other neurons.
-In effect, every population then represents in actuality an infinity of discrete identical neurons.
+We then group neurons of the same type, and make the approximation that they have identical potentials and the same interaction weights with other neurons.
+In effect, every population then represents an infinite number of discrete identical neurons.
 
-The result of this is that the population dynamics can be written with the same equation as a finite network of neurons, but it already represents the $N \to \infty$ limit.
+This results in writing population dynamics with the same equation as a finite network of neurons.
+However, the equation *already represents* the $N \to \infty$ limit.
 We write the potential of the populations $H_a(t)$ and the population connectivity $\tilde J_{ab}$, where $a,b \in \{1, \cdots, N_\text{pop}\}^2$ in @eq-pop-dynamics.
@@
 We don't show the derivation of population dynamics here (see @GerKis14, chapter 12.3 for a detailed analysis), but we point out the implicit scaling behavior of $\tilde J_{ab}$.
-Letting $i$ be a neuron of population $a$ and $j$ be a neuron of population $b$, the connectivity between the two neurons is $J_{ij}$ and the population connectivity scales with the number of neurons in the presynaptic population $|b|$.
+Letting $i$ be a neuron of population $a$ and $j$ be a neuron of population $b$, the connectivity between the two neurons is $J_{ij}$, and the population connectivity scales with the number of neurons in the presynaptic population $|b|$.
 Furthermore, we make the reasonable assumption that $J_{ij}$ scales as $J^0_{ij} / N$, where $J^0_{ij} \sim \bO(1)$.
+We then have
 
 $$
-\tilde J_{ab} = \frac{|b|}{N} J^0_{ij}
+\tilde J_{ab} = \frac{|b|}{N} J^0_{ij}.
 $$
@@
 
-When taken in isolation, the previous developments might sound a little bit silly, though they serve to give an impression of the problem at hand, and an interpretation of the rest of this work.
-For a finite number of populations $N_\text{pop}$, the population dynamics is perfectly defined, however what happens as $N_\text{pop} \to \infty$ ?
+When taken in isolation, the previous developments might sound a little bit silly, though they serve to give an impression of the problem at hand, and an interpretation for the rest of this work.
+For a finite number of populations $N_\text{pop}$, the population dynamics is well-defined, however, what happens as $N_\text{pop} \to \infty$ ?
 The core of the question lies in the fact that we can easily *enumerate* a finite number of populations, but the sum $\sum_{b=1}^{N_\text{pop}}$ must have a clearly defined interpretation when this number tends to infinity.
@@
 When we discretize the $[0,1]^2$ space, we are effectively performing a coarse-graining approximation and simulating a finite number of populations.
-For every bin localized by $(v_1, v_2)$, we are sampling the connectivity kernel $w_U((v_1, v_2), \cdot)$ and 2D neural field $h_U((v_1, v_2), t)$ (see @fig-grid-sampling).
-In the rest of the chapter, we refer to these 2D bins as "square populations", and to the corresponding 1D bins as "segment populations".
+For every bin localized by $(v_1, v_2)$, we are sampling the connectivity kernel $w_U(\cdot, (v_1, v_2))$ and 2D neural field $h_U((v_1, v_2), t)$ (see @fig-grid-sampling).
+In the rest of the chapter, we refer to these 2D bins as "square populations", and to the corresponding 1D bins obtained by the mapping $S$ as "segment populations".
 
-![Every point of the grid effectively performs coarse-graining, and defines a point where we sample the connectivity kernel $w_U$ and the 2D neural field $h_U$](figures/photo_2023-06-10_21-13-23.jpg){#fig-grid-sampling}
+![(TODO, reorder args of $w_U$) Every point of the grid effectively performs coarse-graining, and defines a point where we sample the connectivity kernel $w_U$ and the 2D neural field $h_U$](figures/photo_2023-06-10_21-13-23.jpg){#fig-grid-sampling}
 
@@
 
-Therefore, we can see why the composition of $w_U$ with a possibly non-differentiable mapping $S$ might be problematic.
-Since the only thing we know about $S$ is that it is measurable and bijective, *there is no guarantee that the regularity of the kernel is conserved through the composition with the mapping $S$.*
+For these reasons, we can see why the composition of $w_U$ with a possibly non-differentiable mapping $S$ might be problematic.
+Since the only thing we know about $S$ is that it is measurable and bijective, *there is no guarantee that the regularity of the kernel is conserved through composition with the mapping $S$.*
 
@@
 
-Perhaps the linear projection or CDF examples are a bit too naive, and we need to look for more complex examples.
+Perhaps these examples are a bit too naive, and we need to look for more complex mappings.
 We might ask the question of whether a diffeomorphism (a differentiable bijection with a differentiable inverse) even exists between the unit square and the unit segment.
@@
 
-1. Cantor functions are bijections between $[0,1]$ and $[0,1]^2$. By the non-existence theorem, these functions cannot be continuous.
+1. Cantor functions are bijections between $[0,1]$ and $[0,1]^2$. By Netto's theorem, these functions cannot be continuous.
 2.
@@
 
-In light of these results, we might be tempted to search for expressions for Cantor-type functions, since having a bijection would allow us to have a direct mapping between the square and the segment populations.
+In light of these results, we might be tempted to search for expressions of Cantor-type functions, since having a bijection would allow us to have a direct mapping between the square and the segment populations.
 Unfortunately, these functions are discontinuous which breaks the regularity of the kernel.
 
-Peano functions seem like good candidates for the mapping $S^{-1}$, and are extensively studied in the literature (we refer for instance to [@sagan_space-filling_1994;@Sagan1993]).
+Peano functions seem like good candidates for the mapping $S^{-1}$, and are extensively studied in the literature around space-filling curves (we refer for instance to [@sagan_space-filling_1994]).
 They have the nice property that they are continuous and surjective, so we the segment populations can "reach" the entirety of the square populations.
 All things considered, non-injectivity is not a problem, since it just means that two segment populations can "refer" to the same square population, which does not break the regularity of the kernel, since the two segment populations can just have their own neighborhoods.
-Furthermore, we will see that in our examples, the non-injectivity can be ignored because the mapping is quasi-bijective.
+Furthermore, we will see that in our examples, the non-injectivity can be ignored because the mapping is "quasi-bijective".
 
@@
 
-We consider $(v_1, v_2) \in [0,1]^2$. Given $n \in \mathbb N$, we write a (truncated) binary expansion of $v_1$ and $v_2$.
+Previously, we motivated $S$ as the limit of functions $S^n$ acting on finer and finer grids.
+Let us consider $(v_1, v_2) \in [0,1]^2$.
+Given $n \in \mathbb N$, we write a (truncated) binary expansion of $v_1$ and $v_2$.
 
@@
 
-The intuition behind this expansion is we take a segment, split it in two, write 0 to go left, write one to go right, then recurse.
-Then combine the bits of $v^{(n)}_1$ and $v^{(n)}_2$ to get the position in the $[0,1]^2$ embedding.
+The intuition and illustration @fig-recurse-binary behind this expansion is that we locate each component by recursively splitting the segment in two sub-segments of equal length.
+At every step, taking the left or right segments defines the sequence of binary digits.
+<!-- Then combine the bits of $v^{(n)}_1$ and $v^{(n)}_2$ to get the position in the $[0,1]^2$ embedding. -->
 
-![Splitting $[0,1]$ recursively gives the binary bit expansion of $v_1$](figures/photo_2023-06-10_17-21-28.jpg)
+![(TODO rename bins to segments) Splitting $[0,1]$ recursively gives the binary bit expansion of $v_1$](figures/photo_2023-06-10_17-21-28.jpg){#fig-recurse-binary}
 
@@
 In general, each bit $b^k$ of the image would be a binary function of the $n$ bits of $v_1$ and the $n$ bits of $v_2$, implicitly $b_k = b_k(b^1_1, b^1_2, \cdots, b^1_n, b^2_1, b^2_2, \cdots, b^2_n)$.
-With this, we can formulate any mapping of $4^n$ distinct elements (here : squares of size $2^{-n} \times 2^{-n}$) onto $4^n$ elements (here : segments of length $4^{-n}$).
+With this, we can formulate any mapping between sets of $4^n$ distinct elements.
 
-The intuition behind the use of $2n$ bits for the image is that we take $2^n \times 2^n = 4^n$ squares in $[0,1]^2$ and map those to $4^n$ segments in $[0,1]$.
+The intuition behind the use of $2n$ bits for the image is that we take $2^n \times 2^n = 4^n$ squares of size $2^{-n} \times 2^{-n}$ in $[0,1]^2$, and map those to $4^n$ segments of size $4^{-n}$ in $[0,1]$.
 Informally, we map $n+n$ bits onto $2n$ bits.
 
-In the following, we will restrict ourselves to mappings where the functions $b_k(\{b^1_l\}_{l\in\{1,\cdots,n\}}, \{b^2_l\}_{l\in\{1,\cdots,n\}})$ are defined in such a way that each bit of the inputs is used once, and only once.
-Therefore, for each bit $k$ there is a function $b_k(b^{\mu_k}_{l_k})$ acting on only one bit of the input.
+In the following, we will restrict ourselves to mappings where the functions $b_k$ are defined in such a way that each bit of the inputs $v_1$ and $v_2$ is used only once.
+Therefore, for each bit $k$ of the output, there is a function $b_k(b^{\mu_k}_{l_k})$ acting on only one bit of the input.
 $\mu_k \in \{1,2\}$ is the input component and $l_k \in \{1,\cdots,n\}$ the bit number corresponding to $k$.
@@
 
-For finite $n$, we write the inverse of the column mapping in @eq-column-inverse.
+For finite $n$, we write the inverse of the column mapping:
 
@@
 v^{(n)}_1 = C^{-1}_1(\alpha^{(n)}) &= 0.b_1 b_2 \cdots b_n = \sum_{k=1}^n b_{k} 2^{-k} \\
-v^{(n)}_2 = C^{-1}_2(\alpha^{(n)}) &= 0.b_{n+1} b_{n+2} \cdots b_{2n} = \sum_{k=1}^n b_{n+k} 2^{-k}
+v^{(n)}_2 = C^{-1}_2(\alpha^{(n)}) &= 0.b_{n+1} b_{n+2} \cdots b_{2n} = \sum_{k=1}^n b_{n+k} 2^{-k}.
 \end{aligned}
@@
 In such (contiguous) arrays, the values are laid out in memory in "row-major" order (C-style) or "column-major" order (Fortran-style); and we can go from the 1D memory representation to a 2D matrix representation by using "reshape" operations.
-Effectively, this is how this mapping is implemented in code, using `numpy.ravel_multi_index` (2D to 1D) and `numpy.unravel_index` (1D to 2D).
+Effectively, this is how this mapping is implemented in code, using [`numpy.ravel_multi_index`](https://numpy.org/doc/stable/reference/generated/numpy.ravel_multi_index.html) (2D to 1D) and [`numpy.unravel_index`](https://numpy.org/doc/stable/reference/generated/numpy.unravel_index.html) (1D to 2D).
 
@@
 One way of ordering the 2D texture data in the 1D memory is by using the Z-order curve (also known as "Morton mapping" or "Lebesgue curve" @Sagan1994).
-In this way, points close in 2D *tend* to be tightly packed in 1D.
-We can define the Z-order curve (Z-mapping) in the following way[^zcurve] :
+In this way, points close in 1D *tend* to be tightly packed in 2D.
+We can define the Z-order curve (Z-mapping) in the following way:[^zcurve]
 
@@
 
-And its inverse is :
+And its inverse is:
 
@@
 v^{(n)}_1 = Z^{-1}_1(\alpha^{(n)}) &= 0.b_1 b_3 \cdots b_{2n-1} = \sum_{k=1}^n b_{2k-1} 2^{-k} \\
-v^{(n)}_2 = Z^{-1}_2(\alpha^{(n)}) &= 0.b_2 b_4 \cdots b_{2n} = \sum_{k=1}^n b_{2k} 2^{-k}
+v^{(n)}_2 = Z^{-1}_2(\alpha^{(n)}) &= 0.b_2 b_4 \cdots b_{2n} = \sum_{k=1}^n b_{2k} 2^{-k}.
 \end{aligned}
@@
 
-@fig-z is a visualization for the Z-mapping when $n=2$, and shows how it is composed of the "Z" shapes giving it its name.
+@fig-z is a visualization of the Z-mapping when $n=2$, and shows how it is composed of the "Z" shapes giving it its name.
 
@@
 #| label: fig-anti-z
-#| fig-cap: Anti-Z mapping for $n=2$.
+#| fig-cap: Anti-Z mapping for $n=2,3,4$.
 Numbers represent the ordering of the populations in the $[0,1]$ embedding.
 We don't plot the trace, because it has many crossovers and overlaps which makes the plot hard to read.
+#| fig-subcap: 
+#|   - $n=2$
+#|   - $n=3$
+#|   - $n=4$
+#| layout-ncol: 3
 #| echo: false
 fig, ax = plot_trace(AntiZMapping, nrec=2, do_trace=False)
+display(fig)
+fig, ax = plot_trace(AntiZMapping, nrec=3, do_trace=False)
+display(fig)
+fig, ax = plot_trace(AntiZMapping, nrec=4, do_trace=False, do_numbers=False)
 display(fig)
@@
 @fig-anti-z gives some insight into the problem of this mapping.
-For a fixed input $(v_1,v_2)$, we see that the output point position jumps around and that these jumps do not decay, they are always of order 1.
+For a fixed input $(v_1,v_2)$, we see that the output point position jumps around when $n$ is changed, and that these jumps seem to always be of order 1.
 The intuition behind pointwise convergence is that as $n$ increases, the image of a point by $S^n$ is continually refined.
 
-Another way of looking at this is to consider how $\alpha^{(n)}$ fluctuates as the precision of the input increases with $n \to \infty$.
+Consider how $\alpha^{(n)}$ fluctuates as the precision of the input increases with $n \to \infty$.
 As we refine the input with fluctuations of order $2^{-n}$, the output $\alpha^{(n)} = b^1_n 4^{-1} + b^2_n 4^{-2} + b^1_{n-1} 4^{-3} + b^2_{n-1} 4^{-4} \cdots$ fluctuates (at least) with amplitude $\sim 4^{-2}$ independantly of $n$.
@@
 
-The limit is not bijective, because the bits of $v_2$ are "lost" in the $n\to\infty$ limit, and the inverse $C_2^{-1}(\alpha)$ is not well-defined, since it "takes bits from $\infty$ to $2\infty$".
+The limit is not bijective, because the bits of $v_2$ are "lost" in the $n\to\infty$ limit, and the inverse $C_2^{-1}(\alpha)$ is not well-defined, since it "takes bits between $\infty$ and $2\infty$".
 
-The Z-mapping however seems to converge to a fractal curve.
-Numerically, there seems to be an infinite number of discontinuities, but the size of these discontinuities is overwhelmingly small, and *on average* the size of these numerical discontinuities vanishes as $n\to\infty$.
+The Z-mapping seems to converge to a fractal curve.
+Numerically, there seems to be an infinite number of discontinuities, but the size of these discontinuities is overwhelmingly small, so that *on average* they vanish as $n\to\infty$.
 We won't go further into detail about this in this section (in @sec-zcurve-lebesgue we develop the convergence to a space-filling fractal curve), but this gives some intuition for the arguments presented in @sec-locality.
@@
 We can evidently define mappings where we interleave the bits in a different order, such as $0.b^2_1 b^1_1 b^2_2 b^1_2 \cdots$; or mappings where we reorder the $k$ bits of the input, as in $0.b^1_2 b^2_2 b^1_1 b^2_1 b^1_4 b^2_4 b^1_3 b^2_3 \cdots$, but the conditions restrict the reordering to be permutations of neighbouring bits.
-In essence, these mappings are not fundamentally different, and the Z-mapping can be seen as a "canonical representation" of the pointwise convergent mappings with bijective limits.
+In essence, these mappings are not fundamentally different, and the Z-mapping can be seen as a "canonical representation" of the pointwise convergent mappings with bijective limits, which can be expressed as the permutation of binary digits.
 
@@
 
-## Coarse-graining the neural field in $[0,1]$
+## Coarse-graining the neural field in $[0,1]$ {#sec-coarse-graining}
 
@@
 In the previous sections, we have discussed what mappings have good properties, and have given a few hints as to how we can implement them numerically.
-Since the mappings we presented can be described as a binary expansion, they naturally lend themselves to numerical approximations.
-Furthermore, this is made especially easy since the mappings we defined simply are permutations of bits (in the $n < \infty$ case), and so we can write a one-to-one mapping between (unsigned integers).
+Since the mappings we presented can be written as a binary expansion, they naturally lend themselves to numerical approximations.
+Furthermore, this is made especially easy since the mappings we defined simply are permutations of bits (in the $n < \infty$ case), and so we can write a one-to-one correspondence between unsigned integers.
 
@@
 We however note that using binning methods, the Monte-Carlo simulation can still be applied.
-The intuition behind this is that sampling populations inside a small square of size $2^{-n} \times 2^{-n}$ gives rise to the notion of an "average connectivity kernel", and in our case the notion of "mean patterns".
+The intuition behind this is that sampled populations inside a small square of size $2^{-n} \times 2^{-n}$ give rise to the notion of an "average connectivity kernel", and in our case the notion of "mean patterns".
 More details are given in @sec-mean-patterns.
@@
 
-### Iterative application of $S$ : from $[0,1]^3$ to $[0,1]$
+### Iterative application of $S$ : from $[0,1]^3$ to $[0,1]$ {#sec-simulations-3d-2d-1d}
 

### 2023-06-15 manuscript/chapters/4_locality.qmd
pwoofwead !!!1

@@
 In this chapter, we will further explore these results, and attempt to formulate a measure of this "locality".
-We start by showing that bijectivity and measurability is a sufficient condition for equivalent dynamics.
+We start by showing that bijectivity and measurability are sufficient conditions for equivalent dynamics.
 Then, we apply the locality measure to the special cases of the Column and Z-mappings, and show how this notion can discriminate between the two.
@@
 
-In the previous chapter, we have introduced Cantor and Peano functions, and argued that if have to chose between the two, that the non-injective Peano functions are preferable, because by Netto's theorem, they are the only ones that can be space-filling and reach all the square populations.
+In the previous chapter, we introduced Cantor and Peano functions, and argued that if have to choose between the two, then the non-injective Peano functions are preferable, because, by Netto's theorem, they are the only ones that can be space-filling and reach all the square populations.
 Despite this, let us make the simplifying assumption that $S$ is a bijective function.
-However, as to not break Netto's theorem, we do not make any assumptions regarding its continuity.
+However, to not break Netto's theorem, we do not make any assumptions regarding its continuity.
 
@@
 
-We start this the one-dimensional neural field, then substitute @eq-analytical-equivalence-tildeh and $\vec{v} = S^{-1}(\alpha),\, \vec{u} = S^{-1}(\beta)$.
+We start by recalling the expression of the one-dimensional neural field, then substitute @eq-analytical-equivalence-tildeh and $\vec{v} = S^{-1}(\alpha),\, \vec{u} = S^{-1}(\beta)$.
 
@@
 \partial_t \tilde h(\alpha, t) &= -\tilde h(\alpha, t) + \int_{[0,1]} \tilde w(\alpha, \beta) \phi(\tilde h(\beta, t)) \left[\lambda \circ S^{-1}\right](\d \beta) \\
+\iff&\\
 \partial_t h_U(S^{-1}(\alpha), t) &= 
@@
 Therefore, if $h_U(\vec v, t)$ solves the neural field equation in $[0,1]^2$, then $\tilde h(t) = h_U(t) \circ S^{-1}$ solves the equivalent neural field equation in $[0,1]$.
-The same reasoning applies in reverse: if $\tilde h(t, \alpha)$ solves the neural field equation in $[0,1]$, then $h_U(t) = \tilde h \circ S$ solves the neural field in $[0,1]^2$.
+The same reasoning applies in reverse: if $\tilde h(\alpha, t)$ solves the neural field equation in $[0,1]$, then $h_U(t) = \tilde h \circ S$ solves the neural field in $[0,1]^2$.
 
@@
 The previous chapter gave the intuition that local mappings map populations close in 1D to populations close in 2D.
-In order to verify this, we might try to compare the distance of all segment populations to the population at $\alpha=0$, to the distance of all square populations to the population at $S^{-1}(\alpha=0)$.
-This is done in @fig-distance-2d-1d, where the distance is taken to be the $\ell^2$ norm.
+To verify this, we might try to compare the distance between all segment populations and the population at $\alpha=0$, to the distance of all square populations to the population at $S^{-1}(\alpha=0)$.
+This is done in @fig-distance-2d-1d, where the 2D distance is taken to be the $\ell^2$ norm.
 If the distances matched perfectly, then we would see a line with slope $\sqrt 2$ in the graph.
 
-![(TODO random, column and zcurve only) Distance between the first population and all other populations, in 1D and 2D.
+![(TODO random, column and zcurve only, rename neuron to population) Distance between the first population and all other populations, in 1D and 2D.
 The dotted line has a slope $\sqrt 2$.](figures/distance_2d_1d.png){#fig-distance-2d-1d}
@@
 
-Denoting $\alpha_i,\,i \in \{1, \cdots, 4^n\}$ the positions of the grid in the $[0,1]$ embedding, we define the total variation as the projection of all the corresponding 2D segments:[^tv]
+Denoting $\alpha_i,\,i \in \{1, \cdots, 4^n\}$ the positions of the grid in the $[0,1]$ embedding, we define the total variation as the projection of all the corresponding 2D jumps onto the axes:[^tv]
 
@@
 Additionally, we make use of the $\ell^1$ norm because it lends itself to the nice geometric intuition of "projecting the jumps" on the axes.
-This choice of the norm is irrelevant however, because analytically all norms are equivalent up to a constant.
+This choice of the norm is irrelevant, however, because analytically all norms are equivalent up to a constant.
 
@@
 
-Dissapointingly, all mappings seem to exhibit the same growth of total variation.
+Disappointingly, all mappings seem to exhibit the same growth of total variation.
 It seems like our notion of total variation is not adequate to help discriminate between the mappings.
@@
 
-<!-- In some sence, having the constraint that the mapping traces a path in the bounded space $[0,1]^2$ that can go through each point of the grid only once leaves little flexibility.
-For fixed $n$, there are only finitely many "large jumps" the mapping can make, and the large number of "small jumps" still add up, so that the total variation still grows at approximatively the same rate.
+<!-- In some sense, having the constraint that the mapping traces a path in the bounded space $[0,1]^2$ that can go through each point of the grid only once leaves little flexibility.
+For fixed $n$, there are only finitely many "large jumps" the mapping can make, and the large number of "small jumps" still add up, so that the total variation still grows at approximately the same rate.
 -->
 
-### Locality as vanishing average binned variation
+### Locality as vanishing average binned variation {#sec-Vn}
 
@@
 
-![(TODO the variation for the lower one shoud be lesser) Intuition behind the average binned variation.
+![(TODO the variation for the lower one should be lesser) Intuition behind the average binned variation.
 To simplify the visualization, activity levels are used in lieu of the 2D positions.](figures/photo_2023-06-15_00-44-43.jpg){width=400px #fig-sketch-binned-variation}
@@
 variation inside the bins, strong version.
-basically we take the max over all binning methods that have bins smaller than $\delta$.
+basically, we take the max over all binning methods that have bins smaller than $\delta$.
 Let $P_\delta$ be a permutation of $[0,1]$ with $n_{P_\delta}$ elements $\{\alpha_1 \leq \alpha_2 \leq \cdots \alpha_{n_{P_\delta}}\}$, where $\alpha_{i}-\alpha_{i-1} \leq \delta$, and $\mathcal{P}_\delta$ the set of such permutations.
@@
 we need to convince people this is not a big mistake, since $S^n \to S$, and we can take both limits (binning and $S^n$) "at the same time".
-This might actually be adressed by a similar todo in @sec-mappings-binary
+This might actually be addressed by a similar todo in @sec-mappings-binary
 
-Before we move to analytical derivations applying the average variation to mappings, we can first check numerically if it is gives the expected results; that is $V_n$ vanishes a $n\to\infty$ for local mappings, whereas $V_n$ is lower bounded by a constant for non-local mappings.
-@fig-vnum shows this is indeed the case, rejoice !
+Before we move to analytical derivations applying the average variation to mappings, we can first check numerically if it gives the expected results; that is $V_n$ vanishes a $n\to\infty$ for local mappings, whereas $V_n$ is lower bounded by a constant for non-local mappings.
+@fig-vnum shows this is indeed the case, rejoice!
 
@@
 
-In order to compute (a bound on) $V_n$ for some mappings, we consider the binary expansion of $\alpha$ and $\alpha^\prime$ which appear in the supremum.
+To compute (a bound on) $V_n$ for some mappings, we consider the binary expansion of $\alpha$ and $\alpha^\prime$ which appear in the supremum.
 For clarity and to represent actual numerical implementations, we furthermore truncate the binary expansion to $2m$ bits (on most machines, $2m=64$ bits).
@@
 \begin{aligned}
-|Z^{-1}_1(\alpha) - Z^{-1}_1(\alpha^\prime)| &= 0.0 \cdots 0 (b_n b_{n+2} \cdots b_{2m-1} - b^\prime_n b^\prime_{n+2} \cdots b^\prime_{2m-1}) \\
-&= 2^{-n/2} \sum_{k=1}^{\tfrac{2m-n+1}{2}} (b_{n+2k-1} - b^\prime_{n+2k-1}) 2^{-k} \\
+|Z^{-1}_1(\alpha) - Z^{-1}_1(\alpha^\prime)| &= \left|0.0 \cdots 0 (b_n b_{n+2} \cdots b_{2m-1} - b^\prime_n b^\prime_{n+2} \cdots b^\prime_{2m-1})\right| \\
+&= \left|2^{-n/2} \sum_{k=1}^{\tfrac{2m-n+1}{2}} (b_{n+2k-1} - b^\prime_{n+2k-1}) 2^{-k}\right| \\
 &\leq 2^{-n/2} \\
-|Z^{-1}_2(\alpha) - Z^{-1}_2(\alpha^\prime)| &= 2^{-n/2} \sum_{k=1}^{\tfrac{2m-n+1}{2}} (b_{n+2k} - b^\prime_{n+2k}) 2^{-k} \\
+|Z^{-1}_2(\alpha) - Z^{-1}_2(\alpha^\prime)| &= \left|2^{-n/2} \sum_{k=1}^{\tfrac{2m-n+1}{2}} (b_{n+2k} - b^\prime_{n+2k}) 2^{-k}\right| \\
 &\leq 2^{-n/2}
@@
 
-We now consider the column mapping. Since here we consider finite-precision $\alpha$ and $\alpha^\prime$, the invere is well-defined and we have
+We now consider the Column mapping. Since here we consider finite-precision $\alpha$ and $\alpha^\prime$, the invere is well-defined and we have
 
@@
 \begin{aligned}
-|C^{-1}_1(\alpha) - C^{-1}_1(\alpha^\prime)| &= 2^{-n} \sum_{k=1}^{m-n+1} (b_{n+k-1} - b^\prime_{n+k-1}) 2^{-k} \\
+|C^{-1}_1(\alpha) - C^{-1}_1(\alpha^\prime)| &= \left| 2^{-n} \sum_{k=1}^{m-n+1} (b_{n+k-1} - b^\prime_{n+k-1}) 2^{-k} \right| \\
 &\leq 2^{-n}.
@@
 
-However now problems arise when we compute the averge binned variation, since the second component is not well-behaved.
+However, now problems arise when we compute the averge binned variation, since the second component is not well-behaved.
 In particular, since we take the supremum on $\alpha,\alpha^\prime \in \left[\tfrac{i-1}{2^n},\tfrac{i}{2^n}\right]^2$, we can pick the specific values such that $b_k=1$ and $b^\prime_k=0$ for all $k \in \{m+1, \cdots 2m\}$.
@@
 \begin{aligned}
-|C^{-1}_2(\alpha) - C^{-1}_2(\alpha^\prime)| &= \sum_{k=m+1}^{2m} (b_k - b^\prime_k) 2^{-(k-m)} \\
+|C^{-1}_2(\alpha) - C^{-1}_2(\alpha^\prime)| &= \left| \sum_{k=m+1}^{2m} (b_k - b^\prime_k) 2^{-(k-m)} \right| \\
 &= \sum_{k=m+1}^{2m} 2^{-(k-m)} \\
@@
 
-## Numerical feasibility
+## Numerical convergence of the finite-$n$ estimations
 
-numerically, we have finitely many bits, and cannot fully represent the fractal. we therefore need the finite approximations to be already good.
+In @sec-coarse-graining we discussed how we approximate the fractal mappings numerically, and introduced the downsampling procedure, in which we average the populations in segments of length $2^{-n}$, and end up with $2^n$ "binned" segment populations, compared to the $4^n$ we started with.
 
-locality might be a side-effect from bijections
+By doing this, we were able to numerically estimate the integral on the one-dimensional fractal space, and showed that local mappings exhibited the property that they conserved the dynamics between two spaces of different dimensions.
+@sec-Vn introduced the analytical metric $V_n$ of this locality.
+Assuming that $V_n \xrightarrow{n\to\infty} 0$, is this enough to prove that the numerical estimation of the integral goes to its analytical value?
 
-Before we make the proof of numerical feasibility.
+The intuition behind the proof of convergence is that the variation, coming from the coarse-graining, inside each numerical bin vanishes as we increase the number of iterations $n$, even though the structure becomes fractal.
 
-In the evaluation of the evolution equation, the integral term in each coordinate $\alpha$ is problematic.
-
-Let us define (recall) the numerical and analytical integral (we leave a $w_U(\cdot, \beta)$ as implicit function of the first argument $\alpha$).
-The intuition behind this proof of convergence is that the variation inside of each numerical bin vanishes as we increase the number of iterations $n$ (even though the structure becomes fractal !).
-The variation comes from the coarse-graining, where we go from $4^n$ to $2^n$ samples.
-
-We define $\tilde h_i(t) = \tilde h(t, \tfrac{i-1}{2^n}) = \tilde h(t, \beta_i),\,\beta_i=\tfrac{i-1}{2^n},\,i\in\{1,\cdots,2^n\}$.
+We define the one-dimensional field on the grid:
 
 $$
+\tilde h_i(t) = \tilde h(\tfrac{i-1}{2^n}, t) = \tilde h(\beta_i, t),\,\beta_i=\tfrac{i-1}{2^n},\,i\in\{1,\cdots,2^n\}.
+$$
+
+Let us first recall the numerical and analytical integral. We leave a $\tilde w(\cdot, \beta)$ as an implicit function of the first argument $\alpha$. 
+$$
 \begin{aligned}
-\mathrm{NI}_n &= \frac{1}{2^n} \sum_{i=1}^{2^n} \tilde w(\cdot, \beta_i) \phi(\tilde h_i) \\
-\mathrm{AI} &= \int_{[0,1]} \tilde w(\cdot, \beta) \phi(\tilde h(t, \beta)) \lambda \circ S^{-1}(\d \beta)
+\mathrm{NI}_n &= \frac{1}{2^n} \sum_{i=1}^{2^n} \tilde w(\cdot, \beta_i) \phi(\tilde h_i(t)) \\
+\mathrm{AI} &= \int_{[0,1]} \tilde w(\cdot, \beta) \phi(\tilde h(\beta, t)) \left[\lambda \circ S^{-1}\right](\d \beta)
 \end{aligned}
@@
 
-We wish to prove that $|\mathrm{NI}_n - \mathrm{AI}| \xrightarrow{n \to \infty} 0$ (implicitly : this is valid for all $\alpha \in [0,1]$).
+We wish to prove that $|\mathrm{NI}_n - \mathrm{AI}| \xrightarrow{n \to \infty} 0$, and implicitly, that this is valid for all $\alpha \in [0,1]$.
 
@@
 \begin{aligned}
-|\mathrm{NI}_n - \mathrm{AI}| &= \left| \frac{1}{2^n} \sum_{i=1}^{2^n} \tilde w(\cdot, \beta_i) \phi(\tilde h_i) - \int_{0}^{1} \tilde w(\cdot, \beta) \phi(\tilde h(t, \beta)) \lambda \circ S^{-1}(\d \beta) \right| \\
-&= \left| \frac{1}{2^n} \sum_{i=1}^{2^n} \tilde w(\cdot, \beta_i) \phi(\tilde h_i) - \sum_{i=1}^{2^n} \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \tilde w(\cdot, \beta) \phi(\tilde h(t, \beta)) \lambda \circ S^{-1}(\d \beta) \right| \\
-&= \left| \sum_{i=1}^{2^n} \frac{1}{2^n} \tilde w(\cdot, \beta_i) \phi(\tilde h_i) - \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \tilde w(\cdot, \beta) \phi(\tilde h(t, \beta)) \lambda \circ S^{-1}(\d \beta) \right| \\
-&\leq \sum_{i=1}^{2^n} \left| \frac{1}{2^n} \tilde w(\cdot, \beta_i) \phi(\tilde h_i) - \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \tilde w(\cdot, \beta) \phi(\tilde h(t, \beta)) \lambda \circ S^{-1}(\d \beta) \right| \\
+|\mathrm{NI}_n - \mathrm{AI}| &= \left| \frac{1}{2^n} \sum_{i=1}^{2^n} \tilde w(\cdot, \beta_i) \phi(\tilde h_i(t)) - \int_{0}^{1} \tilde w(\cdot, \beta) \phi(\tilde h(\beta, t)) \left[\lambda \circ S^{-1}\right](\d \beta) \right| \\
+&= \left| \frac{1}{2^n} \sum_{i=1}^{2^n} \tilde w(\cdot, \beta_i) \phi(\tilde h_i(t)) - \sum_{i=1}^{2^n} \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \tilde w(\cdot, \beta) \phi(\tilde h(\beta, t)) \left[\lambda \circ S^{-1}\right](\d \beta) \right| \\
+&= \left| \sum_{i=1}^{2^n} \frac{1}{2^n} \tilde w(\cdot, \beta_i) \phi(\tilde h_i(t)) - \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \tilde w(\cdot, \beta) \phi(\tilde h(\beta, t)) \left[\lambda \circ S^{-1}\right](\d \beta) \right| \\
+&\leq \sum_{i=1}^{2^n} \left| \frac{1}{2^n} \tilde w(\cdot, \beta_i) \phi(\tilde h_i(t)) - \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \tilde w(\cdot, \beta) \phi(\tilde h(\beta, t)) \left[\lambda \circ S^{-1}\right](\d \beta) \right| \\
 \end{aligned}
@@
 We now want to prove that evaluating the integrand at $\beta_i = \tfrac{i-1}{2^n}$ gives a good approximation of the integral.
-Since $\int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \lambda \circ S^{-1}(\d \beta) = \tfrac{1}{2^n}$ [TODO MATH : is this true ?
-this is obvious when we consider the binary expansion of $S^{-1}$], we write
+Because $S$ is measurable and bijective, we have $\int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \left[\lambda \circ S^{-1}\right](\d \beta) = \tfrac{1}{2^n}$, and we write
 
@@
 \begin{aligned}
-|\mathrm{NI}_n - \mathrm{AI}| &\leq \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \left[ \tilde w(\cdot, \beta_i) \phi(\tilde h_i) - \tilde w(\cdot, \beta) \phi(\tilde h(t, \beta)) \right] \lambda \circ S^{-1}(\d \beta) \right| \\
-&= \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \left[ w_U(\cdot, S^{-1}(\beta_i)) \phi(h_U(t, S^{-1}(\beta_i))) - w_U(\cdot, S^{-1}(\beta)) \phi(h_U(t, S^{-1}(\beta))) \right] \lambda \circ S^{-1}(\d \beta) \right|
+|\mathrm{NI}_n - \mathrm{AI}| &\leq \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \Delta(\beta) \left[\lambda \circ S^{-1}\right](\d \beta) \right| \\
+\text{where}\ \Delta(\beta) &= \tilde w(\cdot, \beta_i) \phi(\tilde h_i(t)) - \tilde w(\cdot, \beta) \phi(\tilde h(\beta, t))\\
+&= w_U(\cdot, S^{-1}(\beta_i)) \phi(h_U(S^{-1}(\beta_i), t)) - w_U(\cdot, S^{-1}(\beta)) \phi(h_U(S^{-1}(\beta), t))
 \end{aligned}
@@
 
-The main source of variance inside the integrand is the term $S^{-1}$, since $w_U$ and $h_U$ are "nice" functions.
-We express this regularity by the added assumption that $w_U(\cdot, \vec z) h_U(t, \vec z)$ is Lipschitz in $\vec z$, with Lipschitz constant $L(\cdot)$, and $\norm{\cdot}$ the corresponding norm on $[0,1]^2$.
-[TODO : argue on the niceness of $I_rec$, add $\phi$]
+The main source of variance inside the integrand $\Delta(\beta)$ is the term $S^{-1}$, since $w_U$ and $h_U$ are "nice" functions.
+We express this regularity by the added assumption that $w_U(\cdot, \vec u) \phi(h_U(t, \vec u))$ is Lipschitz in $\vec u$, with Lipschitz constant $L(\cdot)$, and $\norm{\cdot}$ the corresponding norm on $[0,1]^2$.
+We argue this is not a strong assumption, because this is just expressing the regularity of the integrand in the integral over $[0,1]^2$.
 
-Once we have applied the Lipschitz property, we make use of our intuition of "locality" : points close in $[0,1]$ should be (on average) mapped to points close in $[0,1]^2$.
-Therefore, taking the sup inside each $2^{-n}$ segment does not introduce a too large error.
+A more pragmatic justification of this is to notice that the recurrent currents are expressed as:
 
-[TODO MATH : prove that it is indeed Lipschitz.
-The main problem I've enountered with this is that these functions are $\textrm{regular function} \circ \textrm{CDF}^{-1}$, and the inverse CDF goes to infinity at 0 and 1].
+$$
+I^\text{rec}_U(\vec v) = w_U(\vec v, \vec u) \phi(h_U(t, \vec u)),
+$$
+
+and that in @sec-cyling-nf, numerical simulations showed that $I^\text{rec}_U(\vec v)$ is a well-behaved function.
+
+<!-- [TODO MATH : prove that it is indeed Lipschitz.
+The main problem I've encountered with this is that these functions are $\textrm{regular function} \circ \textrm{CDF}^{-1}$, and the inverse CDF goes to infinity at 0 and 1].
+-->
+
+Applying the Lipschitz property yields the following inequality on the integrand:
 
@@
 \begin{aligned}
-|\mathrm{NI}_n - \mathrm{AI}| &\leq L(\cdot) \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \lambda \circ S^{-1}(\d \beta) \right| \\
-&\leq L(\cdot) \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \sup_{\beta \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \lambda \circ S^{-1}(\d \beta^\prime) \right| \\
+\Delta(\beta) &= w_U(\cdot, S^{-1}(\beta_i)) \phi(h_U(S^{-1}(\beta_i), t)) - w_U(\cdot, S^{-1}(\beta)) \phi(h_U(S^{-1}(\beta), t)) \\
+&\leq L(\cdot) \norm{S^{-1}(\beta_i) - S^{-1}(\beta)}
+\end{aligned}
+$$
+
+We now make use of our intuition of "locality": points close in $[0,1]$ should be (on average) mapped to points close in $[0,1]^2$.
+Therefore, taking the sup inside each $2^{-n}$ segment does not introduce a too large error.
+
+$$
+\begin{aligned}
+|\mathrm{NI}_n - \mathrm{AI}| &\leq L(\cdot) \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \left[\lambda \circ S^{-1}\right](\d \beta) \right| \\
+&\leq L(\cdot) \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \sup_{\beta \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \left[\lambda \circ S^{-1}\right](\d \beta^\prime) \right| \\
 &= L(\cdot) \frac{1}{2^n} \sum_{i=1}^{2^n} \sup_{\beta \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \\
@@
 
-ironing out details : here we considered $[0,1]^2 \mapsto [0,1]$, but the mappings can be generalized to any number of dimensions.
+We finish this thesis by tying up the loose ends.
+In the abstract, we wrote that any $p$-dimensional neural field equation is equivalent to a neural field equation in $[0,1]$, but we only gave demonstrations for $[0,1]^2$ to $[0,1]$.
+@sec-simulations-3d-2d-1d showed that it is possible to iteratively apply the mapping $Z : [0,1]^2 \mapsto [0,1]$ to reduce dimensionality of the neural field, but it might be more practical to have a direct mapping $Z_p : [0,1]^p \mapsto [0,1]$.
 
-Example for the Z-mapping. Let $\vec \alpha = (\alpha_1, \cdots, \alpha_p)$, and $\alpha_\mu = \sum_{k=1}^{n} b^\mu_k 2^{-k}$. Then
+Let us give a generalization of the Z-mapping.
+Defining $\vec{v^{(n)}} = (v^{(n)}_1, \cdots, v^{(n)}_p)$, and the finite binary expansions $v^{(n)}_\mu = \sum_{k=1}^{n} b^\mu_k 2^{-k}$, we can define $Z^n_p$ as
 
 $$
-Z^n(\alpha_1, \cdots, \alpha_p) = \sum_{k=1}^{n} \sum_{\mu=1}^p b^{\mu}_k 2^{-(p(k-1)+\mu)}
+\begin{aligned}
+Z_p^n(\vec{v^{(n)}}) &= \sum_{k=1}^{n} \sum_{\mu=1}^p b^{\mu}_k 2^{-(p(k-1)+\mu)} \\
+&= 0.b^1_1 b^2_1 \cdots b^p_1 b^1_2 b^2_2 \cdots b^p_{n-1} b^1_n b^2_n \cdots b^p_n
+\end{aligned}
 $$
 
-also should work with any neural field, here low-rank cycling was just for the purpose of demonstration
+The same arguments apply to show that this generalized Z-mapping converges pointwise to $Z_p$, has the locality property, and that $V_n(Z_p) \xrightarrow{n\to\infty} 0$.
+Then the proof of convergence, albeit more laborious, can be repeated using the same procedure.

### 2023-06-15 manuscript/index.qmd
id rather not lose my changes now lmao
pwoofwead !!!1

@@
-# Index {.unnumbered .hidden}
+::: {.content-visible when-format="html"}
+::: {.content-hidden}
+# Home {.unnumbered}
+:::
 
-\[TODO : a nice illustration for this, maybe some cool animation ?\]
+![Smooth neural fields in $p$ dimensions can be mapped to equivalent neural fields in one dimension, and the dynamics are equivalent.](chapters/figures/photo_2023-06-15_13-51-39.jpg)
+:::

### 2023-06-16 manuscript/appendices/B_zcurve.qmd
dkfjwen fhjk !!!!!

@@
-::: content-hidden
+::: {.content-hidden}
 {{< include ../macros.tex >}}
 \def\C{\mathcal{C}}
+\def\P{\mathcal{P}}
+\def\B{\mathcal{B}}
 :::
 
-# On the quasi-bijectivity and continuity of the Z-mapping
+# The Z-curve is not injective {#sec-zcurve-noninjective}
 
-We study the Z-curve $Z^{-1}$
+Let us study, in a bit more detail, the properties of the Z-curve $Z^{-1}$ introduced in @sec-local, and, by extension, the properties of the Z-mapping $Z$.
 
-## The Z-curve is not injective {#sec-zcurve-noninjective}
+## Ambiguity of binary expansions
 
-this is common misconception, and people write it is a bijection, when really it is not
+Although it might seem that the Z-curve defines a bijection (this is often falsely relayed in online resources, and was even initially believed by Cantor when he discovered this function, before being corrected by Dedekind @fleron1994cantorhistory), this is not the case.
 
-The Z-curve : [0,1] -> [0,1]^2 is not injective since the binary expansion is ambiguous
+The problem with the definition is the ambiguity of (binary) expansions. If we allow infinite sum expansions, then the two representations of $\tfrac 12$ are equivalent:
 
-x = 0.10000... = 0.011111...
+$$
+\begin{aligned}
+\frac 12 &= 0.1000 \cdots = \sum_{l=1}^\infty b_l 2^{-l},\ \text{where}\ b_1=1, b_{l>1}=0 \\
+&= 0.0111 \cdots = \sum_{l=1}^\infty b^\prime_l 2^{-l},\ \text{where}\ b^\prime_1=0, b^\prime_{l>1}=1
+\end{aligned}
+$$
 
-=> alpha = 0.1y0y0y0  !=   alpha = 0.0y1y1y1y1
+Although $Z^n$ are still bijections, this ambiguity is problematic for the $n\to\infty$ limit, since we can find multiple pre-images for the same $\alpha$ mapped by $Z$.
+This illustrate this, let us take a finite integer $m$, and write a counter-example of bijectivity.
+Similarly to the main text, we define the binary expansions of $v_1$ and $v_2$:
 
-but the point on which the expansion is ambiguous form a set of zero measure, so the integral is still ok
+$$
+\begin{aligned}
+v_1 &= \sum_{l=1}^\infty b^1_l 2^{-l} = 0.b^1_1b^1_2 \cdots b^1_{m} b^1_{m+1} \cdots \\
+v_2 &= \sum_{l=1}^\infty b^2_l 2^{-l} = 0.b^2_1b^2_2 \cdots b^2_{m} b^2_{m+1} \cdots
+\end{aligned}
+$$
 
-on the set [0,1] \ {problems}, Z-curve is bijective
+We now consider the $l > m$ binary digits of $v_1$ to give ambiguous representations of $v_1^{(m)} + 2^{-m-1}$, where $v_1^{(m)}$ is the finite-$m$ truncation of $v_1$.
 
-injectivity is the property that f(x) != f(y)  iff  x != y,  (equiv f(x) == f(y) iff x == y)
+$$
+\begin{aligned}
+v_1 &= 0.b^1_1b^1_2 \cdots b^1_{m} 1000 \cdots \\&\qquad\implies \alpha = 0.b^1_1 b^2_1 \cdots b^1_m b^2_m 1 b^2_{m+1} 0 b^2_{m+2} 0 b^2_{m+3} \cdots \\
+&= 0.b^1_1b^1_2 \cdots b^1_{m} 0111 \cdots \\&\qquad\implies \alpha^\prime = 0.b^1_1 b^2_1 \cdots b^1_m b^2_m 0 b^2_{m+1} 1 b^2_{m+2} 1 b^2_{m+3} \cdots\\
+\end{aligned}
+$$
 
-Z^{-1} defines a space-filling curve
+This shows the injectivity is broken, the same point $(v_1, v_2)$ can be mapped to by (at least) two different $\alpha \neq \alpha^\prime$.
+The Z-curve $Z^{-1}$ therefore no longer defines a one-to-one mapping, and is not injective.
 
-not a problem because we are just repeating the same population twice
+For a point $(v_1, v_2) \in [0,1]^2$ which can be written as a finite binary expansion (that is, a $1$ followed by an infinite string of $0$'s), there might as much as 4 pre-images in $[0,1]$:
 
+$$
+\begin{aligned}
+(v_1, v_2) &= (0.b^1_1 \cdots b^1_{m_1} 1000 \cdots, 0.b^2_1 \cdots b^2_{m_2} 1000 \cdots) \\
+&= (0.b^1_1 \cdots b^1_{m_1} 1000 \cdots, 0.b^2_1 \cdots b^2_{m_2} 0111 \cdots) \\
+&= (0.b^1_1 \cdots b^1_{m_1} 0111 \cdots, 0.b^2_1 \cdots b^2_{m_2} 1000 \cdots) \\
+&= (0.b^1_1 \cdots b^1_{m_1} 0111 \cdots, 0.b^2_1 \cdots b^2_{m_2} 0111 \cdots) \\
+\end{aligned}
+$$
+
+has the following pre-images (supposing without loss of generality $m_1 \leq m_2$):
+
+$$
+\begin{aligned}
+\alpha_1 &= 0.b^1_1 b^2_1 \cdots b^1_{m_1} b^2_{m_1} 1 b^2_{m_1+1} 0 b^2_{m_1+2} 0 \cdots b^2_{m_2} 010000 \cdots \\
+\alpha_2 &= 0.b^1_1 b^2_1 \cdots b^1_{m_1} b^2_{m_1} 1 b^2_{m_1+1} 0 b^2_{m_1+2} 0 \cdots b^2_{m_2} 010101 \cdots \\
+\alpha_3 &= 0.b^1_1 b^2_1 \cdots b^1_{m_1} b^2_{m_1} 0 b^2_{m_1+1} 1 b^2_{m_1+2} 1 \cdots b^2_{m_2} 101010 \cdots \\
+\alpha_4 &= 0.b^1_1 b^2_1 \cdots b^1_{m_1} b^2_{m_1} 0 b^2_{m_1+1} 1 b^2_{m_1+2} 1 \cdots b^2_{m_2} 101111 \cdots
+\end{aligned}
+$$
+
+## The set of points with multiples preimages in $[0,1]$ is dense in $[0,1]^2$
+
+The Z-curve is not injective from $[0,1]$ into $[0,1]^2$, and furthermore, we cannot ignore the points for which it is not injective, because the set of points with multiple preimages in $[0,1]$ is dense in $[0,1]^2$.
+We begin the study by considering only the first coordinate $v_1$, since the same arguments apply to $v_2$.
+
+The set $\P$ is defined as the set of $v_1$ which have multiple pre-images in $[0,1]$:
+
+$$
+\P = \{ 0.b_1 b_2 \cdots b_m 1000 \cdots \ |\ m \in \mathbb{N} \}.
+$$ {#eq-P}
+
+We wish to prove that $\P$ is dense in $[0,1]$.
+For this, we will prove that any point $v_1 \in [0,1]$ is either in $\P$, or is the limit point of $\P$.
+If $v_1 \in \P$, we are done.
+Else, we consider $v_1 \notin \P$.
+
+We define an expansion of $v_1 = \sum_{l=1}^\infty b_l 2^{-l}$ (we note $b_l$ are not unique, but this does not matter for the rest of this proof).
+We also define the truncation $v_1^{(m)} = \sum_{l=1}^m b_l 2^{-l}$, such that $v_1^{(m)} \xrightarrow{m\to\infty} v_1$.
+The trick is now to define the sequence of $v_1^\prime{^{(m)}}$ as:
+
+$$
+v_1^\prime{^{(m)}} = v_1^{(m)} + \frac{1}{2^{m+1}} \in \P.
+$$
+
+Therefore, the sequence $\{v_1^\prime{^{(m)}}\}$ is a subset of $\P$, and its limit is:
+
+$$
+\lim_{m\to\infty} v_1^\prime{^{(m)}} = \lim_{m\to\infty} v_1^{(m)} + \lim_{m\to\infty} \frac{1}{2^{m+1}} = v_1.
+$$
+
+Therefore, $v_1$ is a limit point of $\P$. This concludes the proof that $\P$ is dense in $[0,1]$.
+
+By extension, the same arguments apply to $v_2$, and the set $\P \times \P$ is dense in $[0,1]\times [0,1]$. Therefore, the set of points with multiple preimages in $[0,1]$ is dense in $[0,1]^2$.
+
+<!-- 
+For this, we adopt a counting argument and look at the scaling behavior of $\P_n$, the finite-$n$ truncations of $\P$.
+
+Let us take $n > m$, and the finite-$n$ truncations $v_1^{(n)}$ of $v_1 \in \P$.
+Because of the ambiguity of representation, we need to consider two different finite-$n$ truncations, that both converge to $v_1$ as $n\to\infty$.
+This defines the sequence of sets $\P_n$ as:
+
+$$
+\P_n = \{ 0.\underbrace{b_1 b_2 \cdots b_m}_{m\ \text{bits}} \underbrace{1000 \cdots 0}_{n-m\ \text{bits}} \ |\ m < n \} \cup \{b_1 b_2 \cdots b_m 0111\cdots 1\ |\ m < n\}.
+$$ {#eq-Pn}
+
+At finite $n$, we can define the complement, that is, the set of non-problematic points as:
+
+$$
+\begin{aligned}
+\P_n^c &= \B_n \backslash \P_n \\
+\text{where}\ \B_n &= \left\{ \frac{k}{2^n} \ |\ k \in \{0, \cdots, 2^{n}-1\}\right\}
+\end{aligned}
+$$ {#eq-Pnc}
+
+We now argue based on the size of $\P$ and $\P^c$.
+From its definition, we directly see that $|\B_n| = 2^n$.
+To compute $|\P_n|$, it is easier to rewrite $\P_n$ as the union of $m$ *disjoint* sets $\P_{n,m}$:
+
+$$
+\begin{aligned}
+\P_n &= \bigcup_{m=1}^n \P_{n,m} \\
+&= \bigcup_{m=1}^n \left\{ \frac{k}{2^m} + \frac{1}{2^{m+1}}, \frac{k}{2^m} + \frac{2^{m+1}-1}{2^{m+1}}\ |\ k \in \{0, \cdots, 2^{m} - 1\},  \right\}.
+\end{aligned}
+$$
+
+Since each set $\P_{n,m}$ is of size $|\P_{n,m}| = 2 \times 2^{m}$, the size of $|\P_n|$ can be computed as:
+
+$$
+|\P_n| = \sum_{m=1}^n |\P_{n,m}| = 2 \sum_{m=1}^n 2^m = 2 (2^n - 1)
+$$
+
+uh oh... -->
+<!-- 
 ## $Z^{-1}$ defines a space-filling curve on a subset of $[0,1]$ {#sec-zcurve-lebesgue}
@@
 talk about the fact we have other space filling curves, such as hilbert curve or peano curve, that are continuous and surjective [0,1] -> [0,1]^2.
-we use the Z-curve because it's easy to implement numerically
+we use the Z-curve because it's easy to implement numerically -->

### 2023-06-16 manuscript/chapters/3_mappings.qmd
dkfjwen fhjk !!!!!

@@
 They have the nice property that they are continuous and surjective, so we the segment populations can "reach" the entirety of the square populations.
-All things considered, non-injectivity is not a problem, since it just means that two segment populations can "refer" to the same square population, which does not break the regularity of the kernel, since the two segment populations can just have their own neighborhoods.
-Furthermore, we will see that in our examples, the non-injectivity can be ignored because the mapping is "quasi-bijective".
+All things considered, non-injectivity is not a problem, since it just means that two segment populations can "refer" to the same square population (see @fig-non-injection), which does not break the regularity of the kernel, since the two segment populations can just have their own neighbourhoods.
 
-![Non-injectivity of $S^{-1}$ means that two segment populations can refer to the same square population.](figures/photo_2023-06-12_00-46-26.jpg)
+![Non-injectivity of $S^{-1}$ means that two segment populations can refer to the same square population.](figures/photo_2023-06-12_00-46-26.jpg){#fig-non-injection}
 
@@
 In essence, the dynamics in the 2D embedding are already known, and we ask the question of whether it is possible to write an equivalent neural field 1D (in the sense that it has identical latent trajectories).
-Without repeating the intuition given in @sec-intro-simnf, if neighboring populations in 1D have similar potentials, then this allows us to write a neural field equation in $[0,1]$.
+Without repeating the intuition given in @sec-intro-simnf, if neighbouring populations in 1D have similar potentials, then this allows us to write a neural field equation in $[0,1]$.
 Therefore we would like to have the property that populations close in 1D are (inversely) mapped to populations close in 2D, which already have similar potentials.
@@
 @fig-anti-z gives some insight into the problem of this mapping.
-For a fixed input $(v_1,v_2)$, we see that the output point position jumps around when $n$ is changed, and that these jumps seem to always be of order 1.
+For a fixed input $(v_1,v_2)$, we see that the output point position jumps around when $n$ changes, and that these jumps seem to always be of order 1.
 The intuition behind pointwise convergence is that as $n$ increases, the image of a point by $S^n$ is continually refined.
@@
 Numerically, there seems to be an infinite number of discontinuities, but the size of these discontinuities is overwhelmingly small, so that *on average* they vanish as $n\to\infty$.
-We won't go further into detail about this in this section (in @sec-zcurve-lebesgue we develop the convergence to a space-filling fractal curve), but this gives some intuition for the arguments presented in @sec-locality.
+We won't go further into detail about this (we refer to @Sagan1994 for an in-depth study), but this gives some intuition for the arguments presented in @sec-locality.
 
@@
 [^zmapping-notbijective]: Despite what can often be found in online resources, the Z-mapping is *not* a bijection because that would violate Netto's theorem.
+For every point in $[0,1]^2$, there are at most 4 distinct preimages in $[0,1]$.
 The reason for this is the ambiguity in the writing of binary expansions, and in @sec-zcurve-noninjective we give proof that the Z-mapping is not injective in the $n\to\infty$ limit.
-However, the Z-mapping is *quasi-bijective*, in the sense that the set of points for which the Z-mapping is not injective has zero measure, and can therefore be ignored.
 

### 2023-06-18 manuscript/chapters/1_intro.qmd
valentin proofread and comments

@@
 
-Early recordings of the cat's somatosensory visual cortices suggested that the cortical sheet is organized in vertical columns of functionally similar neurons @Mou57.
+Early recordings of cat's somatosensory @Mou57 and visual @HubWie62 cortices suggested that the cortical sheet is organized in vertical columns of functionally similar neurons.
 
@@
 
-As models of spatiotemporal neural population dynamics on the cortical sheet, neural field models can be used to explain experimentally observed patterns of cortical activity, such as travelling waves in visual [@SatNau12;@MulRey14], somatosensory [@PetHah03;@FerBol06], motor [@RubRob06;@TakKim15], and hippocampal [@LubSia09;@PatFuj12;@PatSch13] cortices (see @MulCha18 for a review), and are used to model large-scale brain signals, such as electroencephalography (EEG) recordings @Bre17.
-See also @DipRan18 for an example of neural field modelling of calcium imaging recording of the visual cortex.
+As models of spatiotemporal neural population dynamics on the cortical sheet, neural field models can be used to explain experimentally observed patterns of cortical activity, such as travelling waves in visual [@SatNau12;@MulRey14], somatosensory [@PetHah03;@FerBol06], motor [@RubRob06;@TakKim15], and hippocampal [@LubSia09;@PatFuj12;@PatSch13] cortices (see @MulCha18 for a review), and are used to model large-scale brain signals, such as electroencephalography (EEG) recordings @Bre17 (see also @DipRan18 for an example of neural field modelling of calcium imaging recording of the visual cortex).
 
@@
 The goal of this Master Thesis is to show that the answer to the second question is highly nontrivial because even the dimensionality of the embedding space, for some given spatiotemporal dynamics, is not clearly defined.
-This means that it is possible to find two embeddings (and connectivity kernels) of different dimensions that give rise to identical dynamics; *contradicting the intuition that the embedding dimension should at least be as high as the dimensionality of the dynamics* (the number of variables needed to fully describe the dynamical system).
+This means that it is possible to find two embeddings (and connectivity kernels) of different dimensions that give rise to identical dynamics.
+<!-- ; *contradicting the intuition that the embedding dimension should at least be as high as the dimensionality of the dynamics* (the number of variables needed to fully describe the dynamical system).
+-->
 
-In particular, we show that for any two-dimensional neural field equation (that is, with two spatial dimensions), there is a one-dimensional neural field equation on the interval, where the embedding space is simply the interval $[0,1]$, from which the solution to the two-dimensional equation can be fully reconstructed.
+In particular, we show that for any two-dimensional neural field equation (that is, with two spatial dimensions), there is a one-dimensional neural field equation, where the embedding space is simply the interval $[0,1]$, from which the solution to the two-dimensional equation can be fully reconstructed.
 The mapping from the two-dimensional equation to the one-dimensional equation is done via well-known measurable bijections between the square $[0,1]^2$ and the interval $[0,1]$.
 Importantly, these bijections are not (and can't be) diffeomorphisms.
-Also, we show that although the connectivity kernel of the one-dimensional equation seems to have a fractal structure, it is sufficiently regular for its solution to be numerically approximated using standard mesh-based simulations.
+Also, we show that although the connectivity kernel of the one-dimensional equation seems to have a fractal structure, it is sufficiently regular for its solution to be numerically approximated using standard grid-based simulations.
 
@@
 
-Let us now give a short introduction to the mathematical formulation of neural fields.
-Neural field equations emerge from the structure of networks of rate neurons when the number of neurons $N$ becomes large.
-Positions $\vec z_i \in \Rp$ are attributed to each neuron, which defines the *embedding space*.
-Each neuron has a potential $h_i(t)$ that evolves in time, and the weight with which the neurons interact with each other is given by the components of the connectivity matrix $J_{ij} \in \R^{N \times N}$.
+Let us now give a short introduction to the mathematical theory of neural fields.
+Neural field equations emerge from the spatial structure of networks of rate neurons (or Poisson spiking neurons [@Ger95;@CheDua19]) when the number of neurons $N$ becomes large.
+Positions $\vec z_i \in \Rp$ are attributed to each neuron, where $\Rp$ is the *embedding space*.
+Each neuron has a potential $h_i(t)$ that evolves in time, and the weight with which the neurons interact with each other is given by the entries of the connectivity matrix $J_{ij} \in \R^{N \times N}$.
 Additionally, the activation function $\phi(h)$ describes the firing rate of a given neuron with potential $h$.
@@
 Before we write down the neural field equation which describes how the field evolves in time, we need to introduce the neuron density $\rho(\vec z)$, which measures the number of neurons per unit volume of the embedding space.
-The neural field equation is then written as such:
+The neural field equation writes:
 
@@
 
-![(TODO add phi, add z instead of xi) The connectivity kernel defines the influence between neural densities in the embedding space](figures/connectivity_kernel.png)
+![(TODO add phi, add z instead of xi) The connectivity kernel defines the influence between localized neural populations in the embedding space](figures/connectivity_kernel.png)
 
@@
 This work was inspired by recent mathematical development in mean-field theories involving the theory of graphons @JabPoy21.
-Their main result is that under some boundedness assumptions of the connectivity between the components of a system, the equation of evolution incorporates only an integral over $[0,1]$.
+Informally, these recent results suggest that, even in the absence of any explicit spatial structure, as long as weights scale as $1/N$, the dynamics of large networks can be described, as $N \to \infty$, by a neural field equation where the embedding is simply the interval $[0,1]$, which motivates the search for a formulation of a one-dimensional neural embedding.
+<!-- Their main result is that under some boundedness assumptions of the connectivity between the components of a system, the equation of evolution incorporates only an integral over $[0,1]$.
 This is indeed the case in the toy model described next chapter, where weights scale as $J_{ij} \sim \frac 1N$, which motivates the search for a formulation of a one-dimensional neural embedding.
+-->
 
-The main idea is to define a measurable bijective mapping $S : \Rp \mapsto [0,1]$ and a measure $\lambda : [0,1]^2 \mapsto \mathbb{R}^+$, such that we can write a one-dimensional neural field	$\tilde h(\alpha, t)$ with a connectivity kernel loosely defined as $\tilde w = w \circ S^{-1}$.
+The main idea is to define a measurable bijective mapping $S : \Rp \mapsto [0,1]$ and a measure $\lambda : \Rp \mapsto \mathbb{R}^+$, such that we can write a one-dimensional neural field	$\tilde h(\alpha, t)$ with a connectivity kernel loosely defined as $\tilde w = w \circ S^{-1}$.
 
@@
 
-This change of variable in the kernel seems counter-intuitive: how is it possible to map a higher dimension to a lower dimension without losing information?
-And even if mathematically such a change of variables is possible, is it even feasible to simulate it numerically?
+This change of variable in the kernel seems counter-intuitive: how is it possible to map dynamics in the embedding space $\Rp$ to dynamics in the lower dimensional embedding space $[0,1]$ without losing information?
+And even if mathematically such a change of variables is possible, is it even feasible to numerically simulate the dynamics in the lower dimensional embedding space?
 
@@
 A neuron-based simulation of $N$ neurons.
-Each neuron samples the probability distribution $\rho(z_1, \cdots, z_p)$ independantly.
+The location of each neuron is an independent sample from the distribution $\rho(z_1, \cdots, z_p)$.
 This case then just reduces to simulating the network of rate neurons.
@@
 Suppose we take bins in the $\Rp$ space, and map each bin to a random position in $[0,1]$.
-Intuitively, this won't work (in the sense that the dynamics in $[0,1]$ will not be the same as the dynamics in $\Rp$), because the hypothesis of continuity is not longer true: the mapping is random, so the nearby populations in $[0,1]$ will no longer have similar potentials, and the error of the mean-field approximation will no longer vanish as we take increasingly many bins.
+Intuitively, this won't work (in the sense that the dynamics in $[0,1]$ will not be the same as the dynamics in $\Rp$), because the hypothesis of continuity is not longer true: the mapping is random, so the nearby neurons in $[0,1]$ will no longer have similar potentials, and the error of the mean-field approximation will no longer vanish as we take increasingly many (and smaller) bins.
+<!-- NOTE : was "populations" before, this is explained later in the thesis. but OK, for the introduction, maybe we should use neurons
+the mapping is random, so the nearby populations in $[0,1]$ -->
 

### 2023-06-18 manuscript/chapters/2_neural_fields.qmd
dkfjwen fhjk !!!!!
valentin proofread and comments

@@
 
-The toy model we introduce is an instance of a low-rank rate neural network.
+The toy model we introduce is an instance of a low-rank neural network.
 Such models hold their name from the form of their connectivity matrix, which has a rank of (at most) $p$.
@@
 
-The geometric view of these networks is that the recurrent currents lie in a $p$-dimensional subspace of $\R^N$ spanned by the vectors $\{\vec{F_1}, \cdots, \vec{F_p}\}$ (where $\vec{F_\mu}=(F_{\mu 1}, \cdots F_{\mu N})$), which therefore form a "natural" embedding.
+The geometric view of these networks is that the recurrent currents lie in a $p$-dimensional subspace of $\R^N$ spanned by the vectors $\{\vec{F_1}, \cdots, \vec{F_p}\}$ (where $\vec{F_\mu}=(F_{\mu 1}, \cdots F_{\mu N})$), which therefore define a "natural" linear embedding (in the sense of @JazOst21) of the neural population activity.
 
-Recalling the notation from the introduction, $h_i(t)$ are the neuron potentials (with initial condition $h_i(0)$), $\phi : \R \mapsto \R^+$ is the monotonic increasing activation function, and we write the evolution equation for this network of neurons:
+Recalling the notation from the introduction, the $h_i(t)$ are neuron potentials (with initial condition $h_i(0)$), $\phi : \R \mapsto \R^+$ is the monotonic increasing activation function, and we write the evolution equation for this network of neurons:
 
@@
 
-The exponential decay term describes how, in isolation, a neuron's potential tends to zero, biologically corresponding to a depolarized membrane.
+The exponential decay term describes how, in isolation, a neuron's potential tends to zero, biologically corresponding to the decay of the membrane potential to its value at rest.
 The recurrent current term $I^\text{rec}_i(t)$ describes the "currents" received by neuron $i$ from all the other neurons in the network.
@@
 
-One last remark is that in our model, we omit to add any source of external currents: from the initial condition, the network is left to evolve in isolation.
+One last remark is that in our model, there is no external current; we only study the autonomous dynamics of the system.
 
@@
 
-Let us now specify the low-rank model introduced in the previous section, and use the model introduced in @schmutz2023convergence.
+Let us now specify the low-rank model introduced in the previous section, and use the model introduced in @schmutz2023convergence, which can be seen as a simple example of a "Gaussian mixture low-rank network" @Beiran2021.
 This paper defines the low-rank vectors $F_{\mu i}$ such that each component independently samples a standard Gaussian (zero mean, unit variance).
-In other words, every neuron $i$ samples a vector $\vec{F_i} = (F_{1i}, \cdots, F_{pi})$ from the $p$-dimensional gaussian ball.
+In other words, every neuron $i$ samples a vector $\vec{F_i} = (F_{1i}, \cdots, F_{pi})$ from the $p$-dimensional gaussian.
 We write
@@
 $$
-G_{\mu i} = \tilde\phi(z_{\mu i}) \stackrel{\text{def}}{=} \frac{\phi(z_{\mu i}) - \avg{z_{\mu i}}}{\mathrm{Var}[\phi(z_{\mu i})]}.
+G_{\mu i} = \tilde\phi(z_{\mu i}) \stackrel{\text{def}}{=} \frac{\phi(z_{\mu i}) - \avg{\phi(z_{\mu i})}}{\mathrm{Var}[\phi(z_{\mu i})]}.
 $$
 
-This choice be elucidated in a few paragraphs.
+The motivation for this choice will be presented in a few paragraphs.
 Finally, the activation function remains to be defined.
-For simplicity, we take it to be the logistic function, although the following results are not sensitive to this choice.
+For simplicity, we take it to be the logistic function, although the following results are not sensitive to this choice,
 
 $$
-\phi(h) = \frac{1}{1 + \mathrm{e}^h}
+\phi(h) = \frac{1}{1 + \mathrm{e}^h}.
 $$
@@
 $$
-\dot h_i(t) = -h_i(t) + \frac 1N \sum_{\mu=1}^p \sum_{j=1}^{N} z_{\mu i} \tilde \phi(z_{\mu j}) \phi(h_j(t)).
+\dot h_i(t) = -h_i(t) + \frac 1N \sum_{\mu=1}^p \sum_{j=1}^{N} z_{\mu i} \tilde \phi(z_{\mu j}) \phi(h_j(t)) =: \mathrm{RHS}(h_1, \cdots, h_N).
 $$ {#eq-toymodel-evolution}
@@
 
+<!--
 $$
 \dot h_i(t)\ \text{evaluated at}\ h_i^\star = -h^\star_i(t) + \frac 1N \sum_{\mu=1}^p \sum_{j=1}^{N} z_{\mu i} \tilde \phi(z_{\mu j}) \phi(h^\star_j(t)) = 0
+$$ 
+-->
+
+$$
+\mathrm{RHS}(h^\star_1, \cdots, h^\star_N) = -h^\star_i + \frac 1N \sum_{\mu=1}^p \sum_{j=1}^{N} z_{\mu i} \tilde \phi(z_{\mu j}) \phi(h^\star_j) = 0
 $$
@@
 In the introduction, we introduced the convergence of a network of neurons to a smooth neural field when the embedding space is well-chosen.
-For our toy model, it seems natural to try to embed the neurons according to their positions on the $p$-dimensional Gaussian ball.
+For our toy model, it seems natural to try to embed the neurons according to their positions on the $p$-dimensional Gaussian.
 
@@
 
-The results of the simulation show that the random ordering of the neurons does not yield a smooth embedding, on contrary of the smooth surface that appears when we order the neurons by their position $z_{1i}$.
+The results of the simulation show that the random ordering of the neurons does not yield a smooth embedding, in contrast to the smooth surface that appears when we order the neurons by their position $z_{1i}$.
 The simulation serves as a demonstration that the neurons converge to the stable fixed point $h^\star_i = z_{1i}$, as well as confirming that the zero fixed point is unstable.
@@
 
-### Sampling method (Monte Carlo integration)
+### Sampling method (Monte Carlo integration) {#sec-mc}
 
 Due to the formulation of the neural field weighed by a probability distribution $\rho(y_1, \cdots, y_p)$, a natural way to estimate the integral is by Monte Carlo integration.
-The essence of this method is the use of the Central Limit Theorem, where we take $N$ independent samples $\vec{y_i} = (y_{1i}, \cdots, y_{pi})$ from the distribution $\rho$, and use them to estimate the integrand.
+The essence of this method is to use the Central Limit Theorem, where we take $N$ independent samples $\vec{y_i} = (y_{1i}, \cdots, y_{pi})$ from the distribution $\rho$, and use them to estimate the integrand.
 
@@
 
-The attentive reader might have noticed by now that the Monte-Carlo method just reduces to simulating the network of neurons @eq-toymodel-evolution.
+<!-- The attentive reader might have noticed by now that the Monte-Carlo method just reduces to simulating the network of neurons @eq-toymodel-evolution.
 Effectively, the network of neurons converges to the neural field with rate $\bO(1/\sqrt{N})$.
@@
 One small nuance we should add, is that strictly speaking, Monte-Carlo integration should resample the integrand at every timestep, but the network of neurons samples the neurons once, and propagates these points through time.
+-->
 
@@
 
-Overlap variables were introduced by @Gerstner_1992 in the context of the study of the Hopfield network, which can be seen as a low-rank network with discrete variables instead of our continuous $\vec{z}$.
+Overlap variables in continuous time variables were introduced by @Gerstner_1992 to study spiking networks models of associative memory (Hopfield-type networks), which can be seen as a low-rank network with discrete spatial structure instead of our continuous $\vec{z}$.
 The overlap variables measure the correlation between the state of the network (the value of the potential $h(\vec z, t)$ at each point in space) and the pattern fixed point.
@@
 
-We cite @veltz2009localglobal, section 4.3 for the reduction of general low-rank networks to $p$-dimensional closed systems, and give the equations applied to our toy model.
+We refer to @veltz2009localglobal, section 4.3 for the reduction of general low-rank networks to $p$-dimensional closed systems, and give the equations applied to our toy model.
 Defining the projections $\kappa_\mu(t)$ of the neural field onto the patterns
@@
 
-In @sec-networks-of-neurons-to-neural-fields we introduced a simple toy model a neural field with a natural embedding on the $p$-dimensional Gaussian ball.
+In @sec-networks-of-neurons-to-neural-fields we introduced a simple toy model of a neural field with a natural embedding on the $p$-dimensional Gaussian.
 Analytical derivations accompanied by numerical simulations showed the dynamics of this model can be summarized by the convergence to pattern fixed points.
@@
 We define "rolling" as the response of a neuron to shift towards the "next" pattern given its state matching a current pattern.
-This is done by adding a shift $\mu+1$ in the connectivity kernel, with the convention $p+1=1$.
+This is done by adding a shift $\mu+1$ in the connectivity kernel, with the convention $p+1=1$ (see @Ami89, Chapter 5).
 
@@
 
-![(TODO animation in 3D, rename non-cycling to original) Intuition for the cycling behavior by considering the recurrent currents](figures/cycling_recurrent.png){#fig-cycling-irec}
+![(TODO animation in 3D, rename non-cycling to original, number from one) Intuition for the cycling behavior by considering the recurrent currents](figures/cycling_recurrent.png){#fig-cycling-irec}
 
@@
 
-With these considerations, we simulate a $p=2$ cycling neural field with $\delta=6$ and initial condition $h(z_1, z_2, t < 0) = z_1$ in @fig-cycling-anim.
+With these considerations, we simulate a $p=2$ cycling neural field with $\delta=6$ (since the membrane time constant has been fixed to one, this means that $\delta = 6\tau = 6$) and initial condition $h(z_1, z_2, t < 0) = z_1$ in @fig-cycling-anim.
 The latent trajectory is estimated from the simulation results, and animated simultaneously with the neural field.
@@
 
-This system of equations can be solved by steps of $\delta$, iteratively plugging in the previous solution to solve each cycle.
+<!-- This system of equations can be solved by steps of $\delta$, iteratively plugging in the previous solution to solve each cycle. -->
 
@@
 
-Defining the change of variables $v_\mu=\CDF(z_\mu), u_\mu=\CDF(y_\mu)$, we can define the uniform neural field and uniform connectivity kernel as
+Defining the change of variables $v_\mu=\CDF(z_\mu), u_\mu=\CDF(y_\mu)$, we can define the neural field $h_U$ and the connectivity kernel $w_U$ for the uniform distribution as
 

### 2023-06-18 manuscript/chapters/3_mappings.qmd
valentin proofread and comments

@@
 
-Let us for a moment consider another type of model for networks of neurons, that of homogeneous population dynamics, that describes how $N_\text{pop}$ different populations interact with each other.
-These types of models emerge when networks of neurons exhibit a finite number of "neuron types".
+Let us for a moment consider another type of model for networks of neurons, that of multiple homogeneous population dynamics, describing how $N_\text{pop}$ different populations interact with each other.
+These types of models emerge, for example, when networks of neurons exhibit a finite number of "neuron types".
 We then group neurons of the same type, and make the approximation that they have identical potentials and the same interaction weights with other neurons.
-In effect, every population then represents an infinite number of discrete identical neurons.
+As a consequence, every population then represents a homogeneous population of identical neurons.
 
@@
 We hereby showed (having taken many writing liberties, but this is just to give an idea) how the neural field reduces back to the original population dynamics, which was in itself already the $N\to\infty$ limit of the network.
-The case of population dynamics therefore shows a trivial example of an embedding in $[0,1]$.
+The case of multi-population dynamics therefore shows a trivial example of an embedding in $[0,1]$.
 
@@
 
-When taken in isolation, the previous developments might sound a little bit silly, though they serve to give an impression of the problem at hand, and an interpretation for the rest of this work.
-For a finite number of populations $N_\text{pop}$, the population dynamics is well-defined, however, what happens as $N_\text{pop} \to \infty$ ?
+When taken in isolation, the previous developments might sound a little bit silly, though they serve to give a feeling for the problem at hand, and a motivation for the rest of this work.
+For a finite number of populations $N_\text{pop}$, the population dynamics is well-defined; however, what happens as $N_\text{pop} \to \infty$ ?
 The core of the question lies in the fact that we can easily *enumerate* a finite number of populations, but the sum $\sum_{b=1}^{N_\text{pop}}$ must have a clearly defined interpretation when this number tends to infinity.
@@
 In light of the toy model presented in @sec-nf, we can reinterpret the density $\rho$ of the neural field equation.
-Every point $\vec z \in \R^p$ (let us consider $p=2$ for the rest) describes an infinite population of homogeneous neurons, associated with a probability (density) $\rho(\vec z)$.
+Every point $\vec z \in \R^p$ (let us consider $p=2$ in the following) describes an infinite population of homogeneous neurons, associated with a probability (density) $\rho(\vec z)$.
 Therefore, in order to write an equivalent neural field in $[0,1]$, we need a way to *enumerate* the populations.
-This is done using a mapping $S : [0,1]^2 \mapsto [0,1]$, where we traverse the $[0,1]^2$ space by following the order in the image $[0,1]$.
+This is done using a mapping $S : [0,1]^2 \mapsto [0,1]$, where we cover the $[0,1]^2$ space by following the order assigned to the populations in the image $[0,1]$.
 
@@
 As we increase $N_\text{pop}$ by taking finer and finer grids, the image of the grids defines a sequence of mappings $S^n: \text{grid in}\ [0,1]^2 \mapsto \text{grid in}\ [0,1]$, which we aim to study in the rest of this chapter.
+The main question is:
 
@@
 We start by writing the neural field equation in the $[0,1]$ space. 
-For this, we define a measurable and bijective mapping $S : [0,1]^2 \rightarrow [0, 1]$.
-Furthermore, we let $\lambda : [0,1]^2 \rightarrow [0, \infty]$ be a measure on the unit square (for instance, the Lebesgue measure, which plays the role of a uniform distribution).
-Taking the composition $\lambda \circ S^{-1}$ then defines a measure on $[0,1]$, which keeps track of how 2D surfaces are transformed to 1D lengths through the mapping $S$.
+For $\lambda: [0,1]^2 \to [0, \infty]$ a measurable function (e.g.
+the Lebesgue measure, which is nothing but the uniform probability distribution), we let the composition $\lambda \circ S^{-1}$ define a measure on $[0,1]$, which keeps track of how 2D surfaces are transformed to (possibly disjoint) 1D segments through the mapping $S$.
 
-The connectivity kernel $\tilde w$ defines the interactions between the populations in the $[0,1]$ embedding, which we define relating to the uniform kernel on $[0,1]^2$ as :
+The connectivity kernel $\tilde w$ defines the interactions between the populations in the $[0,1]$ embedding, which we define relating to the kernel $w_U$ on $[0,1]^2$, defined in @sec-mapping-cdf as:
 
@@
 
-The definition of the connectivity kernel in @eq-kernel-1d might seem innocent, but the key element here is the inclusion of the inverse $S^{-1}$.
+The definition of the connectivity kernel in @eq-kernel-1d might seem innocent, but the key element here is the inverse $S^{-1}$.
 We recall that in the toy model, the original connectivity kernel is written as $w(\vec z, \vec y) = \sum_{\mu=1}^p \tilde \phi(y_\mu) z_\mu$.
-This kernel is highly regular, by which we mean that by its continuity in each of its arguments, populations nearby in the embedding will "sense" similar connectivity functions.
+This kernel is highly regular, by which we mean that, by its continuity in each of its arguments, populations nearby in the $\Rp$ will have similar input weights.
 Since it is also differentiable, we can write this as a Taylor expansion in the first argument :
@@
 
-When we do the change of variables with the CDF, we don't encounter problems because the Gaussian (inverse) CDF is differentiable and its derivative is continuous[^inversecdf]; therefore $w_U(\vec v, \vec u) = w(\textrm{CDF}^{-1}(\vec v), \textrm{CDF}^{-1}(\vec u))$ is still highly regular.
+When we do the change of variables with the CDF, we don't encounter problems because the Gaussian (inverse) CDF is differentiable and its derivative is continuous;[^inversecdf] therefore $w_U(\vec v, \vec u) = w(\textrm{CDF}^{-1}(\vec v), \textrm{CDF}^{-1}(\vec u))$ is still highly regular.
 
@@
 This mapping is obviously continuous and differentiable in both components, and surjectively maps the unit square to the unit segment.
-
 However, the obvious problem with this approach is that linear projections are not invertible.
@@
 In other words, $P^{-1}(\alpha)$ defines a vertical line in the $[0,1]^2$ embedding.
-When we then take the composition $w_U \circ P^{-1}$, the regularity is destroyed, in the sense that the same population $\alpha$ in the 1D embedding will "sense" all the connectivity kernels on the vertical line.
+When we then take the composition $w_U \circ P^{-1}$, the "locality" is destroyed, in the sense that the same population $\alpha$ in the 1D embedding will "sense" all the connectivity kernels on the vertical line.
 
@@
 
-However, this again fails because the joint cumulative function is not bijective, and isolines of the CDF define diagonals in the uniform embedding space $[0,1]^2$.[^jointcdf]
+However, this again fails because the joint cumulative function is not bijective, and isolines of the CDF define continuous curves in the uniform embedding space $[0,1]^2$.[^jointcdf]
 
@@
 
-Since in the composition $w_U \circ S^{-1}$ we work with the inverse mapping, it might be judicious to directly consider functions from $[0,1]$ into $[0,1]^2$.
+Since in the composition $w_U \circ S^{-1}$ we work with the inverse mapping, we contextualize our study by first considering the types of functions from $[0,1]$ into $[0,1]^2$.
 We have already established that there is no homeomorphism between those spaces, and need to relax one of the conditions.
+Since bijections are functions that are both injective and surjective, we can represent Netto's theorem as a 3-way Venn diagram in @fig-venn.
 
+![(TODO) Venn diagram reprensenting the properties of functions $[0,1] \mapsto [0,1]^2$ which are allowed by Netto's theorem.](figures/photo_2023-06-18_21-32-45.jpg){width=400px #fig-venn}
+
+The 3-way intersection of the Venn diagram represents that functions which are both bijective (surjective and injective) and continuous do not exist.
+
+For the three remaining 2-way intersections, we cite, without going into detail, the types of functions which they represent (see [@kharazishvili_strange_2017] for Cantor and Peano functions, [@sagan_space-filling_1994] for Peano and Jordan functions):
+
+* Cantor functions are bijections $[0,1] \mapsto [0,1]^2$. However, they are not continuous, and thereby break the continuity of the kernel.
+* Jordan curves are continuous injections $[0,1] \mapsto [0,1]^2$. Since they are not surjective, these curves cannot entirely "fill" the $[0,1]^2$ space, and we will not use them.
+<!-- Nicole : see my notebook about Jordan functions and their nonzero Jordan measure -->
+* Peano curves are continuous surjections $[0,1] \mapsto [0,1]^2$.
+Their non-injectivity however is not a problem, since it just means that two segment populations can "refer" to the same square population (see @fig-non-injection).
+Therefore, Peano functions seem like good candidates for the mapping $S^{-1}$.
+
+<!-- 
 Functions of $[0,1]$ into $[0,1]^2$ are extensively discussed in chapter 1 of @kharazishvili_strange_2017, and we cite two classes of such functions :
@@
 In light of these results, we might be tempted to search for expressions of Cantor-type functions, since having a bijection would allow us to have a direct mapping between the square and the segment populations.
-Unfortunately, these functions are discontinuous which breaks the regularity of the kernel.
+Problematically, these functions are discontinuous and thereby break the continuity of the kernel.
 
 Peano functions seem like good candidates for the mapping $S^{-1}$, and are extensively studied in the literature around space-filling curves (we refer for instance to [@sagan_space-filling_1994]).
-They have the nice property that they are continuous and surjective, so we the segment populations can "reach" the entirety of the square populations.
+They have the nice property that they are continuous and surjective, so the segment populations can "reach" the entirety of the square populations.
 All things considered, non-injectivity is not a problem, since it just means that two segment populations can "refer" to the same square population (see @fig-non-injection), which does not break the regularity of the kernel,  since the two segment populations can just have their own neighbourhoods.
+-->
 
@@
 
-We define the mapping $S^n$ that takes the $n$-bit truncations $v^{(n)}_1$ and $v^{(n)}_2$ as input, and outputs a $2n$ bit truncation $\alpha^{(n)}$ of $\alpha=S(v_1, v_2)$, where $S$ is the $n\to\infty$ limit of the sequence $S^n$.
+We define the mapping $S^n$ that takes the $n$-bit truncations $v^{(n)}_1$ and $v^{(n)}_2$ as input, and outputs a $2n$ bit truncation $\alpha^{(n)}$ of $\alpha=S(v_1, v_2)$, where $S$ is the $n\to\infty$ pointwise limit of the sequence $S^n$ (when the sequence $S^n$ converges).
 
@@
 
-In general, each bit $b^k$ of the image would be a binary function of the $n$ bits of $v_1$ and the $n$ bits of $v_2$, implicitly $b_k = b_k(b^1_1, b^1_2, \cdots, b^1_n, b^2_1, b^2_2, \cdots, b^2_n)$.
+In general, each bit $b_k$ of the image would be a binary function of the $n$ bits of $v_1$ and the $n$ bits of $v_2$, i.e. $b_k = b_k(b^1_1, b^1_2, \cdots, b^1_n, b^2_1, b^2_2, \cdots, b^2_n)$.
 With this, we can formulate any mapping between sets of $4^n$ distinct elements.
@@
 
-These simplifications might seem too restrictive, however, we will show in the rest of this thesis that this formulation is flexible enough to express mappings that have good properties such as "locality", all while being simple enough to help build intuitions.
+<!-- These simplifications might seem too restrictive, however, -->
+We will show in the rest of this thesis that this simple formulation is flexible enough to express mappings that have good properties such as "locality", while being simple enough to help build intuitions.
 
@@
 
-#### A link with space-filling curves
+<!-- #### A geometric view relating to space-filling curves
+Let's not mention space-filling curves, because the Column mapping is not space-filling (no continuity 1D -> 2D), and because we would need a whole lot more formalism, see @Sagan1994
+ -->
+#### A geometric view relating to curves in $[0,1]^2$
 
@@
 
-The intuition behind this mapping is we enumerate the populations on a grid column by column, letting the index of the rows vary the fastest.
+The intuition behind this mapping is that we enumerate the populations on a grid column by column, letting the index of the rows vary the fastest.
 We visualize this mapping in @fig-col, by coloring the square populations by the position of the corresponding $[0,1]$ population, and additionally drawing lines between the points to help guide the eye.
@@
 
-#### "Local" space-filling curves {#sec-local}
+#### "Local" curves {#sec-local}
 
@@
 
-The attentive reader might however object to this treatment of locality, since in some sense the Column mapping is local, but from 2D to 1D.
-This can be seen from the definition of $\alpha^{(n)}=C(v^{(n)}_1,v^{(n)}_2) = 0.b^1_1 b^1_2 \cdots b^1_n b^2_1 b^2_2 \cdots b^2_n$, since small variations of order $2^{-n}$ in $v^{(n)}_1$ result in small variations also of order $2^{-n}$ in $\alpha^{(n)}$.
-Similarly, small variations of order $2^{-n}$ in $v^{(n)}_1$ result in even smaller variations also of order $2^{-2n}$ in $\alpha^{(n)}$ !
-We therefore seem to have two notions of locality: from 1D to 2D, and from 2D to 1D; and the Column mapping shows that they are not equivalent.
+Note that, in some sense, the Column mapping is local, from 2D to 1D.
+This can be seen from the definition of $\alpha^{(n)}=C(v^{(n)}_1,v^{(n)}_2) = 0.b^1_1 b^1_2 \cdots b^1_n b^2_1 b^2_2 \cdots b^2_n$: small variations of order $2^{-n}$ in $v^{(n)}_1$ result in small variations also of order $2^{-n}$ in $\alpha^{(n)}$.
+Small variations of order $2^{-n}$ in $v^{(n)}_1$ result in even smaller variations also of order $2^{-2n}$ in $\alpha^{(n)}$ !
+We therefore have two notions of locality: from 1D to 2D, and from 2D to 1D; and the Column mapping shows that they are not equivalent.
 
-However, we argue that in our case only the notion of locality from 1D to 2D matters.
-In essence, the dynamics in the 2D embedding are already known, and we ask the question of whether it is possible to write an equivalent neural field 1D (in the sense that it has identical latent trajectories).
+We argue that in our case only the notion of locality from 1D to 2D matters.
+In essence, the dynamics in the 2D embedding are already known, and we ask the question of whether it is possible to write an equivalent neural field in 1D.
+<!-- (in the sense that it has identical latent trajectories) -->
 Without repeating the intuition given in @sec-intro-simnf, if neighbouring populations in 1D have similar potentials, then this allows us to write a neural field equation in $[0,1]$.
@@
 We discussed how this hypothetical mapping maps each position in $[0,1]^2$ into a "random position" in $[0,1]$.
-A finite-$n$ formulation would be that the mapping corresponds to a *random permutation* of the bits.
-It is unclear how to formally write this mapping in a way that is well-defined for a given $n$, since by definition, it is random.
+A finite-$n$ formulation would be that the mapping corresponds to a *random permutation* of the bits, therefore, its limit is obviously not well-defined.
+<!-- It is unclear how to formally write this mapping in a way that is well-defined for a given $n$, since by definition, it is random. -->
 For a given position of the input $(v_1,v_2)$, we would like that the image is stable in the $n\to\infty$ limit.
@@
 This is the condition of "pointwise convergence" of the mappings.
-We might be tempted to write that any mapping for which we can write the binary expansion $S^n$ is pointwise convergent, however, we show by the example of the "anti-Z" mapping that this is not the case.
-Let us define for finite $n$
+Let us introduce another mapping, which will help to give a better intuition of what pointwise convergence means in our context. We define, for finite $n$, the "anti-Z" mapping and its inverse:
+<!-- We might be tempted to write that any mapping for which we can write the binary expansion $S^n$ is pointwise convergent, however, we show by the example of the "anti-Z" mapping that this is not the case.
+Let us define for finite $n$ -->
 
@@
 v^{(n)}_1 &= A^{-1}_1(\alpha^{(n)}) = 0.b_{2n-1} b_{2n-3} \cdots b_{1} = \sum_{k=1}^n b_{2(n-k)-1} 2^{-k} \\
-v^{(n)}_2 &= A^{-1}_2(\alpha^{(n)}) = 0.b_{2n} b_{2n-2} \cdots b_{2} = \sum_{k=1}^n b_{2(n+1-k)} 2^{-k}
+v^{(n)}_2 &= A^{-1}_2(\alpha^{(n)}) = 0.b_{2n} b_{2n-2} \cdots b_{2} = \sum_{k=1}^n b_{2(n+1-k)} 2^{-k}.
 \end{aligned}
@@
 
-#### Conservation of bijectivity in the $n\to\infty$ limit {#sec-ninfty-bijection}
+#### Bijectivity in the $n\to\infty$ limit {#sec-ninfty-bijection}
 
 However, the behavior of the Column mapping and that of the Z-mapping are very different in the $n\to\infty$ limit.
-Up until now, we have avoided this issue by considering the finite-$n$ approximation of the mappings, which by construction is bijective.
+Up until now, we have avoided this issue by considering the finite-$n$ approximation of the mappings, which by construction is numerically bijective (that is, bijective on the discretization of $[0,1]^2$ and $[0,1]$ corresponding to finite $n$).
 We illustrate the infinite-$n$ (numerically, large $n$) limit of these two mappings in @fig-n-to-infty.
@@
 
-We see that the Column mapping converges to projection on $v_1$, in other words, $\alpha=C(v_1, v_2) = v_1$, showing that despite every $C^{n}$ being bijective, the limit of the sequence is not bijective.
+We see that the Column mapping converges to projection on $v_1$, in other words, $\alpha=C(v_1, v_2) = v_1$, showing that despite every $C^{n}$ being numerically bijective, the limit of the sequence is not bijective.
 This becomes clear when looking at the binary expansion of $C$ :
@@
 The Z-mapping seems to converge to a fractal curve.
-Numerically, there seems to be an infinite number of discontinuities, but the size of these discontinuities is overwhelmingly small, so that *on average* they vanish as $n\to\infty$.
-We won't go further into detail about this (we refer to @Sagan1994 for an in-depth study), but this gives some intuition for the arguments presented in @sec-locality.
+Numerically, there seems to be an infinite number of discontinuities, but the sizes of these discontinuities are, for the most part, small, so that *on average* they vanish as $n\to\infty$.
+We won't go further into detail about this,[^zmapping-notbijective] but this gives some intuition for the arguments presented in @sec-locality.
 
-We finish this discussion on the limit of the sequence of mappings by asking if there are other mappings similar to the Z-mapping, in the sense that they are pointwise convergent and the limit is bijective.[^zmapping-notbijective] In an informal way, we established two conditions for this in the context of the mappings described in @sec-mappings-binary:
+We finish this discussion on the limit of the sequence of mappings by asking if there are other mappings similar to the Z-mapping, in the sense that they are pointwise convergent and the limit is bijective.
+In an informal way, we established two conditions for this in the context of the mappings described in @sec-mappings-binary:
 
-[^zmapping-notbijective]: Despite what can often be found in online resources, the Z-mapping is *not* a bijection because that would violate Netto's theorem.
+[^zmapping-notbijective]:
+  We refer to @Sagan1994 for an in-depth study, and in particular chapter 6, in which the author shows that the Z-curve converges to the continuous and surjective Lebesgue curve, which is a Peano-type function.<br>
+  Despite what can often be found in online resources, the Z-curve is *not* a bijection because the added injectivity would violate Netto's theorem.
 For every point in $[0,1]^2$, there are at most 4 distinct preimages in $[0,1]$.
-The reason for this is the ambiguity in the writing of binary expansions, and in @sec-zcurve-noninjective we give proof that the Z-mapping is not injective in the $n\to\infty$ limit.
+The reason for this is the ambiguity in the writing of binary expansions, and in @sec-zcurve-noninjective we give proof that the Z-curve is not injective in the $n\to\infty$ limit.<br>
+  For the following, we will consider the Z-curve to be bijective by disallowing expansions of the type $0.1000\cdots=0.0111\cdots$.
+  <!-- We note that in @Sagan1994 many other Peano curves (continuous surjections $[0,1]\mapsto [0,1]^2$) are presented (albeit more difficult to implement numerically). -->
 
@@
 We can evidently define mappings where we interleave the bits in a different order, such as $0.b^2_1 b^1_1 b^2_2 b^1_2 \cdots$; or mappings where we reorder the $k$ bits of the input, as in $0.b^1_2 b^2_2 b^1_1 b^2_1 b^1_4 b^2_4 b^1_3 b^2_3 \cdots$, but the conditions restrict the reordering to be permutations of neighbouring bits.
-In essence, these mappings are not fundamentally different, and the Z-mapping can be seen as a "canonical representation" of the pointwise convergent mappings with bijective limits, which can be expressed as the permutation of binary digits.
+In essence, these mappings are not fundamentally different, and the Z-mapping can be seen as a "standard example" of the pointwise convergent mappings with bijective limits, which can be expressed as the permutation of binary digits.
 
-### Visualizating the mappings of the dynamic 2D neural field
+### Visualizating of the mappings of the dynamic 2D neural field
 
@@
 Before we argue about how to simulate the dynamics of the 1D neural field, let us simulate the dynamics of the 2D neural field, then at each timestep, we map the 2D neural field to the 1D embedding.
-This will give us a reference of how the 1D neural field is supposed to look under the condition that its latent dynamics are identical.
+This will give us a reference of how the 1D neural field is supposed to look under the condition that its dynamics are identical.
 
@@
 Since the mappings we presented can be written as a binary expansion, they naturally lend themselves to numerical approximations.
-Furthermore, this is made especially easy since the mappings we defined simply are permutations of bits (in the $n < \infty$ case), and so we can write a one-to-one correspondence between unsigned integers.
+Furthermore, this is made especially easy since the mappings we defined are simply permutations of bits (in the $n < \infty$ case), and so we can write a one-to-one correspondence between unsigned integers.
 
@@
 
-### Numerical approximations of a fractal structure
+### Naive simulations of the neural field on $[0,1]$
 
-In the continuation of the simulation methods established in the @sec-simnf, we adopt a grid discretization scheme, as it naturally lends itself to be applied with the mappings.
-We however note that using binning methods, the Monte-Carlo simulation can still be applied.
+In the continuation of the simulation methods established in the @sec-simnf, we adopt a grid discretization scheme to simulate the neural field on $[0,1]$, as the grid naturally yields square populations, which can directly be used with the mappings introduced previously.[^note-mc]
+
+[^note-mc]: In order to simulate the mappings using Monte-Carlo schemes, binning methods can be applied.
 The intuition behind this is that sampled populations inside a small square of size $2^{-n} \times 2^{-n}$ give rise to the notion of an "average connectivity kernel", and in our case the notion of "mean patterns".
@@
 
+### Coarse-graining the one-dimensional neural field
+
 In some sense, we would like to "take into account" the fact that a mapping of the square populations has been performed.
-Let us consider what happens to the mapping when $n$ is large, and in particular how $S^{n/2}$ and $S^{n}$ behave.
+Simulations of the neural field equation on $[0,1]^2$ are done by using $4^n$ square populations on a grid, but, as long as we have $4^n$ corresponding segment populations in $[0,1]$, the dynamics will trivially be the same, *independantly of the mapping* $S^n$.
+Therefore, we need a way to reduce the number of segment populations in $[0,1]$.
+
+We do this using "coarse-graining", in which we average $2^n$ consecutive segment populations, each of length $4^{-n}$, in order to obtain $2^n$ coarse-grained segment populations, each of length $2^{-n}$.
+This operation is illustrated in @fig-averaging.
+
+![(TODO) Averaging $2^n$ of the $4^n$ segments populations of $S^n$ results in the $2^n = 4^{n/2}$ segments populations of $S^{n/2}$](figures/photo_2023-06-13_12-15-32.jpg){width=400px #fig-averaging}
+
+Without coarse-graining, the mesh size along each spatial dimension is $2^{-n}$ in 2D, but directly applying the mapping results instead in a mesh size of $4^{-n}$ in 1D, which is much finer than the $2^{-n}$ per dimension.
+However, after coarse-graining, the effective mesh size in 1D becomes $2^{-n}$, which matches the 2D mesh size.
+Therefore, we are are not compensating the reduction of dimensionality by an increase in the numerical resolution.
+
+<!-- Let us consider what happens to the mapping when $n$ is large, and in particular how $S^{n/2}$ and $S^{n}$ behave.
 Since the mappings are pointwise convergent and we are considering large $n$, we argue that both $S^{n/2}$ and $S^{n}$ are good approximations of the true mapping $S$.
@@
 Another way of thinking about this is that $S^n$ is considered to be the "mathematically exact" fractal structure and $S^{n/2}$ the numerical approximation.
+The way we go from $S^n$ to $S^{n/2}$ is by averaging the segment populations by batches of $2^n$, as schematized in @fig-averaging.
+-->
 
-The way we go from $S^n$ to $S^{n/2}$ is by averaging the segment populations by batches of $2^n$, as schematized in @fig-averaging. We call this operation "downsampling" the $S^n$ mapping.
-
-![(TODO) Averaging $2^n$ of the $4^n$ segments populations of $S^n$ results in the $2^n = 4^{n/2}$ segments populations of $S^{n/2}$](figures/photo_2023-06-13_12-15-32.jpg){width=400px #fig-averaging}
-
-The geometrical intuition we can give for this is how the averaging of the segment populations leads to sampling the connectivity kernel at the "average" positions in the $[0,1]^2$ embedding.
-In @fig-averaging-geom, we colour-code the square populations corresponding to the segment populations before averaging.
+A geometrical intuition we can give for the coarse-graining is how the averaging of the segment populations leads to sampling the connectivity kernel at the "average" positions in the $[0,1]^2$ embedding.
+In @fig-averaging-geom, we colour-code the square populations corresponding to the segment populations before coarse-graining.
 
@@
 
-(TODO) The corresponding square populations corresponding to the segment populations before downsampling.
+(TODO) The corresponding square populations corresponding to the segment populations before coarse-graining.
 :::
 
-One question that may arise is why we take $n/2$ instead of $n-1$.
+In summary, coarse-graining allows us to compare the dynamics of a neural field in $[0,1]^2$ simulated using $2^n \times 2^n = 4^n$ square populations; with the dynamics of a neural field in $[0,1]$ simulated using $2^n$ segment populations.
+
+<!-- One question that may arise is why we take $n/2$ instead of $n-1$.
 The reason for this is that we want the number of averaged segment populations to scale with $n$.
@@
 
+For Nicole: This is adressed by the argument of mesh size !!
+-->
+
 ## Simulations of neural fields in 3D, 2D and 1D
@@
 
-![(TODO animation) Simulations of downsampling with different mappings from the $p=2$ embedding.](figures/coarsegraining.png){width=500px}
+![(TODO animation) Simulations of coarse-graining with different mappings from the $p=2$ embedding.](figures/coarsegraining.png){width=500px}
 

### 2023-06-18 manuscript/chapters/4_locality.qmd
dkfjwen fhjk !!!!!
valentin proofread and comments

@@
 The previous chapter gave a detailed description of the mappings $S : [0,1]^2 \mapsto [0,1]$, as well as a few numerical considerations.
-We found that mappings that yield equivalent dynamics crucially seem to have in common the property of *locality*, that is that they conserve the regularity between the domain and the image, even if the image seems to have a fractal structure.
-Intuitively, we saw that the locality seems to emerge from the mapping having overwhelmingly small variations.
+We found that mappings that yield equivalent dynamics crucially seem to have in common the property of *locality* (from 1D to 2D), even if the image seems to have a fractal structure.
+Intuitively, we saw that the locality seems to emerge from the mapping having for the most part small discontinuities.
 
 In this chapter, we will further explore these results, and attempt to formulate a measure of this "locality".
-We start by showing that bijectivity and measurability are sufficient conditions for equivalent dynamics.
-Then, we apply the locality measure to the special cases of the Column and Z-mappings, and show how this notion can discriminate between the two.
-Finally, we show that the formulated measure is sufficient to ensure the numerical estimation of the integral on the fractal structure converges to the analytical integral; therefore, such embeddings are feasible to find and simulate numerically.
+As a preamble, we show that bijectivity and measurability are sufficient conditions for  the 2D and 1D neural field dynamics to be equivalent.
+Then, we apply our measure of "locality" to the special cases of the Column and Z-mappings, and show how this notion can discriminate between the two mappings.
+Finally, we show that our measure of locality is sufficient to ensure that the numerical estimation of the integral converges to the analytical integral; therefore, such embeddings can be simulated numerically.
 
@@
 
-In the previous chapter, we introduced Cantor and Peano functions, and argued that if have to choose between the two, then the non-injective Peano functions are preferable, because, by Netto's theorem, they are the only ones that can be space-filling and reach all the square populations.
-Despite this, let us make the simplifying assumption that $S$ is a bijective function.
-However, to not break Netto's theorem, we do not make any assumptions regarding its continuity.
+<!-- In the previous chapter, we introduced Cantor and Peano functions, and argued that if have to choose between the two, then the non-injective Peano functions are preferable, because, by Netto's theorem, they are the only ones that can be space-filling and reach all the square populations.
+-->
+<!-- Let us make the assumption that $S$ is a bijective function. However, to not break Netto's theorem, we do not make any assumptions regarding its continuity.
 
-We furthermore assume that the mapping $S$ is measurable.
+Nicole: let's not mention Netto's theorem here.
+It applies to S^{-1}, and it becomes difficult to reason about when we talk about S.
+I don't think the proof works strictly in our case, because the Z-mapping is not a bijection, but let us assume uwuwuwu
+-->
+
+Let us assume that the mapping $S$ is bijective and measurable.
 This is not a strong assumption, since the limit of pointwise convergent measurable functions is also measurable.
-In our case, we have written the finite-$n$ expansions $S^n$, which by being sums of piecewise constant functions, are also measurable.
-Measurability of $S$ allows us to write $\lambda \circ S^{-1}$, since by definition the preimage of a measurable function is measurable.
+In our case, we have written the finite-$n$ expansions $S^n$, which, by being sums of piecewise constant functions, are also measurable.
+The measurability of $S$ allows us to write $\lambda \circ S^{-1}$, the pushforward measure (or image measure) of $\lambda$ under the mapping $S$.
 
@@
 
-In the last line, we have used the bijectivity of $S$. The Lebesgue measure $\lambda : [0,1]^2 \mapsto \R^+$ simply plays the role of a uniform distribution on $[0,1]^2$, and we simply write
+In the last line, we used a basic property of the pushfoward measure (see @Bogachev2007, Theorem 3.6.1, p.
+190).
+The Lebesgue measure $\lambda : [0,1]^2 \mapsto \R^+$ simply plays the role of a uniform distribution on $[0,1]^2$, and we simply write
 
@@
 
-[TODO : uniqueness of the solution ?]
+We do not address the question of well-posedness of the neural field equations in this thesis; for proofs of the well-posedness of neural field equations, we refer the reader to [@faugeras2008absolute;@faugeras2009persistent;@veltz2009localglobal].
+
+<!-- TODO : we may argue that the Z-curve is not bijective. This complicates the proof. Though, the pre-image of (v_1, v_2) is composed of at most 4 points, and the measurability compensates. -->
 
@@
 
-### The total variation attempt
+<!-- ### The total variation attempt
 
@@
   \mathrm{TV}(S^{-1}) = \sup_{P} \sum_{i=1}^{N_P} \norm{S^{-1}(\alpha_i) - S^{-1}(\alpha_{i-1})}_1.
-  $$ -->
+  $$
   We don't use the more complex definition, because our version of total variation is just used for illustration purposes.
@@
 
-<!-- In some sense, having the constraint that the mapping traces a path in the bounded space $[0,1]^2$ that can go through each point of the grid only once leaves little flexibility.
+In some sense, having the constraint that the mapping traces a path in the bounded space $[0,1]^2$ that can go through each point of the grid only once leaves little flexibility.
 For fixed $n$, there are only finitely many "large jumps" the mapping can make, and the large number of "small jumps" still add up, so that the total variation still grows at approximately the same rate.
@@
 
-When we do the downsampling of the mappings, we expressed that the locality helps by bounding the variation inside of each averaged bin.
-Therefore, what we would like to express, is that on average, the variation inside of each bin before downsampling is small.
+The previous chapter gave the intuition that local mappings map populations close in 1D to populations close in 2D.
+To verify this, we might try to compare the distance between all segment populations and the population at $\alpha=0$, to the distance of all square populations to the population at $S^{-1}(\alpha=0)$.
+This is done in @fig-distance-2d-1d, where the 2D distance is taken to be the $\ell^2$ norm.
+If the distances matched perfectly, then we would see a line with slope $\sqrt 2$ in the graph.
+
+![(TODO random, column and zcurve only, rename neuron to population) Distance between the first population and all other populations, in 1D and 2D.
+The dotted line has a slope $\sqrt 2$.](figures/distance_2d_1d.png){#fig-distance-2d-1d}
+
+For control, we also demonstrate that the random mapping has no locality: there is no correlation between the positions in 1D and the positions in 2D.
+Both Column and Z-mappings show a clear trend, the distances in 2D are correlated to the distances in 1D.
+The notable difference is that distances for the Column mapping seem to "oscillate" wildly, while the Z-mapping seems much more regular and matches the straight line more closely.
+
+With this view, we understand that the coarse-grained performed in @sec-coarse-graining was making use of that fact that locality helps by bounding the variation inside of each averaged bin.
+Therefore, we would like for mathematically formulate that on average, the variation inside of each bin before coarse-graining is small.
 @fig-sketch-binned-variation gives an intuition of this.
@@
 $$ {#eq-Vn}
+
+<!-- Nicole: V_n -> 0 is implied by continuity -->
 
@@
 
-TODO : I have removed the definition involving the sup over all permutations, because we don't end up using it. Still mention for completeness ?
+<!-- TODO : I have removed the definition involving the sup over all permutations, because we don't end up using it. Still mention for completeness ?
+
+Valentin: For the master thesis, it is ok to not mention it.
+-->
 
@@
 \begin{aligned}
-Z^{-1}_1(\alpha) &= 0.\underbrace{i_{n-1} i_{n-3}}_\text{$n/2$ bits} \cdots i_1 b_n b_{n+2} \cdots b_{2m-1} \\
-Z^{-1}_2(\alpha) &= 0.\underbrace{i_{n-2} i_{n-4}}_\text{$n/2$ bits} \cdots i_0 b_{n+1} b_{n+3} \cdots b_{2m}
-\end{aligned}
+Z^{-1}_1(\alpha) &= 0.\underbrace{i_{n-1} i_{n-3} \cdots i_1}_\text{$n/2$ bits} b_n b_{n+2} \cdots b_{2m-1} \\
+Z^{-1}_2(\alpha) &= 0.\underbrace{i_{n-2} i_{n-4} \cdots i_0}_\text{$n/2$ bits} b_{n+1} b_{n+3} \cdots b_{2m}
+\end{aligned}of this locality
 $$
@@
 \begin{aligned}
-Z^{-1}_1(\alpha) &= 0.\underbrace{i_{n-1} i_{n-3}}_\text{$(n+1)/2$ bits} \cdots i_0 b_n b_{n+2} \cdots b_{2m-1} \\
-Z^{-1}_2(\alpha) &= 0.\underbrace{i_{n-2} i_{n-4}}_\text{$(n-1)/2$ bits} \cdots i_1 b_{n+1} b_{n+3} \cdots b_{2m}
+Z^{-1}_1(\alpha) &= 0.\underbrace{i_{n-1} i_{n-3} \cdots i_0}_\text{$(n+1)/2$ bits} b_n b_{n+2} \cdots b_{2m-1} \\
+Z^{-1}_2(\alpha) &= 0.\underbrace{i_{n-2} i_{n-4} \cdots i_1}_\text{$(n-1)/2$ bits} b_{n+1} b_{n+3} \cdots b_{2m}
 \end{aligned}
@@
 
-In @sec-coarse-graining we discussed how we approximate the fractal mappings numerically, and introduced the downsampling procedure, in which we average the populations in segments of length $2^{-n}$, and end up with $2^n$ "binned" segment populations, compared to the $4^n$ we started with.
+<!-- In @sec-coarse-graining we discussed how we approximate the fractal mappings numerically, and introduced the coarse-graining procedure, in which we average the populations in segments of length $2^{-n}$, and end up with $2^n$ "binned" segment populations, compared to the $4^n$ we started with.
 
@@
 Assuming that $V_n \xrightarrow{n\to\infty} 0$, is this enough to prove that the numerical estimation of the integral goes to its analytical value?
+-->
 
-The intuition behind the proof of convergence is that the variation, coming from the coarse-graining, inside each numerical bin vanishes as we increase the number of iterations $n$, even though the structure becomes fractal.
+In @sec-Vn, we have introduced the average variation metric $V_n$.
+Assuming that $V_n \xrightarrow{n\to\infty} 0$ (which expresses the locality property), does this imply that the numerical estimation of the integral in the neural field equation in $[0,1]$ converges to its true value?
+
+The intuition behind the proof of convergence is that the variation, coming from the coarse-graining, inside each numerical bin vanishes as we increase the number of iterations $n$, even though the the neural field $\tilde{h}$ and the connectivity kernel $\tilde{w}$ become fractal.
 
@@
 
-Let us first recall the numerical and analytical integral. We leave a $\tilde w(\cdot, \beta)$ as an implicit function of the first argument $\alpha$. 
+Let us first recall the numerical and analytical integral. For any $\alpha \in [0,1]$, let us define:
 $$
 \begin{aligned}
-\mathrm{NI}_n &= \frac{1}{2^n} \sum_{i=1}^{2^n} \tilde w(\cdot, \beta_i) \phi(\tilde h_i(t)) \\
-\mathrm{AI} &= \int_{[0,1]} \tilde w(\cdot, \beta) \phi(\tilde h(\beta, t)) \left[\lambda \circ S^{-1}\right](\d \beta)
+\mathrm{NI}_n &= \frac{1}{2^n} \sum_{i=1}^{2^n} \tilde w(\alpha, \beta_i) \phi(\tilde h_i(t)) \\
+\mathrm{AI} &= \int_{[0,1]} \tilde w(\alpha, \beta) \phi(\tilde h(\beta, t)) \left[\lambda \circ S^{-1}\right](\d \beta).
 \end{aligned}
@@
 
-We wish to prove that $|\mathrm{NI}_n - \mathrm{AI}| \xrightarrow{n \to \infty} 0$, and implicitly, that this is valid for all $\alpha \in [0,1]$.
+We wish to prove that $|\mathrm{NI}_n - \mathrm{AI}| \xrightarrow{n \to \infty} 0$.
 
@@
 \begin{aligned}
-|\mathrm{NI}_n - \mathrm{AI}| &= \left| \frac{1}{2^n} \sum_{i=1}^{2^n} \tilde w(\cdot, \beta_i) \phi(\tilde h_i(t)) - \int_{0}^{1} \tilde w(\cdot, \beta) \phi(\tilde h(\beta, t)) \left[\lambda \circ S^{-1}\right](\d \beta) \right| \\
-&= \left| \frac{1}{2^n} \sum_{i=1}^{2^n} \tilde w(\cdot, \beta_i) \phi(\tilde h_i(t)) - \sum_{i=1}^{2^n} \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \tilde w(\cdot, \beta) \phi(\tilde h(\beta, t)) \left[\lambda \circ S^{-1}\right](\d \beta) \right| \\
-&= \left| \sum_{i=1}^{2^n} \frac{1}{2^n} \tilde w(\cdot, \beta_i) \phi(\tilde h_i(t)) - \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \tilde w(\cdot, \beta) \phi(\tilde h(\beta, t)) \left[\lambda \circ S^{-1}\right](\d \beta) \right| \\
-&\leq \sum_{i=1}^{2^n} \left| \frac{1}{2^n} \tilde w(\cdot, \beta_i) \phi(\tilde h_i(t)) - \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \tilde w(\cdot, \beta) \phi(\tilde h(\beta, t)) \left[\lambda \circ S^{-1}\right](\d \beta) \right| \\
+|\mathrm{NI}_n - \mathrm{AI}| &= \left| \frac{1}{2^n} \sum_{i=1}^{2^n} \tilde w(\alpha, \beta_i) \phi(\tilde h_i(t)) - \int_{0}^{1} \tilde w(\alpha, \beta) \phi(\tilde h(\beta, t)) \left[\lambda \circ S^{-1}\right](\d \beta) \right| \\
+&= \left| \frac{1}{2^n} \sum_{i=1}^{2^n} \tilde w(\alpha, \beta_i) \phi(\tilde h_i(t)) - \sum_{i=1}^{2^n} \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \tilde w(\alpha, \beta) \phi(\tilde h(\beta, t)) \left[\lambda \circ S^{-1}\right](\d \beta) \right| \\
+&= \left| \sum_{i=1}^{2^n} \frac{1}{2^n} \tilde w(\alpha, \beta_i) \phi(\tilde h_i(t)) - \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \tilde w(\alpha, \beta) \phi(\tilde h(\beta, t)) \left[\lambda \circ S^{-1}\right](\d \beta) \right| \\
+&\leq \sum_{i=1}^{2^n} \left| \frac{1}{2^n} \tilde w(\alpha, \beta_i) \phi(\tilde h_i(t)) - \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \tilde w(\alpha, \beta) \phi(\tilde h(\beta, t)) \left[\lambda \circ S^{-1}\right](\d \beta) \right| \\
 \end{aligned}
@@
 We now want to prove that evaluating the integrand at $\beta_i = \tfrac{i-1}{2^n}$ gives a good approximation of the integral.
-Because $S$ is measurable and bijective, we have $\int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \left[\lambda \circ S^{-1}\right](\d \beta) = \tfrac{1}{2^n}$, and we write
+Let us assume that $S$ is the Z-mapping.
+Then, $S$ is measure-preserving (if $\lambda$ is the Lebesgue measure on $[0,1]^2$, then $\lambda \circ S^{-1}$ is the Lebesgue measure on $[0,1]$).
+This implies that we have $\int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \left[\lambda \circ S^{-1}\right](\d \beta) = \tfrac{1}{2^n}$, and we write
 
@@
 |\mathrm{NI}_n - \mathrm{AI}| &\leq \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \Delta(\beta) \left[\lambda \circ S^{-1}\right](\d \beta) \right| \\
-\text{where}\ \Delta(\beta) &= \tilde w(\cdot, \beta_i) \phi(\tilde h_i(t)) - \tilde w(\cdot, \beta) \phi(\tilde h(\beta, t))\\
-&= w_U(\cdot, S^{-1}(\beta_i)) \phi(h_U(S^{-1}(\beta_i), t)) - w_U(\cdot, S^{-1}(\beta)) \phi(h_U(S^{-1}(\beta), t))
+\text{where}\ \Delta(\beta) &= \tilde w(\alpha, \beta_i) \phi(\tilde h_i(t)) - \tilde w(\alpha, \beta) \phi(\tilde h(\beta, t))\\
+&= w_U(\alpha, S^{-1}(\beta_i)) \phi(h_U(S^{-1}(\beta_i), t)) - w_U(\alpha, S^{-1}(\beta)) \phi(h_U(S^{-1}(\beta), t))
 \end{aligned}
@@
 The main source of variance inside the integrand $\Delta(\beta)$ is the term $S^{-1}$, since $w_U$ and $h_U$ are "nice" functions.
-We express this regularity by the added assumption that $w_U(\cdot, \vec u) \phi(h_U(t, \vec u))$ is Lipschitz in $\vec u$, with Lipschitz constant $L(\cdot)$, and $\norm{\cdot}$ the corresponding norm on $[0,1]^2$.
+We express this regularity by the added assumption that $w_U(\alpha, \vec u) \phi(h_U(t, \vec u))$ is Lipschitz in $\vec u$, with Lipschitz constant $L(\alpha)$, and $\norm{\alpha}$ the corresponding norm on $[0,1]^2$.
 We argue this is not a strong assumption, because this is just expressing the regularity of the integrand in the integral over $[0,1]^2$.
@@
 
-Applying the Lipschitz property yields the following inequality on the integrand:
+Applying the Lipschitz assumption yields the following inequality on the integrand:
 
@@
 \begin{aligned}
-\Delta(\beta) &= w_U(\cdot, S^{-1}(\beta_i)) \phi(h_U(S^{-1}(\beta_i), t)) - w_U(\cdot, S^{-1}(\beta)) \phi(h_U(S^{-1}(\beta), t)) \\
-&\leq L(\cdot) \norm{S^{-1}(\beta_i) - S^{-1}(\beta)}
+\Delta(\beta) &= w_U(\alpha, S^{-1}(\beta_i)) \phi(h_U(S^{-1}(\beta_i), t)) - w_U(\alpha, S^{-1}(\beta)) \phi(h_U(S^{-1}(\beta), t)) \\
+&\leq L(\alpha) \norm{S^{-1}(\beta_i) - S^{-1}(\beta)}
 \end{aligned}
@@
 \begin{aligned}
-|\mathrm{NI}_n - \mathrm{AI}| &\leq L(\cdot) \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \left[\lambda \circ S^{-1}\right](\d \beta) \right| \\
-&\leq L(\cdot) \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \sup_{\beta \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \left[\lambda \circ S^{-1}\right](\d \beta^\prime) \right| \\
-&= L(\cdot) \frac{1}{2^n} \sum_{i=1}^{2^n} \sup_{\beta \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \\
-&\leq L(\cdot) \frac{1}{2^n} \sum_{i=1}^{2^n} \sup_{\beta, \beta^\prime \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]^2} \norm{S^{-1}(\beta^\prime) - S^{-1}(\beta)} \\
-&= L(\cdot) V_n \\
+|\mathrm{NI}_n - \mathrm{AI}| &\leq L(\alpha) \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \left[\lambda \circ S^{-1}\right](\d \beta) \right| \\
+&\leq L(\alpha) \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \sup_{\beta \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \left[\lambda \circ S^{-1}\right](\d \beta^\prime) \right| \\
+&= L(\alpha) \frac{1}{2^n} \sum_{i=1}^{2^n} \sup_{\beta \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \\
+&\leq L(\alpha) \frac{1}{2^n} \sum_{i=1}^{2^n} \sup_{\beta, \beta^\prime \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]^2} \norm{S^{-1}(\beta^\prime) - S^{-1}(\beta)} \\
+&= L(\alpha) V_n \\
 &\xrightarrow{n \to \infty} 0

### 2023-06-19 manuscript/chapters/3_mappings.qmd
little kitty be vibing

@@
 All this mathematical formalism expresses, in essence, that the mapping $S^n$ is just a *reordering* of the input bits.
-$S^n$ takes $4^n$ points embedded in $[0,1]^2$, and reorders them onto $4^n$ points in $[0,1]$.
+$S^n$ takes $4^n$ points embedded in $[0,1]^2$, and reorders them onto $4^n$ points in $[0,1]$.[^edges]
+
+[^edges]:
+  We note that strictly speaking, $S^n$ maps $(0., 0.)$ onto $0$, and the point
+  $$
+  S^n(1., 1.) = S^n(0.111\cdots, 0.111\cdots) = 0.111\cdots = 1
+  $$
+  is never reached at finite $n$.
+Numerically, we encounter the problem that the inverse of the Gaussian CDF, used to compute the connectivity, goes to infinity when $v_\mu$ (one of the components of $(v_1, v_2) = S^{-1}(\alpha)$) goes to $0$ or $1$.
+To avoid this, we offset the squares (by $2^{-n-1}$ along each dimension), so that geometrically their position is not indexed by their lower-left corner, but rather their center.
+Analytically, this is not a problem, because the kernel defined in @eq-hu-wu diverges only linearly, compared to the exponential cutoff of the Gaussian distribution, so that the integrand still vanishes at infinity.
+
+<!-- TODONE NOTE : talk about how we center the bins instead of taking the corners.
+this is not super important in the $n\to\infty$ limit, but numerically at 0 and 1 the $\mathrm{CDF}^{-1}$ goes to infty.
+at infinity, the connectivity kernel is infinite but we have zero neurons (zero measure, linear growth of kernel vs exponential tail of the gaussian) : effectively the recurrent current is zero and we can safely ignore these effects.
+-->
 
 Therefore, there is a one-to-one mapping between the points on the square and the points on the segment, and the mapping $S^n$ is bijective.
+In the following, we will, for ease of notation and when it is clear in the context, drop the index $n$ when writing $S^n$.
+This is often justified (particularly with the Z-mapping and Column mapping introduced later), because the finite-$n$ truncations $(v_1^{(n)}, v_2^{(n)})$ result in finite truncations of $S$, therefore $S(v_1^{(n)}, v_2^{(n)}) = S^n(v_1^{(n)}, v_2^{(n)})$.
+
+<!-- TODONE NOTE : might need to talk about how $S(v^{(n)}_1,v^{(n)}_2) = S^n(v^{(n)}_1,v^{(n)}_2)$ ?
+so basically the limit $n\to\infty$ is the limit where we take infinite precision in the input.
+Or we can drop the $n$ on the inputs, $\alpha^{(n)} = S^n(v^{(n)}_1,v^{(n)}_2) = S^n(v_1,v_2)$ (i think this true just for pointwise convergent mappings) -->
 
@@
 We will show in the rest of this thesis that this simple formulation is flexible enough to express mappings that have good properties such as "locality", while being simple enough to help build intuitions.
-
-TODO NOTE : might need to talk about how $S(v^{(n)}_1,v^{(n)}_2) = S^n(v^{(n)}_1,v^{(n)}_2)$ ?
-so basically the limit $n\to\infty$ is the limit where we take infinite precision in the input.
-Or we can drop the $n$ on the inputs, $\alpha^{(n)} = S^n(v^{(n)}_1,v^{(n)}_2) = S^n(v_1,v_2)$ (i think this true just for pointwise convergent mappings)
-
-TODO NOTE : talk about how we center the bins instead of taking the corners.
-this is not super important in the $n\to\infty$ limit, but numerically at 0 and 1 the $\mathrm{CDF}^{-1}$ goes to infty.
-at infinity, the connectivity kernel is infinite but we have zero neurons (zero measure, linear growth of kernel vs exponential tail of the gaussian) : effectively the recurrent current is zero and we can safely ignore these effects.
 
@@
 [^zmapping-notbijective]:
-  We refer to @Sagan1994 for an in-depth study, and in particular chapter 6, in which the author shows that the Z-curve converges to the continuous and surjective Lebesgue curve, which is a Peano-type function.<br>
-  Despite what can often be found in online resources, the Z-curve is *not* a bijection because the added injectivity would violate Netto's theorem.
+  We refer to @Sagan1994 for an in-depth study, and in particular chapter 6, in which the author shows that the Z-curve converges to the Lebesgue curve.<br> 
+  <!-- which is a Peano-type function (continuous and surjective from $[0,1]$ into $[0,1]^2$).<br> -->
+  Despite what can often be found in online resources, the Z-curve is *not* a bijection.
+  <!-- because the added injectivity would violate Netto's theorem. -->
   For every point in $[0,1]^2$, there are at most 4 distinct preimages in $[0,1]$.
@@
   <!-- We note that in @Sagan1994 many other Peano curves (continuous surjections $[0,1]\mapsto [0,1]^2$) are presented (albeit more difficult to implement numerically). -->
+  <!-- Nicole: I don't say that the Lebesgue curve is continuous, because that opens up a whole other can of worms that I don't want to go into -->
 
@@
 ## Coarse-graining the neural field in $[0,1]$ {#sec-coarse-graining}
-
-### Practical aspects of implementing the mappings
-
-In the previous sections, we have discussed what mappings have good properties, and have given a few hints as to how we can implement them numerically.
-Since the mappings we presented can be written as a binary expansion, they naturally lend themselves to numerical approximations.
-Furthermore, this is made especially easy since the mappings we defined are simply permutations of bits (in the $n < \infty$ case), and so we can write a one-to-one correspondence between unsigned integers.
-
-Without going into too much detail on the programming, we have implemented the mappings as all inheriting from the same abstract class, and each mapping simply has to define a method `indices_to_indices2d(indices: array[int, 1]) -> array[int, 2]` to go from 1D to 2D ($S^{-1}$), and a method `indices2d_to_indices(indices2d: array[int, 2]) -> array[int, 1]` to go from 2D to 1D.[^indices] The abstract class simply defines additional quality-of-life functionality, which makes use of these two methods.
-
-[^indices]: In the source code, only the method `indices(v12: array[float, 2]) -> array[int, 1]` is defined.
-This is because some mappings are implemented (for instance, the `RecursiveLocalMapping`, that we don't discuss here), which don't lend themselves to the simple permutation of bits formulation.
 
@@
 
-TODO small intro
+In this section, we show numerical simulations of the neural field in $[0,1]$, obtained from mapping higher-dimensional neural fields in $[0,1]^p,\,p\geq 2$.
+We see that the coarse-graining procedure works when applied from $[0,1]^2$ to $[0,1]$.
+Then, we demonstrate that it can be applied iteratively, to map from $[0,1]^3$ to $[0,1]^2$, and from $[0,1]^2$ to $[0,1]$.
+
+### Practical aspects of implementing the simulations of the one-dimensional neural field
+
+We give a quick overview of the numerical details that go into implementing the methods presented above.
+Since the mappings (Z-mapping, Column mapping) can be written as binary expansions, they naturally lend themselves to numerical approximations.
+Furthermore, this is made especially easy since the mappings we defined are simply permutations of bits (in the $n < \infty$ case), and so we can write a one-to-one correspondence between unsigned integers.
+
+We have implemented the mappings as all inheriting from the same abstract class, and each mapping simply has to define a method `indices_to_indices2d(indices: array[int, 1]) -> array[int, 2]` to go from 1D to 2D ($S^{-1}$), and a method `indices2d_to_indices(indices2d: array[int, 2]) -> array[int, 1]` to go from 2D to 1D.[^indices] The abstract class simply defines additional quality-of-life functionality, which makes use of these two methods.
+
+[^indices]: In the source code, only the method `indices(v12: array[float, 2]) -> array[int, 1]` is defined.
+This is because some mappings are implemented (for instance, the [`RecursiveLocalMapping`](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_mapping.py#L244), that we don't discuss here), which don't lend themselves to the simple permutation of bits formulation.
+
+Finally, the coarse-graining is easily implemented as the combination of a reshape operation (to bring all consecutive sequences of $2^n$ populations together), followed by averaging (see the [implementation here](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_rnn.py#L186))
 
@@
 
-TODO : we run simulations and compare latent trajectories.
+To demonstrate the coarse-graining, we start with the cycling neural field presented in @sec-cycling-nf.
+The kernel $\tilde w$ of the one-dimensional neural field is numerically approximated using the coarse-graining procedure of different mappings.
+Simulations are presented in @fig-cg-2d-1d (we again set $\delta=6$ and the initial condition $h(z_1, z_2, t < 0) = z_1$), showing the activity levels, recurrent currents, and the latent trajectory $(\kappa_1(t), \kappa_2(t))$.
+In order, we discuss:
+
+1.
+The random mapping can be numerically evaluated as a trivial case of a mapping that does not have locality.
+The coarse-graining averages out all the structure of the activity, and the recurrent currents are zero, because the contributions from excitatory and inhibitory populations cancel out.
+This results in the erasure of the cycling behavior in the latent trajectory.
+2.
+The Column mapping shows the projection of the initial condition onto the $z_1$ axis, yielding the sigmoid in the $[0,1]$ embedding.
+Again, the current currents cancel out and yield zero, and the latent trajectory quickly decays towards $(\kappa_1, \kappa_2) = (0,0)$.
+3.
+The Z-mapping shows that despite the fractal structure of the activity, there is still regularity in the $[0,1]$ embedding, resulting in non-zero recurrent currents.
+In consequence, the latent trajectory shows a clear cycling behavior, that closely matches that of the 2D neural field.
+
+<!-- TODONE : we run simulations and compare latent trajectories.
 first show only column and z-mapping.
@@
 
-![(TODO animation) Simulations of coarse-graining with different mappings from the $p=2$ embedding.](figures/coarsegraining.png){width=500px}
+Nicole: we don't show the other mappings, because they are not as nice examples, and essentially they work the same
+-->
 
-### Iterative application of $S$ : from $[0,1]^3$ to $[0,1]$ {#sec-simulations-3d-2d-1d}
+![(TODO animation, higher N, remove all the other mappings) Simulations of coarse-graining with different mappings from the $p=2$ embedding.](figures/coarsegraining.png){width=500px #fig-cg-2d-1d}
 
-TODO : take cycling p=3
+### Iterative application of $S$ : from $[0,1]^3$ to $[0,1]^2$, then to $[0,1]$ {#sec-simulations-3d-2d-1d}
 
-#### Continuity on one axis and fractal on the other: from $[0,1]^3$ to $[0,1]^2$
+<!-- #### Continuity on one axis and fractal on the other: from $[0,1]^3$ to $[0,1]^2$ -->
 
-TODO : we do z-mapping on each slice $(v_2, v_3)$ of $v_1$, and downsample.
-we end up with a 2d neural field, direction $v_1$ is continuous, direction $\alpha_{23}$ is fractal.
-now the field is not continuous anymore, it is just sufficiently regular to be L1.
-this shows "local mappings conserve regularity"
+In the previous section, we showed that the coarse-graining procedure works from $[0,1]^2$ to $[0,1]$.
+However, one might argue that the reason why it worked, is not the mapping, but also the continuity of the neural field in $[0,1]^2$.
+To show the generality of our results, we show that coarse-graining works, even when the original neural field is fractal.
+
+For this, we again consider the same cycling field, but with $p=3$, such that the natural embedding space is $\R^3$, with the equivalent formulation on $[0,1]^3$.
+We again consider a grid with a mesh size of $2^{-n}$ along each spatial dimension, resulting in $2^n \times 2^n \times 2^n = 8^n$ "cube populations", at positions $(v_1, v_2, v_3) \in [0,1]^3$.
+
+To go from $[0,1]^3$ to $[0,1]^2$, we consider a slice $(v_2, v_3)$ at constant $v_1$, map $(v_2, v_3)$ to $\alpha_{23}$, then perform coarse-graining, which brings down the number of square populations in $[0,1]^2$ from $8^n$ to $4^n$.
+The result of this is an embedding of the neural field on $(v_1, \alpha_{23}) \in [0,1]^2$.
+The corresponding neural field is continuous in its $v_1$ component, but fractal in $\alpha_{23}$.
+This is visualized in @fig-cg-3d-2d-1d, where we see that in the state $h(z_1, z_2, z_3) = z_1$, the corresponding 2D neural field is constant along the $\alpha_{23}$ component, since the slice $(v_2, v_3)$ has a constant activity level.
+However, as soon as $h$ evolves to a new state, non-orthogonal to $z_2$ and $z_3$, the fractal structure becomes clear, because the slice $(v_2, v_3)$ is no longer constant.
 
 ::: {.content-visible when-format="html"}
-![(TODO 3D -> 2D only) Equivalence of dynamics, from 3D to 2D.](figures/embedding_3d_2d_1d.mp4)
+![Equivalence of dynamics, from 3D to 2D to 1D.](figures/embedding_3d_2d_1d.mp4){#fig-cg-3d-2d-1d}
 :::
@@
 ::: {.content-visible when-format="pdf"}
-![(TODO PDF, 3D -> 2D only) Equivalence of dynamics, from 3D to 2D.](figures/lol7.jpg)
+![(TODO PDF. initial state and final state) Equivalence of dynamics, from 3D to 2D to 1D.](figures/lol7.jpg){#fig-cg-3d-2d-1d}
 :::
 
-![(TODO 3D -> 2D only) Effective sampling points after averaging.](figures/mapping_3d_2d_1d.png){width=400px}
+We close this chapter by repeating the process to go from $[0,1]^2$ to $[0,1]$, mapping $(v_1, \alpha_{23})$ to $\alpha_{123}$, and coarse-graining.
+In total, we have coarse-grained two times: once from $[0,1]^3$ ($8^n$ cube populations) to $[0,1]^2$ ($4^n$ square populations), and again from $[0,1]^2$ to $[0,1]$ ($2^n$ segment populations).
+The corresponding 1D neural field is now fully fractal, and again exhibits the same dynamics.
+This shows that "local mappings conserve regularity": we started with a fully continous neural field in $[0,1]^3$, and obtained a partially fractal neural field in $[0,1]^2$.
+Despite the fractal structure, the 2D neural field is still regular enough that the dynamics are conserved, but more importantly that it can *again* be mapped to an equivalent 1D neural field.
+In @fig-averaging-geom-3d, we repeat a similar geometric view as @fig-averaging-geom: we see the Z-mapping in the slice $(v_2, v_3)$, how after coarse-graining the kernel is still effectively "sampled" at averaged locations, and crucially, how repeating the coarse-graining still leads to average locations that span the entire $[0,1]^3$ box (instead of being constrained to a plane embedded in $[0,1]^3$, as would have been the case with the Column mapping).
 
-#### From fractal to fractal : $[0,1]^2$ to $[0,1]$
-
-TODO : we can re-apply the z-mapping in $(v_1, \alpha_{23})$, we get a fractal 1d embedding. completely fractal, but the dynamics still work, because the mapping conserves regularity.
-
-::: {.content-visible when-format="html"}
-![Equivalence of dynamics, from 3D to 2D to 1D.](figures/embedding_3d_2d_1d.mp4)
-:::
-
-::: {.content-visible when-format="pdf"}
-![(TODO PDF) Equivalence of dynamics, from 3D to 2D to 1D.](figures/lol7.jpg)
-:::
-
-![(TODO) Effective sampling points after averaging.](figures/mapping_3d_2d_1d.png){width=400px}
+![(TODO what $n$) Effective sampling points after averaging.](figures/mapping_3d_2d_1d.png){width=400px #fig-averaging-geom-3d}

### 2023-06-19 manuscript/chapters/4_locality.qmd
little kitty be vibing

@@
 
-<!-- TODO : I have removed the definition involving the sup over all permutations, because we don't end up using it. Still mention for completeness ?
+<!-- TODONE : I have removed the definition involving the sup over all permutations, because we don't end up using it. Still mention for completeness ?
 
@@
 
-TODO NOTE : in the definition of $V$, we use the converged version of $S$, instead of a finite truncation $S^n$.
+<!-- TODONE : in the definition of $V$, we use the converged version of $S$, instead of a finite truncation $S^n$.
 we need to convince people this is not a big mistake, since $S^n \to S$, and we can take both limits (binning and $S^n$) "at the same time".
-This might actually be addressed by a similar todo in @sec-mappings-binary
+This is adressed by a footnote in @sec-mappings-binary -->
 

### 2023-06-19 manuscript/appendices/B_mean_patterns.qmd
all is good :3

::: {.content-hidden}
{{< include ../macros.tex >}}
\def\A{\mathcal{A}}
\def\B{\mathcal{B}}
:::

# Binned sampling of populations and visualization of kernels {#sec-mean-patterns}

## Mean field derivation

Let us start by taking $N$ samples of the distribution $\rho(z_1, z_2)$. As established in @sec-mc, this defines a network of neurons $i \in \{1,\cdots,N\}$ with connectivity matrix $J_{ij} = w_U(\vec{v_i}, \vec{v_j})$. We recall the evolution equation for the network of neurons:

$$
\dot h_i(t) = -h_i(t) + \sum_{j=1}^N J_{ij} \phi(h_j(t)).
$$ {#eq-evolution-network}

For now, we do not use the fact that in our setting, $J_{ij}$ is low-rank.
Applying the finite-$n$ mapping $S^n$,[^details] the neurons $\A = \{i_1, \cdots, i_{|\A |}\}$ will end up in the same 1D bin at position $\alpha$. There are 1D $4^n$ bins, each of length $4^{-n}$.

[^details]: We are glossing over the fact that $S^n : [0,1]^2 \mapsto [0,1]$, but the positions $\vec{z_i} \in \R^2$. In the code, we can deal with these by either applying the inverse CDF (as described in @sec-mapping-cdf), or considering a bounding box $[\min_i(z_{1,i}), \max_i(z_{1,i})] \times [\min_i(z_{2,i}), \max_i(z_{2,i})]$ of the sampled points, which is then rescaled to $[0,1] \times [0,1]$ (see implementation for [`Box`](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_mapping.py#L24)). Both approaches are equivalent, in the sense that they give rise to the same dynamics as $n$ becomes large, although the obtained connectivity matrix $\tilde J_{\alpha\beta}$ may not be the identical.

As explained in the main text, we average the neurons together in a bin, and we write the corresponding potential of the segment population at location $\alpha$:

$$
H_\alpha(t) = \frac{1}{|\A|} \sum_{i \in \A} h_i(t).
$$ {#eq-H}

The segment populations interact via the connectivity kernel $\tilde J_{\alpha\beta}$, and we write the equation of evolution as:

$$
\dot H_\alpha(t) = -H_\alpha(t) + \sum_{\beta=1}^{4^n} \tilde J_{\alpha\beta} \phi(H_\beta(t)).
$$ {#eq-H-evolution}

Since the neurons in $\A$ receive recurrent currents from neurons in $\B$, it might seem natural to take $\tilde J_{\alpha\beta}$ to be the average connectivity between the neurons in $\A$ and $\B$. Furthermore, the neurons in $\A$ receive recurrent currents from all the neurons in $\B$, which adds a weighing term $|\B|$:

$$
\tilde J_{\alpha\beta} = |\B| \frac{1}{|\B| |\A|} \sum_{i \in \A} \sum_{j \in \B} J_{ij} = \frac{1}{|\A|} \sum_{i \in \A} \sum_{j \in \B} J_{ij}.
$$ {#eq-Jab}

We now verify that our intuition is correct. We substitute the definition of the mean field @eq-H and the our guess @eq-Jab into @eq-H-evolution:

$$
\begin{aligned}
\frac{1}{|\A|} \sum_{i \in \A} h_i(t) &= \frac{1}{|\A|} \sum_{i \in \A} h_i(t) + \sum_{\beta=1}^{4^n} \frac{1}{|\A|} \sum_{i \in \A} \sum_{j \in \B} J_{ij} \phi(H_\beta(t)) \\
\iff&\\
\sum_{i \in \A} h_i(t) &= \sum_{i \in \A} h_i(t) + \sum_{i \in \A} \underbrace{\sum_{\beta=1}^{4^n} \sum_{j \in \B}}_{=\sum_{j=1}^N} J_{ij} \phi(H_\beta(t)) \\
&= \sum_{i \in \A} h_i(t) + \sum_{i \in \A} \sum_{j=1}^N J_{ij} \phi(H_\beta(t)).
\end{aligned}
$$

We now make the mean-field approximation by dropping $\sum_{i \in \A}$, that is, that all the neurons inside of each bin are the same (or at least, sufficiently similar so that the mean-field is not a big mistake). Doing so, we recover @eq-evolution-network, which concludes the proof.

## Sampling populations conserves the low-rank structure

In @eq-Jab, we defined the connectivity of the binned segment populations as the average connectivity of the corresponding neurons sampled in the 2D embedding. In the case of low-rank $J_{ij}$, we can show the resulting $\tilde J_{\alpha\beta}$ also has a low-rank structure.

$$
\begin{aligned}
\tilde J_{\alpha\beta} &= \frac{1}{|\A|} \sum_{i \in \A} \sum_{j \in \B} J_{ij} \\
&= \frac{1}{|\A|} \sum_{i \in \A} \sum_{j \in \B} \frac 1N \sum_{\mu=1}^p F_{\mu i} G_{\mu j} \\
&= \frac{1}{|\A| N} \sum_{\mu=1}^p \left(\sum_{i \in \A} F_{\mu i} \right) \left(\sum_{j \in \B} G_{\mu j} \right)
\end{aligned}
$$

The form of the connectivity motivates the definition of *mean low-rank vectors*:

$$
\tilde F_{\mu\alpha} = \frac{1}{|\A|} \sum_{i \in \A} F_{\mu i},\quad \tilde G_{\mu\beta} = \frac{1}{|\B|} \sum_{j \in \B} G_{\mu j}.
$$ {#eq-mean-lr-vectors}

Additionally, we note that $N = \sum_{\beta^\prime} |\B^\prime|$, and substituting, we find that:

$$
\tilde J_{\alpha\beta} = \frac{|\B|}{\sum_{\beta^\prime} |\B^\prime|} \sum_{\mu=1}^p \tilde F_{\mu\alpha} \tilde G_{\mu\beta},
$$ {#eq-Jab-mean-lr}

which shows that $\tilde J_{\alpha\beta}$ has a low-rank structure.[^selfconnections]

[^selfconnections]:
  We note that if we exclude self-connections (as is done in @schmutz2023convergence), that is we redefine $J_{ij} \mapsto (1 - \mathrm{Id}_{ij}) J_{ij}$, then we can show that $\tilde J_{\alpha\beta}$ becomes
  $$
  \tilde J_{\alpha\beta} = \frac{|\B|}{\sum_{\beta^\prime} |\B^\prime|} \sum_{\mu=1}^p \tilde F_{\mu\alpha} \tilde G_{\mu\beta} - \mathrm{Id}_{\alpha\beta} \frac{|\B|}{\sum_{\beta^\prime} |\B^\prime|} \sum_{\mu=1}^p \sum_{i \in \A} \frac{F_{\mu i}}{|\A|} \frac{G_{\mu i}}{|\A|}.
  $$
  We see that the correction is still of order $\bO(\tfrac{1}{N})$, and therefore can be again ignored in the $N \to \infty$ limit.

@eq-Jab-mean-lr also nicely complements the intuition that the more neurons are in each bin, the more "important" this bin should be. This is expressed through the weights $\frac{|\B|}{\sum_{\beta^\prime} |\B^\prime|}$.

<!-- The weighting also shows up if we express the overlap with respect to the mean-field $H_\alpha(t)$:
$$
\tilde m_\mu(t) = \frac{1}{\sum_{\alpha} |\A|} \sum_\alpha |\A| G_{\mu,\alpha} \phi(H_\alpha(t))
$$ -->

We have hereby shown that sampling neurons in the 2D embeddings, and binning the correponding mappings (downsampling), yields the concept of "mean patterns" from @eq-mean-lr-vectors.

## Numerical examples of methods for estimating the $[0,1]$ connectivity kernel

For the sake of illustration, let us compare the connectivity kernels obtained using different methods in @fig-kernels.

* @fig-kernels-samplingR2 is the method described in this appendix, based on sampling $N$ neurons in the PDF space $\R^2$, applying the mapping on the obtained positions $\vec{z_i}$, the applying the binning (downsampling). We see, through the empty rows in the connectivity, that due to finite-$N$ effect, locations where $\rho$ is small are not represented, and, as a consequence, no neurons are present in the corresponding 1D bin. When this happens, we set these rows to zero.
* To obtain @fig-kernels-gridR2, we make the approximation that the support of $\rho(z_1, z_2)$ is compact in $[-4,4]\times[-4,4]$, which of course is not true, but in the case of a Gaussian distribution this is a reasonable approximation, because $\CDF(4) \approx 0.999968$. Defining $\tilde \rho = \rho \circ S^{-1}$, the grid is numerically renormalized so that despite the compact support approximation, the probabilities sum up to one, and the connectivity matrix is modified so that the columns are weighed by $\tilde \rho(\beta)$:
  $$
  \tilde J_{\alpha\beta} = \tilde \rho(\beta) \sum_{\mu=1}^p \tilde F_{\mu \alpha} \tilde G_{\mu \beta}
  $$
  We see that compared to @fig-kernels-samplingR2, using the grid method allows us to "sample" the locations where the density is small, and this results in less "whited-out" rows. The vertical bands of white correspond to the reweighing of the columns with $\tilde \rho(\beta)$.
* Finally, @fig-kernels-grid01 shows the kernel obtained from applying the grid method in the CDF $[0,1]^2$ described in @sec-mapping-cdf. We see that the connectivity seems much more "uniform", because each point of the grid has "equal weight" in the CDF space.

::: {#fig-kernels layout-ncol=3}

![Sampling neurons $\vec{z_i} \sim \rho(z_1, z_2)$ and binning the corresponding mappings](../chapters/figures/sampling_R2_nogrid_kernel.png){#fig-kernels-samplingR2}

![Grid method in a subset of the PDF space $[-4,4]\times[-4,4] \subset \R^2$](../chapters/figures/sampling_R2_kernel.png){#fig-kernels-gridR2}

![Grid method in the CDF space $[0,1]^2$](../chapters/figures/sampling_R2_01_kernel.png){#fig-kernels-grid01}

(TODO) Comparison of methods for computing the connectivity matrices $\tilde J_{\alpha\beta}$ approximating the kernel $\tilde w(\alpha,\beta)$.
:::

### 2023-06-19 manuscript/chapters/1_intro.qmd
all is good :3

@@
 
-![(TODO add phi, add z instead of xi) Every neuron in a rate neural network has a position in the $p$-dimensional embedding space](figures/neural_field_discrete.png)
+![(TODO add phi, add z instead of xi, arrow only in one direction) Every neuron in a rate neural network has a position in the $p$-dimensional embedding space](figures/neural_field_discrete.png)
 
@@
 
-[^1]: Rigorous proofs of convergence are still an active area of study @CheDua19 and are not the subject of this thesis.
+[^1]: Rigorous proofs of convergence are still an active area of study [@CheDua19;@AGATHENERINE202286] and are not the subject of this thesis.
 

### 2023-06-19 manuscript/chapters/2_neural_fields.qmd
little kitty be vibing
all is good :3

@@
 {{< include ../macros.tex >}}
+
+```{python}
+import matplotlib as mpl, matplotlib.pyplot as plt, matplotlib.colors as mplcolors, matplotlib.patheffects as pe, matplotlib.ticker as ticker
+import numpy as np
+import sys
+sys.path.append('../../notebooks')  # import neurodyn
+from neurodyn import *
+from matplotlib_tufte import *
+setup()
+
+mpl.use('svg')
+
+def mktrans(fig, ax):
+	"""Makes figure and axes transparent"""
+	fig.patch.set_facecolor('#FFFFFF00')
+	ax.patch.set_facecolor('#FFFFFF00')
+```
 :::
@@
 This corresponds to a sharper difference between the "inactive" (low potential) and the "active" (high potential) neurons.
-3. The spectrum of $K$ is composed of $p$ eigenvalues that depend on the fixed point, and $N-p$ eigenvalues $\lambda_\perp = -1$ corresponding to the orthogonal complement of the pattern fixed points.
+3.
+The spectrum of $K$ is composed of $p$ eigenvalues (labeled $\lambda_\mu$ for all $\mu\in\{1,\cdots,p\}$) that depend on the fixed point, and $N-p$ eigenvalues $\lambda_\perp = -1$ corresponding to the orthogonal complement of the pattern fixed points.
 
@@
 
-![(TODO add spectrum for zero fixed point) Spectrum of $K$ for $N=64$ for the pattern fixed point](figures/spectrum_N=64.png){#fig-spectrum}
+```{python}
+#| label: fig-spectrum
+#| fig-cap: Spectrum of $K$ for $N=64$ and $p=2$.
+The numerical estimation (black) of the eigenvalues is close to the analytical ($N \to \infty$) derivation (red).
+Eigenvalues above the dotted gray line(positive values) correspond to unstable fixed points, and below (negative values) correspond to stable fixed points.
+Only the
+#| fig-subcap:
+#|   - Pattern (stable) fixed point
+#|   - Zero (unstable) fixed point
+#| layout-ncol: 2
+#| echo: false
+
+dsigmoid = lambda x: sigmoid(x) * (1 - sigmoid(x))
+
+def plot_spectrum(hstar_str: str):
+	params = LowRankRNNParams.new_valentin(p=2, N=256, phi=sigmoid, I_ext=zero, exclude_self_connections=False)
+	if hstar_str == 'pattern':
+		hstar = params.F[:, 0]
+	elif hstar_str == 'zero':
+		hstar = np.zeros(params.N)
+	dphi = dsigmoid(hstar)
+	l = 1/params.N * np.einsum('im,im->m', params.F, params.G * dphi[:, None]) - 1
+	l.sort()
+	J = np.einsum('im,jm->ij', params.F, params.G * dphi[:, None]) / params.N - np.eye(params.N)
+
+	fig, ax = plt.subplots()
+	mktrans(fig, ax)
+	ax.axhline(0, color='gray', linestyle='--')
+	ax.plot(np.sort(np.linalg.eigvals(J).real)[::-1][:8], marker='o', clip_on=False)
+	ax.plot([7,8,8.5], np.sort(np.linalg.eigvals(J).real)[::-1][8:11], linestyle='--', color='k', clip_on=False)
+	ax.plot(l[::-1], color='r', marker='o', clip_on=False)
+	ax.set_ylim((-1, 0.5))
+	ax.set_xlim((0, 8.5))
+	ax.set_ylabel('Eigenvalues')
+	breathe()
+	return fig, ax
+
+# pattern fixed points
+fig, ax = plot_spectrum('pattern')
+display(fig)
+
+# zero fixed point
+fig, ax = plot_spectrum('zero')
+display(fig)
+```
 
@@
 \end{aligned}
-$$
+$$ {#eq-hu-wu}
 
@@
 
-![(TODO) A grid in $[0,1]^2$ and the corresponding samples in $\R^2$.
+![(TODO not color) A grid in $[0,1]^2$ and the corresponding samples in $\R^2$.
 The numeric density of points in $\R^2$ approaches the normal distribution as the grid becomes finer.](figures/sampling_R2_01.png){#fig-sampling-R2-012}

### 2023-06-19 manuscript/chapters/3_mappings.qmd
all is good :3

@@
 For these reasons, we can see why the composition of $w_U$ with a possibly non-differentiable mapping $S$ might be problematic.
-Since the only thing we know about $S$ is that it is measurable and bijective, *there is no guarantee that the regularity of the kernel is conserved through composition with the mapping $S$.*
+Since the only thing we know about $S$ is that it is measurable and bijective, *there is no guarantee that the regularity (continuity and the differentiability) of the kernel is conserved through composition with the mapping $S$.*
 
@@
 ### Mappings of $[0,1]$ into $[0,1]^2$
+
+TOOD : merge with previous
 
@@
 [^edges]:
-  We note that strictly speaking, $S^n$ maps $(0., 0.)$ onto $0$, and the point
-  $$
-  S^n(1., 1.) = S^n(0.111\cdots, 0.111\cdots) = 0.111\cdots = 1
-  $$
-  is never reached at finite $n$.
-Numerically, we encounter the problem that the inverse of the Gaussian CDF, used to compute the connectivity, goes to infinity when $v_\mu$ (one of the components of $(v_1, v_2) = S^{-1}(\alpha)$) goes to $0$ or $1$.
+  We note that in our definition, $S^n$ maps $(0., 0.)$ onto $0$.
+Numerically, we encounter the problem that the inverse of the Gaussian CDF, used to compute the connectivity, goes to infinity when $v_\mu$ (one of the components of $(v_1, v_2) = S^{-1}(\alpha)$) goes to $0$ (or $1$).
 To avoid this, we offset the squares (by $2^{-n-1}$ along each dimension), so that geometrically their position is not indexed by their lower-left corner, but rather their center.
 Analytically, this is not a problem, because the kernel defined in @eq-hu-wu diverges only linearly, compared to the exponential cutoff of the Gaussian distribution, so that the integrand still vanishes at infinity.
-
-<!-- TODONE NOTE : talk about how we center the bins instead of taking the corners.
-this is not super important in the $n\to\infty$ limit, but numerically at 0 and 1 the $\mathrm{CDF}^{-1}$ goes to infty.
-at infinity, the connectivity kernel is infinite but we have zero neurons (zero measure, linear growth of kernel vs exponential tail of the gaussian) : effectively the recurrent current is zero and we can safely ignore these effects.
--->
 
@@
 This is often justified (particularly with the Z-mapping and Column mapping introduced later), because the finite-$n$ truncations $(v_1^{(n)}, v_2^{(n)})$ result in finite truncations of $S$, therefore $S(v_1^{(n)}, v_2^{(n)}) = S^n(v_1^{(n)}, v_2^{(n)})$.
-
-<!-- TODONE NOTE : might need to talk about how $S(v^{(n)}_1,v^{(n)}_2) = S^n(v^{(n)}_1,v^{(n)}_2)$ ?
-so basically the limit $n\to\infty$ is the limit where we take infinite precision in the input.
-Or we can drop the $n$ on the inputs, $\alpha^{(n)} = S^n(v^{(n)}_1,v^{(n)}_2) = S^n(v_1,v_2)$ (i think this true just for pointwise convergent mappings) -->
 
@@
 A problem in GPU computing is that of texture locality: sampling textures often involves reading data corresponding to a small 2D region, and therefore to improve speed (minimize cache misses), the 2D data should be packed close together in the 1D memory.
-One way of ordering the 2D texture data in the 1D memory is by using the Z-order curve (also known as "Morton mapping" or "Lebesgue curve" @Sagan1994).
+One way of ordering the 2D texture data in the 1D memory is by using the Z-order curve (also known as "Morton mapping", and similar to the "Lebesgue curve" @Sagan1994).
 In this way, points close in 1D *tend* to be tightly packed in 2D.
@@
 We also see that the Column mapping is not local in this sense, since there is a large variation along the vertical direction, which seems to always be "of order one".
-This can be explained by looking at the expression for $C^{-1}_2(\alpha^{(n)}) = 0.b_{n+1} b_{n+2} \cdots b_{2n}$ : small variations of the order of $2^{-n}$ in the value of $\alpha$ always result in variations of order one in $C^{-1}_2$.
+This can be explained by looking at the expression for $C^{-1}_2(\alpha^{(n)}) = 0.b_{n+1} b_{n+2} \cdots b_{2n}$ : small variations of the order of $2^{-n}$ in the value of $\alpha$ always result in variations of order one in $C^{-1}_2$.e
 
@@
 Numerically, there seems to be an infinite number of discontinuities, but the sizes of these discontinuities are, for the most part, small, so that *on average* they vanish as $n\to\infty$.
-We won't go further into detail about this,[^zmapping-notbijective] but this gives some intuition for the arguments presented in @sec-locality.
+We won't go further into detail about this, but this gives some intuition for the arguments presented in @sec-locality.
 
@@
 
+<!--
 [^zmapping-notbijective]:
   We refer to @Sagan1994 for an in-depth study, and in particular chapter 6, in which the author shows that the Z-curve converges to the Lebesgue curve.<br> 
-  <!-- which is a Peano-type function (continuous and surjective from $[0,1]$ into $[0,1]^2$).<br> -->
+  <!-- which is a Peano-type function (continuous and surjective from $[0,1]$ into $[0,1]^2$).<br> 
   Despite what can often be found in online resources, the Z-curve is *not* a bijection.
-  <!-- because the added injectivity would violate Netto's theorem. -->
+  <!-- because the added injectivity would violate Netto's theorem. 
   For every point in $[0,1]^2$, there are at most 4 distinct preimages in $[0,1]$.
@@
   For the following, we will consider the Z-curve to be bijective by disallowing expansions of the type $0.1000\cdots=0.0111\cdots$.
-  <!-- We note that in @Sagan1994 many other Peano curves (continuous surjections $[0,1]\mapsto [0,1]^2$) are presented (albeit more difficult to implement numerically). -->
-  <!-- Nicole: I don't say that the Lebesgue curve is continuous, because that opens up a whole other can of worms that I don't want to go into -->
+  <!-- We note that in @Sagan1994 many other Peano curves (continuous surjections $[0,1]\mapsto [0,1]^2$) are presented (albeit more difficult to implement numerically). 
+  <!-- Nicole: I don't say that the Lebesgue curve is continuous, because that opens up a whole other can of worms that I don't want to go into 
+<!-- Nicole: OK, after discussion with valentin, our definition of the Z-mapping is not exactly the same as Lebesgue curve.
+  Z-mapping might seem to approximate the Lebesgue curve, but it is not the same.
+  Lebesgue is defined on the Cantor set (the cantor set avoids the discontinuities, e.g. around alpha=0.5), then extended linearly, thus continuous
+  Our Z-mapping is defined through the bits OBTAINED FROM THE INDICATOR FUNCTIONS; and so it works on NON-AMBIGUOUS binary representations. Thus, our Z-mapping is a bijection.
+  Our Z-mapping is also not continuous, consider alpha=0.5. So we don't have a problem with Netto's theorem.
+-->
 
@@
 
-### Visualizating of the mappings of the dynamic 2D neural field
+<!-- ### Visualizing of the mappings of the dynamic 2D neural field
 
@@
 
-![(TODO) Animation showing the image of the dynamics 2D neural field through different mappings $S : [0,1]^2 \mapsto [0,1]$](figures/photo_2023-06-10_17-21-19.jpg)
+![(TODO) Animation showing the image of the dynamics 2D neural field through different mappings $S : [0,1]^2 \mapsto [0,1]$](figures/photo_2023-06-10_17-21-19.jpg) -->
 

### 2023-06-19 manuscript/chapters/4_locality.qmd
all is good :3

@@
 We do not address the question of well-posedness of the neural field equations in this thesis; for proofs of the well-posedness of neural field equations, we refer the reader to [@faugeras2008absolute;@faugeras2009persistent;@veltz2009localglobal].
-
-<!-- TODO : we may argue that the Z-curve is not bijective. This complicates the proof. Though, the pre-image of (v_1, v_2) is composed of at most 4 points, and the measurability compensates. -->
 
@@
 
-<!-- Nicole: V_n -> 0 is implied by continuity -->
+The definition of locality can essentially be seen as a weakened version of continuity; since from definition the image of a small neighbourhood through a continuous function is also a small neighbourhood.
+Therefore, continuity of a function $f$ implies that $V_n(f) \xrightarrow{n \to \infty} 0$.
+The reverse is obviously not true, since for instance the metric $V_n$ vanishes when applied to a step function, which is only piecewise continuous.
 
@@
 
-![(TODO change title, change name of mappings) (Numerator of the) average binned variation as a function of $n$.
+![(TODO change title, change name of mappings, remove the theory, but keep upper bound) (Numerator of the) average binned variation as a function of $n$.
 The local mappings have a growth slower than the upper bound $2\times 2^n$.](figures/v_num.png){width=500px #fig-vnum}
@@
 
-Let us consider the Z-mapping, and the two terms compositing its inverse. In the case of $\alpha, \alpha^\prime$ above, we have if $n$ is even :
+Let us consider the Z-mapping, and the two terms compositing its inverse. In the case of $\alpha, \alpha^\prime$ above, we have if $n$ is even:
 
@@
 Z^{-1}_2(\alpha) &= 0.\underbrace{i_{n-2} i_{n-4} \cdots i_0}_\text{$n/2$ bits} b_{n+1} b_{n+3} \cdots b_{2m}
-\end{aligned}of this locality
+\end{aligned}
 $$
@@
 
-<!-- [TODO MATH : prove that it is indeed Lipschitz.
-The main problem I've encountered with this is that these functions are $\textrm{regular function} \circ \textrm{CDF}^{-1}$, and the inverse CDF goes to infinity at 0 and 1].
--->
-
 Applying the Lipschitz assumption yields the following inequality on the integrand:

### 2023-06-19 manuscript/chapters/3_mappings.qmd
continuity

@@
 
-### Mappings of $[0,1]$ into $[0,1]^2$
+@fig-venn shows a visual representation of Netto's theorem.
+The 3-way intersection of the Venn diagram represents that functions which are both bijective (surjective and injective) and continuous do not exist.
+For the three remaining 2-way intersections, we annotate, without going into detail, the types of functions which they represent (see [@kharazishvili_strange_2017] for Cantor and Peano functions, [@sagan_space-filling_1994] for Peano and Jordan functions).
 
-TOOD : merge with previous
+<!-- ### Mappings of $[0,1]$ into $[0,1]^2$
 
@@
 Since bijections are functions that are both injective and surjective, we can represent Netto's theorem as a 3-way Venn diagram in @fig-venn.
+-->
 
@@
 
-The 3-way intersection of the Venn diagram represents that functions which are both bijective (surjective and injective) and continuous do not exist.
-
-For the three remaining 2-way intersections, we cite, without going into detail, the types of functions which they represent (see [@kharazishvili_strange_2017] for Cantor and Peano functions, [@sagan_space-filling_1994] for Peano and Jordan functions):
-
+<!--
 * Cantor functions are bijections $[0,1] \mapsto [0,1]^2$. However, they are not continuous, and thereby break the continuity of the kernel.
+<!-- Nicole: ACTUALLY NO THIS IS NOT A PROBLEM !! OUR Z-MAPPING IS NOT CONTINUOUS, BUT LOCALITY STILL SAVES US, SO THAT IT IS CONTINUOUS "ON AVERAGE"
 * Jordan curves are continuous injections $[0,1] \mapsto [0,1]^2$. Since they are not surjective, these curves cannot entirely "fill" the $[0,1]^2$ space, and we will not use them.
-<!-- Nicole : see my notebook about Jordan functions and their nonzero Jordan measure -->
+<!-- Nicole : see my notebook about Jordan functions and their nonzero Jordan measure
 * Peano curves are continuous surjections $[0,1] \mapsto [0,1]^2$.
@@
 Therefore, Peano functions seem like good candidates for the mapping $S^{-1}$.
+-->
 

### 2023-06-19 manuscript/chapters/4_locality.qmd
continuity

@@
 
-The definition of locality can essentially be seen as a weakened version of continuity; since from definition the image of a small neighbourhood through a continuous function is also a small neighbourhood.
-Therefore, continuity of a function $f$ implies that $V_n(f) \xrightarrow{n \to \infty} 0$.
+The definition of locality can essentially be seen as a weakened version of continuity, or "continuity on average".
+A continuous function $f$ evidently verifies $V_n(f) \xrightarrow{n \to \infty} 0$, since small neighbourhoods are mapped to small neighbourhoods.
 The reverse is obviously not true, since for instance the metric $V_n$ vanishes when applied to a step function, which is only piecewise continuous.

### 2023-06-19 manuscript/chapters/2_neural_fields.qmd
*clicks* noice

@@
 We should however note that the results obtained in @sec-fixedpoints are valid in the $N \to \infty$ limit.
-A numerical analysis shows in @fig-spectrum that already at $N=64$, the analytic eigenvalues match the numerical eigenvalues closely, and this correspondence improves when we take larger $N$ (typically we will for the rest of this thesis take $N > 10^3$).
+A numerical analysis shows in @fig-spectrum that at $N=250$, the numerical eigenvalues approximate the analytical eigenvalues, and this correspondence improves when we take larger $N$ (typically we will for the rest of this thesis take $N > 10^3$).
 
@@
 #| label: fig-spectrum
-#| fig-cap: Spectrum of $K$ for $N=64$ and $p=2$.
-The numerical estimation (black) of the eigenvalues is close to the analytical ($N \to \infty$) derivation (red).
-Eigenvalues above the dotted gray line(positive values) correspond to unstable fixed points, and below (negative values) correspond to stable fixed points.
-Only the
+#| fig-cap: Spectrum of $K$ for $N=250$ and $p=2$.
+The numerical estimation (black, eigenvalues $\lambda_\mu$ and $\lambda_\perp=-1$) of the eigenvalues is close to the analytical ($N \to \infty$) derivation (red, only eigenvalues $\lambda_\mu$).
+Eigenvalues above the dotted gray line (positive values) correspond to unstable fixed points, and below (negative values) correspond to stable fixed points.
 #| fig-subcap:
@@
 def plot_spectrum(hstar_str: str):
-	params = LowRankRNNParams.new_valentin(p=2, N=256, phi=sigmoid, I_ext=zero, exclude_self_connections=False)
-	if hstar_str == 'pattern':
-		hstar = params.F[:, 0]
-	elif hstar_str == 'zero':
-		hstar = np.zeros(params.N)
+	params = LowRankRNNParams.new_valentin(p=2, N=250, phi=sigmoid, I_ext=zero, exclude_self_connections=False)
+	if hstar_str == 'pattern': hstar = params.F[:, 0]
+	elif hstar_str == 'zero': hstar = np.zeros(params.N)
 	dphi = dsigmoid(hstar)
-	l = 1/params.N * np.einsum('im,im->m', params.F, params.G * dphi[:, None]) - 1
-	l.sort()
 	J = np.einsum('im,jm->ij', params.F, params.G * dphi[:, None]) / params.N - np.eye(params.N)
+	eig_num = np.sort(np.linalg.eigvals(J).real)[::-1]
+	eig_ana = { 'zero': 0.19061, 'pattern': -0.28090 }[hstar_str]  # see appendix A
 
@@
 	ax.axhline(0, color='gray', linestyle='--')
-	ax.plot(np.sort(np.linalg.eigvals(J).real)[::-1][:8], marker='o', clip_on=False)
-	ax.plot([7,8,8.5], np.sort(np.linalg.eigvals(J).real)[::-1][8:11], linestyle='--', color='k', clip_on=False)
-	ax.plot(l[::-1], color='r', marker='o', clip_on=False)
+	ax.plot(eig_num[:8], marker='o', clip_on=False)
+	ax.plot([7,8,8.5], eig_num[8:11], linestyle='--', color='k', clip_on=False)
+	ax.plot(np.arange(params.p), np.full(params.p, eig_ana), color='r', marker='o', clip_on=False)
 	ax.set_ylim((-1, 0.5))
@@
 	ax.set_ylabel('Eigenvalues')
+	ax.set_xlabel('nth eigenvalue')
 	breathe()
 	return fig, ax
+
+# %matplotlib inline
 
@@
 ::: {.content-visible when-format="html"}
-![(TODO remove overlap, change notation, color for activity) Simulation of the Gaussian network of neurons with $p=1$.
-Every neuron is represented by a dot, of which the height represents its activity level.](figures/embedding_d=1_h0=0_phi=sigmoid.mp4){width=400px #fig-p1}
+![Simulation of the Gaussian network of neurons with $p=1$.
+Every neuron is represented by a dot, of which the height represents its activity level, and the colorbar doubles as a y-axis.](figures/embedding_d=1.mp4){#fig-p1}
 :::
@@
 ::: {.content-visible when-format="pdf"}
-![(TODO PDF remove overlap, change notation, color for activity) Simulation of the Gaussian network of neurons with $p=1$.
-Every neuron is represented by a dot, of which the height represents its activity level.](figures/lol7.jpg){width=400px #fig-p1}
+![(TODO) Simulation of the Gaussian network of neurons with $p=1$.
+Every neuron is represented by a dot, of which the height represents its activity level, and the colorbar doubles as a y-axis.](figures/lol7.jpg){#fig-p1}
 :::
@@
 
-## A cycling neural field {#sec-cyling-nf}
+## A cycling neural field {#sec-cycling-nf}
 

### 2023-06-19 manuscript/chapters/3_mappings.qmd
*clicks* noice

@@
 
-![Non-injectivity of $S^{-1}$ means that two segment populations can refer to the same square population.](figures/photo_2023-06-12_00-46-26.jpg){#fig-non-injection}
+<!-- ![Non-injectivity of $S^{-1}$ means that two segment populations can refer to the same square population.](figures/photo_2023-06-12_00-46-26.jpg){#fig-non-injection}
+
+Nicole: removed. too confusing
+-->
 
@@
 A problem in GPU computing is that of texture locality: sampling textures often involves reading data corresponding to a small 2D region, and therefore to improve speed (minimize cache misses), the 2D data should be packed close together in the 1D memory.
-One way of ordering the 2D texture data in the 1D memory is by using the Z-order curve (also known as "Morton mapping", and similar to the "Lebesgue curve" @Sagan1994).
+One way of ordering the 2D texture data in the 1D memory is by using the Z-order curve (also known as "Morton mapping", and similar to the "Lebesgue curve" @sagan_space-filling_1994, chapter 6).
 In this way, points close in 1D *tend* to be tightly packed in 2D.
@@
 
-![(TODO) Averaging $2^n$ of the $4^n$ segments populations of $S^n$ results in the $2^n = 4^{n/2}$ segments populations of $S^{n/2}$](figures/photo_2023-06-13_12-15-32.jpg){width=400px #fig-averaging}
+![(TODO) Averaging $2^n$ of the $4^n$ segments populations of $S^n$ results in the $2^n$ segments populations](figures/photo_2023-06-13_12-15-32.jpg){width=400px #fig-averaging}
 

### 2023-06-19 manuscript/chapters/4_locality.qmd
*clicks* noice

@@
 I don't think the proof works strictly in our case, because the Z-mapping is not a bijection, but let us assume uwuwuwu
+
+Nicole again: our Z-mapping is not defined in the same way as the Lebesgue curve, because we use indicator functions to take bit representations.
+We are stepwise functions.
+Our Z-mapping is a bijection, and not continuous.
 -->
@@
 
-and that in @sec-cyling-nf, numerical simulations showed that $I^\text{rec}_U(\vec v)$ is a well-behaved function.
+and that in @sec-cycling-nf, numerical simulations showed that $I^\text{rec}_U(\vec v)$ is a well-behaved function.
 

### 2023-06-19 notebooks/neurodyn/_anim.py

@@ -8,12 +8,12 @@ from ._rnn import LowRankRNN, Result
 from ._plot import plot_overlap_trajectory, scale_lightness
 from ._overlap import overlap
 
-__all__ = ['animate1d', 'animate2d']
+__all__ = ['animate2d']
 
 def midpoints(arr):
 	return (arr[:-1] + arr[1:]) / 2
 
-def animate1d(rnn: LowRankRNN, res: Result, outpath: Path, time_stride: int = 1, Nmax = 1500):
+def animate1d_old(rnn: LowRankRNN, res: Result, outpath: Path, time_stride: int = 1, Nmax = 1500):
 	idt = 0
 	activity = rnn.phi(res.h[:Nmax])
 	cmap = mpl.colormaps['RdBu_r']

### 2023-06-19 notebooks/neurodyn/_plot.py

@@ -1,6 +1,7 @@
 """Plotting utilities"""
 
 import matplotlib as mpl, matplotlib.pyplot as plt, matplotlib.ticker as ticker, matplotlib.tri as tri, matplotlib.collections as collections
+from matplotlib.colors import LinearSegmentedColormap
 import numpy as np
 import colorsys
 from matplotlib.axes import Axes
@@ -13,7 +14,8 @@ __all__ = [
 	'plot_overlap_phase2D',
 	'plot_2D_embedding_contour', 'plot_2D_embedding_scatter',
 	'plot_2D_to_1D_mapping',
-	'add_headers', 'scale_lightness'
+	'add_headers', 'scale_lightness',
+	'cmap_bi', 'cmap_bi_r', 'cmap_trans',
 ]
 
 def _unwrap_figax(figax: tuple[Figure, Axes] | None = None, **kwargs) -> tuple[Figure, Axes]:
@@ -149,6 +151,9 @@ def plot_2D_to_1D_mapping(
 
 	return (fig, axes), scat1d, scat2d
 
+cmap_bi = LinearSegmentedColormap.from_list("", list(reversed(["#ff0080","#ff0080","#a349a4","#0000ff","#0000ff"])))
+cmap_bi_r = LinearSegmentedColormap.from_list("", ["#ff0080","#ff0080","#a349a4","#0000ff","#0000ff"])
+cmap_trans = LinearSegmentedColormap.from_list("", ["#55CDFC","#55CDFC","#FFFFFF","#FFFFFF","#F7A8B8","#F7A8B8"])
 
 def add_headers(
 	fig=None, axes=None,

### 2023-06-20 manuscript/chapters/2_neural_fields.qmd
finish figures for chap. 2

@@
 ![Simulation of the Gaussian network of neurons with $p=1$.
-Every neuron is represented by a dot, of which the height represents its activity level, and the colorbar doubles as a y-axis.](figures/embedding_d=1.mp4){#fig-p1}
+Every neuron is represented by a dot, of which the height represents its activity level, and the colorbar doubles as a y-axis.](figures/embedding_p=1.mp4){#fig-p1}
 :::
@@
 ::: {.content-visible when-format="html"}
-![(TODO remove overlap, change notation, scatter only) Simulation of the Gaussian network of neurons with $p=2$.
-Every neuron is represented by a dot, of which the height represents its activity level.](figures/embedding_d=2_h0=0.mp4){width=500px #fig-p2}
+![Simulation of the Gaussian network of neurons with $p=2$.
+Every neuron is represented by a dot, of which the height represents its activity level.
+Projections are shown by gray points.](figures/embedding_p=2.mp4){width=500px #fig-p2}
 :::
@@
 ::: {.content-visible when-format="pdf"}
-![(TODO PDF remove overlap, change notation, scatter only) Simulation of the Gaussian network of neurons with $p=2$.
-Every neuron is represented by a dot, of which the height represents its activity level.](figures/lol7.jpg){width=500px #fig-p2}
+![(TODO PDF) Simulation of the Gaussian network of neurons with $p=2$.
+Every neuron is represented by a dot, of which the height represents its activity level.
+Projections are shown by gray points.](figures/lol7.jpg){width=500px #fig-p2}
 :::
@@
 
-![(TODO animation in 3D, rename non-cycling to original, number from one) Intuition for the cycling behavior by considering the recurrent currents](figures/cycling_recurrent.png){#fig-cycling-irec}
+```{python}
+#| label: fig-cycling-irec
+#| fig-cap: Intuition for the cycling behavior.
+We consider the network in the state $h(z_1,z_2)=z_1$, and look at the recurrent currents.
+The recurrent currents align with the activity for the original neural field, and the effective change in potential is zero everywhere.
+For the cycling neural field, the recurrent currents are orthogonal to the activity, such that with the decay term, the effective change leads to a rotation in the counter-clockwise direction, towards the state $h(z_1,z_2)=z_2$.
+#| echo: false
+
+from matplotlib.image import NonUniformImage
+
+nrec = 6
+rnn_nocyc = LowRankRNN.new_sampled_valentin(
+	phi=sigmoid, mapping_samples=ReshapeMapping.new_nrec(nrec=nrec).inverse_samples(),
+	I_ext=zero, exclude_self_connections=False
+)
+rnn_cyc = LowRankCyclingRNN.new_sampled_valentin(
+	phi=sigmoid, mapping_samples=ReshapeMapping.new_nrec(nrec=nrec).inverse_samples(),
+	I_ext=zero, exclude_self_connections=False, delta=10, shift=1
+)
+rnn_cyc.h_lagging = lambda t, h: h  # HACK : make the recurrent currents work for t=0
+
+fig, axes = plt.subplots(ncols=3, nrows=2, figsize=(3*4-3, 2*3), layout='constrained')
+
+for i, rnn in enumerate([rnn_nocyc, rnn_cyc]):
+	axes[i, 0].set_aspect('equal')
+	axes[i, 0].set_xlabel('$z_1$')
+	axes[i, 0].set_ylabel('$z_2$')
+	shape = (2**nrec, 2**nrec)
+	im = axes[i, 0].pcolormesh(
+		rnn.F[:, 0].reshape(shape), rnn.F[:, 1].reshape(shape), rnn.phi(rnn.F[:, 0]).reshape(shape),
+		cmap=cmap_bi, vmin=0, vmax=1,
+		rasterized=True  # enable rasterization to prevent 50Mo SVG files, lol
+	)
+	breathe(axes[i, 0])
+	if i == 1:
+		plt.colorbar(im, ax=axes[i, 0], orientation='horizontal', label='$\\phi(h(z_1,z_2))$')
+
+	axes[i, 1].set_aspect('equal')
+	axes[i, 1].set_xlabel('$z_1$')
+	axes[i, 1].set_ylabel('$z_2$')
+	im = axes[i, 1].pcolormesh(
+		rnn.F[:, 0].reshape(shape), rnn.F[:, 1].reshape(shape), rnn.I_rec(0, rnn.F[:, 0]).reshape(shape),
+		cmap='coolwarm', vmin=-2, vmax=2,
+		rasterized=True  # enable rasterization to prevent 50Mo SVG files, lol
+	)
+	breathe(axes[i, 1])
+	if i == 1:
+		plt.colorbar(im, ax=axes[i, 1], orientation='horizontal', label='$I^\mathrm{rec}(z_1, z_2)$')
+
+	axes[i, 2].set_aspect('equal')
+	axes[i, 2].set_xlabel('$z_1$')
+	axes[i, 2].set_ylabel('$z_2$')
+	im = axes[i, 2].pcolormesh(
+		rnn.F[:, 0].reshape(shape), rnn.F[:, 1].reshape(shape), rnn.dh(0, rnn.F[:, 0]).reshape(shape),
+		cmap=cmap_pan, vmin=-2, vmax=2, shading='gouraud',
+		rasterized=True  # enable rasterization to prevent 50Mo SVG files, lol
+	)
+	breathe(axes[i, 2])
+	if i == 1:
+		plt.colorbar(im, ax=axes[i, 2], orientation='horizontal', label='$\partial_t h(z_1, z_2)$')
+
+add_headers(axes=axes.flat, row_headers=['Original neural field', 'Cycling neural field'], fontsize='medium', row_pad=8)
+add_headers(axes=axes.flat, col_headers=['Neural activity', 'Recurrent currents', 'Effective change'], fontsize='medium', col_pad=8)
+
+for ax in axes.flat:
+	mktrans(fig, ax)
+
+# %matplotlib inline
+display(fig)
+```
 
@@
 ![A cycling neural field with $\delta=6$ and initial condition $h(z_1, z_2, t < 0) = z_1$.
-The latent trajectory is also computed, and demonstrates the oscillatory behaviour.](figures/embedding_d=2_h0=0_cycling_delta=6.mp4){#fig-cycling-anim}
+The latent trajectory is also computed, and demonstrates the oscillatory behaviour.](figures/embedding_p=2_cycling.mp4){#fig-cycling-anim}
 :::
@@
 
-![(TODO) CDF of the Gaussian distribution](figures/1920px-Normal_Distribution_CDF.svg.png){width=400px #fig-gaussian-cdf}
+```{python}
+#| label: fig-gaussian-cdf
+#| fig-cap: CDF of the Gaussian distribution
+#| echo: false
+#| 
+from scipy import stats
+
+z = np.linspace(-4.5, 4.5, 200)
+fig, ax = plt.subplots()
+ax.plot(z, stats.norm.cdf(z), clip_on=False)
+# ax.set_xlim((-4, 4))
+ax.set_ylim((0, 1))
+ax.set(xlabel='$z$', ylabel='$\\mathrm{CDF}(z)$')
+ax.set_yticks([0.0, 0.25, 0.5, 0.75, 1.0])
+ax.grid(clip_on=False)
+# ax.grid(clip_on=False, axis='y')
+# breathe(ax, which='x')
+# breathe(ax, which='y', pad_frac=0.1)
+breathe(ax)
+mktrans(fig, ax)
+display(fig)
+```
 
@@
 Defining the neural field in $[0,1]^p$ also addresses the problem that we tried to hide in @sec-gridmethod relating to the difficulties of defining a grid on $\Rp$.
-With the integral on $[0,1]^p$, we can simply use a uniform grid, which maps back to samples in $\Rp$ by using the inverse CDF, as shown in @fig-sampling-R2-012.
+With the integral on $[0,1]^p$, we can simply use a uniform grid, which maps back to samples in $\Rp$ by using the inverse CDF, as shown in @fig-grid-correspondance.
 
-![(TODO not color) A grid in $[0,1]^2$ and the corresponding samples in $\R^2$.
-The numeric density of points in $\R^2$ approaches the normal distribution as the grid becomes finer.](figures/sampling_R2_01.png){#fig-sampling-R2-012}
+```{python}
+#| label: fig-grid-correspondance
+#| fig-cap: A grid in $[0,1]^2$ and the corresponding samples in $\R^2$.
+The numeric density of points in $\R^2$ approaches the normal distribution as the grid becomes finer (here, the mesh size is $2^{-4}$ along each dimension).
+#| echo: false
+
+nrec = 4
+mapping = ColumnMapping.new_nrec(nrec=nrec)
+Y = mapping.inverse_samples()
+Z = stats.norm.ppf(Y)
+
+fig, axes = plt.subplots(ncols=2, figsize=(8, 3))
+axes[0].scatter(*Y.T, c='k', s=15)
+axes[1].scatter(*Z.T, c='k', s=15)
+axes[0].set_xlim((0,1))
+axes[0].set_ylim((0,1))
+axes[0].set_aspect('equal')
+axes[1].set_aspect('equal')
+axes[0].set_title('In $[0,1]^2$ (CDF space)', fontsize='medium')
+axes[1].set_title('In $\\mathbb{R}^2$ (PDF space)', fontsize='medium')
+breathe(axes[0])
+breathe(axes[1])
+mktrans(fig, axes[0])
+mktrans(fig, axes[1])
+display(fig)
+```
 

### 2023-06-20 notebooks/neurodyn/_plot.py

@@ -15,7 +15,7 @@ __all__ = [
 	'plot_2D_embedding_contour', 'plot_2D_embedding_scatter',
 	'plot_2D_to_1D_mapping',
 	'add_headers', 'scale_lightness',
-	'cmap_bi', 'cmap_bi_r', 'cmap_trans',
+	'cmap_bi', 'cmap_bi_r', 'cmap_pan', 'cmap_pan_r', 'cmap_trans',
 ]
 
 def _unwrap_figax(figax: tuple[Figure, Axes] | None = None, **kwargs) -> tuple[Figure, Axes]:
@@ -153,6 +153,8 @@ def plot_2D_to_1D_mapping(
 
 cmap_bi = LinearSegmentedColormap.from_list("", list(reversed(["#ff0080","#ff0080","#a349a4","#0000ff","#0000ff"])))
 cmap_bi_r = LinearSegmentedColormap.from_list("", ["#ff0080","#ff0080","#a349a4","#0000ff","#0000ff"])
+cmap_pan = LinearSegmentedColormap.from_list("", list(reversed(["#fa1a8c","#fcd900","#3aafff"])))
+cmap_pan_r = LinearSegmentedColormap.from_list("", list((["#fa1a8c","#fcd900","#3aafff"])))
 cmap_trans = LinearSegmentedColormap.from_list("", ["#55CDFC","#55CDFC","#FFFFFF","#FFFFFF","#F7A8B8","#F7A8B8"])
 
 def add_headers(

### 2023-06-20 manuscript/appendices/B_mean_patterns.qmd
so... many.... figures !

@@
 \def\B{\mathcal{B}}
+
+```{python}
+import matplotlib as mpl, matplotlib.pyplot as plt, matplotlib.colors as mplcolors, matplotlib.patheffects as pe, matplotlib.ticker as ticker
+import numpy as np
+import sys
+sys.path.append('../../notebooks')  # import neurodyn
+from neurodyn import *
+from matplotlib_tufte import *
+setup()
+
+mpl.use('svg')
+
+def mktrans(fig, ax):
+	"""Makes figure and axes transparent"""
+	fig.patch.set_facecolor('#FFFFFF00')
+	ax.patch.set_facecolor('#FFFFFF00')
+```
 :::
@@
 
-* @fig-kernels-samplingR2 is the method described in this appendix, based on sampling $N$ neurons in the PDF space $\R^2$, applying the mapping on the obtained positions $\vec{z_i}$, the applying the binning (downsampling).
+* @fig-kernels-1 is the method described in this appendix, based on sampling $N$ neurons in the PDF space $\R^2$, applying the mapping on the obtained positions $\vec{z_i}$, the applying the binning (downsampling).
 We see, through the empty rows in the connectivity, that due to finite-$N$ effect, locations where $\rho$ is small are not represented, and, as a consequence, no neurons are present in the corresponding 1D bin.
 When this happens, we set these rows to zero.
-* To obtain @fig-kernels-gridR2, we make the approximation that the support of $\rho(z_1, z_2)$ is compact in $[-4,4]\times[-4,4]$, which of course is not true, but in the case of a Gaussian distribution this is a reasonable approximation, because $\CDF(4) \approx 0.999968$.
+* To obtain @fig-kernels-2, we make the approximation that the support of $\rho(z_1, z_2)$ is compact in $[-4,4]\times[-4,4]$, which of course is not true, but in the case of a Gaussian distribution this is a reasonable approximation, because $\CDF(4) \approx 0.999968$.
 Defining $\tilde \rho = \rho \circ S^{-1}$, the grid is numerically renormalized so that despite the compact support approximation, the probabilities sum up to one, and the connectivity matrix is modified so that the columns are weighed by $\tilde \rho(\beta)$:
@@
   $$
-  We see that compared to @fig-kernels-samplingR2, using the grid method allows us to "sample" the locations where the density is small, and this results in less "whited-out" rows.
+  We see that compared to @fig-kernels-1, using the grid method allows us to "sample" the locations where the density is small, and this results in less "whited-out" rows.
 The vertical bands of white correspond to the reweighing of the columns with $\tilde \rho(\beta)$.
-* Finally, @fig-kernels-grid01 shows the kernel obtained from applying the grid method in the CDF $[0,1]^2$ described in @sec-mapping-cdf.
+* Finally, @fig-kernels-3 shows the kernel obtained from applying the grid method in the CDF $[0,1]^2$ described in @sec-mapping-cdf.
 We see that the connectivity seems much more "uniform", because each point of the grid has "equal weight" in the CDF space.
 
-::: {#fig-kernels layout-ncol=3}
+```{python}
+#| label: fig-kernels
+#| fig-cap: Comparison of methods for computing the connectivity matrices $\tilde J_{\alpha\beta}$ approximating the kernel $\tilde w(\alpha,\beta)$.
+#| fig-subcap: 
+#|   - Sampling neurons $\vec{z_i} \sim \rho(z_1, z_2)$ and binning the corresponding mappings.
+#|   - Grid method in a subset of the PDF space $[-4,4]\times[-4,4] \subset \R^2$.
+#|   - Grid method in the CDF space $[0,1]^2$.
+#| layout-ncol: 3
+#| echo: false
 
-![Sampling neurons $\vec{z_i} \sim \rho(z_1, z_2)$ and binning the corresponding mappings](../chapters/figures/sampling_R2_nogrid_kernel.png){#fig-kernels-samplingR2}
+import matplotlib.colors as mplcolors
 
-![Grid method in a subset of the PDF space $[-4,4]\times[-4,4] \subset \R^2$](../chapters/figures/sampling_R2_kernel.png){#fig-kernels-gridR2}
+nrec = 5
 
-![Grid method in the CDF space $[0,1]^2$](../chapters/figures/sampling_R2_01_kernel.png){#fig-kernels-grid01}
+def plot_Jab(Jab):
+	amp = np.abs(J_ab).max()
+	fig, ax = plt.subplots(dpi=200)
+	im = ax.imshow(J_ab,
+		cmap='seismic', norm=mplcolors.TwoSlopeNorm(vcenter=0.0, vmin=-amp, vmax=amp),
+		extent=((0,1,0,1)), rasterized=True, origin='upper'
+		)
+	ax.set(
+		xlim=(0,1), ylim=(0,1), aspect='equal'
+	)
+	fig.colorbar(im)
+	mktrans(fig, ax)
+	breathe(ax)
+	return fig, ax
 
-(TODO) Comparison of methods for computing the connectivity matrices $\tilde J_{\alpha\beta}$ approximating the kernel $\tilde w(\alpha,\beta)$.
-:::
+J_ab = LowRankRNNParams.new_valentin(
+	p=2, N=100_000, phi=sigmoid, I_ext=zero, exclude_self_connections=False
+).to_binmapped(ZMapping.new_nrec(nrec=nrec)).to_dense().J
+fig, ax = plot_Jab(J_ab)
+display(fig)
+
+phi = sigmoid
+rho = lambda z: 1/np.sqrt(2*np.pi)*np.exp(-np.linalg.norm(z, axis=1)**2/2)
+z0 = np.random.default_rng(42).normal(loc=0, scale=1, size=1_000_000)
+phi_z0 = phi(z0)
+a, c = np.mean(phi_z0), np.var(phi_z0)
+def phi_tilde(z):
+	return (phi(z) - a) / c
+mapping = ZMapping.new_nrec(nrec=nrec)  # 4**n = 4**4 = 256 boxes along each axis
+Z = mapping.inverse_samples(bbox=Box(xmin=-4, xmax=4, ymin=-4, ymax=4), centered=True)
+J_ab = LowRankRNNParams(
+	phi=sigmoid, I_ext=zero, exclude_self_connections=False,
+	F=Z, G=phi_tilde(Z) * rho(Z)[:, None]/rho(Z).sum() * len(Z)
+).to_dense().J
+fig, ax = plot_Jab(J_ab)
+display(fig)
+
+mapping = ZMapping.new_nrec(nrec=nrec)
+V = mapping.inverse_samples()
+J_ab = LowRankRNNParams.new_sampled_valentin(
+	phi=sigmoid, mapping_samples=V, exclude_self_connections=False, I_ext=zero
+).to_dense().J
+fig, ax = plot_Jab(J_ab)
+display(fig)
+```

### 2023-06-20 manuscript/chapters/2_neural_fields.qmd
so... many.... figures !

@@
 		rnn.F[:, 0].reshape(shape), rnn.F[:, 1].reshape(shape), rnn.I_rec(0, rnn.F[:, 0]).reshape(shape),
-		cmap='coolwarm', vmin=-2, vmax=2,
+		cmap=cmap_pan, vmin=-2, vmax=2,
 		rasterized=True  # enable rasterization to prevent 50Mo SVG files, lol
@@
 		rnn.F[:, 0].reshape(shape), rnn.F[:, 1].reshape(shape), rnn.dh(0, rnn.F[:, 0]).reshape(shape),
-		cmap=cmap_pan, vmin=-2, vmax=2, shading='gouraud',
+		cmap='coolwarm', vmin=-2, vmax=2, shading='gouraud',
 		rasterized=True  # enable rasterization to prevent 50Mo SVG files, lol

### 2023-06-20 manuscript/chapters/3_mappings.qmd
so... many.... figures !

@@
 mpl.use('svg')
+
+def mktrans(fig, ax):
+	"""Makes figure and axes transparent"""
+	fig.patch.set_facecolor('#FFFFFF00')
+	ax.patch.set_facecolor('#FFFFFF00')
 ```
@@
 
-::: {#fig-averaging-geom layout-ncol=2}
+```{python}
+#| label: fig-averaging-geom
+#| fig-cap: Square populations corresponding to the segment populations before coarse-graining are colored identically when they are the same bin of 1D length $2^{-n}$. Here we show $n=3$.
+#| fig-subcap:
+#|   - Z-mapping
+#|   - Column mapping
+#| layout-ncol: 2
+#| echo: false
 
-![Z-mapping](figures/averaging_z.png)
+from itertools import cycle
 
-![Column mapping](figures/averaging_column.png)
+def plot_averaging_geom(mapping_cls, nrec=3):
+	mapping = mapping_cls.new_nrec(nrec)
+	F = mapping.inverse_samples(centered=True)
+	F_ = F.reshape((2**nrec, 2**nrec, 2))  # bins before coarse-graining
+	values = mapping(F)
 
-(TODO) The corresponding square populations corresponding to the segment populations before coarse-graining.
-:::
+	fig, ax = plt.subplots()
+	ax.plot(*F.T, color='gray')
+	colors = cycle(['#750087','#001997','#06bd01','#ffff00','#fe941e','#ff1e26','#fcaec5','#74d4ea','#603814','#010101','#fbf7f4'])
+	for i in range(2**nrec): ax.scatter(*F_[i, :, :].T, zorder=10, color=next(colors), ec='k')
+	ax.set(
+		xlim=(0,1), ylim=(0,1),
+		xlabel='$v_1$', ylabel='$v_2$', aspect='equal',
+		xticks=[0, 0.25, 0.5, 0.75, 1], yticks=[0, 0.25, 0.5, 0.75, 1])
+	mktrans(fig, ax)
+	breathe(ax)
+	return fig, ax
+
+fig, ax = plot_averaging_geom(ZMapping, nrec=3)
+display(fig)
+
+fig, ax = plot_averaging_geom(ColumnMapping, nrec=3)
+display(fig)
+```
 
@@
 
-1.
-The random mapping can be numerically evaluated as a trivial case of a mapping that does not have locality.
+* @fig-cg-2d-1d-random: The random mapping can be numerically evaluated as a trivial case of a mapping that does not have locality.
 The coarse-graining averages out all the structure of the activity, and the recurrent currents are zero, because the contributions from excitatory and inhibitory populations cancel out.
 This results in the erasure of the cycling behavior in the latent trajectory.
-2.
-The Column mapping shows the projection of the initial condition onto the $z_1$ axis, yielding the sigmoid in the $[0,1]$ embedding.
+* @fig-cg-2d-1d-column: The Column mapping shows the projection of the initial condition onto the $z_1$ axis, yielding the sigmoid in the $[0,1]$ embedding.
 Again, the current currents cancel out and yield zero, and the latent trajectory quickly decays towards $(\kappa_1, \kappa_2) = (0,0)$.
-3.
-The Z-mapping shows that despite the fractal structure of the activity, there is still regularity in the $[0,1]$ embedding, resulting in non-zero recurrent currents.
-In consequence, the latent trajectory shows a clear cycling behavior, that closely matches that of the 2D neural field.
+* @fig-cg-2d-1d-z: The Z-mapping shows that despite the fractal structure of the activity, there is still regularity in the $[0,1]$ embedding, resulting in non-zero recurrent currents.
+In consequence, the latent trajectory shows a clear cycling behavior, that closely matches that of the 2D neural field in @fig-cg-2d-1d-original.
 
@@
 
-![(TODO animation, higher N, remove all the other mappings) Simulations of coarse-graining with different mappings from the $p=2$ embedding.](figures/coarsegraining.png){width=500px #fig-cg-2d-1d}
+::: {#fig-cg-2d-1d}
+
+![Neural field in the 2D embedding](figures/coarse_graining_p=2_original.mp4){#fig-cg-2d-1d-original}
+
+![Random mapping followed by coarse-graining](figures/coarse_graining_p=2_random.mp4){#fig-cg-2d-1d-random}
+
+![Column mapping followed by coarse-graining](figures/coarse_graining_p=2_column.mp4){#fig-cg-2d-1d-column}
+
+![Z-mapping followed by coarse-graining](figures/coarse_graining_p=2_z.mp4){#fig-cg-2d-1d-z}
+
+The cycling 2D neural field ($4^8$ square populations) is mapped to $[0,1]$ using different mappings, then coarse-grained (resulting in $2^8$ segment populations).
+:::
 
@@
 Despite the fractal structure, the 2D neural field is still regular enough that the dynamics are conserved, but more importantly that it can *again* be mapped to an equivalent 1D neural field.
-In @fig-averaging-geom-3d, we repeat a similar geometric view as @fig-averaging-geom: we see the Z-mapping in the slice $(v_2, v_3)$, how after coarse-graining the kernel is still effectively "sampled" at averaged locations, and crucially, how repeating the coarse-graining still leads to average locations that span the entire $[0,1]^3$ box (instead of being constrained to a plane embedded in $[0,1]^3$, as would have been the case with the Column mapping).
+<!-- In @fig-averaging-geom-3d, we repeat a similar geometric view as @fig-averaging-geom: we see the Z-mapping in the slice $(v_2, v_3)$, how after coarse-graining the kernel is still effectively "sampled" at averaged locations, and crucially, how repeating the coarse-graining still leads to average locations that span the entire $[0,1]^3$ box (instead of being constrained to a plane embedded in $[0,1]^3$, as would have been the case with the Column mapping).
+-->
 
-![(TODO what $n$) Effective sampling points after averaging.](figures/mapping_3d_2d_1d.png){width=400px #fig-averaging-geom-3d}
+<!-- ![(TODO what $n$) Effective sampling points after averaging.](figures/mapping_3d_2d_1d.png){width=400px #fig-averaging-geom-3d} -->

### 2023-06-20 manuscript/chapters/4_locality.qmd
so... many.... figures !

@@
 
-![(TODO random, column and zcurve only, rename neuron to population) Distance between the first population and all other populations, in 1D and 2D.
-The dotted line has a slope $\sqrt 2$.](figures/distance_2d_1d.png){#fig-distance-2d-1d}
+```{python}
+#| label: fig-distance-2d-1d
+#| fig-cap: Distance between the first population and all other populations, in 1D and 2D. The dotted line has a slope $\sqrt 2$.
+#| echo: false
+
+mapping_clss = {
+	'Random mapping': RandomMapping,
+	'Column mapping': ColumnMapping,
+	'Z-mapping': ZMapping,
+}
+
+fig, axes = plt.subplots(ncols=len(mapping_clss), figsize=(len(mapping_clss)*3, 3))
+
+isfirst = True
+for ax, (name, mapping_cls) in zip(axes, mapping_clss.items()):
+	F = mapping_cls.new_nrec(6).inverse_samples()
+	a = np.arange(len(F)) / len(F)
+	a_ = np.linspace(-0.02, 1.02, 3)
+	ax.plot(a, np.linalg.norm(F[0] - F, axis=1), linewidth=0.5)
+	ax.plot(a_, a_*np.sqrt(2), '--', clip_on=False) 
+	ax.set_xlabel('Distance in 1D')
+	if isfirst:
+		isfirst = False
+		ax.set_ylabel('Distance in 2D')
+	else:
+		ax.yaxis.set_visible(False)
+		ax.spines.left.set_visible(False)
+	ax.label_outer()
+	ax.set_title(name)
+	ax.set_xlim((0,1))
+	ax.set_ylim((0, 1.6))
+	mktrans(fig, ax)
+	breathe(ax)
+
+# %matplotlib inline
+display(fig)
+```
 
@@
 
-![(TODO change title, change name of mappings, remove the theory, but keep upper bound) (Numerator of the) average binned variation as a function of $n$.
-The local mappings have a growth slower than the upper bound $2\times 2^n$.](figures/v_num.png){width=500px #fig-vnum}
+```{python}
+#| label: fig-vnum
+#| fig-cap: (Numerator of the) average binned variation as a function of $n$. The local mappings have a growth slower than the upper bound $2\times 2^n$.
+#| echo: false
+
+def V(F: np.ndarray) -> np.ndarray:
+	nrec = int(np.emath.logn(4, len(F)))
+	F_ = F.reshape((2**nrec, 2**nrec, 2))
+	# F_ = np.roll(F, 2**nrec).reshape((2**nrec, 2**nrec, 2))  # offset the bins halfway to test
+	return (F_.max(axis=1) - F_.min(axis=1)).sum(axis=0)
+
+mapping_clss = {'Random': RandomMapping, 'Column': ColumnMapping, 'Z': ZMapping}
+mapping_Vs = { key: [] for key in mapping_clss }
+nrecs = np.array([1,2,3,4,5,6,7,8,9])
+
+for name, mapping_cls in mapping_clss.items():
+	for nrec in nrecs:
+		F = mapping_cls.new_nrec(nrec).inverse_samples(centered=False)
+		mapping_Vs[name].append(V(F))
+
+import itertools
+fig, ax = plt.subplots(figsize=(4, 3))
+marker = itertools.cycle(('s', 'd', '^', 'o', 'P'))
+colors = itertools.cycle(('C2', 'C1', 'C0'))
+ax.plot(nrecs, 2**(nrecs+1), color='k', linestyle='--', label='upper bound', clip_on=False, zorder=10)
+for name, Vs in mapping_Vs.items():
+	ax.plot(nrecs, np.array(Vs).sum(axis=-1), marker=next(marker), color=next(colors), label=name, clip_on=False)
+ax.set_xlabel('$n$')
+ax.set_yscale('log', base=2)
+ax.set_autoscaley_on(False)
+# ax.grid(True, clip_on=False)
+# ax.legend(loc='upper left', bbox_to_anchor=(1,1))
+ax.legend(loc='upper left', borderpad=0, borderaxespad=0)
+ax.set_xlim((1, 9))
+ax.set_xticks([1,3,5,7,9])
+ax.set_ylim((1, 2**12))
+breathe(ax, which='y')
+ax.spines.bottom.set_position(('data', 0.5))  # breathe does not work on log plots
+mktrans(fig, ax)
+display(fig)
+```
 

### 2023-06-21 manuscript/chapters/1_intro.qmd
guess who just did rm ../docs -r but accidentally had their inkscape svg figures saved in docs instead of manuscript ???? IT MEEEE AM SO SILLYYYYYYYYYY

@@
 
-![(TODO add phi, add z instead of xi, arrow only in one direction) Every neuron in a rate neural network has a position in the $p$-dimensional embedding space](figures/neural_field_discrete.png)
+![Every neuron in a rate neural network has a position in the $p$-dimensional embedding space](figures/fig-rnn.svg){width=500px}
 
@@
 $$ {#eq-nf}
-
-![(TODO add phi, add z instead of xi) The connectivity kernel defines the influence between localized neural populations in the embedding space](figures/connectivity_kernel.png)
+<!-- 
+![(TODO add phi, add z instead of xi) The connectivity kernel defines the influence between localized neural populations in the embedding space](figures/connectivity_kernel.png) -->
 

### 2023-06-21 manuscript/chapters/3_mappings.qmd
guess who just did rm ../docs -r but accidentally had their inkscape svg figures saved in docs instead of manuscript ???? IT MEEEE AM SO SILLYYYYYYYYYY

@@
 
-![(TODO) Embedding of the population dynamics in $[0,1]$](figures/photo_2023-06-10_19-26-49.jpg){#fig-popdynamics-01}
+![Embedding of the population dynamics in $[0,1]$](figures/fig-multipop.svg){width=500px #fig-popdynamics-01}
 
@@
 
-![(TODO, reorder args of $w_U$) Every point of the grid effectively performs coarse-graining, and defines a point where we sample the connectivity kernel $w_U$ and the 2D neural field $h_U$](figures/photo_2023-06-10_21-13-23.jpg){#fig-grid-sampling}
+![Every point of the grid effectively performs coarse-graining, and defines a point where we sample the connectivity kernel $w_U$ and the 2D neural field $h_U$](figures/fig-smallsquare.svg){width=500px #fig-grid-sampling}
 
@@
 
-![(TODO) Venn diagram reprensenting the properties of functions $[0,1] \mapsto [0,1]^2$ which are allowed by Netto's theorem.](figures/photo_2023-06-18_21-32-45.jpg){width=400px #fig-venn}
+![Venn diagram reprensenting the properties of functions $[0,1] \mapsto [0,1]^2$ which are allowed by Netto's theorem.](figures/fig-venn.svg){width=400px #fig-venn}
 
@@
 
-![(TODO rename bins to segments) Splitting $[0,1]$ recursively gives the binary bit expansion of $v_1$](figures/photo_2023-06-10_17-21-28.jpg){#fig-recurse-binary}
+![Splitting $[0,1]$ recursively with the indicator function gives the binary bit expansion of $v_1$](figures/fig-binary.svg){width=400px #fig-recurse-binary}
 
@@
 
-![(TODO) Averaging $2^n$ of the $4^n$ segments populations of $S^n$ results in the $2^n$ segments populations](figures/photo_2023-06-13_12-15-32.jpg){width=400px #fig-averaging}
+![Averaging $2^n$ of the $4^n$ segments populations of $S^n$ results in the $2^n$ segments populations](figures/fig-cg.svg){width=400px #fig-averaging}
 

### 2023-06-21 manuscript/chapters/4_locality.qmd
finished figures !!! finally

@@
 
-![(TODO the variation for the lower one should be lesser) Intuition behind the average binned variation.
-To simplify the visualization, activity levels are used in lieu of the 2D positions.](figures/photo_2023-06-15_00-44-43.jpg){width=400px #fig-sketch-binned-variation}
+![Intuition behind the average binned variation.
+To simplify the visualization, activity levels are used in lieu of the 2D positions.
+4 bins are shown with a different colour.
+The horizontal line represents the average activity in the bin.
+From top to bottom: random, column, Z mapping.](figures/fig-variation.svg){width=600px #fig-sketch-binned-variation}
 

### 2023-06-21 notebooks/neurodyn/_plot.py

@@ -16,6 +16,7 @@ __all__ = [
 	'plot_2D_to_1D_mapping',
 	'add_headers', 'scale_lightness',
 	'cmap_bi', 'cmap_bi_r', 'cmap_pan', 'cmap_pan_r', 'cmap_trans',
+	'mktrans',
 ]
 
 def _unwrap_figax(figax: tuple[Figure, Axes] | None = None, **kwargs) -> tuple[Figure, Axes]:
@@ -216,3 +217,9 @@ def scale_lightness(rgb, scale_l):
 	h, l, s = colorsys.rgb_to_hls(*rgb)
 	# manipulate h, l, s values and return as rgb
 	return colorsys.hls_to_rgb(h, min(1, l * scale_l), s = s)
+
+
+def mktrans(fig, ax):
+	"""Makes figure and axes transparent"""
+	fig.patch.set_facecolor('#FFFFFF00')
+	ax.patch.set_facecolor('#FFFFFF00')

### 2023-06-21 manuscript/index.qmd
cover page

--- 2023-05-08 manuscript/index.qmd
@@
-# Index {.unnumbered .hidden}
+::: {.content-visible when-format="html"}
+::: {.content-hidden}
+# Cover {.unnumbered}
+:::
 
-\[TODO : a nice illustration for this, maybe some cool animation ?\]
+![Smooth neural fields in $p$ dimensions can be mapped to equivalent neural fields in one dimension, and the dynamics are equivalent.](chapters/figures/fig-cover.svg)
+:::

### 2023-06-21 manuscript/appendices/A_fixed_points.qmd
so... many.... figures !
proofread appendix

@@
 
-We recall the equation for the low-rank network of neurons :
+We recall the equation for the low-rank network of neurons:
 
 $$
-\dot h_i(t) = -h_i(t) + \frac 1N \sum_{\mu=1}^p \sum_{j=1}^N F_{\mu i} G_{\mu j} \phi(h_j(t))
+\dot h_i(t) = -h_i(t) + \frac 1N \sum_{\mu=1}^p \sum_{j=1}^N F_{\mu i} G_{\mu j} \phi(h_j(t)).
 $$
@@
 
-We note that in the main text, the distribution is a multivariate normal with unit covariance and zero mean, such that it factors out as $\rho(z_1, \cdots, z_p) = \prod_{\mu=1}^p \mathcal{N}(z_\mu)$.
+In the main text, the distribution is a multivariate normal with unit variance, no covariance and zero mean, such that it factors out as $\rho(z_1, \cdots, z_p) = \prod_{\mu=1}^p \mathcal{N}(z_\mu)$.
 Furthermore, the functions are component-wise $f_\mu(z_1, \cdots, z_p) = z_\mu$ and $g_\mu(z_1, \cdots, z_p) = \tilde \phi(z_\mu) = \frac{\phi(z_\mu) - \avg{\phi(z_\mu)}}{\mathrm{Var}[\phi(z_\mu)]}$.
@@
 $$
-\avg{g_\mu(\vec{z})} = \avg{\tilde \phi(z_\mu)} = \avg{\frac{\phi(z_\mu) - \avg{\phi(z_\mu)}}{\mathrm{Var}[\phi(z_\mu)]}} = 0
+\avg{g_\mu(\vec{z})} = \avg{\tilde \phi(z_\mu)} = \avg{\frac{\phi(z_\mu) - \avg{\phi(z_\mu)}}{\mathrm{Var}[\phi(z_\mu)]}} = 0.
 $$
@@
 Another set of fixed points is $h^\star_i = F_{\nu i}$ for all $\nu = 1, \cdots, p$.
-We refer to these as the "patterns" in the context of learning the network "remembers" (converges to) these vectors.
+We refer to these as the "patterns", since in the context of learning the network "remembers" (converges to) these vectors.
 We verify the fixed point similarly as before.
@@
 
-As $N \to \infty$, the fixed points $h_i^\star$ become functions $h^\star(z_1, \cdots, z_p)$ on the $\Rp$ space :
+As $N \to \infty$, the fixed points $h_i^\star$ become functions $h^\star(z_1, \cdots, z_p)$ on the $\Rp$ space:
 
@@
 
-We now study the stability of the fixed points. In the original equation, we substitute $h_i(t) = h_i^\star + \delta_i(t)$, where $\delta_i(t)$ is a small perturbation.
+We now study the stability of the fixed points. In the evolution equation, we substitute $h_i(t) = h_i^\star + \delta_i(t)$, where $\delta_i(t)$ is a small perturbation.
 
@@
 
-Before we study the spectrum of $K$, we cite the following result: if $\lambda$ is an eigenvalue of $A$, then $\lambda-1$ is an eigenvalue of $A-\mathrm{Id}$.
+Before we study the spectrum of $K$, we write the following result: if $\lambda$ is an eigenvalue of $A$, then $\lambda-1$ is an eigenvalue of $A-\mathrm{Id}$.
 This follows simply from the equation for the eigenvalues $\lambda_i$ and eigenvectors $\vec{v_i}$ of $A$.
 If $A \vec{v_i} = \lambda_i \vec{v_i}$, then $(A-\mathrm{Id})\vec{v_i}=\lambda_i \vec{v_i} - \vec{v_i} = (\lambda_i - 1)\vec{v_i}$.
-Therefore to study $K$ it suffices to study the eigenvalues of $J^\prime$.
+Therefore, to study $K$, it suffices to study the eigenvalues of $J^\prime$.
 
@@
 The term $\avg{g^\prime_\nu(z_\nu) f_\nu(z_\nu)}$ does not vanish in general because of the coupling between the components.
-The integral is too complicated to be analytically solved, but we provide a numerical estimation in our case ($\rho(z_1, \cdots, z_p)=\prod_{\mu=1}^p\mathcal{N}(z_\mu)$, $f_\mu(z_\mu)=z_\mu$, $g_\mu(z_\mu)=\tilde\phi(z_\mu)$) and with fixed points $h_i^\star=0$ and $h_i^\star=F_{\nu i}$.
+The integral is too complicated to be analytically solved, but we provide a numerical estimation in our case ($\rho(z_1, \cdots, z_p)=\prod_{\mu=1}^p\mathcal{N}(z_\mu)$, $f_\mu(z_\mu)=z_\mu$, $g_\mu(z_\mu)=\tilde\phi(z_\mu)$) for fixed points $h_i^\star=0$ and $h_i^\star=F_{\nu i}$.
 
@@
 
-We now consider a Taylor expansion of the rate function $\phi(h)$ up to order 3, and cite results obtained from symbolic computation and known moments of the Gaussian distribution.
+We now consider a Taylor expansion of the activation function $\phi(h)$ up to order 3, and cite results obtained from symbolic computation and known moments of the Gaussian distribution.
 
@@
 
-What a mess ! We summarize the results from the Taylor expansion :
+What a mess! We summarize the results from the Taylor expansion:
 
 * When $\phi$ is linear, we see that all the eigenvalues are zero, therefore the first order stability is inconclusive.
-Numerically, we see that when $\phi$ is linear, the convergence to the fixed points is very slow.
+Simulations show when $\phi$ is linear, the convergence to the fixed points is very slow (the spectrum is negative, but the eigenvalues $\lambda_\nu$ are weakly negative, and tend to zero when $N \to \infty$).
 * When $\phi$ is quadratic, the zero fixed point is stable, and the pattern fixed point is unstable, regardless of the sign of the coefficients.
@@
 
-[TODO FIG : might be nice to have a phase plot of $c_1, c_3$ and the value of the eigenvalue to illustrate the previous illustration]
+<!-- [TODO FIG: might be nice to have a phase plot of $c_1, c_3$ and the value of the eigenvalue to illustrate the previous illustration] -->
 
 The study of the Taylor expansion therefore instructs us that stability is improved when $\phi$ grows rapidly around $h=0$, and that even powers (symmetries of $\phi$ around the y-axis) deteriorate the stability of the pattern fixed point.
-This helps us understand why the logistic function $\phi(h)=\frac{1}{1+\mathrm{e}^{-h}}=\tfrac 12 + \tfrac h4 - \tfrac{h^3}{48} + \bO(h^5)$ has good convergence properties : we have that $c_1$ is large enough, $c_3$ is negative, and $c_2$ is zero.
+This helps us understand why the logistic function $\phi(h)=\frac{1}{1+\mathrm{e}^{-h}}=\tfrac 12 + \tfrac h4 - \tfrac{h^3}{48} + \bO(h^5)$ has good convergence properties: we have that $c_1$ is large enough, $c_3$ is negative, and $c_2$ is zero.
 
@@
 In conclusion, the spectrum of $K$ is composed of $p$ eigenvalues $\lambda_\nu = \avg{g_\nu(z_\nu) f_\nu(z_\nu) \partial \phi(h^\star)}-1$, and $N-p$ eigenvalues $\lambda_\perp = -1$.
-For the network considered in this work, the interpretation of this is that components of the field aligned with the patterns $z_\nu$ converge to the fixed point, and the orthogonal components decay as $t \to \infty$.
+For the network considered in this work, the interpretation is that components of the field aligned with the patterns $z_\nu$ converge to the fixed point, and the orthogonal components decay as $t \to \infty$.
 The fixed point $h_i^\star=0$ is therefore unstable, and the pattern fixed points $h_i^\star = z_{\nu i}$ are stable.
-We also note that since the normal distribution is centred around zero and symmetric, $h_i^\star = -z_{\nu i}$ are also stable fixed points.
+We finally note that since the normal distribution is centred around zero and symmetric, $h_i^\star = -z_{\nu i}$ are also stable fixed points.

### 2023-06-21 manuscript/appendices/B_mean_patterns.qmd
proofread appendix

@@
 For now, we do not use the fact that in our setting, $J_{ij}$ is low-rank.
-Applying the finite-$n$ mapping $S^n$,[^details] the neurons $\A = \{i_1, \cdots, i_{|\A |}\}$ will end up in the same 1D bin at position $\alpha$. There are 1D $4^n$ bins, each of length $4^{-n}$.
+Applying the finite-$n$ mapping $S^n$,[^details] the neurons $\A = \{i_1, \cdots, i_{|\A |}\}$ will end up in the same 1D bin at position $\alpha$.
+We stress that $|\A| \neq 2^n$ in general, because we sample the distribution, instead of applying a grid.
+There are, however, still $4^n$ 1D bins, each of length $4^{-n}$.
 
@@
 
-As explained in the main text, we average the neurons together in a bin, and we write the corresponding potential of the segment population at location $\alpha$:
+As explained in the main text, we average the segment together in a bin, and we write the corresponding potential of the segment population at location $\alpha$:
 
@@
 
-We now make the mean-field approximation by dropping $\sum_{i \in \A}$, that is, that all the neurons inside of each bin are the same (or at least, sufficiently similar so that the mean-field is not a big mistake).
+We now make the mean-field approximation by dropping $\sum_{i \in \A}$, that is, that all the neurons inside of each bin are the same (or at least, sufficiently similar so that the mean-field approximation is not a big mistake, in the sense that there is a lot of variation inside each bin, see @sec-locality).
 Doing so, we recover @eq-evolution-network, which concludes the proof.
@@
 
-@eq-Jab-mean-lr also nicely complements the intuition that the more neurons are in each bin, the more "important" this bin should be.
+@eq-Jab-mean-lr also nicely complements the intuition that the more neurons there are in each bin, the more "important" this bin should be.
 This is expressed through the weights $\frac{|\B|}{\sum_{\beta^\prime} |\B^\prime|}$.
@@
 For the sake of illustration, let us compare the connectivity kernels obtained using different methods in @fig-kernels.
+These three approximations are equivalent, since they yield the same dynamics in the $N \to \infty$ limit.
 
-* @fig-kernels-1 is the method described in this appendix, based on sampling $N$ neurons in the PDF space $\R^2$, applying the mapping on the obtained positions $\vec{z_i}$, the applying the binning (downsampling).
-We see, through the empty rows in the connectivity, that due to finite-$N$ effect, locations where $\rho$ is small are not represented, and, as a consequence, no neurons are present in the corresponding 1D bin.
+* @fig-kernels-1 is the method described in this appendix, based on sampling $N$ neurons in the PDF space $\R^2$, applying the mapping on the obtained positions $\vec{z_i}$, and finally applying the binning (this is coarse-graining).
+We see, through the empty rows in the connectivity, that due to finite-$N$ effects, locations where $\rho$ is small are not represented, and, as a consequence, no neurons are present in the corresponding 1D bin.
 When this happens, we set these rows to zero.
@@
   $$
-  We see that compared to @fig-kernels-1, using the grid method allows us to "sample" the locations where the density is small, and this results in less "whited-out" rows.
-The vertical bands of white correspond to the reweighing of the columns with $\tilde \rho(\beta)$.
-* Finally, @fig-kernels-3 shows the kernel obtained from applying the grid method in the CDF $[0,1]^2$ described in @sec-mapping-cdf.
+  We see that compared to @fig-kernels-1, using the grid method allows us to "sample" the locations where the density is small, and this results in less "whited-out" *rows*.
+Still, vertical bands of very fade colour correspond to the reweighing of the *columns* where $\tilde \rho(\beta)$ is small.
+* Finally, @fig-kernels-3 shows the kernel obtained from applying the grid method in the CDF space $[0,1]^2$, as described in @sec-mapping-cdf.
 We see that the connectivity seems much more "uniform", because each point of the grid has "equal weight" in the CDF space.

### 2023-06-21 manuscript/chapters/2_neural_fields.qmd
plop

@@
 
-### Gaussian $\mathcal{N}(\vec 0, \mathbb{1}_N)$ low-rank network of neurons
+### Gaussian low-rank network of neurons
 
@@
 
+<!-- :::: {.content-visible when-format="pdf" layout-nrow=2}
+:::{#fig-p1 layout-nrow=2}
+![First frame](figures/embedding_p=1.mp4-1.jpg)
+
+![Last frame](figures/embedding_p=1.mp4-2.jpg)
+
+Simulation of the Gaussian network of neurons with $p=1$.
+Every neuron is represented by a dot, of which the height represents its activity level, and the colorbar doubles as a y-axis.
+The first and the last frame of the animation are shown, [available here](https://ninivert.github.io/lcnthesis/chapters/figures/embedding_p=1.mp4).
+:::
+:::: -->
+
 ::: {.content-visible when-format="pdf"}
-![(TODO) Simulation of the Gaussian network of neurons with $p=1$.
-Every neuron is represented by a dot, of which the height represents its activity level, and the colorbar doubles as a y-axis.](figures/lol7.jpg){#fig-p1}
+```{python}
+#| label: fig-p1
+#| layout-nrow: 2
+#| fig-cap: Simulation of the Gaussian network of neurons with $p=1$.
+Every neuron is represented by a dot, of which the height represents its activity level, and the colorbar doubles as a y-axis.
+The first and the last frame of the animation are shown, [available here](https://ninivert.github.io/lcnthesis/chapters/figures/embedding_p=1.mp4).
+#| fig-subcap:
+#|   - First frame
+#|   - Last frame
+#| echo: false
+# Note: I have to do this wizardry because Quarto *really* doesn't seem to like nested divs with conditional content...
+from IPython.display import Image
+display(Image('figures/embedding_p=1.mp4-1.jpg'))
+display(Image('figures/embedding_p=1.mp4-2.jpg'))
+```
 :::
@@
 ::: {.content-visible when-format="pdf"}
-![(TODO PDF) Simulation of the Gaussian network of neurons with $p=2$.
+```{python}
+#| label: fig-p2
+#| layout-ncol: 2
+#| fig-cap: Simulation of the Gaussian network of neurons with $p=2$.
 Every neuron is represented by a dot, of which the height represents its activity level.
-Projections are shown by gray points.](figures/lol7.jpg){width=500px #fig-p2}
+Projections are shown by gray points.
+The first and the last frame of the animation are shown, [available here](https://ninivert.github.io/lcnthesis/chapters/figures/embedding_p=2.mp4).
+#| fig-subcap:
+#|   - First frame
+#|   - Last frame
+#| echo: false
+# Note: I have to do this wizardry because Quarto *really* doesn't seem to like nested divs with conditional content...
+from IPython.display import Image
+display(Image('figures/embedding_p=2.mp4-1.jpg'))
+display(Image('figures/embedding_p=2.mp4-2.jpg'))
+```
 :::
@@
 ::: {.content-visible when-format="pdf"}
-![(TODO PDF) A cycling neural field with $\delta=6$ and initial condition $h(z_1, z_2, t < 0) = z_1$.
-The latent trajectory is also computed, and demonstrates the oscillatory behaviour.](figures/lol7.jpg){#fig-cycling-anim}
+```{python}
+#| label: fig-cycling-anim
+#| layout-nrow: 2
+#| fig-cap: A cycling neural field with $\delta=6$ and initial condition $h(z_1, z_2, t < 0) = z_1$.
+The latent trajectory is also computed, and demonstrates the oscillatory behaviour.
+The first and the last frame of the animation are shown, [available here](https://ninivert.github.io/lcnthesis/chapters/figures/embedding_p=2_cycling.mp4).
+#| fig-subcap:
+#|   - First frame
+#|   - Last frame
+#| echo: false
+# Note: I have to do this wizardry because Quarto *really* doesn't seem to like nested divs with conditional content...
+from IPython.display import Image
+display(Image('figures/embedding_p=2_cycling.mp4-1.jpg'))
+display(Image('figures/embedding_p=2_cycling.mp4-2.jpg'))
+```
 :::

### 2023-06-21 manuscript/chapters/3_mappings.qmd
plop

@@
 [^indices]: In the source code, only the method `indices(v12: array[float, 2]) -> array[int, 1]` is defined.
-This is because some mappings are implemented (for instance, the [`RecursiveLocalMapping`](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_mapping.py#L244), that we don't discuss here), which don't lend themselves to the simple permutation of bits formulation.
+This is because some mappings are implemented (for instance, the [`RecursiveLocalMapping`](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_mapping.py#L244), that we don't discuss here), that don't lend themselves to the simple permutation of bits formulation.
 
@@
 
-* @fig-cg-2d-1d-random: The random mapping can be numerically evaluated as a trivial case of a mapping that does not have locality.
+* The random mapping can be numerically evaluated as a trivial case of a mapping that does not have locality.
 The coarse-graining averages out all the structure of the activity, and the recurrent currents are zero, because the contributions from excitatory and inhibitory populations cancel out.
 This results in the erasure of the cycling behavior in the latent trajectory.
-* @fig-cg-2d-1d-column: The Column mapping shows the projection of the initial condition onto the $z_1$ axis, yielding the sigmoid in the $[0,1]$ embedding.
+* The Column mapping shows the projection of the initial condition onto the $z_1$ axis, yielding the sigmoid in the $[0,1]$ embedding.
 Again, the current currents cancel out and yield zero, and the latent trajectory quickly decays towards $(\kappa_1, \kappa_2) = (0,0)$.
-* @fig-cg-2d-1d-z: The Z-mapping shows that despite the fractal structure of the activity, there is still regularity in the $[0,1]$ embedding, resulting in non-zero recurrent currents.
-In consequence, the latent trajectory shows a clear cycling behavior, that closely matches that of the 2D neural field in @fig-cg-2d-1d-original.
+* The Z-mapping shows that despite the fractal structure of the activity, there is still regularity in the $[0,1]$ embedding, resulting in non-zero recurrent currents.
+In consequence, the latent trajectory shows a clear cycling behavior, that closely matches that of the 2D neural field.
 
@@
 
-::: {#fig-cg-2d-1d}
+::: {.content-visible when-format="html"}
+
+:::::: {#fig-cg-2d-1d}
 
@@
 The cycling 2D neural field ($4^8$ square populations) is mapped to $[0,1]$ using different mappings, then coarse-grained (resulting in $2^8$ segment populations).
+::::::
+
+:::
+
+::: {.content-visible when-format="pdf"}
+```{python}
+#| label: fig-cg-2d-1d
+#| layout-ncol: 2
+#| layout-nrow: 4
+#| fig-cap: The cycling 2D neural field ($4^8$ square populations) is mapped to $[0,1]$ using different mappings, then coarse-grained (resulting in $2^8$ segment populations).
+#| fig-subcap:
+#|   - Neural field in the 2D embedding. First frame, animation [available here](https://ninivert.github.io/lcnthesis/chapters/figures/coarse_graining_p=2_original.mp4).
+#|   - Neural field in the 2D embedding. Last frame, animation [available here](https://ninivert.github.io/lcnthesis/chapters/figures/coarse_graining_p=2_original.mp4).
+#|   - Random mapping followed by coarse-graining. First frame, animation [available here](https://ninivert.github.io/lcnthesis/chapters/figures/coarse_graining_p=2_random.mp4).
+#|   - Random mapping followed by coarse-graining. Last frame, animation [available here](https://ninivert.github.io/lcnthesis/chapters/figures/coarse_graining_p=2_random.mp4).
+#|   - Column mapping followed by coarse-graining. First frame, animation [available here](https://ninivert.github.io/lcnthesis/chapters/figures/coarse_graining_p=2_column.mp4).
+#|   - Column mapping followed by coarse-graining. Last frame, animation [available here](https://ninivert.github.io/lcnthesis/chapters/figures/coarse_graining_p=2_column.mp4).
+#|   - Z-mapping followed by coarse-graining. First frame, animation [available here](https://ninivert.github.io/lcnthesis/chapters/figures/coarse_graining_p=2_z.mp4).
+#|   - Z-mapping followed by coarse-graining. Last frame, animation [available here](https://ninivert.github.io/lcnthesis/chapters/figures/coarse_graining_p=2_z.mp4).
+#| echo: false
+# Note: I have to do this wizardry because Quarto *really* doesn't seem to like nested divs with conditional content...
+from IPython.display import Image
+display(Image('figures/coarse_graining_p=2_original.mp4-1.jpg'))
+display(Image('figures/coarse_graining_p=2_original.mp4-2.jpg'))
+display(Image('figures/coarse_graining_p=2_random.mp4-1.jpg'))
+display(Image('figures/coarse_graining_p=2_random.mp4-2.jpg'))
+display(Image('figures/coarse_graining_p=2_column.mp4-1.jpg'))
+display(Image('figures/coarse_graining_p=2_column.mp4-2.jpg'))
+display(Image('figures/coarse_graining_p=2_z.mp4-1.jpg'))
+display(Image('figures/coarse_graining_p=2_z.mp4-2.jpg'))
+```
 :::
@@
 ::: {.content-visible when-format="pdf"}
-![(TODO PDF. initial state and final state) Equivalence of dynamics, from 3D to 2D to 1D.](figures/lol7.jpg){#fig-cg-3d-2d-1d}
+```{python}
+#| label: fig-cg-3d-2d-1d
+#| layout-nrow: 2
+#| fig-cap: Equivalence of dynamics, from 3D to 2D to 1D.
+The first and the last frame of the animation are shown, [available here](https://ninivert.github.io/lcnthesis/chapters/figures/embedding_3d_2d_1d.mp4).
+#| fig-subcap:
+#|   - First frame
+#|   - Last frame
+#| echo: false
+# Note: I have to do this wizardry because Quarto *really* doesn't seem to like nested divs with conditional content...
+from IPython.display import Image
+display(Image('figures/embedding_3d_2d_1d.mp4-1.jpg'))
+display(Image('figures/embedding_3d_2d_1d.mp4-2.jpg'))
+```
 :::

### 2023-06-21 manuscript/index.qmd
plop

--- 2023-05-08 manuscript/index.qmd
@@
-# Index {.unnumbered .hidden}
+::: {.content-visible when-format="html"}
 
-\[TODO : a nice illustration for this, maybe some cool animation ?\]
+::: {.content-hidden}
+# Abstract {.unnumbered}
+:::
+
+:::
+
+::: {.content-visible when-format="html"}
+**Abstract**
+:::
+
+Early recordings of cat's somatosensory and visual cortices suggested that the cortical sheet is organized in vertical columns of functionally similar neurons.
+The columnar organization of the cortex motivated the design of spatially structured models of neural population dynamics, namely neural field models, where the spatial dimensions corresponded to the two dimensions of the cortical sheet.
+Later it has been shown that not only the physical position of the neurons can be used to parametrize their function, but also their location in more abstract "embedding spaces", for example a space of "neuron celltypes".
+
+Decoupling the embedding space from the physical position of neurons on the cortical sheet begs the question of uniqueness of the embedding space.
+Given an embedding space and neural dynamics thereon, are there other embedding spaces which can express the same behavior?
+Is there a "natural" choice for the embedding space?
+
+In this work, we show that the answer to the second question is highly nontrivial because given some neural dynamics, we can find multiple equivalent embedding spaces in different dimensions.
+We illustrate this using a toy example of a 2 (or 3)-dimensional embedding that is mapped to an equivalent 1-dimensional embedding.
+This is done by applying measurable bijective mappings $S : [0,1]^2 \mapsto [0,1]$, which lead, as we show, to well-behaved numerical simulation schemes.
+In this, we exploit the key concept of "locality", which expresses the property of (possibly fractal) neural fields, in which neighbouring neurons in the embedding space *tend* to have similar functions.
+
+![Smooth neural fields in $p$ dimensions can be mapped to equivalent neural fields in one dimension, and the dynamics are equivalent.](chapters/figures/fig-cover.svg)

### 2023-06-21 manuscript/chapters/figures/video_to_firstlast.sh

@@ -1,0 +1,11 @@
+#!/bin/bash
+# extracts first and last frame of every video (files ending in *.mp4)
+# output image has max 400px height
+# this is done for the PDF render
+
+for filename in *.mp4; do
+	echo "$filename";
+	ffmpeg -y -i $filename -sseof -1 -i $filename -map 0:v -vframes 1 -q:v 1 -qmin 1 "$filename-1.jpg" -map 1:v -q:v 1 -qmin 1 -update 1 "$filename-2.jpg";
+	# convert "$filename-1.jpg" -resize 'x500>' "$filename-1.jpg";
+	# convert "$filename-2.jpg" -resize 'x500>' "$filename-2.jpg";
+done

### 2023-06-21 manuscript/acknowledgements.qmd
goodbye i will now sleep for 3 days

# Acknowledgements {.unnumbered}

I would like to express my most sincere gratitude towards Dr. Valentin Schmutz, who supervised this thesis and provided key insights on the theoretical aspects, as well as invaluable feedback during the editing process.

My gratitude extends to the members of the LCN directed by Prof. Wulfram Gerstner, for their helpful comments during the midterm presentation, and for maintaining a welcoming atmosphere in the lab. Additionally, I would like to thank Louis Henry Pezon for his patience in sharing his knowledge on low-rank networks.

Thank you to my girlfriend Solène, for her unwavering love and support, and for teaching me about the Z-mapping, which proved to be essential in the writing of this thesis. To my friends and family for encouraging me and being there when I needed it most.

### 2023-06-21 manuscript/appendices/B_mean_patterns.qmd
goodbye i will now sleep for 3 days

@@
 [^details]: We are glossing over the fact that $S^n : [0,1]^2 \mapsto [0,1]$, but the positions $\vec{z_i} \in \R^2$.
-In the code, we can deal with these by either applying the inverse CDF (as described in @sec-mapping-cdf), or considering a bounding box $[\min_i(z_{1,i}), \max_i(z_{1,i})] \times [\min_i(z_{2,i}), \max_i(z_{2,i})]$ of the sampled points, which is then rescaled to $[0,1] \times [0,1]$ (see implementation for [`Box`](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_mapping.py#L24)).
+In the code, we can deal with this by either applying the inverse CDF (as described in @sec-mapping-cdf), or considering a bounding box $[\min_i(z_{1,i}), \max_i(z_{1,i})] \times [\min_i(z_{2,i}), \max_i(z_{2,i})]$ of the sampled points, which is then rescaled to $[0,1] \times [0,1]$ (see implementation for [`Box`](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_mapping.py#L24)).
 Both approaches are equivalent, in the sense that they give rise to the same dynamics as $n$ becomes large, although the obtained connectivity matrix $\tilde J_{\alpha\beta}$ may not be the identical.
@@
   $$
-  \tilde J_{\alpha\beta} = \tilde \rho(\beta) \sum_{\mu=1}^p \tilde F_{\mu \alpha} \tilde G_{\mu \beta}
+  \tilde J_{\alpha\beta} = \tilde \rho(\beta) \sum_{\mu=1}^p \tilde F_{\mu \alpha} \tilde G_{\mu \beta}.
   $$

### 2023-06-21 manuscript/chapters/1_intro.qmd
finished figures !!! finally
goodbye i will now sleep for 3 days

@@
 
-Early recordings of cat's somatosensory @Mou57 and visual @HubWie62 cortices suggested that the cortical sheet is organized in vertical columns of functionally similar neurons.
+Early recordings of cats' somatosensory @Mou57 and visual @HubWie62 cortices suggested that the cortical sheet is organized in vertical columns of functionally similar neurons.
 
-The putative columnar organization of the cortex motivated the design of spatially structured models of neural population dynamics, namely neural field models [@WilCow73;@Nun74;@Ama77], where the spatial dimensions corresponded to the two dimensions of the cortical sheet.
+The columnar organization of the cortex motivated the design of spatially structured models of neural population dynamics, namely neural field models [@WilCow73;@Nun74;@Ama77], where the spatial dimensions corresponded to the two dimensions of the cortical sheet.
 
@@
 More recent recording methods, especially in rodents, have provided evidence of low-dimensional organization of neuronal activity that does not depend on the physical locations of neurons on the cortical sheet.
-For example, contrary to the cat's visual cortex, the orientation selectivity of pyramidal neurons in a rat's visual cortex does not depend on their locations, that is, orientation selectivity is heterogeneous at any given location @OhkChu05.
-Moreover, the activity of interneurons in any small volume of the visual cortex is also heterogeneous and seems to be structured by a low-dimensional transcriptomic manifold of fine cell subtypes @BugDuf22.
+For example, contrary to cats' visual cortex, the orientation selectivity of pyramidal neurons in a rat's visual cortex does not depend on their locations, that is, orientation selectivity is heterogeneous at any given location @OhkChu05.
+Moreover, the activity of interneurons in any small volume of the visual cortex is also heterogeneous and seems to be structured by a low-dimensional manifold of fine cell subtypes @BugDuf22.
 The local functional heterogeneity of neuronal activity in the cortex challenges the old concept of functionally homogeneous cortical columns and, with it, classical neural field models.
@@
 The goal of this Master Thesis is to show that the answer to the second question is highly nontrivial because even the dimensionality of the embedding space, for some given spatiotemporal dynamics, is not clearly defined.
-This means that it is possible to find two embeddings (and connectivity kernels) of different dimensions that give rise to identical dynamics.
+This means that it is possible to find two embeddings (each associated with a connectivity kernel) of different dimensions that give rise to identical dynamics.
 <!-- ; *contradicting the intuition that the embedding dimension should at least be as high as the dimensionality of the dynamics* (the number of variables needed to fully describe the dynamical system).
@@
 
-![Every neuron in a rate neural network has a position in the $p$-dimensional embedding space](figures/fig-rnn.svg){width=500px}
+![Every neuron in the neural network has a position in the $p$-dimensional embedding space.](figures/fig-rnn.svg){width=500px}
 
@@
 This work was inspired by recent mathematical development in mean-field theories involving the theory of graphons @JabPoy21.
-Informally, these recent results suggest that, even in the absence of any explicit spatial structure, as long as weights scale as $1/N$, the dynamics of large networks can be described, as $N \to \infty$, by a neural field equation where the embedding is simply the interval $[0,1]$, which motivates the search for a formulation of a one-dimensional neural embedding.
+Informally, these recent results suggest that, as long as weights scale as $1/N$, the dynamics of large networks can be described, as $N \to \infty$, by a neural field equation where the embedding is simply the interval $[0,1]$, which motivates the search for a formulation of a one-dimensional neural embedding.
 <!-- Their main result is that under some boundedness assumptions of the connectivity between the components of a system, the equation of evolution incorporates only an integral over $[0,1]$.
@@
 2.
-A discretization-based simulation.
+A grid-based simulation.
 In the $\Rp$ space, we discretize each dimension, forming a grid of $p$-dimensional bins.
@@
 Intuitively, this won't work (in the sense that the dynamics in $[0,1]$ will not be the same as the dynamics in $\Rp$), because the hypothesis of continuity is not longer true: the mapping is random, so the nearby neurons in $[0,1]$ will no longer have similar potentials, and the error of the mean-field approximation will no longer vanish as we take increasingly many (and smaller) bins.
+This is visualized in @fig-intuition-locality.
 <!-- NOTE : was "populations" before, this is explained later in the thesis. but OK, for the introduction, maybe we should use neurons
@@
 
-![(TODO) Good mappings from $\Rp$ to $[0,1]$ should maintain some sense of "locality"](figures/photo_2023-06-13_18-15-50.jpg){width=500px}
+![Good mappings from $\Rp$ to $[0,1]$ should maintain some sense of "locality".](figures/fig-locality.svg){width=600px #fig-intuition-locality}
 

### 2023-06-21 manuscript/chapters/2_neural_fields.qmd
goodbye i will now sleep for 3 days

@@
 Such models hold their name from the form of their connectivity matrix, which has a rank of (at most) $p$.
-We write the connectivity matrix as
+We write the connectivity matrix as follows:
 
@@
 
-The geometric view of these networks is that the recurrent currents lie in a $p$-dimensional subspace of $\R^N$ spanned by the vectors $\{\vec{F_1}, \cdots, \vec{F_p}\}$ (where $\vec{F_\mu}=(F_{\mu 1}, \cdots F_{\mu N})$), which therefore define a "natural" linear embedding (in the sense of @JazOst21) of the neural population activity.
+The geometric view of these networks is that the recurrent currents lie in a $p$-dimensional subspace of $\R^N$ spanned by the vectors $\{\vec{F_1}, \cdots, \vec{F_\mu},\cdots, \vec{F_p}\}$ (where each $\vec{F_\mu}=(F_{\mu 1}, \cdots F_{\mu N})$), which therefore define a "natural" linear embedding (in the sense of @JazOst21) of the neural population activity.
 
@@
 Instead of numerically computing the full connectivity matrix $J$ (which would require storing $N^2 = 50,000^2 \sim 10^{10}$ entries, taking about 40 GiB if we use 32-bit floating point numbers...
-yikes !), we take advantage of its low-rank structure to compute recurrent currents, therefore only storing $p \times N$ numbers instead of $N^2$.[^code2]
+yikes!), we take advantage of its low-rank structure to compute recurrent currents, therefore only storing $p \times N$ numbers instead of $N^2$.[^code2]
 
@@
 In the example given here, the network coincidentally converged again to the first pattern fixed point.
-The network is in the $\mu=1$ pattern, which we can see because the neural field is constant in the $z_2$ direction.
+We see this because the neural field is constant in the $z_2$ direction.
 We write the fixed point $h^\star(z_1, z_2, t) = z_1$.
@@
 Due to the formulation of the neural field weighed by a probability distribution $\rho(y_1, \cdots, y_p)$, a natural way to estimate the integral is by Monte Carlo integration.
-The essence of this method is to use the Central Limit Theorem, where we take $N$ independent samples $\vec{y_i} = (y_{1i}, \cdots, y_{pi})$ from the distribution $\rho$, and use them to estimate the integrand.
+The essence of this method is to use the Central Limit Theorem, where we take $N$ independent samples $\vec{y_i} = (y_{1i}, \cdots, y_{pi})$ from the distribution $\rho$, and use them to estimate the integrand $I(\vec z, \vec y)$.
 
@@
 
-Overlap variables in continuous time variables were introduced by @Gerstner_1992 to study spiking networks models of associative memory (Hopfield-type networks), which can be seen as a low-rank network with discrete spatial structure instead of our continuous $\vec{z}$.
+Overlap variables in continuous time were introduced by @Gerstner_1992 to study spiking networks models of associative memory (Hopfield-type networks), which can be seen as a low-rank network with discrete spatial structure instead of our continuous $\vec{z}$.
 The overlap variables measure the correlation between the state of the network (the value of the potential $h(\vec z, t)$ at each point in space) and the pattern fixed point.
@@
 m_\mu(t) &= \frac 1N \sum_{i=1}^N \tilde \phi(z_{\mu i}) \phi(h_i(t)) \\
-&\xrightarrow{N \to \infty} \int_{\Rp} \tilde \phi(z_\mu) \phi(h(\vec z, t)) \mathcal{N}^p(\d \vec z)
+&\xrightarrow{N \to \infty} \int_{\Rp} \tilde \phi(z_\mu) \phi(h(\vec z, t)) \mathcal{N}^p(\d \vec z).
 \end{aligned}
@@
 $$
-h(\vec z, t) = h^\perp(\vec z, t) + \sum_{\mu=1}^p \kappa_\mu(t) z_\mu
+h(\vec z, t) = h^\perp(\vec z, t) + \sum_{\mu=1}^p \kappa_\mu(t) z_\mu.
 $$

### 2023-06-21 manuscript/chapters/3_mappings.qmd
goodbye i will now sleep for 3 days

@@
 We then group neurons of the same type, and make the approximation that they have identical potentials and the same interaction weights with other neurons.
-As a consequence, every population then represents a homogeneous population of identical neurons.
+As a consequence, every population represents a homogeneous population of identical neurons.
 
@@
 
-![Embedding of the population dynamics in $[0,1]$](figures/fig-multipop.svg){width=500px #fig-popdynamics-01}
+![Embedding of the population dynamics in $[0,1]$.](figures/fig-multipop.svg){width=500px #fig-popdynamics-01}
 
@@
 &= -H(\alpha, t) + \sum_{b=1}^{N_\text{pop}} p_b w(\alpha, b) \phi(H_b(t)) \\
-\iff \partial_t H_a(t) &= -H(a, t) + \sum_{b=1}^{N_\text{pop}} p_b w(a, b) \phi(H_b(t)) \\
-&= -H(a, t) + \sum_{b=1}^{N_\text{pop}} p_b J^0_{ab} \phi(H_b(t)) \\
-&= -H(a, t) + \sum_{b=1}^{N_\text{pop}} \tilde J^0_{ab} \phi(H_b(t)) 
+\iff \partial_t H_a(t) &= -H_a(t) + \sum_{b=1}^{N_\text{pop}} p_b w(a, b) \phi(H_b(t)) \\
+&= -H_a(t) + \sum_{b=1}^{N_\text{pop}} p_b J^0_{ab} \phi(H_b(t)) \\
+&= -H_a(t) + \sum_{b=1}^{N_\text{pop}} \tilde J_{ab} \phi(H_b(t)) 
 \end{aligned}
@@
 When we discretize the $[0,1]^2$ space, we are effectively performing a coarse-graining approximation and simulating a finite number of populations.
-For every bin localized by $(v_1, v_2)$, we are sampling the connectivity kernel $w_U(\cdot, (v_1, v_2))$ and 2D neural field $h_U((v_1, v_2), t)$ (see @fig-grid-sampling).
+For every square bin localized by $(v_1, v_2)$, we are sampling the connectivity kernel $w_U((v_1, v_2), \cdot)$ and the 2D neural field $h_U((v_1, v_2), t)$ (see @fig-grid-sampling).
 In the rest of the chapter, we refer to these 2D bins as "square populations", and to the corresponding 1D bins obtained by the mapping $S$ as "segment populations".
 
-![Every point of the grid effectively performs coarse-graining, and defines a point where we sample the connectivity kernel $w_U$ and the 2D neural field $h_U$](figures/fig-smallsquare.svg){width=500px #fig-grid-sampling}
+![Every point of the grid effectively performs coarse-graining, and defines a point where we sample the connectivity kernel $w_U$ and the 2D neural field $h_U$.](figures/fig-smallsquare.svg){width=500px #fig-grid-sampling}
 
@@
 
-The connectivity kernel $\tilde w$ defines the interactions between the populations in the $[0,1]$ embedding, which we define relating to the kernel $w_U$ on $[0,1]^2$, defined in @sec-mapping-cdf as:
+The connectivity kernel $\tilde w$ defines the interactions between the populations in the $[0,1]$ embedding, which we define relating to the kernel $w_U$, defined in @sec-mapping-cdf as:
 
@@
 The definition of the connectivity kernel in @eq-kernel-1d might seem innocent, but the key element here is the inverse $S^{-1}$.
-We recall that in the toy model, the original connectivity kernel is written as $w(\vec z, \vec y) = \sum_{\mu=1}^p \tilde \phi(y_\mu) z_\mu$.
+We recall that in the toy model, the original connectivity kernel is written as $w(\vec z, \vec y) = \sum_{\mu=1}^p z_\mu \tilde \phi(y_\mu)$.
 This kernel is highly regular, by which we mean that, by its continuity in each of its arguments, populations nearby in the $\Rp$ will have similar input weights.
-Since it is also differentiable, we can write this as a Taylor expansion in the first argument :
+Since it is also differentiable, we can write this as a Taylor expansion in the first argument:
 
@@
 &= w(\vec z, \vec y) + \sum_{\mu=1}^p \epsilon_\mu \phi(y_\mu) + \bO(\norm{\vec \epsilon}^2) \\
-&\approx w(\vec z, \vec y) 
+&\approx w(\vec z, \vec y).
 \end{aligned}
@@
 However, the obvious problem with this approach is that linear projections are not invertible.
-Hence, should we nevertheless define an inverse $P^{-1}$ that computes the preimage of $\alpha \in [0,1]$, we have that
+Hence, should we nevertheless define an inverse $P^{-1}$ that computes the preimage of $\alpha \in [0,1]$, we have that:
 
@@
 In other words, $P^{-1}(\alpha)$ defines a vertical line in the $[0,1]^2$ embedding.
-When we then take the composition $w_U \circ P^{-1}$, the "locality" is destroyed, in the sense that the same population $\alpha$ in the 1D embedding will "sense" all the connectivity kernels on the vertical line.
+When we take the composition $w_U \circ P^{-1}$, the "locality" is destroyed, in the sense that the same population $\alpha$ in the 1D embedding will "sense" all the connectivity kernels on the vertical line.
 
@@
 
-The intuition and illustration @fig-recurse-binary behind this expansion is that we locate each component by recursively splitting the segment in two sub-segments of equal length.
+@fig-recurse-binary illustrates this expansion: we locate each component bit by recursively splitting the segment in two sub-segments of equal length.
 At every step, taking the left or right segments defines the sequence of binary digits.
@@
 In the following, we will restrict ourselves to mappings where the functions $b_k$ are defined in such a way that each bit of the inputs $v_1$ and $v_2$ is used only once.
-Therefore, for each bit $k$ of the output, there is a function $b_k(b^{\mu_k}_{l_k})$ acting on only one bit of the input.
-$\mu_k \in \{1,2\}$ is the input component and $l_k \in \{1,\cdots,n\}$ the bit number corresponding to $k$.
+Therefore, for each bit $k$ of the output, there is a function $b_k(b^{\mu_k}_{l_k})$ acting on only one bit of the input, where $\mu_k \in \{1,2\}$ is the input component and $l_k \in \{1,\cdots,n\}$ the bit number corresponding to $k$.
 Since this function is binary, the only allowed definitions are
@@
 Numerically, we encounter the problem that the inverse of the Gaussian CDF, used to compute the connectivity, goes to infinity when $v_\mu$ (one of the components of $(v_1, v_2) = S^{-1}(\alpha)$) goes to $0$ (or $1$).
-To avoid this, we offset the squares (by $2^{-n-1}$ along each dimension), so that geometrically their position is not indexed by their lower-left corner, but rather their center.
+To avoid this, we offset the squares (by $2^{-n-1}$ along each dimension), so that geometrically their position is not indexed by their lower-left corner, but rather their center (this is represented in @fig-grid-sampling).
 Analytically, this is not a problem, because the kernel defined in @eq-hu-wu diverges only linearly, compared to the exponential cutoff of the Gaussian distribution, so that the integrand still vanishes at infinity.
@@
 A problem in GPU computing is that of texture locality: sampling textures often involves reading data corresponding to a small 2D region, and therefore to improve speed (minimize cache misses), the 2D data should be packed close together in the 1D memory.
-One way of ordering the 2D texture data in the 1D memory is by using the Z-order curve (also known as "Morton mapping", and similar to the "Lebesgue curve" @sagan_space-filling_1994, chapter 6).
+One way of ordering the 2D texture data in the 1D memory is by using the Z-order curve (also known as "Morton mapping", and similar to the "Lebesgue curve", see @sagan_space-filling_1994, Chapter 6).
 In this way, points close in 1D *tend* to be tightly packed in 2D.
-We can define the Z-order curve (Z-mapping) in the following way:[^zcurve]
+We define the Z-mapping in the following way:[^zcurve]
 
-[^zcurve]: The Z-curve is often defined in a transpose manner, putting first the bits of $v_2$ (vertical component) then the bits of $v_1$, instead of $v_1$ then $v_2$.
+[^zcurve]: The Z-order curve is often defined in a transpose manner, putting first the bits of $v_2$ (vertical component) then the bits of $v_1$ (horizontal component), instead of $v_1$ then $v_2$.
 By doing this, we get the "upright" Z shapes, as [shown here](https://en.wikipedia.org/wiki/Z-order_curve#/media/File:Four-level_Z.svg).
@@
 We also see that the Column mapping is not local in this sense, since there is a large variation along the vertical direction, which seems to always be "of order one".
-This can be explained by looking at the expression for $C^{-1}_2(\alpha^{(n)}) = 0.b_{n+1} b_{n+2} \cdots b_{2n}$ : small variations of the order of $2^{-n}$ in the value of $\alpha$ always result in variations of order one in $C^{-1}_2$.e
+This can be explained by looking at the expression for $C^{-1}_2(\alpha^{(n)}) = 0.b_{n+1} b_{n+2} \cdots b_{2n}$: small variations of the order of $2^{-n}$ in the value of $\alpha$ always result in variations of order one in $C^{-1}_2$.
 
@@
 #| label: fig-locality
-#| fig-cap: Locality of the mappings. The 2D populations corresponding to three small segments of 1D populations of length $\tfrac{1}{16}$ are shown. The figures are generated with $n=6$.
+#| fig-cap: Locality of the mappings. The 2D populations corresponding to three small segments (each of length $\tfrac{1}{16}$) of 1D populations are shown. The figures are generated with $n=6$.
 #| echo: false
@@
 Without repeating the intuition given in @sec-intro-simnf, if neighbouring populations in 1D have similar potentials, then this allows us to write a neural field equation in $[0,1]$.
-Therefore we would like to have the property that populations close in 1D are (inversely) mapped to populations close in 2D, which already have similar potentials.
+Therefore, we would like to have the property that populations close in 1D are (inversely) mapped to populations close in 2D, which already have similar potentials.
 
@@
 
-### Sequence of mappings and their limit
+### Sequences of mappings and their limit
 
@@
 Consider how $\alpha^{(n)}$ fluctuates as the precision of the input increases with $n \to \infty$.
-As we refine the input with fluctuations of order $2^{-n}$, the output $\alpha^{(n)} = b^1_n 4^{-1} + b^2_n 4^{-2} + b^1_{n-1} 4^{-3} + b^2_{n-1} 4^{-4} \cdots$ fluctuates (at least) with amplitude $\sim 4^{-2}$ independantly of $n$.
-Therefore the output never converges, and the mapping $A^{n}$ is not pointwise convergent.
+As we refine the input with fluctuations of order $2^{-n}$ (we change only the bits that come after $b^1_n$ and $b^2_n$), the output $\alpha^{(n)} = b^1_n 4^{-1} + b^2_n 4^{-2} + b^1_{n-1} 4^{-3} + b^2_{n-1} 4^{-4} \cdots$ fluctuates (at least) with amplitude $\sim 4^{-2}$, independantly of $n$.
+Therefore the output never converges, and the anti-Z mapping is not pointwise convergent.
 
@@
 
-We see that the Column mapping converges to projection on $v_1$, in other words, $\alpha=C(v_1, v_2) = v_1$, showing that despite every $C^{n}$ being numerically bijective, the limit of the sequence is not bijective.
-This becomes clear when looking at the binary expansion of $C$ :
+We see that the Column mapping converges to a projection on $v_1$, in other words, $\alpha=C(v_1, v_2) = v_1$, showing that despite every $C^{n}$ being numerically bijective, the limit of the sequence is not bijective.
+This becomes clear when looking at the binary expansion of $C$:
 
@@
 
-The Z-mapping seems to converge to a fractal curve.
+The Z-mapping seems to converge to a fractal structure.
 Numerically, there seems to be an infinite number of discontinuities, but the sizes of these discontinuities are, for the most part, small, so that *on average* they vanish as $n\to\infty$.
@@
 
-We now address the elephant in the room: how do we numerically simulate a neural field defined on -- what appears to be -- a fractal curve? Naively, we could just proceed as follows :
+We now address the elephant in the room: how do we numerically simulate a neural field defined on -- what appears to be -- a fractal curve? Naively, we could just proceed as follows:
 
@@
    $$
-   at every square location (where we denote for simplicity of notation $S(i)=S(\vec{v_i})$).
-3. Simulate the neural field in $[0,1]$ on the corresponding discretized space with discretization step $4^{-n}$.
+   for every pair of square locations (where we denote for simplicity of notation $S(i)=S(\vec{v_i})$).
+3. Simulate the neural field in $[0,1]$ on the corresponding discretized space, with a mesh size $4^{-n}$.
 
-The problem with this approach is that since the mappings are just *permutations* of the square populations, the resulting connectivity matrix $\tilde J_{\alpha_i,\beta_j}$ is just a permutation of the rows and columns of $J_{ij}$, and will therefore trivially lead to the identical neural field dynamics.
+The problem with this approach is that since the mappings are just *permutations* of the square populations, the resulting connectivity matrix $\tilde J_{\alpha_i,\beta_j}$ is just a permutation of the rows and columns of $J_{ij}$, and will therefore trivially lead to identical neural field dynamics.
 
@@
 
-A geometrical intuition we can give for the coarse-graining is how the averaging of the segment populations leads to sampling the connectivity kernel at the "average" positions in the $[0,1]^2$ embedding.
-In @fig-averaging-geom, we colour-code the square populations corresponding to the segment populations before coarse-graining.
+A geometrical intuition we can give for the coarse-graining is how the averaging of the segment populations leads to sampling the connectivity kernel at "average" positions in the $[0,1]^2$ embedding.
+In @fig-averaging-geom, we colour-code the square populations corresponding to the binned segment populations before coarse-graining.
 
@@
 
-In this section, we show numerical simulations of the neural field in $[0,1]$, obtained from mapping higher-dimensional neural fields in $[0,1]^p,\,p\geq 2$.
+In this section, we show numerical simulations of the neural field in $[0,1]$, obtained from applying mappings on higher-dimensional neural fields in $[0,1]^p$.
 We see that the coarse-graining procedure works when applied from $[0,1]^2$ to $[0,1]$.
@@
 We give a quick overview of the numerical details that go into implementing the methods presented above.
-Since the mappings (Z-mapping, Column mapping) can be written as binary expansions, they naturally lend themselves to numerical approximations.
+Since the mappings (Z-mapping, Column mapping) can be written as binary expansions, they naturally lend themselves to numerical approximations (see [`ZMapping`](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_mapping.py#L513), [`ColumnMapping`](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_mapping.py#L397)).
 Furthermore, this is made especially easy since the mappings we defined are simply permutations of bits (in the $n < \infty$ case), and so we can write a one-to-one correspondence between unsigned integers.
 
-We have implemented the mappings as all inheriting from the same abstract class, and each mapping simply has to define a method `indices_to_indices2d(indices: array[int, 1]) -> array[int, 2]` to go from 1D to 2D ($S^{-1}$), and a method `indices2d_to_indices(indices2d: array[int, 2]) -> array[int, 1]` to go from 2D to 1D.[^indices] The abstract class simply defines additional quality-of-life functionality, which makes use of these two methods.
+We have implemented the mappings as all inheriting from the same abstract class, and each mapping simply has to define a method `indices_to_indices2d(indices: array[int, 1]) -> array[int, 2]` to go from 1D to 2D, and a method `indices2d_to_indices(indices2d: array[int, 2]) -> array[int, 1]` to go from 2D to 1D.[^indices] The abstract class simply defines additional quality-of-life functionality, which makes use of these two methods.
 
@@
 
-Finally, the coarse-graining is easily implemented as the combination of a reshape operation (to bring all consecutive sequences of $2^n$ populations together), followed by averaging (see the [implementation here](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_rnn.py#L186))
+Finally, the coarse-graining is easily implemented as the combination of a reshape operation (to bring all consecutive sequences of $2^n$ segment populations together), followed by averaging (see the [implementation here](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_rnn.py#L186))
 
-### Application of $S$ : from $[0,1]^2$ to $[0,1]$
+### Application of $S$: from $[0,1]^2$ to $[0,1]$
 
-To demonstrate the coarse-graining, we start with the cycling neural field presented in @sec-cycling-nf.
+We start with the cycling neural field presented in @sec-cycling-nf.
 The kernel $\tilde w$ of the one-dimensional neural field is numerically approximated using the coarse-graining procedure of different mappings.
@@
 
-### Iterative application of $S$ : from $[0,1]^3$ to $[0,1]^2$, then to $[0,1]$ {#sec-simulations-3d-2d-1d}
+### Iterative application of $S$: from $[0,1]^3$ to $[0,1]^2$, then to $[0,1]$ {#sec-simulations-3d-2d-1d}
 
@@
 In the previous section, we showed that the coarse-graining procedure works from $[0,1]^2$ to $[0,1]$.
-However, one might argue that the reason why it worked, is not the mapping, but also the continuity of the neural field in $[0,1]^2$.
+However, one might argue that the reason why it worked, is not the mapping, but rather the continuity of the neural field in $[0,1]^2$.
 To show the generality of our results, we show that coarse-graining works, even when the original neural field is fractal.
@@
 
-We close this chapter by repeating the process to go from $[0,1]^2$ to $[0,1]$, mapping $(v_1, \alpha_{23})$ to $\alpha_{123}$, and coarse-graining.
+We close this chapter by repeating this process to go from $[0,1]^2$ to $[0,1]$, mapping $(v_1, \alpha_{23})$ to $\alpha_{123}$, and coarse-graining.
 In total, we have coarse-grained two times: once from $[0,1]^3$ ($8^n$ cube populations) to $[0,1]^2$ ($4^n$ square populations), and again from $[0,1]^2$ to $[0,1]$ ($2^n$ segment populations).

### 2023-06-21 manuscript/chapters/4_locality.qmd
plop
goodbye i will now sleep for 3 days

@@
 
-Therefore, if $h_U(\vec v, t)$ solves the neural field equation in $[0,1]^2$, then $\tilde h(t) = h_U(t) \circ S^{-1}$ solves the equivalent neural field equation in $[0,1]$.
-The same reasoning applies in reverse: if $\tilde h(\alpha, t)$ solves the neural field equation in $[0,1]$, then $h_U(t) = \tilde h \circ S$ solves the neural field in $[0,1]^2$.
+Therefore, if $h_U(\vec v, t)$ solves the neural field equation in $[0,1]^2$, then $\tilde h(\alpha, t) = (h_U(\cdot, t) \circ S^{-1})(\alpha)$ solves the equivalent neural field equation in $[0,1]$.
+The same reasoning applies in reverse: if $\tilde h(\alpha, t)$ solves the neural field equation in $[0,1]$, then $h_U(\vec v, t) = (\tilde h(\cdot, t) \circ S)(\vec v)$ solves the neural field in $[0,1]^2$.
 
@@
 #| label: fig-distance-2d-1d
-#| fig-cap: Distance between the first population and all other populations, in 1D and 2D. The dotted line has a slope $\sqrt 2$.
+#| fig-cap: Distance between the first population and all other populations, in 1D and 2D. The dotted line has a slope $\sqrt 2$. This figure is generated using $n=6$.
 #| echo: false
@@
 
-With this view, we understand that the coarse-grained performed in @sec-coarse-graining was making use of that fact that locality helps by bounding the variation inside of each averaged bin.
+With this view, we understand that the coarse-graining performed in @sec-coarse-graining makes use of that fact that locality helps by bounding the variation inside of each averaged bin.
 Therefore, we would like for mathematically formulate that on average, the variation inside of each bin before coarse-graining is small.
@@
 To simplify the visualization, activity levels are used in lieu of the 2D positions.
-4 bins are shown with a different colour.
-The horizontal line represents the average activity in the bin.
-From top to bottom: random, column, Z mapping.](figures/fig-variation.svg){width=600px #fig-sketch-binned-variation}
+4 bins are each shown with a different colour.
+The horizontal line represents the average activity in each bin.
+From top to bottom: Random, Column, Z-mapping.](figures/fig-variation.svg){width=600px #fig-sketch-binned-variation}
 
@@
 A continuous function $f$ evidently verifies $V_n(f) \xrightarrow{n \to \infty} 0$, since small neighbourhoods are mapped to small neighbourhoods.
-The reverse is obviously not true, since for instance the metric $V_n$ vanishes when applied to a step function, which is only piecewise continuous.
+The reverse is not true, since for instance the metric $V_n$ vanishes when applied to a step function, which is only piecewise continuous.
 
@@
 
-Before we move to analytical derivations applying the average variation to mappings, we can first check numerically if it gives the expected results; that is $V_n$ vanishes a $n\to\infty$ for local mappings, whereas $V_n$ is lower bounded by a constant for non-local mappings.
+Before we move to analytical derivations applying the average variation to mappings, we can first check numerically if it gives the expected results; that is $V_n$ vanishes when $n\to\infty$ for local mappings, whereas $V_n$ should be lower bounded by a constant for non-local mappings.
 @fig-vnum shows this is indeed the case, rejoice!
@@
 #| label: fig-vnum
-#| fig-cap: (Numerator of the) average binned variation as a function of $n$. The local mappings have a growth slower than the upper bound $2\times 2^n$.
+#| fig-cap: (Numerator of the) average binned variation as a function of $n$.
+The Z-mapping has a growth slower than the upper bound $2\times 2^n$.
+The Column mapping grows at the same rate as the upper bound, and the Random mapping saturates it.
 #| echo: false
@@
 Z^{-1}_1(\alpha) &= 0.\underbrace{i_{n-1} i_{n-3} \cdots i_1}_\text{$n/2$ bits} b_n b_{n+2} \cdots b_{2m-1} \\
-Z^{-1}_2(\alpha) &= 0.\underbrace{i_{n-2} i_{n-4} \cdots i_0}_\text{$n/2$ bits} b_{n+1} b_{n+3} \cdots b_{2m}
+Z^{-1}_2(\alpha) &= 0.\underbrace{i_{n-2} i_{n-4} \cdots i_0}_\text{$n/2$ bits} b_{n+1} b_{n+3} \cdots b_{2m}.
 \end{aligned}
@@
 
-If $n$ is odd, we have :
+If $n$ is odd, we have:
 
@@
 Z^{-1}_1(\alpha) &= 0.\underbrace{i_{n-1} i_{n-3} \cdots i_0}_\text{$(n+1)/2$ bits} b_n b_{n+2} \cdots b_{2m-1} \\
-Z^{-1}_2(\alpha) &= 0.\underbrace{i_{n-2} i_{n-4} \cdots i_1}_\text{$(n-1)/2$ bits} b_{n+1} b_{n+3} \cdots b_{2m}
+Z^{-1}_2(\alpha) &= 0.\underbrace{i_{n-2} i_{n-4} \cdots i_1}_\text{$(n-1)/2$ bits} b_{n+1} b_{n+3} \cdots b_{2m}.
 \end{aligned}
@@
 
-We only consider even $n$ in the following, as the proof remains very similar.
+We only consider even $n$ in the following, as the proof for the latter remains very similar.
 
@@
 
-Putting both together, we can write :
+Putting both together, we can write:
 
@@
 \end{aligned}\\
-&\geq \frac 12 + \bO(2^{-n})
+&\geq \frac 12 + \bO(2^{-n}).
 \end{aligned}
@@
 $$
-V_n(C^{-1}) > \frac 12 \implies \lim_{n \to \infty} V_n > \frac 12
+V_n(C^{-1}) > \frac 12 \implies \lim_{n \to \infty} V_n(C^{-1}) > \frac 12
 $$
@@
 In @sec-Vn, we have introduced the average variation metric $V_n$.
-Assuming that $V_n \xrightarrow{n\to\infty} 0$ (which expresses the locality property), does this imply that the numerical estimation of the integral in the neural field equation in $[0,1]$ converges to its true value?
+Assuming that $V_n(S^{-1}) \xrightarrow{n\to\infty} 0$ (which expresses the locality property), does this imply that the numerical estimation of the integral in the neural field equation in $[0,1]$ converges to its true value?
 
-The intuition behind the proof of convergence is that the variation, coming from the coarse-graining, inside each numerical bin vanishes as we increase the number of iterations $n$, even though the the neural field $\tilde{h}$ and the connectivity kernel $\tilde{w}$ become fractal.
+The intuition behind the proof of convergence is that the variation, coming from the coarse-graining, inside each numerical bin vanishes as we increase the number of iterations $n$, even though the neural field $\tilde{h}$ and the connectivity kernel $\tilde{w}$ become fractal.
 
@@
 \text{where}\ \Delta(\beta) &= \tilde w(\alpha, \beta_i) \phi(\tilde h_i(t)) - \tilde w(\alpha, \beta) \phi(\tilde h(\beta, t))\\
-&= w_U(\alpha, S^{-1}(\beta_i)) \phi(h_U(S^{-1}(\beta_i), t)) - w_U(\alpha, S^{-1}(\beta)) \phi(h_U(S^{-1}(\beta), t))
+&= w_U(\alpha, S^{-1}(\beta_i)) \phi(h_U(S^{-1}(\beta_i), t)) - w_U(\alpha, S^{-1}(\beta)) \phi(h_U(S^{-1}(\beta), t)).
 \end{aligned}
@@
 The main source of variance inside the integrand $\Delta(\beta)$ is the term $S^{-1}$, since $w_U$ and $h_U$ are "nice" functions.
-We express this regularity by the added assumption that $w_U(\alpha, \vec u) \phi(h_U(t, \vec u))$ is Lipschitz in $\vec u$, with Lipschitz constant $L(\alpha)$, and $\norm{\alpha}$ the corresponding norm on $[0,1]^2$.
+We express this regularity by the added assumption that $w_U(\alpha, \vec u) \phi(h_U(t, \vec u))$ is Lipschitz in $\vec u$, with Lipschitz constant $L(\alpha)$, and $\norm{\alpha}$ is the corresponding norm on $[0,1]^2$.
 We argue this is not a strong assumption, because this is just expressing the regularity of the integrand in the integral over $[0,1]^2$.
 
-A more pragmatic justification of this is to notice that the recurrent currents are expressed as:
+A more pragmatic justification of this is to notice that the "density of recurrent currents" is expressed as:
 
 $$
-I^\text{rec}_U(\vec v) = w_U(\vec v, \vec u) \phi(h_U(t, \vec u)),
+I^\text{rec}_U(\vec v, \vec u) = w_U(\vec v, \vec u) \phi(h_U(t, \vec u)),
 $$
 
-and that in @sec-cycling-nf, numerical simulations showed that $I^\text{rec}_U(\vec v)$ is a well-behaved function.
+and that in @sec-cycling-nf, numerical simulations showed that $I^\text{rec}_U(\vec v, \vec u)$ (the integrand of the integral on $[0,1]^2$) is a well-behaved function.
 
@@
 \Delta(\beta) &= w_U(\alpha, S^{-1}(\beta_i)) \phi(h_U(S^{-1}(\beta_i), t)) - w_U(\alpha, S^{-1}(\beta)) \phi(h_U(S^{-1}(\beta), t)) \\
-&\leq L(\alpha) \norm{S^{-1}(\beta_i) - S^{-1}(\beta)}
+&\leq L(\alpha) \norm{S^{-1}(\beta_i) - S^{-1}(\beta)}.
 \end{aligned}
@@
 &\leq L(\alpha) \frac{1}{2^n} \sum_{i=1}^{2^n} \sup_{\beta, \beta^\prime \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]^2} \norm{S^{-1}(\beta^\prime) - S^{-1}(\beta)} \\
-&= L(\alpha) V_n \\
+&= L(\alpha) V_n(S^{-1}) \\
 &\xrightarrow{n \to \infty} 0
@@
 Z_p^n(\vec{v^{(n)}}) &= \sum_{k=1}^{n} \sum_{\mu=1}^p b^{\mu}_k 2^{-(p(k-1)+\mu)} \\
-&= 0.b^1_1 b^2_1 \cdots b^p_1 b^1_2 b^2_2 \cdots b^p_{n-1} b^1_n b^2_n \cdots b^p_n
+&= 0.b^1_1 b^2_1 \cdots b^p_1 b^1_2 b^2_2 \cdots b^p_{n-1} b^1_n b^2_n \cdots b^p_n.
 \end{aligned}
@@
 
-The same arguments apply to show that this generalized Z-mapping converges pointwise to $Z_p$, has the locality property, and that $V_n(Z_p) \xrightarrow{n\to\infty} 0$.
+The same arguments apply to show that this generalized Z-mapping converges pointwise to $Z_p$, has the locality property, and that $V_n(Z_p^{-1}) \xrightarrow{n\to\infty} 0$.
 Then the proof of convergence, albeit more laborious, can be repeated using the same procedure.

### 2023-06-21 manuscript/index.qmd
goodbye i will now sleep for 3 days

@@
 
-Early recordings of cat's somatosensory and visual cortices suggested that the cortical sheet is organized in vertical columns of functionally similar neurons.
+Early recordings of cats' somatosensory and visual cortices suggested that the cortical sheet is organized in vertical columns of functionally similar neurons.
 The columnar organization of the cortex motivated the design of spatially structured models of neural population dynamics, namely neural field models, where the spatial dimensions corresponded to the two dimensions of the cortical sheet.

### 2023-06-21 manuscript/postprocess.sh

@@ -1,0 +1,3 @@
+#!/bin/sh
+# fix broken 404 page
+cd ../docs/ && sed -i 's/\/site_libs/site_libs/g' 404.html

### 2023-06-25 manuscript/chapters/3_mappings.qmd
fixed typos
added mean position, fixed typo

@@
 Create a grid of $2^n \times 2^n = 4^n$ squares of size $2^{-n} \times 2^{-n}$ in the $[0,1]\times[0,1]$ embedding.
-Each square is located by a position $\vec{v_i} = (v_{1,i},v_{2,i}),\,i\in\{1,\cdots\,2^n\}$.
+Each square is located by a position $\vec{v_i} = (v_{1,i},v_{2,i}),\,i\in\{1,\cdots\,4^n\}$.
 2. Compute a discretized version of the 1D connectivity kernel
@@
 #| label: fig-averaging-geom
-#| fig-cap: Square populations corresponding to the segment populations before coarse-graining are colored identically when they are the same bin of 1D length $2^{-n}$. Here we show $n=3$.
+#| fig-cap: Square populations corresponding to the segment populations before coarse-graining are colored identically when they are the same bin of 1D length $2^{-n}$.
+A cross marks the "average position" in each bin.
+Here we show $n=3$.
 #| fig-subcap:
@@
 	colors = cycle(['#750087','#001997','#06bd01','#ffff00','#fe941e','#ff1e26','#fcaec5','#74d4ea','#603814','#010101','#fbf7f4'])
-	for i in range(2**nrec): ax.scatter(*F_[i, :, :].T, zorder=10, color=next(colors), ec='k')
+	for i in range(2**nrec):
+		c = next(colors)
+		ax.scatter(*F_[i, :, :].T, zorder=10, color=c, ec='k')
+		ax.scatter(*F_[i, :, :].mean(axis=0), zorder=10, s=80, marker='X', ec='k', color=c)
 	ax.set(

### 2023-07-28 manuscript/chapters/1_intro.qmd
fixed typos + improved eigenvalue plot

@@
 $$ {#eq-nf}
+
+In this equation, as well as for the rest of this work, we will use the notation $\rho(\mathrm d \vec y) := \rho(\vec y) \mathrm d \vec y$ to emphasize that the integral is weighted by the neuron density.
+
 <!-- 
@@
 
-The main idea is to define a measurable bijective mapping $S : \Rp \mapsto [0,1]$ and a measure $\lambda : \Rp \mapsto \mathbb{R}^+$, such that we can write a one-dimensional neural field	$\tilde h(\alpha, t)$ with a connectivity kernel loosely defined as $\tilde w = w \circ S^{-1}$.
+The main idea is to define a measurable bijective mapping $S : \Rp \mapsto [0,1]$ and a measure $\lambda : \Rp \mapsto \mathbb{R}^+$, such that we can write a one-dimensional neural field	$\tilde h(\alpha, t)$ with a connectivity kernel loosely defined as $\tilde w(\alpha, \beta) = w(S^{-1}(\alpha), S^{-1}(\beta))$.
 

### 2023-07-28 manuscript/chapters/2_neural_fields.qmd
added mean position, fixed typo
fixed typos + improved eigenvalue plot

@@
 The toy model we introduce is an instance of a low-rank neural network.
-Such models hold their name from the form of their connectivity matrix, which has a rank of (at most) $p$.
+Such models hold their name from the form of their connectivity matrix, which has a rank $p$.
 We write the connectivity matrix as follows:
@@
 
-The geometric view of these networks is that the recurrent currents lie in a $p$-dimensional subspace of $\R^N$ spanned by the vectors $\{\vec{F_1}, \cdots, \vec{F_\mu},\cdots, \vec{F_p}\}$ (where each $\vec{F_\mu}=(F_{\mu 1}, \cdots F_{\mu N})$), which therefore define a "natural" linear embedding (in the sense of @JazOst21) of the neural population activity.
+The geometric view of these networks is that the recurrent currents lie in a $p$-dimensional subspace of $\R^N$ spanned by the linearly independant vectors $\{\vec{F_1}, \cdots, \vec{F_\mu},\cdots, \vec{F_p}\}$ (where each $\vec{F_\mu}=(F_{\mu 1}, \cdots F_{\mu N})$), which therefore define a "natural" linear embedding (in the sense of @JazOst21) of the neural population activity.
 
@@
 
-One last remark is that in our model, there is no external current; we only study the autonomous dynamics of the system.
+In our model we make the simplifying assumption that there is no external current (we only study the autonomous dynamics of the system), but this does not hurt the generalizability of our results.
 
@@
 $$
-\phi(h) = \frac{1}{1 + \mathrm{e}^h}.
+\phi(h) = \frac{1}{1 + \mathrm{e}^{-h}}.
 $$
@@
 
-We note that contrary to the model introduced in @schmutz2023convergence, we ignore the self-connections term, because it introduces a vanishing correction of order $\bO(1/N)$.
+We note that contrary to the model introduced in @schmutz2023convergence, which excludes the self-connection term by effectively setting $J_{ii} = 0$, we do not do this, because it introduces a vanishing correction of order $\bO(1/N)$.
 
@@
 #| label: fig-spectrum
-#| fig-cap: Spectrum of $K$ for $N=250$ and $p=2$.
-The numerical estimation (black, eigenvalues $\lambda_\mu$ and $\lambda_\perp=-1$) of the eigenvalues is close to the analytical ($N \to \infty$) derivation (red, only eigenvalues $\lambda_\mu$).
+#| fig-cap: Spectrum of $K$ for $N=1024$ and $p=5$.
+Numerical estimations (black, every line represents the sorted spectrum of one random sample of $K$, 20 samples total) of the eigenvalues are close to the analytical ($N \to \infty$) derivation (red, only eigenvalues $\lambda_\mu$).
 Eigenvalues above the dotted gray line (positive values) correspond to unstable fixed points, and below (negative values) correspond to stable fixed points.
+The first $p$ eigenvalues correspond to $\lambda_\mu$ ($\mu\in\{1,\cdots,p\}$), and the remaining eigenvalues correspond to $\lambda_\perp=-1$.
 #| fig-subcap:
@@
 
+N = 1024
+p = 5
+numeig = 20  # number of stored numerical eigenvalues
 dsigmoid = lambda x: sigmoid(x) * (1 - sigmoid(x))
@@
 def plot_spectrum(hstar_str: str):
-	params = LowRankRNNParams.new_valentin(p=2, N=250, phi=sigmoid, I_ext=zero, exclude_self_connections=False)
-	if hstar_str == 'pattern': hstar = params.F[:, 0]
-	elif hstar_str == 'zero': hstar = np.zeros(params.N)
+	nruns = 20
+	eig_nums = np.zeros((nruns, numeig))
+
+	for i, seed in enumerate(range(nruns)):
+		params = LowRankRNNParams.new_valentin(p=p, N=N, phi=sigmoid, I_ext=zero, exclude_self_connections=False, random_state=seed)
+		if hstar_str == 'pattern': hstar = params.F
+		elif hstar_str == 'zero': hstar = np.zeros_like(params.F)
 		dphi = dsigmoid(hstar)
-	J = np.einsum('im,jm->ij', params.F, params.G * dphi[:, None]) / params.N - np.eye(params.N)
-	eig_num = np.sort(np.linalg.eigvals(J).real)[::-1]
+		J = np.einsum('im,jm->ij', params.F, params.G * dphi) / params.N - np.eye(params.N)
+		eig_nums[i] = np.sort(np.linalg.eigvals(J).real)[::-1][:numeig]
+
+	eig_num = eig_nums.mean(axis=0)
+	eig_num_std = eig_nums.std(axis=0, ddof=1) / np.sqrt(nruns)  # std of the mean estimator
 	eig_ana = { 'zero': 0.19061, 'pattern': -0.28090 }[hstar_str]  # see appendix A
@@
 	mktrans(fig, ax)
+	# numerical estimations
 	ax.axhline(0, color='gray', linestyle='--')
-	ax.plot(eig_num[:8], marker='o', clip_on=False)
-	ax.plot([7,8,8.5], eig_num[8:11], linestyle='--', color='k', clip_on=False)
-	ax.plot(np.arange(params.p), np.full(params.p, eig_ana), color='r', marker='o', clip_on=False)
+	# ax.plot(eig_num[:8], marker='o', clip_on=False)
+	# ax.errorbar(np.arange(8), eig_num[:8], yerr=eig_num_std[:8], marker='o', clip_on=False, linestyle=':')
+	ax.plot(eig_nums[:, :8].T, marker='.', color='k', alpha=0.2, clip_on=False, linestyle=':')
+	ax.plot([7,8,8.5], eig_num[8:11], linestyle='--', color='k', clip_on=False)  # dashed line "fadeout"
+	# analytical estimations
+	ax.plot(np.arange(params.p), np.full(params.p, eig_ana), color='r', marker='o', linestyle=':', clip_on=False)
 	ax.set_ylim((-1, 0.5))
@@
 w(\vec z, \vec y) &= \sum_{\mu=1}^p z_\mu \tilde\phi(y_\mu) \\
-\mathcal{N}^p(\d \vec y) &= \frac{\mathrm{exp}(\frac12\sum_{\mu=1}^p z_\mu^2)}{(2\pi)^{p/2}} \mathrm{d} y_1 \cdots \mathrm d y_p.
+\mathcal{N}^p(\d \vec y) &= \frac{\mathrm{exp}(\frac12\sum_{\mu=1}^p y_\mu^2)}{(2\pi)^{p/2}} \mathrm{d} y_1 \cdots \mathrm d y_p.
 \end{aligned}
@@
 
-Due to the formulation of the neural field weighed by a probability distribution $\rho(y_1, \cdots, y_p)$, a natural way to estimate the integral is by Monte Carlo integration.
+Due to the formulation of the neural field weighted by a probability distribution $\rho(y_1, \cdots, y_p)$, a natural way to estimate the integral is by Monte Carlo integration.
 The essence of this method is to use the Central Limit Theorem, where we take $N$ independent samples $\vec{y_i} = (y_{1i}, \cdots, y_{pi})$ from the distribution $\rho$, and use them to estimate the integrand $I(\vec z, \vec y)$.
@@
 We note however that @eq-nf-cycling-p no longer is a partial differential equation, but rather a delayed differential equation, which are in general much more complicated to solve.
-A simplifying assumption that we make is we extend the initial condition back in time, such that $h(\vec z, t < 0) = h(\vec z, 0)$ is a constant function.
-Numerically, this also means that we have to store a history of the neural fields instead of only the current one.
+We make a simplifying assumption to extend the initial condition back in time, such that $h(\vec z, t < 0) = h(\vec z, 0)$ is a constant function.
+Numerically, the delayed partial differential equation also means that we have to store a history of the neural fields instead of only the current one.
 Since the [`solve_ivp`](https://docs.scipy.org/doc/scipy/reference/generated/scipy.integrate.solve_ivp.html) method dynamically adapts the timestep of integration, we use linear interpolation between two known points in the stored neural field states to estimate $h(\vec z, t-\delta)$ (see [`lagging.py`](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_lagging.py) and [its use](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_rnn.py#L334)).

### 2023-07-28 manuscript/chapters/2_neural_fields.qmd
whoopsie typooo

@@
 We should however note that the results obtained in @sec-fixedpoints are valid in the $N \to \infty$ limit.
-A numerical analysis shows in @fig-spectrum that at $N=250$, the numerical eigenvalues approximate the analytical eigenvalues, and this correspondence improves when we take larger $N$ (typically we will for the rest of this thesis take $N > 10^3$).
+A numerical analysis shows in @fig-spectrum that at $N=1024$, the numerical eigenvalues approximate the analytical eigenvalues, and this correspondence improves when we take larger $N$ (typically we will for the rest of this thesis take $N > 10^4$).
 

### 2024-01-27 paper/equivalence-of-neural-field-dynamics-with-different-embedding-dimensionality.tex
first paper commit

%% NETN Template file

%% Documentclass:
\documentclass[NETN,manuscript]{stjour-new}

%% Or,
%% For production only, need .otf fonts, private MIT logo files,
%% and need to use lualatex:
%% \documentclass[finalfonts,NETN]{stjour}

%%%%%%%%%%% Article Set-Up %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Article Type:

%% Choose one of these options, (Default is Research).
%% Research
%% Methods 
%% Data
%% Review
%% Perspectives

\articletype{Research}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% author definitions should be placed here:

\def\R{\mathbb R}
\def\RR{\R^2}
\def\Rp{\R^p}
\def\d{\mathrm d}
\def\Jab{\tilde J_{\alpha\beta}}
\def\Fab{\tilde F_{\alpha\beta}}
\def\Gab{\tilde G_{\alpha\beta}}
\def\Jij{J_{ij}}
\def\Fij{F_{ij}}
\def\gij{G_{ij}}
\def\wu{w_\textrm{U}}
\def\hu{h_\textrm{U}}
\def\bO{\mathcal{O}}
\def\CDF{\mathrm{CDF}}
\newcommand{\avg}[1]{\langle{#1}\rangle}
\newcommand{\norm}[1]{\lVert{#1}\rVert}
% \def\avg#1{\langle{#1}\rangle}
\renewcommand{\vec}[1]{\boldsymbol{#1}}

\begin{document}

\title{Equivalence of neural field dynamics with different embedding dimensionality}
% \subtitle{<Subtitle Here>} %% Optional subtitle

%% If shortened title for running head is needed so that the article title can fit
%%   in the running head, use [] argument, ie,
%%
%%   \title[Shortened Title for Running Head]{Title of Article}
%%   \subtitle{Subtitle Here}

%%% Author/Affil
%% Since we use \affil{} in the argument of the \author command, we
%% need to supply another version of the author names, without \affil{}
%% to be used for running heads:

\author{Nicole Vadot\affil{1}, Valentin Schmutz\affil{1}, \and Wulfram Gerstner \affil{1}}

\affiliation{1}{LCN, EPFL, Lausanne, Switzerland}

%ie.
%\affiliation{1}{Gatsby Computational Neuroscience Unit, University
%College London, London, United Kingdom}

% \affiliation{2}{Another Department, Institution, City, Country}

%ie
%\affiliation{2}{Center for Studies in
%Behavioral Neurobiology, Concordia University, Montreal, Quebec,
%Canada}

\correspondingauthor{Nicole Vadot}{nicole.vadot@epfl.ch}
\correspondingauthor{Valentin Schmutz}{valentin.schmutz@epfl.ch}
\correspondingauthor{Wulfram Gerstner}{wulfram.gerstner@epfl.ch}

% ie,
%\correspondingauthor{Ritwik K. Niyogi}{ritwik.niyogi@gatsby.ucl.ac.uk}

\keywords{Neural fields, Embedding, Dimensionality}

%ie
%\keywords{Work, Leisure, Normative, Microscopic,  Reinforcement Learning, Economics}

\begin{abstract}
In neuroscience, neural field equations model interactions between neurons positioned in an "embedding"; a physical  (for instance, that of the cortical sheet) or abstract space (for instance, that of neuron celltypes) in which the dynamics are highly regular in space and time.

This work contributes to the formalization of the search for neural field models, and in particular treats the question of the embedding space uniqueness by studying its dimensionality. We show that given some neural dynamics, we can find multiple equivalent embedding spaces in different dimensions, in the sense that they have the same dynamics.

The core of this work lies in the study of bijective mappings $S$ between $[0,1]^2$ and $[0,1]$, which allows mapping a two-dimensional neural field to an equivalent one-dimensional neural field. This procedure is easily generalized to map between embeddings in arbitrary number of dimensions.

We illustrate this work through extensive numerical simulations and animations of a low-rank toy model. In particular, the coarse-graining procedure is introduced, in which numerical simulations of $2^n$ (where $n$ is an integer quantifying the resolution of the simulation grid) neural populations in a one-dimensional embedding converge to the same dynamics obtained from simulations of $2^n \times 2^n$ neural populations in a two-dimensional embedding. We additionally introduce a measure of locality, which quantifies how the regularity ("smoothness") of the neural fields is conserved through the mappings $S$. In particular, we use a subset composed of binary mappings, and borrow the Z-order curve from the computer graphics field, which we show is a "local" mapping. We finally prove that this locality guarantees that the one-dimensional neural field equations can be approximated using classical grid-based methods, and therefore the numerical simulations yield the correct equivalent dynamics.
\end{abstract}

% \begin{authorsummary}
% Author Summary here.
% \end{authorsummary}

\section{Introduction}

\subsection{Neural field models in neuroscience}

valentin's introduction with many citations. chap 1.1

\subsection{Motivation}

question of uniqueness, discretization error, mean field limit and locality, link with numerical simualtions. chap 1.4, chap 3.1

argue over smoothness of the kernel (see presentation given for SPC)

\subsection{Definitions}

math definitions of neural fields and embedding spaces, reduction of 2d->1d and CDF, and the mappings

chap 1.1, chap 2.1, 2.5

\subsection{Summary of findings}

\section{Study of mappings from 2D to 1D}

\subsection{$S$ as the limit of a sequence of mappings}

motivation, why do we do this

\subsection{The Z-mapping and Column mapping}

introduce the mappings and give plots

\subsection{Behavior and locality measures}

measure of locality, give examples applied to the mappings.

see how it scales, and say what we expect

\subsection{Numerical demonstrations}

introduce coarse-graining procedure (applying the smoothness hypothesis numerically)

introduce the low-rank cycling model

\section{Proof of results}

\subsection{Analytical equivalence given $S$ measurable and bijective}

\subsection{Numerical convergence}

\section{Conclusion}

\acknowledgments

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% End of Article

% \acknowledgments
% \section{Supporting Information} (optional)
% \section{Competing Interests} (optional)
% \bibliography{<name of .bib file>}

% \acknowledgments
% The Funder and award ID information you input at submission will be introduced by the publisher under a Funding Information head during production. 
% Please use this space for any additional acknowledgements and verbiage required by your funders.

% \section{Supporting Information}
% This is an optional section. Please use this space to provide information about any supporting information referred to in your manuscript.

% \section{Competing Interests}
% This is an optional section. If you declared a conflict of interest when you submitted your manuscript, please  use this space to provide details about this conflict.

% \section{Technical Terms}

% All NETN article types require Technical Terms.

% Identify approximately 10 key terms that are mentioned in your article and whose usage and definition may not be familiar across the broad readership of the journal. 
% Provide brief (20-word or less) definitions for each term, avoiding in these definitions the use of jargon, or highly technical or specialized language. 
% When the article is typeset, the Technical Terms will appear in the margins at or near their first mention in the text.

% In your manuscript, bold the first occurrence of each \textbf{Technical Term} and then provide a list of the terms and their definitions at the end of the manuscript after the references.

%%%%%%%%%%%%%%%%%%%%%%%
%% The bibliography

%% The bibliography is made using only the entries that you cite using 
%% \cite{}, or one of the Natbib citation entries, like \citep{}, \citet{} etc.

\bibliography{references}

%ie.,
%\bibliography{bibsamp}

% \section{Technical Terms}

% \textbf{Technical Term} a key term that is mentioned in an NETN article and whose usage and definition may not be familiar across the broad readership of the journal.

% \textbf{Technical Term} a key term that is mentioned in an NETN article and whose usage and definition may not be familiar across the broad readership of the journal.

% \textbf{Technical Term} a key term that is mentioned in an NETN article and whose usage and definition may not be familiar across the broad readership of the journal.

% \textbf{Technical Term} a key term that is mentioned in an NETN article and whose usage and definition may not be familiar across the broad readership of the journal.

\end{document}

The examples below may be helpful when you are looking
for a quick example. They are taken from NETNsample.tex
so you can see their results in NETNSample.pdf

\begin{boxedtext}{box title}
Text
\end{boxedtext}

or, put figure in boxed text:
\begin{boxedtext}{box title}
Text
\begin{figure}
illustration
\caption{caption here}
\end{figure}
\end{boxedtext}

or, put table in boxed text:
\begin{boxedtext}{box title}
Text
\begin{table}
\caption{caption here}
tabular...
\end{table}
\end{boxedtext}

or, put anything in whitebox within boxed text:
\begin{boxedtext}{Showing the use of whitebox} 
Going beyond the examination of shared topological features across..

\whitebox{
From $\mathcal{W}$, we can estimate the variability in the fluctuations of the functional connection between nodes $i$ and $j$ over time as:
\begin{equation}
s_{ij}=\sqrt{\frac{1}{T-L}\sum_{t=1}^{T-L+1}(W_{ij}(t) - m_{ij})}
\end{equation}
where $m_{ij}=\frac{1}{T-L+1}\sum_{t=1}^{T-L+1}W_{ij}(t)$ is the mean
dynamic functional connectivity over time. 
}
...differences across a range of network classes.
\end{boxedtext}

==================
More samples:

\section{Sample Section}
Text here. Text here. Text here. Text here.
Text here. Text here. Text here. Text here.
Text here. Text here. Text here. Text here.
Text here. Text here. Text here. Text here.

\subsection{Sample Subsection}
Text here. Text here. Text here. Text here.
Text here. Text here. Text here. Text here.
Text here. Text here. Text here. Text here.
Text here. Text here. Text here. Text here.

\subsubsection{Sample Subsubsection}
Text here. Text here. Text here. Text here.
Text here. Text here. Text here. Text here.
Text here. Text here. Text here. Text here.
Text here. Text here. Text here. Text here.

\section{Sample equations}
\begin{equation}
\label{eq:rhoCHT}
\rho^{\pi}= \frac{RI + \mathbb{E}_{\pi([L,\tau_L]|\textrm{post})}
\left[C_L(\taupav+\tau_L) \right]   +
\displaystyle{\int_{0}^{P}}{dw~ \mathbb{E}_{\pi_{w_L}}}
\Biggl[\/\sum_{n_{L|[\textrm{pre},w]}}C_L(\tau_L)
\Biggr]            }      {P +
\mathbb{E}_{\pi([L,\tau_L] |\textrm{post})}[\tau_{L}] +\taupav +
\displaystyle{ \int_{0}^{P}}{dw~ \mathbb{E}_{\pi_{w_L}}}   
\Biggl[\sum_{n_{L|[\textrm{pre},w]}}\tau_L\Biggr]  
}
\end{equation}
As long as
$RI - K_LP > 
\frac{1}{\beta}$
\begin{equation}
%\def\theequation{5.1}
\left.\begin{array}{lrcl}
&\rho^{\pi} &=&  \displaystyle\frac{\beta ( RI + K_L \taupav )-1} {\beta
(P+\taupav )}    \\[12pt]
\hbox{and}\hbox to .25in{\hfill}&\mathbb{E}[\tau_L | \text{post}] &=&\displaystyle \frac{P+\taupav}{\beta ( RI -
K_LP)-1}  
\label{eq:analytical_linear}
\end{array}\right\}\hbox to 1.25in{\hfill}
\end{equation}

\subsection{Simple code sample}

\begin{code}
\begin{verbatim}
procedure bubbleSort( A : list of sortable items )
    n = length(A)
    repeat
       newn = 0
       for i = 1 to n-1 inclusive do
          if A[i-1] > A[i] then
             swap(A[i-1], A[i])
             newn = i
          end if
       end for
       n = newn
    until n = 0
end procedure
\end{verbatim}
\end{code}

\subsection{Algorithm environment}

%% \begin{algorithm} takes option [p][b][t][h],  or some combination, like \begin{figure}
%% See documentation for algorithmic.sty for more information on formatting algorithms.

\begin{algorithm}[h]
\caption{A sample in an algorithm environment.}
\begin{algorithmic}
\If {$i\geq maxval$}
    \State $i\gets 0$
\Else
    \If {$i+k\leq maxval$}
        \State $i\gets i+k$
    \EndIf
\EndIf
\end{algorithmic}
\end{algorithm}

\section{Technical Terms}
If you've been asked by the journal editor to provide technical terms and definitions, please bold the first occurrence of each term in the text and provide a list of the terms and their definitions at the end of the manuscript after the references. Please only supply technical terms if you've been instructed to do so by a journal editor.

\section{Itemized Lists}

\subsection{Roman list:}

\begin{enumerate}
\item[(i)] at high 
payoffs, subjects work almost continuously, engaging in little leisure
in between work bouts; 
\item[(ii)] at low payoffs, they 
engage in leisure all at once, in long bouts after working, rather
than distributing the same amount of leisure time into multiple short
leisure bouts; 
\item[(iii)] subjects work continuously for the entire price duration, as long as
the price is not very long (as shown by an analysis conducted by Y-AB, to be published separately); %(\textbf{Figure \ref{fig:task_data}D}).
\item[(iv)] the duration of leisure bouts is variable.
\end{enumerate}

\subsection{Numbered list:}

\begin{enumerate}
\item at high 
payoffs, subjects work almost continuously, engaging in little leisure
inbetween work bouts; 
\item at low payoffs, they 
engage in leisure all at once, in long bouts after working, rather
than distributing the same amount of leisure time into multiple short
leisure bouts; 
\item subjects work continuously for the entire price duration, as long as
the price is not very long (as shown by an analysis conducted by Y-AB, to be published separately); %(\textbf{Figure \ref{fig:task_data}D}).
\item the duration of leisure bouts is variable.
\end{enumerate}

\subsection{Bulleted list:}

\begin{itemize}
\item at high 
payoffs, subjects work almost continuously, engaging in little leisure
inbetween work bouts; 
\item at low payoffs, they 
engage in leisure all at once, in long bouts after working, rather
than distributing the same amount of leisure time into multiple short
leisure bouts; 
\item subjects work continuously for the entire price duration, as long as
the price is not very long (as shown by an analysis conducted by Y-AB, to be published separately); %(\textbf{Figure \ref{fig:task_data}D}).
\item the duration of leisure bouts is variable.
\end{itemize}

\section{Natbib citation mark up}

\subsection{Single citations}
\noindent
\begin{tabular}{ll}
\bf Type&\bf Results\\
\hline
\verb+\citet{jon90}+&Jones et al. (1990)\\
\verb+\citet[chap. 2]{jon90}+&Jones et al. (1990, chap. 2)\\
    \verb+\citep{jon90}+	    &   	(Jones et al., 1990)\\
    \verb+\citep[chap. 2]{jon90}+ 	&    	(Jones et al., 1990, chap. 2)\\
    \verb+\citep[see][]{jon90}+ 	 &    	(see Jones et al., 1990)\\
    \verb+\citep[see][chap. 2]{jon90}+ 	&    	(see Jones et al., 1990, chap. 2)\\
    \verb+\citet*{jon90}+ 	    &    	Jones, Baker, and Williams (1990)\\
    \verb+\citep*{jon90}+	    &    	(Jones, Baker, and Williams,
    1990) \\
\end{tabular}

For example, here are some sample citations:
citet:\citet{anderson}, citep:\citep{antibayes}, and
cite*: \citet*{anderson}.

\subsection{Multiple citations}
Multiple citations may be made by including more than one citation
key in the \verb+\cite+ command argument.

\noindent
\begin{tabular}{ll}
\bf Type&\bf Results\\
\hline
\verb+\citet{jon90,jam91}+&Jones et al. (1990); James et al. (1991)\\
\verb+\citep{jon90,jam91}+&(Jones et al., 1990; James et al. 1991)\\
\verb+\citep{jon90,jon91}+&(Jones et al., 1990, 1991)\\
\verb+\citep{jon90a,jon90b}+&(Jones et al., 1990a,b)\\
\end{tabular}

For example, multiple citations from the CompPsychSample bibliography:
citet:\citet{anderson,antibayes}, citep:\citep{anderson,antibayes}.
As you see, the citations are automatically hyperlinked to their
reference in the bibliography.

\newpage

\section{Sample figures}

\begin{figure}[h] 
\centerline{\includegraphics[width=\textwidth]{Fig1}}
\caption{(Colour online) \textbf{Task and key features of the
 data.} \\
 A) Cumulative handling time (CHT) task. Grey bars denote work
(depressing a lever), white gaps show leisure. The subject must
 accumulate work up to a total period of time called the
\emph{price} ($P$) in order to obtain a single reward (black dot) of subjective reward
intensity $RI$. The trial duration is $25\times \mathrm{price}$ (plus
$2$s each time the price is attained, during which the lever is retracted so it cannot
work; not shown).
}
\label{fig:task_data}
\end{figure}

\begin{figure}[ht] 
\widefigure{\fullpagewidth}{Fig1}
\caption{(Colour online) \textbf{Task and key features of the
 data.} \\
A) Cumulative handling time (CHT) task. Grey bars denote work
(depressing a lever), white gaps show leisure. The subject must
accumulate work up to a total period of time called the
\emph{price} ($P$) in order to obtain a single reward (black dot) of subjective reward
intensity $RI$. The trial duration is $25\times \mathrm{price}$ (plus
$2$s each time the price is attained, during which the lever is retracted so it cannot
work; not shown).
}
\label{newfig:task_data}
\end{figure}

\clearpage
\section{Sample tables}

\begin{table}[!ht]
\caption{Time of the Transition Between Phase 1 and Phase 2$^{a}$}
\label{tab:label}
\centering
\begin{tabular}{lc}
\hline
 Run  & Time (min)  \\
\hline
  $l1$  & 260   \\
  $l2$  & 300   \\
  $l3$  & 340   \\
  $h1$  & 270   \\
  $h2$  & 250   \\
  $h3$  & 380   \\
  $r1$  & 370   \\
  $r2$  & 390   \\
\hline
\multicolumn{2}{l}{$^{a}$Table note text here.}
\end{tabular}
\end{table}

\begin{table}[ht]
\widecaption{Sample table taken from [treu03]\label{tbl-1}}
\begin{widetable}
\advance\tabcolsep-1pt
\small
\begin{tabular}{ccrrccccccccc}
\hline
\bf 
POS &\bf  chip &\multicolumn1c{\bf ID} &\multicolumn1c{\bf X}
&\multicolumn1c{\bf Y} &\bf
RA &\bf DEC &\bf IAU$\pm$ $\delta$ IAU &\bf
IAP1$\pm$ $\delta$ IAP1 &\bf IAP2 $\pm$ $\delta$
IAP2 &\bf star &\bf E &\bf Comment\\
\hline
0 & 2 & 1 & 1370.99 & 57.35\rlap{$^a$}    &   6.651120 &  17.131149 &
21.344$\pm$0.006\rlap{$^b$}  & 2 4.385$\pm$0.016 & 23.528$\pm$0.013 & 0.0 & 9 & -    \\
0 & 2 & 2 & 1476.62 & 8.03     &   6.651480 &  17.129572 & 21.641$\pm$0.005  & 2 3.141$\pm$0.007 & 22.007$\pm$0.004 & 0.0 & 9 & -    \\
0 & 2 & 3 & 1079.62 & 28.92    &   6.652430 &  17.135000 & 23.953$\pm$0.030  & 2 4.890$\pm$0.023 & 24.240$\pm$0.023 & 0.0 & - & -    \\
0 & 2 & 4 & 114.58  & 21.22    &   6.655560 &  17.148020 & 23.801$\pm$0.025  & 2 5.039$\pm$0.026 & 24.112$\pm$0.021 & 0.0 & - & -    \\
0 & 2 & 5 & 46.78   & 19.46    &   6.655800 &  17.148932 & 23.012$\pm$0.012  & 2 3.924$\pm$0.012 & 23.282$\pm$0.011 & 0.0 & - & -    \\
0 & 2 & 6 & 1441.84 & 16.16    &   6.651480 &  17.130072 & 24.393$\pm$0.045  & 2 6.099$\pm$0.062 & 25.119$\pm$0.049 & 0.0 & - & -    \\
0 & 2 & 7 & 205.43  & 3.96     &   6.655520 &  17.146742 & 24.424$\pm$0.032  & 2 5.028$\pm$0.025 & 24.597$\pm$0.027 & 0.0 & - & -    \\
0 & 2 & 8 & 1321.63 & 9.76     &   6.651950 &  17.131672 &
22.189$\pm$0.011  & 2 4.743$\pm$0.021 & 23.298$\pm$0.011 & 0.0 & 4 &
edge \\
\hline
\multicolumn{13}{l}{%
Table 2 is published in its entirety in the electronic
edition of the {\it Astrophysical Journal}.}\\[3pt]
\multicolumn{13}{l}{%
$^a$ Sample footnote for table 2.}\\[3pt]
\multicolumn{13}{l}{%
$^b$ Another sample footnote for table 2.}
\end{tabular}
\end{widetable}
\end{table}

\begin{table}[p]
\rotatebox{90}{\vbox{\hsize=\textheight
\caption{Here is a caption for a table that is found in landscape
mode.}
\begin{tabular}{ccrrccccccccc}
\hline
\bf 
POS &\bf  chip &\multicolumn1c{\bf ID} &\multicolumn1c{\bf X}
&\multicolumn1c{\bf Y} &\bf
RA &\bf DEC &\bf IAU$\pm$ $\delta$ IAU &\bf
IAP1$\pm$ $\delta$ IAP1 &\bf IAP2 $\pm$ $\delta$
IAP2 &\bf star &\bf E &\bf Comment\\
\hline
0 & 2 & 1 & 1370.99 & 57.35\rlap{$^a$}    &   6.651120 &  17.131149 &
21.344$\pm$0.006\rlap{$^b$}  & 2 4.385$\pm$0.016 & 23.528$\pm$0.013 & 0.0 & 9 & -    \\
0 & 2 & 2 & 1476.62 & 8.03     &   6.651480 &  17.129572 & 21.641$\pm$0.005  & 2 3.141$\pm$0.007 & 22.007$\pm$0.004 & 0.0 & 9 & -    \\
0 & 2 & 3 & 1079.62 & 28.92    &   6.652430 &  17.135000 & 23.953$\pm$0.030  & 2 4.890$\pm$0.023 & 24.240$\pm$0.023 & 0.0 & - & -    \\
0 & 2 & 4 & 114.58  & 21.22    &   6.655560 &  17.148020 & 23.801$\pm$0.025  & 2 5.039$\pm$0.026 & 24.112$\pm$0.021 & 0.0 & - & -    \\
0 & 2 & 5 & 46.78   & 19.46    &   6.655800 &  17.148932 & 23.012$\pm$0.012  & 2 3.924$\pm$0.012 & 23.282$\pm$0.011 & 0.0 & - & -    \\
0 & 2 & 6 & 1441.84 & 16.16    &   6.651480 &  17.130072 & 24.393$\pm$0.045  & 2 6.099$\pm$0.062 & 25.119$\pm$0.049 & 0.0 & - & -    \\
0 & 2 & 7 & 205.43  & 3.96     &   6.655520 &  17.146742 & 24.424$\pm$0.032  & 2 5.028$\pm$0.025 & 24.597$\pm$0.027 & 0.0 & - & -    \\
0 & 2 & 8 & 1321.63 & 9.76     &   6.651950 &  17.131672 &
22.189$\pm$0.011  & 2 4.743$\pm$0.021 & 23.298$\pm$0.011 & 0.0 & 4 &
edge \\
\hline
\multicolumn{13}{l}{%
Table 2 is published in its entirety in the electronic
edition of the {\it Astrophysical Journal}.}\\[3pt]
\multicolumn{13}{l}{%
$^a$ Sample footnote for table 2.}\\[3pt]
\multicolumn{13}{l}{%
$^b$ Another sample footnote for table 2.}
\end{tabular}
}}
\end{table}
\clearpage

\vglue 3in
Example of table continuing over pages:

\begin{center}
\begin{longtable}{ccc@{}}
\caption{ApJ costs from 1991 to 2013
\label{tab:table}} \\[2pt]
\hline
\bf Year & \bf Subscription & \bf Publication \\
 & \bf cost &\bf charges\\
 & \bf(\$) & \bf (\$/page)\\
\hline
\endfirsthead

\multicolumn3c{Table \thetable, \it continued from previous page.}\\[6pt]
\multicolumn3c{ApJ costs from 1991 to 2013}\\[2pt]
\hline
\bf Year & \bf Subscription & \bf Publication \\
 & \bf cost &\bf charges\\
 & \bf(\$) & \bf (\$/page)\\
\hline
\endhead
\\\hline
\\[-8pt]
\multicolumn{3}{r}{\it Table continued on next page}\\ 
\endfoot

\hline
\endlastfoot

1991 & 600 & 100 \\
1992 & 650 & 105 \\
1993 & 550 & 103 \\
1994 & 450 & 110 \\
1995 & 410 & 112 \\
1996 & 400 & 114 \\
1997 & 525 & 115 \\
1998 & 590 & 116 \\
1999 & 575 & 115 \\
2000 & 450 & 103 \\
2001 & 490 &  90 \\
2002 & 500 &  88 \\
2003 & 450 &  90 \\
2004 & 460 &  88 \\
2005 & 440 &  79 \\
2006 & 350 &  77 \\
2007 & 325 &  70 \\
2008 & 320 &  65 \\
2009 & 190 &  68 \\
2010 & 280 &  70 \\
2011 & 275 &  68 \\
2012 & 150 &  56 \\
2013 & 140 &  55 \\
\end{longtable}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%
%% The bibliography

%% The bibliography is made using only
%% the entries that you cite using \cite{}, or one of the Natbib citation
%% entries, like \citep{}, \citet{} etc.

\bibliography{bibsamp}

NO APPENDICES allowed in the Network Neuroscience Style.

Please submit Supporting Information documents as PDFs in a format ready to publish.

### 2024-02-01 paper/equivalence-of-neural-field-dynamics-with-different-embedding-dimensionality.tex
introduction wip

@@
 %% and need to use lualatex:
-%% \documentclass[finalfonts,NETN]{stjour}
+% \documentclass[finalfonts,NETN]{stjour-new}
 
@@
 
-\title{Equivalence of neural field dynamics with different embedding dimensionality}
+% TODO: what title is better ?
+% \title{Equivalence of neural field dynamics with different embedding dimensionality}
+\title{Neural fields with different embedding dimensionality can express identical dynamics}
+% \title{Neural dynamics can be equivalently embedded in neural fields of different dimensions}
+
 % \subtitle{<Subtitle Here>} %% Optional subtitle
@@
 \begin{abstract}
-In neuroscience, neural field equations model interactions between neurons positioned in an "embedding"; a physical  (for instance, that of the cortical sheet) or abstract space (for instance, that of neuron celltypes) in which the dynamics are highly regular in space and time.
+Classical neural field models describe interactions between neurons positioned in an embedding; often a physical space (for instance, that of the cortical sheet) or, later, abstract spaces (for instance, that of neuron celltypes), in which the dynamics are highly regular in space and time.
+Recent improvements in recordings of neural activity have shined the light on the smoothness of the neural dynamics, when placed in an appropriate embedding space.
 
-This work contributes to the formalization of the search for neural field models, and in particular treats the question of the embedding space uniqueness by studying its dimensionality.
-We show that given some neural dynamics, we can find multiple equivalent embedding spaces in different dimensions, in the sense that they have the same dynamics.
+Decoupling the embedding space from the physical position of neurons begs the question of uniqueness of the embedding space.
+Given an embedding space and neural dynamics theron, are there other embedding spaces which can express the same behavior?
 
+We show that embedding spaces are not unique, and more suprisingly, that we can find multiple embedding spaces in different dimensions, which all model identical neural dynamics.
 The core of this work lies in the study of bijective mappings $S$ between $[0,1]^2$ and $[0,1]$, which allows mapping a two-dimensional neural field to an equivalent one-dimensional neural field.
 This procedure is easily generalized to map between embeddings in arbitrary number of dimensions.
-
-We illustrate this work through extensive numerical simulations and animations of a low-rank toy model.
-In particular, the coarse-graining procedure is introduced, in which numerical simulations of $2^n$ (where $n$ is an integer quantifying the resolution of the simulation grid) neural populations in a one-dimensional embedding converge to the same dynamics obtained from simulations of $2^n \times 2^n$ neural populations in a two-dimensional embedding.
-We additionally introduce a measure of locality, which quantifies how the regularity ("smoothness") of the neural fields is conserved through the mappings $S$.
-In particular, we use a subset composed of binary mappings, and borrow the Z-order curve from the computer graphics field, which we show is a "local" mapping.
-We finally prove that this locality guarantees that the one-dimensional neural field equations can be approximated using classical grid-based methods, and therefore the numerical simulations yield the correct equivalent dynamics.
+We introduce a measure of locality, which quantifies how the regularity of the neural fields in conserved through the mappings $S$.
+This allows us to show that the resulting neural field is well-behaved, and can therefore be approximated using classical grid-based methods.
+To illustrate these results through numerical models, we describe a coarse-graining procedure enforcing the locality of the numerically applied mappings.
 \end{abstract}
@@
 
+% NOTE: the goal here is to give an intuitive understanding of the problem,
+% kind of in the same "story-telling"-way that I did in the master thesis.
+% This also allows to introduce some of the notation that I use later on.
+
 \subsection{Neural field models in neuroscience}
 
-valentin's introduction with many citations. chap 1.1
+Early recordings of cats' somatosensory \citep{Mou57} and visual \citep{HubWie62} cortices suggested that the cortical sheet is organized in vertical columns of functionally similar neurons.
 
-\subsection{Motivation}
+The columnar organization of the cortex motivated the design of spatially structured models of neural population dynamics, namely neural field models \citep{WilCow73,Nun74,Ama77}, where the spatial dimensions corresponded to the two dimensions of the cortical sheet.
+
+As models of spatiotemporal neural population dynamics on the cortical sheet, neural field models can be used to explain experimentally observed patterns of cortical activity, such as travelling waves in visual \citep{SatNau12,MulRey14}, somatosensory \citep{PetHah03,FerBol06}, motor \citep{RubRob06,TakKim15}, and hippocampal \citep{LubSia09,PatFuj12,PatSch13} cortices (see \citep{MulCha18} for a review), and are used to model large-scale brain signals, such as electroencephalography (EEG) recordings \citep{Bre17} (see also \citep{DipRan18} for an example of neural field modelling of calcium imaging recording of the visual cortex).
+
+More recent recording methods, especially in rodents, have provided evidence of low-dimensional organization of neuronal activity that does not depend on the physical locations of neurons on the cortical sheet.
+For example, contrary to cats' visual cortex, the orientation selectivity of pyramidal neurons in a rat's visual cortex does not depend on their locations, that is, orientation selectivity is heterogeneous at any given location \citep{OhkChu05}.
+Moreover, the activity of interneurons in any small volume of the visual cortex is also heterogeneous and seems to be structured by a low-dimensional manifold of fine cell subtypes \citep{BugDuf22}.
+The local functional heterogeneity of neuronal activity in the cortex challenges the old concept of functionally homogeneous cortical columns and, with it, classical neural field models.
+
+While, historically, the spatial dimensions in neural field models have been the two dimensions of the cortical sheet (with the notable exception of the ring models for orientation selectivity in the visual cortex \citep{BenBar95} and the head-direction system \citep{Zha96}), the "space" in neural field models does not need to represent physical space but can represent any suitable abstract embedding space.
+Neural field models with abstract embedding spaces could constitute a generalization of classical neural field theory for cortical networks that are not solely structured by distances of neurons on the cortical sheet.
+
+Moreover, this work is motivated by recent mathematical developments \citep{JabPoy21} in mean-field theories involving the theory of graphons.
+Informally, these results suggest that, as long as the connectivity scales as $1/N$ (where $N$ is the number of neurons), the dynamics of large networks can be described in the $N \to \infty$ limit by a neural field equation where the embedding is simply the interval $[0,1]$.
+In the following, we will describe a method of explicitly writing such a one-dimensional neural field.
+
+\subsection{Formulating a one-dimensional neural field}
+
+% If we consider neural fields in abstract embedding spaces, which embedding space should we choose?
+More fundamentally, given some spatiotemporal dynamics, is there a "natural" choice for the embedding space?
+
+% We show that the answer to the second question is highly nontrivial because even the dimensionality of the embedding space, for some given spatiotemporal dynamics, is not clearly defined.
+This means that it is possible to find two embeddings (each associated with a connectivity kernel) of different dimensions that give rise to identical dynamics.
+
+% TODO: do we talk about the tau=1, R=1 ?
+% TODO: do we talk about external currents ?
+
+Let us write a neural field associated with a $p$-dimensional embedding $\Rp$.
+$h(\vec z, t)$ describes the potential of the neural population at position $\vec z=(z_1,\cdots,z_p)$ in the embedding.
+The connectivity kernel $w(\vec y, \vec z)$ describes the amplitude of the recurrent currents from the population at $\vec z$ to the population at $\vec y$.
+$\rho(\vec z)$ measures the number of neurons per unit volume of the embedding space, and we use the notation $\rho(\mathrm d \vec y) := \rho(\vec y) \mathrm d \vec y$ to emphasize that the integral is weighted by the neuron density.
+The activation function $\phi(h)$ describes the firing rate of a given neuron (population) with potential $h$.
+The neural field then evolves according to \autoref{eq:nf-in-rp}:
+
+\begin{equation} \label{eq:nf-in-rp}
+\partial_t h(\vec z, t) = \underbrace{-h(\vec z, t)}_\text{exponential decay} + \underbrace{\int_{\mathbb{R}^p} w(\vec z, \vec y) \phi(h(\vec y, t)) \rho(\mathrm d \vec y)}_\text{recurrent current}.
+\end{equation}
+
+The embedding space $\Rp$ can be mapped to $[0,1]^p$ via bijective functions between the two spaces.
+For instance, if the neuron density factorizes as $\rho(z_1, \cdots, z_p)= \rho_1(z_1) \cdots \rho_p(z_p)$, then the associated cumulative density functions can be used to perform a change of variables, resulting in the an equivalent "uniform" neural field equation:
+
+\begin{equation} \label{eq:nf-in-01p}
+\partial_t h_U(t, \vec v) = -h_U(t, \vec v) + \int_{[0,1]^p} w_U(\vec v, \vec u) \phi(h_U(t, \vec u)) \mathrm d \vec u.
+\end{equation}
+
+If the neuron density does not factorize, then one can arrive at the same formulation by redefining the connectivity kernel to include the neuron density and the determinant of the transformation.
+[TODO: this sentence might need some work]
+
+In particular for $p=2$, we show that for any two-dimensional neural field equation (that is, with two spatial dimensions), there is a one-dimensional neural field equation, where the embedding space is simply the interval $[0,1]$, from which the solution to the two-dimensional equation can be fully reconstructed.
+The mapping from the two-dimensional equation to the one-dimensional equation is done via well-known measurable bijections between the square $[0,1]^2$ and the interval $[0,1]$.
+Importantly, these bijections are not (and can't be, due to Netto's theorem [TODO: citation?]) diffeomorphisms.
+% Also, we show that although the connectivity kernel of the one-dimensional equation seems to have a fractal structure, it is sufficiently regular for its solution to be numerically approximated using standard grid-based simulations.
+The main idea is to define a measurable bijective mapping $S : [0,1]^2 \mapsto [0,1]$ and a measure $\lambda : [0,1]^2 \mapsto \mathbb{R}^+$, such that we can write a one-dimensional neural field	$\tilde h(\alpha, t)$ with a connectivity kernel defined as $\tilde w(\alpha, \beta) = w_U(S^{-1}(\alpha), S^{-1}(\beta))$.
+The question of well-posedness of the neural field in \autoref{eq:nf-in-01} is defered in \autoref{sec:proof-equivalence}.
+
+\begin{equation} \label{eq:nf-in-01}
+\partial_t \tilde h(\alpha, t) = -\tilde h(\alpha, t) + \int_{[0,1]} \tilde w(\alpha, \beta) \phi(\tilde h(\beta, t)) \left[\lambda \circ S^{-1}\right](\d \beta)
+\end{equation}
+
+\subsection{Regularity of the mapping}
+
+[chap 1.4] introduce a random mapping to introduce locality
+
+figure 1.2
 
@@
 
+"discretization error"
+
 argue over smoothness of the kernel (see presentation given for SPC)
-
-\subsection{Definitions}
-
-math definitions of neural fields and embedding spaces, reduction of 2d->1d and CDF, and the mappings
 
@@
 
-\subsection{Summary of findings}
+Our results suggest that the notion of (spatial) dimensionality in neural field dynamics is not well-defined if not associated with a constraint on the regularity of the neural fields (and the connectivity kernel).
+Thereby, by studying a simple toy model, this work illustrates the importance of the analytic notion of regularity when neural population dynamics over abstract continuous spaces are considered.
 
@@
 
+chap. 2.1,
+
+we stress that this toy model serves only for illustration purposes, and our results apply to any neural field
+
 introduce coarse-graining procedure (applying the smoothness hypothesis numerically)
@@
 \subsection{Analytical equivalence given $S$ measurable and bijective}
+\label{sec:proof-equivalence}
 
@@
 
-The examples below may be helpful when you are looking
-for a quick example. They are taken from NETNsample.tex
-so you can see their results in NETNSample.pdf
+% The examples below may be helpful when you are looking
+% for a quick example. They are taken from NETNsample.tex
+% so you can see their results in NETNSample.pdf
 
-\begin{boxedtext}{box title}
-Text
-\end{boxedtext}
+% \begin{boxedtext}{box title}
+% Text
+% \end{boxedtext}
 
-or, put figure in boxed text:
-\begin{boxedtext}{box title}
-Text
-\begin{figure}
-illustration
-\caption{caption here}
-\end{figure}
-\end{boxedtext}
+% or, put figure in boxed text:
+% \begin{boxedtext}{box title}
+% Text
+% \begin{figure}
+% illustration
+% \caption{caption here}
+% \end{figure}
+% \end{boxedtext}
 
-or, put table in boxed text:
-\begin{boxedtext}{box title}
-Text
-\begin{table}
-\caption{caption here}
-tabular...
-\end{table}
-\end{boxedtext}
+% or, put table in boxed text:
+% \begin{boxedtext}{box title}
+% Text
+% \begin{table}
+% \caption{caption here}
+% tabular...
+% \end{table}
+% \end{boxedtext}
 
-or, put anything in whitebox within boxed text:
-\begin{boxedtext}{Showing the use of whitebox} 
-Going beyond the examination of shared topological features across..
+% or, put anything in whitebox within boxed text:
+% \begin{boxedtext}{Showing the use of whitebox} 
+% Going beyond the examination of shared topological features across..
 
-\whitebox{
-From $\mathcal{W}$, we can estimate the variability in the fluctuations of the functional connection between nodes $i$ and $j$ over time as:
-\begin{equation}
-s_{ij}=\sqrt{\frac{1}{T-L}\sum_{t=1}^{T-L+1}(W_{ij}(t) - m_{ij})}
-\end{equation}
-where $m_{ij}=\frac{1}{T-L+1}\sum_{t=1}^{T-L+1}W_{ij}(t)$ is the mean
-dynamic functional connectivity over time. 
-}
-...differences across a range of network classes.
-\end{boxedtext}
+% \whitebox{
+% From $\mathcal{W}$, we can estimate the variability in the fluctuations of the functional connection between nodes $i$ and $j$ over time as:
+% \begin{equation}
+% s_{ij}=\sqrt{\frac{1}{T-L}\sum_{t=1}^{T-L+1}(W_{ij}(t) - m_{ij})}
+% \end{equation}
+% where $m_{ij}=\frac{1}{T-L+1}\sum_{t=1}^{T-L+1}W_{ij}(t)$ is the mean
+% dynamic functional connectivity over time. 
+% }
+% ...differences across a range of network classes.
+% \end{boxedtext}
 
-==================
-More samples:
+% ==================
+% More samples:
 
-\section{Sample Section}
-Text here. Text here. Text here. Text here.
-Text here. Text here. Text here. Text here.
-Text here. Text here. Text here. Text here.
-Text here. Text here. Text here. Text here.
+% \section{Sample Section}
+% Text here. Text here. Text here. Text here.
+% Text here. Text here. Text here. Text here.
+% Text here. Text here. Text here. Text here.
+% Text here. Text here. Text here. Text here.
 
-\subsection{Sample Subsection}
-Text here. Text here. Text here. Text here.
-Text here. Text here. Text here. Text here.
-Text here. Text here. Text here. Text here.
-Text here. Text here. Text here. Text here.
+% \subsection{Sample Subsection}
+% Text here. Text here. Text here. Text here.
+% Text here. Text here. Text here. Text here.
+% Text here. Text here. Text here. Text here.
+% Text here. Text here. Text here. Text here.
 
-\subsubsection{Sample Subsubsection}
-Text here. Text here. Text here. Text here.
-Text here. Text here. Text here. Text here.
-Text here. Text here. Text here. Text here.
-Text here. Text here. Text here. Text here.
+% \subsubsection{Sample Subsubsection}
+% Text here. Text here. Text here. Text here.
+% Text here. Text here. Text here. Text here.
+% Text here. Text here. Text here. Text here.
+% Text here. Text here. Text here. Text here.
 
-\section{Sample equations}
-\begin{equation}
-\label{eq:rhoCHT}
-\rho^{\pi}= \frac{RI + \mathbb{E}_{\pi([L,\tau_L]|\textrm{post})}
-\left[C_L(\taupav+\tau_L) \right]   +
-\displaystyle{\int_{0}^{P}}{dw~ \mathbb{E}_{\pi_{w_L}}}
-\Biggl[\/\sum_{n_{L|[\textrm{pre},w]}}C_L(\tau_L)
-\Biggr]            }      {P +
-\mathbb{E}_{\pi([L,\tau_L] |\textrm{post})}[\tau_{L}] +\taupav +
-\displaystyle{ \int_{0}^{P}}{dw~ \mathbb{E}_{\pi_{w_L}}}   
-\Biggl[\sum_{n_{L|[\textrm{pre},w]}}\tau_L\Biggr]  
-}
-\end{equation}
-As long as
-$RI - K_LP > 
-\frac{1}{\beta}$
-\begin{equation}
-%\def\theequation{5.1}
-\left.\begin{array}{lrcl}
-&\rho^{\pi} &=&  \displaystyle\frac{\beta ( RI + K_L \taupav )-1} {\beta
-(P+\taupav )}    \\[12pt]
-\hbox{and}\hbox to .25in{\hfill}&\mathbb{E}[\tau_L | \text{post}] &=&\displaystyle \frac{P+\taupav}{\beta ( RI -
-K_LP)-1}  
-\label{eq:analytical_linear}
-\end{array}\right\}\hbox to 1.25in{\hfill}
-\end{equation}
+% \section{Sample equations}
+% \begin{equation}
+% \label{eq:rhoCHT}
+% \rho^{\pi}= \frac{RI + \mathbb{E}_{\pi([L,\tau_L]|\textrm{post})}
+% \left[C_L(\taupav+\tau_L) \right]   +
+% \displaystyle{\int_{0}^{P}}{dw~ \mathbb{E}_{\pi_{w_L}}}
+% \Biggl[\/\sum_{n_{L|[\textrm{pre},w]}}C_L(\tau_L)
+% \Biggr]            }      {P +
+% \mathbb{E}_{\pi([L,\tau_L] |\textrm{post})}[\tau_{L}] +\taupav +
+% \displaystyle{ \int_{0}^{P}}{dw~ \mathbb{E}_{\pi_{w_L}}}   
+% \Biggl[\sum_{n_{L|[\textrm{pre},w]}}\tau_L\Biggr]  
+% }
+% \end{equation}
+% As long as
+% $RI - K_LP > 
+% \frac{1}{\beta}$
+% \begin{equation}
+% %\def\theequation{5.1}
+% \left.\begin{array}{lrcl}
+% &\rho^{\pi} &=&  \displaystyle\frac{\beta ( RI + K_L \taupav )-1} {\beta
+% (P+\taupav )}    \\[12pt]
+% \hbox{and}\hbox to .25in{\hfill}&\mathbb{E}[\tau_L | \text{post}] &=&\displaystyle \frac{P+\taupav}{\beta ( RI -
+% K_LP)-1}  
+% \label{eq:analytical_linear}
+% \end{array}\right\}\hbox to 1.25in{\hfill}
+% \end{equation}
 
-\subsection{Simple code sample}
+% \subsection{Simple code sample}
 
-\begin{code}
-\begin{verbatim}
-procedure bubbleSort( A : list of sortable items )
-    n = length(A)
-    repeat
-       newn = 0
-       for i = 1 to n-1 inclusive do
-          if A[i-1] > A[i] then
-             swap(A[i-1], A[i])
-             newn = i
-          end if
-       end for
-       n = newn
-    until n = 0
-end procedure
-\end{verbatim}
-\end{code}
+% \begin{code}
+% \begin{verbatim}
+% procedure bubbleSort( A : list of sortable items )
+%     n = length(A)
+%     repeat
+%        newn = 0
+%        for i = 1 to n-1 inclusive do
+%           if A[i-1] > A[i] then
+%              swap(A[i-1], A[i])
+%              newn = i
+%           end if
+%        end for
+%        n = newn
+%     until n = 0
+% end procedure
+% \end{verbatim}
+% \end{code}
 
-\subsection{Algorithm environment}
+% \subsection{Algorithm environment}
 
-%% \begin{algorithm} takes option [p][b][t][h],  or some combination, like \begin{figure}
-%% See documentation for algorithmic.sty for more information on formatting algorithms.
+% %% \begin{algorithm} takes option [p][b][t][h],  or some combination, like \begin{figure}
+% %% See documentation for algorithmic.sty for more information on formatting algorithms.
 
-\begin{algorithm}[h]
-\caption{A sample in an algorithm environment.}
-\begin{algorithmic}
-\If {$i\geq maxval$}
-    \State $i\gets 0$
-\Else
-    \If {$i+k\leq maxval$}
-        \State $i\gets i+k$
-    \EndIf
-\EndIf
-\end{algorithmic}
-\end{algorithm}
+% \begin{algorithm}[h]
+% \caption{A sample in an algorithm environment.}
+% \begin{algorithmic}
+% \If {$i\geq maxval$}
+%     \State $i\gets 0$
+% \Else
+%     \If {$i+k\leq maxval$}
+%         \State $i\gets i+k$
+%     \EndIf
+% \EndIf
+% \end{algorithmic}
+% \end{algorithm}
 
-\section{Technical Terms}
-If you've been asked by the journal editor to provide technical terms and definitions, please bold the first occurrence of each term in the text and provide a list of the terms and their definitions at the end of the manuscript after the references.
+% \section{Technical Terms}
+% If you've been asked by the journal editor to provide technical terms and definitions, please bold the first occurrence of each term in the text and provide a list of the terms and their definitions at the end of the manuscript after the references.
 Please only supply technical terms if you've been instructed to do so by a journal editor.
 
-\section{Itemized Lists}
+% \section{Itemized Lists}
 
-\subsection{Roman list:}
+% \subsection{Roman list:}
 
-\begin{enumerate}
-\item[(i)] at high 
-payoffs, subjects work almost continuously, engaging in little leisure
-in between work bouts; 
-\item[(ii)] at low payoffs, they 
-engage in leisure all at once, in long bouts after working, rather
-than distributing the same amount of leisure time into multiple short
-leisure bouts; 
-\item[(iii)] subjects work continuously for the entire price duration, as long as
-the price is not very long (as shown by an analysis conducted by Y-AB, to be published separately); %(\textbf{Figure \ref{fig:task_data}D}).
-\item[(iv)] the duration of leisure bouts is variable.
-\end{enumerate}
+% \begin{enumerate}
+% \item[(i)] at high 
+% payoffs, subjects work almost continuously, engaging in little leisure
+% in between work bouts; 
+% \item[(ii)] at low payoffs, they 
+% engage in leisure all at once, in long bouts after working, rather
+% than distributing the same amount of leisure time into multiple short
+% leisure bouts; 
+% \item[(iii)] subjects work continuously for the entire price duration, as long as
+% the price is not very long (as shown by an analysis conducted by Y-AB, to be published separately); %(\textbf{Figure \ref{fig:task_data}D}).
+% \item[(iv)] the duration of leisure bouts is variable.
+% \end{enumerate}
 
-\subsection{Numbered list:}
+% \subsection{Numbered list:}
 
-\begin{enumerate}
-\item at high 
-payoffs, subjects work almost continuously, engaging in little leisure
-inbetween work bouts; 
-\item at low payoffs, they 
-engage in leisure all at once, in long bouts after working, rather
-than distributing the same amount of leisure time into multiple short
-leisure bouts; 
-\item subjects work continuously for the entire price duration, as long as
-the price is not very long (as shown by an analysis conducted by Y-AB, to be published separately); %(\textbf{Figure \ref{fig:task_data}D}).
-\item the duration of leisure bouts is variable.
-\end{enumerate}
+% \begin{enumerate}
+% \item at high 
+% payoffs, subjects work almost continuously, engaging in little leisure
+% inbetween work bouts; 
+% \item at low payoffs, they 
+% engage in leisure all at once, in long bouts after working, rather
+% than distributing the same amount of leisure time into multiple short
+% leisure bouts; 
+% \item subjects work continuously for the entire price duration, as long as
+% the price is not very long (as shown by an analysis conducted by Y-AB, to be published separately); %(\textbf{Figure \ref{fig:task_data}D}).
+% \item the duration of leisure bouts is variable.
+% \end{enumerate}
 
-\subsection{Bulleted list:}
+% \subsection{Bulleted list:}
 
-\begin{itemize}
-\item at high 
-payoffs, subjects work almost continuously, engaging in little leisure
-inbetween work bouts; 
-\item at low payoffs, they 
-engage in leisure all at once, in long bouts after working, rather
-than distributing the same amount of leisure time into multiple short
-leisure bouts; 
-\item subjects work continuously for the entire price duration, as long as
-the price is not very long (as shown by an analysis conducted by Y-AB, to be published separately); %(\textbf{Figure \ref{fig:task_data}D}).
-\item the duration of leisure bouts is variable.
-\end{itemize}
+% \begin{itemize}
+% \item at high 
+% payoffs, subjects work almost continuously, engaging in little leisure
+% inbetween work bouts; 
+% \item at low payoffs, they 
+% engage in leisure all at once, in long bouts after working, rather
+% than distributing the same amount of leisure time into multiple short
+% leisure bouts; 
+% \item subjects work continuously for the entire price duration, as long as
+% the price is not very long (as shown by an analysis conducted by Y-AB, to be published separately); %(\textbf{Figure \ref{fig:task_data}D}).
+% \item the duration of leisure bouts is variable.
+% \end{itemize}
 
-\section{Natbib citation mark up}
+% \section{Natbib citation mark up}
 
-\subsection{Single citations}
-\noindent
-\begin{tabular}{ll}
-\bf Type&\bf Results\\
-\hline
-\verb+\citet{jon90}+&Jones et al. (1990)\\
-\verb+\citet[chap. 2]{jon90}+&Jones et al. (1990, chap. 2)\\
-    \verb+\citep{jon90}+	    &   	(Jones et al., 1990)\\
-    \verb+\citep[chap. 2]{jon90}+ 	&    	(Jones et al., 1990, chap. 2)\\
-    \verb+\citep[see][]{jon90}+ 	 &    	(see Jones et al., 1990)\\
-    \verb+\citep[see][chap. 2]{jon90}+ 	&    	(see Jones et al., 1990, chap. 2)\\
-    \verb+\citet*{jon90}+ 	    &    	Jones, Baker, and Williams (1990)\\
-    \verb+\citep*{jon90}+	    &    	(Jones, Baker, and Williams,
-    1990) \\
-\end{tabular}
+% \subsection{Single citations}
+% \noindent
+% \begin{tabular}{ll}
+% \bf Type&\bf Results\\
+% \hline
+% \verb+\citet{jon90}+&Jones et al. (1990)\\
+% \verb+\citet[chap. 2]{jon90}+&Jones et al. (1990, chap. 2)\\
+%     \verb+\citep{jon90}+	    &   	(Jones et al., 1990)\\
+%     \verb+\citep[chap. 2]{jon90}+ 	&    	(Jones et al., 1990, chap. 2)\\
+%     \verb+\citep[see][]{jon90}+ 	 &    	(see Jones et al., 1990)\\
+%     \verb+\citep[see][chap. 2]{jon90}+ 	&    	(see Jones et al., 1990, chap. 2)\\
+%     \verb+\citet*{jon90}+ 	    &    	Jones, Baker, and Williams (1990)\\
+%     \verb+\citep*{jon90}+	    &    	(Jones, Baker, and Williams,
+%     1990) \\
+% \end{tabular}
 
-For example, here are some sample citations:
-citet:\citet{anderson}, citep:\citep{antibayes}, and
-cite*: \citet*{anderson}.
+% For example, here are some sample citations:
+% citet:\citet{anderson}, citep:\citep{antibayes}, and
+% cite*: \citet*{anderson}.
 
-\subsection{Multiple citations}
-Multiple citations may be made by including more than one citation
-key in the \verb+\cite+ command argument.
+% \subsection{Multiple citations}
+% Multiple citations may be made by including more than one citation
+% key in the \verb+\cite+ command argument.
 
-\noindent
-\begin{tabular}{ll}
-\bf Type&\bf Results\\
-\hline
-\verb+\citet{jon90,jam91}+&Jones et al. (1990); James et al. (1991)\\
-\verb+\citep{jon90,jam91}+&(Jones et al., 1990; James et al. 1991)\\
-\verb+\citep{jon90,jon91}+&(Jones et al., 1990, 1991)\\
-\verb+\citep{jon90a,jon90b}+&(Jones et al., 1990a,b)\\
-\end{tabular}
+% \noindent
+% \begin{tabular}{ll}
+% \bf Type&\bf Results\\
+% \hline
+% \verb+\citet{jon90,jam91}+&Jones et al. (1990); James et al. (1991)\\
+% \verb+\citep{jon90,jam91}+&(Jones et al., 1990; James et al. 1991)\\
+% \verb+\citep{jon90,jon91}+&(Jones et al., 1990, 1991)\\
+% \verb+\citep{jon90a,jon90b}+&(Jones et al., 1990a,b)\\
+% \end{tabular}
 
-For example, multiple citations from the CompPsychSample bibliography:
-citet:\citet{anderson,antibayes}, citep:\citep{anderson,antibayes}.
-As you see, the citations are automatically hyperlinked to their
-reference in the bibliography.
+% For example, multiple citations from the CompPsychSample bibliography:
+% citet:\citet{anderson,antibayes}, citep:\citep{anderson,antibayes}.
+% As you see, the citations are automatically hyperlinked to their
+% reference in the bibliography.
 
-\newpage
+% \newpage
 
-\section{Sample figures}
+% \section{Sample figures}
 
-\begin{figure}[h] 
-\centerline{\includegraphics[width=\textwidth]{Fig1}}
-\caption{(Colour online) \textbf{Task and key features of the
- data.} \\
- A) Cumulative handling time (CHT) task. Grey bars denote work
-(depressing a lever), white gaps show leisure. The subject must
- accumulate work up to a total period of time called the
-\emph{price} ($P$) in order to obtain a single reward (black dot) of subjective reward
-intensity $RI$. The trial duration is $25\times \mathrm{price}$ (plus
-$2$s each time the price is attained, during which the lever is retracted so it cannot
-work; not shown).
-}
-\label{fig:task_data}
-\end{figure}
+% \begin{figure}[h] 
+% \centerline{\includegraphics[width=\textwidth]{Fig1}}
+% \caption{(Colour online) \textbf{Task and key features of the
+%  data.} \\
+%  A) Cumulative handling time (CHT) task. Grey bars denote work
+% (depressing a lever), white gaps show leisure. The subject must
+%  accumulate work up to a total period of time called the
+% \emph{price} ($P$) in order to obtain a single reward (black dot) of subjective reward
+% intensity $RI$. The trial duration is $25\times \mathrm{price}$ (plus
+% $2$s each time the price is attained, during which the lever is retracted so it cannot
+% work; not shown).
+% }
+% \label{fig:task_data}
+% \end{figure}
 
-\begin{figure}[ht] 
-\widefigure{\fullpagewidth}{Fig1}
-\caption{(Colour online) \textbf{Task and key features of the
- data.} \\
-A) Cumulative handling time (CHT) task. Grey bars denote work
-(depressing a lever), white gaps show leisure. The subject must
-accumulate work up to a total period of time called the
-\emph{price} ($P$) in order to obtain a single reward (black dot) of subjective reward
-intensity $RI$. The trial duration is $25\times \mathrm{price}$ (plus
-$2$s each time the price is attained, during which the lever is retracted so it cannot
-work; not shown).
-}
-\label{newfig:task_data}
-\end{figure}
+% \begin{figure}[ht] 
+% \widefigure{\fullpagewidth}{Fig1}
+% \caption{(Colour online) \textbf{Task and key features of the
+%  data.} \\
+% A) Cumulative handling time (CHT) task. Grey bars denote work
+% (depressing a lever), white gaps show leisure. The subject must
+% accumulate work up to a total period of time called the
+% \emph{price} ($P$) in order to obtain a single reward (black dot) of subjective reward
+% intensity $RI$. The trial duration is $25\times \mathrm{price}$ (plus
+% $2$s each time the price is attained, during which the lever is retracted so it cannot
+% work; not shown).
+% }
+% \label{newfig:task_data}
+% \end{figure}
 
-\clearpage
-\section{Sample tables}
+% \clearpage
+% \section{Sample tables}
 
-\begin{table}[!ht]
-\caption{Time of the Transition Between Phase 1 and Phase 2$^{a}$}
-\label{tab:label}
-\centering
-\begin{tabular}{lc}
-\hline
- Run  & Time (min)  \\
-\hline
-  $l1$  & 260   \\
-  $l2$  & 300   \\
-  $l3$  & 340   \\
-  $h1$  & 270   \\
-  $h2$  & 250   \\
-  $h3$  & 380   \\
-  $r1$  & 370   \\
-  $r2$  & 390   \\
-\hline
-\multicolumn{2}{l}{$^{a}$Table note text here.}
-\end{tabular}
-\end{table}
+% \begin{table}[!ht]
+% \caption{Time of the Transition Between Phase 1 and Phase 2$^{a}$}
+% \label{tab:label}
+% \centering
+% \begin{tabular}{lc}
+% \hline
+%  Run  & Time (min)  \\
+% \hline
+%   $l1$  & 260   \\
+%   $l2$  & 300   \\
+%   $l3$  & 340   \\
+%   $h1$  & 270   \\
+%   $h2$  & 250   \\
+%   $h3$  & 380   \\
+%   $r1$  & 370   \\
+%   $r2$  & 390   \\
+% \hline
+% \multicolumn{2}{l}{$^{a}$Table note text here.}
+% \end{tabular}
+% \end{table}
 
-\begin{table}[ht]
-\widecaption{Sample table taken from [treu03]\label{tbl-1}}
-\begin{widetable}
-\advance\tabcolsep-1pt
-\small
-\begin{tabular}{ccrrccccccccc}
-\hline
-\bf 
-POS &\bf  chip &\multicolumn1c{\bf ID} &\multicolumn1c{\bf X}
-&\multicolumn1c{\bf Y} &\bf
-RA &\bf DEC &\bf IAU$\pm$ $\delta$ IAU &\bf
-IAP1$\pm$ $\delta$ IAP1 &\bf IAP2 $\pm$ $\delta$
-IAP2 &\bf star &\bf E &\bf Comment\\
-\hline
-0 & 2 & 1 & 1370.99 & 57.35\rlap{$^a$}    &   6.651120 &  17.131149 &
-21.344$\pm$0.006\rlap{$^b$}  & 2 4.385$\pm$0.016 & 23.528$\pm$0.013 & 0.0 & 9 & -    \\
-0 & 2 & 2 & 1476.62 & 8.03     &   6.651480 &  17.129572 & 21.641$\pm$0.005  & 2 3.141$\pm$0.007 & 22.007$\pm$0.004 & 0.0 & 9 & -    \\
-0 & 2 & 3 & 1079.62 & 28.92    &   6.652430 &  17.135000 & 23.953$\pm$0.030  & 2 4.890$\pm$0.023 & 24.240$\pm$0.023 & 0.0 & - & -    \\
-0 & 2 & 4 & 114.58  & 21.22    &   6.655560 &  17.148020 & 23.801$\pm$0.025  & 2 5.039$\pm$0.026 & 24.112$\pm$0.021 & 0.0 & - & -    \\
-0 & 2 & 5 & 46.78   & 19.46    &   6.655800 &  17.148932 & 23.012$\pm$0.012  & 2 3.924$\pm$0.012 & 23.282$\pm$0.011 & 0.0 & - & -    \\
-0 & 2 & 6 & 1441.84 & 16.16    &   6.651480 &  17.130072 & 24.393$\pm$0.045  & 2 6.099$\pm$0.062 & 25.119$\pm$0.049 & 0.0 & - & -    \\
-0 & 2 & 7 & 205.43  & 3.96     &   6.655520 &  17.146742 & 24.424$\pm$0.032  & 2 5.028$\pm$0.025 & 24.597$\pm$0.027 & 0.0 & - & -    \\
-0 & 2 & 8 & 1321.63 & 9.76     &   6.651950 &  17.131672 &
-22.189$\pm$0.011  & 2 4.743$\pm$0.021 & 23.298$\pm$0.011 & 0.0 & 4 &
-edge \\
-\hline
-\multicolumn{13}{l}{%
-Table 2 is published in its entirety in the electronic
-edition of the {\it Astrophysical Journal}.}\\[3pt]
-\multicolumn{13}{l}{%
-$^a$ Sample footnote for table 2.}\\[3pt]
-\multicolumn{13}{l}{%
-$^b$ Another sample footnote for table 2.}
-\end{tabular}
-\end{widetable}
-\end{table}
+% \begin{table}[ht]
+% \widecaption{Sample table taken from [treu03]\label{tbl-1}}
+% \begin{widetable}
+% \advance\tabcolsep-1pt
+% \small
+% \begin{tabular}{ccrrccccccccc}
+% \hline
+% \bf 
+% POS &\bf  chip &\multicolumn1c{\bf ID} &\multicolumn1c{\bf X}
+% &\multicolumn1c{\bf Y} &\bf
+% RA &\bf DEC &\bf IAU$\pm$ $\delta$ IAU &\bf
+% IAP1$\pm$ $\delta$ IAP1 &\bf IAP2 $\pm$ $\delta$
+% IAP2 &\bf star &\bf E &\bf Comment\\
+% \hline
+% 0 & 2 & 1 & 1370.99 & 57.35\rlap{$^a$}    &   6.651120 &  17.131149 &
+% 21.344$\pm$0.006\rlap{$^b$}  & 2 4.385$\pm$0.016 & 23.528$\pm$0.013 & 0.0 & 9 & -    \\
+% 0 & 2 & 2 & 1476.62 & 8.03     &   6.651480 &  17.129572 & 21.641$\pm$0.005  & 2 3.141$\pm$0.007 & 22.007$\pm$0.004 & 0.0 & 9 & -    \\
+% 0 & 2 & 3 & 1079.62 & 28.92    &   6.652430 &  17.135000 & 23.953$\pm$0.030  & 2 4.890$\pm$0.023 & 24.240$\pm$0.023 & 0.0 & - & -    \\
+% 0 & 2 & 4 & 114.58  & 21.22    &   6.655560 &  17.148020 & 23.801$\pm$0.025  & 2 5.039$\pm$0.026 & 24.112$\pm$0.021 & 0.0 & - & -    \\
+% 0 & 2 & 5 & 46.78   & 19.46    &   6.655800 &  17.148932 & 23.012$\pm$0.012  & 2 3.924$\pm$0.012 & 23.282$\pm$0.011 & 0.0 & - & -    \\
+% 0 & 2 & 6 & 1441.84 & 16.16    &   6.651480 &  17.130072 & 24.393$\pm$0.045  & 2 6.099$\pm$0.062 & 25.119$\pm$0.049 & 0.0 & - & -    \\
+% 0 & 2 & 7 & 205.43  & 3.96     &   6.655520 &  17.146742 & 24.424$\pm$0.032  & 2 5.028$\pm$0.025 & 24.597$\pm$0.027 & 0.0 & - & -    \\
+% 0 & 2 & 8 & 1321.63 & 9.76     &   6.651950 &  17.131672 &
+% 22.189$\pm$0.011  & 2 4.743$\pm$0.021 & 23.298$\pm$0.011 & 0.0 & 4 &
+% edge \\
+% \hline
+% \multicolumn{13}{l}{%
+% Table 2 is published in its entirety in the electronic
+% edition of the {\it Astrophysical Journal}.}\\[3pt]
+% \multicolumn{13}{l}{%
+% $^a$ Sample footnote for table 2.}\\[3pt]
+% \multicolumn{13}{l}{%
+% $^b$ Another sample footnote for table 2.}
+% \end{tabular}
+% \end{widetable}
+% \end{table}
 
-\begin{table}[p]
-\rotatebox{90}{\vbox{\hsize=\textheight
-\caption{Here is a caption for a table that is found in landscape
-mode.}
-\begin{tabular}{ccrrccccccccc}
-\hline
-\bf 
-POS &\bf  chip &\multicolumn1c{\bf ID} &\multicolumn1c{\bf X}
-&\multicolumn1c{\bf Y} &\bf
-RA &\bf DEC &\bf IAU$\pm$ $\delta$ IAU &\bf
-IAP1$\pm$ $\delta$ IAP1 &\bf IAP2 $\pm$ $\delta$
-IAP2 &\bf star &\bf E &\bf Comment\\
-\hline
-0 & 2 & 1 & 1370.99 & 57.35\rlap{$^a$}    &   6.651120 &  17.131149 &
-21.344$\pm$0.006\rlap{$^b$}  & 2 4.385$\pm$0.016 & 23.528$\pm$0.013 & 0.0 & 9 & -    \\
-0 & 2 & 2 & 1476.62 & 8.03     &   6.651480 &  17.129572 & 21.641$\pm$0.005  & 2 3.141$\pm$0.007 & 22.007$\pm$0.004 & 0.0 & 9 & -    \\
-0 & 2 & 3 & 1079.62 & 28.92    &   6.652430 &  17.135000 & 23.953$\pm$0.030  & 2 4.890$\pm$0.023 & 24.240$\pm$0.023 & 0.0 & - & -    \\
-0 & 2 & 4 & 114.58  & 21.22    &   6.655560 &  17.148020 & 23.801$\pm$0.025  & 2 5.039$\pm$0.026 & 24.112$\pm$0.021 & 0.0 & - & -    \\
-0 & 2 & 5 & 46.78   & 19.46    &   6.655800 &  17.148932 & 23.012$\pm$0.012  & 2 3.924$\pm$0.012 & 23.282$\pm$0.011 & 0.0 & - & -    \\
-0 & 2 & 6 & 1441.84 & 16.16    &   6.651480 &  17.130072 & 24.393$\pm$0.045  & 2 6.099$\pm$0.062 & 25.119$\pm$0.049 & 0.0 & - & -    \\
-0 & 2 & 7 & 205.43  & 3.96     &   6.655520 &  17.146742 & 24.424$\pm$0.032  & 2 5.028$\pm$0.025 & 24.597$\pm$0.027 & 0.0 & - & -    \\
-0 & 2 & 8 & 1321.63 & 9.76     &   6.651950 &  17.131672 &
-22.189$\pm$0.011  & 2 4.743$\pm$0.021 & 23.298$\pm$0.011 & 0.0 & 4 &
-edge \\
-\hline
-\multicolumn{13}{l}{%
-Table 2 is published in its entirety in the electronic
-edition of the {\it Astrophysical Journal}.}\\[3pt]
-\multicolumn{13}{l}{%
-$^a$ Sample footnote for table 2.}\\[3pt]
-\multicolumn{13}{l}{%
-$^b$ Another sample footnote for table 2.}
-\end{tabular}
-}}
-\end{table}
-\clearpage
+% \begin{table}[p]
+% \rotatebox{90}{\vbox{\hsize=\textheight
+% \caption{Here is a caption for a table that is found in landscape
+% mode.}
+% \begin{tabular}{ccrrccccccccc}
+% \hline
+% \bf 
+% POS &\bf  chip &\multicolumn1c{\bf ID} &\multicolumn1c{\bf X}
+% &\multicolumn1c{\bf Y} &\bf
+% RA &\bf DEC &\bf IAU$\pm$ $\delta$ IAU &\bf
+% IAP1$\pm$ $\delta$ IAP1 &\bf IAP2 $\pm$ $\delta$
+% IAP2 &\bf star &\bf E &\bf Comment\\
+% \hline
+% 0 & 2 & 1 & 1370.99 & 57.35\rlap{$^a$}    &   6.651120 &  17.131149 &
+% 21.344$\pm$0.006\rlap{$^b$}  & 2 4.385$\pm$0.016 & 23.528$\pm$0.013 & 0.0 & 9 & -    \\
+% 0 & 2 & 2 & 1476.62 & 8.03     &   6.651480 &  17.129572 & 21.641$\pm$0.005  & 2 3.141$\pm$0.007 & 22.007$\pm$0.004 & 0.0 & 9 & -    \\
+% 0 & 2 & 3 & 1079.62 & 28.92    &   6.652430 &  17.135000 & 23.953$\pm$0.030  & 2 4.890$\pm$0.023 & 24.240$\pm$0.023 & 0.0 & - & -    \\
+% 0 & 2 & 4 & 114.58  & 21.22    &   6.655560 &  17.148020 & 23.801$\pm$0.025  & 2 5.039$\pm$0.026 & 24.112$\pm$0.021 & 0.0 & - & -    \\
+% 0 & 2 & 5 & 46.78   & 19.46    &   6.655800 &  17.148932 & 23.012$\pm$0.012  & 2 3.924$\pm$0.012 & 23.282$\pm$0.011 & 0.0 & - & -    \\
+% 0 & 2 & 6 & 1441.84 & 16.16    &   6.651480 &  17.130072 & 24.393$\pm$0.045  & 2 6.099$\pm$0.062 & 25.119$\pm$0.049 & 0.0 & - & -    \\
+% 0 & 2 & 7 & 205.43  & 3.96     &   6.655520 &  17.146742 & 24.424$\pm$0.032  & 2 5.028$\pm$0.025 & 24.597$\pm$0.027 & 0.0 & - & -    \\
+% 0 & 2 & 8 & 1321.63 & 9.76     &   6.651950 &  17.131672 &
+% 22.189$\pm$0.011  & 2 4.743$\pm$0.021 & 23.298$\pm$0.011 & 0.0 & 4 &
+% edge \\
+% \hline
+% \multicolumn{13}{l}{%
+% Table 2 is published in its entirety in the electronic
+% edition of the {\it Astrophysical Journal}.}\\[3pt]
+% \multicolumn{13}{l}{%
+% $^a$ Sample footnote for table 2.}\\[3pt]
+% \multicolumn{13}{l}{%
+% $^b$ Another sample footnote for table 2.}
+% \end{tabular}
+% }}
+% \end{table}
+% \clearpage
 
-\vglue 3in
-Example of table continuing over pages:
+% \vglue 3in
+% Example of table continuing over pages:
 
-\begin{center}
-\begin{longtable}{ccc@{}}
-\caption{ApJ costs from 1991 to 2013
-\label{tab:table}} \\[2pt]
-\hline
-\bf Year & \bf Subscription & \bf Publication \\
- & \bf cost &\bf charges\\
- & \bf(\$) & \bf (\$/page)\\
-\hline
-\endfirsthead
+% \begin{center}
+% \begin{longtable}{ccc@{}}
+% \caption{ApJ costs from 1991 to 2013
+% \label{tab:table}} \\[2pt]
+% \hline
+% \bf Year & \bf Subscription & \bf Publication \\
+%  & \bf cost &\bf charges\\
+%  & \bf(\$) & \bf (\$/page)\\
+% \hline
+% \endfirsthead
 
-\multicolumn3c{Table \thetable, \it continued from previous page.}\\[6pt]
-\multicolumn3c{ApJ costs from 1991 to 2013}\\[2pt]
-\hline
-\bf Year & \bf Subscription & \bf Publication \\
- & \bf cost &\bf charges\\
- & \bf(\$) & \bf (\$/page)\\
-\hline
-\endhead
-\\\hline
-\\[-8pt]
-\multicolumn{3}{r}{\it Table continued on next page}\\ 
-\endfoot
+% \multicolumn3c{Table \thetable, \it continued from previous page.}\\[6pt]
+% \multicolumn3c{ApJ costs from 1991 to 2013}\\[2pt]
+% \hline
+% \bf Year & \bf Subscription & \bf Publication \\
+%  & \bf cost &\bf charges\\
+%  & \bf(\$) & \bf (\$/page)\\
+% \hline
+% \endhead
+% \\\hline
+% \\[-8pt]
+% \multicolumn{3}{r}{\it Table continued on next page}\\ 
+% \endfoot
 
-\hline
-\endlastfoot
+% \hline
+% \endlastfoot
 
-1991 & 600 & 100 \\
-1992 & 650 & 105 \\
-1993 & 550 & 103 \\
-1994 & 450 & 110 \\
-1995 & 410 & 112 \\
-1996 & 400 & 114 \\
-1997 & 525 & 115 \\
-1998 & 590 & 116 \\
-1999 & 575 & 115 \\
-2000 & 450 & 103 \\
-2001 & 490 &  90 \\
-2002 & 500 &  88 \\
-2003 & 450 &  90 \\
-2004 & 460 &  88 \\
-2005 & 440 &  79 \\
-2006 & 350 &  77 \\
-2007 & 325 &  70 \\
-2008 & 320 &  65 \\
-2009 & 190 &  68 \\
-2010 & 280 &  70 \\
-2011 & 275 &  68 \\
-2012 & 150 &  56 \\
-2013 & 140 &  55 \\
-\end{longtable}
-\end{center}
+% 1991 & 600 & 100 \\
+% 1992 & 650 & 105 \\
+% 1993 & 550 & 103 \\
+% 1994 & 450 & 110 \\
+% 1995 & 410 & 112 \\
+% 1996 & 400 & 114 \\
+% 1997 & 525 & 115 \\
+% 1998 & 590 & 116 \\
+% 1999 & 575 & 115 \\
+% 2000 & 450 & 103 \\
+% 2001 & 490 &  90 \\
+% 2002 & 500 &  88 \\
+% 2003 & 450 &  90 \\
+% 2004 & 460 &  88 \\
+% 2005 & 440 &  79 \\
+% 2006 & 350 &  77 \\
+% 2007 & 325 &  70 \\
+% 2008 & 320 &  65 \\
+% 2009 & 190 &  68 \\
+% 2010 & 280 &  70 \\
+% 2011 & 275 &  68 \\
+% 2012 & 150 &  56 \\
+% 2013 & 140 &  55 \\
+% \end{longtable}
+% \end{center}
 
-%%%%%%%%%%%%%%%%%%%%%%%
-%% The bibliography
+% %%%%%%%%%%%%%%%%%%%%%%%
+% %% The bibliography
 
-%% The bibliography is made using only
-%% the entries that you cite using \cite{}, or one of the Natbib citation
-%% entries, like \citep{}, \citet{} etc.
+% %% The bibliography is made using only
+% %% the entries that you cite using \cite{}, or one of the Natbib citation
+% %% entries, like \citep{}, \citet{} etc.
 
-\bibliography{bibsamp}
+% \bibliography{bibsamp}
 
-NO APPENDICES allowed in the Network Neuroscience Style.
+% NO APPENDICES allowed in the Network Neuroscience Style.
 
-Please submit Supporting Information documents as PDFs in a format ready to publish.
+% Please submit Supporting Information documents as PDFs in a format ready to publish.

### 2024-02-05 paper/equivalence-of-neural-field-dynamics-with-different-embedding-dimensionality.tex
revamp abstract and introduction

@@
 Classical neural field models describe interactions between neurons positioned in an embedding; often a physical space (for instance, that of the cortical sheet) or, later, abstract spaces (for instance, that of neuron celltypes), in which the dynamics are highly regular in space and time.
-Recent improvements in recordings of neural activity have shined the light on the smoothness of the neural dynamics, when placed in an appropriate embedding space.
+Recent improvements in recordings of neural activity have further demonstrated the smoothness of the neural dynamics, when placed in an appropriate embedding space.
 
@@
 
-\subsection{Neural field models in neuroscience}
+% \subsection{Neural field models in neuroscience}
 
@@
 Informally, these results suggest that, as long as the connectivity scales as $1/N$ (where $N$ is the number of neurons), the dynamics of large networks can be described in the $N \to \infty$ limit by a neural field equation where the embedding is simply the interval $[0,1]$.
-In the following, we will describe a method of explicitly writing such a one-dimensional neural field.
+The main contribution of this paper is the derivation of a method, allowing to explicitly write such a one-dimensional neural field, starting from a known higher-dimensional neural field.
 
-\subsection{Formulating a one-dimensional neural field}
+% \subsection{Formulating a one-dimensional neural field}
 
@@
 The connectivity kernel $w(\vec y, \vec z)$ describes the amplitude of the recurrent currents from the population at $\vec z$ to the population at $\vec y$.
+Importantly, $w(\vec y, \vec z)$ is a ``smooth'' function in both arguments.
+This is a natural consequence of the derivation of neural field equations (rigorous proofs of convergence of networks to neural fields are still an active area of study, and we may cite \citep{CheDua19,AGATHENERINE202286} for recent developments).
 $\rho(\vec z)$ measures the number of neurons per unit volume of the embedding space, and we use the notation $\rho(\mathrm d \vec y) := \rho(\vec y) \mathrm d \vec y$ to emphasize that the integral is weighted by the neuron density.
@@
 
-The embedding space $\Rp$ can be mapped to $[0,1]^p$ via bijective functions between the two spaces.
+The embedding space $\Rp$ (or any $p$-dimensional space) can be mapped to $[0,1]^p$ via bijective functions between the two spaces.
 For instance, if the neuron density factorizes as $\rho(z_1, \cdots, z_p)= \rho_1(z_1) \cdots \rho_p(z_p)$, then the associated cumulative density functions can be used to perform a change of variables, resulting in the an equivalent "uniform" neural field equation:
@@
 The mapping from the two-dimensional equation to the one-dimensional equation is done via well-known measurable bijections between the square $[0,1]^2$ and the interval $[0,1]$.
-Importantly, these bijections are not (and can't be, due to Netto's theorem [TODO: citation?]) diffeomorphisms.
+Importantly, these bijections are not diffeomorphisms.
 % Also, we show that although the connectivity kernel of the one-dimensional equation seems to have a fractal structure, it is sufficiently regular for its solution to be numerically approximated using standard grid-based simulations.
 The main idea is to define a measurable bijective mapping $S : [0,1]^2 \mapsto [0,1]$ and a measure $\lambda : [0,1]^2 \mapsto \mathbb{R}^+$, such that we can write a one-dimensional neural field	$\tilde h(\alpha, t)$ with a connectivity kernel defined as $\tilde w(\alpha, \beta) = w_U(S^{-1}(\alpha), S^{-1}(\beta))$.
-The question of well-posedness of the neural field in \autoref{eq:nf-in-01} is defered in \autoref{sec:proof-equivalence}.
+Doing this, we arrive at the following one-dimensional neural field equation:
+% The question of well-posedness of the neural field in \autoref{eq:nf-in-01} is defered in \autoref{sec:proof-equivalence}.
 
@@
 
-\subsection{Regularity of the mapping}
+The difficulty in writing the one-dimensional neural field, is that the spatial structure of the original neural field in $[0,1]^2$ may be lost.
+The intuition is that we would like populations close in $[0,1]^2$ to be bijectively mapped to populations close in $[0,1]$.
+However, since $S^{-1}$ is a bijection, it cannot, by Netto's theorem, also be continuous.
+The core of our argument lies in the fact that we can relax the requirement that the spatial embedding is continuous, instead, we only require it is ``continuous on average''.
+We quantify this by introducing "locality" $V(S^{-1})$, which measures if small neighbourhoods in $[0,1]^2$ are mapped to small neighbourhoods in $[0,1]$ by the mapping $S^{-1}$.
+Using locality, we then prove that the mean-field approximation is still valid, implying that we can simulate the one-dimensional neural field using classical grid-based methods.
 
-[chap 1.4] introduce a random mapping to introduce locality
+% question of uniqueness, discretization error, mean field limit and locality, link with numerical simualtions. chap 1.4, chap 3.1
 
-figure 1.2
+% "discretization error"
 
-question of uniqueness, discretization error, mean field limit and locality, link with numerical simualtions. chap 1.4, chap 3.1
-
-"discretization error"
-
-argue over smoothness of the kernel (see presentation given for SPC)
-
-chap 1.1, chap 2.1, 2.5
+% chap 1.1, chap 2.1, 2.5
 

### 2024-02-06 paper/equivalence-of-neural-field-dynamics-with-different-embedding-dimensionality.tex
small improvements to intro

@@
 
-Let us write a neural field associated with a $p$-dimensional embedding $\Rp$.
+We begin our work by considering a neural field associated with a $p$-dimensional embedding $\Rp$.
 $h(\vec z, t)$ describes the potential of the neural population at position $\vec z=(z_1,\cdots,z_p)$ in the embedding.
@@
 This is a natural consequence of the derivation of neural field equations (rigorous proofs of convergence of networks to neural fields are still an active area of study, and we may cite \citep{CheDua19,AGATHENERINE202286} for recent developments).
-$\rho(\vec z)$ measures the number of neurons per unit volume of the embedding space, and we use the notation $\rho(\mathrm d \vec y) := \rho(\vec y) \mathrm d \vec y$ to emphasize that the integral is weighted by the neuron density.
-The activation function $\phi(h)$ describes the firing rate of a given neuron (population) with potential $h$.
+$\rho(\vec z)$ measures the number of neurons per unit volume of the embedding space, effectively weighing the recurrent currents.
+The monotonic increasing activation function $\phi(h)$ describes the firing rate of a given neuron (population) with potential $h$.
 The neural field then evolves according to \autoref{eq:nf-in-rp}:
@@
 \begin{equation} \label{eq:nf-in-rp}
-\partial_t h(\vec z, t) = \underbrace{-h(\vec z, t)}_\text{exponential decay} + \underbrace{\int_{\mathbb{R}^p} w(\vec z, \vec y) \phi(h(\vec y, t)) \rho(\mathrm d \vec y)}_\text{recurrent current}.
+\partial_t h(\vec z, t) = \underbrace{-h(\vec z, t)}_\text{exponential decay} + \underbrace{\int_{\mathbb{R}^p} w(\vec z, \vec y) \phi(h(\vec y, t)) \rho(\vec y) \mathrm d \vec y}_\text{recurrent current}.
 \end{equation}
 
-The embedding space $\Rp$ (or any $p$-dimensional space) can be mapped to $[0,1]^p$ via bijective functions between the two spaces.
+Without loss of generality, a simple bijective mapping can be applied to reduce any $p$-dimensional neural field to a neural field on $[0,1]^p$, and the redefining the connectivity kernel to absorb the density and determinant of the transformation, we arrive at \autoref{eq:nf-in-01p}, a general form for any neural field equation in $p$ dimensions:
+
+% The embedding space $\Rp$ (or any $p$-dimensional space) can be mapped to $[0,1]^p$ via bijective functions between the two spaces.
 For instance, if the neuron density factorizes as $\rho(z_1, \cdots, z_p)= \rho_1(z_1) \cdots \rho_p(z_p)$, then the associated cumulative density functions can be used to perform a change of variables, resulting in the an equivalent "uniform" neural field equation:
@@
 
-If the neuron density does not factorize, then one can arrive at the same formulation by redefining the connectivity kernel to include the neuron density and the determinant of the transformation.
+This form highlights the central role of the connectivity kernel $w_U$, which together with the initial condition $h_U(t=0, \vec v)$, uniquely determines the neural dynamics.
+
+% If the neuron density does not factorize, then one can arrive at the same formulation by redefining the connectivity kernel to include the neuron density and the determinant of the transformation.
 [TODO: this sentence might need some work]
@@
 The core of our argument lies in the fact that we can relax the requirement that the spatial embedding is continuous, instead, we only require it is ``continuous on average''.
-We quantify this by introducing "locality" $V(S^{-1})$, which measures if small neighbourhoods in $[0,1]^2$ are mapped to small neighbourhoods in $[0,1]$ by the mapping $S^{-1}$.
+We quantify this by introducing ``locality'' $V(S^{-1})$, which measures if small neighbourhoods in $[0,1]^2$ are mapped to small neighbourhoods in $[0,1]$ by the mapping $S^{-1}$.
 Using locality, we then prove that the mean-field approximation is still valid, implying that we can simulate the one-dimensional neural field using classical grid-based methods.
@@
 
-Our results suggest that the notion of (spatial) dimensionality in neural field dynamics is not well-defined if not associated with a constraint on the regularity of the neural fields (and the connectivity kernel).
+Our results suggest that the notion of spatial dimensionality in neural field dynamics is not well-defined if not associated with a constraint on the regularity of the neural fields (and the connectivity kernel).
 Thereby, by studying a simple toy model, this work illustrates the importance of the analytic notion of regularity when neural population dynamics over abstract continuous spaces are considered.

### 2024-02-07 paper/equivalence-of-neural-field-dynamics-with-different-embedding-dimensionality.tex
up and to S as pointwise limit

@@
 
-Without loss of generality, a simple bijective mapping can be applied to reduce any $p$-dimensional neural field to a neural field on $[0,1]^p$, and the redefining the connectivity kernel to absorb the density and determinant of the transformation, we arrive at \autoref{eq:nf-in-01p}, a general form for any neural field equation in $p$ dimensions:
+Without loss of generality, a simple bijective mapping can be applied to reduce any $p$-dimensional neural field to a neural field on $[0,1]^p$, additionally redefining the connectivity kernel to absorb the density and determinant of the transformation, we arrive at \autoref{eq:nf-in-01p}, a general form for any neural field equation in $p$ dimensions:
 
@@
 \begin{equation} \label{eq:nf-in-01}
-\partial_t \tilde h(\alpha, t) = -\tilde h(\alpha, t) + \int_{[0,1]} \tilde w(\alpha, \beta) \phi(\tilde h(\beta, t)) \left[\lambda \circ S^{-1}\right](\d \beta)
+\partial_t \tilde h(\alpha, t) = -\tilde h(\alpha, t) + \int_{[0,1]} \tilde w(\alpha, \beta) \phi(\tilde h(\beta, t)) \left[\lambda \circ S^{-1}\right](\d \beta).
 \end{equation}
@@
 
-\section{Study of mappings from 2D to 1D}
+\section{Mappings $S$ from $[0,1]^2$ to $[0,1]$}
 
-\subsection{$S$ as the limit of a sequence of mappings}
+\subsection{Motivation}
 
-motivation, why do we do this
+We motivate the upcoming study of the mappings $S$ by considering a grid discretization of the original two-dimensional space $[0,1]^2$.
+For this, we subdivide each axis into $2^n$ equal segments, where $n \in \mathbb{N}$ describes the grid resolution.
+Each segment has a length of $2^{-n}$, resulting in a grid constituted of $2^n \times 2^n = 4^n$ squares in $[0,1]^2$.
+To each square bin indexed $i \in \{1,\cdots,4^n\}$, we associate a position $\vec{v_i} = (v_{i,1}, v_{i,2})$ located within the bounds of the bin.
+Since everytime $n$ is incremented by one, every axis is subdivided into two smaller equal parts, the positions $v_{i,1}$ and $v_{i,2}$ naturally lend themselves to a binary notation.
+We also introduce $v_\mu^{(n)}$ to denote the $n$-bit truncation of the position $v_\mu \in [0,1]$, where $\mu = 1,2$.
+Additionally, we write the error of the trunctation as a small corrective factor $0 \leq c^{(n)} < 2^{-n}$.
+Using this notation and dropping the index $i$ in favor of readability, we write the components of $\vec{v} \in [0,1]^2$ as:
 
-\subsection{The Z-mapping and Column mapping}
+\begin{equation} \begin{aligned} \label{eq:positions-v}
+v_1 &= v_1^{(n)} + c_1^{(n)} = \sum_{l=1}^n b_1^l 2^{-l} + c_1^{(n)} = 0.b_1^1 b_1^2 \cdots b_1^n + c_1^{(n)}, \\
+v_2 &= v_2^{(n)} + c_2^{(n)} = \sum_{l=1}^n b_2^l 2^{-l} + c_2^{(n)} = 0.b_2^1 b_2^2 \cdots b_2^n + c_2^{(n)}.
+\end{aligned}\end{equation}
 
-introduce the mappings and give plots
+It can easily be seen that $v_\mu^{(n)} \to v_\mu$ as $n\to\infty$. In general, the bits of the $v_\mu$ can be expressed as:
 
-\subsection{Behavior and locality measures}
+\begin{equation*}
+b_\mu^l = \mathrm{Ind} \left\{ 2^{l-1}v_\mu - \left\lfloor 2^{l-1}v_\mu \right\rfloor \right\}
+\end{equation*}
 
-measure of locality, give examples applied to the mappings.
+[TODO: add a figure here of the discretization here ? do we need to clarify explicitly the bounds of each square bin ?]
 
-see how it scales, and say what we expect
+Sampling the original connectivity kernel at the points $\vec{v_i}$ and $\vec{v_j}$ yields a numerical connectivity matrix $J_{ij} := w_U(\vec{v_i}, \vec{v_j})$, which can be used to simulate the resulting population dynamics equation:
+
+\begin{equation} \label{eq:population-dynamics}
+\frac{\d}{\d t} h_i(t) = -h_i(t) + \sum_{j=1}^{4^n} J_{ij} \phi(h_j(t)), \quad \text{for all } i \in \{1,\cdots,4^n\}.
+\end{equation}
+
+\autoref{eq:population-dynamics} can be seen as the discretized version of \autoref{eq:nf-in-01p}, where we associate $h_i(t) := h_U(\vec{v_i}, t)$.
+Indeed, by construction of the original neural field, the discrete dynamics converge to the true solution of the neural field, because the error of the mean-field approximation in $[0,1]^2$ vanishes as we take finer grids, letting $n \to \infty$.
+
+We would now like to derive a bijective mapping $S$, such that the same argument applies on the neural field in $[0,1]$.
+Intuively, we want $S$ to conserve the regularity of the neural field on $[0,1]^2$, since if the mean-field argument also carries over to the one-dimensional neural field, it seems likely that it will express identical dynamics.
+
+\subsection{$S$ as the pointwise limit of a sequence of bijections $S^n$}
+
+Let us define a sequence of bijective mappings $S^n$, acting on the $n$-bit truncations of $\vec{v_i}$, as well as the sequence $\alpha_i^{(n)} \in [0,1]$ of positions resulting from the image of $\vec{v_i^{(n)}}$.
+In the following, we again drop the index $i$ for clarity.
+
+\begin{equation} \begin{aligned}
+S^n &: (v_1^{(n)}, v_2^{(n)}) = (0.b_1^1 b_1^2 \cdots b_1^n, 0.b_2^1 b_2^2 \cdots b_2^n) \\
+&\mapsto \alpha^{(n)} = 0.b^1 b^2 \cdots b^{2n}
+\end{aligned} \end{equation}
+
+We note that we write $\alpha^{(n)}$ using $2n$ bits, which is intuitively the lowest number of bits required so that $S^n$ remains a bijection.
+More than $2n$ bits could be allowed while maintaining bijectivity, but we omit this possibility, since they wouldn't strictly be required.
+Effectively, $S^n$ maps between $2\times n$ input bits and $2n$ output bits.
+In this formulation, it becomes clear that in order to be a bijection, every bit of the input must be used once and only once to form the output.
+Setting aside bit inversions, this means that $S^n$ defines a permutation of the $2n$ input bits.
+[TODO: are more details necessary ?
+I think this is clear enough, though in the thesis manuscript I got over this point in more detail]
+
+If it exists, we then we define $\alpha = S(v_1, v_2)$ as the pointwise limit of the sequence $\{S^n(v_1, v_2)\}_{n \in \mathbb N}$ (Note: We may define an extension of $S^n$ acting on an arbitrary $(v_1, v_2) \in [0,1]^2$ as $S^n(v_1, v_2) := S^n(v_1^{(n)}, v_2^{(n)})$.
+This extension is obviously not bijective, but in the $n \to \infty$ limit the finite-$n$ truncation converges, such that the pointwise limit may be a bijection).
+
+We will in the rest of this paper show that this simple formulation in terms of permutation of input bits is flexible enough to express mappings that have good properties such as ``locality'', while being simple enough to help build intuitions.
+
+---
+
+let us think of what happens when we apply the mapping on a discrete grid of squares in [0,1]², and where the squares end up on the segment.
+the squares are bijectively mapped to segments on the unit line, i.e.
+it defines a permutation of the squares onto the segments, or as defining an "ordering" of the square bins in $[0,1]^2$.
+This is done using a mapping S : [0, 1]2 7 → [0, 1],
+where we cover the [0, 1]2 space by following the order assigned to the populations in
+the image [0, 1].
+considering finer and finer grids, we look/see/observe where the squares end up, and may define a sequence of mappings $S^n$ acting on finer and finer grids.
+if the sequence converges pointwise, we then define $S$ as the limit of the sequence of mappings $S^n$.
+
+here is how we mathematically write down / formalism for the mappings $S^n$.
+special case where the bins are uniformly spaced along each axis, and we obtain the next iteration by subdividing each bin of the axis into two equally smaller bins.
+along each axis, every bin can be located by a tuple of coordinates.
+a particularly elegant way to locate the square bins is to use binary notation (base 2), such that at iteration n, every coordinate has n bits of information.
+therefore every square can be uniquely located by a total of 2n bits.
+
+similarly, we subdivide the unit segment into 4 every time, such that the number of segments matches the number of squares.
+
+then
+
+\begin{equation} \begin{aligned}
+S^n &: (v_1^{(n)}, v_2^{(n)}) = (0.b_1^1 b_2^1 \cdots b_n^1, 0.b_1^2 b_2^2 \cdots b_n^2) \\
+&\mapsto \alpha^{(n)} = 0.b_1 b_2 \cdots b_{2n}
+\end{aligned} \end{equation}
+
+\section{Locality of the Z-mapping and the Column mapping}
+
+% \subsection{$S$ as the limit of a sequence of mappings}
+
+% motivation, why do we do this
+
+% \subsection{The Z-mapping and Column mapping}
+
+% introduce the mappings and give plots
+
+% \subsection{Behavior and locality measures}
+
+% measure of locality, give examples applied to the mappings.
+
+% see how it scales, and say what we expect
 

### 2024-02-07 paper/equivalence-of-neural-field-dynamics-with-different-embedding-dimensionality.tex
aegument of pointwise convergence

@@
 This extension is obviously not bijective, but in the $n \to \infty$ limit the finite-$n$ truncation converges, such that the pointwise limit may be a bijection).
+In terms of the current formulation, pointwise convergence requires that the least signficiant bits (LSB) of the input be mapped to the LSB of the output, and identically the most significant bits (MSB) of the input be mapped to the MSB of the output [TODO: do we need a proof of this ?].
+This ensures that small corrections in the input $\vec{v}$ results in small corrections in the output $\alpha$.
 

### 2024-02-20 paper/manuscript.tex
PLOS compliant text

--- 2024-02-07 paper/equivalence-of-neural-field-dynamics-with-different-embedding-dimensionality.tex
@@
-%% NETN Template file
+% Template for PLoS
+% Version 3.6 Aug 2022
+%
+% % % % % % % % % % % % % % % % % % % % % %
+%
+% -- IMPORTANT NOTE
+%
+% This template contains comments intended 
+% to minimize problems and delays during our production 
+% process. Please follow the template instructions
+% whenever possible.
+%
+% % % % % % % % % % % % % % % % % % % % % % % 
+%
+% Once your paper is accepted for publication, 
+% PLEASE REMOVE ALL TRACKED CHANGES in this file 
+% and leave only the final text of your manuscript. 
+% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
+% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
+%
+%
+% There are no restrictions on package use within the LaTeX files except that no packages listed in the template may be deleted.
+%
+% Please do not include colors or graphics in the text.
+%
+% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
+%
+% % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- FIGURES AND TABLES
+%
+% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
+%
+% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
+% - Figures should be uploaded separately from your manuscript file. 
+% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
+% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
+% For figure citations, please use "Fig" instead of "Figure".
+% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
+%
+% Tables should be cell-based and may not contain:
+% - spacing/line breaks within cells to alter layout or alignment
+% - do not nest tabular environments (no tabular environments within tabular environments)
+% - no graphics or colored text (cell background color/shading OK)
+% See http://journals.plos.org/plosone/s/tables for table guidelines.
+%
+% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
+%
+% IMPORTANT
+% Below are a few tips to help format your equations and other special characters according to our specifications.
+For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
+%
+% For inline equations, please be sure to include all portions of an equation in the math environment.
+For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
+%
+% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
+%
+% Please add line breaks to long display equations when possible in order to fit size of the column. 
+%
+% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
+%
+% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
+%
+% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
+%
+% % % % % % % % % % % % % % % % % % % % % % % % 
+%
+% Please contact latex@plos.org with any questions.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
 
-%% Documentclass:
-\documentclass[NETN,manuscript]{stjour-new}
+\documentclass[10pt,letterpaper]{article}
+\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
 
-%% Or,
-%% For production only, need .otf fonts, private MIT logo files,
-%% and need to use lualatex:
-% \documentclass[finalfonts,NETN]{stjour-new}
+% amsmath and amssymb packages, useful for mathematical formulas and symbols
+\usepackage{amsmath,amssymb}
 
-%%%%%%%%%%% Article Set-Up %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Article Type:
+% Use adjustwidth environment to exceed column width (see example table in text)
+\usepackage{changepage}
 
-%% Choose one of these options, (Default is Research).
-%% Research
-%% Methods 
-%% Data
-%% Review
-%% Perspectives
+% textcomp package and marvosym package for additional characters
+\usepackage{textcomp,marvosym}
 
-\articletype{Research}
+% cite package, to clean up citations in the main text. Do not remove.
+\usepackage{cite}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% author definitions should be placed here:
+% Use nameref to cite supporting information files (see Supporting Information section for more info)
+\usepackage{nameref,hyperref}
+
+% line numbers
+\usepackage[right]{lineno}
+
+% ligatures disabled
+\usepackage[nopatch=eqnum]{microtype}
+\DisableLigatures[f]{encoding = *, family = * }
+
+% color can be used to apply background shading to table cells only
+\usepackage[table]{xcolor}
+
+% array package and thick rules for tables
+\usepackage{array}
+
+% create "+" rule type for thick vertical lines
+\newcolumntype{+}{!{\vrule width 2pt}}
+
+% create \thickcline for thick horizontal lines of variable length
+\newlength\savedwidth
+\newcommand\thickcline[1]{%
+  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+  \cline{#1}%
+  \noalign{\vskip\arrayrulewidth}%
+  \noalign{\global\arrayrulewidth\savedwidth}%
+}
+
+% \thickhline command for thick horizontal lines that span the table
+\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+\hline
+\noalign{\global\arrayrulewidth\savedwidth}}
+
+% Remove comment for double spacing
+%\usepackage{setspace} 
+%\doublespacing
+
+% Text layout
+\raggedright
+\setlength{\parindent}{0.5cm}
+\textwidth 5.25in 
+\textheight 8.75in
+
+% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
+% Captions will be left justified
+\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
+\renewcommand{\figurename}{Fig}
+
+% Use the PLoS provided BiBTeX style
+\bibliographystyle{plos2015}
+
+% Remove brackets from numbering in List of References
+\makeatletter
+\renewcommand{\@biblabel}[1]{\quad#1.}
+\makeatother
+
+% Header and Footer with logo
+\usepackage{lastpage,fancyhdr,graphicx}
+\usepackage{epstopdf}
+%\pagestyle{myheadings}
+\pagestyle{fancy}
+\fancyhf{}
+%\setlength{\headheight}{27.023pt}
+%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
+\rfoot{\thepage/\pageref{LastPage}}
+\renewcommand{\headrulewidth}{0pt}
+\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
+\fancyheadoffset[L]{2.25in}
+\fancyfootoffset[L]{2.25in}
+\lfoot{\today}
+
+%% Include all macros below
+
+\newcommand{\lorem}{{\bf LOREM}}
+\newcommand{\ipsum}{{\bf IPSUM}}
+
+%% User macros
 
@@
 
+%% END MACROS SECTION
+
 \begin{document}
+\vspace*{0.2in}
 
+% Title must be 250 characters or less.
+\begin{flushleft}
 % TODO: what title is better ?
 % \title{Equivalence of neural field dynamics with different embedding dimensionality}
-\title{Neural fields with different embedding dimensionality can express identical dynamics}
+% \title{Neural fields with different embedding dimensionality can express identical dynamics}
 % \title{Neural dynamics can be equivalently embedded in neural fields of different dimensions}
+{\Large
+\textbf\newline{Neural fields with different embedding dimensionality can express identical dynamics} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+}
+\newline
+% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
+\\
+Nicole Vadot\textsuperscript{1},
+Valentin Schmutz\textsuperscript{1},
+Wulfram Gerstner\textsubscript{1}
+% Name1 Surname\textsuperscript{1,2\Yinyang},
+% Name2 Surname\textsuperscript{2\Yinyang},
+% Name3 Surname\textsuperscript{2,3\textcurrency},
+% Name4 Surname\textsuperscript{2},
+% Name5 Surname\textsuperscript{2\ddag},
+% Name6 Surname\textsuperscript{2\ddag},
+% Name7 Surname\textsuperscript{1,2,3*},
+% with the Lorem Ipsum Consortium\textsuperscript{\textpilcrow}
+\\
+\bigskip
+\textbf{1} LCN, EPFL, Lausanne, Vaud, Switzerland
+% \\
+% \textbf{2} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+% \\
+% \textbf{3} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+% \\
+\bigskip
 
-% \subtitle{<Subtitle Here>} %% Optional subtitle
+% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
+% 
+% Remove or comment out the author notes below if they aren't used.
+%
+% Primary Equal Contribution Note
+% \Yinyang These authors contributed equally to this work.
 
-%% If shortened title for running head is needed so that the article title can fit
-%%   in the running head, use [] argument, ie,
-%%
-%%   \title[Shortened Title for Running Head]{Title of Article}
-%%   \subtitle{Subtitle Here}
+% Additional Equal Contribution Note
+% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
+% \ddag These authors also contributed equally to this work.
 
-%%% Author/Affil
-%% Since we use \affil{} in the argument of the \author command, we
-%% need to supply another version of the author names, without \affil{}
-%% to be used for running heads:
+% Current address notes
+% \textcurrency Current Address: Dept/Program/Center, Institution Name, City, State, Country % change symbol to "\textcurrency a" if more than one current address note
+% \textcurrency b Insert second current address 
+% \textcurrency c Insert third current address
 
-\author{Nicole Vadot\affil{1}, Valentin Schmutz\affil{1}, \and Wulfram Gerstner \affil{1}}
+% Deceased author note
+% \dag Deceased
 
-\affiliation{1}{LCN, EPFL, Lausanne, Switzerland}
+% Group/Consortium Author Note
+% \textpilcrow Membership list can be found in the Acknowledgments section.
 
-%ie.
-%\affiliation{1}{Gatsby Computational Neuroscience Unit, University
-%College London, London, United Kingdom}
+% Use the asterisk to denote corresponding authorship and provide email address in note below.
+% * correspondingauthor@institute.edu
 
-% \affiliation{2}{Another Department, Institution, City, Country}
-
-%ie
-%\affiliation{2}{Center for Studies in
-%Behavioral Neurobiology, Concordia University, Montreal, Quebec,
-%Canada}
-
-\correspondingauthor{Nicole Vadot}{nicole.vadot@epfl.ch}
-\correspondingauthor{Valentin Schmutz}{valentin.schmutz@epfl.ch}
-\correspondingauthor{Wulfram Gerstner}{wulfram.gerstner@epfl.ch}
-
-% ie,
-%\correspondingauthor{Ritwik K. Niyogi}{ritwik.niyogi@gatsby.ucl.ac.uk}
-
-\keywords{Neural fields, Embedding, Dimensionality}
-
-%ie
-%\keywords{Work, Leisure, Normative, Microscopic,  Reinforcement Learning, Economics}
-
-\begin{abstract}
+\end{flushleft}
+% Please keep the abstract below 300 words
+\section*{Abstract}
 Classical neural field models describe interactions between neurons positioned in an embedding; often a physical space (for instance, that of the cortical sheet) or, later, abstract spaces (for instance, that of neuron celltypes), in which the dynamics are highly regular in space and time.
@@
 To illustrate these results through numerical models, we describe a coarse-graining procedure enforcing the locality of the numerically applied mappings.
-\end{abstract}
 
-% \begin{authorsummary}
-% Author Summary here.
-% \end{authorsummary}
+% Please keep the Author Summary between 150 and 200 words
+% Use first person. PLOS ONE authors please skip this step. 
+% Author Summary not valid for PLOS ONE submissions.   
+\section*{Author summary}
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Curabitur eget porta erat.
+Morbi consectetur est vel gravida pretium.
+Suspendisse ut dui eu ante cursus gravida non sed sem.
+Nullam sapien tellus, commodo id velit id, eleifend volutpat quam.
+Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
+Nunc pellentesque pretium diam, quis maximus dolor faucibus id.
+Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
+Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
+
+\linenumbers
 
@@
 
-Early recordings of cats' somatosensory \citep{Mou57} and visual \citep{HubWie62} cortices suggested that the cortical sheet is organized in vertical columns of functionally similar neurons.
+Early recordings of cats' somatosensory \cite{Mou57} and visual \cite{HubWie62} cortices suggested that the cortical sheet is organized in vertical columns of functionally similar neurons.
 
-The columnar organization of the cortex motivated the design of spatially structured models of neural population dynamics, namely neural field models \citep{WilCow73,Nun74,Ama77}, where the spatial dimensions corresponded to the two dimensions of the cortical sheet.
+The columnar organization of the cortex motivated the design of spatially structured models of neural population dynamics, namely neural field models \cite{WilCow73,Nun74,Ama77}, where the spatial dimensions corresponded to the two dimensions of the cortical sheet.
 
-As models of spatiotemporal neural population dynamics on the cortical sheet, neural field models can be used to explain experimentally observed patterns of cortical activity, such as travelling waves in visual \citep{SatNau12,MulRey14}, somatosensory \citep{PetHah03,FerBol06}, motor \citep{RubRob06,TakKim15}, and hippocampal \citep{LubSia09,PatFuj12,PatSch13} cortices (see \citep{MulCha18} for a review), and are used to model large-scale brain signals, such as electroencephalography (EEG) recordings \citep{Bre17} (see also \citep{DipRan18} for an example of neural field modelling of calcium imaging recording of the visual cortex).
+As models of spatiotemporal neural population dynamics on the cortical sheet, neural field models can be used to explain experimentally observed patterns of cortical activity, such as travelling waves in visual \cite{SatNau12,MulRey14}, somatosensory \cite{PetHah03,FerBol06}, motor \cite{RubRob06,TakKim15}, and hippocampal \cite{LubSia09,PatFuj12,PatSch13} cortices (see \cite{MulCha18} for a review), and are used to model large-scale brain signals, such as electroencephalography (EEG) recordings \cite{Bre17} (see also \cite{DipRan18} for an example of neural field modelling of calcium imaging recording of the visual cortex).
 
 More recent recording methods, especially in rodents, have provided evidence of low-dimensional organization of neuronal activity that does not depend on the physical locations of neurons on the cortical sheet.
-For example, contrary to cats' visual cortex, the orientation selectivity of pyramidal neurons in a rat's visual cortex does not depend on their locations, that is, orientation selectivity is heterogeneous at any given location \citep{OhkChu05}.
-Moreover, the activity of interneurons in any small volume of the visual cortex is also heterogeneous and seems to be structured by a low-dimensional manifold of fine cell subtypes \citep{BugDuf22}.
+For example, contrary to cats' visual cortex, the orientation selectivity of pyramidal neurons in a rat's visual cortex does not depend on their locations, that is, orientation selectivity is heterogeneous at any given location \cite{OhkChu05}.
+Moreover, the activity of interneurons in any small volume of the visual cortex is also heterogeneous and seems to be structured by a low-dimensional manifold of fine cell subtypes \cite{BugDuf22}.
 The local functional heterogeneity of neuronal activity in the cortex challenges the old concept of functionally homogeneous cortical columns and, with it, classical neural field models.
 
-While, historically, the spatial dimensions in neural field models have been the two dimensions of the cortical sheet (with the notable exception of the ring models for orientation selectivity in the visual cortex \citep{BenBar95} and the head-direction system \citep{Zha96}), the "space" in neural field models does not need to represent physical space but can represent any suitable abstract embedding space.
+While, historically, the spatial dimensions in neural field models have been the two dimensions of the cortical sheet (with the notable exception of the ring models for orientation selectivity in the visual cortex \cite{BenBar95} and the head-direction system \cite{Zha96}), the "space" in neural field models does not need to represent physical space but can represent any suitable abstract embedding space.
 Neural field models with abstract embedding spaces could constitute a generalization of classical neural field theory for cortical networks that are not solely structured by distances of neurons on the cortical sheet.
 
-Moreover, this work is motivated by recent mathematical developments \citep{JabPoy21} in mean-field theories involving the theory of graphons.
+Moreover, this work is motivated by recent mathematical developments \cite{JabPoy21} in mean-field theories involving the theory of graphons.
 Informally, these results suggest that, as long as the connectivity scales as $1/N$ (where $N$ is the number of neurons), the dynamics of large networks can be described in the $N \to \infty$ limit by a neural field equation where the embedding is simply the interval $[0,1]$.
@@
 Importantly, $w(\vec y, \vec z)$ is a ``smooth'' function in both arguments.
-This is a natural consequence of the derivation of neural field equations (rigorous proofs of convergence of networks to neural fields are still an active area of study, and we may cite \citep{CheDua19,AGATHENERINE202286} for recent developments).
+This is a natural consequence of the derivation of neural field equations (rigorous proofs of convergence of networks to neural fields are still an active area of study, and we may cite \cite{CheDua19,AGATHENERINE202286} for recent developments).
 $\rho(\vec z)$ measures the number of neurons per unit volume of the embedding space, effectively weighing the recurrent currents.
@@
 
-Let us define a sequence of bijective mappings $S^n$, acting on the $n$-bit truncations of $\vec{v_i}$, as well as the sequence $\alpha_i^{(n)} \in [0,1]$ of positions resulting from the image of $\vec{v_i^{(n)}}$.
+Let us define a sequence of bijective mappings $S^n$, acting on the $n$-bit truncations of $\vec{v_i}$, as well as the sequence $\alpha_i^n \in [0,1]$ of positions resulting from the image of $\vec{v_i^{(n)}}$.
 In the following, we again drop the index $i$ for clarity.
@@
 S^n &: (v_1^{(n)}, v_2^{(n)}) = (0.b_1^1 b_1^2 \cdots b_1^n, 0.b_2^1 b_2^2 \cdots b_2^n) \\
-&\mapsto \alpha^{(n)} = 0.b^1 b^2 \cdots b^{2n}
+&\mapsto \alpha^n = 0.b^1 b^2 \cdots b^{2n}
 \end{aligned} \end{equation}
 
-We note that we write $\alpha^{(n)}$ using $2n$ bits, which is intuitively the lowest number of bits required so that $S^n$ remains a bijection.
+We note that we write $\alpha^n$ using $2n$ bits, which is intuitively the lowest number of bits required so that $S^n$ remains a bijection.
 More than $2n$ bits could be allowed while maintaining bijectivity, but we omit this possibility, since they wouldn't strictly be required.
@@
 
-\acknowledgments
+% Use "Eq" instead of "Equation" for equation citations.
+% \section*{Introduction}
+% Lorem ipsum dolor sit~\cite{bib1} amet, consectetur adipiscing elit.
+Curabitur eget porta erat.
+Morbi consectetur est vel gravida pretium.
+Suspendisse ut dui eu ante cursus gravida non sed sem.
+Nullam Eq~(\ref{eq:schemeP}) sapien tellus, commodo id velit id, eleifend volutpat quam.
+Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
+Nunc pellentesque pretium diam, quis maximus dolor faucibus id.~\cite{bib2} Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
+Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% End of Article
+% \begin{eqnarray}
+% \label{eq:schemeP}
+% 	\mathrm{P_Y} = \underbrace{H(Y_n) - H(Y_n|\mathbf{V}^{Y}_{n})}_{S_Y} + \underbrace{H(Y_n|\mathbf{V}^{Y}_{n})- H(Y_n|\mathbf{V}^{X,Y}_{n})}_{T_{X\rightarrow Y}},
+% \end{eqnarray}
 
-% \acknowledgments
-% \section{Supporting Information} (optional)
-% \section{Competing Interests} (optional)
-% \bibliography{<name of .bib file>}
+% \section*{Materials and methods}
+% \subsection*{Etiam eget sapien nibh}
 
-% \acknowledgments
-% The Funder and award ID information you input at submission will be introduced by the publisher under a Funding Information head during production. 
-% Please use this space for any additional acknowledgements and verbiage required by your funders.
+% % For figure citations, please use "Fig" instead of "Figure".
+% Nulla mi mi, Fig~\ref{fig1} venenatis sed ipsum varius, volutpat euismod diam.
+Proin rutrum vel massa non gravida.
+Quisque tempor sem et dignissim rutrum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Morbi at justo vitae nulla elementum commodo eu id massa.
+In vitae diam ac augue semper tincidunt eu ut eros.
+Fusce fringilla erat porttitor lectus cursus, \nameref{S1_Video} vel sagittis arcu lobortis.
+Aliquam in enim semper, aliquam massa id, cursus neque.
+Praesent faucibus semper libero.
 
-% \section{Supporting Information}
-% This is an optional section. Please use this space to provide information about any supporting information referred to in your manuscript.
+% % Place figure captions after the first paragraph in which they are cited.
+% \begin{figure}[!h]
+% \caption{{\bf Bold the figure title.}
+% Figure caption text here, please use this space for the figure panel descriptions instead of using subfigure commands. A: Lorem ipsum dolor sit amet. B: Consectetur adipiscing elit.}
+% \label{fig1}
+% \end{figure}
 
-% \section{Competing Interests}
-% This is an optional section. If you declared a conflict of interest when you submitted your manuscript, please  use this space to provide details about this conflict.
+% % Results and Discussion can be combined.
+% \section*{Results}
+% Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat euismod diam.
+Proin rutrum vel massa non gravida.
+Quisque tempor sem et dignissim rutrum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Morbi at justo vitae nulla elementum commodo eu id massa.
+In vitae diam ac augue semper tincidunt eu ut eros.
+Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis.
+Aliquam in enim semper, aliquam massa id, cursus neque.
+Praesent faucibus semper libero.
 
-% \section{Technical Terms}
+% % Place tables after the first paragraph in which they are cited.
+% \begin{table}[!ht]
+% \begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
+% \centering
+% \caption{
+% {\bf Table caption Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.}}
+% \begin{tabular}{|l+l|l|l|l|l|l|l|}
+% \hline
+% \multicolumn{4}{|l|}{\bf Heading1} & \multicolumn{4}{|l|}{\bf Heading2}\\ \thickhline
+% $cell1 row1$ & cell2 row 1 & cell3 row 1 & cell4 row 1 & cell5 row 1 & cell6 row 1 & cell7 row 1 & cell8 row 1\\ \hline
+% $cell1 row2$ & cell2 row 2 & cell3 row 2 & cell4 row 2 & cell5 row 2 & cell6 row 2 & cell7 row 2 & cell8 row 2\\ \hline
+% $cell1 row3$ & cell2 row 3 & cell3 row 3 & cell4 row 3 & cell5 row 3 & cell6 row 3 & cell7 row 3 & cell8 row 3\\ \hline
+% \end{tabular}
+% \begin{flushleft} Table notes Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl.
+Ut ornare mauris tellus, vel dapibus arcu suscipit sed.
+% \end{flushleft}
+% \label{table1}
+% \end{adjustwidth}
+% \end{table}
 
-% All NETN article types require Technical Terms.
+% %PLOS does not support heading levels beyond the 3rd (no 4th level headings).
+% \subsection*{\lorem\ and \ipsum\ nunc blandit a tortor}
+% \subsubsection*{3rd level heading} 
+% Maecenas convallis mauris sit amet sem ultrices gravida.
+Etiam eget sapien nibh.
+Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
+Curabitur fringilla pulvinar lectus consectetur pellentesque.
+Quisque augue sem, tincidunt sit amet feugiat eget, ullamcorper sed velit.
+Sed non aliquet felis.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Mauris commodo justo ac dui pretium imperdiet.
+Sed suscipit iaculis mi at feugiat.
 
-% Identify approximately 10 key terms that are mentioned in your article and whose usage and definition may not be familiar across the broad readership of the journal. 
-% Provide brief (20-word or less) definitions for each term, avoiding in these definitions the use of jargon, or highly technical or specialized language. 
-% When the article is typeset, the Technical Terms will appear in the margins at or near their first mention in the text.
+% \begin{enumerate}
+% 	\item{react}
+% 	\item{diffuse free particles}
+% 	\item{increment time by dt and go to 1}
+% \end{enumerate}
 
-% In your manuscript, bold the first occurrence of each \textbf{Technical Term} and then provide a list of the terms and their definitions at the end of the manuscript after the references.
+% \subsection*{Sed ac quam id nisi malesuada congue}
 
-%%%%%%%%%%%%%%%%%%%%%%%
-%% The bibliography
+% Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.
+Proin rutrum vel massa non gravida.
+Quisque tempor sem et dignissim rutrum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Morbi at justo vitae nulla elementum commodo eu id massa.
+In vitae diam ac augue semper tincidunt eu ut eros.
+Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis.
+Aliquam in enim semper, aliquam massa id, cursus neque.
+Praesent faucibus semper libero.
 
-%% The bibliography is made using only the entries that you cite using 
-%% \cite{}, or one of the Natbib citation entries, like \citep{}, \citet{} etc.
+% \begin{itemize}
+% 	\item First bulleted item.
+% 	\item Second bulleted item.
+% 	\item Third bulleted item.
+% \end{itemize}
 
-\bibliography{references}
+% \section*{Discussion}
+% Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat euismod diam.
+Proin rutrum vel massa non gravida.
+Quisque tempor sem et dignissim rutrum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Morbi at justo vitae nulla elementum commodo eu id massa.
+In vitae diam ac augue semper tincidunt eu ut eros.
+Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis.
+Aliquam in enim semper, aliquam massa id, cursus neque.
+Praesent faucibus semper libero~\cite{bib3}.
 
-%ie.,
-%\bibliography{bibsamp}
+% \section*{Conclusion}
 
-% \section{Technical Terms}
+% CO\textsubscript{2} Maecenas convallis mauris sit amet sem ultrices gravida.
+Etiam eget sapien nibh.
+Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
+Curabitur fringilla pulvinar lectus consectetur pellentesque.
+Quisque augue sem, tincidunt sit amet feugiat eget, ullamcorper sed velit.
 
-% \textbf{Technical Term} a key term that is mentioned in an NETN article and whose usage and definition may not be familiar across the broad readership of the journal.
+% Sed non aliquet felis.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Mauris commodo justo ac dui pretium imperdiet.
+Sed suscipit iaculis mi at feugiat.
+Ut neque ipsum, luctus id lacus ut, laoreet scelerisque urna.
+Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl.
+Ut ornare mauris tellus, vel dapibus arcu suscipit sed.
+Nam condimentum sem eget mollis euismod.
+Nullam dui urna, gravida venenatis dui et, tincidunt sodales ex.
+Nunc est dui, sodales sed mauris nec, auctor sagittis leo.
+Aliquam tincidunt, ex in facilisis elementum, libero lectus luctus est, non vulputate nisl augue at dolor.
+For more information, see \nameref{S1_Appendix}.
 
-% \textbf{Technical Term} a key term that is mentioned in an NETN article and whose usage and definition may not be familiar across the broad readership of the journal.
+% \section*{Supporting information}
 
-% \textbf{Technical Term} a key term that is mentioned in an NETN article and whose usage and definition may not be familiar across the broad readership of the journal.
+% % Include only the SI item label in the paragraph heading. Use the \nameref{label} command to cite SI items in the text.
+% \paragraph*{S1 Fig.}
+% \label{S1_Fig}
+% {\bf Bold the title sentence.} Add descriptive text after the title of the item (optional).
 
-% \textbf{Technical Term} a key term that is mentioned in an NETN article and whose usage and definition may not be familiar across the broad readership of the journal.
+% \paragraph*{S2 Fig.}
+% \label{S2_Fig}
+% {\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida.
+Etiam eget sapien nibh.
+Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
+Curabitur fringilla pulvinar lectus consectetur pellentesque.
+
+% \paragraph*{S1 File.}
+% \label{S1_File}
+% {\bf Lorem ipsum.}  Maecenas convallis mauris sit amet sem ultrices gravida.
+Etiam eget sapien nibh.
+Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
+Curabitur fringilla pulvinar lectus consectetur pellentesque.
+
+% \paragraph*{S1 Video.}
+% \label{S1_Video}
+% {\bf Lorem ipsum.}  Maecenas convallis mauris sit amet sem ultrices gravida.
+Etiam eget sapien nibh.
+Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
+Curabitur fringilla pulvinar lectus consectetur pellentesque.
+
+% \paragraph*{S1 Appendix.}
+% \label{S1_Appendix}
+% {\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida.
+Etiam eget sapien nibh.
+Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
+Curabitur fringilla pulvinar lectus consectetur pellentesque.
+
+% \paragraph*{S1 Table.}
+% \label{S1_Table}
+% {\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida.
+Etiam eget sapien nibh.
+Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
+Curabitur fringilla pulvinar lectus consectetur pellentesque.
+
+% \section*{Acknowledgments}
+% Cras egestas velit mauris, eu mollis turpis pellentesque sit amet.
+Interdum et malesuada fames ac ante ipsum primis in faucibus.
+Nam id pretium nisi.
+Sed ac quam id nisi malesuada congue.
+Sed interdum aliquet augue, at pellentesque quam rhoncus vitae.
+
+\nolinenumbers
+
+% Either type in your references using
+% \begin{thebibliography}{}
+% \bibitem{}
+% Text
+% \end{thebibliography}
+%
+% or
+%
+% Compile your BiBTeX database using our plos2015.bst
+% style file and paste the contents of your .bbl file
+% here. See http://journals.plos.org/plosone/s/latex for 
+% step-by-step instructions.
+%
+
+% bibliography automatically substituted by gpp preprocessor,
+% run `make` (see also: Makefile)
+\ifdef{SUBMISSION}\include{"manuscript-tmp.bbl"}\else\bibliography{references}\endif
 
 \end{document}
-
-% The examples below may be helpful when you are looking
-% for a quick example. They are taken from NETNsample.tex
-% so you can see their results in NETNSample.pdf
-
-% \begin{boxedtext}{box title}
-% Text
-% \end{boxedtext}
-
-% or, put figure in boxed text:
-% \begin{boxedtext}{box title}
-% Text
-% \begin{figure}
-% illustration
-% \caption{caption here}
-% \end{figure}
-% \end{boxedtext}
-
-% or, put table in boxed text:
-% \begin{boxedtext}{box title}
-% Text
-% \begin{table}
-% \caption{caption here}
-% tabular...
-% \end{table}
-% \end{boxedtext}
-
-% or, put anything in whitebox within boxed text:
-% \begin{boxedtext}{Showing the use of whitebox} 
-% Going beyond the examination of shared topological features across..
-
-% \whitebox{
-% From $\mathcal{W}$, we can estimate the variability in the fluctuations of the functional connection between nodes $i$ and $j$ over time as:
-% \begin{equation}
-% s_{ij}=\sqrt{\frac{1}{T-L}\sum_{t=1}^{T-L+1}(W_{ij}(t) - m_{ij})}
-% \end{equation}
-% where $m_{ij}=\frac{1}{T-L+1}\sum_{t=1}^{T-L+1}W_{ij}(t)$ is the mean
-% dynamic functional connectivity over time. 
-% }
-% ...differences across a range of network classes.
-% \end{boxedtext}
-
-% ==================
-% More samples:
-
-% \section{Sample Section}
-% Text here. Text here. Text here. Text here.
-% Text here. Text here. Text here. Text here.
-% Text here. Text here. Text here. Text here.
-% Text here. Text here. Text here. Text here.
-
-% \subsection{Sample Subsection}
-% Text here. Text here. Text here. Text here.
-% Text here. Text here. Text here. Text here.
-% Text here. Text here. Text here. Text here.
-% Text here. Text here. Text here. Text here.
-
-% \subsubsection{Sample Subsubsection}
-% Text here. Text here. Text here. Text here.
-% Text here. Text here. Text here. Text here.
-% Text here. Text here. Text here. Text here.
-% Text here. Text here. Text here. Text here.
-
-% \section{Sample equations}
-% \begin{equation}
-% \label{eq:rhoCHT}
-% \rho^{\pi}= \frac{RI + \mathbb{E}_{\pi([L,\tau_L]|\textrm{post})}
-% \left[C_L(\taupav+\tau_L) \right]   +
-% \displaystyle{\int_{0}^{P}}{dw~ \mathbb{E}_{\pi_{w_L}}}
-% \Biggl[\/\sum_{n_{L|[\textrm{pre},w]}}C_L(\tau_L)
-% \Biggr]            }      {P +
-% \mathbb{E}_{\pi([L,\tau_L] |\textrm{post})}[\tau_{L}] +\taupav +
-% \displaystyle{ \int_{0}^{P}}{dw~ \mathbb{E}_{\pi_{w_L}}}   
-% \Biggl[\sum_{n_{L|[\textrm{pre},w]}}\tau_L\Biggr]  
-% }
-% \end{equation}
-% As long as
-% $RI - K_LP > 
-% \frac{1}{\beta}$
-% \begin{equation}
-% %\def\theequation{5.1}
-% \left.\begin{array}{lrcl}
-% &\rho^{\pi} &=&  \displaystyle\frac{\beta ( RI + K_L \taupav )-1} {\beta
-% (P+\taupav )}    \\[12pt]
-% \hbox{and}\hbox to .25in{\hfill}&\mathbb{E}[\tau_L | \text{post}] &=&\displaystyle \frac{P+\taupav}{\beta ( RI -
-% K_LP)-1}  
-% \label{eq:analytical_linear}
-% \end{array}\right\}\hbox to 1.25in{\hfill}
-% \end{equation}
-
-% \subsection{Simple code sample}
-
-% \begin{code}
-% \begin{verbatim}
-% procedure bubbleSort( A : list of sortable items )
-%     n = length(A)
-%     repeat
-%        newn = 0
-%        for i = 1 to n-1 inclusive do
-%           if A[i-1] > A[i] then
-%              swap(A[i-1], A[i])
-%              newn = i
-%           end if
-%        end for
-%        n = newn
-%     until n = 0
-% end procedure
-% \end{verbatim}
-% \end{code}
-
-% \subsection{Algorithm environment}
-
-% %% \begin{algorithm} takes option [p][b][t][h],  or some combination, like \begin{figure}
-% %% See documentation for algorithmic.sty for more information on formatting algorithms.
-
-% \begin{algorithm}[h]
-% \caption{A sample in an algorithm environment.}
-% \begin{algorithmic}
-% \If {$i\geq maxval$}
-%     \State $i\gets 0$
-% \Else
-%     \If {$i+k\leq maxval$}
-%         \State $i\gets i+k$
-%     \EndIf
-% \EndIf
-% \end{algorithmic}
-% \end{algorithm}
-
-% \section{Technical Terms}
-% If you've been asked by the journal editor to provide technical terms and definitions, please bold the first occurrence of each term in the text and provide a list of the terms and their definitions at the end of the manuscript after the references.
-Please only supply technical terms if you've been instructed to do so by a journal editor.
-
-% \section{Itemized Lists}
-
-% \subsection{Roman list:}
-
-% \begin{enumerate}
-% \item[(i)] at high 
-% payoffs, subjects work almost continuously, engaging in little leisure
-% in between work bouts; 
-% \item[(ii)] at low payoffs, they 
-% engage in leisure all at once, in long bouts after working, rather
-% than distributing the same amount of leisure time into multiple short
-% leisure bouts; 
-% \item[(iii)] subjects work continuously for the entire price duration, as long as
-% the price is not very long (as shown by an analysis conducted by Y-AB, to be published separately); %(\textbf{Figure \ref{fig:task_data}D}).
-% \item[(iv)] the duration of leisure bouts is variable.
-% \end{enumerate}
-
-% \subsection{Numbered list:}
-
-% \begin{enumerate}
-% \item at high 
-% payoffs, subjects work almost continuously, engaging in little leisure
-% inbetween work bouts; 
-% \item at low payoffs, they 
-% engage in leisure all at once, in long bouts after working, rather
-% than distributing the same amount of leisure time into multiple short
-% leisure bouts; 
-% \item subjects work continuously for the entire price duration, as long as
-% the price is not very long (as shown by an analysis conducted by Y-AB, to be published separately); %(\textbf{Figure \ref{fig:task_data}D}).
-% \item the duration of leisure bouts is variable.
-% \end{enumerate}
-
-% \subsection{Bulleted list:}
-
-% \begin{itemize}
-% \item at high 
-% payoffs, subjects work almost continuously, engaging in little leisure
-% inbetween work bouts; 
-% \item at low payoffs, they 
-% engage in leisure all at once, in long bouts after working, rather
-% than distributing the same amount of leisure time into multiple short
-% leisure bouts; 
-% \item subjects work continuously for the entire price duration, as long as
-% the price is not very long (as shown by an analysis conducted by Y-AB, to be published separately); %(\textbf{Figure \ref{fig:task_data}D}).
-% \item the duration of leisure bouts is variable.
-% \end{itemize}
-
-% \section{Natbib citation mark up}
-
-% \subsection{Single citations}
-% \noindent
-% \begin{tabular}{ll}
-% \bf Type&\bf Results\\
-% \hline
-% \verb+\citet{jon90}+&Jones et al. (1990)\\
-% \verb+\citet[chap. 2]{jon90}+&Jones et al. (1990, chap. 2)\\
-%     \verb+\citep{jon90}+	    &   	(Jones et al., 1990)\\
-%     \verb+\citep[chap. 2]{jon90}+ 	&    	(Jones et al., 1990, chap. 2)\\
-%     \verb+\citep[see][]{jon90}+ 	 &    	(see Jones et al., 1990)\\
-%     \verb+\citep[see][chap. 2]{jon90}+ 	&    	(see Jones et al., 1990, chap. 2)\\
-%     \verb+\citet*{jon90}+ 	    &    	Jones, Baker, and Williams (1990)\\
-%     \verb+\citep*{jon90}+	    &    	(Jones, Baker, and Williams,
-%     1990) \\
-% \end{tabular}
-
-% For example, here are some sample citations:
-% citet:\citet{anderson}, citep:\citep{antibayes}, and
-% cite*: \citet*{anderson}.
-
-% \subsection{Multiple citations}
-% Multiple citations may be made by including more than one citation
-% key in the \verb+\cite+ command argument.
-
-% \noindent
-% \begin{tabular}{ll}
-% \bf Type&\bf Results\\
-% \hline
-% \verb+\citet{jon90,jam91}+&Jones et al. (1990); James et al. (1991)\\
-% \verb+\citep{jon90,jam91}+&(Jones et al., 1990; James et al. 1991)\\
-% \verb+\citep{jon90,jon91}+&(Jones et al., 1990, 1991)\\
-% \verb+\citep{jon90a,jon90b}+&(Jones et al., 1990a,b)\\
-% \end{tabular}
-
-% For example, multiple citations from the CompPsychSample bibliography:
-% citet:\citet{anderson,antibayes}, citep:\citep{anderson,antibayes}.
-% As you see, the citations are automatically hyperlinked to their
-% reference in the bibliography.
-
-% \newpage
-
-% \section{Sample figures}
-
-% \begin{figure}[h] 
-% \centerline{\includegraphics[width=\textwidth]{Fig1}}
-% \caption{(Colour online) \textbf{Task and key features of the
-%  data.} \\
-%  A) Cumulative handling time (CHT) task. Grey bars denote work
-% (depressing a lever), white gaps show leisure. The subject must
-%  accumulate work up to a total period of time called the
-% \emph{price} ($P$) in order to obtain a single reward (black dot) of subjective reward
-% intensity $RI$. The trial duration is $25\times \mathrm{price}$ (plus
-% $2$s each time the price is attained, during which the lever is retracted so it cannot
-% work; not shown).
-% }
-% \label{fig:task_data}
-% \end{figure}
-
-% \begin{figure}[ht] 
-% \widefigure{\fullpagewidth}{Fig1}
-% \caption{(Colour online) \textbf{Task and key features of the
-%  data.} \\
-% A) Cumulative handling time (CHT) task. Grey bars denote work
-% (depressing a lever), white gaps show leisure. The subject must
-% accumulate work up to a total period of time called the
-% \emph{price} ($P$) in order to obtain a single reward (black dot) of subjective reward
-% intensity $RI$. The trial duration is $25\times \mathrm{price}$ (plus
-% $2$s each time the price is attained, during which the lever is retracted so it cannot
-% work; not shown).
-% }
-% \label{newfig:task_data}
-% \end{figure}
-
-% \clearpage
-% \section{Sample tables}
-
-% \begin{table}[!ht]
-% \caption{Time of the Transition Between Phase 1 and Phase 2$^{a}$}
-% \label{tab:label}
-% \centering
-% \begin{tabular}{lc}
-% \hline
-%  Run  & Time (min)  \\
-% \hline
-%   $l1$  & 260   \\
-%   $l2$  & 300   \\
-%   $l3$  & 340   \\
-%   $h1$  & 270   \\
-%   $h2$  & 250   \\
-%   $h3$  & 380   \\
-%   $r1$  & 370   \\
-%   $r2$  & 390   \\
-% \hline
-% \multicolumn{2}{l}{$^{a}$Table note text here.}
-% \end{tabular}
-% \end{table}
-
-% \begin{table}[ht]
-% \widecaption{Sample table taken from [treu03]\label{tbl-1}}
-% \begin{widetable}
-% \advance\tabcolsep-1pt
-% \small
-% \begin{tabular}{ccrrccccccccc}
-% \hline
-% \bf 
-% POS &\bf  chip &\multicolumn1c{\bf ID} &\multicolumn1c{\bf X}
-% &\multicolumn1c{\bf Y} &\bf
-% RA &\bf DEC &\bf IAU$\pm$ $\delta$ IAU &\bf
-% IAP1$\pm$ $\delta$ IAP1 &\bf IAP2 $\pm$ $\delta$
-% IAP2 &\bf star &\bf E &\bf Comment\\
-% \hline
-% 0 & 2 & 1 & 1370.99 & 57.35\rlap{$^a$}    &   6.651120 &  17.131149 &
-% 21.344$\pm$0.006\rlap{$^b$}  & 2 4.385$\pm$0.016 & 23.528$\pm$0.013 & 0.0 & 9 & -    \\
-% 0 & 2 & 2 & 1476.62 & 8.03     &   6.651480 &  17.129572 & 21.641$\pm$0.005  & 2 3.141$\pm$0.007 & 22.007$\pm$0.004 & 0.0 & 9 & -    \\
-% 0 & 2 & 3 & 1079.62 & 28.92    &   6.652430 &  17.135000 & 23.953$\pm$0.030  & 2 4.890$\pm$0.023 & 24.240$\pm$0.023 & 0.0 & - & -    \\
-% 0 & 2 & 4 & 114.58  & 21.22    &   6.655560 &  17.148020 & 23.801$\pm$0.025  & 2 5.039$\pm$0.026 & 24.112$\pm$0.021 & 0.0 & - & -    \\
-% 0 & 2 & 5 & 46.78   & 19.46    &   6.655800 &  17.148932 & 23.012$\pm$0.012  & 2 3.924$\pm$0.012 & 23.282$\pm$0.011 & 0.0 & - & -    \\
-% 0 & 2 & 6 & 1441.84 & 16.16    &   6.651480 &  17.130072 & 24.393$\pm$0.045  & 2 6.099$\pm$0.062 & 25.119$\pm$0.049 & 0.0 & - & -    \\
-% 0 & 2 & 7 & 205.43  & 3.96     &   6.655520 &  17.146742 & 24.424$\pm$0.032  & 2 5.028$\pm$0.025 & 24.597$\pm$0.027 & 0.0 & - & -    \\
-% 0 & 2 & 8 & 1321.63 & 9.76     &   6.651950 &  17.131672 &
-% 22.189$\pm$0.011  & 2 4.743$\pm$0.021 & 23.298$\pm$0.011 & 0.0 & 4 &
-% edge \\
-% \hline
-% \multicolumn{13}{l}{%
-% Table 2 is published in its entirety in the electronic
-% edition of the {\it Astrophysical Journal}.}\\[3pt]
-% \multicolumn{13}{l}{%
-% $^a$ Sample footnote for table 2.}\\[3pt]
-% \multicolumn{13}{l}{%
-% $^b$ Another sample footnote for table 2.}
-% \end{tabular}
-% \end{widetable}
-% \end{table}
-
-% \begin{table}[p]
-% \rotatebox{90}{\vbox{\hsize=\textheight
-% \caption{Here is a caption for a table that is found in landscape
-% mode.}
-% \begin{tabular}{ccrrccccccccc}
-% \hline
-% \bf 
-% POS &\bf  chip &\multicolumn1c{\bf ID} &\multicolumn1c{\bf X}
-% &\multicolumn1c{\bf Y} &\bf
-% RA &\bf DEC &\bf IAU$\pm$ $\delta$ IAU &\bf
-% IAP1$\pm$ $\delta$ IAP1 &\bf IAP2 $\pm$ $\delta$
-% IAP2 &\bf star &\bf E &\bf Comment\\
-% \hline
-% 0 & 2 & 1 & 1370.99 & 57.35\rlap{$^a$}    &   6.651120 &  17.131149 &
-% 21.344$\pm$0.006\rlap{$^b$}  & 2 4.385$\pm$0.016 & 23.528$\pm$0.013 & 0.0 & 9 & -    \\
-% 0 & 2 & 2 & 1476.62 & 8.03     &   6.651480 &  17.129572 & 21.641$\pm$0.005  & 2 3.141$\pm$0.007 & 22.007$\pm$0.004 & 0.0 & 9 & -    \\
-% 0 & 2 & 3 & 1079.62 & 28.92    &   6.652430 &  17.135000 & 23.953$\pm$0.030  & 2 4.890$\pm$0.023 & 24.240$\pm$0.023 & 0.0 & - & -    \\
-% 0 & 2 & 4 & 114.58  & 21.22    &   6.655560 &  17.148020 & 23.801$\pm$0.025  & 2 5.039$\pm$0.026 & 24.112$\pm$0.021 & 0.0 & - & -    \\
-% 0 & 2 & 5 & 46.78   & 19.46    &   6.655800 &  17.148932 & 23.012$\pm$0.012  & 2 3.924$\pm$0.012 & 23.282$\pm$0.011 & 0.0 & - & -    \\
-% 0 & 2 & 6 & 1441.84 & 16.16    &   6.651480 &  17.130072 & 24.393$\pm$0.045  & 2 6.099$\pm$0.062 & 25.119$\pm$0.049 & 0.0 & - & -    \\
-% 0 & 2 & 7 & 205.43  & 3.96     &   6.655520 &  17.146742 & 24.424$\pm$0.032  & 2 5.028$\pm$0.025 & 24.597$\pm$0.027 & 0.0 & - & -    \\
-% 0 & 2 & 8 & 1321.63 & 9.76     &   6.651950 &  17.131672 &
-% 22.189$\pm$0.011  & 2 4.743$\pm$0.021 & 23.298$\pm$0.011 & 0.0 & 4 &
-% edge \\
-% \hline
-% \multicolumn{13}{l}{%
-% Table 2 is published in its entirety in the electronic
-% edition of the {\it Astrophysical Journal}.}\\[3pt]
-% \multicolumn{13}{l}{%
-% $^a$ Sample footnote for table 2.}\\[3pt]
-% \multicolumn{13}{l}{%
-% $^b$ Another sample footnote for table 2.}
-% \end{tabular}
-% }}
-% \end{table}
-% \clearpage
-
-% \vglue 3in
-% Example of table continuing over pages:
-
-% \begin{center}
-% \begin{longtable}{ccc@{}}
-% \caption{ApJ costs from 1991 to 2013
-% \label{tab:table}} \\[2pt]
-% \hline
-% \bf Year & \bf Subscription & \bf Publication \\
-%  & \bf cost &\bf charges\\
-%  & \bf(\$) & \bf (\$/page)\\
-% \hline
-% \endfirsthead
-
-% \multicolumn3c{Table \thetable, \it continued from previous page.}\\[6pt]
-% \multicolumn3c{ApJ costs from 1991 to 2013}\\[2pt]
-% \hline
-% \bf Year & \bf Subscription & \bf Publication \\
-%  & \bf cost &\bf charges\\
-%  & \bf(\$) & \bf (\$/page)\\
-% \hline
-% \endhead
-% \\\hline
-% \\[-8pt]
-% \multicolumn{3}{r}{\it Table continued on next page}\\ 
-% \endfoot
-
-% \hline
-% \endlastfoot
-
-% 1991 & 600 & 100 \\
-% 1992 & 650 & 105 \\
-% 1993 & 550 & 103 \\
-% 1994 & 450 & 110 \\
-% 1995 & 410 & 112 \\
-% 1996 & 400 & 114 \\
-% 1997 & 525 & 115 \\
-% 1998 & 590 & 116 \\
-% 1999 & 575 & 115 \\
-% 2000 & 450 & 103 \\
-% 2001 & 490 &  90 \\
-% 2002 & 500 &  88 \\
-% 2003 & 450 &  90 \\
-% 2004 & 460 &  88 \\
-% 2005 & 440 &  79 \\
-% 2006 & 350 &  77 \\
-% 2007 & 325 &  70 \\
-% 2008 & 320 &  65 \\
-% 2009 & 190 &  68 \\
-% 2010 & 280 &  70 \\
-% 2011 & 275 &  68 \\
-% 2012 & 150 &  56 \\
-% 2013 & 140 &  55 \\
-% \end{longtable}
-% \end{center}
-
-% %%%%%%%%%%%%%%%%%%%%%%%
-% %% The bibliography
-
-% %% The bibliography is made using only
-% %% the entries that you cite using \cite{}, or one of the Natbib citation
-% %% entries, like \citep{}, \citet{} etc.
-
-% \bibliography{bibsamp}
-
-% NO APPENDICES allowed in the Network Neuroscience Style.
-
-% Please submit Supporting Information documents as PDFs in a format ready to publish.

### 2024-02-21 paper/manuscript.tex
finish section sequence of mappings

@@
 
+\renewcommand{\equationautorefname}{Eq.}
+\renewcommand{\tableautorefname}{Tab.}
 \def\R{\mathbb R}
@@
 
+[TODO: add the pretty figure here]
+
 % Please keep the Author Summary between 150 and 200 words
@@
 \section*{Author summary}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Curabitur eget porta erat.
-Morbi consectetur est vel gravida pretium.
-Suspendisse ut dui eu ante cursus gravida non sed sem.
-Nullam sapien tellus, commodo id velit id, eleifend volutpat quam.
-Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
-Nunc pellentesque pretium diam, quis maximus dolor faucibus id.
-Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
-Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
-
+[TODO]
 \linenumbers
@@
 
-While, historically, the spatial dimensions in neural field models have been the two dimensions of the cortical sheet (with the notable exception of the ring models for orientation selectivity in the visual cortex \cite{BenBar95} and the head-direction system \cite{Zha96}), the "space" in neural field models does not need to represent physical space but can represent any suitable abstract embedding space.
+While, historically, the spatial dimensions in neural field models have been the two dimensions of the cortical sheet (with the notable exception of the ring models for orientation selectivity in the visual cortex \cite{BenBar95} and the head-direction system \cite{Zha96}), the ``space'' in neural field models does not need to represent physical space but can represent any suitable abstract embedding space.
 Neural field models with abstract embedding spaces could constitute a generalization of classical neural field theory for cortical networks that are not solely structured by distances of neurons on the cortical sheet.
@@
 
-Without loss of generality, a simple bijective mapping can be applied to reduce any $p$-dimensional neural field to a neural field on $[0,1]^p$, additionally redefining the connectivity kernel to absorb the density and determinant of the transformation, we arrive at \autoref{eq:nf-in-01p}, a general form for any neural field equation in $p$ dimensions:
+Without loss of generality, a simple bijective mapping can be applied to reduce any $p$-dimensional neural field to a neural field on $[0,1]^p$.
+Additionally redefining the connectivity kernel to absorb the density and determinant of the transformation, we arrive at \autoref{eq:nf-in-01p}, a general form for any neural field equation in $p$ dimensions:
 
@@
 
-In particular for $p=2$, we show that for any two-dimensional neural field equation (that is, with two spatial dimensions), there is a one-dimensional neural field equation, where the embedding space is simply the interval $[0,1]$, from which the solution to the two-dimensional equation can be fully reconstructed.
+In particular for $p=2$, we show that for any two-dimensional neural field equation (that is, with two spatial dimensions), there is a one-dimensional neural field equation (where the embedding space is simply the interval $[0,1]$), from which the solution to the two-dimensional equation can be fully reconstructed.
 The mapping from the two-dimensional equation to the one-dimensional equation is done via well-known measurable bijections between the square $[0,1]^2$ and the interval $[0,1]$.
@@
 The core of our argument lies in the fact that we can relax the requirement that the spatial embedding is continuous, instead, we only require it is ``continuous on average''.
-We quantify this by introducing ``locality'' $V(S^{-1})$, which measures if small neighbourhoods in $[0,1]^2$ are mapped to small neighbourhoods in $[0,1]$ by the mapping $S^{-1}$.
+We quantify this by introducing the ``locality'' $V(S^{-1})$ of the mapping $S^{-1}$, which measures if small neighbourhoods in $[0,1]^2$ are mapped to small neighbourhoods in $[0,1]$.
 Using locality, we then prove that the mean-field approximation is still valid, implying that we can simulate the one-dimensional neural field using classical grid-based methods.
@@
 
-\subsection{Motivation}
+\subsection{Motivation for a grid-based approach}
 
 We motivate the upcoming study of the mappings $S$ by considering a grid discretization of the original two-dimensional space $[0,1]^2$.
-For this, we subdivide each axis into $2^n$ equal segments, where $n \in \mathbb{N}$ describes the grid resolution.
-Each segment has a length of $2^{-n}$, resulting in a grid constituted of $2^n \times 2^n = 4^n$ squares in $[0,1]^2$.
+For this, we subdivide each axis into $2^n$ equal segments (each segment has a length of $2^{-n}$), where $n \in \mathbb{N}$ describes the grid resolution.
+This results in a grid constituted of $2^n \times 2^n = 4^n$ squares in $[0,1]^2$.
 To each square bin indexed $i \in \{1,\cdots,4^n\}$, we associate a position $\vec{v_i} = (v_{i,1}, v_{i,2})$ located within the bounds of the bin.
 Since everytime $n$ is incremented by one, every axis is subdivided into two smaller equal parts, the positions $v_{i,1}$ and $v_{i,2}$ naturally lend themselves to a binary notation.
-We also introduce $v_\mu^{(n)}$ to denote the $n$-bit truncation of the position $v_\mu \in [0,1]$, where $\mu = 1,2$.
+We introduce $v_\mu^{(n)}$ to denote the $n$-bit truncation of the position $v_\mu \in [0,1]$, where $\mu = 1,2$ indicates the axis of the dimension.
 Additionally, we write the error of the trunctation as a small corrective factor $0 \leq c^{(n)} < 2^{-n}$.
@@
 
-It can easily be seen that $v_\mu^{(n)} \to v_\mu$ as $n\to\infty$. In general, the bits of the $v_\mu$ can be expressed as:
+Since $c_\mu^{(n)} < 2^n$, it can easily be seen that $v_\mu^{(n)} \to v_\mu$ as $n\to\infty$. In general, the bits of the $v_\mu$ can be expressed as:
 
@@
 
-[TODO: add a figure here of the discretization here ? do we need to clarify explicitly the bounds of each square bin ?]
-
-Sampling the original connectivity kernel at the points $\vec{v_i}$ and $\vec{v_j}$ yields a numerical connectivity matrix $J_{ij} := w_U(\vec{v_i}, \vec{v_j})$, which can be used to simulate the resulting population dynamics equation:
+Sampling the original connectivity kernel at the points $\vec{v_i}$ and $\vec{v_j}$ yields a numerical connectivity matrix $J_{ij} := w_U(\vec{v_i}, \vec{v_j})$, illustrated in Fig.
+[TODO: add a figure here of the discretization here and the sampling on a grid, and the limit of n to infty.
+do we need to clarify explicitly the bounds of each square bin ?], which can be used to simulate the resulting population dynamics equation:
 
@@
 \autoref{eq:population-dynamics} can be seen as the discretized version of \autoref{eq:nf-in-01p}, where we associate $h_i(t) := h_U(\vec{v_i}, t)$.
-Indeed, by construction of the original neural field, the discrete dynamics converge to the true solution of the neural field, because the error of the mean-field approximation in $[0,1]^2$ vanishes as we take finer grids, letting $n \to \infty$.
+Indeed, by a mean-field argument on the original smooth neural field, the discrete dynamics converge to the true solution of the neural field, because the error of the mean-field approximation in $[0,1]^2$ vanishes as we take finer grids, letting $n \to \infty$.
 
 We would now like to derive a bijective mapping $S$, such that the same argument applies on the neural field in $[0,1]$.
-Intuively, we want $S$ to conserve the regularity of the neural field on $[0,1]^2$, since if the mean-field argument also carries over to the one-dimensional neural field, it seems likely that it will express identical dynamics.
+Intuively, we want $S$ to conserve the regularity of the neural field on $[0,1]^2$, since if the mean-field argument also carries over to the one-dimensional neural field, it seems natural that it will express identical dynamics.
 
@@
 
-We note that we write $\alpha^n$ using $2n$ bits, which is intuitively the lowest number of bits required so that $S^n$ remains a bijection.
+We write $\alpha^n$ using the $2n$ bits $\{b^1, b^2, \cdots, b^{2n}\}$, which is intuitively the lowest number of bits required so that $S^n$ remains a bijection, since it takes $2n$ bits of information as argument.
 More than $2n$ bits could be allowed while maintaining bijectivity, but we omit this possibility, since they wouldn't strictly be required.
-Effectively, $S^n$ maps between $2\times n$ input bits and $2n$ output bits.
-In this formulation, it becomes clear that in order to be a bijection, every bit of the input must be used once and only once to form the output.
-Setting aside bit inversions, this means that $S^n$ defines a permutation of the $2n$ input bits.
+In this formulation, it becomes clear that in order to be a bijection, every bit of the input must be used once and only once to form the image.
+Setting aside bit inversions, this means that we let $S^n$ define a permutation of the $2n$ input bits.
 [TODO: are more details necessary ?
@@
 
-If it exists, we then we define $\alpha = S(v_1, v_2)$ as the pointwise limit of the sequence $\{S^n(v_1, v_2)\}_{n \in \mathbb N}$ (Note: We may define an extension of $S^n$ acting on an arbitrary $(v_1, v_2) \in [0,1]^2$ as $S^n(v_1, v_2) := S^n(v_1^{(n)}, v_2^{(n)})$.
-This extension is obviously not bijective, but in the $n \to \infty$ limit the finite-$n$ truncation converges, such that the pointwise limit may be a bijection).
-In terms of the current formulation, pointwise convergence requires that the least signficiant bits (LSB) of the input be mapped to the LSB of the output, and identically the most significant bits (MSB) of the input be mapped to the MSB of the output [TODO: do we need a proof of this ?].
-This ensures that small corrections in the input $\vec{v}$ results in small corrections in the output $\alpha$.
+For any finite $n$ and any $(v_1,v_2) \in [0,1]^2$, we may extend the definition of $S^n$ with $S^n(v_1, v_2) := S^n(v_1^{(n)}, v_2^{(n)})$ such that it acts on the entire space $[0,1]^2$ instead of the finite set of grid points.
+While this extension is obviously not bijective (it maps a grid cell to a single point), in the $n \to \infty$ limit the finite-$n$ truncation converges, such that the pointwise limit could still be a bijection.
+In \autoref{sec:mapping-examples} we formulate two examples, where in one case the extension converges to a bijection, and in the other it doesn't.
 
-We will in the rest of this paper show that this simple formulation in terms of permutation of input bits is flexible enough to express mappings that have good properties such as ``locality'', while being simple enough to help build intuitions.
+Given $(v_1, v_2) \in [0,1]^2$ and the sequence $\{S^n(v_1, v_2)\}_{n \in \mathbb N}$, we then we define $\alpha = S(v_1, v_2)$ as its limit, if it exists.
+Informally, pointwise convergence of $\{S^n\}_{n\in\mathbb N}$ requires that the least signficiant bits (LSB) of the input be mapped to the LSB of the image, ensuring that small corrections in the input $\vec{v}$ result in small corrections in the image $\alpha$.
+% TODO: I would like to talk briefly about the MSB intuition
+% and identically the most significant bits (MSB) of the input be mapped to the MSB of the image. ``no information is lost''.
 
----
+To illustrate this, let us give an example of the sequence $\{A^n\}_{n \in \mathbb N}$ that does not converge pointwise:
 
-let us think of what happens when we apply the mapping on a discrete grid of squares in [0,1]², and where the squares end up on the segment.
-the squares are bijectively mapped to segments on the unit line, i.e.
-it defines a permutation of the squares onto the segments, or as defining an "ordering" of the square bins in $[0,1]^2$.
-This is done using a mapping S : [0, 1]2 7 → [0, 1],
-where we cover the [0, 1]2 space by following the order assigned to the populations in
-the image [0, 1].
-considering finer and finer grids, we look/see/observe where the squares end up, and may define a sequence of mappings $S^n$ acting on finer and finer grids.
-if the sequence converges pointwise, we then define $S$ as the limit of the sequence of mappings $S^n$.
+\begin{equation*}
+\alpha^{(n)} = A^n(v_1^{(n)},v_2^{(n)}) = 0.b_1^n b_2^n b_1^{n-1} b_2^{n-1} \cdots b_1^1 b_2^1 = \sum_{k=1}^n b_1^{n+1-k} 2^{1-2k} + b_2^{n+1-k} 2^{-2k}.
+\end{equation*}
 
-here is how we mathematically write down / formalism for the mappings $S^n$.
-special case where the bins are uniformly spaced along each axis, and we obtain the next iteration by subdividing each bin of the axis into two equally smaller bins.
-along each axis, every bin can be located by a tuple of coordinates.
-a particularly elegant way to locate the square bins is to use binary notation (base 2), such that at iteration n, every coordinate has n bits of information.
-therefore every square can be uniquely located by a total of 2n bits.
+The permutation between the bits of $\alpha^{(n)}$, and the bits of $(v_1^{(n)}, v_2^{(n)})$ is illustrated in Fig.
+[TODO: arrow connecting the bits together], and shows how this example does not respect the bit ordering described previously.
+For finite $n$, $A^n$ is clearly bijective and has a well-defined inverse $A^n_\textrm{inv}$:
 
-similarly, we subdivide the unit segment into 4 every time, such that the number of segments matches the number of squares.
+\begin{equation*} \begin{aligned}
+v_1^{(n)} &= A^n_{\textrm{inv},1}(\alpha^{(n)}) = 0.b^{2n-1} b^{2n-3} \cdots b^1 = \sum_{k=1}^n b^{2(n-k)-1} 2^{-k}, \\
+v_2^{(n)} &= A^n_{\textrm{inv},2}(\alpha^{(n)}) = 0.b^{2n-2} b^{2n-4} \cdots b^2 = \sum_{k=1}^n b^{2(n+1-k)} 2^{-k}.
+\end{aligned} \end{equation*}
 
-then
+However, for any $(v_1,v_2)$, the sequence $\{A^n(v_1, v_2)\}_{n \in \mathbb N}$ does not converge.
+In Fig.
+[TODO: fig 3.8 of the thesis] we graph the mapping $A^n$, which demonstrates that the image of any point $(v_1,v_2)$ in $[0,1]^2$ diverges as $n$ increases.
+As $n$ increases, $A^n$ ``refines'' its input with fluctuations of order $2^{-n}$ (because it works with truncations of $(v_1, v_2)$, which have error $c^{(n)} < 2^{-n}$), but the image $\alpha^{(n)}=b_1^n 4^{-1} + b_2^n 4^{-2} + b_1^{n-1} 4^{-3} + b_2^{n-1} 4^{-4} \cdots$ fluctuates (at least) with fixed amplitude $\sim 4^{-2}$, independantly of $n$.
+Therefore the image never converges, and $\{A_n\}_{n \in \mathbb N}$ is not pointwise convergent.
 
-\begin{equation} \begin{aligned}
-S^n &: (v_1^{(n)}, v_2^{(n)}) = (0.b_1^1 b_2^1 \cdots b_n^1, 0.b_1^2 b_2^2 \cdots b_n^2) \\
-&\mapsto \alpha^{(n)} = 0.b_1 b_2 \cdots b_{2n}
-\end{aligned} \end{equation}
+Within the context of sequences of mappings acting as permutations on the bits of continually refined grid points, the criterion of pointwise convergence restricts the general ``form'' that those mappings can take.
+In the following section, we give the standard examples of the Column and the Z-mapping, showing that our formulation is flexible enough to express mappings that have good properties such as ``locality'', while being simple enough to help build intuitions.
 
-\section{Locality of the Z-mapping and the Column mapping}
+\subsection{Two examples of sequences $\{S^n\}_{n \in \mathbb N}$: the Column and the Z-mapping} \label{sec:mapping-examples}
 
-% \subsection{$S$ as the limit of a sequence of mappings}
+introduce the Column mapping and the Z-mapping, show the figures
 
-% motivation, why do we do this
+\section{The measure of locality}
 
-% \subsection{The Z-mapping and Column mapping}
+where do the populations end up
 
-% introduce the mappings and give plots
-
-% \subsection{Behavior and locality measures}
-
-% measure of locality, give examples applied to the mappings.
+\subsection{Locality of the Z-mapping and the Column mapping}
 
@@
 
-\subsection{Numerical demonstrations}
+\section{Numerical demonstrations}
 
@@
 
+goal of this section : "it seems that it works for the Z-mapping"
+
 \section{Proof of results}
@@
 \section{Conclusion}
+
+% TODO: are all bijective limits local ? does local imply bijection ?
+% in this case, we have shown that bijection + locality => we can safely do the mapping (bijection for the analytical approach, locality for the numerical approach)
 

### 2024-02-22 paper/manuscript.tex
finish column and z mapping defs

@@
 This results in a grid constituted of $2^n \times 2^n = 4^n$ squares in $[0,1]^2$.
-To each square bin indexed $i \in \{1,\cdots,4^n\}$, we associate a position $\vec{v_i} = (v_{i,1}, v_{i,2})$ located within the bounds of the bin.
-Since everytime $n$ is incremented by one, every axis is subdivided into two smaller equal parts, the positions $v_{i,1}$ and $v_{i,2}$ naturally lend themselves to a binary notation.
-We introduce $v_\mu^{(n)}$ to denote the $n$-bit truncation of the position $v_\mu \in [0,1]$, where $\mu = 1,2$ indicates the axis of the dimension.
+To each square bin indexed $i \in \{1,\cdots,4^n\}$, we associate a position $\vec{v_i^{(n)}} = (v_{i,1}^{(n)}, v_{i,2}^{(n)})$ located at the origin of the bin.
+Since everytime $n$ is incremented by one, every axis is subdivided into two smaller equal parts, the positions $v_{i,1}^{(n)}$ and $v_{i,2}^{(n)}$ naturally lend themselves to a binary notation.
+
+For the component $v_\mu \in [0,1]$ of an arbitrary position $\vec{v} \in [0,1]^2$, we introduce $v_\mu^{(n)}$ to denote its $n$-bit truncation, where $\mu = 1,2$ (and in general for $p$-dimensional grids, $\mu=1, \cdots, p$) indicates the axis of the dimension.
 Additionally, we write the error of the trunctation as a small corrective factor $0 \leq c^{(n)} < 2^{-n}$.
-Using this notation and dropping the index $i$ in favor of readability, we write the components of $\vec{v} \in [0,1]^2$ as:
+In bit notation, this yields:
 
@@
 
-Since $c_\mu^{(n)} < 2^n$, it can easily be seen that $v_\mu^{(n)} \to v_\mu$ as $n\to\infty$. In general, the bits of the $v_\mu$ can be expressed as:
+Since $c_\mu^{(n)} < 2^n$, it can easily be seen that $v_\mu^{(n)} \to v_\mu$ as $n\to\infty$, that is, as we take finer and finer grids, the truncation tends to the true value of $\vec{v}$.
+In general, the bits of the component $v_\mu$ can be expressed as:
 
@@
 
-Sampling the original connectivity kernel at the points $\vec{v_i}$ and $\vec{v_j}$ yields a numerical connectivity matrix $J_{ij} := w_U(\vec{v_i}, \vec{v_j})$, illustrated in Fig.
+Sampling the original connectivity kernel at the grid points $\vec{v_i}$ and $\vec{v_j}$ yields a numerical connectivity matrix $J_{ij} := w_U(\vec{v_i}, \vec{v_j})$, illustrated in Fig.
 [TODO: add a figure here of the discretization here and the sampling on a grid, and the limit of n to infty.
@@
 
-Let us define a sequence of bijective mappings $S^n$, acting on the $n$-bit truncations of $\vec{v_i}$, as well as the sequence $\alpha_i^n \in [0,1]$ of positions resulting from the image of $\vec{v_i^{(n)}}$.
-In the following, we again drop the index $i$ for clarity.
+Let us define a sequence of bijective mappings $\{S^n\}_{n \in \mathbb N}$, acting on the $n$-bit truncations $\vec{v^{(n)}}$ of any $\vec{v} \in [0,1]^2$, as well as the corresponding sequence of images $\{\alpha^n := S^n(\vec{v^{(n)}})\}_{n \in \mathbb N}$.
+$S^n$ defines the following bijection between points on the two-dimensional grid and the one-dimensional line:
 
-\begin{equation} \begin{aligned}
-S^n &: (v_1^{(n)}, v_2^{(n)}) = (0.b_1^1 b_1^2 \cdots b_1^n, 0.b_2^1 b_2^2 \cdots b_2^n) \\
-&\mapsto \alpha^n = 0.b^1 b^2 \cdots b^{2n}
+\begin{equation} \label{eq:def-S} \begin{aligned}
+S^n :& \{\tfrac{j-1}{2^n}\;|\;j\in\{1,\cdots,2^n\}\}^2 \mapsto \{\tfrac{i-1}{4^n}\;|\;i\in\{1,\cdots,4^n\}\},\\
+\text{such that }&(v_1^{(n)}, v_2^{(n)}) = (0.b_1^1 b_1^2 \cdots b_1^n, 0.b_2^1 b_2^2 \cdots b_2^n) \mapsto \alpha^n = 0.b^1 b^2 \cdots b^{2n}.
 \end{aligned} \end{equation}
+
+% NOTE: it is important to write \alpha^n, and not \alpha^{(n)}, since in all generality the n-bit trunction of \alpha is NOT equal to \alpha^n (example: the column mapping \alpha^n = C(v))
 
@@
 For any finite $n$ and any $(v_1,v_2) \in [0,1]^2$, we may extend the definition of $S^n$ with $S^n(v_1, v_2) := S^n(v_1^{(n)}, v_2^{(n)})$ such that it acts on the entire space $[0,1]^2$ instead of the finite set of grid points.
-While this extension is obviously not bijective (it maps a grid cell to a single point), in the $n \to \infty$ limit the finite-$n$ truncation converges, such that the pointwise limit could still be a bijection.
+While this extension is obviously not bijective (all points in a grid cell are ``rounded'' to the corresponding grid position), in the $n \to \infty$ limit the finite-$n$ truncation converges, such that the pointwise limit could still be a bijection.
 In \autoref{sec:mapping-examples} we formulate two examples, where in one case the extension converges to a bijection, and in the other it doesn't.
@@
 Given $(v_1, v_2) \in [0,1]^2$ and the sequence $\{S^n(v_1, v_2)\}_{n \in \mathbb N}$, we then we define $\alpha = S(v_1, v_2)$ as its limit, if it exists.
-Informally, pointwise convergence of $\{S^n\}_{n\in\mathbb N}$ requires that the least signficiant bits (LSB) of the input be mapped to the LSB of the image, ensuring that small corrections in the input $\vec{v}$ result in small corrections in the image $\alpha$.
-% TODO: I would like to talk briefly about the MSB intuition
-% and identically the most significant bits (MSB) of the input be mapped to the MSB of the image. ``no information is lost''.
+Informally, pointwise convergence of $\{S^n\}_{n\in\mathbb N}$ requires that the least signficiant bits of the input be mapped to the least significant bits of the image, ensuring that small corrections in the input $\vec{v}$ result in small corrections in the image $\alpha$.
 
@@
 \begin{equation*}
-\alpha^{(n)} = A^n(v_1^{(n)},v_2^{(n)}) = 0.b_1^n b_2^n b_1^{n-1} b_2^{n-1} \cdots b_1^1 b_2^1 = \sum_{k=1}^n b_1^{n+1-k} 2^{1-2k} + b_2^{n+1-k} 2^{-2k}.
+\alpha_A^n = A^n(v_1^{(n)},v_2^{(n)}) = 0.b_1^n b_2^n b_1^{n-1} b_2^{n-1} \cdots b_1^1 b_2^1 = \sum_{k=1}^n b_1^{n+1-k} 2^{1-2k} + b_2^{n+1-k} 2^{-2k}.
 \end{equation*}
 
-The permutation between the bits of $\alpha^{(n)}$, and the bits of $(v_1^{(n)}, v_2^{(n)})$ is illustrated in Fig.
+The permutation between the bits of $\alpha^n$, and the bits of $(v_1^{(n)}, v_2^{(n)})$ is illustrated in Fig.
 [TODO: arrow connecting the bits together], and shows how this example does not respect the bit ordering described previously.
@@
 \begin{equation*} \begin{aligned}
-v_1^{(n)} &= A^n_{\textrm{inv},1}(\alpha^{(n)}) = 0.b^{2n-1} b^{2n-3} \cdots b^1 = \sum_{k=1}^n b^{2(n-k)-1} 2^{-k}, \\
-v_2^{(n)} &= A^n_{\textrm{inv},2}(\alpha^{(n)}) = 0.b^{2n-2} b^{2n-4} \cdots b^2 = \sum_{k=1}^n b^{2(n+1-k)} 2^{-k}.
+v_1^{(n)} &= A^n_{\textrm{inv},1}(\alpha^n) = 0.b_A^{2n-1} b_A^{2n-3} \cdots b_A^1 = \sum_{k=1}^n b_A^{2(n-k)-1} 2^{-k}, \\
+v_2^{(n)} &= A^n_{\textrm{inv},2}(\alpha^n) = 0.b_A^{2n} b_A^{2n-2} \cdots b_A^2 = \sum_{k=1}^n b_A^{2(n+1-k)} 2^{-k}.
 \end{aligned} \end{equation*}
@@
 [TODO: fig 3.8 of the thesis] we graph the mapping $A^n$, which demonstrates that the image of any point $(v_1,v_2)$ in $[0,1]^2$ diverges as $n$ increases.
-As $n$ increases, $A^n$ ``refines'' its input with fluctuations of order $2^{-n}$ (because it works with truncations of $(v_1, v_2)$, which have error $c^{(n)} < 2^{-n}$), but the image $\alpha^{(n)}=b_1^n 4^{-1} + b_2^n 4^{-2} + b_1^{n-1} 4^{-3} + b_2^{n-1} 4^{-4} \cdots$ fluctuates (at least) with fixed amplitude $\sim 4^{-2}$, independantly of $n$.
+As $n$ increases, $A^n$ ``refines'' its input with fluctuations of order $2^{-n}$ (because it works with truncations of $(v_1, v_2)$, which have error $c^{(n)} < 2^{-n}$), but the image $\alpha^n=b_1^n 2^{-1} + b_2^n 2^{-2} + b_1^{n-1} 2^{-3} + b_2^{n-1} 2^{-4} \cdots$ fluctuates (at least) with fixed amplitude $\sim 2^{-2}$, independantly of $n$.
 Therefore the image never converges, and $\{A_n\}_{n \in \mathbb N}$ is not pointwise convergent.
@@
 
-\subsection{Two examples of sequences $\{S^n\}_{n \in \mathbb N}$: the Column and the Z-mapping} \label{sec:mapping-examples}
+\subsection{Two examples of sequences of mappings: the Column and the Z-mapping} \label{sec:mapping-examples}
 
-introduce the Column mapping and the Z-mapping, show the figures
+Let us provide two examples of sequence of mappings that are pointwise convergent, and act as permutation of the grid point bits.
+We first define the sequence of Column mappings $\{C^n\}_{n\in\mathbb N}$ as the concatenation of the input bits of $v_1^{(n)}$, followed by the bits of $v_2^{(n)}$.
+We define the image $\alpha_C^n = 0.b_C^1 b_C^2 \cdots b_C^{2n}$ as:
 
-\section{The measure of locality}
+\begin{equation} \label{eq:def-column-mapping}
+\alpha_C^n := C^n(v_1^{(n)}, v_2^{(n)}) = 0.b_1^1 b_1^2 \cdots b_1^n b_2^1 b_2^2 \cdots b_2^n = \sum_{k=1}^n b_1^k 2^{-k} + b_2^k 2^{-(k+n)}.
+\end{equation}
 
-where do the populations end up
+Instead of concatenation, we may also interleave the bits of $v_1^{(n)}$ and $v_2^{(n)}$, defining the sequence of Z-mappings $\{Z^n\}_{n\in\mathbb N}$.
+We define the image $\alpha_Z^n = 0.b_Z^1 b_Z^2 \cdots b_Z^{2n}$ as:
 
-\subsection{Locality of the Z-mapping and the Column mapping}
+\begin{equation} \label{eq:def-z-mapping}
+\alpha_Z^n := Z^n(v_1^{(n)}, v_2^{(n)}) = 0.b_1^1 b_2^1 b_1^2 b_2^2 \cdots b_1^n b_2^n = \sum_{k=1}^n b_1^k 2^{1-2k} + b_2^k 2^{-2k}.
+\end{equation}
 
-% see how it scales, and say what we expect
+We graph both mappings in Fig. [TODO]. At finite $n$, both mappings can be inversed to retrieve the truncations:
 
-\section{Numerical demonstrations}
+\begin{equation} \begin{aligned}
+C^n_{\textrm{inv},1}(\alpha_C^n) &= 0.b_C^1 b_C^2 \cdots b_C^n = \sum_{k=1}^n b_C^k 2^{-k} = v_1^{(n)}, \\
+C^n_{\textrm{inv},2}(\alpha_C^n) &= 0.b_C^{n+1} b_C^{n+2} \cdots b_C^{2n} = \sum_{k=1}^n b_C^{n+k} 2^{-k} = v_2^{(n)}, \\
+Z^n_{\textrm{inv},1}(\alpha_Z^n) &= 0.b_Z^1 b_Z^3 \cdots b_Z^{2n-1} = \sum_{k=1}^n b_Z^{2k-1} 2^{-k} = v_1^{(n)},\\
+Z^n_{\textrm{inv},2}(\alpha_Z^n) &= 0.b_Z^2 b_Z^4 \cdots b_Z^{2n} = \sum_{k=1}^n b_Z^{2k} 2^{-k} = v_2^{(n)}.
+\end{aligned} \end{equation}
+
+Both mappings converge pointwise, since for any $(v_1,v_2) \in [0,1]^2$, the limit exists, which we write as $C(v_1,v_2)$ and $Z(v_1,v_2)$ respectively, and we may find the following upper bounds:
+
+\begin{equation} \begin{aligned}
+& C(v_1,v_2) = \sum_{k=1}^\infty b_1^k 2^{-k} = v_1, \;\text{and}\; |C^n(v_1,v_2)-C(v_1,v_2)| \leq 2^{-n},\\
+& Z(v_1,v_2) = \sum_{k=1}^\infty b_1^k 2^{-k} + b_2^k 2^{-(k+n)}, \;\text{and}\; |Z^n(v_1,v_2)-Z(v_1,v_2)|\leq 2^{-(2n-1)}.
+\end{aligned} \end{equation}
+
+Importantly, the limit $C(v_1,v_2)=v_1$ is a projection, and is therefore not bijective.
+This result is not suprising, since in the definition of $C^n$ (\autoref{eq:def-column-mapping}) the information on the bits of $v_2$ appears only after the bit $n$, and gets discarded in the $n \to \infty$ limit.
+The Z-mapping does not suffer from the same problem, since informally the most significant bits of the input get mapped to the most significant bits of the image (no information gets lost), and we can conveniently take the $n \to \infty$ limit of $Z_{\textrm{inv},1}^n$ and $Z_{\textrm{inv},2}^n$:
+
+\begin{equation} \begin{aligned}
+&Z_{\textrm{inv,1}}^n(\alpha^n) \xrightarrow{n \to \infty} Z^{-1}_1(\alpha) = 0.b_Z^1 b_Z^3 \cdots = \sum_{k=1}^\infty b_Z^{2k-1} 2^{-k} = v_1,\\
+&Z_{\textrm{inv,2}}^n(\alpha^n) \xrightarrow{n \to \infty} Z^{-1}_2(\alpha) = 0.b_Z^2 b_Z^4 \cdots = \sum_{k=1}^\infty b_Z^{2k} 2^{-k} = v_2.
+\end{aligned} \end{equation}
+
+\section{Numerical simulations of mappings from two-dimensional to one-dimensional neural fields}
+
+\subsection{Coarse-graining the images to numerically validate the mean-field approximation}
+
+\subsection{The cycling gaussian low-rank neural field toy model}
 
@@
 goal of this section : "it seems that it works for the Z-mapping"
+
+\section{The measure of locality}
+
+where do the populations end up
 

### 2024-02-23 paper/manuscript.tex
henlo yes i did the coarse-graining stuffs

@@
 \usepackage{array}
+
+\ifndef{SUBMISSION}\usepackage{graphicx}\endif
 
@@
 \section*{Abstract}
+
+\ifndef{SUBMISSION}
+\includegraphics[width=\textwidth]{../figures/cover.pdf}
+\endif
+
 Classical neural field models describe interactions between neurons positioned in an embedding; often a physical space (for instance, that of the cortical sheet) or, later, abstract spaces (for instance, that of neuron celltypes), in which the dynamics are highly regular in space and time.
@@
 
-[TODO: add the pretty figure here]
-
 % Please keep the Author Summary between 150 and 200 words
@@
 \section*{Author summary}
-[TODO]
+[TODO?]
 \linenumbers
@@
 
-\subsection{Motivation for a grid-based approach}
+\subsection{Motivation for a grid-based approach} \label{sec:grid-motivation}
 
@@
 
-Sampling the original connectivity kernel at the grid points $\vec{v_i}$ and $\vec{v_j}$ yields a numerical connectivity matrix $J_{ij} := w_U(\vec{v_i}, \vec{v_j})$, illustrated in Fig.
-[TODO: add a figure here of the discretization here and the sampling on a grid, and the limit of n to infty.
-do we need to clarify explicitly the bounds of each square bin ?], which can be used to simulate the resulting population dynamics equation:
+Sampling the original connectivity kernel at the grid points $\vec{v_i}$ and $\vec{u_j}$ yields a numerical connectivity matrix $J_{ij} := w_U(\vec{v_i}, \vec{u_j})$, illustrated in Fig.
+[TODO], which can be used to simulate the resulting population dynamics equation:
 
@@
 \end{equation}
+
+\ifndef{SUBMISSION}
+\includegraphics[height=20em]{../figures/discretization.png}
+
+[TODO: the discretization here and the sampling on a grid, and the limit of n to infty. do we need to clarify explicitly the bounds of each square bin ?]
+\endif
 
@@
 Let us define a sequence of bijective mappings $\{S^n\}_{n \in \mathbb N}$, acting on the $n$-bit truncations $\vec{v^{(n)}}$ of any $\vec{v} \in [0,1]^2$, as well as the corresponding sequence of images $\{\alpha^n := S^n(\vec{v^{(n)}})\}_{n \in \mathbb N}$.
-$S^n$ defines the following bijection between points on the two-dimensional grid and the one-dimensional line:
+$S^n$ defines the following bijection between points on the two-dimensional and one-dimensional grids:
 
 \begin{equation} \label{eq:def-S} \begin{aligned}
-S^n :& \{\tfrac{j-1}{2^n}\;|\;j\in\{1,\cdots,2^n\}\}^2 \mapsto \{\tfrac{i-1}{4^n}\;|\;i\in\{1,\cdots,4^n\}\},\\
+S^n :& \{\tfrac{l-1}{2^n}\;|\;l\in\{1,\cdots,2^n\}\}^2 \mapsto \{\tfrac{i-1}{4^n}\;|\;i\in\{1,\cdots,4^n\}\},\\
 \text{such that }&(v_1^{(n)}, v_2^{(n)}) = (0.b_1^1 b_1^2 \cdots b_1^n, 0.b_2^1 b_2^2 \cdots b_2^n) \mapsto \alpha^n = 0.b^1 b^2 \cdots b^{2n}.
@@
 In Fig.
-[TODO: fig 3.8 of the thesis] we graph the mapping $A^n$, which demonstrates that the image of any point $(v_1,v_2)$ in $[0,1]^2$ diverges as $n$ increases.
+[TODO] we graph the mapping $A^n$, which demonstrates that the image of any point $(v_1,v_2)$ in $[0,1]^2$ diverges as $n$ increases.
 As $n$ increases, $A^n$ ``refines'' its input with fluctuations of order $2^{-n}$ (because it works with truncations of $(v_1, v_2)$, which have error $c^{(n)} < 2^{-n}$), but the image $\alpha^n=b_1^n 2^{-1} + b_2^n 2^{-2} + b_1^{n-1} 2^{-3} + b_2^{n-1} 2^{-4} \cdots$ fluctuates (at least) with fixed amplitude $\sim 2^{-2}$, independantly of $n$.
 Therefore the image never converges, and $\{A_n\}_{n \in \mathbb N}$ is not pointwise convergent.
+
+\ifndef{SUBMISSION}
+\includegraphics[width=\linewidth]{../figures/mapping-A.pdf}
+[TODO: remove the numbers on this figure, dot the grid nodes]
+\endif
 
@@
 
+\ifndef{SUBMISSION}
+\includegraphics[height=20em]{../figures/mapping-col-Z.pdf}
+
+[TODO: add labels, show n=3, etc.]
+\endif
+
 Both mappings converge pointwise, since for any $(v_1,v_2) \in [0,1]^2$, the limit exists, which we write as $C(v_1,v_2)$ and $Z(v_1,v_2)$ respectively, and we may find the following upper bounds:
@@
 
+In the interest of developing an insight on the properties of the two mappings introduced previously, let us now develop a numerical grid-based simulation.
+As established in the \autoref{sec:grid-motivation}, the original two-dimensional neural field can be simulated by sampling the connectivity kernel at pairs of discrete grid points $(\vec{v_i},\vec{u_j})$, discretizing the neural field to a set of $4^n$ scalars $\{h_i(t) \;|\; i=1,\cdots,4^n\}$.
+
+Naively, we could proceed in the same way to simulate the corresponding one-dimensional neural field, by sampling the connectivity kernel $\tilde w(\alpha_i, \beta_j) = w_U(S^{-1}(\alpha_i), S^{-1}(\beta_j)) = w_U(\vec{v_i}, \vec{u_j})$, where the pairs of discrete one-dimensional points $(\alpha_i,\beta_j)$ are obtained by mapping the two-dimensional grid-points $(\vec{v_i},\vec{u_j})$.
+
+Critically, since $S$ defines a bijection, it acts as a permutation on the discrete connectivity matrix indices.
+Furthermore the population dynamics in \autoref{eq:population-dynamics} are invariant to permutation in the indices, and the naively sampled connectivity matrix corresponding to the one-dimensional neural field trivially yields identical dynamics.
+
+In essence, since both connectivity matrices of the one- and two-dimensional neural fields have the same size, we are compensating for the reduction in dimensionality by increasing the one-dimensional grid resolution.
+In two dimensions, the $2^n \times 2^n = 4^n$ grid points are spaced with $2^{-n}$ along every axis, but after the mapping, those $4^n$ grid points are spaced every $4^{-n}$, since there are twice as many bits to describe the position along the unique axis.
+This is illustrated in Fig.
+[TODO].
+
+\ifndef{SUBMISSION}
+\includegraphics[width=\linewidth]{../figures/S-viz.export.pdf}
+[TODO: adapt this figure better, but gives the idea]
+\endif
+
+% We then loosely have the following correspondence between the discrete connectivity matrices for the one- and two-dimensional neural fields:
+
+% \begin{equation} \begin{aligned}
+% \tilde w(\alpha_i,\beta_j) &= \tilde J_{\alpha_i,\beta_j} \\
+% &= \tilde J_{S(\vec{v_i},\vec{u_j})} \\
+% &= J_{ij} \;\text{by definition} \\
+% &= w_U(\vec{v_i},\vec{u_j})
+% \end{aligned}\end{equation}
+
 \subsection{Coarse-graining the images to numerically validate the mean-field approximation}
+
+Coarse-graining adresses the problem of numerically enforcing the mean-field approximation.
+The mean-field approximation is done on the original two-dimensional neural field when sampling the connectivity matrix $J_{ij}$, as all neurons inside the same grid cells are assumed to be the same.
+This is not the case when defining the connectivity of the one-dimensional neural field as the permutation of rows and columns of $J_{ij}$.
+
+In order to perform the mean-field approximation on the one-dimensional neural field, we average together $2^n$ consecutive one-dimensional bins, each of length $4^{-n}$, resulting in $2^n$ coarse-grained bins.
+This is illustrated in Fig.
+[TODO] Specifically, for every coarse-grained bin $a=1,\cdots,2^n$, the set of consecutive positions $\mathcal{C}_a = \{\tfrac{a-1}{2^n} + \tfrac{l-1}{4^n} \;|\; l \in 1, \cdots, 2^n\}$, gets mapped to the two-dimensional grid points and averaged:
+
+\begin{equation*}
+\vec{\bar v_a} := \textrm{mean}(\{S^{-1}(\alpha_l) \;|\; \alpha_l \in \mathcal{C}_a\}).
+\end{equation*}
+
+\ifndef{SUBMISSION}
+\includegraphics[height=15em]{../figures/coarse-graining.pdf}
+
+[TODO: add the indices a,b and l on this figure, for clarity.]
+\endif
+
+The obtained coarse-grained points are then used to sample the connectivity kernel, and form the connectivity matrix of the one-dimensional neural field:
+
+\begin{equation}
+\tilde J_{ab} := w_U(\vec{\bar v_a}, \vec{\bar v_b}).
+\end{equation}
 
@@
 we stress that this toy model serves only for illustration purposes, and our results apply to any neural field
-
-introduce coarse-graining procedure (applying the smoothness hypothesis numerically)
 
@@
 % in this case, we have shown that bijection + locality => we can safely do the mapping (bijection for the analytical approach, locality for the numerical approach)
+% TODO: mention the "standard example" of the Z-mapping
 

### 2024-02-26 paper/manuscript.tex
locality

@@
 Without loss of generality, a simple bijective mapping can be applied to reduce any $p$-dimensional neural field to a neural field on $[0,1]^p$.
-Additionally redefining the connectivity kernel to absorb the density and determinant of the transformation, we arrive at \autoref{eq:nf-in-01p}, a general form for any neural field equation in $p$ dimensions:
+Additionally redefining the connectivity kernel to absorb the density and determinant of the transformation (see an example in section 2.5 of \cite{nvadot2023mathesis}), we arrive at \autoref{eq:nf-in-01p}, a general form for any neural field equation in $p$ dimensions:
 
@@
 The difficulty in writing the one-dimensional neural field, is that the spatial structure of the original neural field in $[0,1]^2$ may be lost.
+Fig.
+[TODO] illustrates this problem: if the mapping is seemingly random, the regularity of the orginal neural field is destroyed.
+Contrarily, if the mapping is ``local'', the resulting neural field might still be well-defined and exhibit identical dynamics, despite the apparent discontinuities.
 The intuition is that we would like populations close in $[0,1]^2$ to be bijectively mapped to populations close in $[0,1]$.
@@
 The core of our argument lies in the fact that we can relax the requirement that the spatial embedding is continuous, instead, we only require it is ``continuous on average''.
-We quantify this by introducing the ``locality'' $V(S^{-1})$ of the mapping $S^{-1}$, which measures if small neighbourhoods in $[0,1]^2$ are mapped to small neighbourhoods in $[0,1]$.
+We quantify this by introducing the ``locality'' $V(S^{-1})$ of the mapping $S^{-1}$, which measures if small neighbourhoods in $[0,1]$ are mapped to small neighbourhoods in $[0,1]^2$.
 Using locality, we then prove that the mean-field approximation is still valid, implying that we can simulate the one-dimensional neural field using classical grid-based methods.
+
+\ifndef{SUBMISSION}
+\includegraphics[height=15em]{../figures/locality.png}
+
+[TODO: clean up this graphic, cite it in text. we need to introduce a random mapping maybe ?]
+\endif
 
@@
 
-[TODO: add labels, show n=3, etc.]
+[TODO: add labels, show n=3, etc. Comment on the recursive and fractal structure of the Z-mapping, and the projection for the Column mapping.]
 \endif
@@
 \includegraphics[width=\linewidth]{../figures/S-viz.export.pdf}
-[TODO: adapt this figure better, but gives the idea]
+[TODO: adapt this figure better, but gives the idea. We see the $4^n$ bins in 2D, and $4^n$ bins in 1D, but twice as many bits]
 \endif
@@
 
-\subsection{Coarse-graining the images to numerically validate the mean-field approximation}
+\subsection{Coarse-graining the images to numerically validate the mean-field approximation} \label{sec:coarse-graining}
 
@@
 This is illustrated in Fig.
-[TODO] Specifically, for every coarse-grained bin $a=1,\cdots,2^n$, the set of consecutive positions $\mathcal{C}_a = \{\tfrac{a-1}{2^n} + \tfrac{l-1}{4^n} \;|\; l \in 1, \cdots, 2^n\}$, gets mapped to the two-dimensional grid points and averaged:
+[TODO] Specifically, for every coarse-grained bin $a=1,\cdots,2^n$, the set of consecutive positions $\mathcal{C}_a = \{\tfrac{a-1}{2^n} + \tfrac{l-1}{4^n} \;|\; l \in 1, \cdots, 2^n\}$, gets mapped to the two-dimensional grid points and averaged, intuitively computing the ``center of mass'' of $\mathcal{C}_a$:
 
@@
 
-\begin{equation}
+\begin{equation} \label{eq:def-tilde-J}
 \tilde J_{ab} := w_U(\vec{\bar v_a}, \vec{\bar v_b}).
@@
 
-chap. 2.1,
+By using the grid-based approach and the coarse-graining procedure, it is now possible to numerically simulate the two-dimensional neural field, and the mapping to the one-dimensional neural field.
+We introduce a toy model for illustration purposes, but we stress that our results apply to any neural field model.
 
-we stress that this toy model serves only for illustration purposes, and our results apply to any neural field
+We extend the model introduced in \cite{schmutz2023convergence}, which can be seen as a simple example of a ``Gaussian mixture low-rank network'' \cite{Beiran2021}.
+Given the $p$-dimensional embedding $\vec z = (z_1, \cdots, z_p) \in \Rp$, the neural field $h(\vec z, t)$ evolves as:
 
-introduce the low-rank cycling model
+\begin{equation*} \begin{aligned}
+\partial_t h(\vec z, t) &= -h(\vec z, t) + \int_{\Rp} w(\vec z, \vec y) \phi(h(\vec y, t)) \mathcal N^p(\vec y) \mathrm d \vec y, \\
+\text{with } w(\vec z, \vec y) &= \sum_{\mu=1}^p z_\mu \tilde\phi(y_\mu) \text{ the connectivity kernel}, \\
+\text{where } \tilde\phi(y) &= \frac{\phi(y) - \mathrm E_{y\sim\mathcal N}[\phi(y)]}{\mathrm{Var}_{y\sim\mathcal N}[\phi(y)]}, \quad \mathcal N(y) = \frac{1}{\sqrt{2\pi}} \mathrm{e}^{-\tfrac 12 y^2}, \\
+\text{and } \phi(y) &= \frac{1}{1+\mathrm{e}^{-y}}.
+\end{aligned} \end{equation*}
 
-goal of this section : "it seems that it works for the Z-mapping"
+The geometric view (see illustation in Fig.
+[TODO]) of this neural field is that the neuron populations are scattered in a $p$-dimensional Gaussian ball.
+With the exception of the zero fixed point, there are $p$ ``patterns'' (in this case, the basis functions $f_\mu(\vec z) = z_\mu,\,\mu=1,\cdots,p$) towards which the neural field $h(\vec z,t)$ may converge (we refer to \cite{nvadot2023mathesis}, appendix A for a derivation of the fixed points of this model, and a study of their stability).
+We stress that one should not confuse the embedding dimension $p$ with the rank of the connectivity kernel, which in this case also coincides with $p$.
+
+\ifndef{SUBMISSION}
+\includegraphics[height=15em]{../figures/lowrankgaussian.pdf}
+
+[TODO: does this need cleaning up (e.g. the Computer Modern font) ?]
+\endif
+
+This example is made particularly easy to study, because the kernel and dynamics both have the rank $p$.
+Firstly, the kernel's low-rank structure makes numerical simulations cheap, as we store $2p$ low-rank vectors of size $4^n$ instead of the full $4^n \times 4^n$ connectivity matrix (see the [TODO: source code url] for the full implementation details).
+Secondly, by projecting the neural field at any time $t$ onto the $p$ patterns, we can track the full state of the network (setting aside the orthogonal component, which is decoupled and decays exponentially) by following the trajectory in the latent space with the latent variables:
+% NOTE: in reality, the latent variables even form a closed system, but this is off-topic for this article
+
+\begin{equation} \label{eq:def-kappa}
+\kappa_\mu(t) = \int_{\Rp} y_\mu h(\vec y, t) \mathcal N^p(\vec y) \mathrm d \vec y,\quad \mu=1,\cdots,p.
+\end{equation}
+
+To illustrate the latent variables, suppose the neural field has converged to the pattern $h(\vec z, t \to \infty) = z_\mu$.
+Then the latent variable $\kappa_\mu = 1$, whereas all other latent variables are zero, $\kappa_\mu=0,\,\mu\neq\nu$.
+
+We extend this low-rank Gaussian model by making the dynamics cyclic: instead of converging to a single fixed point, we force the neural field to cycle between the fixed points.
+This modification makes any differences between numerically evaluated trajectories more apparent, which will be useful when asserting that the dynamics are conserved after the mapping to one dimension.
+To introduce a cycling behavior, the term $\delta > 0$ ``delays'' the integral term, and the ``rolling'' $z_\mu \mapsto z_{\mu+1}$ (with the convention that $p+1=1$) pushes the recurrent currents to the pattern $\mu+1$ when the neural field is close to the pattern $\mu$.
+The dynamics of the low-rank Gaussian cycling neural field in \autoref{eq:def-cycling-nf} are illustrated in Fig.
+[TODO]:
+
+\begin{equation} \label{eq:def-cycling-nf}
+\partial_t h(\vec z, t) = -h(\vec z, t) + \int_{\Rp} \sum_{\mu=1}^p z_{\mu+1} \tilde\phi(y_\mu) \phi(h(\vec y, t+\delta)) \mathcal N^p(\vec y) \mathrm d \vec y.
+\end{equation}
+
+\ifndef{SUBMISSION}
+\includegraphics[width=\linewidth]{../figures/cycling-illustration.png}
+
+[TODO: clean this up]
+\endif
+
+\subsection{Simulations of the Column and the Z-mappings applied to the toy model}
+
+In this section, we show numerical simulations of the neural field in $[0,1]$, obtained from applying mappings on higher-dimensional neural fields in $[0,1]^p$.
+We see that the coarse-graining procedure works when applied from $[0,1]^2$ to $[0,1]$.
+Then, we demonstrate that it can be applied iteratively, to map from $[0,1]^3$ to $[0,1]^2$, and from $[0,1]^2$ to $[0,1]$.
+
+We start with simulating the cycling neural field \autoref{eq:def-cycling-nf} using the grid-based approach.
+The connectivity matrix defined in \autoref{eq:def-tilde-J} is computed by applying different mappings to the matrix $J_{ij}$ (that is, permuting the rows and columns), followed by the coarse-graining procedure to reduce to a $2^n \times 2^n$ matrix.
+Setting $\delta=6$ and the initial condition $h(z_1, z_2, t < 0) = z_1$, simulations are presented in Fig.
+[TODO] showing the activity levels, recurrent currents, and the latent trajectory $(\kappa_1(t), \kappa_2(t))$.
+In order, we discuss:
+\begin{itemize}
+\item The original two-dimensional neural field evolves while oscillating between the patterns $\mu=1$ and $\mu=2$.
+The neural field is initialized in the $\mu=1$ state, and as a consequence $\kappa_1(0)=1$ and $\kappa_2(0)=0$.
+The oscillation can better be seen in the latent trajectory $(\kappa_1(t), \kappa_2(t))$, evolving somewhat between the points $(1,0)$ and $(0,1)$.
+\item A random mapping can be numerically evaluated as a random permutation of the rows and columns of $J_{ij}$, and is used as the ``worst-case'' example, since all regularity of the original neural field is obviously destroyed.
+The coarse-graining averages out all the structure of the activity, and the recurrent currents are zero, because the contributions from excitatory and inhibitory populations cancel out.
+This results in the erasure of the cycling behavior in the latent trajectory, and the latent trajectory stays at $(0,0)$.
+\item The Column mapping shows the projection of the initial condition onto the $z_1$ axis, yielding the sigmoidal shape in the $[0,1]$ embedding.
+Problematically, the current currents corresponding to the pattern $\mu=2$ cancel out and yield zero, and the latent trajectory quickly decays towards $(0,0)$.
+\item The Z-mapping shows that despite the fractal structure of the activity, there is still regularity in the $[0,1]$ embedding, resulting in non-zero recurrent currents.
+In consequence, the latent trajectory shows a clear cycling behavior, that closely matches that of the two-dimensional neural field.
+\end{itemize}
+
+\ifndef{SUBMISSION}
+\includegraphics[height=30em]{../figures/simulation-snapshots.png}
+
+[TODO: these are simulation snapshots. clean this up, adapt the legends, upload the videos]
+\endif
+
+These results suggest that the Z-mapping has the desired properties of conserving the regularity of the original smooth kernel.
+To further stress this, we demonstrate that we can repeat the dimensionality reduction of the embedding space, by iteratively applying the Z-mapping and coarse-graining.
+Starting with a three-dimensional neural field (setting $p=3$ in \autoref{eq:def-cycling-nf}), we can simulate the mapping from $[0,1]^3$ to $[0,1]^2$ (by applying, at constant $z_1$, the mapping $Z(z_2,z_3) = \alpha_{23}$), then from $[0,1]^2$ to $[0,1]$ (applying the mapping $Z(\alpha_{23},z_1)=\alpha_{123}$).
+This is shown in Fig.
+[TODO], and we assert equivalence of dynamics by studying the numerical equality of the latent trajectories.
+This demonstrates that for the mapping to yield identical dynamics, the starting kernel only has to be ``continuous on average''.
+
+\ifndef{SUBMISSION}
+\includegraphics[height=20em]{../figures/simulation-321-snapshots.png}
+
+[TODO: legend, link video]
+\endif
 
@@
 
-where do the populations end up
+In the previous sections we derived the Column and the Z-mapping, and showed that, at least numerically, the Z-mapping seems to be a ``local'' mapping, in the sense that small neighbourhoods in $[0,1]$ are mapped by $S^{-1}$ to small neighbourhoods in $[0,1]^2$.
+When writing the one-dimensional neural field in \autoref{eq:nf-in-01} with the kernel $\tilde w(\alpha,\beta) = w_U(S^{-1}(\alpha), S^{-1}(\beta))$, are neural populations close in the one-dimensional embedding also close in the two-dimensional embedding?
+
+% TODO: rework this paragraph, i'm tired
+% Locality is closely tied to the coarse-graining procedure introduced in \autoref{sec:coarse-graining}, because local mappings will yield populations that are close in $[0,1]^2$, and coarse-graining will not introduce much error; the mean-field approximation is in this case valid.
+Contrarily, when the mapping is not local, the coarse-graing will average together positions that are far apart, resulting in inaccurate samplings of the connectivity kernel.
+For the Column mapping, coarse-graining removes all variation on the vertical axis, reducing the sampling of the kernel $w_U$ to a line: the mapping is not local.
+The Z-mapping is local: even after coarse-graining, the sampling points span the full two-dimensional as $n \to \infty$.
+
+Locality is closely tied to the coarse-graining procedure introduced in \autoref{sec:coarse-graining}. 
+In Fig. [TODO], we colour-code the two-dimensional populations corresponding to the same $\mathcal{C}_a$ before coarse-graining.
+We see that the Z-mapping yields populations close in $[0,1]^2$, such that coarse-graining will not introduce much error and the kernel $w_U$ is sampled uniformly as $n \to \infty$: the mean-field approximation for the resulting one-dimensional neural field is valid.
+On the contrary, the coarse-grained populations of the Column mapping lie on a line in the two-dimensional embedding (it ``changes too fast''), resulting in non-uniform sampling of the kernel $w_U$: this mapping is not local and we cannot properly write the one-dimensional neural field.
+
+\ifndef{SUBMISSION}
+\includegraphics[height=20em]{../figures/where-are-the-coarse-grained-positions.png}
+
+[TODO: clean this up]
+\endif
+
+\subsection{Definition of locality}
+
+Given $n \in \mathbb N$, let us define the locality $V^n(S^n_\textrm{inv})$ as the average error made by the coarse-graining procedure.
+For this, we compute the maximum difference in positions of the two-dimensional embedding $\norm{S^n_\textrm{inv}(\alpha) - S^n_\textrm{inv}(\alpha^\prime)}_1$, associated to the one-dimensional populations in the bounds $\left[\tfrac{a-1}{2^n}, \tfrac{a}{2^n}\right]$ of the coarse-graining bin $\mathcal C_a$:
+
+\begin{equation} \label{eq:def-V}
+V^n(S^n_\textrm{inv}) = \frac{1}{2^n} \sum_{a=1}^{2^n} \sup_{\alpha, \alpha^\prime \in \left[\frac{a-1}{2^n}, \frac{a}{2^n} \right]} \norm{S^n_\textrm{inv}(\alpha) - S^n_\textrm{inv}(\alpha^\prime)}_1
+\end{equation}
+
+Within the context of a sequence of mappings, we are specifically interested in the $n \to \infty$ limit, that is, whether the error of the mean-field approximation vanishes as the grid becomes infinitely fine.
+For the random, Column and Z-mappings, we graph the scaling behavior of the numerator (the total error of the mean-field approximation) in Fig.
+[TODO].
+As expected, we see that the total error of the Z-mapping scales much slower than $2^n$, whereas that of the Column (and random) mapping scales roughly as $2^n$.
+
+\ifndef{SUBMISSION}
+\includegraphics[height=15em]{../figures/locality-scaling.png}
+
+[TODO: cleanup]
+\endif
+
+\subsection{Computing the locality of the Column and the Z-mapping}
+
+In this section we analytically derive the locality of the Column and the Z-mapping, showing that $V^n(Z^n_\textrm{inv}) \xrightarrow{n \to \infty} 0$, but that $V^n(C^n_\textrm{inv}) > 0$ for all $n \in \mathbb{N}$.
+
+We consider the binary expansion of $\alpha$ and $\alpha^\prime$ which appear in the supremum of \autoref{eq:def-V}.
+For clarity of notation, we furthermore truncate the binary expansion to $2m$ bits (on most machines, $2m=64$ bits, but analytically $m$ may be arbitrarily large):
+
+\begin{equation*} \begin{aligned}
+\alpha &= 0. b^1 b^2 \cdots b^{2m-1} b^{2m},\\
+\alpha^\prime &= 0. b^{\prime 1} b^{\prime 2} \cdots b^{\prime 2m-1} b^{\prime 2m}. \\
+\end{aligned} \end{equation*}
+
+Furthermore, since $\alpha \in \left[\frac{i-1}{2^n}, \frac{i}{2^n} \right],\, i \in \{1, \cdots, 2^n\}$, we may express the first $n$ (where $n < 2m$) bits of $\alpha$ with the bits of $i=\sum_{k=0}^{n-1} i^k 2^k$ (ignoring $i=2^n$ for the sake of simplicity, which would introduce one more bit, $n+1$ in total):
+
+\begin{equation*} \begin{aligned}
+\alpha &= 0. i^{n-1} i^{n-2} \cdots i^1 i^0 b^n b^{n+1} \cdots b^{2m-1} b^{2m},\\
+\alpha^\prime &= 0. i^{n-1} i^{n-2} \cdots i^1 i^0 b^{\prime n} b^{\prime n+1} \cdots b^{\prime 2m-1} b^{\prime 2m}.
+\end{aligned} \end{equation*}
+
+\paragraph{Z-mapping.} Let us consider the Z-mapping. If $n$ is even, we write with $\alpha$ defined above (and similarly for $\alpha^\prime$):
+
+$$
+\begin{aligned}
+Z^n_{\textrm{inv},1}(\alpha) &= 0.\underbrace{i^{n-1} i^{n-3} \cdots i_1}_\text{$n/2$ bits} b^n b^{n+2} \cdots b^{2m-1}, \\
+Z^n_{\textrm{inv},2}(\alpha) &= 0.\underbrace{i^{n-2} i^{n-4} \cdots i_0}_\text{$n/2$ bits} b^{n+1} b^{n+3} \cdots b^{2m}.
+\end{aligned}
+$$
+
+If $n$ is odd, we have:
+
+$$
+\begin{aligned}
+Z^n_{\textrm{inv},1}(\alpha) &= 0.\underbrace{i^{n-1} i^{n-3} \cdots i^0}_\text{$(n+1)/2$ bits} b^n b^{n+2} \cdots b^{2m-1}, \\
+Z^n_{\textrm{inv},2}(\alpha) &= 0.\underbrace{i^{n-2} i^{n-4} \cdots i^1}_\text{$(n-1)/2$ bits} b^{n+1} b^{n+3} \cdots b^{2m}.
+\end{aligned}
+$$
+
+We only consider even $n$ in the following, as the proof for the latter case remains very similar. The key is that the first $n/2$ bits of the difference of the inverses cancel out:
+
+$$
+\begin{aligned}
+|Z^n_{\textrm{inv},1}(\alpha) - Z^n_{\textrm{inv},1}(\alpha^\prime)| &= \left|0.0 \cdots 0 (b^n b^{n+2} \cdots b^{2m-1} - b^{\prime n} b^{\prime n+2} \cdots b^{\prime 2m-1})\right| \\
+&= \left|2^{-n/2} \sum_{k=1}^{\tfrac{2m-n+1}{2}} (b^{n+2k-1} - b^{\prime n+2k-1}) 2^{-k}\right| \\
+&\leq 2^{-n/2} \\
+|Z^n_{\textrm{inv},2}(\alpha) - Z^n_{\textrm{inv},2}(\alpha^\prime)| &= \left|2^{-n/2} \sum_{k=1}^{\tfrac{2m-n+1}{2}} (b^{n+2k} - b^{\prime n+2k}) 2^{-k}\right| \\
+&\leq 2^{-n/2}
+\end{aligned}
+$$
+
+Putting everything together, we get the following vanishing upper bound:
+
+$$
+V^n(Z^n_\textrm{inv}) \leq \frac{1}{2^n} \sum_{i=1}^{2^n} 2^{-n/2} + 2^{-n/2} = 2^{-n/2+1} \xrightarrow{n \to \infty} 0
+$$
+
+\paragraph{Column mapping.} We now consider the Column mapping:
+
+$$
+\begin{aligned}
+C^n_{\textrm{inv},1}(\alpha) &= 0.i^{n-1} i^{n-2} \cdots i^1 i^0 b^n b^{n+1} \cdots b^m \\
+C^n_{\textrm{inv},2}(\alpha) &= 0.b^{m+1} b^{m+2} \cdots b^{2m}.
+\end{aligned}
+$$
+
+The first component can be bounded using the same trick as for the Z-mapping above:
+
+$$
+\begin{aligned}
+|C^n_{\textrm{inv},1}(\alpha) - C^n_{\textrm{inv},1}(\alpha^\prime)| &= \left| 2^{-n} \sum_{k=1}^{m-n+1} (b^{n+k-1} - b^{\prime n+k-1}) 2^{-k} \right| \\
+&\leq 2^{-n}.
+\end{aligned}
+$$
+
+However, problems arise due to the second component not being well-behaved.
+In particular, since we take the supremum on $\alpha,\alpha^\prime \in \left[\tfrac{i-1}{2^n},\tfrac{i}{2^n}\right]^2$, we can pick the specific values such that $b^k=1$ and $b^{\prime k}=0$ for all $k \in \{m+1, \cdots 2m\}$:
+
+$$
+\begin{aligned}
+|C^n_{\textrm{inv},2}(\alpha) - C^n_{\textrm{inv},2}(\alpha^\prime)| &= \left| \sum_{k=m+1}^{2m} (b^k - b^{\prime k}) 2^{-(k-m)} \right| \\
+&= \sum_{k=m+1}^{2m} 2^{-(k-m)} \\
+&> \frac 12.
+\end{aligned}
+$$
+
+Putting both together, we can write:
+
+$$
+\begin{aligned}
+V^n(C^n_\textrm{inv}) &= \frac{1}{2^n}\sum_{i=1}^{2^n} \sup_{\alpha, \alpha^\prime \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]^2} |C^n_{\textrm{inv},1}(\alpha^\prime) - C^n_{\textrm{inv},1}(\alpha)| + |C^n_{\textrm{inv},2}(\alpha^\prime) - C^n_{\textrm{inv},2}(\alpha)| \\
+&\geq
+\begin{aligned}[t]
+	\frac{1}{2^n}\sum_{i=1}^{2^n}&
+	\sup_{\alpha, \alpha^\prime \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]^2} |C^n_{\textrm{inv},1}(\alpha^\prime) - C^n_{\textrm{inv},1}(\alpha)|\\
+	&+ \sup_{\alpha, \alpha^\prime \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]^2} |C^n_{\textrm{inv},2}(\alpha^\prime) - C^n_{\textrm{inv},2}(\alpha)|
+\end{aligned}\\
+&\geq \frac 12 + \bO(2^{-n}).
+\end{aligned}
+$$
+
+We following upper bound therefore shows that the locality measure cannot converge to zero as $n \to \infty$:
+
+$$
+V^n(C^n_\textrm{inv}) > \frac 12.
+$$
 
@@
 
+\subsection{Locality garantees convergence of grid-based simulations as $n \to \infty$}.
+
 \subsection{Analytical equivalence given $S$ measurable and bijective}
 \label{sec:proof-equivalence}
-
-\subsection{Numerical convergence}
 
@@
 % in this case, we have shown that bijection + locality => we can safely do the mapping (bijection for the analytical approach, locality for the numerical approach)
-% TODO: mention the "standard example" of the Z-mapping
+% TODO: mention the "standard example" of the Z-mapping.
+This framework is surely not unique, for instance the reclocal bijection cannot easily be expressed as the permutation of bits, but it still works in the same way as the Z-mapping (fractal formulation), and numerical simulations are OK
 

### 2024-02-28 paper/manuscript.tex
review manuscript up to section 3

@@
 Valentin Schmutz\textsuperscript{1},
-Wulfram Gerstner\textsubscript{1}
+Wulfram Gerstner\textsuperscript{1}
 % Name1 Surname\textsuperscript{1,2\Yinyang},
@@
 
-As models of spatiotemporal neural population dynamics on the cortical sheet, neural field models can be used to explain experimentally observed patterns of cortical activity, such as travelling waves in visual \cite{SatNau12,MulRey14}, somatosensory \cite{PetHah03,FerBol06}, motor \cite{RubRob06,TakKim15}, and hippocampal \cite{LubSia09,PatFuj12,PatSch13} cortices (see \cite{MulCha18} for a review), and are used to model large-scale brain signals, such as electroencephalography (EEG) recordings \cite{Bre17} (see also \cite{DipRan18} for an example of neural field modelling of calcium imaging recording of the visual cortex).
+As models of spatiotemporal neural population dynamics on the cortical sheet, neural field models can be used to explain experimentally observed patterns of cortical activity, such as travelling waves in visual \cite{SatNau12,MulRey14}, somatosensory \cite{PetHah03,FerBol06}, motor \cite{RubRob06,TakKim15}, and hippocampal \cite{LubSia09,PatFuj12,PatSch13} cortices (see \cite{MulCha18} for a review), and are used to model large-scale brain signals, such as electroencephalography (EEG) recordings \cite{Bre17} (see also \cite{DipRan18} for an example of neural field modeling of calcium imaging recordings of the visual cortex).
 
@@
 
-While, historically, the spatial dimensions in neural field models have been the two dimensions of the cortical sheet (with the notable exception of the ring models for orientation selectivity in the visual cortex \cite{BenBar95} and the head-direction system \cite{Zha96}), the ``space'' in neural field models does not need to represent physical space but can represent any suitable abstract embedding space.
+While, historically, the spatial dimensions in neural field models have been the two dimensions of the cortical sheet (with the notable exception of the ring models for orientation selectivity in the visual cortex \cite{BenBar95} and the head-direction system \cite{Zha96}), the ``embedding space'' in neural field models does not need to represent physical space but can represent any suitable abstract space.
 Neural field models with abstract embedding spaces could constitute a generalization of classical neural field theory for cortical networks that are not solely structured by distances of neurons on the cortical sheet.
@@
 
+We do not address the question of well-posedness of the neural field equations in this paper; for proofs of the well-posedness of neural field equations, we refer the reader to \cite{faugeras2008absolute,faugeras2009persistent,veltz2009localglobal}.
+
 Without loss of generality, a simple bijective mapping can be applied to reduce any $p$-dimensional neural field to a neural field on $[0,1]^p$.
@@
 \begin{equation} \label{eq:nf-in-01p}
-\partial_t h_U(t, \vec v) = -h_U(t, \vec v) + \int_{[0,1]^p} w_U(\vec v, \vec u) \phi(h_U(t, \vec u)) \mathrm d \vec u.
+\partial_t h_U(\vec v, t) = -h_U(\vec v, t) + \int_{[0,1]^p} w_U(\vec v, \vec u) \phi(h_U(\vec u, t)) \mathrm d \vec u.
 \end{equation}
 
-This form highlights the central role of the connectivity kernel $w_U$, which together with the initial condition $h_U(t=0, \vec v)$, uniquely determines the neural dynamics.
+This form highlights the central role of the connectivity kernel $w_U$, which together with the initial condition $h_U(\vec v, t=0)$, uniquely determines the neural dynamics.
 
@@
 The mapping from the two-dimensional equation to the one-dimensional equation is done via well-known measurable bijections between the square $[0,1]^2$ and the interval $[0,1]$.
-Importantly, these bijections are not diffeomorphisms.
+Importantly, these bijections cannot be diffeomorphisms.
 % Also, we show that although the connectivity kernel of the one-dimensional equation seems to have a fractal structure, it is sufficiently regular for its solution to be numerically approximated using standard grid-based simulations.
-The main idea is to define a measurable bijective mapping $S : [0,1]^2 \mapsto [0,1]$ and a measure $\lambda : [0,1]^2 \mapsto \mathbb{R}^+$, such that we can write a one-dimensional neural field	$\tilde h(\alpha, t)$ with a connectivity kernel defined as $\tilde w(\alpha, \beta) = w_U(S^{-1}(\alpha), S^{-1}(\beta))$.
+
+The main idea is to define a measurable bijective mapping $S : [0,1]^2 \mapsto [0,1]$ and the Lebesgue measure $\lambda : [0,1]^2 \mapsto \mathbb{R}^+$, such that we can write a one-dimensional neural field	$\tilde h(\alpha, t)$ with a connectivity kernel defined as $\tilde w(\alpha, \beta) = w_U(S^{-1}(\alpha), S^{-1}(\beta))$.
 Doing this, we arrive at the following one-dimensional neural field equation:
@@
 
-The difficulty in writing the one-dimensional neural field, is that the spatial structure of the original neural field in $[0,1]^2$ may be lost.
+The difficulty in writing the one-dimensional neural field is that the spatial structure of the original neural field in $[0,1]^2$ may be lost.
 Fig.
-[TODO] illustrates this problem: if the mapping is seemingly random, the regularity of the orginal neural field is destroyed.
+[TODO] illustrates this problem: if the mapping is random, the regularity of the orginal neural field is destroyed.
 Contrarily, if the mapping is ``local'', the resulting neural field might still be well-defined and exhibit identical dynamics, despite the apparent discontinuities.
+
 The intuition is that we would like populations close in $[0,1]^2$ to be bijectively mapped to populations close in $[0,1]$.
@@
 We quantify this by introducing the ``locality'' $V(S^{-1})$ of the mapping $S^{-1}$, which measures if small neighbourhoods in $[0,1]$ are mapped to small neighbourhoods in $[0,1]^2$.
-Using locality, we then prove that the mean-field approximation is still valid, implying that we can simulate the one-dimensional neural field using classical grid-based methods.
+Using locality, we then prove that the mean-field approximation applied to the one-dimensinoal neural field is still valid, implying that it can be simulated using classical grid-based methods.
 
@@
 \endif
-
-% question of uniqueness, discretization error, mean field limit and locality, link with numerical simualtions. chap 1.4, chap 3.1
-
-% "discretization error"
-
-% chap 1.1, chap 2.1, 2.5
 
@@
 
-\autoref{eq:population-dynamics} can be seen as the discretized version of \autoref{eq:nf-in-01p}, where we associate $h_i(t) := h_U(\vec{v_i}, t)$.
-Indeed, by a mean-field argument on the original smooth neural field, the discrete dynamics converge to the true solution of the neural field, because the error of the mean-field approximation in $[0,1]^2$ vanishes as we take finer grids, letting $n \to \infty$.
-
-We would now like to derive a bijective mapping $S$, such that the same argument applies on the neural field in $[0,1]$.
+\autoref{eq:population-dynamics} can be seen as the discretized version of \autoref{eq:nf-in-01p}, where we associate the samples $h_i(t) = h_U(\vec{v_i}, t)$.
+Indeed, by a mean-field argument on the original smooth neural field, the discrete population dynamics converge to the true solution of the neural field, because the error of the mean-field approximation in $[0,1]^2$ vanishes as we take finer grids, letting $n \to \infty$.
+
+We would now like to derive a bijective mapping $S$, such that the same argument applies to the neural field in $[0,1]$.
 Intuively, we want $S$ to conserve the regularity of the neural field on $[0,1]^2$, since if the mean-field argument also carries over to the one-dimensional neural field, it seems natural that it will express identical dynamics.
@@
 
-Let us define a sequence of bijective mappings $\{S^n\}_{n \in \mathbb N}$, acting on the $n$-bit truncations $\vec{v^{(n)}}$ of any $\vec{v} \in [0,1]^2$, as well as the corresponding sequence of images $\{\alpha^n := S^n(\vec{v^{(n)}})\}_{n \in \mathbb N}$.
+Let us define a sequence of bijective mappings $\{S^n\}_{n \in \mathbb N}$ acting on the $n$-bit truncations $\vec{v^{(n)}}$ of any $\vec{v} \in [0,1]^2$, as well as the corresponding sequence of images $\{\alpha^n := S^n(\vec{v^{(n)}})\}_{n \in \mathbb N}$.
 $S^n$ defines the following bijection between points on the two-dimensional and one-dimensional grids:
@@
 
-We write $\alpha^n$ using the $2n$ bits $\{b^1, b^2, \cdots, b^{2n}\}$, which is intuitively the lowest number of bits required so that $S^n$ remains a bijection, since it takes $2n$ bits of information as argument.
+We write $\alpha^n$ using the $2n$ bits $\{b^1, b^2, \cdots, b^{2n}\}$, which is intuitively the lowest number of bits required so that $S^n$ remains a bijection, since the mapping takes $2n$ bits of information as argument.
 More than $2n$ bits could be allowed while maintaining bijectivity, but we omit this possibility, since they wouldn't strictly be required.
@@
 
-Given $(v_1, v_2) \in [0,1]^2$ and the sequence $\{S^n(v_1, v_2)\}_{n \in \mathbb N}$, we then we define $\alpha = S(v_1, v_2)$ as its limit, if it exists.
+Given $(v_1, v_2) \in [0,1]^2$ and the sequence $\{S^n(v_1, v_2)\}_{n \in \mathbb N}$, we then define $\alpha = S(v_1, v_2)$ as its limit, if it exists.
 Informally, pointwise convergence of $\{S^n\}_{n\in\mathbb N}$ requires that the least signficiant bits of the input be mapped to the least significant bits of the image, ensuring that small corrections in the input $\vec{v}$ result in small corrections in the image $\alpha$.
@@
 
-The permutation between the bits of $\alpha^n$, and the bits of $(v_1^{(n)}, v_2^{(n)})$ is illustrated in Fig.
+The permutation between the bits of $\alpha^n_A$, and the bits of $(v_1^{(n)}, v_2^{(n)})$ is illustrated in Fig.
 [TODO: arrow connecting the bits together], and shows how this example does not respect the bit ordering described previously.
@@
 
-Let us provide two examples of sequence of mappings that are pointwise convergent, and act as permutation of the grid point bits.
+Let us provide two examples of sequences of mappings that are pointwise convergent, and act as permutations of the grid point bits.
 We first define the sequence of Column mappings $\{C^n\}_{n\in\mathbb N}$ as the concatenation of the input bits of $v_1^{(n)}$, followed by the bits of $v_2^{(n)}$.
@@
 \end{aligned} \end{equation}
+
+We finish by noting that by definition of the sequence of Z-mappings and the corresponding inverses, $2n$-bit truncations of $Z^{-1}(\alpha)$ (where $\alpha \in [0,1]$) give identical results to $Z^n_\textrm{inv}(\alpha^n)$ applied to the truncated $\alpha^n$.
+Where context allows it, we will therefore drop $n$ and directly write the limit $Z^{-1}$.
 
@@
 
-Given $n \in \mathbb N$, let us define the locality $V^n(S^n_\textrm{inv})$ as the average error made by the coarse-graining procedure.
-For this, we compute the maximum difference in positions of the two-dimensional embedding $\norm{S^n_\textrm{inv}(\alpha) - S^n_\textrm{inv}(\alpha^\prime)}_1$, associated to the one-dimensional populations in the bounds $\left[\tfrac{a-1}{2^n}, \tfrac{a}{2^n}\right]$ of the coarse-graining bin $\mathcal C_a$:
+Given $n \in \mathbb N$, let us define the locality $V_n(S^{-1})$ as the average error made by the coarse-graining procedure.
+For this, we compute the maximum difference in positions of the two-dimensional embedding $\norm{S^{-1}(\alpha) - S^{-1}(\alpha^\prime)}_1$, associated to the one-dimensional populations in the bounds $\left[\tfrac{a-1}{2^n}, \tfrac{a}{2^n}\right]$ of the coarse-graining bin $\mathcal C_a$:
 
 \begin{equation} \label{eq:def-V}
-V^n(S^n_\textrm{inv}) = \frac{1}{2^n} \sum_{a=1}^{2^n} \sup_{\alpha, \alpha^\prime \in \left[\frac{a-1}{2^n}, \frac{a}{2^n} \right]} \norm{S^n_\textrm{inv}(\alpha) - S^n_\textrm{inv}(\alpha^\prime)}_1
+V_n(S^{-1}) = \frac{1}{2^n} \sum_{a=1}^{2^n} \sup_{\alpha, \alpha^\prime \in \left[\frac{a-1}{2^n}, \frac{a}{2^n} \right]} \norm{S^{-1}(\alpha) - S^{-1}(\alpha^\prime)}_1
 \end{equation}
@@
 Within the context of a sequence of mappings, we are specifically interested in the $n \to \infty$ limit, that is, whether the error of the mean-field approximation vanishes as the grid becomes infinitely fine.
+We then say a mapping is local if and only if $V_n(S^{-1}) \to 0$ as $n \to \infty$.
 For the random, Column and Z-mappings, we graph the scaling behavior of the numerator (the total error of the mean-field approximation) in Fig.
@@
 
+% TODO: continuity and continuity on average
+
 \ifndef{SUBMISSION}
@@
 \subsection{Computing the locality of the Column and the Z-mapping}
-
-In this section we analytically derive the locality of the Column and the Z-mapping, showing that $V^n(Z^n_\textrm{inv}) \xrightarrow{n \to \infty} 0$, but that $V^n(C^n_\textrm{inv}) > 0$ for all $n \in \mathbb{N}$.
+[TODO: should this section be moved to an appendix ?]
+
+In this section we analytically derive the locality of the Column and the Z-mapping, showing that $V_n(Z^{-1}) \xrightarrow{n \to \infty} 0$, but that $V_n(C^{-1}) > 0$ for all $n \in \mathbb{N}$.
 
@@
 \begin{aligned}
-Z^n_{\textrm{inv},1}(\alpha) &= 0.\underbrace{i^{n-1} i^{n-3} \cdots i_1}_\text{$n/2$ bits} b^n b^{n+2} \cdots b^{2m-1}, \\
-Z^n_{\textrm{inv},2}(\alpha) &= 0.\underbrace{i^{n-2} i^{n-4} \cdots i_0}_\text{$n/2$ bits} b^{n+1} b^{n+3} \cdots b^{2m}.
+Z^{-1}_1(\alpha) &= 0.\underbrace{i^{n-1} i^{n-3} \cdots i_1}_\text{$n/2$ bits} b^n b^{n+2} \cdots b^{2m-1}, \\
+Z^{-1}_2(\alpha) &= 0.\underbrace{i^{n-2} i^{n-4} \cdots i_0}_\text{$n/2$ bits} b^{n+1} b^{n+3} \cdots b^{2m}.
 \end{aligned}
@@
 \begin{aligned}
-Z^n_{\textrm{inv},1}(\alpha) &= 0.\underbrace{i^{n-1} i^{n-3} \cdots i^0}_\text{$(n+1)/2$ bits} b^n b^{n+2} \cdots b^{2m-1}, \\
-Z^n_{\textrm{inv},2}(\alpha) &= 0.\underbrace{i^{n-2} i^{n-4} \cdots i^1}_\text{$(n-1)/2$ bits} b^{n+1} b^{n+3} \cdots b^{2m}.
+Z^{-1}_1(\alpha) &= 0.\underbrace{i^{n-1} i^{n-3} \cdots i^0}_\text{$(n+1)/2$ bits} b^n b^{n+2} \cdots b^{2m-1}, \\
+Z^{-1}_2(\alpha) &= 0.\underbrace{i^{n-2} i^{n-4} \cdots i^1}_\text{$(n-1)/2$ bits} b^{n+1} b^{n+3} \cdots b^{2m}.
 \end{aligned}
@@
 \begin{aligned}
-|Z^n_{\textrm{inv},1}(\alpha) - Z^n_{\textrm{inv},1}(\alpha^\prime)| &= \left|0.0 \cdots 0 (b^n b^{n+2} \cdots b^{2m-1} - b^{\prime n} b^{\prime n+2} \cdots b^{\prime 2m-1})\right| \\
+|Z^{-1}_1(\alpha) - Z^{-1}_1(\alpha^\prime)| &= \left|0.0 \cdots 0 (b^n b^{n+2} \cdots b^{2m-1} - b^{\prime n} b^{\prime n+2} \cdots b^{\prime 2m-1})\right| \\
 &= \left|2^{-n/2} \sum_{k=1}^{\tfrac{2m-n+1}{2}} (b^{n+2k-1} - b^{\prime n+2k-1}) 2^{-k}\right| \\
 &\leq 2^{-n/2} \\
-|Z^n_{\textrm{inv},2}(\alpha) - Z^n_{\textrm{inv},2}(\alpha^\prime)| &= \left|2^{-n/2} \sum_{k=1}^{\tfrac{2m-n+1}{2}} (b^{n+2k} - b^{\prime n+2k}) 2^{-k}\right| \\
+|Z^{-1}_2(\alpha) - Z^{-1}_2(\alpha^\prime)| &= \left|2^{-n/2} \sum_{k=1}^{\tfrac{2m-n+1}{2}} (b^{n+2k} - b^{\prime n+2k}) 2^{-k}\right| \\
 &\leq 2^{-n/2}
@@
 $$
-V^n(Z^n_\textrm{inv}) \leq \frac{1}{2^n} \sum_{i=1}^{2^n} 2^{-n/2} + 2^{-n/2} = 2^{-n/2+1} \xrightarrow{n \to \infty} 0
-$$
-
-\paragraph{Column mapping.} We now consider the Column mapping:
-
-$$
-\begin{aligned}
-C^n_{\textrm{inv},1}(\alpha) &= 0.i^{n-1} i^{n-2} \cdots i^1 i^0 b^n b^{n+1} \cdots b^m \\
-C^n_{\textrm{inv},2}(\alpha) &= 0.b^{m+1} b^{m+2} \cdots b^{2m}.
+V_n(Z^{-1}) \leq \frac{1}{2^n} \sum_{i=1}^{2^n} 2^{-n/2} + 2^{-n/2} = 2^{-n/2+1} \xrightarrow{n \to \infty} 0
+$$
+
+\paragraph{Column mapping.} We now consider the Column mapping.
+Although the inverse of the Column mapping is not well-defined as $n \to \infty$, we can still study how the it behaves on finite truncations of the input:
+
+$$
+\begin{aligned}
+C^{-1}_1(\alpha) &= 0.i^{n-1} i^{n-2} \cdots i^1 i^0 b^n b^{n+1} \cdots b^m, \\
+C^{-1}_2(\alpha) &= 0.b^{m+1} b^{m+2} \cdots b^{2m}.
 \end{aligned}
@@
 \begin{aligned}
-|C^n_{\textrm{inv},1}(\alpha) - C^n_{\textrm{inv},1}(\alpha^\prime)| &= \left| 2^{-n} \sum_{k=1}^{m-n+1} (b^{n+k-1} - b^{\prime n+k-1}) 2^{-k} \right| \\
+|C^{-1}_1(\alpha) - C^{-1}_1(\alpha^\prime)| &= \left| 2^{-n} \sum_{k=1}^{m-n+1} (b^{n+k-1} - b^{\prime n+k-1}) 2^{-k} \right| \\
 &\leq 2^{-n}.
@@
 \begin{aligned}
-|C^n_{\textrm{inv},2}(\alpha) - C^n_{\textrm{inv},2}(\alpha^\prime)| &= \left| \sum_{k=m+1}^{2m} (b^k - b^{\prime k}) 2^{-(k-m)} \right| \\
+|C^{-1}_2(\alpha) - C^{-1}_2(\alpha^\prime)| &= \left| \sum_{k=m+1}^{2m} (b^k - b^{\prime k}) 2^{-(k-m)} \right| \\
 &= \sum_{k=m+1}^{2m} 2^{-(k-m)} \\
@@
 
-Putting both together, we can write:
-
-$$
-\begin{aligned}
-V^n(C^n_\textrm{inv}) &= \frac{1}{2^n}\sum_{i=1}^{2^n} \sup_{\alpha, \alpha^\prime \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]^2} |C^n_{\textrm{inv},1}(\alpha^\prime) - C^n_{\textrm{inv},1}(\alpha)| + |C^n_{\textrm{inv},2}(\alpha^\prime) - C^n_{\textrm{inv},2}(\alpha)| \\
+Putting both together, we find the following upper bound, showing that the locality measure cannot converge to zero as $n \to \infty$:
+
+$$
+\begin{aligned}
+V_n(C^{-1}) &= \frac{1}{2^n}\sum_{i=1}^{2^n} \sup_{\alpha, \alpha^\prime \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]^2} |C^{-1}_1(\alpha^\prime) - C^{-1}_1(\alpha)| + |C^{-1}_2(\alpha^\prime) - C^{-1}_2(\alpha)| \\
 &\geq
@@
 	\frac{1}{2^n}\sum_{i=1}^{2^n}&
-	\sup_{\alpha, \alpha^\prime \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]^2} |C^n_{\textrm{inv},1}(\alpha^\prime) - C^n_{\textrm{inv},1}(\alpha)|\\
-	&+ \sup_{\alpha, \alpha^\prime \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]^2} |C^n_{\textrm{inv},2}(\alpha^\prime) - C^n_{\textrm{inv},2}(\alpha)|
+	\sup_{\alpha, \alpha^\prime \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]^2} |C^{-1}_1(\alpha^\prime) - C^{-1}_1(\alpha)|\\
+	&+ \sup_{\alpha, \alpha^\prime \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]^2} |C^{-1}_2(\alpha^\prime) - C^{-1}_2(\alpha)|
 \end{aligned}\\
@@
 
-We following upper bound therefore shows that the locality measure cannot converge to zero as $n \to \infty$:
-
-$$
-V^n(C^n_\textrm{inv}) > \frac 12.
-$$
-
 \section{Proof of results}
 
-\subsection{Locality garantees convergence of grid-based simulations as $n \to \infty$}.
+\subsection{Locality garantees convergence of grid-based simulations as $n \to \infty$}
+
+In the previous sections we have made use of numerical simulations and coarse-graining, without analytic justification.
+Setting $S$ to the Z-mapping (or any other mapping with the locality property), does the numerical estimation of the integral in the neural field equation \autoref{eq:nf-in-01} converge to the true analytical value, as $n \to \infty$?
+
+The intuition behind the proof of convergence is that the variation coming from the coarse-graining of each bin $\mathcal C_a$ vanishes as we increase $n$, even though the neural field $\tilde h(\alpha, t)$ and the connectivity kernel $\tilde w(\alpha, \beta)$ become fractal.
+
+We define the one-dimensional field, sampled on the one-dimensional grid of size $2^n$:
+
+$$
+\tilde h_i(t) = \tilde h(\tfrac{i-1}{2^n}, t) = \tilde h(\beta_i, t),\,\beta_i=\tfrac{i-1}{2^n},\,i\in\{1,\cdots,2^n\}.
+$$
+
+For any $\alpha \in [0,1]$, let us define the numerical and analytical integral:
+$$
+\begin{aligned}
+\mathrm{NI}_n &= \frac{1}{2^n} \sum_{i=1}^{2^n} \tilde w(\alpha, \beta_i) \phi(\tilde h_i(t)), \\
+\mathrm{AI} &= \int_{[0,1]} \tilde w(\alpha, \beta) \phi(\tilde h(\beta, t)) \left[\lambda \circ S^{-1}\right](\d \beta).
+\end{aligned}
+$$
+
+We wish to prove that $|\mathrm{NI}_n - \mathrm{AI}| \xrightarrow{n \to \infty} 0$.
+
+The first part of the proof involves splitting the integral on $[0,1]$ into $2^n$ integrals over segments of length $2^{-n}$, and using the triangle inequality:
+
+$$
+\begin{aligned}
+|\mathrm{NI}_n - \mathrm{AI}| &= \left| \frac{1}{2^n} \sum_{i=1}^{2^n} \tilde w(\alpha, \beta_i) \phi(\tilde h_i(t)) - \int_{0}^{1} \tilde w(\alpha, \beta) \phi(\tilde h(\beta, t)) \left[\lambda \circ S^{-1}\right](\d \beta) \right| \\
+&= \left| \frac{1}{2^n} \sum_{i=1}^{2^n} \tilde w(\alpha, \beta_i) \phi(\tilde h_i(t)) - \sum_{i=1}^{2^n} \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \tilde w(\alpha, \beta) \phi(\tilde h(\beta, t)) \left[\lambda \circ S^{-1}\right](\d \beta) \right| \\
+&= \left| \sum_{i=1}^{2^n} \frac{1}{2^n} \tilde w(\alpha, \beta_i) \phi(\tilde h_i(t)) - \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \tilde w(\alpha, \beta) \phi(\tilde h(\beta, t)) \left[\lambda \circ S^{-1}\right](\d \beta) \right| \\
+&\leq \sum_{i=1}^{2^n} \left| \frac{1}{2^n} \tilde w(\alpha, \beta_i) \phi(\tilde h_i(t)) - \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \tilde w(\alpha, \beta) \phi(\tilde h(\beta, t)) \left[\lambda \circ S^{-1}\right](\d \beta) \right|.
+\end{aligned}
+$$
+
+We now want to prove that the integrand evaluated at each $\beta_i = \tfrac{i-1}{2^n}$ gives a good approximation of the integral.
+Because the Z-mapping is measure-preserving (see \autoref{sec:proof-equivalence}) and $\lambda$ is the Lebesgue measure on $[0,1]^2$, $\lambda \circ S^{-1}$ is a Lebesgue measure on $[0,1]$, and therefore $\int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \left[\lambda \circ S^{-1}\right](\d \beta) = \tfrac{1}{2^n}$.
+This allows us to write:
+
+$$
+\begin{aligned}
+|\mathrm{NI}_n - \mathrm{AI}| &\leq \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \Delta(\beta) \left[\lambda \circ S^{-1}\right](\d \beta) \right|, \\
+\text{where}\ \Delta(\beta) &= \tilde w(\alpha, \beta_i) \phi(\tilde h_i(t)) - \tilde w(\alpha, \beta) \phi(\tilde h(\beta, t))\\
+&= w_U(S^{-1}(\alpha), S^{-1}(\beta_i)) \phi(h_U(S^{-1}(\beta_i), t)) \\ &\qquad- w_U(S^{-1}(\alpha), S^{-1}(\beta)) \phi(h_U(S^{-1}(\beta), t)).
+\end{aligned}
+$$
+
+The main source of variance inside the integrand $\Delta(\beta)$ is the term $S^{-1}$, since $w_U$ and $h_U$ are ``nice'' functions.
+We express this regularity by the added assumption that $w_U(S^{-1}(\alpha), \vec u) \phi(h_U(t, \vec u))$ is Lipschitz in $\vec u$, with Lipschitz constant $L(\alpha)$.
+We argue this is not a strong assumption, because this is just expressing the regularity of the integrand in the original neural field equation.
+
+% A more pragmatic justification of this is to notice that the ``density of recurrent currents'' is expressed as:
+
+% $$
+% I^\text{rec}_U(\vec v, \vec u) = w_U(\vec v, \vec u) \phi(h_U(t, \vec u)),
+% $$
+
+% and that in @sec-cycling-nf, numerical simulations showed that $I^\text{rec}_U(\vec v, \vec u)$ (the integrand of the integral on $[0,1]^2$) is a well-behaved function.
+
+Applying the Lipschitz assumption yields the following inequality on the integrand:
+
+$$
+\begin{aligned}
+\Delta(\beta) &= w_U(\alpha, S^{-1}(\beta_i)) \phi(h_U(S^{-1}(\beta_i), t)) - w_U(\alpha, S^{-1}(\beta)) \phi(h_U(S^{-1}(\beta), t)) \\
+&\leq L(\alpha) \norm{S^{-1}(\beta_i) - S^{-1}(\beta)}.
+\end{aligned}
+$$
+
+We now make use of our intuition of locality: points close in $[0,1]$ should be (on average) mapped to points close in $[0,1]^2$.
+Therefore, taking the sup inside each $2^{-n}$ segment does not introduce a too large error:
+
+$$
+\begin{aligned}
+|\mathrm{NI}_n - \mathrm{AI}| &\leq L(\alpha) \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \left[\lambda \circ S^{-1}\right](\d \beta) \right| \\
+&\leq L(\alpha) \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \sup_{\beta \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \left[\lambda \circ S^{-1}\right](\d \beta^\prime) \right| \\
+&= L(\alpha) \frac{1}{2^n} \sum_{i=1}^{2^n} \sup_{\beta \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \\
+&\leq L(\alpha) \frac{1}{2^n} \sum_{i=1}^{2^n} \sup_{\beta, \beta^\prime \in \left[\tfrac{i-1}{2^n}, \tfrac{i}{2^n}\right]^2} \norm{S^{-1}(\beta^\prime) - S^{-1}(\beta)} \\
+&= L(\alpha) V_n(S^{-1}) \\
+&\xrightarrow{n \to \infty} 0.
+\end{aligned}
+$$
+
+We see that it is the property of locality associated to the mapping that allowed us to write the convergence of the numerical integral to the analytical value.
+In doing so, we exploited the fact that the variance in the image of each small segment in $[0,1]$ is small on average, and that the errors vanish as we take finer and finer grids.
 
@@
 \label{sec:proof-equivalence}
+[TODO: should this section be moved to an appendix ?]
+
+In this section we prove that the analytical formulation of the one-dimensional neural field holds in the case that the mapping $S$ is bijective and measurable.
+In the our case of sequences of mappings acting on grid points, the finite-$n$ expansions $S^n$ are sums of piecewise constant functions, and therefore measureable.
+Since the limit of pointwise convergent measurable functions is also measurable, this implies that $S$ is measurable.
+The measurability of $S$ allows us to write $\lambda \circ S^{-1}$, the pushforward measure (or image measure) of $\lambda$ under the mapping $S$.
+
+With these two assumptions, let us prove that by defining the initial condition and connectivity kernel
+$$
+\begin{aligned}
+\tilde h(\alpha, t=0) &= h_U(S^{-1}(\alpha), t=0)\\
+\tilde w(\alpha, \beta) &= w_U(S^{-1}(\alpha), S^{-1}(\beta)),
+\end{aligned}
+$$
+
+then the solution $h_U(\vec v, t)$ of the neural field \autoref{eq:nf-in-01p} in $[0,1]^2$ uniquely defines the solution of the neural field equation in $[0,1]$ by taking
+
+\begin{equation} \label{eq:tilde-h-subst}
+\tilde h(\alpha, t) = h_U(S^{-1}(\alpha), t).
+\end{equation}
+
+Substituting \autoref{eq:tilde-h-subst}, the definitions of $\tilde h(\alpha, t)$ and $\tilde w(\alpha, \beta)$ into \autoref{eq:nf-in-01}, and denoting $\vec{v} = S^{-1}(\alpha)$, $\vec{u} = S^{-1}(\beta)$, we get:
+
+$$
+\begin{aligned}
+\partial_t h_U(S^{-1}(\alpha), t) &= 
+\begin{aligned}[t]
+	&-h_U(S^{-1}(\alpha), t)\\&+ \int_{[0,1]} w(S^{-1}(\alpha), S^{-1}(\beta)) \phi(h_U(S^{-1}(\beta), t)) \left[\lambda \circ S^{-1}\right](\d \beta)
+\end{aligned}\\
+&= \begin{aligned}[t]
+	&-h_U(\vec v, t)\\&+ \int_{[0,1]} w_U(\vec v, S^{-1}(\beta)) \phi(h_U(S^{-1}(\beta), t)) \left[\lambda \circ S^{-1}\right](\d \beta)
+\end{aligned}\\
+&= -h_U(\vec v, t) + \int_{[0,1]^2} w_U(\vec v, \vec u) \phi(h_U(\vec u, t)) \lambda(\d \vec u)
+\end{aligned}
+$$
+
+In the last line, we used a basic property of the pushfoward measure (see \cite{Bogachev2007}, Theorem 3.6.1, p.
+190).
+The Lebesgue measure $\lambda$ simply plays the role of a uniform distribution on $[0,1]^2$, further simplifying into:
+
+$$
+\partial_t h_U(\vec v, t) = -h_U(\vec v, t) + \int_{[0,1]^2} w(\vec v, \vec u) \phi(h_U(\vec u, t)) \d \vec u.
+$$
+
+Therefore, if $h_U(\vec v, t)$ solves the neural field equation in $[0,1]^2$, then $\tilde h(\alpha, t) = (h_U(\cdot, t) \circ S^{-1})(\alpha)$ solves the equivalent neural field equation in $[0,1]$.
+The same reasoning applies in reverse: if $\tilde h(\alpha, t)$ solves the neural field equation in $[0,1]$, then $h_U(\vec v, t) = (\tilde h(\cdot, t) \circ S)(\vec v)$ solves the neural field in $[0,1]^2$.
 
